### 2025-03-11 paper-cardinalR.Rmd
Initial commit of the cardinalR paper setup

---
title: "cardinalR: Generating interesting high-dimensional data structures"
abstract: >
  A high-dimensional dataset is where each observation is described by many features, or dimensions. Such a dataset might contain various types of structures that have complex geometric properties, such as nonlinear manifolds, clusters, or sparse distributions. We can generate data containing a variety of structures using mathematical functions and statistical distributions. Sampling from a multivariate normal distribution will generate data in an elliptical shape. Using a trigonometric function we can generate a spiral. A torus function can create a donut shape. High-dimensional data structures are useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization. Their controlled complexity allows researchers to understand challenges posed in data analysis and helps to develop robust analytical methods across diverse scientific fields like bioinformatics, machine learning, and forensic science. Functions to generate a large variety of structures in high dimensions are organized into the R package `cardinalR`, along with some already generated examples.

draft: true
author:  
  - name: Jayani P.G. Lakshika
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email:  \email{jayani.piyadigamage@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: application
creative_commons: CC BY
date: "`r Sys.Date()`"
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
output: 
 rjtools::rjournal_web_article:
    css: "style.css"
    keep_md: true
bibliography: RJreferences.bib
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(ggplot2)
library(palmerpenguins)
library(kableExtra)
```

# Introduction

Interactive data graphics provides plots that allow users to interact them. One of the most basic types of interaction is through tooltips, where users are provided additional information about elements in the plot by moving the cursor over the plot.

This paper will first review some R packages on interactive graphics and their tooltip implementations. A new package \CRANpkg{ToOoOlTiPs} that provides customized tooltips for plot, is introduced. Some example plots will then be given to showcase how these tooltips help users to better read the graphics.

# Background

Some packages on interactive graphics include \CRANpkg{plotly} [@plotly] that interfaces with Javascript for web-based interactive graphics, \CRANpkg{crosstalk} [@crosstalk] that specializes cross-linking elements across individual graphics.  The recent R Journal paper \CRANpkg{tsibbletalk} [@RJ-2021-050] provides a good example of including interactive graphics into an article for the journal. It has both a set of linked plots, and also an animated gif example, illustrating linking between time series plots and feature summaries.

# Customizing tooltip design with \pkg{ToOoOlTiPs}

\pkg{ToOoOlTiPs} is a packages for customizing tooltips in interactive graphics, it features these possibilities.

# A gallery of tooltips examples

The \CRANpkg{palmerpenguins} data [@palmerpenguins]  features three penguin species which has a lovely illustration by Alison Horst in Figure \@ref(fig:penguins-alison).

```{r penguins-alison, out.width = "100%", out.height = "30%", fig.cap = "Artwork by \\@allison\\_horst", fig.alt="A picture of three different penguins with their species: Chinstrap, Gentoo, and Adelie. "}
knitr::include_graphics("figures/penguins.png")
```

Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:penguins-tab-interactive)', '\\@ref(tab:penguins-tab-static)'))` prints at the first few rows of the `penguins` data:

```{r penguins-tab-interactive, eval = knitr::is_html_output(), layout = "l-body-outset"}
knitr::kable(head(penguins), format = "html", caption = "A basic table")
```

```{r penguins-tab-static, eval = knitr::is_latex_output()}
knitr::kable(head(penguins), format = "latex", caption = "A basic table") %>% 
  kableExtra::kable_styling(font_size = 7)
```

Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(fig:penguins-plotly)', '\\@ref(fig:penguins-ggplot)'))` shows an `r ifelse(knitr::is_html_output(), 'interactive', '')` plot of the penguins data, made using the `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\CRANpkg{plotly}', '\\CRANpkg{ggplot2}'))` package.

```{r penguins-plotly, echo = TRUE, fig.height = 5, fig.cap="A basic interactive plot made with the plotly package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. When hovering on a point, a tooltip will show the exact value of the bill depth and length for that point, along with the species name.", include=knitr::is_html_output(), eval=knitr::is_html_output(), fig.alt = "A scatterplot of bill length against bill depth, both measured in millimetre. The three species are shown in different colours and loosely forms three clusters. Adelie has small bill length and large bill depth, Gentoo has small bill depth but large bill length, and Chinstrap has relatively large bill depth and bill length."}
p <- penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
ggplotly(p)
```

```{r penguins-ggplot, echo = TRUE, fig.height = 5, fig.cap="A basic non-interactive plot made with the ggplot2 package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. Visit the online article to access the interactive version made with the plotly package.", include=knitr::is_latex_output(), eval=knitr::is_latex_output()}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
```

# Summary

We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}.

### 2025-03-11 paper-cardinalR.tex
Initial commit of the cardinalR paper setup

% !TeX root = RJwrapper.tex
\title{ToOoOlTiPs: An R Package for Customizable Tooltips in Interactive Graphics}

\author{by Quietest Quokka and Bounciest Bilby}

\maketitle

\abstract{%
An abstract of less than 150 words.
}

\section{Introduction}\label{introduction}

Interactive data graphics provides plots that allow users to interact them. One of the most basic types of interaction is through tooltips, where users are provided additional information about elements in the plot by moving the cursor over the plot.

This paper will first review some R packages on interactive graphics and their tooltip implementations. A new package \CRANpkg{ToOoOlTiPs} that provides customized tooltips for plot, is introduced. Some example plots will then be given to showcase how these tooltips help users to better read the graphics.

\section{Background}\label{background}

Some packages on interactive graphics include \CRANpkg{plotly} \citep{plotly} that interfaces with Javascript for web-based interactive graphics, \CRANpkg{crosstalk} \citep{crosstalk} that specializes cross-linking elements across individual graphics. The recent R Journal paper \CRANpkg{tsibbletalk} \citep{RJ-2021-050} provides a good example of including interactive graphics into an article for the journal. It has both a set of linked plots, and also an animated gif example, illustrating linking between time series plots and feature summaries.

\section{\texorpdfstring{Customizing tooltip design with \pkg{ToOoOlTiPs}}{Customizing tooltip design with }}\label{customizing-tooltip-design-with}

\pkg{ToOoOlTiPs} is a packages for customizing tooltips in interactive graphics, it features these possibilities.

\section{A gallery of tooltips examples}\label{a-gallery-of-tooltips-examples}

The \CRANpkg{palmerpenguins} data \citep{palmerpenguins} features three penguin species which has a lovely illustration by Alison Horst in Figure \ref{fig:penguins-alison}.

\begin{figure}
\includegraphics[width=1\linewidth,height=0.3\textheight,alt={A picture of three different penguins with their species: Chinstrap, Gentoo, and Adelie. }]{figures/penguins} \caption{Artwork by \@allison\_horst}\label{fig:penguins-alison}
\end{figure}

Table \ref{tab:penguins-tab-static} prints at the first few rows of the \texttt{penguins} data:

\begin{table}
\centering
\caption{\label{tab:penguins-tab-static}A basic table}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{l|l|r|r|r|r|l|r}
\hline
species & island & bill\_length\_mm & bill\_depth\_mm & flipper\_length\_mm & body\_mass\_g & sex & year\\
\hline
Adelie & Torgersen & 39.1 & 18.7 & 181 & 3750 & male & 2007\\
\hline
Adelie & Torgersen & 39.5 & 17.4 & 186 & 3800 & female & 2007\\
\hline
Adelie & Torgersen & 40.3 & 18.0 & 195 & 3250 & female & 2007\\
\hline
Adelie & Torgersen & NA & NA & NA & NA & NA & 2007\\
\hline
Adelie & Torgersen & 36.7 & 19.3 & 193 & 3450 & female & 2007\\
\hline
Adelie & Torgersen & 39.3 & 20.6 & 190 & 3650 & male & 2007\\
\hline
\end{tabular}
\end{table}

Figure \ref{fig:penguins-ggplot} shows an plot of the penguins data, made using the \CRANpkg{ggplot2} package.

\begin{verbatim}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
\end{verbatim}

\begin{figure}
\centering
\includegraphics{paper-cardinalR_files/figure-latex/penguins-ggplot-1.pdf}
\caption{\label{fig:penguins-ggplot}A basic non-interactive plot made with the ggplot2 package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. Visit the online article to access the interactive version made with the plotly package.}
\end{figure}

\section{Summary}\label{summary}

We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}.

\bibliography{RJreferences.bib}

\address{%
Quietest Quokka\\
University of Little Mates\\%
Department of Letter Q\\ Somewhere, Australia\\
%
\url{https://www.britannica.com/animal/quokka}\\%
\textit{ORCiD: \href{https://orcid.org/0000-1721-1511-1101}{0000-1721-1511-1101}}\\%
\href{mailto:qquo@ulm.edu}{\nolinkurl{qquo@ulm.edu}}%
}

\address{%
Bounciest Bilby\\
University of Little MatesUniversity of Aussie Animals\\%
Department of Letter Q, Somewhere, Australia\\ Department of Marsupials, Somewhere, Australia\\
%
\url{https://www.britannica.com/animal/bilby}\\%
\textit{ORCiD: \href{https://orcid.org/0000-0002-0912-0225}{0000-0002-0912-0225}}\\%
\href{mailto:bbil@ulm.edu}{\nolinkurl{bbil@ulm.edu}}%
}

### 2025-03-11 paper-cardinalR.R

@@ -1,0 +1,38 @@
+# Generated by `rjournal_pdf_article()` using `knitr::purl()`: do not edit by hand
+# Please edit paper-cardinalR.Rmd to modify this file
+
+## ----setup, include=FALSE-----------------------------------------------------
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(plotly)
+library(ggplot2)
+library(palmerpenguins)
+library(kableExtra)
+
+
+## ----penguins-alison, out.width = "100%", out.height = "30%", fig.cap = "Artwork by \\@allison\\_horst", fig.alt="A picture of three different penguins with their species: Chinstrap, Gentoo, and Adelie. "----
+knitr::include_graphics("figures/penguins.png")
+
+
+## ----penguins-tab-interactive, eval = knitr::is_html_output(), layout = "l-body-outset"----
+# knitr::kable(head(penguins), format = "html", caption = "A basic table")
+
+
+## ----penguins-tab-static, eval = knitr::is_latex_output()---------------------
+knitr::kable(head(penguins), format = "latex", caption = "A basic table") %>% 
+  kableExtra::kable_styling(font_size = 7)
+
+
+## ----penguins-plotly, echo = TRUE, fig.height = 5, fig.cap="A basic interactive plot made with the plotly package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. When hovering on a point, a tooltip will show the exact value of the bill depth and length for that point, along with the species name.", include=knitr::is_html_output(), eval=knitr::is_html_output(), fig.alt = "A scatterplot of bill length against bill depth, both measured in millimetre. The three species are shown in different colours and loosely forms three clusters. Adelie has small bill length and large bill depth, Gentoo has small bill depth but large bill length, and Chinstrap has relatively large bill depth and bill length."----
+# p <- penguins %>%
+#   ggplot(aes(x = bill_depth_mm, y = bill_length_mm,
+#              color = species)) +
+#   geom_point()
+# ggplotly(p)
+
+
[…]

### 2025-03-11 paper-cardinalR.Rmd
Add main set up

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-library(plotly)
-library(ggplot2)
-library(palmerpenguins)
-library(kableExtra)
+options(repos = "https://cloud.r-project.org") ## set up CRAN mirror
+knitr::opts_chunk$set(
+  cache=FALSE, 
+  message=FALSE, 
+  warning=FALSE,
+  out.width = "100%")
+
+```
+
+```{r load-libraries}
+
 ```
@@
 
-Interactive data graphics provides plots that allow users to interact them.
-One of the most basic types of interaction is through tooltips, where users are provided additional information about elements in the plot by moving the cursor over the plot.
-
-This paper will first review some R packages on interactive graphics and their tooltip implementations.
-A new package \CRANpkg{ToOoOlTiPs} that provides customized tooltips for plot, is introduced.
-Some example plots will then be given to showcase how these tooltips help users to better read the graphics.
-
 # Background
 
-Some packages on interactive graphics include \CRANpkg{plotly} [@plotly] that interfaces with Javascript for web-based interactive graphics, \CRANpkg{crosstalk} [@crosstalk] that specializes cross-linking elements across individual graphics.
-The recent R Journal paper \CRANpkg{tsibbletalk} [@RJ-2021-050] provides a good example of including interactive graphics into an article for the journal.
-It has both a set of linked plots, and also an animated gif example, illustrating linking between time series plots and feature summaries.
+# Methods
 
-# Customizing tooltip design with \pkg{ToOoOlTiPs}
+# Discussion
 
-\pkg{ToOoOlTiPs} is a packages for customizing tooltips in interactive graphics, it features these possibilities.
+# Code
 
-# A gallery of tooltips examples
+The code is available at <https://github.com/JayaniLakshika/cardinalR>, and source material for this paper is available at <https://github.com/JayaniLakshika/paper-cardinalR>.
 
-The \CRANpkg{palmerpenguins} data [@palmerpenguins]  features three penguin species which has a lovely illustration by Alison Horst in Figure \@ref(fig:penguins-alison).
+# Acknowledgements
 
-```{r penguins-alison, out.width = "100%", out.height = "30%", fig.cap = "Artwork by \\@allison\\_horst", fig.alt="A picture of three different penguins with their species: Chinstrap, Gentoo, and Adelie.
-"}
-knitr::include_graphics("figures/penguins.png")
-```
-
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:penguins-tab-interactive)', '\\@ref(tab:penguins-tab-static)'))` prints at the first few rows of the `penguins` data:
-
-```{r penguins-tab-interactive, eval = knitr::is_html_output(), layout = "l-body-outset"}
-knitr::kable(head(penguins), format = "html", caption = "A basic table")
-```
-
-```{r penguins-tab-static, eval = knitr::is_latex_output()}
-knitr::kable(head(penguins), format = "latex", caption = "A basic table") %>% 
-  kableExtra::kable_styling(font_size = 7)
-```
-
-Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(fig:penguins-plotly)', '\\@ref(fig:penguins-ggplot)'))` shows an `r ifelse(knitr::is_html_output(), 'interactive', '')` plot of the penguins data, made using the `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\CRANpkg{plotly}', '\\CRANpkg{ggplot2}'))` package.
-
-```{r penguins-plotly, echo = TRUE, fig.height = 5, fig.cap="A basic interactive plot made with the plotly package on palmer penguin data.
-Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis.
-When hovering on a point, a tooltip will show the exact value of the bill depth and length for that point, along with the species name.", include=knitr::is_html_output(), eval=knitr::is_html_output(), fig.alt = "A scatterplot of bill length against bill depth, both measured in millimetre.
-The three species are shown in different colours and loosely forms three clusters.
-Adelie has small bill length and large bill depth, Gentoo has small bill depth but large bill length, and Chinstrap has relatively large bill depth and bill length."}
-p <- penguins %>% 
-  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
-             color = species)) + 
-  geom_point()
-ggplotly(p)
-```
-
-```{r penguins-ggplot, echo = TRUE, fig.height = 5, fig.cap="A basic non-interactive plot made with the ggplot2 package on palmer penguin data.
-Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis.
-Visit the online article to access the interactive version made with the plotly package.", include=knitr::is_latex_output(), eval=knitr::is_latex_output()}
-penguins %>% 
-  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
-             color = species)) + 
-  geom_point()
-```
-
-# Summary
-
-We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}.
+This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template.

### 2025-03-30 paper-cardinalR.Rmd
Paper's initial workflow

@@
 knitr::opts_chunk$set(
+  echo = FALSE,
   cache=FALSE, 
@@
 ```{r load-libraries}
-
+library(tidyverse)
+library(knitr)
 ```
@@
 
-# Background
+# Implementation
 
-# Methods
+## Installation
+
+The package can be installed from CRAN using
+
+> ```install.packages("cardinalR")```
+
+and from GitHub using
+
+> ```remotes::install_github("jayanilakshika/cardinalR")```
+
+to install the development version.
+
+## Web site
+
+More documentation of the package can be found at the web site https://jayanilakshika.github.io/cardinalR/.
+
+## Data sets
+
+The `cardinalR` package comes with several data sets that load with the package.
+These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:datasets-tb-html)', '\\@ref(tab:datasets-tb-pdf)'))`.
+
+```{r}
+datasets_tb <- tibble(dt = c("mobius_clust_data", 
+                             "mobius_clust_tsne_param1", 
+                             "mobius_clust_tsne_param2",
+                             "mobius_clust_tsne_param3",
+                             "mobius_clust_umap_param1",
+                             "mobius_clust_umap_param2",
+                             "mobius_clust_umap_param3"), 
+                      text = c("",
+                               "", 
+                               "",
+                               "",
+                               "",
+                               "", ""))
+```
+
+```{r datasets-tb-html, eval=is_html_output()}
+datasets_tb |> 
+  kable(caption = "cardinalR data sets", col.names = c("data", "explanation")) 
+```
+
+```{r datasets-tb-pdf, eval=is_latex_output()}
+datasets_tb |> 
+  kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+## Functions
 

### 2025-04-02 paper-cardinalR.Rmd
Added plan for example
Start writing on Swiss roll data

@@
 
+Research gap with `geozoo`, `snedata` packages...
+
 # Implementation
@@
                                "",
-                               "", ""))
+                               "", 
+                               ""))
 ```
@@
 
+### Swiss Roll
+
+To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function `generate_swiss_roll(n, p, noise_sd)`, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
+
+The function generates `n` points in a `p`-dimensional space, where the first two dimensions (`X_1, X_2`) define the primary Swiss roll shape using a parametric equation:
+
+$$
+X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad \text{where } t \sim U(0, 3\pi)
+$$
+
+The third dimension (`X_3`) introduces variation perpendicular to the roll, sampled uniformly from \([-1,1]\).
+Additional dimensions (`X_4` to `X_p`) extend the data structure by applying a **sinusoidal transformation** of the parameter `t`, ensuring continuity in higher-dimensional spaces:
+
+$$
+X_i = \frac{\sin(i t)}{i}, \quad \text{for } i \geq 4.
+$$
+
+This transformation ensures a gradual decay in variance across dimensions, mimicking real-world high-dimensional structures where later dimensions often capture subtler variations.
+Optionally, Gaussian noise (`\mathcal{N}(0, \sigma^2)`) can be added to all dimensions using `noise_sd` to simulate measurement variability.
+
+# Examples
+
+Add one or two datasets and evaluate how it will be useful.... (one with NLDR, one with clustering)
+
 # Discussion

### 2025-04-02 paper-cardinalR.Rmd
Updated Swiss roll p-D

@@
 
-To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function `generate_swiss_roll(n, p, noise_sd)`, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
+To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function `generate_swiss_roll(n, p)`, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
 
@@
 This transformation ensures a gradual decay in variance across dimensions, mimicking real-world high-dimensional structures where later dimensions often capture subtler variations.
-Optionally, Gaussian noise (`\mathcal{N}(0, \sigma^2)`) can be added to all dimensions using `noise_sd` to simulate measurement variability.
 

### 2025-04-15 paper-cardinalR.Rmd
Add to render both html and pdf versions

@@
     keep_md: true
-bibliography: RJreferences.bib
+bibliography: paper-cardinalR.bib
 editor_options: 

### 2025-04-17 paper-cardinalR.Rmd
Added plan for writing
Initial tabel of branching data

@@
 
+### Main function
+
+### Branching
+
+```{r}
+branch_tb <- tibble(fun = c("gen_expbranches",
+                             "gen_linearbranches",
+                             "gen_curvybranches", 
+                             "gen_orglinearbranches", 
+                             "gen_orgcurvybranches"), 
+                      exp = c("",
+                               "", 
+                               "",
+                               "",
+                               ""))
+```
+
+```{r branching-tb-html, eval=is_html_output()}
+branch_tb |> 
+  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r branching-tb-pdf, eval=is_latex_output()}
+branch_tb |> 
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+### Cone
+
+### Cube
+
+### Gaussian
+
+### Linear
+
+### Mobius
+
+### Polynomial
+
+### Pyramid
+
+### S-curve
+
+### Sphere
+
 ### Swiss Roll
@@
 
+### Odd shapes
+
+### Additional functions
+
 # Examples

### 2025-04-19 paper-cardinalR.Rmd
Added intro

@@
 Functions to generate a large variety of structures in high dimensions are organized into the R package `cardinalR`, along with some already generated examples.
-
 draft: true
@@
     orcid: 0000-0002-0656-9789
-type: application
+type: package
 creative_commons: CC BY
@@
 output: 
- rjtools::rjournal_web_article:
-    css: "style.css"
+ rjtools::rjournal_article:
     keep_md: true
@@
 library(knitr)
+library(kableExtra)
 ```
@@
 
-Research gap with `geozoo`, `snedata` packages...
+Okay, based on your research focus of creating a package for generating high-dimensional data structures with controlled noise and cluster generation based on geometric primitives, here's an introduction that highlights potential research gaps in relation to `geozoo` and `snedata`:
+
+**Introduction:**
+
+The generation of synthetic datasets with well-defined structural properties is crucial for evaluating and benchmarking algorithms across various fields, including machine learning, data mining, and spatial statistics.
+Researchers often require the ability to create data with specific dimensionalities, noise characteristics, and underlying cluster structures to rigorously test the performance and robustness of their methods.
+While existing R packages like `geozoo` and `snedata` offer valuable tools for working with geometric objects and spatial network data, respectively, a gap exists in the comprehensive and flexible generation of high-dimensional data structures with integrated geometric underpinnings and controlled noise injection.
+
+Specifically, `geozoo` provides an extensive collection of geometric objects, primarily focused on fixed, often low-dimensional, shapes.
+While useful for representing and visualizing these forms, it lacks direct functionalities for generating high-dimensional datasets based on these geometries, nor does it inherently incorporate mechanisms for adding controlled noise or creating clustered structures based on geometric primitives with user-defined parameters such as position, scale, and orientation in arbitrary dimensions.
+
+On the other hand, `snedata` focuses on spatial network datasets, which, while inherently spatial, are often represented in lower-dimensional geographic spaces.
+It does not offer general tools for generating high-dimensional non-networked data structures based on flexible geometric layouts or the systematic addition of noise dimensions and background noise.
+Furthermore, the generation of clustered data based on basic geometric shapes with fine-grained control over cluster properties is not a primary focus of this package.
+
+Therefore, a need exists for a dedicated R package that provides a versatile framework for generating synthetic data structures in arbitrary high dimensions, starting from basic geometric primitives.
+Such a package would ideally offer functionalities to: (i) create data structures based on geometric shapes that can be embedded and extended to higher dimensions by adding noise dimensions; (ii) introduce controlled levels of background noise to these structures; and (iii) generate clustered data by leveraging fundamental geometric shapes with user-specified positions, scales, orientations, and sample sizes in any desired dimensionality.
+This would empower researchers to create more realistic and challenging synthetic datasets for evaluating algorithms in various high-dimensional scenarios and spatial contexts.
+
+This paper introduces the `cardinalR` R package, which aims to address these limitations by providing a comprehensive suite of functions for generating customizable high-dimensional data structures based on geometric primitives, with integrated noise injection and flexible cluster generation capabilities.
+By offering these functionalities, `cardinalR` seeks to provide a valuable resource for researchers needing to create synthetic datasets tailored to their specific evaluation needs in the realm of high-dimensional data analysis and beyond.
 

### 2025-04-19 paper-cardinalR.tex
Added intro

@@
 % !TeX root = RJwrapper.tex
-\title{ToOoOlTiPs: An R Package for Customizable Tooltips in Interactive Graphics}
+\title{cardinalR: Generating interesting high-dimensional data structures}
 
-\author{by Quietest Quokka and Bounciest Bilby}
+\author{by Jayani P.G. Lakshika, Dianne Cook, Paul Harrison, Michael Lydeamore, and Thiyanga S. Talagala}
 
@@
 \abstract{%
-An abstract of less than 150 words.
+A high-dimensional dataset is where each observation is described by many features, or dimensions.
+Such a dataset might contain various types of structures that have complex geometric properties, such as nonlinear manifolds, clusters, or sparse distributions.
+We can generate data containing a variety of structures using mathematical functions and statistical distributions.
+Sampling from a multivariate normal distribution will generate data in an elliptical shape.
+Using a trigonometric function we can generate a spiral.
+A torus function can create a donut shape.
+High-dimensional data structures are useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
+Their controlled complexity allows researchers to understand challenges posed in data analysis and helps to develop robust analytical methods across diverse scientific fields like bioinformatics, machine learning, and forensic science.
+Functions to generate a large variety of structures in high dimensions are organized into the R package \texttt{cardinalR}, along with some already generated examples.
 }
@@
 
-Interactive data graphics provides plots that allow users to interact them.
-One of the most basic types of interaction is through tooltips, where users are provided additional information about elements in the plot by moving the cursor over the plot.
+Okay, based on your research focus of creating a package for generating high-dimensional data structures with controlled noise and cluster generation based on geometric primitives, here's an introduction that highlights potential research gaps in relation to \texttt{geozoo} and \texttt{snedata}:
 
-This paper will first review some R packages on interactive graphics and their tooltip implementations.
-A new package \CRANpkg{ToOoOlTiPs} that provides customized tooltips for plot, is introduced.
-Some example plots will then be given to showcase how these tooltips help users to better read the graphics.
+\textbf{Introduction:}
 
-\section{Background}\label{background}
+The generation of synthetic datasets with well-defined structural properties is crucial for evaluating and benchmarking algorithms across various fields, including machine learning, data mining, and spatial statistics.
+Researchers often require the ability to create data with specific dimensionalities, noise characteristics, and underlying cluster structures to rigorously test the performance and robustness of their methods.
+While existing R packages like \texttt{geozoo} and \texttt{snedata} offer valuable tools for working with geometric objects and spatial network data, respectively, a gap exists in the comprehensive and flexible generation of high-dimensional data structures with integrated geometric underpinnings and controlled noise injection.
 
-Some packages on interactive graphics include \CRANpkg{plotly} \citep{plotly} that interfaces with Javascript for web-based interactive graphics, \CRANpkg{crosstalk} \citep{crosstalk} that specializes cross-linking elements across individual graphics.
-The recent R Journal paper \CRANpkg{tsibbletalk} \citep{RJ-2021-050} provides a good example of including interactive graphics into an article for the journal.
-It has both a set of linked plots, and also an animated gif example, illustrating linking between time series plots and feature summaries.
+Specifically, \texttt{geozoo} provides an extensive collection of geometric objects, primarily focused on fixed, often low-dimensional, shapes.
+While useful for representing and visualizing these forms, it lacks direct functionalities for generating high-dimensional datasets based on these geometries, nor does it inherently incorporate mechanisms for adding controlled noise or creating clustered structures based on geometric primitives with user-defined parameters such as position, scale, and orientation in arbitrary dimensions.
 
-\section{\texorpdfstring{Customizing tooltip design with \pkg{ToOoOlTiPs}}{Customizing tooltip design with }}\label{customizing-tooltip-design-with}
+On the other hand, \texttt{snedata} focuses on spatial network datasets, which, while inherently spatial, are often represented in lower-dimensional geographic spaces.
+It does not offer general tools for generating high-dimensional non-networked data structures based on flexible geometric layouts or the systematic addition of noise dimensions and background noise.
+Furthermore, the generation of clustered data based on basic geometric shapes with fine-grained control over cluster properties is not a primary focus of this package.
 
-\pkg{ToOoOlTiPs} is a packages for customizing tooltips in interactive graphics, it features these possibilities.
+Therefore, a need exists for a dedicated R package that provides a versatile framework for generating synthetic data structures in arbitrary high dimensions, starting from basic geometric primitives.
+Such a package would ideally offer functionalities to: (i) create data structures based on geometric shapes that can be embedded and extended to higher dimensions by adding noise dimensions; (ii) introduce controlled levels of background noise to these structures; and (iii) generate clustered data by leveraging fundamental geometric shapes with user-specified positions, scales, orientations, and sample sizes in any desired dimensionality.
+This would empower researchers to create more realistic and challenging synthetic datasets for evaluating algorithms in various high-dimensional scenarios and spatial contexts.
 
-\section{A gallery of tooltips examples}\label{a-gallery-of-tooltips-examples}
+This paper introduces the \texttt{cardinalR} R package, which aims to address these limitations by providing a comprehensive suite of functions for generating customizable high-dimensional data structures based on geometric primitives, with integrated noise injection and flexible cluster generation capabilities.
+By offering these functionalities, \texttt{cardinalR} seeks to provide a valuable resource for researchers needing to create synthetic datasets tailored to their specific evaluation needs in the realm of high-dimensional data analysis and beyond.
 
-The \CRANpkg{palmerpenguins} data \citep{palmerpenguins} features three penguin species which has a lovely illustration by Alison Horst in Figure \ref{fig:penguins-alison}.
+\section{Implementation}\label{implementation}
 
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.3\textheight,alt={A picture of three different penguins with their species: Chinstrap, Gentoo, and Adelie.
-}]{figures/penguins} \caption{Artwork by \@allison\_horst}\label{fig:penguins-alison}
-\end{figure}
+\subsection{Installation}\label{installation}
 
-Table \ref{tab:penguins-tab-static} prints at the first few rows of the \texttt{penguins} data:
+The package can be installed from CRAN using
+
+\begin{quote}
+\texttt{install.packages("cardinalR")}
+\end{quote}
+
+and from GitHub using
+
+\begin{quote}
+\texttt{remotes::install\_github("jayanilakshika/cardinalR")}
+\end{quote}
+
+to install the development version.
+
+\subsection{Web site}\label{web-site}
+
+More documentation of the package can be found at the web site \url{https://jayanilakshika.github.io/cardinalR/}.
+
+\subsection{Data sets}\label{data-sets}
+
+The \texttt{cardinalR} package comes with several data sets that load with the package. These are described in Table \ref{tab:datasets-tb-pdf}.
 
 \begin{table}
+
+\caption{\label{tab:datasets-tb-pdf}cardinalR data sets}
 \centering
-\caption{\label{tab:penguins-tab-static}A basic table}
-\centering
-\fontsize{7}{9}\selectfont
-\begin{tabular}[t]{l|l|r|r|r|r|l|r}
-\hline
-species & island & bill\_length\_mm & bill\_depth\_mm & flipper\_length\_mm & body\_mass\_g & sex & year\\
-\hline
-Adelie & Torgersen & 39.1 & 18.7 & 181 & 3750 & male & 2007\\
-\hline
-Adelie & Torgersen & 39.5 & 17.4 & 186 & 3800 & female & 2007\\
-\hline
-Adelie & Torgersen & 40.3 & 18.0 & 195 & 3250 & female & 2007\\
-\hline
-Adelie & Torgersen & NA & NA & NA & NA & NA & 2007\\
-\hline
-Adelie & Torgersen & 36.7 & 19.3 & 193 & 3450 & female & 2007\\
-\hline
-Adelie & Torgersen & 39.3 & 20.6 & 190 & 3650 & male & 2007\\
-\hline
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+data & explanation\\
+\midrule
+mobius\_clust\_data & \\
+mobius\_clust\_tsne\_param1 & \\
+mobius\_clust\_tsne\_param2 & \\
+mobius\_clust\_tsne\_param3 & \\
+mobius\_clust\_umap\_param1 & \\
+\addlinespace
+mobius\_clust\_umap\_param2 & \\
+mobius\_clust\_umap\_param3 & \\
+\bottomrule
 \end{tabular}
@@
 
-Figure \ref{fig:penguins-ggplot} shows an plot of the penguins data, made using the \CRANpkg{ggplot2} package.
+\subsection{Functions}\label{functions}
 
-\begin{verbatim}
-penguins %>% 
-  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
-             color = species)) + 
-  geom_point()
-\end{verbatim}
+\subsubsection{Main function}\label{main-function}
 
-\begin{figure}
+\subsubsection{Branching}\label{branching}
+
+\begin{table}
+
+\caption{\label{tab:branching-tb-pdf}cardinalR branching data generation functions}
 \centering
-\includegraphics{paper-cardinalR_files/figure-latex/penguins-ggplot-1.pdf}
-\caption{\label{fig:penguins-ggplot}A basic non-interactive plot made with the ggplot2 package on palmer penguin data.
-Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis.
-Visit the online article to access the interactive version made with the plotly package.}
-\end{figure}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_expbranches & \\
+gen\_linearbranches & \\
+gen\_curvybranches & \\
+gen\_orglinearbranches & \\
+gen\_orgcurvybranches & \\
+\bottomrule
+\end{tabular}
+\end{table}
 
-\section{Summary}\label{summary}
+\subsubsection{Cone}\label{cone}
 
-We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}.
+\subsubsection{Cube}\label{cube}
 
-\bibliography{RJreferences.bib}
+\subsubsection{Gaussian}\label{gaussian}
+
+\subsubsection{Linear}\label{linear}
+
+\subsubsection{Mobius}\label{mobius}
+
+\subsubsection{Polynomial}\label{polynomial}
+
+\subsubsection{Pyramid}\label{pyramid}
+
+\subsubsection{S-curve}\label{s-curve}
+
+\subsubsection{Sphere}\label{sphere}
+
+\subsubsection{Swiss Roll}\label{swiss-roll}
+
+To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function \texttt{generate\_swiss\_roll(n,\ p)}, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
+
+The function generates \texttt{n} points in a \texttt{p}-dimensional space, where the first two dimensions (\texttt{X\_1,\ X\_2}) define the primary Swiss roll shape using a parametric equation:
+
+\[
+X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad \text{where } t \sim U(0, 3\pi)
+\]
+
+The third dimension (\texttt{X\_3}) introduces variation perpendicular to the roll, sampled uniformly from \([-1,1]\).
+Additional dimensions (\texttt{X\_4} to \texttt{X\_p}) extend the data structure by applying a \textbf{sinusoidal transformation} of the parameter \texttt{t}, ensuring continuity in higher-dimensional spaces:
+
+\[
+X_i = \frac{\sin(i t)}{i}, \quad \text{for } i \geq 4.
+\]
+
+This transformation ensures a gradual decay in variance across dimensions, mimicking real-world high-dimensional structures where later dimensions often capture subtler variations.
+
+\subsubsection{Odd shapes}\label{odd-shapes}
+
+\subsubsection{Additional functions}\label{additional-functions}
+
+\section{Examples}\label{examples}
+
+Add one or two datasets and evaluate how it will be useful\ldots. (one with NLDR, one with clustering)
+
+\section{Discussion}\label{discussion}
+
+\section{Code}\label{code}
+
+The code is available at \url{https://github.com/JayaniLakshika/cardinalR}, and source material for this paper is available at \url{https://github.com/JayaniLakshika/paper-cardinalR}.
+
+\section{Acknowledgements}\label{acknowledgements}
+
+This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.
+
+\bibliography{paper-cardinalR.bib}
 
 \address{%
-Quietest Quokka\\
-University of Little Mates\\%
-Department of Letter Q\\ Somewhere, Australia\\
+Jayani P.G. Lakshika\\
+Monash University\\%
+Department of Econometrics and Business Statistics, VIC 3800 Australia\\
 %
-\url{https://www.britannica.com/animal/quokka}\\%
-\textit{ORCiD: \href{https://orcid.org/0000-1721-1511-1101}{0000-1721-1511-1101}}\\%
-\href{mailto:qquo@ulm.edu}{\nolinkurl{qquo@ulm.edu}}%
+\url{https://jayanilakshika.netlify.app/}\\%
+\textit{ORCiD: \href{https://orcid.org/0000-0002-6265-6481}{0000-0002-6265-6481}}\\%
+\email{jayani.piyadigamage@monash.edu}%
 }
@@
 \address{%
-Bounciest Bilby\\
-University of Little MatesUniversity of Aussie Animals\\%
-Department of Letter Q, Somewhere, Australia\\ Department of Marsupials, Somewhere, Australia\\
+Dianne Cook\\
+Monash University\\%
+Department of Econometrics and Business Statistics, VIC 3800 Australia\\
 %
-\url{https://www.britannica.com/animal/bilby}\\%
-\textit{ORCiD: \href{https://orcid.org/0000-0002-0912-0225}{0000-0002-0912-0225}}\\%
-\href{mailto:bbil@ulm.edu}{\nolinkurl{bbil@ulm.edu}}%
+\url{http://www.dicook.org/}\\%
+\textit{ORCiD: \href{https://orcid.org/0000-0002-3813-7155}{0000-0002-3813-7155}}\\%
+\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
 }
+
+\address{%
+Paul Harrison\\
+Monash University\\%
+MGBP, BDInstitute, VIC 3800 Australia\\
+%
+%
+\textit{ORCiD: \href{https://orcid.org/0000-0002-3980-268X}{0000-0002-3980-268X}}\\%
+\href{mailto:paul.harrison@monash.edu}{\nolinkurl{paul.harrison@monash.edu}}%
+}
+
+\address{%
+Michael Lydeamore\\
+Monash University\\%
+Department of Econometrics and Business Statistics, VIC 3800 Australia\\
+%
+%
+\textit{ORCiD: \href{https://orcid.org/0000-0001-6515-827X}{0000-0001-6515-827X}}\\%
+\href{mailto:michael.lydeamore@monash.edu}{\nolinkurl{michael.lydeamore@monash.edu}}%
+}
+
+\address{%
+Thiyanga S. Talagala\\
+University of Sri Jayewardenepura\\%
+Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka\\
+%
+\url{https://thiyanga.netlify.app/}\\%
+\textit{ORCiD: \href{https://orcid.org/0000-0002-0656-9789}{0000-0002-0656-9789}}\\%
+\href{mailto:ttalagala@sjp.ac.lk}{\nolinkurl{ttalagala@sjp.ac.lk}}%
+}

### 2025-04-19 paper-cardinalR.Rmd
Set up the web article

@@
 output: 
- rjtools::rjournal_article:
+ rjtools::rjournal_web_article:
+    css: "style.css"
     keep_md: true
@@
 library(knitr)
-library(kableExtra)
 ```
@@
 # Introduction
-
-Okay, based on your research focus of creating a package for generating high-dimensional data structures with controlled noise and cluster generation based on geometric primitives, here's an introduction that highlights potential research gaps in relation to `geozoo` and `snedata`:
-
-**Introduction:**
 

### 2025-04-19 paper-cardinalR.tex
Set up the web article

@@
 \section{Introduction}\label{introduction}
-
-Okay, based on your research focus of creating a package for generating high-dimensional data structures with controlled noise and cluster generation based on geometric primitives, here's an introduction that highlights potential research gaps in relation to \texttt{geozoo} and \texttt{snedata}:
-
-\textbf{Introduction:}
 
@@
 
-\begin{table}
-
-\caption{\label{tab:datasets-tb-pdf}cardinalR data sets}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-data & explanation\\
-\midrule
-mobius\_clust\_data & \\
-mobius\_clust\_tsne\_param1 & \\
-mobius\_clust\_tsne\_param2 & \\
-mobius\_clust\_tsne\_param3 & \\
-mobius\_clust\_umap\_param1 & \\
-\addlinespace
-mobius\_clust\_umap\_param2 & \\
-mobius\_clust\_umap\_param3 & \\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \subsection{Functions}\label{functions}
-
-\subsubsection{Main function}\label{main-function}
-
-\subsubsection{Branching}\label{branching}
-
-\begin{table}
-
-\caption{\label{tab:branching-tb-pdf}cardinalR branching data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_expbranches & \\
-gen\_linearbranches & \\
-gen\_curvybranches & \\
-gen\_orglinearbranches & \\
-gen\_orgcurvybranches & \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsubsection{Cone}\label{cone}
-
-\subsubsection{Cube}\label{cube}
-
-\subsubsection{Gaussian}\label{gaussian}
-
-\subsubsection{Linear}\label{linear}
-
-\subsubsection{Mobius}\label{mobius}
-
-\subsubsection{Polynomial}\label{polynomial}
-
-\subsubsection{Pyramid}\label{pyramid}
-
-\subsubsection{S-curve}\label{s-curve}
-
-\subsubsection{Sphere}\label{sphere}
 
@@
 This transformation ensures a gradual decay in variance across dimensions, mimicking real-world high-dimensional structures where later dimensions often capture subtler variations.
-
-\subsubsection{Odd shapes}\label{odd-shapes}
-
-\subsubsection{Additional functions}\label{additional-functions}
 

### 2025-04-20 paper-cardinalR.Rmd
Added paper outline

--- 2025-04-17 paper-cardinalR.Rmd
@@
 
-Research gap with `geozoo`, `snedata` packages...
+snedata, geozoo, mlbench, splatter
+
+The paper is organized as follows.
+In next section, introduces the implementation of `cardinalR` package on CRAN and GitHub, including demonstration of the package's key functions.
+We illustrate how a clustering data structure affect the Non-linear dimension reductions in **Application** section.
+Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.
 

### 2025-04-20 paper-cardinalR.Rmd
Added intial introduction version

@@
 
-snedata, geozoo, mlbench, splatter
+The generation of synthetic datasets with well-defined structural properties is a cornerstone of evaluating and benchmarking algorithms across diverse fields, including machine learning, data mining, and computational biology.
+Researchers frequently require the ability to create data with specific dimensionalities, noise characteristics, and intricate underlying structures to rigorously test the performance and robustness of their methods.
+The R ecosystem offers a variety of packages for generating synthetic data, each with its strengths and focus.
+For instance, **`geozoo` (Schloerke, [Year of last commit on GitHub])** provides a rich collection of geometric objects, facilitating the creation and analysis of specific shapes, primarily in lower-dimensional spaces.
+**`snedata` (Melville, [Year of last commit on GitHub])** offers tools for generating simplified datasets tailored for evaluating dimensionality reduction techniques like t-SNE, often focusing on revealing low-dimensional embeddings of complex structures.
+Packages like **`mlbench` (Leisch & Dimitriadou, 2010)** provide a collection of well-known benchmark datasets, often with established classification or regression challenges.
+In the realm of single-cell omics, **`splatter` (Zappia et al., 2017)** excels at simulating complex biological data, capturing nuances like batch effects and differential expression.
+
+However, a research gap exists in the comprehensive and flexible generation of high-dimensional data structures that seamlessly integrate fundamental geometric principles with sophisticated noise control and customizable cluster generation.
+While `geozoo` offers geometric primitives, it lacks direct mechanisms for extending these structures to arbitrary high dimensions with controlled noise or for generating clustered data based on these shapes with user-defined parameters like position, scale, and orientation.
+`snedata` focuses on creating abstract datasets for dimensionality reduction evaluation, not on building high-dimensional data directly from geometric layouts with flexible noise and clustering options.
+`mlbench` provides static benchmark datasets, limiting the ability to generate data with specific, user-defined geometric structures and noise profiles in high dimensions.
+Similarly, `splatter` is specialized for biological data simulation and does not offer the general framework for creating diverse geometric structures in arbitrary dimensions with fine-grained control over noise and cluster arrangements based on basic geometric shapes.
+
+To address this gap, this paper introduces the `cardinalR` R package.
+This novel package provides a versatile suite of functions designed to generate customizable data structures in any number of dimensions, starting from fundamental geometric primitives.
+`cardinalR` offers key functionalities that go beyond the capabilities of existing tools by enabling users to: (i) construct high-dimensional datasets based on geometric shapes, with the ability to augment dimensionality through the addition of controlled noise dimensions; (ii) introduce tunable levels of background noise to these structures; and (iii) generate complex clustered data arrangements by leveraging basic geometric forms with precise control over their positions, scales, orientations, and sample sizes in arbitrary dimensional spaces.
+By providing these integrated functionalities, `cardinalR` aims to empower researchers to create more nuanced and challenging synthetic datasets tailored to the specific demands of evaluating algorithms in high-dimensional settings and complex spatial scenarios, bridging the gap between geometric foundations and the flexible generation of complex synthetic data.
+
+This paper will detail the architecture and functionalities of the `cardinalR` package, demonstrating its utility through illustrative examples and highlighting its advantages over existing R packages like `geozoo`, `snedata`, `mlbench`, and `splatter` in the context of generating high-dimensional, geometrically-informed, and noise-controlled synthetic data.
 

### 2025-04-20 paper-cardinalR.Rmd
Updated first intro paragraph

@@
 
-The generation of synthetic datasets with well-defined structural properties is a cornerstone of evaluating and benchmarking algorithms across diverse fields, including machine learning, data mining, and computational biology.
-Researchers frequently require the ability to create data with specific dimensionalities, noise characteristics, and intricate underlying structures to rigorously test the performance and robustness of their methods.
+Generating synthetic datasets with clearly defined geometric properties is essential for evaluating and benchmarking algorithms in various fields, such as machine learning, data mining, and computational biology.
+Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
+
 The R ecosystem offers a variety of packages for generating synthetic data, each with its strengths and focus.

### 2025-04-20 paper-cardinalR.Rmd
Updated second paragraph of intro

@@
 
-The R ecosystem offers a variety of packages for generating synthetic data, each with its strengths and focus.
-For instance, **`geozoo` (Schloerke, [Year of last commit on GitHub])** provides a rich collection of geometric objects, facilitating the creation and analysis of specific shapes, primarily in lower-dimensional spaces.
-**`snedata` (Melville, [Year of last commit on GitHub])** offers tools for generating simplified datasets tailored for evaluating dimensionality reduction techniques like t-SNE, often focusing on revealing low-dimensional embeddings of complex structures.
-Packages like **`mlbench` (Leisch & Dimitriadou, 2010)** provide a collection of well-known benchmark datasets, often with established classification or regression challenges.
-In the realm of single-cell omics, **`splatter` (Zappia et al., 2017)** excels at simulating complex biological data, capturing nuances like batch effects and differential expression.
+There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.
+For example, `geozoo` (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
+Another useful package is `snedata` (@james2025), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
+Additionally, `mlbench` (@friedrich2024) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+In the field of single-cell omics, `splatter` (@luke2017) is particularly simulate complex biological data, effectively capturing nuances such as batch effects and differential expression.
 

### 2025-04-20 paper-cardinalR.Rmd
Added third paragraph for introduction

@@
 By providing these integrated functionalities, `cardinalR` aims to empower researchers to create more nuanced and challenging synthetic datasets tailored to the specific demands of evaluating algorithms in high-dimensional settings and complex spatial scenarios, bridging the gap between geometric foundations and the flexible generation of complex synthetic data.
-
-This paper will detail the architecture and functionalities of the `cardinalR` package, demonstrating its utility through illustrative examples and highlighting its advantages over existing R packages like `geozoo`, `snedata`, `mlbench`, and `splatter` in the context of generating high-dimensional, geometrically-informed, and noise-controlled synthetic data.
 

### 2025-04-20 paper-cardinalR.Rmd
Added intro

@@
 
-However, a research gap exists in the comprehensive and flexible generation of high-dimensional data structures that seamlessly integrate fundamental geometric principles with sophisticated noise control and customizable cluster generation.
-While `geozoo` offers geometric primitives, it lacks direct mechanisms for extending these structures to arbitrary high dimensions with controlled noise or for generating clustered data based on these shapes with user-defined parameters like position, scale, and orientation.
-`snedata` focuses on creating abstract datasets for dimensionality reduction evaluation, not on building high-dimensional data directly from geometric layouts with flexible noise and clustering options.
-`mlbench` provides static benchmark datasets, limiting the ability to generate data with specific, user-defined geometric structures and noise profiles in high dimensions.
-Similarly, `splatter` is specialized for biological data simulation and does not offer the general framework for creating diverse geometric structures in arbitrary dimensions with fine-grained control over noise and cluster arrangements based on basic geometric shapes.
+There is a valuable opportunity to improve the generation of high-dimensional data structures by integrating geometric principles with advanced noise control and customizable clustering.
+The `geozoo` package provides a strong foundation but could be enhanced to support high-dimensional extensions with controlled noise and user-defined parameters for clustering.
+Similarly, while `snedata` focuses on abstract datasets for dimensionality reduction, adding features for generating high-dimensional data from geometric layouts would enhance its usability.
+The `mlbench` package could also benefit from allowing users to create datasets with specific geometric structures and noise profiles.
+Additionally, although `splatter` specializes in biological data simulation, it could be expanded to offer a broader framework for generating diverse geometric structures across dimensions, enabling detailed control over noise and clustering.
+Addressing these areas could lead to more robust high-dimensional data generation tools.
 
 To address this gap, this paper introduces the `cardinalR` R package.
-This novel package provides a versatile suite of functions designed to generate customizable data structures in any number of dimensions, starting from fundamental geometric primitives.
-`cardinalR` offers key functionalities that go beyond the capabilities of existing tools by enabling users to: (i) construct high-dimensional datasets based on geometric shapes, with the ability to augment dimensionality through the addition of controlled noise dimensions; (ii) introduce tunable levels of background noise to these structures; and (iii) generate complex clustered data arrangements by leveraging basic geometric forms with precise control over their positions, scales, orientations, and sample sizes in arbitrary dimensional spaces.
-By providing these integrated functionalities, `cardinalR` aims to empower researchers to create more nuanced and challenging synthetic datasets tailored to the specific demands of evaluating algorithms in high-dimensional settings and complex spatial scenarios, bridging the gap between geometric foundations and the flexible generation of complex synthetic data.
+This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes.
+`cardinalR` offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) create complex clustered data arrangements by using fundamental geometric forms, while maintaining their positions, scales, orientations, and sample sizes in arbitrary dimensional spaces.
+By providing these integrated features, `cardinalR` aims to provide researchers to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
+This bridges the gap between geometric foundations and the flexible generation of complex synthetic data.
 

### 2025-04-20 paper-cardinalR.tex
Added intro for pdf

--- 2025-04-19 paper-cardinalR.tex
@@
 
-Okay, based on your research focus of creating a package for generating high-dimensional data structures with controlled noise and cluster generation based on geometric primitives, here's an introduction that highlights potential research gaps in relation to \texttt{geozoo} and \texttt{snedata}:
+Generating synthetic datasets with clearly defined geometric properties is essential for evaluating and benchmarking algorithms in various fields, such as machine learning, data mining, and computational biology.
+Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
 
-\textbf{Introduction:}
+There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.
+For example, \texttt{geozoo} (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
+Another useful package is \texttt{snedata} (\citet{james2025}), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
+Additionally, \texttt{mlbench} (\citet{friedrich2024}) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+In the field of single-cell omics, \texttt{splatter} (\citet{luke2017}) is particularly simulate complex biological data, effectively capturing nuances such as batch effects and differential expression.
 
-The generation of synthetic datasets with well-defined structural properties is crucial for evaluating and benchmarking algorithms across various fields, including machine learning, data mining, and spatial statistics.
-Researchers often require the ability to create data with specific dimensionalities, noise characteristics, and underlying cluster structures to rigorously test the performance and robustness of their methods.
-While existing R packages like \texttt{geozoo} and \texttt{snedata} offer valuable tools for working with geometric objects and spatial network data, respectively, a gap exists in the comprehensive and flexible generation of high-dimensional data structures with integrated geometric underpinnings and controlled noise injection.
+There is a valuable opportunity to improve the generation of high-dimensional data structures by integrating geometric principles with advanced noise control and customizable clustering.
+The \texttt{geozoo} package provides a strong foundation but could be enhanced to support high-dimensional extensions with controlled noise and user-defined parameters for clustering.
+Similarly, while \texttt{snedata} focuses on abstract datasets for dimensionality reduction, adding features for generating high-dimensional data from geometric layouts would enhance its usability.
+The \texttt{mlbench} package could also benefit from allowing users to create datasets with specific geometric structures and noise profiles.
+Additionally, although \texttt{splatter} specializes in biological data simulation, it could be expanded to offer a broader framework for generating diverse geometric structures across dimensions, enabling detailed control over noise and clustering.
+Addressing these areas could lead to more robust high-dimensional data generation tools.
 
-Specifically, \texttt{geozoo} provides an extensive collection of geometric objects, primarily focused on fixed, often low-dimensional, shapes.
-While useful for representing and visualizing these forms, it lacks direct functionalities for generating high-dimensional datasets based on these geometries, nor does it inherently incorporate mechanisms for adding controlled noise or creating clustered structures based on geometric primitives with user-defined parameters such as position, scale, and orientation in arbitrary dimensions.
+To address this gap, this paper introduces the \texttt{cardinalR} R package.
+This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes.
+\texttt{cardinalR} offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) create complex clustered data arrangements by using fundamental geometric forms, while maintaining their positions, scales, orientations, and sample sizes in arbitrary dimensional spaces.
+By providing these integrated features, \texttt{cardinalR} aims to provide researchers to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
+This bridges the gap between geometric foundations and the flexible generation of complex synthetic data.
 
-On the other hand, \texttt{snedata} focuses on spatial network datasets, which, while inherently spatial, are often represented in lower-dimensional geographic spaces.
-It does not offer general tools for generating high-dimensional non-networked data structures based on flexible geometric layouts or the systematic addition of noise dimensions and background noise.
-Furthermore, the generation of clustered data based on basic geometric shapes with fine-grained control over cluster properties is not a primary focus of this package.
-
-Therefore, a need exists for a dedicated R package that provides a versatile framework for generating synthetic data structures in arbitrary high dimensions, starting from basic geometric primitives.
-Such a package would ideally offer functionalities to: (i) create data structures based on geometric shapes that can be embedded and extended to higher dimensions by adding noise dimensions; (ii) introduce controlled levels of background noise to these structures; and (iii) generate clustered data by leveraging fundamental geometric shapes with user-specified positions, scales, orientations, and sample sizes in any desired dimensionality.
-This would empower researchers to create more realistic and challenging synthetic datasets for evaluating algorithms in various high-dimensional scenarios and spatial contexts.
-
-This paper introduces the \texttt{cardinalR} R package, which aims to address these limitations by providing a comprehensive suite of functions for generating customizable high-dimensional data structures based on geometric primitives, with integrated noise injection and flexible cluster generation capabilities.
-By offering these functionalities, \texttt{cardinalR} seeks to provide a valuable resource for researchers needing to create synthetic datasets tailored to their specific evaluation needs in the realm of high-dimensional data analysis and beyond.
+The paper is organized as follows.
+In next section, introduces the implementation of \texttt{cardinalR} package on CRAN and GitHub, including demonstration of the package's key functions.
+We illustrate how a clustering data structure affect the Non-linear dimension reductions in \textbf{Application} section.
+Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.
 

### 2025-04-20 paper-cardinalR.R

@@ -15,6 +15,7 @@ knitr::opts_chunk$set(
 ## ----load-libraries-----------------------------------------------------------
 library(tidyverse)
 library(knitr)
+library(kableExtra)
 
 
 ## -----------------------------------------------------------------------------
@@ -32,4 +33,41 @@ datasets_tb <- tibble(dt = c("mobius_clust_data",
                                "",
                                "", 
                                ""))
+
+
+## ----datasets-tb-html, eval=is_html_output()----------------------------------
+# datasets_tb |>
+#   kable(caption = "cardinalR data sets", col.names = c("data", "explanation"))
+
+
+## ----datasets-tb-pdf, eval=is_latex_output()----------------------------------
+datasets_tb |> 
+  kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+
+
+## -----------------------------------------------------------------------------
+branch_tb <- tibble(fun = c("gen_expbranches",
+                             "gen_linearbranches",
+                             "gen_curvybranches", 
+                             "gen_orglinearbranches", 
+                             "gen_orgcurvybranches"), 
+                      exp = c("",
+                               "", 
+                               "",
+                               "",
+                               ""))
+
+
+## ----branching-tb-html, eval=is_html_output()---------------------------------
+# branch_tb |>
+#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"))
+
+
+## ----branching-tb-pdf, eval=is_latex_output()---------------------------------
+branch_tb |> 
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
 

### 2025-04-20 paper-cardinalR.Rmd
Added missing citations

@@
 In next section, introduces the implementation of `cardinalR` package on CRAN and GitHub, including demonstration of the package's key functions.
-We illustrate how a clustering data structure affect the Non-linear dimension reductions in **Application** section.
+We illustrate how a clustering data structure affect the Non-linear dimension reductions in **Example** section.
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.
@@
 
-# Examples
+# Example
 
@@
 
-This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template.
+This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.

### 2025-04-20 paper-cardinalR.Rmd
Updated installation section
Updated my name

@@
 author:  
-  - name: Jayani P.G. Lakshika
+  - name: Jayani P. Gamage
     affiliation: Monash University
@@
 ```{r load-libraries}
+library(cardinalR)
 library(tidyverse)
@@
 
-> ```install.packages("cardinalR")```
+```r
+install.packages("cardinalR")
+```
 
-and from GitHub using
+and the development version can be installed from GitHub
 
-> ```remotes::install_github("jayanilakshika/cardinalR")```
-
-to install the development version.
+```r
+devtools::install_github("JayaniLakshika/cardinalR")
+```
 
@@
 
-More documentation of the package can be found at the web site https://jayanilakshika.github.io/cardinalR/.
+More documentation of the package can be found at the web site <https://jayanilakshika.github.io/cardinalR/>.
 
@@
 ### Main function
+
+```{r, echo=TRUE, eval=FALSE}
+gen_multicluster(
+  n = c(200, 300, 500), p = 4, k = 3,
+  loc = matrix(c(
+    0, 0, 0, 0,
+    5, 9, 0, 0,
+    3, 4, 10, 7
+  ), nrow = 4, byrow = TRUE),
+  dim_weights = dim4_weights,
+  scale = c(3, 1, 2),
+  shape = c("gaussian", "bluntedcorn", "unifcube"),
+  rotation = rotations_4d,
+  is_bkg = FALSE
+)
+```
 
@@
 
+### Trigonometric
+
 ### Odd shapes

### 2025-04-20 paper-cardinalR.Rmd
Added initial table for cube

@@
 
+```{r}
+cube_tb <- tibble(fun = c("gen_gridcube",
+                          "gen_unifcube",
+                          "gen_cubehole"), 
+                      exp = c("",
+                              "", 
+                              ""))
+```
+
+```{r cube-tb-html, eval=is_html_output()}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r cube-tb-pdf, eval=is_latex_output()}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ### Gaussian

### 2025-04-20 paper-cardinalR.Rmd
Added tables to different functions

@@
 
+```{r}
+polynomial_tb <- tibble(fun = c("gen_quadratic",
+                                "gen_cubic"), 
+                        exp = c("",
+                                ""))
+```
+
+```{r polynomial-tb-html, eval=is_html_output()}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r polynomial-tb-pdf, eval=is_latex_output()}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ### Pyramid
+
+```{r}
+pyramid_tb <- tibble(fun = c("gen_pyr",
+                             "gen_pyrrect",
+                             "gen_pyrtri",
+                             "gen_pyrstar",
+                             "gen_pyrholes"), 
+                        exp = c("",
+                                "",
+                                "",
+                                "",
+                                ""))
+```
+
+```{r pyramid-tb-html, eval=is_html_output()}
+pyramid_tb |> 
+  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r pyramid-tb-pdf, eval=is_latex_output()}
+pyramid_tb |> 
+  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
 
@@
 
+```{r}
+scurve_tb <- tibble(fun = c("gen_scurve",
+                            "gen_scurvehole"), 
+                        exp = c("",
+                                ""))
+```
+
+```{r scurve-tb-html, eval=is_html_output()}
+scurve_tb |> 
+  kable(caption = "cardinalR S-curve data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r scurve-tb-pdf, eval=is_latex_output()}
+scurve_tb |> 
+  kable(caption = "cardinalR S-curve data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ### Sphere
+
+```{r}
+sphere_tb <- tibble(fun = c("gen_circle",
+                            "gen_curvycycle",
+                            "gen_unifsphere",
+                            "gen_gridedsphere",
+                            "gen_clusteredspheres"), 
+                        exp = c("",
+                                "",
+                                "",
+                                "",
+                                ""))
+```
+
+```{r sphere-tb-html, eval=is_html_output()}
+sphere_tb |> 
+  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r sphere-tb-pdf, eval=is_latex_output()}
+sphere_tb |> 
+  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
 
@@
 
+```{r}
+trigonometric_tb <- tibble(fun = c("gen_crescent",
+                            "gen_curvycylinder",
+                            "gen_sphericalspiral",
+                            "gen_helicalspiral",
+                            "gen_conicspiral",
+                            "gen_nonlinear"), 
+                        exp = c("",
+                                "",
+                                "",
+                                "",
+                                "",
+                                ""))
+```
+
+```{r trigonometric-tb-html, eval=is_html_output()}
+trigonometric_tb |> 
+  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r trigonometric-tb-pdf, eval=is_latex_output()}
+trigonometric_tb |> 
+  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ### Odd shapes

### 2025-04-20 paper-cardinalR.Rmd
Updated table set up

--- 2025-04-20 paper-cardinalR.Rmd
@@
 library(tidyverse)
-library(knitr)
 library(kableExtra)
@@
 
-```{r datasets-tb-html, eval=is_html_output()}
+```{r datasets-tb-html, eval=knitr::is_html_output()}
 datasets_tb |> 
@@
 
-```{r datasets-tb-pdf, eval=is_latex_output()}
+```{r datasets-tb-pdf, eval=knitr::is_latex_output()}
 datasets_tb |> 
@@
 
-```{r branching-tb-html, eval=is_html_output()}
+```{r branching-tb-html, eval=knitr::is_html_output()}
 branch_tb |> 
@@
 
-```{r branching-tb-pdf, eval=is_latex_output()}
+```{r branching-tb-pdf, eval=knitr::is_latex_output()}
 branch_tb |> 
@@
 
+```{r}
+cube_tb <- tibble(fun = c("gen_gridcube",
+                          "gen_unifcube",
+                          "gen_cubehole"), 
+                      exp = c("",
+                              "", 
+                              ""))
+```
+
+```{r cube-tb-html, eval=knitr::is_html_output()}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r cube-tb-pdf, eval=knitr::is_latex_output()}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ### Gaussian
@@
 
+```{r}
+polynomial_tb <- tibble(fun = c("gen_quadratic",
+                                "gen_cubic"), 
+                        exp = c("",
+                                ""))
+```
+
+```{r polynomial-tb-html, eval=knitr::is_html_output()}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r polynomial-tb-pdf, eval=knitr::is_latex_output()}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ### Pyramid
+
+```{r}
+pyramid_tb <- tibble(fun = c("gen_pyr",
+                             "gen_pyrrect",
+                             "gen_pyrtri",
+                             "gen_pyrstar",
+                             "gen_pyrholes"), 
+                        exp = c("",
+                                "",
+                                "",
+                                "",
+                                ""))
+```
+
+```{r pyramid-tb-html, eval=knitr::is_html_output()}
+pyramid_tb |> 
+  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r pyramid-tb-pdf, eval=knitr::is_latex_output()}
+pyramid_tb |> 
+  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
 
@@
 
+```{r}
+scurve_tb <- tibble(fun = c("gen_scurve",
+                            "gen_scurvehole"), 
+                        exp = c("",
+                                ""))
+```
+
+```{r scurve-tb-html, eval=knitr::is_html_output()}
+scurve_tb |> 
+  kable(caption = "cardinalR S-curve data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r scurve-tb-pdf, eval=knitr::is_latex_output()}
+scurve_tb |> 
+  kable(caption = "cardinalR S-curve data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ### Sphere
+
+```{r}
+sphere_tb <- tibble(fun = c("gen_circle",
+                            "gen_curvycycle",
+                            "gen_unifsphere",
+                            "gen_gridedsphere",
+                            "gen_clusteredspheres"), 
+                        exp = c("",
+                                "",
+                                "",
+                                "",
+                                ""))
+```
+
+```{r sphere-tb-html, eval=knitr::is_html_output()}
+sphere_tb |> 
+  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r sphere-tb-pdf, eval=knitr::is_latex_output()}
+sphere_tb |> 
+  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
 
@@
 
+```{r}
+trigonometric_tb <- tibble(fun = c("gen_crescent",
+                            "gen_curvycylinder",
+                            "gen_sphericalspiral",
+                            "gen_helicalspiral",
+                            "gen_conicspiral",
+                            "gen_nonlinear"), 
+                        exp = c("",
+                                "",
+                                "",
+                                "",
+                                "",
+                                ""))
+```
+
+```{r trigonometric-tb-html, eval=knitr::is_html_output()}
+trigonometric_tb |> 
+  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r trigonometric-tb-pdf, eval=knitr::is_latex_output()}
+trigonometric_tb |> 
+  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ### Odd shapes

### 2025-04-20 paper-cardinalR.tex
Updated table set up

@@
 
-\author{by Jayani P.G. Lakshika, Dianne Cook, Paul Harrison, Michael Lydeamore, and Thiyanga S. Talagala}
+\author{by Jayani P. Gamage, Dianne Cook, Paul Harrison, Michael Lydeamore, and Thiyanga S. Talagala}
 
@@
 In next section, introduces the implementation of \texttt{cardinalR} package on CRAN and GitHub, including demonstration of the package's key functions.
-We illustrate how a clustering data structure affect the Non-linear dimension reductions in \textbf{Application} section.
+We illustrate how a clustering data structure affect the Non-linear dimension reductions in \textbf{Example} section.
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.
@@
 
-\begin{quote}
-\texttt{install.packages("cardinalR")}
-\end{quote}
+\begin{verbatim}
+install.packages("cardinalR")
+\end{verbatim}
 
-and from GitHub using
+and the development version can be installed from GitHub
 
-\begin{quote}
-\texttt{remotes::install\_github("jayanilakshika/cardinalR")}
-\end{quote}
-
-to install the development version.
+\begin{verbatim}
+devtools::install_github("JayaniLakshika/cardinalR")
+\end{verbatim}
 
@@
 The \texttt{cardinalR} package comes with several data sets that load with the package. These are described in Table \ref{tab:datasets-tb-pdf}.
+
+\begin{verbatim}
+datasets_tb |> 
+  kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+\end{verbatim}
 
@@
 
+\begin{verbatim}
+gen_multicluster(
+  n = c(200, 300, 500), p = 4, k = 3,
+  loc = matrix(c(
+    0, 0, 0, 0,
+    5, 9, 0, 0,
+    3, 4, 10, 7
+  ), nrow = 4, byrow = TRUE),
+  dim_weights = dim4_weights,
+  scale = c(3, 1, 2),
+  shape = c("gaussian", "bluntedcorn", "unifcube"),
+  rotation = rotations_4d,
+  is_bkg = FALSE
+)
+\end{verbatim}
+
 \subsubsection{Branching}\label{branching}
+
+\begin{table}
+
+\caption{\label{tab:branching-tb-html}cardinalR branching data generation functions}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+Function & Explanation\\
+\hline
+gen\_expbranches & \\
+\hline
+gen\_linearbranches & \\
+\hline
+gen\_curvybranches & \\
+\hline
+gen\_orglinearbranches & \\
+\hline
+gen\_orgcurvybranches & \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+branch_tb |> 
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+\end{verbatim}
 
@@
 
+\begin{table}
+
+\caption{\label{tab:cube-tb-html}cardinalR cube data generation functions}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+Function & Explanation\\
+\hline
+gen\_gridcube & \\
+\hline
+gen\_unifcube & \\
+\hline
+gen\_cubehole & \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+\end{verbatim}
+
+\begin{table}
+
+\caption{\label{tab:cube-tb-pdf}cardinalR cube data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_gridcube & \\
+gen\_unifcube & \\
+gen\_cubehole & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsubsection{Gaussian}\label{gaussian}
@@
 
+\begin{table}
+
+\caption{\label{tab:polynomial-tb-html}cardinalR polynomial data generation functions}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+Function & Explanation\\
+\hline
+gen\_quadratic & \\
+\hline
+gen\_cubic & \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+\end{verbatim}
+
+\begin{table}
+
+\caption{\label{tab:polynomial-tb-pdf}cardinalR polynomial data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_quadratic & \\
+gen\_cubic & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsubsection{Pyramid}\label{pyramid}
+
+\begin{table}
+
+\caption{\label{tab:pyramid-tb-html}cardinalR pyramid data generation functions}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+Function & Explanation\\
+\hline
+gen\_pyr & \\
+\hline
+gen\_pyrrect & \\
+\hline
+gen\_pyrtri & \\
+\hline
+gen\_pyrstar & \\
+\hline
+gen\_pyrholes & \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+pyramid_tb |> 
+  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+\end{verbatim}
+
+\begin{table}
+
+\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_pyr & \\
+gen\_pyrrect & \\
+gen\_pyrtri & \\
+gen\_pyrstar & \\
+gen\_pyrholes & \\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
+\begin{table}
+
+\caption{\label{tab:scurve-tb-pdf}cardinalR S-curve data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_scurve & \\
+gen\_scurvehole & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsubsection{Sphere}\label{sphere}
+
+\begin{table}
+
+\caption{\label{tab:sphere-tb-html}cardinalR sphere data generation functions}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+Function & Explanation\\
+\hline
+gen\_circle & \\
+\hline
+gen\_curvycycle & \\
+\hline
+gen\_unifsphere & \\
+\hline
+gen\_gridedsphere & \\
+\hline
+gen\_clusteredspheres & \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+sphere_tb |> 
+  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+\end{verbatim}
+
+\begin{table}
+
+\caption{\label{tab:sphere-tb-pdf}cardinalR sphere data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_circle & \\
+gen\_curvycycle & \\
+gen\_unifsphere & \\
+gen\_gridedsphere & \\
+gen\_clusteredspheres & \\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
+\subsubsection{Trigonometric}\label{trigonometric}
+
+\begin{table}
+
+\caption{\label{tab:trigonometric-tb-html}cardinalR trigonometric data generation functions}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+Function & Explanation\\
+\hline
+gen\_crescent & \\
+\hline
+gen\_curvycylinder & \\
+\hline
+gen\_sphericalspiral & \\
+\hline
+gen\_helicalspiral & \\
+\hline
+gen\_conicspiral & \\
+\hline
+gen\_nonlinear & \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+trigonometric_tb |> 
+  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+\end{verbatim}
+
+\begin{table}
+
+\caption{\label{tab:trigonometric-tb-pdf}cardinalR trigonometric data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_crescent & \\
+gen\_curvycylinder & \\
+gen\_sphericalspiral & \\
+gen\_helicalspiral & \\
+gen\_conicspiral & \\
+\addlinespace
+gen\_nonlinear & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsubsection{Odd shapes}\label{odd-shapes}
@@
 
-\section{Examples}\label{examples}
+\section{Example}\label{example}
 
@@
 
-This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.
+This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.
 
@@
 \address{%
-Jayani P.G. Lakshika\\
+Jayani P. Gamage\\
 Monash University\\%

### 2025-04-20 paper-cardinalR.Rmd
Added additional function tabel initial

@@
 
+```{r}
+add_fun_tb <- tibble(fun = c("gen_noisedims",
+                            "gen_bkgnoise",
+                            "randomize_rows",
+                            "gen_nproduct",
+                            "gen_nsum",
+                            "gen_wavydims1",
+                            "gen_wavydims2",
+                            "gen_wavydims3",
+                            "gen_rotation",
+                            "gen_globalminmax",
+                            "scale_data"), 
+                        exp = c("",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                ""))
+```
+
+```{r add-tb-html, eval=knitr::is_html_output()}
+add_fun_tb |> 
+  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r add-tb-pdf, eval=knitr::is_latex_output()}
+add_fun_tb |> 
+  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 # Example

### 2025-04-20 paper-cardinalR.tex
Added PDF version

--- 2025-04-20 paper-cardinalR.tex
@@
 
+\begin{verbatim}
+gen_multicluster(
+  n = c(200, 300, 500), p = 4, k = 3,
+  loc = matrix(c(
+    0, 0, 0, 0,
+    5, 9, 0, 0,
+    3, 4, 10, 7
+  ), nrow = 4, byrow = TRUE),
+  dim_weights = dim4_weights,
+  scale = c(3, 1, 2),
+  shape = c("gaussian", "bluntedcorn", "unifcube"),
+  rotation = rotations_4d,
+  is_bkg = FALSE
+)
+\end{verbatim}
+
 \subsubsection{Branching}\label{branching}
@@
 
+\begin{table}
+
+\caption{\label{tab:cube-tb-pdf}cardinalR cube data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_gridcube & \\
+gen\_unifcube & \\
+gen\_cubehole & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsubsection{Gaussian}\label{gaussian}
@@
 
+\begin{table}
+
+\caption{\label{tab:polynomial-tb-pdf}cardinalR polynomial data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_quadratic & \\
+gen\_cubic & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsubsection{Pyramid}\label{pyramid}
+
+\begin{table}
+
+\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_pyr & \\
+gen\_pyrrect & \\
+gen\_pyrtri & \\
+gen\_pyrstar & \\
+gen\_pyrholes & \\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
+\begin{table}
+
+\caption{\label{tab:scurve-tb-pdf}cardinalR S-curve data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_scurve & \\
+gen\_scurvehole & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsubsection{Sphere}\label{sphere}
+
+\begin{table}
+
+\caption{\label{tab:sphere-tb-pdf}cardinalR sphere data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_circle & \\
+gen\_curvycycle & \\
+gen\_unifsphere & \\
+gen\_gridedsphere & \\
+gen\_clusteredspheres & \\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
+\subsubsection{Trigonometric}\label{trigonometric}
+
+\begin{table}
+
+\caption{\label{tab:trigonometric-tb-pdf}cardinalR trigonometric data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_crescent & \\
+gen\_curvycylinder & \\
+gen\_sphericalspiral & \\
+gen\_helicalspiral & \\
+gen\_conicspiral & \\
+\addlinespace
+gen\_nonlinear & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsubsection{Odd shapes}\label{odd-shapes}
@@
 
-\section{Examples}\label{examples}
+\begin{table}
+
+\caption{\label{tab:add-tb-pdf}cardinalR additional functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_noisedims & \\
+gen\_bkgnoise & \\
+randomize\_rows & \\
+gen\_nproduct & \\
+gen\_nsum & \\
+\addlinespace
+gen\_wavydims1 & \\
+gen\_wavydims2 & \\
+gen\_wavydims3 & \\
+gen\_rotation & \\
+gen\_globalminmax & \\
+\addlinespace
+scale\_data & \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\section{Example}\label{example}
 

### 2025-04-20 paper-cardinalR.R

@@ -35,12 +35,12 @@ datasets_tb <- tibble(dt = c("mobius_clust_data",
                                ""))
 
 
-## ----datasets-tb-html, echo=knitr::is_html_output(), eval=knitr::is_html_output()----
+## ----datasets-tb-html, eval=knitr::is_html_output()---------------------------
 # datasets_tb |>
 #   kable(caption = "cardinalR data sets", col.names = c("data", "explanation"))
 
 
-## ----datasets-tb-pdf, echo=knitr::is_latex_output(), eval=knitr::is_latex_output()----
+## ----datasets-tb-pdf, eval=knitr::is_latex_output()---------------------------
 datasets_tb |> 
   kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
   column_spec(1, width = "4cm") |>
@@ -76,12 +76,12 @@ branch_tb <- tibble(fun = c("gen_expbranches",
                                ""))
 
 
-## ----branching-tb-html, echo=knitr::is_html_output()--------------------------
-branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation")) 
+## ----branching-tb-html, eval=knitr::is_html_output()--------------------------
+# branch_tb |>
+#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"))
 
 
-## ----branching-tb-pdf, echo=knitr::is_latex_output()--------------------------
+## ----branching-tb-pdf, eval=knitr::is_latex_output()--------------------------
 branch_tb |> 
   kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
   column_spec(1, width = "4cm") |>
@@ -97,12 +97,12 @@ cube_tb <- tibble(fun = c("gen_gridcube",
                               ""))
 
 
-## ----cube-tb-html, echo=knitr::is_html_output()-------------------------------
-cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation")) 
+## ----cube-tb-html, eval=knitr::is_html_output()-------------------------------
+# cube_tb |>
+#   kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"))
 
 
-## ----cube-tb-pdf, echo=knitr::is_latex_output()-------------------------------
+## ----cube-tb-pdf, eval=knitr::is_latex_output()-------------------------------
 cube_tb |> 
[…]

### 2025-04-21 paper-cardinalR.Rmd
Added plan for the application

--- 2025-04-20 paper-cardinalR.Rmd
@@
 ### Main function
+
+Add main function argument list to a table...
 
@@
 
+Add table of odd shapes..
+
 ### Additional functions
@@
 
-# Example
+# Application
 
-Add one or two datasets and evaluate how it will be useful.... (one with NLDR, one with clustering)
+1) Generate a data structure with different shapes (two non-linear, one gaussian, one linear, one cone in 5-D simplex) and fit different NLDR
+
+2) Add background noise to the same data by linking two or three clusters, then discuss how single and complete linkage affect for the solutions
 

### 2025-04-21 paper-cardinalR.Rmd
Added plan for two applications
Added sample discussion..need to write after completing the application section

--- 2025-04-20 paper-cardinalR.Rmd
@@
 
-# Example
+# Application
 
-Add one or two datasets and evaluate how it will be useful.... (one with NLDR, one with clustering)
+1) Generate a data structure with different shapes (two non-linear, one gaussian, one linear, one cone in 5-D simplex) and fit different NLDR (Then, demonstrate how this data can be used to assess the ability of different dimensionality reduction techniques (e.g., PCA, t-SNE, UMAP) to preserve the underlying structure in a lower-dimensional embedding.)
+
+2) Add noise dimensions to the same data by linking two or three clusters, then discuss how single and complete linkage affect for the solutions (refer Di's book <https://dicook.github.io/mulgar_book/8-hierarchical.html>)
 
 # Discussion
+
+The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.
+The ability to generate synthetic datasets with well-defined underlying structures, coupled with the controlled introduction of nuisance variables, provided a valuable platform for assessing the robustness of downstream unsupervised learning techniques.
+
+Our findings demonstrated that the choice of dimensionality reduction method significantly impacted the ability of hierarchical clustering to recover the true underlying clusters.
+Methods that effectively preserved the global structure of the data, as defined by our generation process, generally led to more accurate and interpretable hierarchical clustering results.
+However, the presence of nuisance variables often confounded the low-dimensional embeddings, making it more challenging for hierarchical clustering to separate truly distinct groups.
+This highlights a critical consideration in real-world data analysis, where unmeasured or latent factors can obscure the signal of interest.
+
+The hierarchical clustering analysis itself, and the choice of linkage criteria, also played a crucial role.
+Different linkage methods revealed varying degrees of sensitivity to the distorted representations caused by the nuisance variables.
+For instance, methods that prioritize compact clusters might have been more susceptible to being misled by variance introduced by nuisance factors, while others focusing on the distance between clusters might have shown more resilience.
+The dendrograms generated by hierarchical clustering provided a visual means to explore the relationships between samples and the potential influence of nuisance, although determining the optimal number of clusters remained a challenge in the presence of these confounding factors.
+
+This application underscores the utility of our data generation package as a powerful tool for controlled experimentation in unsupervised learning.
+By providing the ground truth cluster assignments and the ability to systematically manipulate data characteristics like dimensionality, geometric structure, and the presence of nuisance variables, researchers can gain a deeper understanding of the strengths and limitations of various DR and clustering algorithms.
+This controlled environment allows for a more objective evaluation than often possible with real-world datasets where the underlying structure is unknown.
+
+One limitation of this particular application was the specific type and magnitude of the nuisance variables introduced.
+Future work could explore a wider range of nuisance types (e.g., batch effects, technical noise with specific distributions) and their varying degrees of influence on different DR and clustering methodologies.
+Furthermore, investigating strategies for mitigating the impact of nuisance variables, either during the DR step or within the clustering process itself, would be a valuable extension.
+
+In conclusion, this example demonstrates the critical role of synthetic data generation in dissecting the complex interactions within unsupervised learning pipelines.
+Our package provides a flexible and controlled means to create such data, enabling researchers to systematically evaluate the performance and robustness of dimensionality reduction and clustering algorithms under well-defined conditions, including the presence of confounding factors.
+This capability contributes to a more rigorous and informed approach to high-dimensional data analysis.
 

### 2025-04-21 paper-cardinalR.tex
Added sample discussion..need to write after completing the application section

@@
 \subsubsection{Main function}\label{main-function}
+
+Add main function argument list to a table\ldots{}
 
@@
 
+Add table of odd shapes..
+
 \subsubsection{Additional functions}\label{additional-functions}
@@
 
-\section{Example}\label{example}
+\section{Application}\label{application}
 
-Add one or two datasets and evaluate how it will be useful\ldots. (one with NLDR, one with clustering)
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  Generate a data structure with different shapes (two non-linear, one gaussian, one linear, one cone in 5-D simplex) and fit different NLDR (Then, demonstrate how this data can be used to assess the ability of different dimensionality reduction techniques (e.g., PCA, t-SNE, UMAP) to preserve the underlying structure in a lower-dimensional embedding.)
+\item
+  Add noise dimensions to the same data by linking two or three clusters, then discuss how single and complete linkage affect for the solutions (refer Di's book \url{https://dicook.github.io/mulgar_book/8-hierarchical.html})
+\end{enumerate}
 
 \section{Discussion}\label{discussion}
+
+The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.
+The ability to generate synthetic datasets with well-defined underlying structures, coupled with the controlled introduction of nuisance variables, provided a valuable platform for assessing the robustness of downstream unsupervised learning techniques.
+
+Our findings demonstrated that the choice of dimensionality reduction method significantly impacted the ability of hierarchical clustering to recover the true underlying clusters.
+Methods that effectively preserved the global structure of the data, as defined by our generation process, generally led to more accurate and interpretable hierarchical clustering results.
+However, the presence of nuisance variables often confounded the low-dimensional embeddings, making it more challenging for hierarchical clustering to separate truly distinct groups.
+This highlights a critical consideration in real-world data analysis, where unmeasured or latent factors can obscure the signal of interest.
+
+The hierarchical clustering analysis itself, and the choice of linkage criteria, also played a crucial role.
+Different linkage methods revealed varying degrees of sensitivity to the distorted representations caused by the nuisance variables.
+For instance, methods that prioritize compact clusters might have been more susceptible to being misled by variance introduced by nuisance factors, while others focusing on the distance between clusters might have shown more resilience.
+The dendrograms generated by hierarchical clustering provided a visual means to explore the relationships between samples and the potential influence of nuisance, although determining the optimal number of clusters remained a challenge in the presence of these confounding factors.
+
+This application underscores the utility of our data generation package as a powerful tool for controlled experimentation in unsupervised learning.
+By providing the ground truth cluster assignments and the ability to systematically manipulate data characteristics like dimensionality, geometric structure, and the presence of nuisance variables, researchers can gain a deeper understanding of the strengths and limitations of various DR and clustering algorithms.
+This controlled environment allows for a more objective evaluation than often possible with real-world datasets where the underlying structure is unknown.
+
+One limitation of this particular application was the specific type and magnitude of the nuisance variables introduced.
+Future work could explore a wider range of nuisance types (e.g., batch effects, technical noise with specific distributions) and their varying degrees of influence on different DR and clustering methodologies.
+Furthermore, investigating strategies for mitigating the impact of nuisance variables, either during the DR step or within the clustering process itself, would be a valuable extension.
+
+In conclusion, this example demonstrates the critical role of synthetic data generation in dissecting the complex interactions within unsupervised learning pipelines.
+Our package provides a flexible and controlled means to create such data, enabling researchers to systematically evaluate the performance and robustness of dimensionality reduction and clustering algorithms under well-defined conditions, including the presence of confounding factors.
+This capability contributes to a more rigorous and informed approach to high-dimensional data analysis.
 

### 2025-04-21 paper-cardinalR.Rmd
Added intro to the application

@@
 
-1) Generate a data structure with different shapes (two non-linear, one gaussian, one linear, one cone in 5-D simplex) and fit different NLDR (Then, demonstrate how this data can be used to assess the ability of different dimensionality reduction techniques (e.g., PCA, t-SNE, UMAP) to preserve the underlying structure in a lower-dimensional embedding.)
+<!-- 1) Generate a data structure with different shapes (two non-linear, one gaussian, one linear, one cone in 5-D simplex) and fit different NLDR (Then, demonstrate how this data can be used to assess the ability of different dimensionality reduction techniques (e.g., PCA, t-SNE, UMAP) to preserve the underlying structure in a lower-dimensional embedding.) -->
 
-2) Add noise dimensions to the same data by linking two or three clusters, then discuss how single and complete linkage affect for the solutions (refer Di's book <https://dicook.github.io/mulgar_book/8-hierarchical.html>)
+<!-- 2) Add noise dimensions to the same data by linking two or three clusters, then discuss how single and complete linkage affect for the solutions (refer Di's book <https://dicook.github.io/mulgar_book/8-hierarchical.html>) -->
+
+## Assessing the performance of dimension reduction on different geomatric structures in high-dimensions
+
+We illustrate the use of the package by generating five clusters of different shapes, positioned with equal inter-cluster distances in $4\text{-}D$.
+Using this synthetic dataset, we will evaluate the performance of three popular dimensionality reduction techniques: Principal Component Analysis (PCA), t-distributed Stochastic Neighbor Embedding (tSNE), and Uniform Manifold Approximation and Projection (UMAP).
 
@@
 
-The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.
+<!-- The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.
 The ability to generate synthetic datasets with well-defined underlying structures, coupled with the controlled introduction of nuisance variables, provided a valuable platform for assessing the robustness of downstream unsupervised learning techniques.
+-->
 
-Our findings demonstrated that the choice of dimensionality reduction method significantly impacted the ability of hierarchical clustering to recover the true underlying clusters.
+<!-- Our findings demonstrated that the choice of dimensionality reduction method significantly impacted the ability of hierarchical clustering to recover the true underlying clusters.
 Methods that effectively preserved the global structure of the data, as defined by our generation process, generally led to more accurate and interpretable hierarchical clustering results.
@@
 This highlights a critical consideration in real-world data analysis, where unmeasured or latent factors can obscure the signal of interest.
+-->
 
-The hierarchical clustering analysis itself, and the choice of linkage criteria, also played a crucial role.
+<!-- The hierarchical clustering analysis itself, and the choice of linkage criteria, also played a crucial role.
 Different linkage methods revealed varying degrees of sensitivity to the distorted representations caused by the nuisance variables.
@@
 The dendrograms generated by hierarchical clustering provided a visual means to explore the relationships between samples and the potential influence of nuisance, although determining the optimal number of clusters remained a challenge in the presence of these confounding factors.
+-->
 
-This application underscores the utility of our data generation package as a powerful tool for controlled experimentation in unsupervised learning.
+<!-- This application underscores the utility of our data generation package as a powerful tool for controlled experimentation in unsupervised learning.
 By providing the ground truth cluster assignments and the ability to systematically manipulate data characteristics like dimensionality, geometric structure, and the presence of nuisance variables, researchers can gain a deeper understanding of the strengths and limitations of various DR and clustering algorithms.
 This controlled environment allows for a more objective evaluation than often possible with real-world datasets where the underlying structure is unknown.
+-->
 
-One limitation of this particular application was the specific type and magnitude of the nuisance variables introduced.
+<!-- One limitation of this particular application was the specific type and magnitude of the nuisance variables introduced.
 Future work could explore a wider range of nuisance types (e.g., batch effects, technical noise with specific distributions) and their varying degrees of influence on different DR and clustering methodologies.
 Furthermore, investigating strategies for mitigating the impact of nuisance variables, either during the DR step or within the clustering process itself, would be a valuable extension.
+-->
 
-In conclusion, this example demonstrates the critical role of synthetic data generation in dissecting the complex interactions within unsupervised learning pipelines.
+<!-- In conclusion, this example demonstrates the critical role of synthetic data generation in dissecting the complex interactions within unsupervised learning pipelines.
 Our package provides a flexible and controlled means to create such data, enabling researchers to systematically evaluate the performance and robustness of dimensionality reduction and clustering algorithms under well-defined conditions, including the presence of confounding factors.
 This capability contributes to a more rigorous and informed approach to high-dimensional data analysis.
+-->
 

### 2025-04-25 paper-cardinalR.Rmd
Add intro to the first application

@@
 library(kableExtra)
+library(geozoo)
 ```
@@
 
-## Assessing the performance of dimension reduction on different geomatric structures in high-dimensions
+## Assessing the performance of dimension reduction on different geometric structures in high-dimensions
 
 We illustrate the use of the package by generating five clusters of different shapes, positioned with equal inter-cluster distances in $4\text{-}D$.
-Using this synthetic dataset, we will evaluate the performance of three popular dimensionality reduction techniques: Principal Component Analysis (PCA), t-distributed Stochastic Neighbor Embedding (tSNE), and Uniform Manifold Approximation and Projection (UMAP).
+Using this synthetic dataset, we will evaluate the performance of six popular dimensionality reduction techniques: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 

### 2025-04-25 paper-cardinalR.Rmd
Added subtitle for the second application
Added initial setup for five clusters data

@@
 
+```
+
+```{r set-seed}
+set.seed(20240412)
 ```
@@
 
+```{r gen-five-clust-data}
+
+positions <- geozoo::simplex(p=4)$points
+
+## To generate data
+five_clusts <- gen_multicluster(n = c(450, 300, 150, 250, 350), p = 4, k = 5,
+                       loc = positions,
+                       scale = c(0.4, 0.35, 0.3, 1, 0.3),
+                       shape = c("helicalspiral", "hemisphere", "unifcube", "bluntedcone", "gaussian"),
+                       rotation = NULL,
+                       is_bkg = FALSE)
+```
+
+```{r}
+langevitour::langevitour(five_clusts[, -5])
+```
+
+```{r projections}
+
+```
+
+## Assessing single and complete linkage in the presence of noise dimensions
+
 # Discussion

### 2025-04-25 paper-cardinalR.Rmd
Initial work of adding layouts

@@
 library(geozoo)
+library(patchwork)
+```
+
+```{r}
+#| label: plot-theme
+theme_set(theme_linedraw() +
+   theme(
+     #aspect.ratio = 1,
+     plot.background = element_rect(fill = 'transparent', colour = NA),
+     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+     panel.background = element_rect(fill = 'transparent', 
+                                     colour = NA),
+     panel.grid.major = element_blank(), 
+     panel.grid.minor = element_blank(), 
+     axis.title.x = element_blank(), axis.title.y = element_blank(),
+     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+     legend.background = element_rect(fill = 'transparent', 
+                                      colour = NA),
+     legend.key = element_rect(fill = 'transparent', 
+                               colour = NA),
+     legend.position = "bottom", 
+     legend.title = element_blank(), 
+     legend.text = element_text(size=4),
+     legend.key.height = unit(0.25, 'cm'),
+     legend.key.width = unit(0.25, 'cm')
+   )
+)
+```
+
+```{r}
+interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
+  annotation_custom(grid::textGrob(label = label,
+                                   x = unit(position[1], "npc"), y = unit(position[2], "npc"),
+                                   gp = grid::gpar(cex = cex, col=col)))
+}
 ```
@@
 
+```{r layouts}
+tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds")
+
+nldr1 <- tsne_data |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#000000') +
+  interior_annotation("a", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
+umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr2 <- umap_data |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#000000') +
+  interior_annotation("b", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
+phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds")
+
+nldr3 <- phate_data |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1, colour='#000000') +
+  interior_annotation("c", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
+trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr4 <- trimap_data |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#000000') +
+  interior_annotation("d", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
+pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr5 <- pacmap_data |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#000000') +
+  interior_annotation("e", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
+pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds")
+
+nldr6 <- pca_data |>
+  ggplot(aes(x = pca1,
+             y = pca2))+
+  geom_point(alpha=0.1, size=1, colour='#000000') +
+  interior_annotation("f", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1) 
+```
+
+```{r}
+nldr1 + nldr2 + nldr3 +
+  nldr4 + nldr5 + nldr6 +
+  plot_layout(ncol = 3)
+```
+
 ## Assessing single and complete linkage in the presence of noise dimensions

### 2025-04-25 scripts/five_clusts/02_gen_embeddings.R

@@ -1,0 +1,134 @@
+library(dplyr)
+library(tibble)
+library(readr)
+library(conflicted)
+
+library(Rtsne)
+library(uwot)
+library(phateR)
+library(reticulate)
+
+set.seed(20240110)
+
+use_python("~/miniforge3/envs/pcamp_env/bin/python")
+use_condaenv("pcamp_env")
+
+
+data <- read_rds(here::here("data/five_clusts/five_clusts.rds"))
+data <- data |>
+  dplyr::select(-cluster)
+
+## tSNE (default)
+perplexity <- 30
+
+tSNE_fit <- data |>
+  dplyr::select(where(is.numeric)) |>
+  Rtsne::Rtsne(perplexity = perplexity,
+               pca = FALSE)
+
+tSNE_data <- tSNE_fit$Y |>
+  tibble::as_tibble(.name_repair = "unique")
+names(tSNE_data) <- c("tSNE1", "tSNE2")
+
+write_rds(tSNE_data, file = paste0("data/five_clusts/five_clusts_tsne_perplexity_", perplexity, ".rds"))
+
+## UMAP
+
+n_neighbors <- 15
+min_dist <- 0.1
+
+UMAP_model <- umap(data,
+                  n_neighbors = n_neighbors,
+                  min_dist = min_dist,
+                  n_components =  2,
+                  init ="spca")
+
+UMAP_data <- UMAP_model |>
+  as_tibble(.name_repair = "unique")
+
+names(UMAP_data) <- c("UMAP1", "UMAP2")
+
+## Run only once
+write_rds(UMAP_data, file = paste0("data/five_clusts/five_clusts_umap_n-neigbors_", n_neighbors, "_min-dist_", min_dist, ".rds"))
+
+## PHATE
+knn <- 5
+
+PHATE_data <- phate(data, knn = knn)
+PHATE_data <- as_tibble(PHATE_data$embedding)
+
+names(PHATE_data) <- c("PHATE1", "PHATE2")
+
+write_rds(PHATE_data, file = paste0("data/five_clusts/five_clusts_phate_knn_", knn, ".rds"))
+
+
+## TriMAP
+
+trimap <- reticulate::import("trimap")
+
+data_vector <- unlist(data)
+# Convert the vector into a matrix
+data_matrix <- matrix(data_vector, ncol = NCOL(data))
+
+n_inliers <- as.integer(12)
+n_outliers <- as.integer(4)
+n_random <- as.integer(3)
+
+# Initialize PaCMAP instance
[…]

### 2025-04-25 paper-cardinalR.Rmd
Added application with more close clusters

@@
 positions <- geozoo::simplex(p=4)$points
+positions <- positions * 0.8
 
 ## To generate data
-five_clusts <- gen_multicluster(n = c(450, 300, 150, 250, 350), p = 4, k = 5,
+five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), p = 4, k = 5,
                        loc = positions,

### 2025-04-25 scripts/five_clusts/01_data_generation.R

@@ -3,6 +3,7 @@ library(readr)
 set.seed(20240412)
 
 positions <- geozoo::simplex(p=4)$points
+positions <- positions * 0.8
 
 ## To generate data
 five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), p = 4, k = 5,

### 2025-04-25 paper-cardinalR.Rmd
Added second application data

@@
 
-```{r projections}
+```{r five-clusts-projections}
 
@@
 
+```{r gen-three-clust-data}
+
+positions <- geozoo::simplex(p=4)$points[1:3, ]
+positions <- positions * 0.8
+
+## To generate data
+three_clusts <- gen_multicluster(n = c(750, 500, 250), p = 4, k = 3,
+                       loc = positions,
+                       scale = c(1, 1, 0.3),
+                       shape = c("nonlinear", "pyrtri", "gaussian"),
+                       rotation = NULL,
+                       is_bkg = FALSE)
+
+noise_df <- gen_noisedims(n = 1500, p = 3, m = rep(0, 3), s = rep(0.1, 3))
+names(noise_df) <- paste0("x", 5:7)
+
+three_clusts <- dplyr::bind_cols(three_clusts, noise_df)
+
+three_clusts <- three_clusts |>
+  dplyr::select(c(tidyselect::starts_with("x"), "cluster"))
+
+three_clusts <- normalize_data(data = three_clusts)
+
+```
+
+```{r}
+langevitour::langevitour(three_clusts[, -8])
+```
+
+```{r three-clusts-projections}
+
+```
+
 # Discussion

### 2025-04-25 paper-cardinalR.Rmd
Started working on clustring example

@@
 library(patchwork)
+library(GGally)
+library(ggdendro)
+library(mulgar)
 ```
@@
 
+```{r}
+ggscatmat(three_clusts[, -8]) + 
+  theme_minimal() +
+  xlab("") + ylab("")
+```
+
+```{r}
+p_dist <- dist(three_clusts[, -8])
+p_hcw <- hclust(p_dist, method="ward.D2")
+p_hcs <- hclust(p_dist, method="single")
+
+p_clw <- three_clusts[, -8] |> 
+  mutate(cl = factor(cutree(p_hcw, 3))) |>
+  as.data.frame()
+p_cls <- three_clusts[, -8] |> 
+  mutate(cl = factor(cutree(p_hcs, 3))) |>
+  as.data.frame()
+
+p_w_hfly <- mulgar::hierfly(p_clw, p_hcw, scale=FALSE)
+p_s_hfly <- mulgar::hierfly(p_cls, p_hcs, scale=FALSE)
+```
+
+```{r}
+# Generate the dendrograms in 2D
+p_hcw_dd <- ggdendro::dendro_data(p_hcw)
+pw_dd <- ggplot() +
+  geom_segment(data=p_hcw_dd$segments, 
+               aes(x = x, y = y, 
+                   xend = xend, yend = yend)) + 
+  geom_point(data=p_hcw_dd$labels, aes(x=x, y=y),
+             alpha=0.8) +
+  theme_dendro()
+
+p_hcs_dd <- ggdendro::dendro_data(p_hcs)
+ps_dd <- ggplot() +
+  geom_segment(data=p_hcs_dd$segments, 
+               aes(x = x, y = y, 
+                   xend = xend, yend = yend)) + 
+  geom_point(data=p_hcs_dd$labels, aes(x=x, y=y),
+             alpha=0.8) +
+  theme_dendro()
+```
+
 # Discussion

### 2025-04-25 paper-cardinalR.Rmd
For now let's keep only one example

--- 2025-04-25 paper-cardinalR.Rmd
@@
 
-## Assessing single and complete linkage in the presence of noise dimensions
-
 # Discussion

### 2025-04-25 paper-cardinalR.Rmd
Updated additional functions table

@@
                             "gen_rotation",
-                            "gen_globalminmax",
-                            "scale_data"), 
+                            "normalize_data"), 
                         exp = c("",
-                                "",
                                 "",

### 2025-04-25 paper-cardinalR.Rmd
Updated the application

@@
 library(patchwork)
-library(GGally)
-library(ggdendro)
-library(mulgar)
+library(colorspace)
 ```
@@
 We illustrate the use of the package by generating five clusters of different shapes, positioned with equal inter-cluster distances in $4\text{-}D$.
-Using this synthetic dataset, we will evaluate the performance of six popular dimensionality reduction techniques: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
+Using this synthetic dataset, we will evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 
@@
 
-```{r}
+```{r, eval=FALSE}
 langevitour::langevitour(five_clusts[, -5])
@@
 ```{r layouts}
+clusters <- as.vector(five_clusts[, 5])
+
 tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds")

### 2025-04-26 paper-cardinalR.Rmd
Updated the final para intro
Added datasets description table

@@
 In next section, introduces the implementation of `cardinalR` package on CRAN and GitHub, including demonstration of the package's key functions.
-We illustrate how a clustering data structure affect the Non-linear dimension reductions in **Example** section.
+We illustrate how a clustering data structure affect the dimension reductions in **Application** section.
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.
@@
 ```{r}
-datasets_tb <- tibble(dt = c("mobius_clust_data", 
-                             "mobius_clust_tsne_param1", 
-                             "mobius_clust_tsne_param2",
-                             "mobius_clust_tsne_param3",
-                             "mobius_clust_umap_param1",
-                             "mobius_clust_umap_param2",
-                             "mobius_clust_umap_param3"), 
-                      text = c("",
-                               "", 
-                               "",
-                               "",
-                               "",
-                               "", 
-                               ""))
+datasets_tb <- tibble(dt = c("mobiusgau", 
+                             "mobiusgau_tsne1", 
+                             "mobiusgau_tsne2",
+                             "mobiusgau_tsne3",
+                             "mobiusgau_umap1",
+                             "mobiusgau_umap2",
+                             "mobiusgau_umap3"), 
+                      text = c("Simulated data with a Mobius and a Gaussian in 4-D space.",
+                               "The tSNE embedding with perplexity 15 for mobiusgau.", 
+                               "The tSNE embedding with perplexity 30 for mobiusgau.",
+                               "The tSNE embedding with perplexity 5 for mobiusgau.",
+                               "The UMAP embedding with number of neighbors 15 and minimum distance 0.1 for mobiusgau.",
+                               "The UMAP embedding with number of neighbors 30 and minimum distance 0.08 for mobiusgau.", 
+                               "The UMAP embedding with number of neighbors 5 and minimum distance 0.9 for mobiusgau."))
 ```

### 2025-04-26 paper-cardinalR.Rmd
Updated table 2

--- 2025-04-25 paper-cardinalR.Rmd
@@
 ```{r}
-datasets_tb <- tibble(dt = c("mobius_clust_data", 
-                             "mobius_clust_tsne_param1", 
-                             "mobius_clust_tsne_param2",
-                             "mobius_clust_tsne_param3",
-                             "mobius_clust_umap_param1",
-                             "mobius_clust_umap_param2",
-                             "mobius_clust_umap_param3"), 
-                      text = c("",
-                               "", 
-                               "",
-                               "",
-                               "",
-                               "", 
-                               ""))
+datasets_tb <- tibble(dt = c("mobiusgau", 
+                             "mobiusgau_tsne1", 
+                             "mobiusgau_tsne2",
+                             "mobiusgau_tsne3",
+                             "mobiusgau_umap1",
+                             "mobiusgau_umap2",
+                             "mobiusgau_umap3"), 
+                      text = c("Simulated data with a Mobius and a Gaussian in 4-D space.",
+                               "The tSNE embedding with perplexity $15$ for mobiusgau.", 
+                               "The tSNE embedding with perplexity $30$ for mobiusgau.",
+                               "The tSNE embedding with perplexity $5$ for mobiusgau.",
+                               "The UMAP embedding with number of neighbors $15$ and minimum distance $0.1$ for mobiusgau.",
+                               "The UMAP embedding with number of neighbors $30$ and minimum distance $0.08$ for mobiusgau.", 
+                               "The UMAP embedding with number of neighbors $5$ and minimum distance $0.9$ for mobiusgau."))
 ```

### 2025-04-26 paper-cardinalR.Rmd
Added Table 9

@@
                             "randomize_rows",
+                            "relocate_clusters",
                             "gen_nproduct",
@@
                             "normalize_data"), 
-                        exp = c("",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                ""))
+                        exp = c("Generates additional noise dimensions.",
+                                "Adds background noise.",
+                                "Randomizes the rows.",
+                                "Relocates the clusters.",
+                                "Generates a vector of positive integers whose product is approximately equal to a target value.",
+                                "Generates a vector of positive integers whose summation is approximately equal to a target value.",
+                                "Generates random noise dimensions with wavy pattern generated with theta.",
+                                "Generates random noise dimensions with wavy pattern generated with power functions.",
+                                "Generates random noise dimensions with wavy pattern generated with power and sine functions.",
+                                "Generates rotations.",
+                                "Normalizes data."))
 ```

### 2025-04-26 paper-cardinalR.Rmd
Added explanation for arguments in main function

@@
 
-Add main function argument list to a table...
+```{r}
+main_tb <- tibble(arg = c("n",
+                            "p",
+                            "k",
+                            "loc",
+                            "scale",
+                            "shape",
+                            "rotation",
+                            "is_bkg"), 
+                        exp = c("A numeric vector representing the sample sizes.",
+                                "A numeric value representing the number of dimensions.",
+                                "A numeric value representing the number of clusters.",
+                                "A numeric matrix representing the locations/centroids of clusters.",
+                                "A numeric vector representing the scaling factors of clusters.",
+                                "A character vector representing the shapes of clusters.",
+                                "A numeric list which contains plane and the corresponding angle along that plane for each cluster.",
+                                "A Boolean value representing the background noise should exist or not."))
+```
 
-```{r, echo=TRUE, eval=FALSE}
-gen_multicluster(
-  n = c(200, 300, 500), p = 4, k = 3,
-  loc = matrix(c(
-    0, 0, 0, 0,
-    5, 9, 0, 0,
-    3, 4, 10, 7
-  ), nrow = 4, byrow = TRUE),
-  dim_weights = dim4_weights,
-  scale = c(3, 1, 2),
-  shape = c("gaussian", "bluntedcorn", "unifcube"),
-  rotation = rotations_4d,
-  is_bkg = FALSE
-)
+```{r main-tb-html, eval=knitr::is_html_output()}
+main_tb |> 
+  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Explanation")) 
+```
+
+```{r main-tb-pdf, eval=knitr::is_latex_output()}
+main_tb |> 
+  kable(caption = "The main arguments for `gen_multicluster()`.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
 ```

### 2025-04-26 paper-cardinalR.Rmd
Added data generation

@@
 
-We illustrate the use of the package by generating five clusters of different shapes, positioned with equal inter-cluster distances in $4\text{-}D$.
-Using this synthetic dataset, we will evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
+This section illustrates the use of package by generating a synthetic dataset to evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 
-```{r gen-five-clust-data}
+We generated a dataset of five clusters, positioned with equal inter-cluster distances in $4\text{-}D$ space.
+The five clusters have different geometric structures and each contain different number of points.
+Specifically, the helical spiral cluster includes $2250$ points and was generated with a scale parameter of $0.4$.
+The hemisphere cluster consists of $1500$ points with a scale parameter of $0.35$.
+The uniform cube-shaped cluster contains $750$ points and uses a scale parameter of $0.3$.
+The blunted cone cluster includes $1250$ points, generated with a scale parameter of $1$.
+Finally, the Gaussian-shaped cluster contains $1750$ points and was generated with a scale parameter of $0.3$.
+
+```{r gen-five-clust-data, echo=TRUE}
 

### 2025-04-26 paper-cardinalR.tex
Added data generation

--- 2025-04-20 paper-cardinalR.tex
@@
 In next section, introduces the implementation of \texttt{cardinalR} package on CRAN and GitHub, including demonstration of the package's key functions.
-We illustrate how a clustering data structure affect the Non-linear dimension reductions in \textbf{Example} section.
+We illustrate how a clustering data structure affect the dimension reductions in \textbf{Application} section.
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.
@@
 \midrule
-mobius\_clust\_data & \\
-mobius\_clust\_tsne\_param1 & \\
-mobius\_clust\_tsne\_param2 & \\
-mobius\_clust\_tsne\_param3 & \\
-mobius\_clust\_umap\_param1 & \\
+mobiusgau & Simulated data with a Mobius and a Gaussian in 4-D space.\\
+mobiusgau\_tsne1 & The tSNE embedding with perplexity \$15\$ for mobiusgau.\\
+mobiusgau\_tsne2 & The tSNE embedding with perplexity \$30\$ for mobiusgau.\\
+mobiusgau\_tsne3 & The tSNE embedding with perplexity \$5\$ for mobiusgau.\\
+mobiusgau\_umap1 & The UMAP embedding with number of neighbors \$15\$ and minimum distance \$0.1\$ for mobiusgau.\\
 \addlinespace
-mobius\_clust\_umap\_param2 & \\
-mobius\_clust\_umap\_param3 & \\
+mobiusgau\_umap2 & The UMAP embedding with number of neighbors \$30\$ and minimum distance \$0.08\$ for mobiusgau.\\
+mobiusgau\_umap3 & The UMAP embedding with number of neighbors \$5\$ and minimum distance \$0.9\$ for mobiusgau.\\
 \bottomrule
@@
 
-\begin{verbatim}
-gen_multicluster(
-  n = c(200, 300, 500), p = 4, k = 3,
-  loc = matrix(c(
-    0, 0, 0, 0,
-    5, 9, 0, 0,
-    3, 4, 10, 7
-  ), nrow = 4, byrow = TRUE),
-  dim_weights = dim4_weights,
-  scale = c(3, 1, 2),
-  shape = c("gaussian", "bluntedcorn", "unifcube"),
-  rotation = rotations_4d,
-  is_bkg = FALSE
-)
-\end{verbatim}
+\begin{table}
+
+\caption{\label{tab:main-tb-pdf}The main arguments for `gen_multicluster()`.}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Argument & Explanation\\
+\midrule
+n & A numeric vector representing the sample sizes.\\
+p & A numeric value representing the number of dimensions.\\
+k & A numeric value representing the number of clusters.\\
+loc & A numeric matrix representing the locations/centroids of clusters.\\
+scale & A numeric vector representing the scaling factors of clusters.\\
+\addlinespace
+shape & A character vector representing the shapes of clusters.\\
+rotation & A numeric list which contains plane and the corresponding angle along that plane for each cluster.\\
+is\_bkg & A Boolean value representing the background noise should exist or not.\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 \midrule
-gen\_noisedims & \\
-gen\_bkgnoise & \\
-randomize\_rows & \\
-gen\_nproduct & \\
-gen\_nsum & \\
+gen\_noisedims & Generates additional noise dimensions.\\
+gen\_bkgnoise & Adds background noise.\\
+randomize\_rows & Randomizes the rows.\\
+relocate\_clusters & Relocates the clusters.\\
+gen\_nproduct & Generates a vector of positive integers whose product is approximately equal to a target value.\\
 \addlinespace
-gen\_wavydims1 & \\
-gen\_wavydims2 & \\
-gen\_wavydims3 & \\
-gen\_rotation & \\
-gen\_globalminmax & \\
+gen\_nsum & Generates a vector of positive integers whose summation is approximately equal to a target value.\\
+gen\_wavydims1 & Generates random noise dimensions with wavy pattern generated with theta.\\
+gen\_wavydims2 & Generates random noise dimensions with wavy pattern generated with power functions.\\
+gen\_wavydims3 & Generates random noise dimensions with wavy pattern generated with power and sine functions.\\
+gen\_rotation & Generates rotations.\\
 \addlinespace
-scale\_data & \\
+normalize\_data & Normalizes data.\\
 \bottomrule
@@
 
-\section{Example}\label{example}
+\section{Application}\label{application}
 
-Add one or two datasets and evaluate how it will be useful\ldots. (one with NLDR, one with clustering)
+\subsection{Assessing the performance of dimension reduction on different geometric structures in high-dimensions}\label{assessing-the-performance-of-dimension-reduction-on-different-geometric-structures-in-high-dimensions}
+
+We illustrate the use of the package by generating five clusters of different shapes, positioned with equal inter-cluster distances in \(4\text{-}D\).
+Using this synthetic dataset, we will evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
+
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/unnamed-chunk-13-1}
 

### 2025-04-26 paper-cardinalR.R

@@ -12,27 +12,69 @@ knitr::opts_chunk$set(
 
 
 
+## ----set-seed-----------------------------------------------------------------
+set.seed(20240412)
+
+
 ## ----load-libraries-----------------------------------------------------------
 library(cardinalR)
 library(tidyverse)
 library(kableExtra)
+library(geozoo)
+library(patchwork)
+library(colorspace)
 
 
 ## -----------------------------------------------------------------------------
-datasets_tb <- tibble(dt = c("mobius_clust_data", 
-                             "mobius_clust_tsne_param1", 
-                             "mobius_clust_tsne_param2",
-                             "mobius_clust_tsne_param3",
-                             "mobius_clust_umap_param1",
-                             "mobius_clust_umap_param2",
-                             "mobius_clust_umap_param3"), 
-                      text = c("",
-                               "", 
-                               "",
-                               "",
-                               "",
-                               "", 
-                               ""))
+#| label: plot-theme
+theme_set(theme_linedraw() +
+   theme(
+     #aspect.ratio = 1,
+     plot.background = element_rect(fill = 'transparent', colour = NA),
+     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+     panel.background = element_rect(fill = 'transparent', 
+                                     colour = NA),
+     panel.grid.major = element_blank(), 
+     panel.grid.minor = element_blank(), 
+     axis.title.x = element_blank(), axis.title.y = element_blank(),
+     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+     legend.background = element_rect(fill = 'transparent', 
+                                      colour = NA),
+     legend.key = element_rect(fill = 'transparent', 
+                               colour = NA),
+     legend.position = "bottom", 
+     legend.title = element_blank(), 
+     legend.text = element_text(size=4),
+     legend.key.height = unit(0.25, 'cm'),
+     legend.key.width = unit(0.25, 'cm')
+   )
+)
+
+
+## -----------------------------------------------------------------------------
+interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
+  annotation_custom(grid::textGrob(label = label,
+                                   x = unit(position[1], "npc"), y = unit(position[2], "npc"),
[…]

### 2025-04-26 paper-cardinalR.Rmd
Added interpretation of the application

@@
 
-```{r}
+```{r, label = "fig-nldr-layouts", fig.cap="Six Drs"}
 nldr1 + nldr2 + nldr3 +
@@
 ```
+
+UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure.
+However, PHATE reveals three non-linear clusters, even though two of them do not show non-linearity.
+UMAP, TriMAP, and PaCMAP successfully maintain the local structures of the data.
+In contrast, tSNE divides the non-linear cluster into sub-clusters.
+Also, tSNE fails to preserve the distances between the clusters.
+PCA, on the other hand, preserves the local structures of the clusters, but some clusters are incorrectly merged that should remain distinct.
 

### 2025-04-26 paper-cardinalR.Rmd
Added

@@
 main_tb |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
   column_spec(1, width = "4cm") |>
@@
                        scale = c(0.4, 0.35, 0.3, 1, 0.3),
-                       shape = c("helicalspiral", "hemisphere", "unifcube", "bluntedcone", "gaussian"),
+                       shape = c("helicalspiral", "hemisphere", "unifcube", 
+                                 "bluntedcone", "gaussian"),
                        rotation = NULL,

### 2025-04-26 paper-cardinalR.tex
Added

@@
 
-\caption{\label{tab:main-tb-pdf}The main arguments for `gen_multicluster()`.}
+\caption{\label{tab:main-tb-pdf}The main arguments for gen\_multicluster().}
 \centering
@@
 
-We illustrate the use of the package by generating five clusters of different shapes, positioned with equal inter-cluster distances in \(4\text{-}D\).
-Using this synthetic dataset, we will evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
+This section illustrates the use of package by generating a synthetic dataset to evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
 
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/unnamed-chunk-13-1}
+We generated a dataset of five clusters, positioned with equal inter-cluster distances in \(4\text{-}D\) space.
+The five clusters have different geometric structures and each contain different number of points.
+Specifically, the helical spiral cluster includes \(2250\) points and was generated with a scale parameter of \(0.4\).
+The hemisphere cluster consists of \(1500\) points with a scale parameter of \(0.35\).
+The uniform cube-shaped cluster contains \(750\) points and uses a scale parameter of \(0.3\).
+The blunted cone cluster includes \(1250\) points, generated with a scale parameter of \(1\).
+Finally, the Gaussian-shaped cluster contains \(1750\) points and was generated with a scale parameter of \(0.3\).
+
+\begin{verbatim}
+positions <- geozoo::simplex(p=4)$points
+positions <- positions * 0.8
+
+## To generate data
+five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), p = 4, k = 5,
+                       loc = positions,
+                       scale = c(0.4, 0.35, 0.3, 1, 0.3),
+                       shape = c("helicalspiral", "hemisphere", "unifcube", 
+                                 "bluntedcone", "gaussian"),
+                       rotation = NULL,
+                       is_bkg = FALSE)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six Drs}\label{fig:fig-nldr-layouts}
+\end{figure}
+
+UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure.
+However, PHATE reveals three non-linear clusters, even though two of them do not show non-linearity.
+UMAP, TriMAP, and PaCMAP successfully maintain the local structures of the data.
+In contrast, tSNE divides the non-linear cluster into sub-clusters.
+Also, tSNE fails to preserve the distances between the clusters.
+PCA, on the other hand, preserves the local structures of the clusters, but some clusters are incorrectly merged that should remain distinct.
 

### 2025-04-26 paper-cardinalR.R

@@ -115,7 +115,7 @@ main_tb <- tibble(arg = c("n",
 
 ## ----main-tb-pdf, eval=knitr::is_latex_output()-------------------------------
 main_tb |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
   column_spec(1, width = "4cm") |>
   column_spec(2, width = "8cm")
 
@@ -318,7 +318,7 @@ add_fun_tb |>
   column_spec(2, width = "8cm")
 
 
-## ----gen-five-clust-data------------------------------------------------------
+## ----gen-five-clust-data, echo=TRUE-------------------------------------------
 
 positions <- geozoo::simplex(p=4)$points
 positions <- positions * 0.8
@@ -327,7 +327,8 @@ positions <- positions * 0.8
 five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), p = 4, k = 5,
                        loc = positions,
                        scale = c(0.4, 0.35, 0.3, 1, 0.3),
-                       shape = c("helicalspiral", "hemisphere", "unifcube", "bluntedcone", "gaussian"),
+                       shape = c("helicalspiral", "hemisphere", "unifcube", 
+                                 "bluntedcone", "gaussian"),
                        rotation = NULL,
                        is_bkg = FALSE)
 
@@ -398,7 +399,7 @@ nldr6 <- pca_data |>
   theme(aspect.ratio = 1) 
 
 
-## -----------------------------------------------------------------------------
+## ----label = "fig-nldr-layouts", fig.cap="Six Drs"----------------------------
 nldr1 + nldr2 + nldr3 +
   nldr4 + nldr5 + nldr6 +
   plot_layout(ncol = 3)

### 2025-04-26 paper-cardinalR.Rmd
Added Figure title

@@
 
-```{r, label = "fig-nldr-layouts", fig.cap="Six Drs"}
+```{r, label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
 nldr1 + nldr2 + nldr3 +

### 2025-04-29 paper-cardinalR.Rmd
Updated
Added additional_functions.R

@@
 ```{r}
-interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
-  annotation_custom(grid::textGrob(label = label,
-                                   x = unit(position[1], "npc"), y = unit(position[2], "npc"),
-                                   gp = grid::gpar(cex = cex, col=col)))
-}
+#| label: import-scripts
+source("scripts/additional_functions.R")
 ```
-
 # Introduction
@@
 ### Main function
+
+The main arguments of the ```gen_multicluster()``` function are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:main-tb-html)', '\\@ref(tab:main-tb-pdf)'))`.
 

### 2025-04-29 scripts/additional_functions.R

@@ -1,0 +1,218 @@
+# Add plot label
+
+interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
+  annotation_custom(grid::textGrob(label = label,
+                                   x = unit(position[1], "npc"), y = unit(position[2], "npc"),
+                                   gp = grid::gpar(cex = cex, col=col)))
+}
+
+# Scale high-d data
+
+# Center the data by subtracting the mean of each column
+center_data <- function(data) {
+  apply(data, 2, function(col) col - mean(col))
+}
+
+# Function to scale data manually
+scale_data_manual <- function(data, type_col) {
+  # Step 1: Center the data (mean 0)
+  data_centered <- center_data(data |> select(-all_of(type_col)))
+
+  # Step 2: Calculate the standard deviation of each dimension
+  sds <- apply(data_centered, 2, sd)
+
+  # Step 3: Scale each dimension to have the range [0, 1]
+  data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
+
+  # Step 4: Scale dimensions according to their variation
+  # The dimension with the highest standard deviation is scaled to [-1, 1]
+  # Other dimensions are scaled to smaller ranges based on their standard deviations
+  max_sd <- max(sds)
+
+  # Normalize the standard deviations to get scaling factors
+  scaling_factors <- sds / max_sd
+
+  for (i in seq_along(scaling_factors)) {
+    data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
+  }
+
+  # Combine the scaled data with the 'type' column and return as a tibble
+  data_scaled <- as_tibble(data_scaled) %>%
+    mutate(!!type_col := data[[type_col]])
+
+  return(data_scaled)
+}
+
+# Get projection
+
+get_projection <- function(projection, proj_scale, scaled_data,
+                           scaled_data_model, distance_df_small_edges,
+                           axis_param) {
+
+  projection_scaled <- projection * proj_scale
+  projected <- as.matrix(scaled_data) %*% projection_scaled
+
+  projected_df <- projected |>
+    tibble::as_tibble(.name_repair = "unique") |>
+    dplyr::rename(c("proj1" = "...1",
[…]

### 2025-04-30 paper-cardinalR.Rmd
Added high-D projection for application

@@
 
-```{r five-clusts-projections}
+```{r five-clusts-projections1}
+scaled_data <- scale_data_manual(five_clusts[, -5])
 
+## First projection
+projection <- cbind(
+  c(-0.59692,0.46414,0.42436,-0.10048),
+  c(-0.54999,-0.67673,-0.03631,-0.01193))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.05))
+
+proj_obj1[["cluster"]] <- as.character(five_clusts$cluster)
+
+five_clusts_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.text = element_text(size = 10))
+
+```
+
+```{r five-clusts-projections2}
+## Second projection
+projection <- cbind(
+  c(0.34673,-0.35774,0.66272,0.27298),
+  c(0.24023,0.01170,-0.41886,0.72707))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.07))
+
+proj_obj2[["cluster"]] <- as.character(five_clusts$cluster)
+
+five_clusts_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1.2), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.text = element_text(size = 10))
+```
+
+```{r five-clusts-projections3}
+## Third projection
+projection <- cbind(
+  c(0.47388,-0.20984,-0.56560,0.41644),
+  c(-0.51124,-0.60184,0.06463,0.36627))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.05))
+
+proj_obj3[["cluster"]] <- as.character(five_clusts$cluster)
+
+five_clusts_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1.1), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.text = element_text(size = 10))
+
+```
+
+```{r, label = "fig-highd-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobiusgau` data."}
+#| fig-width: 15
+#| fig-height: 5
+five_clusts_proj1 + five_clusts_proj2 + five_clusts_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-04-30 paper-cardinalR.Rmd
Add Figure title for Figure1

@@
 
-<!-- 1) Generate a data structure with different shapes (two non-linear, one gaussian, one linear, one cone in 5-D simplex) and fit different NLDR (Then, demonstrate how this data can be used to assess the ability of different dimensionality reduction techniques (e.g., PCA, t-SNE, UMAP) to preserve the underlying structure in a lower-dimensional embedding.) -->
-
-<!-- 2) Add noise dimensions to the same data by linking two or three clusters, then discuss how single and complete linkage affect for the solutions (refer Di's book <https://dicook.github.io/mulgar_book/8-hierarchical.html>) -->
-
 ## Assessing the performance of dimension reduction on different geometric structures in high-dimensions
@@
 
-```{r, label = "fig-highd-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobiusgau` data."}
+```{r, label = "fig-highd-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobiusgau` data.
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green."}
 #| fig-width: 15
 #| fig-height: 5
+
 five_clusts_proj1 + five_clusts_proj2 + five_clusts_proj3 +

### 2025-04-30 paper-cardinalR.Rmd
Added the colors for NLDR layouts

@@
 ```{r layouts}
-clusters <- as.vector(five_clusts[, 5])
+clusters <- five_clusts |> pull(cluster)
 
-tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds")
+tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds") |>
+  mutate(cluster = clusters)
 
@@
   ggplot(aes(x = tSNE1,
-             y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#000000') +
+             y = tSNE2,
+             color = cluster))+
+  geom_point(alpha=0.1, size=1) +
   interior_annotation("a", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
+  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(aspect.ratio = 1,
+        legend.text = element_text(size = 10))
 
-umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds") |>
+  mutate(cluster = clusters)
 
@@
   ggplot(aes(x = UMAP1,
-             y = UMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#000000') +
+             y = UMAP2,
+             color = cluster))+
+  geom_point(alpha=0.1, size=1) +
   interior_annotation("b", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
+  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(aspect.ratio = 1,
+        legend.text = element_text(size = 10))
 
-phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds")
+phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds") |>
+  mutate(cluster = clusters)
 
@@
   ggplot(aes(x = PHATE1,
-             y = PHATE2))+
-  geom_point(alpha=0.1, size=1, colour='#000000') +
+             y = PHATE2,
+             color = cluster))+
+  geom_point(alpha=0.1, size=1) +
   interior_annotation("c", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
+  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(aspect.ratio = 1,
+        legend.text = element_text(size = 10))
 
-trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds") |>
+  mutate(cluster = clusters)
 
@@
   ggplot(aes(x = TriMAP1,
-             y = TriMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#000000') +
+             y = TriMAP2,
+             color = cluster))+
+  geom_point(alpha=0.1, size=1) +
   interior_annotation("d", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
+  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(aspect.ratio = 1,
+        legend.text = element_text(size = 10))
 
-pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds") |>
+  mutate(cluster = clusters)
 
@@
   ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#000000') +
+             y = PaCMAP2,
+             color = cluster))+
+  geom_point(alpha=0.1, size=1) +
   interior_annotation("e", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
+  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(aspect.ratio = 1,
+        legend.text = element_text(size = 10))
 
-pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds")
+pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds") |>
+  mutate(cluster = clusters)
 
@@
   ggplot(aes(x = pca1,
-             y = pca2))+
-  geom_point(alpha=0.1, size=1, colour='#000000') +
+             y = pca2,
+             color = cluster))+
+  geom_point(alpha=0.1, size=1) +
   interior_annotation("f", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1) 
+  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(aspect.ratio = 1,
+        legend.text = element_text(size = 10)) 
 ```
@@
 ```{r, label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
+
 nldr1 + nldr2 + nldr3 +
   nldr4 + nldr5 + nldr6 +
-  plot_layout(ncol = 3)
+  plot_layout(ncol = 3, guides = "collect")
 ```

### 2025-04-30 paper-cardinalR.Rmd
Completed trigonometric_tb

@@
 ```
+
 # Introduction
@@
 
+```{r data-expbranches}
+expbranches <- gen_expbranches(n = 1000, p = 4, k = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(expbranches)
+```
+
 ### Cone
+
+```{r data-bluntedcone}
+bluntedcone <- gen_bluntedcone(n = 1000, p = 4, h = 5, ratio = 0.5)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(bluntedcone)
+```
 
@@
                             "gen_nonlinear"), 
-                        exp = c("",
-                                "",
-                                "",
-                                "",
-                                "",
-                                ""))
+                        exp = c("Generate a crescent pattern.",
+                                "Generate a curvy cylinder.",
+                                "Generate a spherical spiral.",
+                                "Generate a helical spiral.",
+                                "Generate a conic spiral.",
+                                "Generate a nonlinear hyperbola."))
 ```

### 2025-04-30 paper-cardinalR.Rmd
Added sphere_tb

@@
                             "gen_gridedsphere",
-                            "gen_clusteredspheres"), 
-                        exp = c("",
-                                "",
-                                "",
-                                "",
-                                ""))
+                            "gen_clusteredspheres",
+                            "gen_hemisphere"), 
+                        exp = c("Generate a circle.",
+                                "Generate a curvy cell cycle.",
+                                "Generate a uniform sphere.",
+                                "Generate a grided sphere.",
+                                "Generate multiple small spheres within a big sphere.",
+                                "Generate a hemisphere."))
 ```

### 2025-04-30 paper-cardinalR.Rmd
Updated scurve_tb
Added pyramid_tb

@@
 ```{r}
-pyramid_tb <- tibble(fun = c("gen_pyr",
-                             "gen_pyrrect",
+pyramid_tb <- tibble(fun = c("gen_pyrrect",
                              "gen_pyrtri",
@@
                              "gen_pyrholes"), 
-                        exp = c("",
-                                "",
-                                "",
-                                "",
-                                ""))
+                        exp = c("Generate a pyramid with a rectangular base, with the option of a sharp or blunted apex.",
+                                "Generate a pyramid with a triangular base, with the option of a sharp or blunted apex.",
+                                "Generate a pyramid with a star-shape base, with the option of a sharp or blunted apex.",
+                                "Generate a pyramid with triangular pyramid shaped holes."))
 ```
@@
                             "gen_scurvehole"), 
-                        exp = c("",
-                                ""))
+                        exp = c("Generate a S-curve.",
+                                "Generate a S-curve with a hole."))
 ```

### 2025-04-30 paper-cardinalR.Rmd
Updated polynomial_tb
Updated cube_tb

@@
                           "gen_cubehole"), 
-                      exp = c("",
-                              "", 
-                              ""))
+                      exp = c("Generate a cube with specified grid points along each axes.",
+                              "Generate a cube with uniform points.", 
+                              "Generate a cube with a hole."))
 ```
@@
                                 "gen_cubic"), 
-                        exp = c("",
-                                ""))
+                        exp = c("Generate a quadratic pattern.",
+                                "Generate a cubic pattern."))
 ```

### 2025-04-30 paper-cardinalR.Rmd
Updated multicluster

@@
                              "gen_orgcurvybranches"), 
-                      exp = c("",
-                               "", 
-                               "",
-                               "",
-                               ""))
+                      exp = c("Generate a structure with exponential shaped branches.",
+                               "Generate a structure with linear shaped branches.", 
+                               "Generate a structure with curvy shaped branches.",
+                               "Generate a structure with linear shaped branches originated in one point.",
+                               "Generate a structure with curvy shaped branches originated in one point."))
 ```
@@
 ```{r data-bluntedcone}
-bluntedcone <- gen_bluntedcone(n = 1000, p = 4, h = 5, ratio = 0.5)
+cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
 ```
@@
 ```{r, eval=FALSE}
-langevitour::langevitour(bluntedcone)
+langevitour::langevitour(cone)
 ```
@@
                        shape = c("helicalspiral", "hemisphere", "unifcube", 
-                                 "bluntedcone", "gaussian"),
+                                 "cone", "gaussian"),
                        rotation = NULL,

### 2025-04-30 paper-cardinalR.Rmd
Add intro to main function

@@
 
-The main arguments of the ```gen_multicluster()``` function are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:main-tb-html)', '\\@ref(tab:main-tb-pdf)'))`.
+The main function of the package is `gen_multicluster()`.
+This function generates clusters of various shapes, allowing users to specify the number of samples in each cluster, as well as their locations, scaling, and rotations across specific dimensions.
+Additionally, users can add background noise into the generated data by using the `is_bkg` option.
+
+The main arguments of the `gen_multicluster()` function are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:main-tb-html)', '\\@ref(tab:main-tb-pdf)'))`.
 

### 2025-04-30 paper-cardinalR.Rmd
Initial work on multiple cluster examples

@@
 The main function of the package is `gen_multicluster()`.
-This function generates clusters of various shapes, allowing users to specify the number of samples in each cluster, as well as their locations, scaling, and rotations across specific dimensions.
+This function generates clusters of various shapes, allowing users to specify the number of points in each cluster, as well as their locations, scaling, and rotations across specific dimensions.
 Additionally, users can add background noise into the generated data by using the `is_bkg` option.
@@
                             "is_bkg"), 
-                        exp = c("A numeric vector representing the sample sizes.",
+                        exp = c("A numeric vector representing the number of points in each cluster.",
                                 "A numeric value representing the number of dimensions.",
@@
 
-### Odd shapes
+### Multiple cluster examples
 
-Add table of odd shapes..
+```{r}
+odd_shapes_tb <- tibble(fun = c("make_mobiusgau",
+                                "make_multigau"), 
+                        exp = c("",
+                                ""))
+```
+
+```{r odd-shape-tb-html, eval=knitr::is_html_output()}
+odd_shapes_tb |> 
+  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r odd-shape-tb-pdf, eval=knitr::is_latex_output()}
+odd_shapes_tb |> 
+  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
 

### 2025-04-30 paper-cardinalR.Rmd
Add explanation for Multiple cluster examples
Added main arguments for branching shape generators

@@
 
+```{r}
+arg_branching_tb <- tibble(arg = c("n",
+                          "p",
+                          "k"), 
+                        exp = c("A numeric value representing the number of points.",
+                                "A numeric value representing the number of dimensions.",
+                                "A numeric value representing the number of clusters."))
+```
+
+```{r arg-branching-tb-html, eval=knitr::is_html_output()}
+arg_branching_tb |> 
+  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Explanation")) 
+```
+
+```{r arg-branching-tb-pdf, eval=knitr::is_latex_output()}
+arg_branching_tb |> 
+  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 ```{r data-expbranches}
@@
 ### Multiple cluster examples
+
+By using the shape generators mentioned above, we can create various examples of multiple clusters.
+The package includes some of these examples, which are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:odd-shape-tb-html)', '\\@ref(tab:odd-shape-tb-pdf)'))`.
 

### 2025-04-30 paper-cardinalR.Rmd
Add refercing to the branching functions

@@
 ### Branching
+
+The
+
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:branching-tb-html)', '\\@ref(tab:branching-tb-pdf)'))` outlines these functions.
+The main arguments of the functions described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:arg-branching-tb-html)', '\\@ref(tab:arg-branching-tb-pdf)'))`.
 

### 2025-04-30 paper-cardinalR.Rmd
Add intro paragraph for branching

@@
 
-The
-
+A branching structure captures trajectories that diverge or bifurcate from a common origin, resembling processes such as cell differentiation in biology.
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:branching-tb-html)', '\\@ref(tab:branching-tb-pdf)'))` outlines these functions.

### 2025-04-30 paper-cardinalR.Rmd
Fixed the issue fo Figure 2

@@
 ```{r layouts}
-clusters <- five_clusts |> pull(cluster)
 
-tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds") |>
-  mutate(cluster = clusters)
+tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds")
 
@@
 
-umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds") |>
-  mutate(cluster = clusters)
+umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds")
 
@@
 
-phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds") |>
-  mutate(cluster = clusters)
+phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds")
 
@@
 
-trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds") |>
-  mutate(cluster = clusters)
+trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
 
@@
 
-pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds") |>
-  mutate(cluster = clusters)
+pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
 
@@
 
-pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds") |>
-  mutate(cluster = clusters)
+pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds")
 
@@
 # Discussion
+
+- Branching: These functions create a controlled environment for testing how effectively various algorithms preserve branching topology and continuity in their low-dimensional embeddings.
 

### 2025-04-30 scripts/five_clusts/02_gen_embeddings.R

@@ -15,6 +15,7 @@ use_condaenv("pcamp_env")
 
 
 data <- read_rds(here::here("data/five_clusts/five_clusts.rds"))
+data_n <- data
 data <- data |>
   dplyr::select(-cluster)
 
@@ -29,6 +30,9 @@ tSNE_fit <- data |>
 tSNE_data <- tSNE_fit$Y |>
   tibble::as_tibble(.name_repair = "unique")
 names(tSNE_data) <- c("tSNE1", "tSNE2")
+
+tSNE_data <- tSNE_data |>
+  dplyr::mutate(cluster = data_n$cluster)
 
 write_rds(tSNE_data, file = paste0("data/five_clusts/five_clusts_tsne_perplexity_", perplexity, ".rds"))
 
@@ -48,6 +52,9 @@ UMAP_data <- UMAP_model |>
 
 names(UMAP_data) <- c("UMAP1", "UMAP2")
 
+UMAP_data <- UMAP_data |>
+  dplyr::mutate(cluster = data_n$cluster)
+
 ## Run only once
 write_rds(UMAP_data, file = paste0("data/five_clusts/five_clusts_umap_n-neigbors_", n_neighbors, "_min-dist_", min_dist, ".rds"))
 
@@ -58,6 +65,9 @@ PHATE_data <- phate(data, knn = knn)
 PHATE_data <- as_tibble(PHATE_data$embedding)
 
 names(PHATE_data) <- c("PHATE1", "PHATE2")
+
+PHATE_data <- PHATE_data |>
+  dplyr::mutate(cluster = data_n$cluster)
 
 write_rds(PHATE_data, file = paste0("data/five_clusts/five_clusts_phate_knn_", knn, ".rds"))
 
@@ -85,6 +95,9 @@ TriMAP_data <- reducer$fit_transform(data_matrix) |>
   as_tibble()
 
 names(TriMAP_data) <- c("TriMAP1", "TriMAP2")
+
+TriMAP_data <- TriMAP_data |>
+  dplyr::mutate(cluster = data_n$cluster)
 
 write_rds(TriMAP_data, file = paste0("data/five_clusts/five_clusts_trimap_n-inliers_", n_inliers, "_n-outliers_", n_outliers, "_n-random_", n_random, ".rds"))
 
@@ -114,6 +127,9 @@ PacMAP_data <- reducer$fit_transform(data_matrix, init = init) |>
 
[…]

### 2025-04-30 scripts/three_clusts/01_data_generation.R

@@ -1,29 +1,0 @@
-library(cardinalR)
-library(readr)
-set.seed(20240412)
-
-positions <- geozoo::simplex(p=4)$points[1:3, ]
-positions <- positions * 0.8
-
-## To generate data
-three_clusts <- gen_multicluster(n = c(750, 500, 250), p = 4, k = 3,
-                       loc = positions,
-                       scale = c(1, 1, 0.3),
-                       shape = c("nonlinear", "pyrtri", "gaussian"),
-                       rotation = NULL,
-                       is_bkg = FALSE)
-
-noise_df <- gen_noisedims(n = 1500, p = 3, m = rep(0, 3), s = rep(0.1, 3))
-names(noise_df) <- paste0("x", 5:7)
-
-three_clusts <- dplyr::bind_cols(three_clusts, noise_df)
-
-three_clusts <- three_clusts |>
-  dplyr::select(c(tidyselect::starts_with("x"), "cluster"))
-
-three_clusts <- normalize_data(data = three_clusts)
-
-langevitour::langevitour(three_clusts[, -8])
-
-## Run only once
-write_rds(three_clusts, "data/three_clusts/three_clusts.rds")

### 2025-04-30 paper-cardinalR.Rmd
Add figure references for the application

@@
 
-We generated a dataset of five clusters, positioned with equal inter-cluster distances in $4\text{-}D$ space.
-The five clusters have different geometric structures and each contain different number of points.
-Specifically, the helical spiral cluster includes $2250$ points and was generated with a scale parameter of $0.4$.
-The hemisphere cluster consists of $1500$ points with a scale parameter of $0.35$.
-The uniform cube-shaped cluster contains $750$ points and uses a scale parameter of $0.3$.
-The blunted cone cluster includes $1250$ points, generated with a scale parameter of $1$.
-Finally, the Gaussian-shaped cluster contains $1750$ points and was generated with a scale parameter of $0.3$.
+The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in $4\text{-}D$ space (Figure \@ref(fig:fig-highd-proj)).
 
@@
 
+The five clusters have different geometric structures and each contain different number of points.
+Specifically, the helical spiral cluster includes $2250$ points and was generated with a scale parameter of $0.4$.
+The hemisphere cluster consists of $1500$ points with a scale parameter of $0.35$.
+The uniform cube-shaped cluster contains $750$ points and uses a scale parameter of $0.3$.
+The blunted cone cluster includes $1250$ points, generated with a scale parameter of $1$.
+Finally, the Gaussian-shaped cluster contains $1750$ points and was generated with a scale parameter of $0.3$.
+
 ```{r layouts}
@@
 
-UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure.
+UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure (Figure \@ref(fig:fig-nldr-layouts)).
 However, PHATE reveals three non-linear clusters, even though two of them do not show non-linearity.

### 2025-04-30 paper-cardinalR.Rmd
Add additional function intro

@@
 
+The package includes various supplementary tools in addition to the shape-generating functions mentioned earlier.
+These tools allow users to generate noise dimensions with a normal distribution and various wavy patterns, create background noise, randomize the rows of the data, reposition clusters, generate a vector whose product and sum are approximately equal to a target value, rotate structures, and normalize the data.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:add-tb-html)', '\\@ref(tab:add-tb-pdf)'))` details these functions.
+
 ```{r}

### 2025-04-30 paper-cardinalR.Rmd
Started adding projections for cone data

@@
 
-```{r data-bluntedcone}
+```{r data-bluntedcone, echo=TRUE}
 cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
@@
 langevitour::langevitour(cone)
+```
+
+```{r cone-proj1}
+scaled_data <- scale_data_manual(cone)
+
+## First projection
+projection <- cbind(
+  c(0.22417,-0.02943,-0.02870,0.08418),
+  c(-0.02715,0.18019,0.05053,0.15254))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.2,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.016))
+
+cone_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.25), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = FALSE)
+
+```
+
+```{r cone-proj2}
+
+## Second projection
+projection <- cbind(
+  c(0.22417,-0.02943,-0.02870,0.08418),
+  c(-0.02715,0.18019,0.05053,0.15254))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.2,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.016))
+
+cone_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.15), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = FALSE)
+
+```
+
+```{r cone-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.22417,-0.02943,-0.02870,0.08418),
+  c(-0.02715,0.18019,0.05053,0.15254))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.2,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.016))
+
+cone_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.15), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-cone-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+cone_proj1 + cone_proj2 + cone_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-05-01 paper-cardinalR.Rmd
Updated cone projection 1

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.2,
+                            axis_param = list(limits = 0.25,
                                               axis_scaled = 3, 
@@
                                               axis_pos_y = -0.2, 
-                                              threshold = 0.016))
+                                              threshold = 0.022))
 
@@
   cex = 2, 
-  axis_text_size = 4,
+  axis_text_size = 5,
   is_color = FALSE)

### 2025-05-01 paper-cardinalR.Rmd
Added projections for cone data

--- 2025-04-30 paper-cardinalR.Rmd
@@
 langevitour::langevitour(cone)
+```
+
+```{r cone-proj1}
+scaled_data <- scale_data_manual(cone)
+
+## First projection
+projection <- cbind(
+  c(0.22417,-0.02943,-0.02870,0.08418),
+  c(-0.02715,0.18019,0.05053,0.15254))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.25,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.022))
+
+cone_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.25), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cone-proj2}
+
+## Second projection
+projection <- cbind(
+    c(0.15591,0.17907,0.03198,0.04496),
+    c(-0.15376,0.10435,0.15766,0.00547))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.1,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.08, 
+                                              axis_pos_y = -0.08, 
+                                              threshold = 0.01))
+
+cone_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.1, 0.1), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cone-proj3}
+
+## Third projection
+projection <- cbind(
+    c(-0.13548,0.13832,0.05991,-0.13544),
+    c(0.08479,-0.10719,-0.01572,-0.20123))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.25,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.016))
+
+cone_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.2), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-cone-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+cone_proj1 + cone_proj2 + cone_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-05-01 paper-cardinalR.Rmd
Added initial mobius projections

@@
 ### Mobius
+
+```{r data-mobius, echo=TRUE}
+mobius <- gen_mobius(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(mobius)
+```
+
+```{r mobius-proj1}
+scaled_data <- scale_data_manual(mobius)
+
+## First projection
+projection <- cbind(
+  c(0.22417,-0.02943,-0.02870,0.08418),
+  c(-0.02715,0.18019,0.05053,0.15254))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.25,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.022))
+
+mobius_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.25), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r mobius-proj2}
+
+## Second projection
+projection <- cbind(
+  c(0.15591,0.17907,0.03198,0.04496),
+  c(-0.15376,0.10435,0.15766,0.00547))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.1,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.08, 
+                                              axis_pos_y = -0.08, 
+                                              threshold = 0.01))
+
+mobius_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.1, 0.1), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r mobius-proj3}
+
+## Third projection
+projection <- cbind(
+  c(-0.13548,0.13832,0.05991,-0.13544),
+  c(0.08479,-0.10719,-0.01572,-0.20123))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.25,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.016))
+
+mobius_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.2), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-mobius-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+mobius_proj1 + mobius_proj2 + mobius_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
 

### 2025-05-01 paper-cardinalR.tex
Added initial mobius projections

@@
 
+The main function of the package is \texttt{gen\_multicluster()}.
+This function generates clusters of various shapes, allowing users to specify the number of points in each cluster, as well as their locations, scaling, and rotations across specific dimensions.
+Additionally, users can add background noise into the generated data by using the \texttt{is\_bkg} option.
+
+The main arguments of the \texttt{gen\_multicluster()} function are shown in Table \ref{tab:main-tb-pdf}.
+
 \begin{table}
@@
 \midrule
-n & A numeric vector representing the sample sizes.\\
+n & A numeric vector representing the number of points in each cluster.\\
 p & A numeric value representing the number of dimensions.\\
@@
 
+A branching structure captures trajectories that diverge or bifurcate from a common origin, resembling processes such as cell differentiation in biology.
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
+Table \ref{tab:branching-tb-pdf} outlines these functions.
+The main arguments of the functions described in Table \ref{tab:arg-branching-tb-pdf}.
+
 \begin{table}
@@
 \midrule
-gen\_expbranches & \\
-gen\_linearbranches & \\
-gen\_curvybranches & \\
-gen\_orglinearbranches & \\
-gen\_orgcurvybranches & \\
+gen\_expbranches & Generate a structure with exponential shaped branches.\\
+gen\_linearbranches & Generate a structure with linear shaped branches.\\
+gen\_curvybranches & Generate a structure with curvy shaped branches.\\
+gen\_orglinearbranches & Generate a structure with linear shaped branches originated in one point.\\
+gen\_orgcurvybranches & Generate a structure with curvy shaped branches originated in one point.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:arg-branching-tb-pdf}The main arguments for branching shape generators.}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Argument & Explanation\\
+\midrule
+n & A numeric value representing the number of points.\\
+p & A numeric value representing the number of dimensions.\\
+k & A numeric value representing the number of clusters.\\
 \bottomrule
@@
 \subsubsection{Cone}\label{cone}
+
+\begin{verbatim}
+cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:fig-cone-proj}
+\end{figure}
 
@@
 \midrule
-gen\_gridcube & \\
-gen\_unifcube & \\
-gen\_cubehole & \\
+gen\_gridcube & Generate a cube with specified grid points along each axes.\\
+gen\_unifcube & Generate a cube with uniform points.\\
+gen\_cubehole & Generate a cube with a hole.\\
 \bottomrule
@@
 \midrule
-gen\_quadratic & \\
-gen\_cubic & \\
+gen\_quadratic & Generate a quadratic pattern.\\
+gen\_cubic & Generate a cubic pattern.\\
 \bottomrule
@@
 \midrule
-gen\_pyr & \\
-gen\_pyrrect & \\
-gen\_pyrtri & \\
-gen\_pyrstar & \\
-gen\_pyrholes & \\
+gen\_pyrrect & Generate a pyramid with a rectangular base, with the option of a sharp or blunted apex.\\
+gen\_pyrtri & Generate a pyramid with a triangular base, with the option of a sharp or blunted apex.\\
+gen\_pyrstar & Generate a pyramid with a star-shape base, with the option of a sharp or blunted apex.\\
+gen\_pyrholes & Generate a pyramid with triangular pyramid shaped holes.\\
 \bottomrule
@@
 \midrule
-gen\_scurve & \\
-gen\_scurvehole & \\
+gen\_scurve & Generate a S-curve.\\
+gen\_scurvehole & Generate a S-curve with a hole.\\
 \bottomrule
@@
 \midrule
-gen\_circle & \\
-gen\_curvycycle & \\
-gen\_unifsphere & \\
-gen\_gridedsphere & \\
-gen\_clusteredspheres & \\
+gen\_circle & Generate a circle.\\
+gen\_curvycycle & Generate a curvy cell cycle.\\
+gen\_unifsphere & Generate a uniform sphere.\\
+gen\_gridedsphere & Generate a grided sphere.\\
+gen\_clusteredspheres & Generate multiple small spheres within a big sphere.\\
+\addlinespace
+gen\_hemisphere & Generate a hemisphere.\\
 \bottomrule
@@
 \midrule
-gen\_crescent & \\
-gen\_curvycylinder & \\
-gen\_sphericalspiral & \\
-gen\_helicalspiral & \\
-gen\_conicspiral & \\
+gen\_crescent & Generate a crescent pattern.\\
+gen\_curvycylinder & Generate a curvy cylinder.\\
+gen\_sphericalspiral & Generate a spherical spiral.\\
+gen\_helicalspiral & Generate a helical spiral.\\
+gen\_conicspiral & Generate a conic spiral.\\
 \addlinespace
-gen\_nonlinear & \\
+gen\_nonlinear & Generate a nonlinear hyperbola.\\
 \bottomrule
@@
 
-\subsubsection{Odd shapes}\label{odd-shapes}
+\subsubsection{Multiple cluster examples}\label{multiple-cluster-examples}
 
-Add table of odd shapes..
+By using the shape generators mentioned above, we can create various examples of multiple clusters.
+The package includes some of these examples, which are described in Table \ref{tab:odd-shape-tb-pdf}.
+
+\begin{table}
+
+\caption{\label{tab:odd-shape-tb-pdf}cardinalR multiple clusters generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+make\_mobiusgau & \\
+make\_multigau & \\
+\bottomrule
+\end{tabular}
+\end{table}
 
 \subsubsection{Additional functions}\label{additional-functions}
+
+The package includes various supplementary tools in addition to the shape-generating functions mentioned earlier.
+These tools allow users to generate noise dimensions with a normal distribution and various wavy patterns, create background noise, randomize the rows of the data, reposition clusters, generate a vector whose product and sum are approximately equal to a target value, rotate structures, and normalize the data.
+Table \ref{tab:add-tb-pdf} details these functions.
 
@@
 
-We generated a dataset of five clusters, positioned with equal inter-cluster distances in \(4\text{-}D\) space.
+The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in \(4\text{-}D\) space (Figure \ref{fig:fig-highd-proj}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `mobiusgau` data.
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:fig-highd-proj}
+\end{figure}
+
 The five clusters have different geometric structures and each contain different number of points.
@@
 
-\begin{verbatim}
-positions <- geozoo::simplex(p=4)$points
-positions <- positions * 0.8
-
-## To generate data
-five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), p = 4, k = 5,
-                       loc = positions,
-                       scale = c(0.4, 0.35, 0.3, 1, 0.3),
-                       shape = c("helicalspiral", "hemisphere", "unifcube", 
-                                 "bluntedcone", "gaussian"),
-                       rotation = NULL,
-                       is_bkg = FALSE)
-\end{verbatim}
-
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six Drs}\label{fig:fig-nldr-layouts}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
 \end{figure}
 
-UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure.
+UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure (Figure \ref{fig:fig-nldr-layouts}).
 However, PHATE reveals three non-linear clusters, even though two of them do not show non-linearity.
@@
 \section{Discussion}\label{discussion}
+
+\begin{itemize}
+\tightlist
+\item
+  Branching: These functions create a controlled environment for testing how effectively various algorithms preserve branching topology and continuity in their low-dimensional embeddings.
+\end{itemize}
 

### 2025-05-01 paper-cardinalR.Rmd
Added mobius projection1

@@
 projection <- cbind(
-  c(0.22417,-0.02943,-0.02870,0.08418),
-  c(-0.02715,0.18019,0.05053,0.15254))
+    c(0.53506,0.00763,0.00097,-0.10901),
+    c(0.00896,0.52496,0.12597,0.08184))
 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.25,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.25), 
+  plot_limits = c(-0.65, 0.7), 
   title = "a1", 
@@
 projection <- cbind(
-  c(0.15591,0.17907,0.03198,0.04496),
-  c(-0.15376,0.10435,0.15766,0.00547))
+    c(-0.06614,0.15134,-0.44278,0.27367),
+    c(-0.50493,0.14388,0.14615,0.03486))
 
@@
 projection <- cbind(
-  c(-0.13548,0.13832,0.05991,-0.13544),
-  c(0.08479,-0.10719,-0.01572,-0.20123))
+    c(0.10704,0.04822,-0.03416,0.53224),
+    c(0.50329,0.07199,-0.16064,-0.11806))
 

### 2025-05-01 paper-cardinalR.Rmd
Added mobius second projection

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.1,
+                            axis_param = list(limits = 0.7,
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.08, 
-                                              axis_pos_y = -0.08, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
                                               threshold = 0.01))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.1, 0.1), 
+  plot_limits = c(-0.7, 0.6), 
   title = "a2", 

### 2025-05-01 paper-cardinalR.Rmd
Added third projection for mobius data

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.25,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.2), 
+  plot_limits = c(-0.5, 0.62), 
   title = "a3", 

### 2025-05-01 paper-cardinalR.Rmd
Added initial projection for linear data

@@
 
+```{r data-linear, echo=TRUE}
+linear <- gen_longlinear(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(linear)
+```
+
+```{r linear-proj1}
+scaled_data <- scale_data_manual(linear)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+linear_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linear-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+linear_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linear-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+linear_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-linear-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linear` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+linear_proj1 + linear_proj2 + linear_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Mobius

### 2025-05-01 paper-cardinalR.Rmd
Added first linear projection

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.8,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.8, 0.8), 
   title = "a1", 

### 2025-05-01 paper-cardinalR.Rmd
Added linear projections

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.9,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.02))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.9, 0.9), 
   title = "a3", 

### 2025-05-01 paper-cardinalR.Rmd
Added initial projection for Gaussian data

@@
 
+```{r data-gau, echo=TRUE}
+gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(gau)
+```
+
+```{r gau-proj1}
+scaled_data <- scale_data_manual(gau)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+gau_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gau-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+gau_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gau-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+gau_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-gau-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gau` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+gau_proj1 + gau_proj2 + gau_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Linear

### 2025-05-01 paper-cardinalR.Rmd
Updated Gau projections

@@
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
                                               threshold = 0.02))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.9), 
+  plot_limits = c(-1, 0.9), 
   title = "a3", 

### 2025-05-01 paper-cardinalR.Rmd
Added initial projection of swiss roll

@@
 
+```{r data-swissroll, echo=TRUE}
+swissroll <- gen_swissroll(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(swissroll)
+```
+
+```{r swissroll-proj1}
+scaled_data <- scale_data_manual(swissroll)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+swissroll_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r swissroll-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+swissroll_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r swissroll-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+swissroll_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-swissroll-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `swissroll` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+swissroll_proj1 + swissroll_proj2 + swissroll_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Trigonometric

### 2025-05-01 paper-cardinalR.Rmd
Added all parameters for swissroll
Added initial trigonometric projections

@@
 ```{r data-swissroll, echo=TRUE}
-swissroll <- gen_swissroll(n = 1000, p = 4)
+swissroll <- gen_swissroll(n = 1000, p = 4, w = c(-1, 1))
 ```
@@
   column_spec(2, width = "8cm")
+```
+
+```{r data-crescent, echo=TRUE}
+crescent <- gen_crescent(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(crescent)
+```
+
+```{r crescent-proj1}
+scaled_data <- scale_data_manual(crescent)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+crescent_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r crescent-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+crescent_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r crescent-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+crescent_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-crescent-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+crescent_proj1 + crescent_proj2 + crescent_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-curvycylinder, echo=TRUE}
+curvycylinder <- gen_curvycylinder(n = 1000, p = 4, h = 10)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(curvycylinder)
+```
+
+```{r curvycylinder-proj1}
+scaled_data <- scale_data_manual(curvycylinder)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+curvycylinder_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycylinder-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+curvycylinder_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycylinder-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+curvycylinder_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-curvycylinder-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-sphericalspiral, echo=TRUE}
+sphericalspiral <- gen_sphericalspiral(n = 1000, p = 4, spins = 1)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(sphericalspiral)
+```
+
+```{r sphericalspiral-proj1}
+scaled_data <- scale_data_manual(sphericalspiral)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+sphericalspiral_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r sphericalspiral-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+sphericalspiral_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r sphericalspiral-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+sphericalspiral_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-sphericalspiral-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `sphericalspiral` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-helicalspiral, echo=TRUE}
+helicalspiral <- gen_helicalspiral(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(helicalspiral)
+```
+
+```{r helicalspiral-proj1}
+scaled_data <- scale_data_manual(helicalspiral)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+helicalspiral_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r helicalspiral-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+helicalspiral_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r helicalspiral-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+helicalspiral_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-helicalspiral-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `helicalspiral` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+helicalspiral_proj1 + helicalspiral_proj2 + helicalspiral_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-conicspiral, echo=TRUE}
+conicspiral <- gen_conicspiral(n = 1000, p = 4, spins = 1)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(conicspiral)
+```
+
+```{r conicspiral-proj1}
+scaled_data <- scale_data_manual(conicspiral)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+conicspiral_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r conicspiral-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+conicspiral_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r conicspiral-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+conicspiral_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-conicspiral-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `conicspiral` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+conicspiral_proj1 + conicspiral_proj2 + conicspiral_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-nonlinear, echo=TRUE}
+nonlinear <- gen_nonlinear(n = 1000, p = 4, hc = 1, non_fac = 0.5)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(nonlinear)
+```
+
+```{r nonlinear-proj1}
+scaled_data <- scale_data_manual(nonlinear)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+nonlinear_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r nonlinear-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+nonlinear_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r nonlinear-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+nonlinear_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-nonlinear-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `nonlinear` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+nonlinear_proj1 + nonlinear_proj2 + nonlinear_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-05-01 paper-cardinalR.Rmd
Added projection for S-curve data

@@
 
+```{r data-scurve, echo=TRUE}
+scurve <- gen_scurve(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(scurve)
+```
+
+```{r scurve-proj1}
+scaled_data <- scale_data_manual(scurve)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+scurve_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r scurve-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+scurve_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r scurve-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+scurve_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-scurve-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+scurve_proj1 + scurve_proj2 + scurve_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-scurvehole, echo=TRUE}
+scurvehole <- gen_scurvehole(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(scurvehole)
+```
+
+```{r scurvehole-proj1}
+scaled_data <- scale_data_manual(scurvehole)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+scurvehole_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r scurvehole-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+scurvehole_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r scurvehole-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+scurvehole_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-scurvehole-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurvehole` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+scurvehole_proj1 + scurvehole_proj2 + scurvehole_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Sphere

### 2025-05-01 paper-cardinalR.Rmd
Added initial sphere projections

@@
 
+```{r data-circle, echo=TRUE}
+circle <- gen_circle(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(circle)
+```
+
+```{r circle-proj1}
+scaled_data <- scale_data_manual(circle)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+circle_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r circle-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+circle_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r circle-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+circle_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-circle-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+circle_proj1 + circle_proj2 + circle_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-curvycycle, echo=TRUE}
+curvycycle <- gen_curvycycle(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(curvycycle)
+```
+
+```{r curvycycle-proj1}
+scaled_data <- scale_data_manual(curvycycle)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+curvycycle_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycycle-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+curvycycle_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycycle-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+curvycycle_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-curvycycle-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycycle` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-unifsphere, echo=TRUE}
+unifsphere <- gen_unifsphere(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(unifsphere)
+```
+
+```{r unifsphere-proj1}
+scaled_data <- scale_data_manual(unifsphere)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+unifsphere_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifsphere-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+unifsphere_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifsphere-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+unifsphere_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-unifsphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `unifsphere` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+unifsphere_proj1 + unifsphere_proj2 + unifsphere_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-gridedsphere, echo=TRUE}
+gridedsphere <- gen_gridedsphere(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(gridedsphere)
+```
+
+```{r gridedsphere-proj1}
+scaled_data <- scale_data_manual(gridedsphere)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+gridedsphere_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridedsphere-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+gridedsphere_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridedsphere-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+gridedsphere_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-gridedsphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridedsphere` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+gridedsphere_proj1 + gridedsphere_proj2 + gridedsphere_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-clusteredspheres, echo=TRUE}
+clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, p = 4, r = c(15, 3),
+                                         loc = 10 / sqrt(3)) |>
+  dplyr::select(-cluster)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(clusteredspheres)
+```
+
+```{r clusteredspheres-proj1}
+scaled_data <- scale_data_manual(clusteredspheres)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+clusteredspheres_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r clusteredspheres-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+clusteredspheres_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r clusteredspheres-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+clusteredspheres_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-clusteredspheres-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `clusteredspheres` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+clusteredspheres_proj1 + clusteredspheres_proj2 + clusteredspheres_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-hemisphere, echo=TRUE}
+hemisphere <- gen_hemisphere(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(hemisphere)
+```
+
+```{r hemisphere-proj1}
+scaled_data <- scale_data_manual(hemisphere)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+hemisphere_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r hemisphere-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+hemisphere_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r hemisphere-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+hemisphere_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-hemisphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `hemisphere` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+hemisphere_proj1 + hemisphere_proj2 + hemisphere_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Swiss Roll

### 2025-05-01 paper-cardinalR.Rmd
Added initial polynomial projections

@@
 
+```{r data-quadratic, echo=TRUE}
+quadratic <- gen_quadratic(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(quadratic)
+```
+
+```{r quadratic-proj1}
+scaled_data <- scale_data_manual(quadratic)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+quadratic_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r quadratic-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+quadratic_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r quadratic-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+quadratic_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-quadratic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+quadratic_proj1 + quadratic_proj2 + quadratic_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-cubic, echo=TRUE}
+cubic <- gen_cubic(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(cubic)
+```
+
+```{r cubic-proj1}
+scaled_data <- scale_data_manual(cubic)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+cubic_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cubic-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+cubic_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cubic-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+cubic_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-cubic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cubic` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+cubic_proj1 + cubic_proj2 + cubic_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Pyramid

### 2025-05-01 paper-cardinalR.Rmd
Added initial cube projections

@@
 
+```{r data-gridcube, echo=TRUE}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(gridcube)
+```
+
+```{r gridcube-proj1}
+scaled_data <- scale_data_manual(gridcube)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+gridcube_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridcube-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+gridcube_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridcube-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+gridcube_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-gridcube-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+gridcube_proj1 + gridcube_proj2 + gridcube_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-unifcube, echo=TRUE}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(unifcube)
+```
+
+```{r unifcube-proj1}
+scaled_data <- scale_data_manual(unifcube)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+unifcube_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifcube-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+unifcube_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifcube-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+unifcube_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-unifcube-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `unifcube` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+unifcube_proj1 + unifcube_proj2 + unifcube_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-cubehole, echo=TRUE}
+cubehole <- gen_cubehole(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(cubehole)
+```
+
+```{r cubehole-proj1}
+scaled_data <- scale_data_manual(cubehole)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+cubehole_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cubehole-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+cubehole_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cubehole-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+cubehole_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-cubehole-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cubehole` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+cubehole_proj1 + cubehole_proj2 + cubehole_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Gaussian

### 2025-05-01 paper-cardinalR.Rmd
Added pyramid projections

@@
 
+```{r data-pyrrect, echo=TRUE}
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(pyrrect)
+```
+
+```{r pyrrect-proj1}
+scaled_data <- scale_data_manual(pyrrect)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrrect_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrrect-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrrect_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrrect-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrrect_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-pyrrect-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+pyrrect_proj1 + pyrrect_proj2 + pyrrect_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-pyrtri, echo=TRUE}
+pyrtri <- gen_pyrtri(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(pyrtri)
+```
+
+```{r pyrtri-proj1}
+scaled_data <- scale_data_manual(pyrtri)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrtri_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrtri-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrtri_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrtri-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrtri_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-pyrtri-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrtri` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-pyrstar, echo=TRUE}
+pyrstar <- gen_pyrstar(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(pyrstar)
+```
+
+```{r pyrstar-proj1}
+scaled_data <- scale_data_manual(pyrstar)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrstar_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrstar-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrstar_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrstar-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrstar_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-pyrstar-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrstar` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+pyrstar_proj1 + pyrstar_proj2 + pyrstar_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-pyrholes, echo=TRUE}
+pyrholes <- gen_pyrholes(n = 1000, p = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(pyrholes)
+```
+
+```{r pyrholes-proj1}
+scaled_data <- scale_data_manual(pyrholes)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrholes_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrholes-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrholes_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrholes-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrholes_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-pyrholes-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrholes` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+pyrholes_proj1 + pyrholes_proj2 + pyrholes_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### S-curve

### 2025-05-01 paper-cardinalR.Rmd
Added initial projection for branches

@@
 
-```{r data-expbranches}
+```{r data-expbranches, echo=TRUE}
 expbranches <- gen_expbranches(n = 1000, p = 4, k = 4)
@@
 langevitour::langevitour(expbranches)
+```
+
+```{r expbranches-proj1}
+scaled_data <- scale_data_manual(expbranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+expbranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r expbranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+expbranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r expbranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+expbranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-expbranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+expbranches_proj1 + expbranches_proj2 + expbranches_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-linearbranches, echo=TRUE}
+linearbranches <- gen_linearbranches(n = 1000, p = 4, k = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(linearbranches)
+```
+
+```{r linearbranches-proj1}
+scaled_data <- scale_data_manual(linearbranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+linearbranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linearbranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+linearbranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linearbranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+linearbranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-linearbranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linearbranches` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+linearbranches_proj1 + linearbranches_proj2 + linearbranches_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-curvybranches, echo=TRUE}
+curvybranches <- gen_curvybranches(n = 1000, p = 4, k = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(curvybranches)
+```
+
+```{r curvybranches-proj1}
+scaled_data <- scale_data_manual(curvybranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+curvybranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvybranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+curvybranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvybranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+curvybranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-curvybranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvybranches` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+curvybranches_proj1 + curvybranches_proj2 + curvybranches_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-orglinearbranches, echo=TRUE}
+orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(orglinearbranches)
+```
+
+```{r orglinearbranches-proj1}
+scaled_data <- scale_data_manual(orglinearbranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+orglinearbranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r orglinearbranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+orglinearbranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r orglinearbranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+orglinearbranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-orglinearbranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+```{r data-orgcurvybranches, echo=TRUE}
+orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(orgcurvybranches)
+```
+
+```{r orgcurvybranches-proj1}
+scaled_data <- scale_data_manual(orgcurvybranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+orgcurvybranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r orgcurvybranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+orgcurvybranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r orgcurvybranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+orgcurvybranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-orgcurvybranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orgcurvybranches` data."}
+#| fig-width: 15
+#| fig-height: 5
+
+orgcurvybranches_proj1 + orgcurvybranches_proj2 + orgcurvybranches_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-05-04 paper-cardinalR.Rmd
Added cone description

@@
 
+To simulate a cone-shaped structure in arbitrary dimensions, we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+
+This function generates `n` points in a `p`-dimensional space, where the last dimension (`X_p`) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
+
+Points along the height axis are drawn from an exponential distribution to increase density near the tip:
+
+$$
+X_p = h_i \sim \text{Exp}(\lambda = 2/h), \quad \text{truncated at } h
+$$
+
+The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter:
+
+$$
+r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) \cdot \frac{h_i}{h}, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1
+$$
+
+For each point, a direction is sampled uniformly from a $(p-1)$-dimensional hypersphere using generalized spherical coordinates (angles).
+The radial coordinates are scaled by $r_i$, which ensures a conical taper.
+Specifically, when $p = 3$, this scaling results in a classic $3\text{-}D$ cone.
+However, for $p > 3$, the introduction of additional angular components allows for a smooth extension into higher dimensions, preserving the conical shape while accommodating the complexities of multi-dimensional geometry.
+
 ```{r data-bluntedcone, echo=TRUE}
@@
 
-To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function `generate_swiss_roll(n, p)`, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
+To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function `gen_swissroll(n, p, w)`, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
 

### 2025-05-04 paper-cardinalR.Rmd
Added updated description for Swiss roll

@@
 
-To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function `gen_swissroll(n, p, w)`, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
-
-The function generates `n` points in a `p`-dimensional space, where the first two dimensions (`X_1, X_2`) define the primary Swiss roll shape using a parametric equation:
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, p, w)`, where `n` is the number of points, `p` is the total number of dimensions, and `w` is the vertical range in the third dimension.
+
+The first three dimensions form the classic $3\text{-}D$ Swiss roll shape:
 
 $$
-X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad \text{where } t \sim U(0, 3\pi)
+X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)
 $$
 
-The third dimension (`X_3`) introduces variation perpendicular to the roll, sampled uniformly from \([-1,1]\).
-Additional dimensions (`X_4` to `X_p`) extend the data structure by applying a **sinusoidal transformation** of the parameter `t`, ensuring continuity in higher-dimensional spaces:
-
-$$
-X_i = \frac{\sin(i t)}{i}, \quad \text{for } i \geq 4.
-$$
-
-This transformation ensures a gradual decay in variance across dimensions, mimicking real-world high-dimensional structures where later dimensions often capture subtler variations.
+For $p > 3$, the remaining $p - 3$ dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.
 

### 2025-05-04 paper-cardinalR.Rmd
Added S-curve explanantion

@@
 
+To simulate an S-curve structure in a higher-dimensional space, we define the function `gen_scurve(n, p)`, which generates `n` observations in `p` dimensions.
+
+The $3\text{-}D$ geometry is constructed by introducing a latent parameter, $\theta \sim U\left(-\frac{3\pi}{2}, \frac{3\pi}{2}\right)$.
+This parameter controls the curvature of the manifold.
+The first three dimensions form the S-curve structure:
+
+$$
+  X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta) \cdot (\cos(\theta) - 1).
+$$
+
+This configuration creates a horizontally curled shape in $(X_1, X_3)$, with additional band thickness in the $X_2$ direction.
+
+For $p > 3$, additional noise dimensions are appended introducing structured, wavy perturbations.
+
 ```{r data-scurve, echo=TRUE}

### 2025-05-04 paper-cardinalR.Rmd
Added S-curve hole description

@@
 
+To simulate a variant of the S-curve structure with a removed region ("hole"), we define the function `gen_scurvehole(n, p)`.
+This function produces `n` observations in a $p$-dimensional space where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
+A subset of observations near a designated anchor point is excluded to introduce a hole in the manifold.
+
+To simulate missing regions on the manifold, a fixed anchor point $(0, 1, 0, \ldots)$ is defined in $p$-dimensional space.
+All observations within a Euclidean distance of $\sqrt{0.3} \approx 0.5477$ from the anchor are removed:
+
+$$
+\text{remove if} \quad \sum_{j=1}^p (X_j - a_j)^2 \leq 0.3
+$$.
+
+This exclusion creates a hole in the manifold centered near the middle vertical region of the S-curve.
+
 ```{r data-scurvehole, echo=TRUE}
@@
 
+- Scurve with a hole allowing for evaluation of how well algorithms handle incomplete manifolds or missing local structure.
+
 <!-- The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.

### 2025-05-04 paper-cardinalR.Rmd
Added gen_circle definition

@@
 
+The function `gen_circle(n, p)` generates a $p$-dimensional dataset of $n$ observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale.
+
+A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$. Coordinates in the first two dimensions represent a perfect circle on the plane:
+
+$$
+X_1 = \cos(\theta), \quad X_2 = \sin(\theta)
+$$
+
+For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
+The first component is a scaling factor that decreases with the dimension index, defined as $\text{scale}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$.
+The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2) \cdot \frac{\pi}{2p}$.
+
+Each additional dimension is computed as:
+
+$$
+X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p
+$$
+
+This structure retains a one-dimensional manifold (the circle) but allows it to be embedded non-linearly and decorrelatively in higher dimensions.
+The decreasing scale ensures that higher dimensions add less variation, preserving the dominance of the circular shape while adding complexity.
+
 ```{r data-circle, echo=TRUE}

### 2025-05-04 paper-cardinalR.Rmd
Added curvy circle

@@
 
+The `gen_curvycycle(n, p)` function generates a $p$-dimensional dataset of $n$ observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations.
+The structure forms a one-dimensional nonlinear cycle embedded in higher-dimensional space.
+
+A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
+The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle." In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin.
+Additionally, $X_3 = \frac{1}{3} \cdot \cos(3\theta)$ introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
+
+Together, these define a periodic, non-trivial, closed curve in $3\text{-}D$ with internal folds that produce a more complex geometry than a standard circle or ellipse.
+
+For dimensions $X_4$ through $X_p$, additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
+The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from 4 to $p$, which means that the amplitude decreases as the dimension increases.
+Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j \cdot \sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2) \cdot \frac{\pi}{2p}$.
+
+This introduces decorrelated, low-variance sinusoidal variations across dimensions, retaining the cyclical nature while embedding the data in higher-dimensional space with a coherent but subtle structure.
+
 ```{r data-curvycycle, echo=TRUE}

### 2025-05-04 paper-cardinalR.Rmd
Added grided sphere description

@@
 
+The `gen_unifsphere(n, p, r)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on the surface of a $3\text{-}D$ sphere of radius $r$, with additional Gaussian noise dimensions added when $p > 3$.
+
+Each observation lies on the surface of a sphere in $3\text{-}D$, constructed by generating $u \sim U(-1, 1)$ which represents the cosine of the polar angle $\phi$ and $\theta \sim U(0, 2\pi)$ which represents the azimuthal angle.
+
+The corresponding Cartesian coordinates are calculated as;
+
+$$
+X_1 = r \cdot \sqrt{1 - u^2} \cdot \cos(\theta),
+$$
+$$
+X_2 = r \cdot \sqrt{1 - u^2} \cdot \sin(\theta),
+$$
+$$
+X_3 = r \cdot u,
+$$
+
+which gives points uniformly distributed on the surface of a $3\text{-}D$ sphere (not within).
+
+For $p > 3$, additional dimensions $X_4$ to $X_p$ are generated as low-variance Gaussian noise:
+
+$$
+X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 4, \dots, p.
+$$
+
+This extends the $3\text{-}D$ spherical manifold into $p$-dimensional space with orthogonal Gaussian noise, preserving the spherical structure while allowing for testing in higher-dimensional settings.
+
 ```{r data-unifsphere, echo=TRUE}
@@
 ```
+
+The `gen_gridedsphere(n, p)` function generates a $p$-dimensional dataset of approximately $n$ observations evenly distributed on the surface of a $3\text{-}D$ unit sphere, with optional Gaussian noise dimensions when $p > 3$.
+
+The base structure consists of a $3\text{-}D$ spherical surface created using a regular grid in spherical coordinates, where $\theta \in [0, 2\pi]$ represents the azimuthal angle (longitude) and $\phi \in [0, \pi]$ denotes the polar angle (co-latitude).
+
+The number of grid steps along each dimension is determined by factoring $n$ into two approximately equal integers via `gen_nproduct(n, p = 2)`.
+
+Each point on the sphere is computed using the spherical-to-Cartesian transformation:
+
+$$
+X_1 = \sin(\phi) \cdot \cos(\theta),
+$$
+$$
+X_2 = \sin(\phi) \cdot \sin(\theta),
+$$
+$$
+X_3 = \cos(\phi).
+$$
+
+This forms a structured grid of points on the unit sphere, ideal for studying how regular geometric manifolds are transformed under nonlinear dimensionality reduction.
+
+For $p > 3$, the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions:
+
+$$
+X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 4, \dots, p.
+$$
+
+These dimensions represent low-variance orthogonal noise, preserving the spherical shape while testing robustness in high-dimensional settings.
 

### 2025-05-05 paper-cardinalR.Rmd
Added hemisphere

@@
 
+The `gen_clusteredspheres(n, k, p, r, loc)` function generates a synthetic dataset of $n_1 + k \cdot n_2$ observations in $p$-dimensional space, consisting of one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location.
+
+#### Geometric Structure
+
+- A **large uniform sphere** centered at the origin:
+  - \(n_1\) points are sampled **uniformly on the surface** of a \(p\)-dimensional sphere of radius \(r_1\).
+  - Generated using `gen_unifsphere(n_1, p, r_1)`.
+
+- \(k\) **smaller uniform spheres**, each with:
+  - \(n_2\) points sampled uniformly on a sphere of radius \(r_2\),
+  - Positioned at different random locations in \(p\)-space:
+    - Each center is drawn from \( \mathcal{N}(0, \texttt{loc}^2 I_p) \),
+    - Ensures separation between clusters.
+
+All points on all spheres are generated using the standard hyperspherical method:
+
+- For each point:
+  - \(u \sim U(-1, 1)\), the cosine of the polar angle,
+  - \(\theta \sim U(0, 2\pi)\), the azimuthal angle (for 3D),
+  - Generalized to \(p\)-dimensions to uniformly sample on the sphere’s surface.
+
+Each observation is classified by cluster, with labels such as "big" for the large central sphere and "small_1" to "small_k" for the smaller spheres.
+
 ```{r data-clusteredspheres, echo=TRUE}
@@
 ```
+
+The `gen_hemisphere(n, p)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on a $4\text{-}D$ hemisphere, optionally extended with Gaussian noise in additional dimensions when $p > 4$.
+
+Each observation is situated on a restricted $4\text{-}D$ spherical surface, defined by spherical coordinates.
+The azimuthal angle $\theta_1 \sim U(0, \pi)$ in the $(x_1, x_2)$ plane, while the elevation angle $\theta_2 \sim U(0, \pi)$ in the $(x_2, x_3)$ plane.
+Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
+
+The coordinates are transformed into $4\text{-}D$ Cartesian space:
+
+$$
+X_1 = \sin(\theta_1) \cdot \cos(\theta_2),
+$$
+$$
+X_2 = \sin(\theta_1) \cdot \sin(\theta_2),
+$$
+$$
+X_3 = \cos(\theta_1) \cdot \cos(\theta_3),
+$$
+$$
+X_4 = \cos(\theta_1) \cdot \sin(\theta_3).
+$$
+
+This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
+
+For $p > 4$, additional dimensions $X_5$ to $X_p$ are added as low-variance Gaussian noise:
+
+$$
+X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \dots, p.
+$$
+
+These dimensions maintain the spherical structure while simulating embedding into higher-dimensional space with small orthogonal perturbations.
 
@@
 
+- clusteredsphere: This structure allows for cluster separation on curved manifolds in high-dimensional space and can be used to test the ability of NLDR methods and clustering algorithms to detect spherical clusters of different sizes and separations.
+
 <!-- The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.

### 2025-05-05 paper-cardinalR.Rmd
Added sphere section

@@
 
-#### Geometric Structure
-
-- A **large uniform sphere** centered at the origin:
-  - \(n_1\) points are sampled **uniformly on the surface** of a \(p\)-dimensional sphere of radius \(r_1\).
-  - Generated using `gen_unifsphere(n_1, p, r_1)`.
-
-- \(k\) **smaller uniform spheres**, each with:
-  - \(n_2\) points sampled uniformly on a sphere of radius \(r_2\),
-  - Positioned at different random locations in \(p\)-space:
-    - Each center is drawn from \( \mathcal{N}(0, \texttt{loc}^2 I_p) \),
-    - Ensures separation between clusters.
-
-All points on all spheres are generated using the standard hyperspherical method:
-
-- For each point:
-  - \(u \sim U(-1, 1)\), the cosine of the polar angle,
-  - \(\theta \sim U(0, 2\pi)\), the azimuthal angle (for 3D),
-  - Generalized to \(p\)-dimensions to uniformly sample on the sphere’s surface.
+A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p$-dimensional sphere with a radius of $r_1$.
+The sampling is executed using the function `gen_unifsphere(n_1, p, r_1)`, which generates the desired points in the specified dimensional space.
+
+In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
+These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $\mathcal{N}(0, \texttt{loc}^2 I_p)$.
+
+Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for 3D).
 

### 2025-05-06 paper-cardinalR.Rmd
Added gen_crescent definition

@@
 
+The `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise.
+
+The first two dimensions define a crescent-shaped curve using a semi-circular arc. Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
+
+The corresponding $2\text{-}D$ coordinates are defined by:
+
+$$
+X_1 = \cos(\theta),
+$$
+$$
+X_2 = \sin(\theta).
+$$
+
+This forms a nonlinear curvilinear structure in the shape of a tilted crescent in the $X_1$-$X_2$ plane.
+
+For $p > 2$, additional dimensions $X_3$ to $X_p$ are generated using `gen_wavydims1()`:
+
+$$
+X_j = s_j \cdot \theta + \varepsilon_j$,\text{ where }$\varepsilon_j \sim \mathcal{N}(0, 0.5^2)$,\text{ and }$s_j$\text{ is a random scale}.
+$$
+
+These noise dimensions are nonlinearly correlated with $\theta$, producing structured, wave-like variation that aligns with the progression along the crescent curve.
+
 ```{r data-crescent, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added gen_curvycylinder

@@
 
+The `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$.
+
+The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension:
+
+Let $\theta \sim \mathcal{U}(0, 3\pi)$ represent a random angle on a circular base and $z \sim \mathcal{U}(0, h)$ represent the height along the cylinder.
+
+The coordinates are defined as:
+
+* $X_1 = \cos(\theta)$ (Circular base, x-axis),
+* $X_2 = \sin(\theta)$ (Circular base, y-axis),
+* $X_3 = z$ (Linear height),
+* $X_4 = \sin(z)$ (Nonlinear curvy variation along height).
+
+This forms a curvy cylindrical surface in $4\text{-}D$, where the fourth dimension bends periodically along the height axis, resembling a helicoid or twisting wave along the cylinder.
+
+For $p > 4$, dimensions $X_5$ through $X_p$ are generated as:
+
+$$
+X_j \sim \mathcal{N}(0, 0.05^2)$,\text{ for }$j = 5, \ldots, p.
+$$
+
+These dimensions represent unstructured, low-variance noise.
+
 ```{r data-curvycylinder, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added the definition for gen_sphericalspiral

@@
 
+The `gen_sphericalspiral(n, p, spins)` function simulates a dataset of $n$ observations that form a spiraling path on a spherical surface in the first four dimensions.
+When extended beyond $4\text{-}D$, structured nonlinear noise dimensions are added to simulate more realistic high-dimensional manifolds.
+
+The first three dimensions represent points on a unit sphere.
+Let $\theta \in [0, 2\pi \times \text{spins}]$ be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
+
+Cartesian coordinates from spherical conversion:
+
+  * $X_1 = \sin(\phi) \cdot \cos(\theta)$
+  * $X_2 = \sin(\phi) \cdot \sin(\theta)$
+  * $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim \mathcal{U}(-0.5, 0.5)$ introduces vertical jitter.
+  * $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
+
+This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
+
+If $p > 4$, the function appends structured, non-linear noise via `gen_wavydims2()`.
+These dimensions are functions of the first coordinate $X_1$, introducing dependencies that preserve some geometric coherence.
+Each added dimension follows the form:
+
+$$
+X_j = s_j \cdot (-1)^{\lfloor j/2 \rfloor} \cdot X_1^{a_j} + \eta_j,
+$$
+
+where:
+$a_j \in \{2, 3, 4, 5\}$ is a randomly chosen polynomial power,
+$s_j \sim \mathcal{U}(0.5, 2)$ is a scale factor,
+$\eta_j \sim \mathcal{U}(-\sigma_j, 2\sigma_j)$, with $\sigma_j \sim \mathcal{U}(0, 0.05)$, adds mild randomness.
+
 ```{r data-sphericalspiral, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Add gen_helicalspiral

@@
 
+The `gen_helicalspiral(n, p)` function generates a dataset of $n$ observations forming a helical spiral, embedded in the first four dimensions of a $p$-dimensional space.
+Additional dimensions (if $p > 4$) are filled with structured noise to mimic high-dimensional complexity.
+
+The first four coordinates follow a $3\text{-}D$ helix with an additional oscillating component. Let $\theta \in [0, \frac{5\pi}{4}]$ be a sequence of angles controlling rotation around a circle.
+
+Cartesian coordinates:
+  * $X_1 = \cos(\theta)$: circular trajectory along the x-axis.
+
+  * $X_2 = \sin(\theta)$: circular trajectory along the y-axis.
+
+  * $X_3 = 0.05 \cdot \theta + \varepsilon_3$, with $\varepsilon_3 \sim \mathcal{U}(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix.
+
+  * $X_4 = 0.1 \cdot \sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
+
+This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
+
+When $p > 4$, the remaining $p - 4$ dimensions are populated using `gen_noisedims()`. These are independent Gaussian noise dimensions:
+
+$$
+X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+$$
+
 ```{r data-helicalspiral, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added gen_conicspiral

@@
 
+The `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of a $p$-dimensional space.
+The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a $3\text{-}D$ conic helix.
+
+The shape is defined by parameter $\theta \in [0, 2\pi \cdot \text{spins}]$, controlling the angular progression of the spiral. The mapping to Cartesian space is as follows:
+
+The Archimedean spiral in the horizontal plane is represented by;
+
+$X_1 = \theta \cdot \cos(\theta)$ for radial expansion in x.
+$X_2 = \theta \cdot \sin(\theta)$ for radial expansion in y.
+
+The growth pattern resembles a cone, with the height increasing according to $X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim \mathcal{U}(-0.1, 0.6).$
+
+Spiral modulation in the fourth dimension is represented by $X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim \mathcal{U}(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
+
+This results in a $3\text{-}D$ spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
+
+For $p > 4$, the dataset includes isotropic Gaussian noise in the remaining dimensions via `gen_noisedims()`:
+
+$$
+X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+$$
+
 ```{r data-conicspiral, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added gen_nonlinear

@@
 
+The `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables.
+The key geometry lies in dimensions 1–4, where:
+
+* $X_1 \sim \mathcal{U}(0.1, 2)$: base variable (avoids zero to prevent division errors).
+* $X_3 \sim \mathcal{U}(0.1, 0.8)$: independent auxiliary variable.
+* $X_2 = \frac{hc}{X_1} + \text{non\_fac} \cdot \sin(X_1)$: non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
+* $X_4 = \cos(\pi \cdot X_1) + \varepsilon$, with $\varepsilon \sim \mathcal{U}(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+
+These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
+
+If $p > 4$, the remaining dimensions are populated with Gaussian noise using `gen_noisedims()`:
+
+$$
+X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+$$
+
 ```{r data-nonlinear, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added Mobius data

@@
 
+The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p$-dimensional space, centered at the origin with user-defined covariance structure.
+
+Each point is independently drawn using the multivariate normal distribution with:
+
+$$
+X_i \sim \mathcal{N}_p(\boldsymbol{0}, \Sigma)
+$$
+
+where $\Sigma$ is a user-defined $p \times p$ positive-definite matrix.
+
 ```{r data-gau, echo=TRUE}
@@
 ### Linear
+
+The `gen_longlinear(n, p)` function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise.
+
+Each variable $X_i$ is created as:
+
+$$
+X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i
+$$
+
+where:
+
+$\text{scale}_i \sim \mathcal{U}(-10, 10)$: randomly chosen direction and stretch.
+
+$\text{shift}_i \sim \mathcal{U}(-300, 300)$: large offset to spread features apart.
+
+$\epsilon \sim \mathcal{N}(0, 0.03 \cdot n)$: Gaussian noise scaled to dataset size to introduce mild irregularity.
 

### 2025-05-06 paper-cardinalR.Rmd
Added Mobius

@@
 
+The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p$-dimensional space.
+This classical non-orientable surface loops back on itself with a half-twist.
+
+The Mobius strip is generated using `geozoo::mobius(n, p = 3)$points`, which samples $n$ points from a parametrization of the surface defined by angle $\theta$ and strip width $w$.
+The cartesian coordinates are;
+
+$$
+X\_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
+$$
+$$
+X\_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
+$$
+$$
+X\_3 = \frac{w}{2} \sin(\theta/2).
+$$
+
+This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable 1-sided surface.
+
+For $p > 3$, the function appends $p - 3$ dimensions with small Gaussian noise:
+
+$$
+x_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+$$
+
 ```{r data-mobius, echo=TRUE}
@@
 
+- Mobius: The core geometric structure is a Mobius strip—a classic one-sided surface with a half-twist—useful for evaluating how well methods capture non-orientability and twisted manifolds.
+
 <!-- The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.

### 2025-05-06 paper-cardinalR.Rmd
Added gen_quadratic

@@
 
+The `gen_quadratic(n, p, range)` function generates a dataset of $n$ points forming a quadratic curve in the first two dimensions of a $p$-dimensional space.
+This $2\text{-}D$ parabolic structure is embedded within a higher-dimensional space with additive noise in the remaining dimensions.
+
+The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation.
+Let $X\_1 \sim \mathcal{U}(\text{range}\[1], \text{range}\[2])$ be the independent variable.
+A raw polynomial basis of degree 2 is used to form $X\_2 = X\_1 - X\_1^2 + \varepsilon\_2$, where $\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.
+
+This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate measurement noise or intrinsic variability.
+
+For $p > 2$, the dataset includes Gaussian noise in the remaining dimensions via `gen_noisedims()`:
+
+$$
+X\_j \sim \mathcal{N}(0, 0.1^2), \quad j = 3, \dots, p.
+$$
+
 ```{r data-quadratic, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added gen_cubic

@@
 
+The `gen_cubic(n, p, range)` function generates a dataset of $n$ points forming a cubic curve in the first two dimensions of a $p$-dimensional space.
+This function creates a more complex curvilinear structure than a simple parabola.
+
+The shape is generated using a third-degree raw polynomial basis expansion.
+Let $X\_1 \sim \mathcal{U}(\text{range}\[1], \text{range}\[2])$ be the base input.
+A cubic transformation with vertical jitter is used to define $X\_2 = X\_1 + X\_1^2 - X\_1^3 + \varepsilon\_2\$, where \$\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.
+
+When $p > 2$, the remaining dimensions consist of independent Gaussian noise features generated via `gen_noisedims()`:
+
+$$
+X\_j \sim \mathcal{N}(0, 0.1^2)\$ for \$j = 3, \dots, p.
+$$
+
 ```{r data-cubic, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Updated gen_gridcube

@@
 
+The `gen_gridcube(n, p)` constructs a regular grid using approximately $n$ points by evenly partitioning each of the $p$ dimensions.
+Specifically, the number of grid points along each axis is determined by `gen_nproduct(n, p)`, which calculates the nearest balanced integer factors whose product is close to $n$.
+The full grid is generated via `expand_grid()`, giving Cartesian combinations of these evenly spaced positions.
+
+Each variable $X\_j$ represents integer grid coordinates (e.g., 1, 2, 3, ...) in dimension $j$:
+
+$$
+X\_j \in {1, \dots, k\_j},\text{ where }\prod\_{j=1}^p k\_j \approx n.
+$$
+
+This results in a hypercube-shaped lattice in $p$-dimensional space.
+
 ```{r data-gridcube, echo=TRUE}
@@
 
+- Grided cube: This function is useful for assessing how algorithms preserve uniformly spaced data in high-dimensional spaces.
+
 <!-- The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.

### 2025-05-06 paper-cardinalR.Rmd
Added gen_unifcube

@@
 
+The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in a higher-dimensional space with additional noise dimensions.
+
+Points are sampled independently from $X\_1, X\_2, X\_3 \sim \mathcal{U}(-0.5, 0.5)$ of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.
+
+For $p > 3$, the remaining dimensions are filled with Gaussian noise:
+
+$$
+X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+$$
+
 ```{r data-unifcube, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added gen_cubehole

@@
 
+The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p$-dimensional space.
+
+The initial points are drawn from a uniform cube centered at the origin: $X\_1, X\_2, X\_3 \sim \mathcal{U}(-0.5, 0.5)$.
+
+The center of the cube is $(0, 0, 0)$. A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$:
+
+Remove if $\sqrt{X\_1^2 + X\_2^2 + X\_3^2} \leq 0.5$.
+
+This results in a shell-like point cloud that fills the cube but avoids the spherical center.
+
+For $p > 3$, Gaussian noise is added to the remaining dimensions (via `gen_unifcube()`):
+
+$$
+X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+$$.
+
 ```{r data-cubehole, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added gen_expbranches

@@
 
+The `gen_expbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ exponential branches in 2D, with optional noise dimensions to embed the structure in a \$p\$-dimensional space.
+These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
+
+Each branch $i$ is constructed using:
+
+$X\_1 \sim \mathcal{U}(-2, 2)$
+$X\_2 = \exp(\pm s\_i \cdot X\_1) + \epsilon$,
+
+where:
+$\epsilon \sim \mathcal{U}(0, 0.1)$ adds local jitter
+$s\_i \in \[0.5, 2]$ is randomly sampled for each branch.
+
+The sign alternates between branches (odd: negative, even: positive exponent).
+This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches: decay pattern (reflected about $x$-axis) and even-numbered branches: growth pattern.
+
+For $p > 2$, Gaussian noise is added to embed the $2\text{-}D$ branches in higher dimensions:
+
+$$
+X\_j \sim \mathcal{N}(0, 0.1^2), \quad j = 3, \dots, p.
+$$
+
+These represent irrelevant or noisy dimensions.
+
 ```{r data-expbranches, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added gen_linearbranches

@@
 
+The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches in a $p$-dimensional space.
+The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.
+
+Each branch is a segment of a line with added jitter to simulate measurement noise.
+The branches differ in direction and location.
+Branches 1 and 2 are initialized with fixed slopes and intercepts.
+The Branch 1 is generated from $X\_1 \sim \mathcal{U}(-2, 8)\$, \$X\_2 = 0.5 \cdot X\_1 + \epsilon$.
+The Branch 2 is generated from $X\_1 \sim \mathcal{U}(-6, 2)\$, \$X\_2 = -0.5 \cdot X\_1 + \epsilon$.
+$\epsilon \sim \mathcal{U}(0, 0.5)$.
+
+Branches 3 to $k$** are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
+
+$X\_1$ values span a short segment ($x\_{start}$ to $x\_{start} + 1$),
+
+$X\_2 = s\_i \cdot (X\_1 - x\_{start}) + y\_{start} + \epsilon$ where $s\_i$ is a sampled slope from a filtered set and $\epsilon \sim \mathcal{U}(0, 0.2)$.
+
+This setup yields $k$ spatially distinct, noisy linear trajectories.
+
+For $p > 2$, the remaining dimensions are pure noise:
+
+$$
+X\_j \sim \mathcal{N}(0, 0.05^2)\$ for \$j = 3, \dots, p.
+$$
+
 ```{r data-linearbranches, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Addded gen_curvybranches

@@
 
+The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in a $p$-dimensional space.
+The underlying geometry lies in the first two dimensions, while the remaining $p - 2$ dimensions contain Gaussian noise.
+
+Branch 1 is generated from $X\_1 \sim \mathcal{U}(0, 1)$, and $X\_2 = 0.1 \cdot X\_1 + 1 \cdot X\_1^2 + \epsilon$, where $\epsilon \sim \mathcal{U}(0, 0.05)$.
+This produces a gently upward-curving parabola in the right half-plane.
+
+Branch 2 is generated from $X\_1 \sim \mathcal{U}(-1, 0)$, and $X\_2 = 0.1 \cdot X\_1 - 2 \cdot X\_1^2 + \epsilon$, where $\epsilon \sim \mathcal{U}(0, 0.05)$.
+This creates a steeper, leftward-facing curve in the left half-plane.
+
+Branches $3$ to $k$** are added iteratively.
+Each new branch begins at a randomly chosen point within a restricted horizontal band: $X\_1 \in \[-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X\_1\$:$X\_1 \in \[x\_{\text{start}}, x\_{\text{start}} + 1]$, and has the structure:
+
+$$
+X_2 = 0.1 \cdot X_1 - s_i \cdot (X_1^2 - x_{\text{start}}) + y_{\text{start}}, \quad s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}.
+$$
+
+Here, $s\_i$ is a sampled scale factor (slope-like term) controlling curvature, and $(x\_{\text{start}}, y\_{\text{start}})$ is the starting point from the existing structure.
+
+This construction yields $k$ spatially connected, nonlinear branches with varying curvature.
+
+For $p > 2$, the remaining dimensions are pure Gaussian noise:
+
+$$
+X_j \sim \mathcal{N}(0, 0.05^2), \quad \text{for } j = 3, \dots, p.
+$$
+
 ```{r data-curvybranches, echo=TRUE}

### 2025-05-06 paper-cardinalR.Rmd
Added branching definitions

@@
 
+The `gen_orglinearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches embedded in a $p$-dimensional space.
+Each branch lies primarily within a distinct $2\text{-}D$ subspace, while the remaining $p - 2$ dimensions contain Gaussian noise.
+
+To construct each branch, a unique or repeated pair of dimensions is selected from the $\binom{p}{2}$ possible $2\text{-}D$ combinations.
+If $k \leq \binom{p}{2}$, combinations are sampled without replacement.
+If $k > \binom{p}{2}$, additional pairs are sampled with replacement to reach $k$ total branches.
+Each selected pair $(x\_{i1}, x\_{i2})$ defines the $2\text{-}D$ plane for branch $i$.
+
+For branch $i$, $n\_i$ points are generated, where $\sum\_{i=1}^k n\_i = n$. The structure follows:
+
+$$
+x_{i1} \sim \mathcal{U}(0, 2), \quad
+x_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim \mathcal{U}(0, 0.5),
+$$
+
+where $s\_i$ is a scale factor controlling the slope.
+When $k \leq \binom{p}{2}$, $s\_i = 1$.
+When sampling with replacement, $s\_i$ is drawn from the set ${1, 1.5, 2, \dots, 8}$ in increments of $0.5$.
+
+For $p > 2$, the remaining dimensions contain independent Gaussian noise:
+
+$$
+x_j \sim \mathcal{N}(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
+$$
+
 ```{r data-orglinearbranches, echo=TRUE}
@@
 ```
+
+The `gen_orgcurvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in a $p$-dimensional space.
+Each branch is constructed in a unique or repeated $2\text{-}D$ subspace of the $p$-dimensional space, with curvature induced by a second-degree polynomial structure.
+The remaining $p - 2$ dimensions contain Gaussian noise.
+
+Let $\binom{p}{2}$ denote the number of unique $2\text{-}D$ subspace combinations.
+When $k \leq \binom{p}{2}$, $k$ distinct subspace pairs $(x\_{i1}, x\_{i2})$ are sampled without replacement.
+Otherwise, combinations are selected with replacement to reach $k$ total branches.
+For each branch $i = 1, \dots, k$, a random scale factor $s\_i$ is used to vary the curvature:
+
+* If $k \leq \binom{p}{2}$: $s\_i = 1 $
+* If $k > \binom{p}{2}$: $s\_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}$
+
+Each branch contains $n\_i$ points such that $\sum\_{i=1}^k n\_i = n$, where the vector $(n\_1, \dots, n\_k)$ is randomly drawn using the helper function `gen_nsum()` to partition $n$.
+
+Within its assigned subspace $(x\_{i1}, x\_{i2})$, branch $i$ is defined by:
+
+$$
+x_{i1} \sim \mathcal{U}(0, 2), \quad
+x_{i2} = -s_i \cdot x_{i1}^2 + \epsilon, \quad \epsilon \sim \mathcal{U}(0, 0.5)
+$$
+
+This forms a smooth downward-opening parabola in the plane defined by $(x\_{i1}, x\_{i2})$, with the degree of curvature controlled by $s\_i$.
+
+For $p > 2$, all remaining dimensions are independent Gaussian noise:
+
+$$
+x_j \sim \mathcal{N}(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
+$$
 

### 2025-05-06 paper-cardinalR.Rmd
Added initial pyr descriptions

@@
 
+The `gen_pyrrect(n, p, h, l_vec, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a rectangular base.
+The pyramid is embedded in a $p$-dimensional space, with a tip at height zero and base at height $h$.
+The shape tapers linearly from the base dimensions $(l\_x, l\_y)$ to a smaller rectangular section with side lengths $(2r\_t, 2r\_t)$ at the tip.
+Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.
+
+Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The final dimension $x\_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$:
+
+$$
+x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)
+$$.
+
+Let $r\_x(z)$ and $r\_y(z)$ denote the half-widths of the rectangular cross-section at height $z$:
+
+$$
+r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}, \quad
+r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}
+$$.
+
+The first three coordinates are then defined as:
+
+$$
+x_1 \sim \mathcal{U}(-r_x(z),\ r_x(z)), \quad
+x_2 \sim \mathcal{U}(-r_y(z),\ r_y(z)), \quad
+x_3 \sim \mathcal{U}(-r_x(z),\ r_x(z))
+$$.
+
+For $p > 3$, the remaining $p - 4$ dimensions (i.e., $x\_4$ to $x\_{p-1}$) are tapered toward zero with decreasing height and generated as:
+
+$$
+x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
+$$
+
+This tapering ensures that the structure narrows in all directions as it approaches the tip.
+
 ```{r data-pyrrect, echo=TRUE}
@@
 
+The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section.
+The structure is embedded in a $p$-dimensional space, with the tip located at height 0 and the base at height $h$.
+The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
+
+Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points.
+The final coordinate \$x\_p\$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:
+
+$$
+x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
+$$
+
+Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$:
+
+$$
+r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
+$$
+
+A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section:
+
+$$
+u, v \sim \mathcal{U}(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
+$$
+
+The first three coordinates (triangle plane) are then:
+
+$$
+x_1 = r(z)(1 - u - v), \quad
+x_2 = r(z) \cdot u, \quad
+x_3 = r(z) \cdot v
+$$
+
+For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip:
+
+$$
+x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
+$$
+
+The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section.
+The structure is embedded in a $p$-dimensional space, with the tip located at height 0 and the base at height $h$.
+The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
+
+Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points.
+The final coordinate $x\_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:
+
+$$
+x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
+$$
+
+Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$:
+
+$$
+r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
+$$
+
+A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section:
+
+$$
+u, v \sim \mathcal{U}(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
+$$
+
+The first three coordinates (triangle plane) are then:
+
+$$
+x_1 = r(z)(1 - u - v), \quad
+x_2 = r(z) \cdot u, \quad
+x_3 = r(z) \cdot v
+$$
+
+For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip:
+
+$$
+x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
+$$
+
+The result is a tapered, high-dimensional triangle-based pyramid with more points near the apex and greater width at the base.
+
 ```{r data-pyrtri, echo=TRUE}
@@
 
+The `gen_pyrstar(n, p, h, rb)` function generates a dataset of \$n\$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
+The pyramid extends vertically from height $z = 0$ (tip) to height $z = h$ (base), embedded in a $p$-dimensional space.
+The distribution concentrates more points near the base, and the hexagonal spread tapers toward the tip.
+
+Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The final coordinate $x\_p$ encodes the height $z$:
+
+$$
+x_p = z \sim \mathcal{U}(0, h).
+$$
+
+The radius at height $z$ scales linearly from zero (tip) to the base radius $r\_b$:
+
+$$
+r(z) = r_b \cdot \left(1 - \frac{z}{h}\right).
+$$
+
+Each point is placed within a regular hexagon in the plane $(x\_1, x\_2)$, using a randomly chosen hexagon sector angle $\theta \in \left{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3} \right}$ and a uniformly random radial scaling factor:
+
+$$
+\theta \sim \text{Uniform sample from 6 hexagon angles}, \quad
+r_{\text{point}} \sim \sqrt{\mathcal{U}(0, 1)}.
+$$
+
+Then, the first two coordinates are:
+
+$$
+x_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta), \quad
+x_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)
+$$
+
+For dimensions $j = 3, \dots, p-1$, values are tapered toward zero near the tip:
+
+$$
+x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
+$$
+
+This results in a smooth, star-like pyramid in $p$-dimensional space, with a broad hexagonal base and narrow tip.
+
 ```{r data-pyrstar, echo=TRUE}
@@
 ```
+
+The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p$-dimensional simplex using a chaotic attractor-like midpoint algorithm.
+The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a "Sierpinski-like pyramid" in high dimensions.
+
+Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T\_0 \in \[0, 1]^p$ drawn from a uniform distribution:
+
+$$
+T_0 \sim \mathcal{U}(0, 1)^p.
+$$
+
+Let $C\_1, C\_2, \dots, C\_{p+1}$ denote the corner vertices of a $p$-dimensional simplex.
+At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T\_{i-1}$ and a randomly selected vertex $C\_k$:
+
+$$
+T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}.
+$$
+
+This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points. The points remain bounded inside the convex hull of the simplex.
+
+The final output is a $n \times p$ matrix where each row represents a point:
+
+$$
+X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}.
+$$
 
@@
 
+- Pyrholes: This structure is useful for testing clustering and NLDR algorithms on non-convex and sparse high-dimensional shapes.
+
 <!-- The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.

### 2025-05-07 paper-cardinalR.Rmd
Removed the subsection of application
Try to fix the issues of compiling the Latex code

@@
 where:
-$\epsilon \sim \mathcal{U}(0, 0.1)$ adds local jitter
-$s\_i \in \[0.5, 2]$ is randomly sampled for each branch.
+$\epsilon \sim \mathcal{U}(0, 0.1)$ adds local jitter,
+
+$s\_i \in [0.5, 2]$ is randomly sampled for each branch.
 
@@
 Branches $3$ to $k$** are added iteratively.
-Each new branch begins at a randomly chosen point within a restricted horizontal band: $X\_1 \in \[-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X\_1\$:$X\_1 \in \[x\_{\text{start}}, x\_{\text{start}} + 1]$, and has the structure:
+Each new branch begins at a randomly chosen point within a restricted horizontal band: $X\_1 \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X\_1$:$X\_1 \in [x\_{\text{start}}, x\_{\text{start}} + 1]$, and has the structure:
 
@@
 The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation.
-Let $X\_1 \sim \mathcal{U}(\text{range}\[1], \text{range}\[2])$ be the independent variable.
+Let $X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the independent variable.
 A raw polynomial basis of degree 2 is used to form $X\_2 = X\_1 - X\_1^2 + \varepsilon\_2$, where $\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.
@@
 The shape is generated using a third-degree raw polynomial basis expansion.
-Let $X\_1 \sim \mathcal{U}(\text{range}\[1], \text{range}\[2])$ be the base input.
+Let $X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the base input.
 A cubic transformation with vertical jitter is used to define $X\_2 = X\_1 + X\_1^2 - X\_1^3 + \varepsilon\_2\$, where \$\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.
@@
 
-Each point is placed within a regular hexagon in the plane $(x\_1, x\_2)$, using a randomly chosen hexagon sector angle $\theta \in \left{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3} \right}$ and a uniformly random radial scaling factor:
+Each point is placed within a regular hexagon in the plane $(x\_1, x\_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor:
 
@@
 
-Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T\_0 \in \[0, 1]^p$ drawn from a uniform distribution:
+Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T\_0 \in [0, 1]^p$ drawn from a uniform distribution:
 
@@
 $$
-X_j = s_j \cdot \theta + \varepsilon_j$,\text{ where }$\varepsilon_j \sim \mathcal{N}(0, 0.5^2)$,\text{ and }$s_j$\text{ is a random scale}.
+X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim \mathcal{N}(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
 $$
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2)$,\text{ for }$j = 5, \ldots, p.
+X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \ldots, p.
 $$
@@
 
-## Assessing the performance of dimension reduction on different geometric structures in high-dimensions
+<!-- ## Assessing the performance of dimension reduction on different geometric structures in high-dimensions -->
 

### 2025-05-07 paper-cardinalR.tex
Try to fix the issues of compiling the Latex code

@@
 
+The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in 2D, with optional noise dimensions to embed the structure in a \$p\$-dimensional space.
+These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
+
+Each branch \(i\) is constructed using:
+
+\(X\_1 \sim \mathcal{U}(-2, 2)\)
+\(X\_2 = \exp(\pm s\_i \cdot X\_1) + \epsilon\),
+
+where:
+\(\epsilon \sim \mathcal{U}(0, 0.1)\) adds local jitter,
+
+\(s\_i \in [0.5, 2]\) is randomly sampled for each branch.
+
+The sign alternates between branches (odd: negative, even: positive exponent).
+This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches: decay pattern (reflected about \(x\)-axis) and even-numbered branches: growth pattern.
+
+For \(p > 2\), Gaussian noise is added to embed the \(2\text{-}D\) branches in higher dimensions:
+
+\[
+X\_j \sim \mathcal{N}(0, 0.1^2), \quad j = 3, \dots, p.
+\]
+
+These represent irrelevant or noisy dimensions.
+
+\begin{verbatim}
+expbranches <- gen_expbranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-expbranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches` data.}\label{fig:fig-expbranches-proj}
+\end{figure}
+
+The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches in a \(p\)-dimensional space.
+The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.
+
+Each branch is a segment of a line with added jitter to simulate measurement noise.
+The branches differ in direction and location.
+Branches 1 and 2 are initialized with fixed slopes and intercepts.
+The Branch 1 is generated from \(X\_1 \sim \mathcal{U}(-2, 8)\$, \$X\_2 = 0.5 \cdot X\_1 + \epsilon\).
+The Branch 2 is generated from \(X\_1 \sim \mathcal{U}(-6, 2)\$, \$X\_2 = -0.5 \cdot X\_1 + \epsilon\).
+\(\epsilon \sim \mathcal{U}(0, 0.5)\).
+
+Branches 3 to \(k\)** are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
+
+\(X\_1\) values span a short segment (\(x\_{start}\) to \(x\_{start} + 1\)),
+
+\(X\_2 = s\_i \cdot (X\_1 - x\_{start}) + y\_{start} + \epsilon\) where \(s\_i\) is a sampled slope from a filtered set and \(\epsilon \sim \mathcal{U}(0, 0.2)\).
+
+This setup yields \(k\) spatially distinct, noisy linear trajectories.
+
+For \(p > 2\), the remaining dimensions are pure noise:
+
+\[
+X\_j \sim \mathcal{N}(0, 0.05^2)\$ for \$j = 3, \dots, p.
+\]
+
+\begin{verbatim}
+linearbranches <- gen_linearbranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-linearbranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linearbranches` data.}\label{fig:fig-linearbranches-proj}
+\end{figure}
+
+The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in a \(p\)-dimensional space.
+The underlying geometry lies in the first two dimensions, while the remaining \(p - 2\) dimensions contain Gaussian noise.
+
+Branch 1 is generated from \(X\_1 \sim \mathcal{U}(0, 1)\), and \(X\_2 = 0.1 \cdot X\_1 + 1 \cdot X\_1^2 + \epsilon\), where \(\epsilon \sim \mathcal{U}(0, 0.05)\).
+This produces a gently upward-curving parabola in the right half-plane.
+
+Branch 2 is generated from \(X\_1 \sim \mathcal{U}(-1, 0)\), and \(X\_2 = 0.1 \cdot X\_1 - 2 \cdot X\_1^2 + \epsilon\), where \(\epsilon \sim \mathcal{U}(0, 0.05)\).
+This creates a steeper, leftward-facing curve in the left half-plane.
+
+Branches \(3\) to \(k\)** are added iteratively.
+Each new branch begins at a randomly chosen point within a restricted horizontal band: \(X\_1 \in [-0.15, 0.15]\) (to ensure connectivity with earlier branches), spans a unit-length interval on \(X\_1\):\(X\_1 \in [x\_{\text{start}}, x\_{\text{start}} + 1]\), and has the structure:
+
+\[
+X_2 = 0.1 \cdot X_1 - s_i \cdot (X_1^2 - x_{\text{start}}) + y_{\text{start}}, \quad s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}.
+\]
+
+Here, \(s\_i\) is a sampled scale factor (slope-like term) controlling curvature, and \((x\_{\text{start}}, y\_{\text{start}})\) is the starting point from the existing structure.
+
+This construction yields \(k\) spatially connected, nonlinear branches with varying curvature.
+
+For \(p > 2\), the remaining dimensions are pure Gaussian noise:
+
+\[
+X_j \sim \mathcal{N}(0, 0.05^2), \quad \text{for } j = 3, \dots, p.
+\]
+
+\begin{verbatim}
+curvybranches <- gen_curvybranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-curvybranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvybranches` data.}\label{fig:fig-curvybranches-proj}
+\end{figure}
+
+The \texttt{gen\_orglinearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches embedded in a \(p\)-dimensional space.
+Each branch lies primarily within a distinct \(2\text{-}D\) subspace, while the remaining \(p - 2\) dimensions contain Gaussian noise.
+
+To construct each branch, a unique or repeated pair of dimensions is selected from the \(\binom{p}{2}\) possible \(2\text{-}D\) combinations.
+If \(k \leq \binom{p}{2}\), combinations are sampled without replacement.
+If \(k > \binom{p}{2}\), additional pairs are sampled with replacement to reach \(k\) total branches.
+Each selected pair \((x\_{i1}, x\_{i2})\) defines the \(2\text{-}D\) plane for branch \(i\).
+
+For branch \(i\), \(n\_i\) points are generated, where \(\sum\_{i=1}^k n\_i = n\). The structure follows:
+
+\[
+x_{i1} \sim \mathcal{U}(0, 2), \quad
+x_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim \mathcal{U}(0, 0.5),
+\]
+
+where \(s\_i\) is a scale factor controlling the slope.
+When \(k \leq \binom{p}{2}\), \(s\_i = 1\).
+When sampling with replacement, \(s\_i\) is drawn from the set \({1, 1.5, 2, \dots, 8}\) in increments of \(0.5\).
+
+For \(p > 2\), the remaining dimensions contain independent Gaussian noise:
+
+\[
+x_j \sim \mathcal{N}(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
+\]
+
+\begin{verbatim}
+orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-orglinearbranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `orglinearbranches` data.}\label{fig:fig-orglinearbranches-proj}
+\end{figure}
+
+The \texttt{gen\_orgcurvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in a \(p\)-dimensional space.
+Each branch is constructed in a unique or repeated \(2\text{-}D\) subspace of the \(p\)-dimensional space, with curvature induced by a second-degree polynomial structure.
+The remaining \(p - 2\) dimensions contain Gaussian noise.
+
+Let \(\binom{p}{2}\) denote the number of unique \(2\text{-}D\) subspace combinations.
+When \(k \leq \binom{p}{2}\), \(k\) distinct subspace pairs \((x\_{i1}, x\_{i2})\) are sampled without replacement.
+Otherwise, combinations are selected with replacement to reach \(k\) total branches.
+For each branch \(i = 1, \dots, k\), a random scale factor \(s\_i\) is used to vary the curvature:
+
+\begin{itemize}
+\tightlist
+\item
+  If \(k \leq \binom{p}{2}\): \$s\_i = 1 \$
+\item
+  If \(k > \binom{p}{2}\): \(s\_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}\)
+\end{itemize}
+
+Each branch contains \(n\_i\) points such that \(\sum\_{i=1}^k n\_i = n\), where the vector \((n\_1, \dots, n\_k)\) is randomly drawn using the helper function \texttt{gen\_nsum()} to partition \(n\).
+
+Within its assigned subspace \((x\_{i1}, x\_{i2})\), branch \(i\) is defined by:
+
+\[
+x_{i1} \sim \mathcal{U}(0, 2), \quad
+x_{i2} = -s_i \cdot x_{i1}^2 + \epsilon, \quad \epsilon \sim \mathcal{U}(0, 0.5)
+\]
+
+This forms a smooth downward-opening parabola in the plane defined by \((x\_{i1}, x\_{i2})\), with the degree of curvature controlled by \(s\_i\).
+
+For \(p > 2\), all remaining dimensions are independent Gaussian noise:
+
+\[
+x_j \sim \mathcal{N}(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
+\]
+
+\begin{verbatim}
+orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-orgcurvybranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `orgcurvybranches` data.}\label{fig:fig-orgcurvybranches-proj}
+\end{figure}
+
 \subsubsection{Cone}\label{cone}
+
+To simulate a cone-shaped structure in arbitrary dimensions, we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+
+This function generates \texttt{n} points in a \texttt{p}-dimensional space, where the last dimension (\texttt{X\_p}) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
+
+Points along the height axis are drawn from an exponential distribution to increase density near the tip:
+
+\[
+X_p = h_i \sim \text{Exp}(\lambda = 2/h), \quad \text{truncated at } h
+\]
+
+The effective radius at height \(h_i\) decreases linearly from base to tip, controlled by a shape ratio parameter:
+
+\[
+r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) \cdot \frac{h_i}{h}, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1
+\]
+
+For each point, a direction is sampled uniformly from a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates (angles).
+The radial coordinates are scaled by \(r_i\), which ensures a conical taper.
+Specifically, when \(p = 3\), this scaling results in a classic \(3\text{-}D\) cone.
+However, for \(p > 3\), the introduction of additional angular components allows for a smooth extension into higher dimensions, preserving the conical shape while accommodating the complexities of multi-dimensional geometry.
 
@@
 
+The \texttt{gen\_gridcube(n,\ p)} constructs a regular grid using approximately \(n\) points by evenly partitioning each of the \(p\) dimensions.
+Specifically, the number of grid points along each axis is determined by \texttt{gen\_nproduct(n,\ p)}, which calculates the nearest balanced integer factors whose product is close to \(n\).
+The full grid is generated via \texttt{expand\_grid()}, giving Cartesian combinations of these evenly spaced positions.
+
+Each variable \(X\_j\) represents integer grid coordinates (e.g., 1, 2, 3, \ldots) in dimension \(j\):
+
+\[
+X\_j \in {1, \dots, k\_j},\text{ where }\prod\_{j=1}^p k\_j \approx n.
+\]
+
+This results in a hypercube-shaped lattice in \(p\)-dimensional space.
+
+\begin{verbatim}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-gridcube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube` data.}\label{fig:fig-gridcube-proj}
+\end{figure}
+
+The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in a higher-dimensional space with additional noise dimensions.
+
+Points are sampled independently from \(X\_1, X\_2, X\_3 \sim \mathcal{U}(-0.5, 0.5)\) of a cube of side length 1.
+This results in a uniform cloud of points filling the unit cube in \(3\text{-}D\) space.
+
+For \(p > 3\), the remaining dimensions are filled with Gaussian noise:
+
+\[
+X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+\]
+
+\begin{verbatim}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-unifcube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `unifcube` data.}\label{fig:fig-unifcube-proj}
+\end{figure}
+
+The \texttt{gen\_cubehole(n,\ p)} function generates a dataset of \(n\) points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in \(p\)-dimensional space.
+
+The initial points are drawn from a uniform cube centered at the origin: \(X\_1, X\_2, X\_3 \sim \mathcal{U}(-0.5, 0.5)\).
+
+The center of the cube is \((0, 0, 0)\). A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\):
+
+Remove if \(\sqrt{X\_1^2 + X\_2^2 + X\_3^2} \leq 0.5\).
+
+This results in a shell-like point cloud that fills the cube but avoids the spherical center.
+
+For \(p > 3\), Gaussian noise is added to the remaining dimensions (via \texttt{gen\_unifcube()}):
+
+\[
+X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+\].
+
+\begin{verbatim}
+cubehole <- gen_cubehole(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubehole-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cubehole` data.}\label{fig:fig-cubehole-proj}
+\end{figure}
+
 \subsubsection{Gaussian}\label{gaussian}
+
+The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\)-dimensional space, centered at the origin with user-defined covariance structure.
+
+Each point is independently drawn using the multivariate normal distribution with:
+
+\[
+X_i \sim \mathcal{N}_p(\boldsymbol{0}, \Sigma)
+\]
+
+where \(\Sigma\) is a user-defined \(p \times p\) positive-definite matrix.
+
+\begin{verbatim}
+gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:fig-gau-proj}
+\end{figure}
 
@@
 
+The \texttt{gen\_longlinear(n,\ p)} function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise.
+
+Each variable \(X_i\) is created as:
+
+\[
+X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i
+\]
+
+where:
+
+\(\text{scale}_i \sim \mathcal{U}(-10, 10)\): randomly chosen direction and stretch.
+
+\(\text{shift}_i \sim \mathcal{U}(-300, 300)\): large offset to spread features apart.
+
+\(\epsilon \sim \mathcal{N}(0, 0.03 \cdot n)\): Gaussian noise scaled to dataset size to introduce mild irregularity.
+
+\begin{verbatim}
+linear <- gen_longlinear(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-linear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linear` data.}\label{fig:fig-linear-proj}
+\end{figure}
+
 \subsubsection{Mobius}\label{mobius}
+
+The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\)-dimensional space.
+This classical non-orientable surface loops back on itself with a half-twist.
+
+The Mobius strip is generated using \texttt{geozoo::mobius(n,\ p\ =\ 3)\$points}, which samples \(n\) points from a parametrization of the surface defined by angle \(\theta\) and strip width \(w\).
+The cartesian coordinates are;
+
+\[
+X\_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
+\]
+\[
+X\_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
+\]
+\[
+X\_3 = \frac{w}{2} \sin(\theta/2).
+\]
+
+This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable 1-sided surface.
+
+For \(p > 3\), the function appends \(p - 3\) dimensions with small Gaussian noise:
+
+\[
+x_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+\]
+
+\begin{verbatim}
+mobius <- gen_mobius(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-mobius-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `mobius` data.}\label{fig:fig-mobius-proj}
+\end{figure}
 
@@
 \end{table}
+
+The \texttt{gen\_quadratic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a quadratic curve in the first two dimensions of a \(p\)-dimensional space.
+This \(2\text{-}D\) parabolic structure is embedded within a higher-dimensional space with additive noise in the remaining dimensions.
+
+The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation.
+Let \(X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])\) be the independent variable.
+A raw polynomial basis of degree 2 is used to form \(X\_2 = X\_1 - X\_1^2 + \varepsilon\_2\), where \(\varepsilon\_2 \sim \mathcal{U}(0, 0.5)\).
+
+This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate measurement noise or intrinsic variability.
+
+For \(p > 2\), the dataset includes Gaussian noise in the remaining dimensions via \texttt{gen\_noisedims()}:
+
+\[
+X\_j \sim \mathcal{N}(0, 0.1^2), \quad j = 3, \dots, p.
+\]
+
+\begin{verbatim}
+quadratic <- gen_quadratic(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-quadratic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` data.}\label{fig:fig-quadratic-proj}
+\end{figure}
+
+The \texttt{gen\_cubic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a cubic curve in the first two dimensions of a \(p\)-dimensional space.
+This function creates a more complex curvilinear structure than a simple parabola.
+
+The shape is generated using a third-degree raw polynomial basis expansion.
+Let \(X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])\) be the base input.
+A cubic transformation with vertical jitter is used to define \(X\_2 = X\_1 + X\_1^2 - X\_1^3 + \varepsilon\_2\$, where \$\varepsilon\_2 \sim \mathcal{U}(0, 0.5)\).
+
+When \(p > 2\), the remaining dimensions consist of independent Gaussian noise features generated via \texttt{gen\_noisedims()}:
+
+\[
+X\_j \sim \mathcal{N}(0, 0.1^2)\$ for \$j = 3, \dots, p.
+\]
+
+\begin{verbatim}
+cubic <- gen_cubic(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cubic` data.}\label{fig:fig-cubic-proj}
+\end{figure}
 
@@
 
+The \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a rectangular base.
+The pyramid is embedded in a \(p\)-dimensional space, with a tip at height zero and base at height \(h\).
+The shape tapers linearly from the base dimensions \((l\_x, l\_y)\) to a smaller rectangular section with side lengths \((2r\_t, 2r\_t)\) at the tip.
+Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.
+
+Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points.
+The final dimension \(x\_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\):
+
+\[
+x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)
+\].
+
+Let \(r\_x(z)\) and \(r\_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\):
+
+\[
+r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}, \quad
+r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}
+\].
+
+The first three coordinates are then defined as:
+
+\[
+x_1 \sim \mathcal{U}(-r_x(z),\ r_x(z)), \quad
+x_2 \sim \mathcal{U}(-r_y(z),\ r_y(z)), \quad
+x_3 \sim \mathcal{U}(-r_x(z),\ r_x(z))
+\].
+
+For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(x\_4\) to \(x\_{p-1}\)) are tapered toward zero with decreasing height and generated as:
+
+\[
+x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
+\]
+
+This tapering ensures that the structure narrows in all directions as it approaches the tip.
+
+\begin{verbatim}
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrrect-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` data.}\label{fig:fig-pyrrect-proj}
+\end{figure}
+
+The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section.
+The structure is embedded in a \(p\)-dimensional space, with the tip located at height 0 and the base at height \(h\).
+The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
+
+Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points.
+The final coordinate \$x\_p\$ encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):
+
+\[
+x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
+\]
+
+Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\):
+
+\[
+r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
+\]
+
+A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section:
+
+\[
+u, v \sim \mathcal{U}(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
+\]
+
+The first three coordinates (triangle plane) are then:
+
+\[
+x_1 = r(z)(1 - u - v), \quad
+x_2 = r(z) \cdot u, \quad
+x_3 = r(z) \cdot v
+\]
+
+For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip:
+
+\[
+x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
+\]
+
+The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section.
+The structure is embedded in a \(p\)-dimensional space, with the tip located at height 0 and the base at height \(h\).
+The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
+
+Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points.
+The final coordinate \(x\_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):
+
+\[
+x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
+\]
+
+Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\):
+
+\[
+r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
+\]
+
+A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section:
+
+\[
+u, v \sim \mathcal{U}(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
+\]
+
+The first three coordinates (triangle plane) are then:
+
+\[
+x_1 = r(z)(1 - u - v), \quad
+x_2 = r(z) \cdot u, \quad
+x_3 = r(z) \cdot v
+\]
+
+For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip:
+
+\[
+x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
+\]
+
+The result is a tapered, high-dimensional triangle-based pyramid with more points near the apex and greater width at the base.
+
+\begin{verbatim}
+pyrtri <- gen_pyrtri(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrtri-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrtri` data.}\label{fig:fig-pyrtri-proj}
+\end{figure}
+
+The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \$n\$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
+The pyramid extends vertically from height \(z = 0\) (tip) to height \(z = h\) (base), embedded in a \(p\)-dimensional space.
+The distribution concentrates more points near the base, and the hexagonal spread tapers toward the tip.
+
+Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points. The final coordinate \(x\_p\) encodes the height \(z\):
+
+\[
+x_p = z \sim \mathcal{U}(0, h).
+\]
+
+The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r\_b\):
+
+\[
+r(z) = r_b \cdot \left(1 - \frac{z}{h}\right).
+\]
+
+Each point is placed within a regular hexagon in the plane \((x\_1, x\_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}\) and a uniformly random radial scaling factor:
+
+\[
+\theta \sim \text{Uniform sample from 6 hexagon angles}, \quad
+r_{\text{point}} \sim \sqrt{\mathcal{U}(0, 1)}.
+\]
+
+Then, the first two coordinates are:
+
+\[
+x_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta), \quad
+x_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)
+\]
+
+For dimensions \(j = 3, \dots, p-1\), values are tapered toward zero near the tip:
+
+\[
+x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
+\]
+
+This results in a smooth, star-like pyramid in \(p\)-dimensional space, with a broad hexagonal base and narrow tip.
+
+\begin{verbatim}
+pyrstar <- gen_pyrstar(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrstar-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrstar` data.}\label{fig:fig-pyrstar-proj}
+\end{figure}
+
+The \texttt{gen\_pyrholes(n,\ p)} function generates \(n\) points embedded in a \(p\)-dimensional simplex using a chaotic attractor-like midpoint algorithm.
+The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a ``Sierpinski-like pyramid'' in high dimensions.
+
+Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points. The generation process begins with an initial point \(T\_0 \in [0, 1]^p\) drawn from a uniform distribution:
+
+\[
+T_0 \sim \mathcal{U}(0, 1)^p.
+\]
+
+Let \(C\_1, C\_2, \dots, C\_{p+1}\) denote the corner vertices of a \(p\)-dimensional simplex.
+At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T\_{i-1}\) and a randomly selected vertex \(C\_k\):
+
+\[
+T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}.
+\]
+
+This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points. The points remain bounded inside the convex hull of the simplex.
+
+The final output is a \(n \times p\) matrix where each row represents a point:
+
+\[
+X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}.
+\]
+
+\begin{verbatim}
+pyrholes <- gen_pyrholes(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrholes-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrholes` data.}\label{fig:fig-pyrholes-proj}
+\end{figure}
+
 \subsubsection{S-curve}\label{s-curve}
@@
 \end{table}
+
+To simulate an S-curve structure in a higher-dimensional space, we define the function \texttt{gen\_scurve(n,\ p)}, which generates \texttt{n} observations in \texttt{p} dimensions.
+
+The \(3\text{-}D\) geometry is constructed by introducing a latent parameter, \(\theta \sim U\left(-\frac{3\pi}{2}, \frac{3\pi}{2}\right)\).
+This parameter controls the curvature of the manifold.
+The first three dimensions form the S-curve structure:
+
+\[
+  X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta) \cdot (\cos(\theta) - 1).
+\]
+
+This configuration creates a horizontally curled shape in \((X_1, X_3)\), with additional band thickness in the \(X_2\) direction.
+
+For \(p > 3\), additional noise dimensions are appended introducing structured, wavy perturbations.
+
+\begin{verbatim}
+scurve <- gen_scurve(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` data.}\label{fig:fig-scurve-proj}
+\end{figure}
+
+To simulate a variant of the S-curve structure with a removed region (``hole''), we define the function \texttt{gen\_scurvehole(n,\ p)}.
+This function produces \texttt{n} observations in a \(p\)-dimensional space where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
+A subset of observations near a designated anchor point is excluded to introduce a hole in the manifold.
+
+To simulate missing regions on the manifold, a fixed anchor point \((0, 1, 0, \ldots)\) is defined in \(p\)-dimensional space.
+All observations within a Euclidean distance of \(\sqrt{0.3} \approx 0.5477\) from the anchor are removed:
+
+\[
+\text{remove if} \quad \sum_{j=1}^p (X_j - a_j)^2 \leq 0.3
+\].
+
+This exclusion creates a hole in the manifold centered near the middle vertical region of the S-curve.
+
+\begin{verbatim}
+scurvehole <- gen_scurvehole(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurvehole-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurvehole` data.}\label{fig:fig-scurvehole-proj}
+\end{figure}
 
@@
 
+The function \texttt{gen\_circle(n,\ p)} generates a \(p\)-dimensional dataset of \(n\) observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale.
+
+A latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\). Coordinates in the first two dimensions represent a perfect circle on the plane:
+
+\[
+X_1 = \cos(\theta), \quad X_2 = \sin(\theta)
+\]
+
+For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
+The first component is a scaling factor that decreases with the dimension index, defined as \(\text{scale}_j = \sqrt{(0.5)^{j-2}}\) for \(j = 3, \dots, p\).
+The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula \(\phi_j = (j - 2) \cdot \frac{\pi}{2p}\).
+
+Each additional dimension is computed as:
+
+\[
+X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p
+\]
+
+This structure retains a one-dimensional manifold (the circle) but allows it to be embedded non-linearly and decorrelatively in higher dimensions.
+The decreasing scale ensures that higher dimensions add less variation, preserving the dominance of the circular shape while adding complexity.
+
+\begin{verbatim}
+circle <- gen_circle(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-circle-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` data.}\label{fig:fig-circle-proj}
+\end{figure}
+
+The \texttt{gen\_curvycycle(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations.
+The structure forms a one-dimensional nonlinear cycle embedded in higher-dimensional space.
+
+A latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
+The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle.'' In this configuration, \(X_1 = \cos(\theta)\) represents horizontal oscillation, while \(X_2 = \sqrt{3}/3 + \sin(\theta)\) introduces a vertical offset to avoid centering the curve at the origin.
+Additionally, \(X_3 = \frac{1}{3} \cdot \cos(3\theta)\) introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
+
+Together, these define a periodic, non-trivial, closed curve in \(3\text{-}D\) with internal folds that produce a more complex geometry than a standard circle or ellipse.
+
+For dimensions \(X_4\) through \(X_p\), additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
+The scaling factor is defined as \(\text{scale}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from 4 to \(p\), which means that the amplitude decreases as the dimension increases.
+Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j \cdot \sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2) \cdot \frac{\pi}{2p}\).
+
+This introduces decorrelated, low-variance sinusoidal variations across dimensions, retaining the cyclical nature while embedding the data in higher-dimensional space with a coherent but subtle structure.
+
+\begin{verbatim}
+curvycycle <- gen_curvycycle(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-curvycycle-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycycle` data.}\label{fig:fig-curvycycle-proj}
+\end{figure}
+
+The \texttt{gen\_unifsphere(n,\ p,\ r)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on the surface of a \(3\text{-}D\) sphere of radius \(r\), with additional Gaussian noise dimensions added when \(p > 3\).
+
+Each observation lies on the surface of a sphere in \(3\text{-}D\), constructed by generating \(u \sim U(-1, 1)\) which represents the cosine of the polar angle \(\phi\) and \(\theta \sim U(0, 2\pi)\) which represents the azimuthal angle.
+
+The corresponding Cartesian coordinates are calculated as;
+
+\[
+X_1 = r \cdot \sqrt{1 - u^2} \cdot \cos(\theta),
+\]
+\[
+X_2 = r \cdot \sqrt{1 - u^2} \cdot \sin(\theta),
+\]
+\[
+X_3 = r \cdot u,
+\]
+
+which gives points uniformly distributed on the surface of a \(3\text{-}D\) sphere (not within).
+
+For \(p > 3\), additional dimensions \(X_4\) to \(X_p\) are generated as low-variance Gaussian noise:
+
+\[
+X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 4, \dots, p.
+\]
+
+This extends the \(3\text{-}D\) spherical manifold into \(p\)-dimensional space with orthogonal Gaussian noise, preserving the spherical structure while allowing for testing in higher-dimensional settings.
+
+\begin{verbatim}
+unifsphere <- gen_unifsphere(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-unifsphere-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `unifsphere` data.}\label{fig:fig-unifsphere-proj}
+\end{figure}
+
+The \texttt{gen\_gridedsphere(n,\ p)} function generates a \(p\)-dimensional dataset of approximately \(n\) observations evenly distributed on the surface of a \(3\text{-}D\) unit sphere, with optional Gaussian noise dimensions when \(p > 3\).
+
+The base structure consists of a \(3\text{-}D\) spherical surface created using a regular grid in spherical coordinates, where \(\theta \in [0, 2\pi]\) represents the azimuthal angle (longitude) and \(\phi \in [0, \pi]\) denotes the polar angle (co-latitude).
+
+The number of grid steps along each dimension is determined by factoring \(n\) into two approximately equal integers via \texttt{gen\_nproduct(n,\ p\ =\ 2)}.
+
+Each point on the sphere is computed using the spherical-to-Cartesian transformation:
+
+\[
+X_1 = \sin(\phi) \cdot \cos(\theta),
+\]
+\[
+X_2 = \sin(\phi) \cdot \sin(\theta),
+\]
+\[
+X_3 = \cos(\phi).
+\]
+
+This forms a structured grid of points on the unit sphere, ideal for studying how regular geometric manifolds are transformed under nonlinear dimensionality reduction.
+
+For \(p > 3\), the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions:
+
+\[
+X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 4, \dots, p.
+\]
+
+These dimensions represent low-variance orthogonal noise, preserving the spherical shape while testing robustness in high-dimensional settings.
+
+\begin{verbatim}
+gridedsphere <- gen_gridedsphere(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-gridedsphere-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridedsphere` data.}\label{fig:fig-gridedsphere-proj}
+\end{figure}
+
+The \texttt{gen\_clusteredspheres(n,\ k,\ p,\ r,\ loc)} function generates a synthetic dataset of \(n_1 + k \cdot n_2\) observations in \(p\)-dimensional space, consisting of one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location.
+
+A large uniform sphere centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\)-dimensional sphere with a radius of \(r_1\).
+The sampling is executed using the function \texttt{gen\_unifsphere(n\_1,\ p,\ r\_1)}, which generates the desired points in the specified dimensional space.
+
+In generation of \(k\) smaller uniform spheres, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
+These spheres are positioned at distinct random locations in \(p\)-space, with the center of each sphere being drawn from a normal distribution \(\mathcal{N}(0, \texttt{loc}^2 I_p)\).
+
+Points on spheres are generated using the standard hyperspherical method, which involves sampling \(u \sim U(-1, 1)\) to determine the cosine of the polar angle, and sampling \(\theta \sim U(0, 2\pi)\) to determine the azimuthal angle (for 3D).
+
+Each observation is classified by cluster, with labels such as ``big'' for the large central sphere and ``small\_1'' to ``small\_k'' for the smaller spheres.
+
+\begin{verbatim}
+clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, p = 4, r = c(15, 3),
+                                         loc = 10 / sqrt(3)) |>
+  dplyr::select(-cluster)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-clusteredspheres-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `clusteredspheres` data.}\label{fig:fig-clusteredspheres-proj}
+\end{figure}
+
+The \texttt{gen\_hemisphere(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on a \(4\text{-}D\) hemisphere, optionally extended with Gaussian noise in additional dimensions when \(p > 4\).
+
+Each observation is situated on a restricted \(4\text{-}D\) spherical surface, defined by spherical coordinates.
+The azimuthal angle \(\theta_1 \sim U(0, \pi)\) in the \((x_1, x_2)\) plane, while the elevation angle \(\theta_2 \sim U(0, \pi)\) in the \((x_2, x_3)\) plane.
+Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
+
+The coordinates are transformed into \(4\text{-}D\) Cartesian space:
+
+\[
+X_1 = \sin(\theta_1) \cdot \cos(\theta_2),
+\]
+\[
+X_2 = \sin(\theta_1) \cdot \sin(\theta_2),
+\]
+\[
+X_3 = \cos(\theta_1) \cdot \cos(\theta_3),
+\]
+\[
+X_4 = \cos(\theta_1) \cdot \sin(\theta_3).
+\]
+
+This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
+
+For \(p > 4\), additional dimensions \(X_5\) to \(X_p\) are added as low-variance Gaussian noise:
+
+\[
+X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \dots, p.
+\]
+
+These dimensions maintain the spherical structure while simulating embedding into higher-dimensional space with small orthogonal perturbations.
+
+\begin{verbatim}
+hemisphere <- gen_hemisphere(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-hemisphere-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `hemisphere` data.}\label{fig:fig-hemisphere-proj}
+\end{figure}
+
 \subsubsection{Swiss Roll}\label{swiss-roll}
 
-To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function \texttt{generate\_swiss\_roll(n,\ p)}, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ p,\ w)}, where \texttt{n} is the number of points, \texttt{p} is the total number of dimensions, and \texttt{w} is the vertical range in the third dimension.
 
-The function generates \texttt{n} points in a \texttt{p}-dimensional space, where the first two dimensions (\texttt{X\_1,\ X\_2}) define the primary Swiss roll shape using a parametric equation:
+The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape:
 
 \[
-X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad \text{where } t \sim U(0, 3\pi)
+X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)
 \]
 
-The third dimension (\texttt{X\_3}) introduces variation perpendicular to the roll, sampled uniformly from \([-1,1]\).
-Additional dimensions (\texttt{X\_4} to \texttt{X\_p}) extend the data structure by applying a \textbf{sinusoidal transformation} of the parameter \texttt{t}, ensuring continuity in higher-dimensional spaces:
+For \(p > 3\), the remaining \(p - 3\) dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.
 
-\[
-X_i = \frac{\sin(i t)}{i}, \quad \text{for } i \geq 4.
-\]
+\begin{verbatim}
+swissroll <- gen_swissroll(n = 1000, p = 4, w = c(-1, 1))
+\end{verbatim}
 
-This transformation ensures a gradual decay in variance across dimensions, mimicking real-world high-dimensional structures where later dimensions often capture subtler variations.
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-swissroll-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `swissroll` data.}\label{fig:fig-swissroll-proj}
+\end{figure}
 
@@
 \end{table}
+
+The \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise.
+
+The first two dimensions define a crescent-shaped curve using a semi-circular arc. Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
+
+The corresponding \(2\text{-}D\) coordinates are defined by:
+
+\[
+X_1 = \cos(\theta),
+\]
+\[
+X_2 = \sin(\theta).
+\]
+
+This forms a nonlinear curvilinear structure in the shape of a tilted crescent in the \(X_1\)-\(X_2\) plane.
+
+For \(p > 2\), additional dimensions \(X_3\) to \(X_p\) are generated using \texttt{gen\_wavydims1()}:
+
+\[
+X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim \mathcal{N}(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
+\]
+
+These noise dimensions are nonlinearly correlated with \(\theta\), producing structured, wave-like variation that aligns with the progression along the crescent curve.
+
+\begin{verbatim}
+crescent <- gen_crescent(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-crescent-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` data.}\label{fig:fig-crescent-proj}
+\end{figure}
+
+The \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\).
+
+The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension:
+
+Let \(\theta \sim \mathcal{U}(0, 3\pi)\) represent a random angle on a circular base and \(z \sim \mathcal{U}(0, h)\) represent the height along the cylinder.
+
+The coordinates are defined as:
+
+\begin{itemize}
+\tightlist
+\item
+  \(X_1 = \cos(\theta)\) (Circular base, x-axis),
+\item
+  \(X_2 = \sin(\theta)\) (Circular base, y-axis),
+\item
+  \(X_3 = z\) (Linear height),
+\item
+  \(X_4 = \sin(z)\) (Nonlinear curvy variation along height).
+\end{itemize}
+
+This forms a curvy cylindrical surface in \(4\text{-}D\), where the fourth dimension bends periodically along the height axis, resembling a helicoid or twisting wave along the cylinder.
+
+For \(p > 4\), dimensions \(X_5\) through \(X_p\) are generated as:
+
+\[
+X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \ldots, p.
+\]
+
+These dimensions represent unstructured, low-variance noise.
+
+\begin{verbatim}
+curvycylinder <- gen_curvycylinder(n = 1000, p = 4, h = 10)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-curvycylinder-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycylinder` data.}\label{fig:fig-curvycylinder-proj}
+\end{figure}
+
+The \texttt{gen\_sphericalspiral(n,\ p,\ spins)} function simulates a dataset of \(n\) observations that form a spiraling path on a spherical surface in the first four dimensions.
+When extended beyond \(4\text{-}D\), structured nonlinear noise dimensions are added to simulate more realistic high-dimensional manifolds.
+
+The first three dimensions represent points on a unit sphere.
+Let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\)be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
+
+Cartesian coordinates from spherical conversion:
+
+\begin{itemize}
+\tightlist
+\item
+  \(X_1 = \sin(\phi) \cdot \cos(\theta)\)
+\item
+  \(X_2 = \sin(\phi) \cdot \sin(\theta)\)
+\item
+  \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim \mathcal{U}(-0.5, 0.5)\) introduces vertical jitter.
+\item
+  \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
+\end{itemize}
+
+This generates a spherical spiral curve embedded in \(4\text{-}D\) space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
+
+If \(p > 4\), the function appends structured, non-linear noise via \texttt{gen\_wavydims2()}.
+These dimensions are functions of the first coordinate \(X_1\), introducing dependencies that preserve some geometric coherence.
+Each added dimension follows the form:
+
+\[
+X_j = s_j \cdot (-1)^{\lfloor j/2 \rfloor} \cdot X_1^{a_j} + \eta_j,
+\]
+
+where:
+\(a_j \in \{2, 3, 4, 5\}\) is a randomly chosen polynomial power,
+\(s_j \sim \mathcal{U}(0.5, 2)\) is a scale factor,
+\(\eta_j \sim \mathcal{U}(-\sigma_j, 2\sigma_j)\), with \(\sigma_j \sim \mathcal{U}(0, 0.05)\), adds mild randomness.
+
+\begin{verbatim}
+sphericalspiral <- gen_sphericalspiral(n = 1000, p = 4, spins = 1)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-sphericalspiral-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `sphericalspiral` data.}\label{fig:fig-sphericalspiral-proj}
+\end{figure}
+
+The \texttt{gen\_helicalspiral(n,\ p)} function generates a dataset of \(n\) observations forming a helical spiral, embedded in the first four dimensions of a \(p\)-dimensional space.
+Additional dimensions (if \(p > 4\)) are filled with structured noise to mimic high-dimensional complexity.
+
+The first four coordinates follow a \(3\text{-}D\) helix with an additional oscillating component. Let \(\theta \in [0, \frac{5\pi}{4}]\) be a sequence of angles controlling rotation around a circle.
+
+Cartesian coordinates:
+* \(X_1 = \cos(\theta)\): circular trajectory along the x-axis.
+
+\begin{itemize}
+\item
+  \(X_2 = \sin(\theta)\): circular trajectory along the y-axis.
+\item
+  \(X_3 = 0.05 \cdot \theta + \varepsilon_3\), with \(\varepsilon_3 \sim \mathcal{U}(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix.
+\item
+  \(X_4 = 0.1 \cdot \sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
+\end{itemize}
+
+This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
+
+When \(p > 4\), the remaining \(p - 4\) dimensions are populated using \texttt{gen\_noisedims()}. These are independent Gaussian noise dimensions:
+
+\[
+X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+\]
+
+\begin{verbatim}
+helicalspiral <- gen_helicalspiral(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-helicalspiral-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `helicalspiral` data.}\label{fig:fig-helicalspiral-proj}
+\end{figure}
+
+The \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of a \(p\)-dimensional space.
+The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a \(3\text{-}D\) conic helix.
+
+The shape is defined by parameter \(\theta \in [0, 2\pi \cdot \text{spins}]\), controlling the angular progression of the spiral. The mapping to Cartesian space is as follows:
+
+The Archimedean spiral in the horizontal plane is represented by;
+
+\(X_1 = \theta \cdot \cos(\theta)\) for radial expansion in x.
+\(X_2 = \theta \cdot \sin(\theta)\) for radial expansion in y.
+
+The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim \mathcal{U}(-0.1, 0.6).\)
+
+Spiral modulation in the fourth dimension is represented by \(X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim \mathcal{U}(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
+
+This results in a \(3\text{-}D\) spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
+
+For \(p > 4\), the dataset includes isotropic Gaussian noise in the remaining dimensions via \texttt{gen\_noisedims()}:
+
+\[
+X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+\]
+
+\begin{verbatim}
+conicspiral <- gen_conicspiral(n = 1000, p = 4, spins = 1)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-conicspiral-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `conicspiral` data.}\label{fig:fig-conicspiral-proj}
+\end{figure}
+
+The \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables.
+The key geometry lies in dimensions 1--4, where:
+
+\begin{itemize}
+\tightlist
+\item
+  \(X_1 \sim \mathcal{U}(0.1, 2)\): base variable (avoids zero to prevent division errors).
+\item
+  \(X_3 \sim \mathcal{U}(0.1, 0.8)\): independent auxiliary variable.
+\item
+  \(X_2 = \frac{hc}{X_1} + \text{non\_fac} \cdot \sin(X_1)\): non-linear combination of \textbf{hyperbolic and sinusoidal transformations}, creating sharp curvature and oscillation.
+\item
+  \(X_4 = \cos(\pi \cdot X_1) + \varepsilon\), with \(\varepsilon \sim \mathcal{U}(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+\end{itemize}
+
+These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
+
+If \(p > 4\), the remaining dimensions are populated with Gaussian noise using \texttt{gen\_noisedims()}:
+
+\[
+X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+\]
+
+\begin{verbatim}
+nonlinear <- gen_nonlinear(n = 1000, p = 4, hc = 1, non_fac = 0.5)
+\end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nonlinear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `nonlinear` data.}\label{fig:fig-nonlinear-proj}
+\end{figure}
 
@@
 
-\subsection{Assessing the performance of dimension reduction on different geometric structures in high-dimensions}\label{assessing-the-performance-of-dimension-reduction-on-different-geometric-structures-in-high-dimensions}
-
 This section illustrates the use of package by generating a synthetic dataset to evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
@@
 The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in \(4\text{-}D\) space (Figure \ref{fig:fig-highd-proj}).
+
+\begin{verbatim}
+positions <- geozoo::simplex(p=4)$points
+positions <- positions * 0.8
+
+## To generate data
+five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), p = 4, k = 5,
+                       loc = positions,
+                       scale = c(0.4, 0.35, 0.3, 1, 0.3),
+                       shape = c("helicalspiral", "hemisphere", "unifcube", 
+                                 "cone", "gaussian"),
+                       rotation = NULL,
+                       is_bkg = FALSE)
+\end{verbatim}
 
@@
 \begin{itemize}
-\tightlist
 \item
   Branching: These functions create a controlled environment for testing how effectively various algorithms preserve branching topology and continuity in their low-dimensional embeddings.
+\item
+  Scurve with a hole allowing for evaluation of how well algorithms handle incomplete manifolds or missing local structure.
+\item
+  clusteredsphere: This structure allows for cluster separation on curved manifolds in high-dimensional space and can be used to test the ability of NLDR methods and clustering algorithms to detect spherical clusters of different sizes and separations.
+\item
+  Mobius: The core geometric structure is a Mobius strip---a classic one-sided surface with a half-twist---useful for evaluating how well methods capture non-orientability and twisted manifolds.
+\item
+  Grided cube: This function is useful for assessing how algorithms preserve uniformly spaced data in high-dimensional spaces.
+\item
+  Pyrholes: This structure is useful for testing clustering and NLDR algorithms on non-convex and sparse high-dimensional shapes.
 \end{itemize}

### 2025-05-07 paper-cardinalR.Rmd
Try to fix the latex error

@@
   \usepackage{array}
+  \usepackage{hyperref}
+  \hypersetup{draft}
 output: 
@@
     keep_md: true
+    includes:
+      in_header: preamble.tex
 bibliography: paper-cardinalR.bib
@@
 
-More documentation of the package can be found at the web site <https://jayanilakshika.github.io/cardinalR/>.
+More documentation of the package can be found at the web site [https://jayanilakshika.github.io/cardinalR/](https://jayanilakshika.github.io/cardinalR/).
 
@@
 
-The code is available at <https://github.com/JayaniLakshika/cardinalR>, and source material for this paper is available at <https://github.com/JayaniLakshika/paper-cardinalR>.
+The code is available at [https://github.com/JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR), and source material for this paper is available at [https://github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
 

### 2025-05-07 paper-cardinalR.Rmd
Updated the new version

@@
 output: 
- rjtools::rjournal_article:
+ rjtools::rjournal_web_article:
     keep_md: true
-    includes:
-      in_header: preamble.tex
 bibliography: paper-cardinalR.bib
@@
 ---
+
+<!-- output:  -->
+<!--  rjtools::rjournal_article: -->
+<!--     keep_md: true -->
+<!--     includes: -->
+<!--       in_header: preamble.tex -->
 

### 2025-05-07 paper-cardinalR.Rmd
Updated the mathematical formula
Updated notification

@@
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  kable(caption = "The main arguments for gen_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
   column_spec(1, width = "4cm") |>
@@
 
-The `gen_expbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ exponential branches in 2D, with optional noise dimensions to embed the structure in a \$p\$-dimensional space.
+The `gen_expbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ exponential branches in $2\text{-}D$, with optional noise dimensions to embed the structure in a $p$-dimensional space.
 These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
@@
 
-$X\_1 \sim \mathcal{U}(-2, 2)$
-$X\_2 = \exp(\pm s\_i \cdot X\_1) + \epsilon$,
+$X_1 \sim \mathcal{U}(-2, 2)$
+$X_2 = \exp(\pm s_i \cdot X_1) + \epsilon$,
 
@@
 
-$s\_i \in [0.5, 2]$ is randomly sampled for each branch.
+$s_i \in [0.5, 2]$ is randomly sampled for each branch.
 
@@
 $$
-X\_j \sim \mathcal{N}(0, 0.1^2), \quad j = 3, \dots, p.
+X_j \sim \mathcal{N}(0, 0.1^2), \quad j = 3, \dots, p.
 $$
@@
 Branches 1 and 2 are initialized with fixed slopes and intercepts.
-The Branch 1 is generated from $X\_1 \sim \mathcal{U}(-2, 8)\$, \$X\_2 = 0.5 \cdot X\_1 + \epsilon$.
-The Branch 2 is generated from $X\_1 \sim \mathcal{U}(-6, 2)\$, \$X\_2 = -0.5 \cdot X\_1 + \epsilon$.
+The Branch 1 is generated from $X_1 \sim \mathcal{U}(-2, 8)\$, \$X_2 = 0.5 \cdot X_1 + \epsilon$.
+The Branch 2 is generated from $X_1 \sim \mathcal{U}(-6, 2)\$, \$X_2 = -0.5 \cdot X_1 + \epsilon$.
 $\epsilon \sim \mathcal{U}(0, 0.5)$.
@@
 
-$X\_1$ values span a short segment ($x\_{start}$ to $x\_{start} + 1$),
-
-$X\_2 = s\_i \cdot (X\_1 - x\_{start}) + y\_{start} + \epsilon$ where $s\_i$ is a sampled slope from a filtered set and $\epsilon \sim \mathcal{U}(0, 0.2)$.
+$X_1$ values span a short segment ($x_{start}$ to $x_{start} + 1$),
+
+$X_2 = s_i \cdot (X_1 - x_{start}) + y_{start} + \epsilon$ where $s_i$ is a sampled slope from a filtered set and $\epsilon \sim \mathcal{U}(0, 0.2)$.
 
@@
 $$
-X\_j \sim \mathcal{N}(0, 0.05^2)\$ for \$j = 3, \dots, p.
+X_j \sim \mathcal{N}(0, 0.05^2)\text{ for }j = 3, \dots, p.
 $$
@@
 
-Branch 1 is generated from $X\_1 \sim \mathcal{U}(0, 1)$, and $X\_2 = 0.1 \cdot X\_1 + 1 \cdot X\_1^2 + \epsilon$, where $\epsilon \sim \mathcal{U}(0, 0.05)$.
+Branch 1 is generated from $X_1 \sim \mathcal{U}(0, 1)$, and $X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon$, where $\epsilon \sim \mathcal{U}(0, 0.05)$.
 This produces a gently upward-curving parabola in the right half-plane.
 
-Branch 2 is generated from $X\_1 \sim \mathcal{U}(-1, 0)$, and $X\_2 = 0.1 \cdot X\_1 - 2 \cdot X\_1^2 + \epsilon$, where $\epsilon \sim \mathcal{U}(0, 0.05)$.
+Branch 2 is generated from $X_{1} \sim \mathcal{U}(-1, 0)$, and $X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon$, where $\epsilon \sim \mathcal{U}(0, 0.05)$.
 This creates a steeper, leftward-facing curve in the left half-plane.
@@
 Branches $3$ to $k$** are added iteratively.
-Each new branch begins at a randomly chosen point within a restricted horizontal band: $X\_1 \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X\_1$:$X\_1 \in [x\_{\text{start}}, x\_{\text{start}} + 1]$, and has the structure:
+Each new branch begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X_{1}$:$X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure:
 
@@
 
-Here, $s\_i$ is a sampled scale factor (slope-like term) controlling curvature, and $(x\_{\text{start}}, y\_{\text{start}})$ is the starting point from the existing structure.
+Here, $s_i$ is a sampled scale factor (slope-like term) controlling curvature, and $(x_{\text{start}}, y_{\text{start}})$ is the starting point from the existing structure.
 
@@
 If $k > \binom{p}{2}$, additional pairs are sampled with replacement to reach $k$ total branches.
-Each selected pair $(x\_{i1}, x\_{i2})$ defines the $2\text{-}D$ plane for branch $i$.
-
-For branch $i$, $n\_i$ points are generated, where $\sum\_{i=1}^k n\_i = n$. The structure follows:
+Each selected pair $(x_{i1}, x_{i2})$ defines the $2\text{-}D$ plane for branch $i$.
+
+For branch $i$, $n_i$ points are generated, where $\sum_{i=1}^k n_i = n$. The structure follows:
 
@@
 
-where $s\_i$ is a scale factor controlling the slope.
-When $k \leq \binom{p}{2}$, $s\_i = 1$.
-When sampling with replacement, $s\_i$ is drawn from the set ${1, 1.5, 2, \dots, 8}$ in increments of $0.5$.
+where $s_i$ is a scale factor controlling the slope. When $k \leq \binom{p}{2}$, $s_i = 1$. When sampling with replacement, $s_i$ is drawn from the set ${1, 1.5, 2, \dots, 8}$ in increments of $0.5$.
 
@@
 Let $\binom{p}{2}$ denote the number of unique $2\text{-}D$ subspace combinations.
-When $k \leq \binom{p}{2}$, $k$ distinct subspace pairs $(x\_{i1}, x\_{i2})$ are sampled without replacement.
+When $k \leq \binom{p}{2}$, $k$ distinct subspace pairs $(x_{i1}, x_{i2})$ are sampled without replacement.
 Otherwise, combinations are selected with replacement to reach $k$ total branches.
-For each branch $i = 1, \dots, k$, a random scale factor $s\_i$ is used to vary the curvature:
-
-* If $k \leq \binom{p}{2}$: $s\_i = 1 $
-* If $k > \binom{p}{2}$: $s\_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}$
-
-Each branch contains $n\_i$ points such that $\sum\_{i=1}^k n\_i = n$, where the vector $(n\_1, \dots, n\_k)$ is randomly drawn using the helper function `gen_nsum()` to partition $n$.
-
-Within its assigned subspace $(x\_{i1}, x\_{i2})$, branch $i$ is defined by:
+For each branch $i = 1, \dots, k$, a random scale factor $s_i$ is used to vary the curvature:
+
+* If $k \leq \binom{p}{2}$: $s_i = 1 $
+* If $k > \binom{p}{2}$: $s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}$
+
+Each branch contains $n_i$ points such that $\sum_{i=1}^k n_i = n$, where the vector $(n_1, \dots, n_k)$ is randomly drawn using the helper function `gen_nsum()` to partition $n$.
+
+Within its assigned subspace $(x_{i1}, x_{i2})$, branch $i$ is defined by:
 
@@
 
-This forms a smooth downward-opening parabola in the plane defined by $(x\_{i1}, x\_{i2})$, with the degree of curvature controlled by $s\_i$.
+This forms a smooth downward-opening parabola in the plane defined by $(x_{i1}, x_{i2})$, with the degree of curvature controlled by $s_i$.
 
@@
 
-Each variable $X\_j$ represents integer grid coordinates (e.g., 1, 2, 3, ...) in dimension $j$:
-
-$$
-X\_j \in {1, \dots, k\_j},\text{ where }\prod\_{j=1}^p k\_j \approx n.
+Each variable $X_j$ represents integer grid coordinates (e.g., 1, 2, 3, ...) in dimension $j$:
+
+$$
+X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n.
 $$
@@
 
-Points are sampled independently from $X\_1, X\_2, X\_3 \sim \mathcal{U}(-0.5, 0.5)$ of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.
+Points are sampled independently from $X_1, X_2, X_3 \sim \mathcal{U}(-0.5, 0.5)$ of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.
 
@@
 $$
-X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
 $$
@@
 
-The initial points are drawn from a uniform cube centered at the origin: $X\_1, X\_2, X\_3 \sim \mathcal{U}(-0.5, 0.5)$.
+The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim \mathcal{U}(-0.5, 0.5)$.
 
@@
 
-Remove if $\sqrt{X\_1^2 + X\_2^2 + X\_3^2} \leq 0.5$.
+Remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$.
 
@@
 $$
-X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
 $$.
@@
 $$
-X\_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
-$$
-$$
-X\_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
-$$
-$$
-X\_3 = \frac{w}{2} \sin(\theta/2).
+X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
+$$
+$$
+X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
+$$
+$$
+X_3 = \frac{w}{2} \sin(\theta/2).
 $$
@@
 The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation.
-Let $X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the independent variable.
-A raw polynomial basis of degree 2 is used to form $X\_2 = X\_1 - X\_1^2 + \varepsilon\_2$, where $\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.
+Let $X_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the independent variable.
+A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim \mathcal{U}(0, 0.5)$.
 
@@
 The shape is generated using a third-degree raw polynomial basis expansion.
-Let $X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the base input.
-A cubic transformation with vertical jitter is used to define $X\_2 = X\_1 + X\_1^2 - X\_1^3 + \varepsilon\_2\$, where \$\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.
+Let $X_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the base input.
+A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\$, where \$\varepsilon_2 \sim \mathcal{U}(0, 0.5)$.
 
@@
 $$
-X\_j \sim \mathcal{N}(0, 0.1^2)\$ for \$j = 3, \dots, p.
+X_j \sim \mathcal{N}(0, 0.1^2)\$ for \$j = 3, \dots, p.
 $$
@@
 The pyramid is embedded in a $p$-dimensional space, with a tip at height zero and base at height $h$.
-The shape tapers linearly from the base dimensions $(l\_x, l\_y)$ to a smaller rectangular section with side lengths $(2r\_t, 2r\_t)$ at the tip.
+The shape tapers linearly from the base dimensions $(l_x, l_y)$ to a smaller rectangular section with side lengths $(2r_t, 2r_t)$ at the tip.
 Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.
 
-Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The final dimension $x\_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$:
+Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final dimension $x_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$:
 
@@
 
-Let $r\_x(z)$ and $r\_y(z)$ denote the half-widths of the rectangular cross-section at height $z$:
+Let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$:
 
@@
 
-For $p > 3$, the remaining $p - 4$ dimensions (i.e., $x\_4$ to $x\_{p-1}$) are tapered toward zero with decreasing height and generated as:
+For $p > 3$, the remaining $p - 4$ dimensions (i.e., $x_4$ to $x_{p-1}$) are tapered toward zero with decreasing height and generated as:
 
@@
 
-Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points.
-The final coordinate \$x\_p\$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:
+Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points.
+The final coordinate \$x_p\$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:
 
@@
 
-Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points.
-The final coordinate $x\_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:
+Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final coordinate $x_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:
 
@@
 
-Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The final coordinate $x\_p$ encodes the height $z$:
+Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final coordinate $x_p$ encodes the height $z$:
 
@@
 
-The radius at height $z$ scales linearly from zero (tip) to the base radius $r\_b$:
+The radius at height $z$ scales linearly from zero (tip) to the base radius $r_b$:
 
@@
 
-Each point is placed within a regular hexagon in the plane $(x\_1, x\_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor:
+Each point is placed within a regular hexagon in the plane $(x_1, x_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor:
 
@@
 
-Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T\_0 \in [0, 1]^p$ drawn from a uniform distribution:
+Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution:
 
@@
 
-Let $C\_1, C\_2, \dots, C\_{p+1}$ denote the corner vertices of a $p$-dimensional simplex.
-At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T\_{i-1}$ and a randomly selected vertex $C\_k$:
+Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p$-dimensional simplex.
+At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$:
 
@@
 $$
-X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
 $$
@@
 * $X_3 \sim \mathcal{U}(0.1, 0.8)$: independent auxiliary variable.
-* $X_2 = \frac{hc}{X_1} + \text{non\_fac} \cdot \sin(X_1)$: non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
+* $X_2 = \frac{hc}{X_1} + \text{non_fac} \cdot \sin(X_1)$: non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
 * $X_4 = \cos(\pi \cdot X_1) + \varepsilon$, with $\varepsilon \sim \mathcal{U}(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.

### 2025-05-08 paper-cardinalR.Rmd
Added the package list
Added package references
Updated the distributions

@@
 
-$X_1 \sim \mathcal{U}(-2, 2)$
+$X_1 \sim U(-2, 2)$
 $X_2 = \exp(\pm s_i \cdot X_1) + \epsilon$,
@@
 where:
-$\epsilon \sim \mathcal{U}(0, 0.1)$ adds local jitter,
+$\epsilon \sim U(0, 0.1)$ adds local jitter,
 
@@
 $$
-X_j \sim \mathcal{N}(0, 0.1^2), \quad j = 3, \dots, p.
+X_j \sim N(0, 0.1^2), \quad j = 3, \dots, p.
 $$
@@
 Branches 1 and 2 are initialized with fixed slopes and intercepts.
-The Branch 1 is generated from $X_1 \sim \mathcal{U}(-2, 8)\$, \$X_2 = 0.5 \cdot X_1 + \epsilon$.
-The Branch 2 is generated from $X_1 \sim \mathcal{U}(-6, 2)\$, \$X_2 = -0.5 \cdot X_1 + \epsilon$.
-$\epsilon \sim \mathcal{U}(0, 0.5)$.
+The Branch 1 is generated from $X_1 \sim U(-2, 8)\$, \$X_2 = 0.5 \cdot X_1 + \epsilon$.
+The Branch 2 is generated from $X_1 \sim U(-6, 2)\$, \$X_2 = -0.5 \cdot X_1 + \epsilon$.
+$\epsilon \sim U(0, 0.5)$.
 
@@
 
-$X_2 = s_i \cdot (X_1 - x_{start}) + y_{start} + \epsilon$ where $s_i$ is a sampled slope from a filtered set and $\epsilon \sim \mathcal{U}(0, 0.2)$.
+$X_2 = s_i \cdot (X_1 - x_{start}) + y_{start} + \epsilon$ where $s_i$ is a sampled slope from a filtered set and $\epsilon \sim U(0, 0.2)$.
 
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2)\text{ for }j = 3, \dots, p.
+X_j \sim N(0, 0.05^2)\text{ for }j = 3, \dots, p.
 $$
@@
 
-Branch 1 is generated from $X_1 \sim \mathcal{U}(0, 1)$, and $X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon$, where $\epsilon \sim \mathcal{U}(0, 0.05)$.
+Branch 1 is generated from $X_1 \sim U(0, 1)$, and $X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
 This produces a gently upward-curving parabola in the right half-plane.
 
-Branch 2 is generated from $X_{1} \sim \mathcal{U}(-1, 0)$, and $X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon$, where $\epsilon \sim \mathcal{U}(0, 0.05)$.
+Branch 2 is generated from $X_{1} \sim U(-1, 0)$, and $X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
 This creates a steeper, leftward-facing curve in the left half-plane.
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2), \quad \text{for } j = 3, \dots, p.
+X_j \sim N(0, 0.05^2), \quad \text{for } j = 3, \dots, p.
 $$
@@
 $$
-x_{i1} \sim \mathcal{U}(0, 2), \quad
-x_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim \mathcal{U}(0, 0.5),
+x_{i1} \sim U(0, 2), \quad
+x_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5),
 $$
@@
 $$
-x_j \sim \mathcal{N}(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
+x_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
 $$
@@
 $$
-x_{i1} \sim \mathcal{U}(0, 2), \quad
-x_{i2} = -s_i \cdot x_{i1}^2 + \epsilon, \quad \epsilon \sim \mathcal{U}(0, 0.5)
+x_{i1} \sim U(0, 2), \quad
+x_{i2} = -s_i \cdot x_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)
 $$
@@
 
-Points are sampled independently from $X_1, X_2, X_3 \sim \mathcal{U}(-0.5, 0.5)$ of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.
+Points are sampled independently from $X_1, X_2, X_3 \sim U(-0.5, 0.5)$ of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.
 
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
 $$
@@
 
-The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim \mathcal{U}(-0.5, 0.5)$.
+The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim U(-0.5, 0.5)$.
 
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
 $$.
@@
 $$
-X_i \sim \mathcal{N}_p(\boldsymbol{0}, \Sigma)
+X_i \sim N_p(\boldsymbol{0}, \Sigma)
 $$
@@
 
-$\text{scale}_i \sim \mathcal{U}(-10, 10)$: randomly chosen direction and stretch.
-
-$\text{shift}_i \sim \mathcal{U}(-300, 300)$: large offset to spread features apart.
-
-$\epsilon \sim \mathcal{N}(0, 0.03 \cdot n)$: Gaussian noise scaled to dataset size to introduce mild irregularity.
+$\text{scale}_i \sim U(-10, 10)$: randomly chosen direction and stretch.
+
+$\text{shift}_i \sim U(-300, 300)$: large offset to spread features apart.
+
+$\epsilon \sim N(0, 0.03 \cdot n)$: Gaussian noise scaled to dataset size to introduce mild irregularity.
 
@@
 $$
-x_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+x_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
 $$
@@
 The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation.
-Let $X_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the independent variable.
-A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim \mathcal{U}(0, 0.5)$.
+Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the independent variable.
+A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
 
@@
 The shape is generated using a third-degree raw polynomial basis expansion.
-Let $X_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the base input.
-A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\$, where \$\varepsilon_2 \sim \mathcal{U}(0, 0.5)$.
+Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the base input.
+A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\$, where \$\varepsilon_2 \sim U(0, 0.5)$.
 
@@
 $$
-X_j \sim \mathcal{N}(0, 0.1^2)\$ for \$j = 3, \dots, p.
+X_j \sim N(0, 0.1^2)\$ for \$j = 3, \dots, p.
 $$
@@
 $$
-x_1 \sim \mathcal{U}(-r_x(z),\ r_x(z)), \quad
-x_2 \sim \mathcal{U}(-r_y(z),\ r_y(z)), \quad
-x_3 \sim \mathcal{U}(-r_x(z),\ r_x(z))
+x_1 \sim U(-r_x(z),\ r_x(z)), \quad
+x_2 \sim U(-r_y(z),\ r_y(z)), \quad
+x_3 \sim U(-r_x(z),\ r_x(z))
 $$.
@@
 $$
-x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
+x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
 $$
@@
 $$
-u, v \sim \mathcal{U}(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
+u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
 $$
@@
 $$
-x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
+x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
 $$
@@
 $$
-x_p = z \sim \mathcal{U}(0, h).
+x_p = z \sim U(0, h).
 $$
@@
 \theta \sim \text{Uniform sample from 6 hexagon angles}, \quad
-r_{\text{point}} \sim \sqrt{\mathcal{U}(0, 1)}.
+r_{\text{point}} \sim \sqrt{U(0, 1)}.
 $$
@@
 $$
-x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
+x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
 $$
@@
 $$
-T_0 \sim \mathcal{U}(0, 1)^p.
+T_0 \sim U(0, 1)^p.
 $$
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 4, \dots, p.
+X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p.
 $$
@@
 In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
-These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $\mathcal{N}(0, \texttt{loc}^2 I_p)$.
+These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
 
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \dots, p.
+X_j \sim N(0, 0.05^2),\text{ for }j = 5, \dots, p.
 $$
@@
 $$
-X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim \mathcal{N}(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
+X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
 $$
@@
 
-Let $\theta \sim \mathcal{U}(0, 3\pi)$ represent a random angle on a circular base and $z \sim \mathcal{U}(0, h)$ represent the height along the cylinder.
+Let $\theta \sim U(0, 3\pi)$ represent a random angle on a circular base and $z \sim U(0, h)$ represent the height along the cylinder.
 
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \ldots, p.
+X_j \sim N(0, 0.05^2),\text{ for }j = 5, \ldots, p.
 $$
@@
   * $X_2 = \sin(\phi) \cdot \sin(\theta)$
-  * $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim \mathcal{U}(-0.5, 0.5)$ introduces vertical jitter.
+  * $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter.
   * $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
@@
 $a_j \in \{2, 3, 4, 5\}$ is a randomly chosen polynomial power,
-$s_j \sim \mathcal{U}(0.5, 2)$ is a scale factor,
-$\eta_j \sim \mathcal{U}(-\sigma_j, 2\sigma_j)$, with $\sigma_j \sim \mathcal{U}(0, 0.05)$, adds mild randomness.
+$s_j \sim U(0.5, 2)$ is a scale factor,
+$\eta_j \sim U(-\sigma_j, 2\sigma_j)$, with $\sigma_j \sim U(0, 0.05)$, adds mild randomness.
 
@@
 
-  * $X_3 = 0.05 \cdot \theta + \varepsilon_3$, with $\varepsilon_3 \sim \mathcal{U}(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix.
+  * $X_3 = 0.05 \cdot \theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix.
 
@@
 $$
-X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
 $$
@@
 
-The growth pattern resembles a cone, with the height increasing according to $X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim \mathcal{U}(-0.1, 0.6).$
-
-Spiral modulation in the fourth dimension is represented by $X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim \mathcal{U}(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
+The growth pattern resembles a cone, with the height increasing according to $X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$
+
+Spiral modulation in the fourth dimension is represented by $X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
 
@@
 
-* $X_1 \sim \mathcal{U}(0.1, 2)$: base variable (avoids zero to prevent division errors).
-* $X_3 \sim \mathcal{U}(0.1, 0.8)$: independent auxiliary variable.
+* $X_1 \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors).
+* $X_3 \sim U(0.1, 0.8)$: independent auxiliary variable.
 * $X_2 = \frac{hc}{X_1} + \text{non_fac} \cdot \sin(X_1)$: non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
-* $X_4 = \cos(\pi \cdot X_1) + \varepsilon$, with $\varepsilon \sim \mathcal{U}(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+* $X_4 = \cos(\pi \cdot X_1) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 
@@
 
-# Discussion
+# Conclusion
 
@@
 
+These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], gtools [@gregory2023], dplyr [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].
 This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.

### 2025-05-15 paper-cardinalR.Rmd
Updated the .bib

@@
 
-The package can be installed from CRAN using
-
-```r
-install.packages("cardinalR")
-```
-
-and the development version can be installed from GitHub
+The development version can be installed from GitHub:
 
@@
 
-## Web site
-
-More documentation of the package can be found at the web site [https://jayanilakshika.github.io/cardinalR/](https://jayanilakshika.github.io/cardinalR/).
+<!-- ## Web site -->
+
+<!-- More documentation of the package can be found at the web site [https://jayanilakshika.github.io/cardinalR/](https://jayanilakshika.github.io/cardinalR/). -->
 
@@
 
-## Functions
+## Usage
 
@@
 
-These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], gtools [@gregory2023], dplyr [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].
+These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].
 This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.

### 2025-05-15 paper-cardinalR.Rmd
Updated text in
Removed data sets section

@@
 
-## Data sets
-
-The `cardinalR` package comes with several data sets that load with the package.
-These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:datasets-tb-html)', '\\@ref(tab:datasets-tb-pdf)'))`.
-
-```{r}
-datasets_tb <- tibble(dt = c("mobiusgau", 
-                             "mobiusgau_tsne1", 
-                             "mobiusgau_tsne2",
-                             "mobiusgau_tsne3",
-                             "mobiusgau_umap1",
-                             "mobiusgau_umap2",
-                             "mobiusgau_umap3"), 
-                      text = c("Simulated data with a Mobius and a Gaussian in 4-D space.",
-                               "The tSNE embedding with perplexity $15$ for mobiusgau.", 
-                               "The tSNE embedding with perplexity $30$ for mobiusgau.",
-                               "The tSNE embedding with perplexity $5$ for mobiusgau.",
-                               "The UMAP embedding with number of neighbors $15$ and minimum distance $0.1$ for mobiusgau.",
-                               "The UMAP embedding with number of neighbors $30$ and minimum distance $0.08$ for mobiusgau.", 
-                               "The UMAP embedding with number of neighbors $5$ and minimum distance $0.9$ for mobiusgau."))
-```
-
-```{r datasets-tb-html, eval=knitr::is_html_output()}
-datasets_tb |> 
-  kable(caption = "cardinalR data sets", col.names = c("data", "explanation")) 
-```
-
-```{r datasets-tb-pdf, eval=knitr::is_latex_output()}
-datasets_tb |> 
-  kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
 ## Usage
@@
 Branches 1 and 2 are initialized with fixed slopes and intercepts.
-The Branch 1 is generated from $X_1 \sim U(-2, 8)\$, \$X_2 = 0.5 \cdot X_1 + \epsilon$.
-The Branch 2 is generated from $X_1 \sim U(-6, 2)\$, \$X_2 = -0.5 \cdot X_1 + \epsilon$.
+The Branch 1 is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5 \cdot X_1 + \epsilon$.
+The Branch 2 is generated from $X_1 \sim U(-6, 2)$, $X_2 = -0.5 \cdot X_1 + \epsilon$.
 $\epsilon \sim U(0, 0.5)$.
 
-Branches 3 to $k$** are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
+Branches 3 to $k$ are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
 
@@
 
-Branches $3$ to $k$** are added iteratively.
+Branches $3$ to $k$ are added iteratively.
 Each new branch begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X_{1}$:$X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure:
@@
 
-* If $k \leq \binom{p}{2}$: $s_i = 1 $
+* If $k \leq \binom{p}{2}$: $s_i = 1$
 * If $k > \binom{p}{2}$: $s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}$
@@
 Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the base input.
-A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\$, where \$\varepsilon_2 \sim U(0, 0.5)$.
+A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
 

### 2025-05-19 paper-cardinalR.Rmd
Fixed the issu of rendering the pdf

@@
   \usepackage{array}
-  \usepackage{hyperref}
-  \hypersetup{draft}
 output: 
- rjtools::rjournal_web_article:
+ rjtools::rjournal_article:
     keep_md: true
@@
 ---
-
-<!-- output:  -->
-<!--  rjtools::rjournal_article: -->
-<!--     keep_md: true -->
-<!--     includes: -->
-<!--       in_header: preamble.tex -->
 
@@
 main_tb |> 
-  kable(caption = "The main arguments for gen_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
   column_spec(1, width = "4cm") |>
@@
 
-* $X_1 \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors).
-* $X_3 \sim U(0.1, 0.8)$: independent auxiliary variable.
-* $X_2 = \frac{hc}{X_1} + \text{non_fac} \cdot \sin(X_1)$: non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
-* $X_4 = \cos(\pi \cdot X_1) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+$X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors).
+$X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable.
+
+<!-- $$ -->
+<!-- X_{2} = \frac{hc}{X_{1}} + \text{non_fac} \cdot \sin(X_{1}) -->
+<!-- $$ -->
+
+:non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
+$X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 

### 2025-05-19 paper-cardinalR.tex
Fixed the issu of rendering the pdf

@@
 
-The package can be installed from CRAN using
-
-\begin{verbatim}
-install.packages("cardinalR")
-\end{verbatim}
-
-and the development version can be installed from GitHub
+The development version can be installed from GitHub:
 
@@
 
-\subsection{Web site}\label{web-site}
-
-More documentation of the package can be found at the web site \url{https://jayanilakshika.github.io/cardinalR/}.
-
-\subsection{Data sets}\label{data-sets}
-
-The \texttt{cardinalR} package comes with several data sets that load with the package. These are described in Table \ref{tab:datasets-tb-pdf}.
-
-\begin{table}
-
-\caption{\label{tab:datasets-tb-pdf}cardinalR data sets}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-data & explanation\\
-\midrule
-mobiusgau & Simulated data with a Mobius and a Gaussian in 4-D space.\\
-mobiusgau\_tsne1 & The tSNE embedding with perplexity \$15\$ for mobiusgau.\\
-mobiusgau\_tsne2 & The tSNE embedding with perplexity \$30\$ for mobiusgau.\\
-mobiusgau\_tsne3 & The tSNE embedding with perplexity \$5\$ for mobiusgau.\\
-mobiusgau\_umap1 & The UMAP embedding with number of neighbors \$15\$ and minimum distance \$0.1\$ for mobiusgau.\\
-\addlinespace
-mobiusgau\_umap2 & The UMAP embedding with number of neighbors \$30\$ and minimum distance \$0.08\$ for mobiusgau.\\
-mobiusgau\_umap3 & The UMAP embedding with number of neighbors \$5\$ and minimum distance \$0.9\$ for mobiusgau.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsection{Functions}\label{functions}
+\subsection{Usage}\label{usage}
 
@@
 
-The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in 2D, with optional noise dimensions to embed the structure in a \$p\$-dimensional space.
+The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in \(2\text{-}D\), with optional noise dimensions to embed the structure in a \(p\)-dimensional space.
 These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
@@
 
-\(X\_1 \sim \mathcal{U}(-2, 2)\)
-\(X\_2 = \exp(\pm s\_i \cdot X\_1) + \epsilon\),
+\(X_1 \sim U(-2, 2)\)
+\(X_2 = \exp(\pm s_i \cdot X_1) + \epsilon\),
 
 where:
-\(\epsilon \sim \mathcal{U}(0, 0.1)\) adds local jitter,
-
-\(s\_i \in [0.5, 2]\) is randomly sampled for each branch.
+\(\epsilon \sim U(0, 0.1)\) adds local jitter,
+
+\(s_i \in [0.5, 2]\) is randomly sampled for each branch.
 
@@
 \[
-X\_j \sim \mathcal{N}(0, 0.1^2), \quad j = 3, \dots, p.
+X_j \sim N(0, 0.1^2), \quad j = 3, \dots, p.
 \]
@@
 Branches 1 and 2 are initialized with fixed slopes and intercepts.
-The Branch 1 is generated from \(X\_1 \sim \mathcal{U}(-2, 8)\$, \$X\_2 = 0.5 \cdot X\_1 + \epsilon\).
-The Branch 2 is generated from \(X\_1 \sim \mathcal{U}(-6, 2)\$, \$X\_2 = -0.5 \cdot X\_1 + \epsilon\).
-\(\epsilon \sim \mathcal{U}(0, 0.5)\).
-
-Branches 3 to \(k\)** are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
-
-\(X\_1\) values span a short segment (\(x\_{start}\) to \(x\_{start} + 1\)),
-
-\(X\_2 = s\_i \cdot (X\_1 - x\_{start}) + y\_{start} + \epsilon\) where \(s\_i\) is a sampled slope from a filtered set and \(\epsilon \sim \mathcal{U}(0, 0.2)\).
+The Branch 1 is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5 \cdot X_1 + \epsilon\).
+The Branch 2 is generated from \(X_1 \sim U(-6, 2)\), \(X_2 = -0.5 \cdot X_1 + \epsilon\).
+\(\epsilon \sim U(0, 0.5)\).
+
+Branches 3 to \(k\) are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
+
+\(X_1\) values span a short segment (\(x_{start}\) to \(x_{start} + 1\)),
+
+\(X_2 = s_i \cdot (X_1 - x_{start}) + y_{start} + \epsilon\) where \(s_i\) is a sampled slope from a filtered set and \(\epsilon \sim U(0, 0.2)\).
 
@@
 \[
-X\_j \sim \mathcal{N}(0, 0.05^2)\$ for \$j = 3, \dots, p.
+X_j \sim N(0, 0.05^2)\text{ for }j = 3, \dots, p.
 \]
@@
 
-Branch 1 is generated from \(X\_1 \sim \mathcal{U}(0, 1)\), and \(X\_2 = 0.1 \cdot X\_1 + 1 \cdot X\_1^2 + \epsilon\), where \(\epsilon \sim \mathcal{U}(0, 0.05)\).
+Branch 1 is generated from \(X_1 \sim U(0, 1)\), and \(X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
 This produces a gently upward-curving parabola in the right half-plane.
 
-Branch 2 is generated from \(X\_1 \sim \mathcal{U}(-1, 0)\), and \(X\_2 = 0.1 \cdot X\_1 - 2 \cdot X\_1^2 + \epsilon\), where \(\epsilon \sim \mathcal{U}(0, 0.05)\).
+Branch 2 is generated from \(X_{1} \sim U(-1, 0)\), and \(X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
 This creates a steeper, leftward-facing curve in the left half-plane.
 
-Branches \(3\) to \(k\)** are added iteratively.
-Each new branch begins at a randomly chosen point within a restricted horizontal band: \(X\_1 \in [-0.15, 0.15]\) (to ensure connectivity with earlier branches), spans a unit-length interval on \(X\_1\):\(X\_1 \in [x\_{\text{start}}, x\_{\text{start}} + 1]\), and has the structure:
+Branches \(3\) to \(k\) are added iteratively.
+Each new branch begins at a randomly chosen point within a restricted horizontal band: \(X_{1} \in [-0.15, 0.15]\) (to ensure connectivity with earlier branches), spans a unit-length interval on \(X_{1}\):\(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), and has the structure:
 
@@
 
-Here, \(s\_i\) is a sampled scale factor (slope-like term) controlling curvature, and \((x\_{\text{start}}, y\_{\text{start}})\) is the starting point from the existing structure.
+Here, \(s_i\) is a sampled scale factor (slope-like term) controlling curvature, and \((x_{\text{start}}, y_{\text{start}})\) is the starting point from the existing structure.
 
@@
 \[
-X_j \sim \mathcal{N}(0, 0.05^2), \quad \text{for } j = 3, \dots, p.
+X_j \sim N(0, 0.05^2), \quad \text{for } j = 3, \dots, p.
 \]
@@
 If \(k > \binom{p}{2}\), additional pairs are sampled with replacement to reach \(k\) total branches.
-Each selected pair \((x\_{i1}, x\_{i2})\) defines the \(2\text{-}D\) plane for branch \(i\).
-
-For branch \(i\), \(n\_i\) points are generated, where \(\sum\_{i=1}^k n\_i = n\). The structure follows:
-
-\[
-x_{i1} \sim \mathcal{U}(0, 2), \quad
-x_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim \mathcal{U}(0, 0.5),
-\]
-
-where \(s\_i\) is a scale factor controlling the slope.
-When \(k \leq \binom{p}{2}\), \(s\_i = 1\).
-When sampling with replacement, \(s\_i\) is drawn from the set \({1, 1.5, 2, \dots, 8}\) in increments of \(0.5\).
+Each selected pair \((x_{i1}, x_{i2})\) defines the \(2\text{-}D\) plane for branch \(i\).
+
+For branch \(i\), \(n_i\) points are generated, where \(\sum_{i=1}^k n_i = n\). The structure follows:
+
+\[
+x_{i1} \sim U(0, 2), \quad
+x_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5),
+\]
+
+where \(s_i\) is a scale factor controlling the slope.
+When \(k \leq \binom{p}{2}\), \(s_i = 1\).
+When sampling with replacement, \(s_i\) is drawn from the set \({1, 1.5, 2, \dots, 8}\) in increments of \(0.5\).
 
@@
 \[
-x_j \sim \mathcal{N}(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
+x_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
 \]
@@
 Let \(\binom{p}{2}\) denote the number of unique \(2\text{-}D\) subspace combinations.
-When \(k \leq \binom{p}{2}\), \(k\) distinct subspace pairs \((x\_{i1}, x\_{i2})\) are sampled without replacement.
+When \(k \leq \binom{p}{2}\), \(k\) distinct subspace pairs \((x_{i1}, x_{i2})\) are sampled without replacement.
 Otherwise, combinations are selected with replacement to reach \(k\) total branches.
-For each branch \(i = 1, \dots, k\), a random scale factor \(s\_i\) is used to vary the curvature:
+For each branch \(i = 1, \dots, k\), a random scale factor \(s_i\) is used to vary the curvature:
 
@@
 \item
-  If \(k \leq \binom{p}{2}\): \$s\_i = 1 \$
-\item
-  If \(k > \binom{p}{2}\): \(s\_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}\)
+  If \(k \leq \binom{p}{2}\): \(s_i = 1\)
+\item
+  If \(k > \binom{p}{2}\): \(s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}\)
 \end{itemize}
 
-Each branch contains \(n\_i\) points such that \(\sum\_{i=1}^k n\_i = n\), where the vector \((n\_1, \dots, n\_k)\) is randomly drawn using the helper function \texttt{gen\_nsum()} to partition \(n\).
-
-Within its assigned subspace \((x\_{i1}, x\_{i2})\), branch \(i\) is defined by:
-
-\[
-x_{i1} \sim \mathcal{U}(0, 2), \quad
-x_{i2} = -s_i \cdot x_{i1}^2 + \epsilon, \quad \epsilon \sim \mathcal{U}(0, 0.5)
-\]
-
-This forms a smooth downward-opening parabola in the plane defined by \((x\_{i1}, x\_{i2})\), with the degree of curvature controlled by \(s\_i\).
+Each branch contains \(n_i\) points such that \(\sum_{i=1}^k n_i = n\), where the vector \((n_1, \dots, n_k)\) is randomly drawn using the helper function \texttt{gen\_nsum()} to partition \(n\).
+
+Within its assigned subspace \((x_{i1}, x_{i2})\), branch \(i\) is defined by:
+
+\[
+x_{i1} \sim U(0, 2), \quad
+x_{i2} = -s_i \cdot x_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)
+\]
+
+This forms a smooth downward-opening parabola in the plane defined by \((x_{i1}, x_{i2})\), with the degree of curvature controlled by \(s_i\).
 
@@
 
-Each variable \(X\_j\) represents integer grid coordinates (e.g., 1, 2, 3, \ldots) in dimension \(j\):
-
-\[
-X\_j \in {1, \dots, k\_j},\text{ where }\prod\_{j=1}^p k\_j \approx n.
+Each variable \(X_j\) represents integer grid coordinates (e.g., 1, 2, 3, \ldots) in dimension \(j\):
+
+\[
+X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n.
 \]
@@
 
-Points are sampled independently from \(X\_1, X\_2, X\_3 \sim \mathcal{U}(-0.5, 0.5)\) of a cube of side length 1.
-This results in a uniform cloud of points filling the unit cube in \(3\text{-}D\) space.
+Points are sampled independently from \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\) of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in \(3\text{-}D\) space.
 
@@
 \[
-X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
 \]
@@
 
-The initial points are drawn from a uniform cube centered at the origin: \(X\_1, X\_2, X\_3 \sim \mathcal{U}(-0.5, 0.5)\).
+The initial points are drawn from a uniform cube centered at the origin: \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\).
 
@@
 
-Remove if \(\sqrt{X\_1^2 + X\_2^2 + X\_3^2} \leq 0.5\).
+Remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\).
 
@@
 \[
-X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
 \].
@@
 \[
-X_i \sim \mathcal{N}_p(\boldsymbol{0}, \Sigma)
+X_i \sim N_p(\boldsymbol{0}, \Sigma)
 \]
@@
 
-\(\text{scale}_i \sim \mathcal{U}(-10, 10)\): randomly chosen direction and stretch.
-
-\(\text{shift}_i \sim \mathcal{U}(-300, 300)\): large offset to spread features apart.
-
-\(\epsilon \sim \mathcal{N}(0, 0.03 \cdot n)\): Gaussian noise scaled to dataset size to introduce mild irregularity.
+\(\text{scale}_i \sim U(-10, 10)\): randomly chosen direction and stretch.
+
+\(\text{shift}_i \sim U(-300, 300)\): large offset to spread features apart.
+
+\(\epsilon \sim N(0, 0.03 \cdot n)\): Gaussian noise scaled to dataset size to introduce mild irregularity.
 
@@
 \[
-X\_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
-\]
-\[
-X\_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
-\]
-\[
-X\_3 = \frac{w}{2} \sin(\theta/2).
+X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
+\]
+\[
+X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
+\]
+\[
+X_3 = \frac{w}{2} \sin(\theta/2).
 \]
@@
 \[
-x_j \sim \mathcal{N}(0, 0.05^2), \quad j = 4, \dots, p.
+x_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
 \]
@@
 The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation.
-Let \(X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])\) be the independent variable.
-A raw polynomial basis of degree 2 is used to form \(X\_2 = X\_1 - X\_1^2 + \varepsilon\_2\), where \(\varepsilon\_2 \sim \mathcal{U}(0, 0.5)\).
+Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the independent variable.
+A raw polynomial basis of degree 2 is used to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
 
@@
 The shape is generated using a third-degree raw polynomial basis expansion.
-Let \(X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])\) be the base input.
-A cubic transformation with vertical jitter is used to define \(X\_2 = X\_1 + X\_1^2 - X\_1^3 + \varepsilon\_2\$, where \$\varepsilon\_2 \sim \mathcal{U}(0, 0.5)\).
+Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the base input.
+A cubic transformation with vertical jitter is used to define \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
 
@@
 \[
-X\_j \sim \mathcal{N}(0, 0.1^2)\$ for \$j = 3, \dots, p.
+X_j \sim N(0, 0.1^2)\$ for \$j = 3, \dots, p.
 \]
@@
 The pyramid is embedded in a \(p\)-dimensional space, with a tip at height zero and base at height \(h\).
-The shape tapers linearly from the base dimensions \((l\_x, l\_y)\) to a smaller rectangular section with side lengths \((2r\_t, 2r\_t)\) at the tip.
+The shape tapers linearly from the base dimensions \((l_x, l_y)\) to a smaller rectangular section with side lengths \((2r_t, 2r_t)\) at the tip.
 Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.
 
-Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points.
-The final dimension \(x\_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\):
+Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The final dimension \(x_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\):
 
@@
 
-Let \(r\_x(z)\) and \(r\_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\):
+Let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\):
 
@@
 \[
-x_1 \sim \mathcal{U}(-r_x(z),\ r_x(z)), \quad
-x_2 \sim \mathcal{U}(-r_y(z),\ r_y(z)), \quad
-x_3 \sim \mathcal{U}(-r_x(z),\ r_x(z))
+x_1 \sim U(-r_x(z),\ r_x(z)), \quad
+x_2 \sim U(-r_y(z),\ r_y(z)), \quad
+x_3 \sim U(-r_x(z),\ r_x(z))
 \].
 
-For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(x\_4\) to \(x\_{p-1}\)) are tapered toward zero with decreasing height and generated as:
-
-\[
-x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
+For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(x_4\) to \(x_{p-1}\)) are tapered toward zero with decreasing height and generated as:
+
+\[
+x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
 \]
@@
 
-Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points.
+Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points.
 The final coordinate \$x\_p\$ encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):
@@
 \[
-u, v \sim \mathcal{U}(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
+u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
 \]
@@
 \[
-x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
+x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
 \]
@@
 
-Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points.
-The final coordinate \(x\_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):
+Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points.
+The final coordinate \(x_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):
 
@@
 
-Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points. The final coordinate \(x\_p\) encodes the height \(z\):
-
-\[
-x_p = z \sim \mathcal{U}(0, h).
-\]
-
-The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r\_b\):
+Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The final coordinate \(x_p\) encodes the height \(z\):
+
+\[
+x_p = z \sim U(0, h).
+\]
+
+The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r_b\):
 
@@
 
-Each point is placed within a regular hexagon in the plane \((x\_1, x\_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}\) and a uniformly random radial scaling factor:
+Each point is placed within a regular hexagon in the plane \((x_1, x_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}\) and a uniformly random radial scaling factor:
 
@@
 \theta \sim \text{Uniform sample from 6 hexagon angles}, \quad
-r_{\text{point}} \sim \sqrt{\mathcal{U}(0, 1)}.
+r_{\text{point}} \sim \sqrt{U(0, 1)}.
 \]
@@
 \[
-x_j \sim \mathcal{U}(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
+x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
 \]
@@
 
-Let \(x\_1, x\_2, \dots, x\_p\) denote the coordinates of the generated points. The generation process begins with an initial point \(T\_0 \in [0, 1]^p\) drawn from a uniform distribution:
-
-\[
-T_0 \sim \mathcal{U}(0, 1)^p.
-\]
-
-Let \(C\_1, C\_2, \dots, C\_{p+1}\) denote the corner vertices of a \(p\)-dimensional simplex.
-At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T\_{i-1}\) and a randomly selected vertex \(C\_k\):
+Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution:
+
+\[
+T_0 \sim U(0, 1)^p.
+\]
+
+Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\)-dimensional simplex.
+At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\):
 
@@
 \[
-X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 4, \dots, p.
+X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p.
 \]
@@
 In generation of \(k\) smaller uniform spheres, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
-These spheres are positioned at distinct random locations in \(p\)-space, with the center of each sphere being drawn from a normal distribution \(\mathcal{N}(0, \texttt{loc}^2 I_p)\).
+These spheres are positioned at distinct random locations in \(p\)-space, with the center of each sphere being drawn from a normal distribution \(N(0, \texttt{loc}^2 I_p)\).
 
@@
 \[
-X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \dots, p.
+X_j \sim N(0, 0.05^2),\text{ for }j = 5, \dots, p.
 \]
@@
 \[
-X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim \mathcal{N}(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
+X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
 \]
@@
 
-Let \(\theta \sim \mathcal{U}(0, 3\pi)\) represent a random angle on a circular base and \(z \sim \mathcal{U}(0, h)\) represent the height along the cylinder.
+Let \(\theta \sim U(0, 3\pi)\) represent a random angle on a circular base and \(z \sim U(0, h)\) represent the height along the cylinder.
 
@@
 \[
-X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \ldots, p.
+X_j \sim N(0, 0.05^2),\text{ for }j = 5, \ldots, p.
 \]
@@
 \item
-  \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim \mathcal{U}(-0.5, 0.5)\) introduces vertical jitter.
+  \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter.
 \item
@@
 \(a_j \in \{2, 3, 4, 5\}\) is a randomly chosen polynomial power,
-\(s_j \sim \mathcal{U}(0.5, 2)\) is a scale factor,
-\(\eta_j \sim \mathcal{U}(-\sigma_j, 2\sigma_j)\), with \(\sigma_j \sim \mathcal{U}(0, 0.05)\), adds mild randomness.
+\(s_j \sim U(0.5, 2)\) is a scale factor,
+\(\eta_j \sim U(-\sigma_j, 2\sigma_j)\), with \(\sigma_j \sim U(0, 0.05)\), adds mild randomness.
 
@@
 \item
-  \(X_3 = 0.05 \cdot \theta + \varepsilon_3\), with \(\varepsilon_3 \sim \mathcal{U}(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix.
+  \(X_3 = 0.05 \cdot \theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix.
 \item
@@
 \[
-X_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
 \]
@@
 
-The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim \mathcal{U}(-0.1, 0.6).\)
-
-Spiral modulation in the fourth dimension is represented by \(X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim \mathcal{U}(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
+The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\)
+
+Spiral modulation in the fourth dimension is represented by \(X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
 
@@
 \[
-X\_j \sim \mathcal{N}(0, 0.05^2), \quad j = 5, \dots, p.
+X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
 \]
@@
 
-\begin{itemize}
-\tightlist
-\item
-  \(X_1 \sim \mathcal{U}(0.1, 2)\): base variable (avoids zero to prevent division errors).
-\item
-  \(X_3 \sim \mathcal{U}(0.1, 0.8)\): independent auxiliary variable.
-\item
-  \(X_2 = \frac{hc}{X_1} + \text{non\_fac} \cdot \sin(X_1)\): non-linear combination of \textbf{hyperbolic and sinusoidal transformations}, creating sharp curvature and oscillation.
-\item
-  \(X_4 = \cos(\pi \cdot X_1) + \varepsilon\), with \(\varepsilon \sim \mathcal{U}(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-\end{itemize}
+\(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors).
+\(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable.
+
+:non-linear combination of \textbf{hyperbolic and sinusoidal transformations}, creating sharp curvature and oscillation.
+\(X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 
@@
 
-\section{Discussion}\label{discussion}
+\section{Conclusion}\label{conclusion}
 
@@
 
+These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{tibble} \citep{kirill2023}, \texttt{gtools} \citep{gregory2023}, \texttt{dplyr} \citep{hadley2023}, \texttt{stats} \citep{core2025}, \texttt{tidyr} \citep{hadley2024}, \texttt{purrr} \citep{hadley2025}, \texttt{mvtnorm} \citep{alan2009}, \texttt{geozoo} \citep{barret2016}, and \texttt{MASS} \citep{venables2002}.
 This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.

### 2025-05-19 paper-cardinalR.R

@@ -58,43 +58,14 @@ source("scripts/additional_functions.R")
 
 
 ## -----------------------------------------------------------------------------
-datasets_tb <- tibble(dt = c("mobiusgau", 
-                             "mobiusgau_tsne1", 
-                             "mobiusgau_tsne2",
-                             "mobiusgau_tsne3",
-                             "mobiusgau_umap1",
-                             "mobiusgau_umap2",
-                             "mobiusgau_umap3"), 
-                      text = c("Simulated data with a Mobius and a Gaussian in 4-D space.",
-                               "The tSNE embedding with perplexity $15$ for mobiusgau.", 
-                               "The tSNE embedding with perplexity $30$ for mobiusgau.",
-                               "The tSNE embedding with perplexity $5$ for mobiusgau.",
-                               "The UMAP embedding with number of neighbors $15$ and minimum distance $0.1$ for mobiusgau.",
-                               "The UMAP embedding with number of neighbors $30$ and minimum distance $0.08$ for mobiusgau.", 
-                               "The UMAP embedding with number of neighbors $5$ and minimum distance $0.9$ for mobiusgau."))
-
-
-## ----datasets-tb-html, eval=knitr::is_html_output()---------------------------
-# datasets_tb |>
-#   kable(caption = "cardinalR data sets", col.names = c("data", "explanation"))
-
-
-## ----datasets-tb-pdf, eval=knitr::is_latex_output()---------------------------
-datasets_tb |> 
-  kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-
-
-## -----------------------------------------------------------------------------
 main_tb <- tibble(arg = c("n",
-                            "p",
-                            "k",
-                            "loc",
-                            "scale",
-                            "shape",
-                            "rotation",
-                            "is_bkg"), 
+                          "p",
+                          "k",
+                          "loc",
+                          "scale",
+                          "shape",
+                          "rotation",
+                          "is_bkg"), 
                         exp = c("A numeric vector representing the number of points in each cluster.",
                                 "A numeric value representing the number of dimensions.",
                                 "A numeric value representing the number of clusters.",

### 2025-05-19 paper-cardinalR.tex
Fixed the mathematical formula issue

@@
 \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable.
-
-:non-linear combination of \textbf{hyperbolic and sinusoidal transformations}, creating sharp curvature and oscillation.
+\(X_{2} = \frac{hc}{X_{1}} + \text{nonfac} \cdot \sin(X_{1})\): non-linear combination of \textbf{hyperbolic and sinusoidal transformations}, creating sharp curvature and oscillation.
 \(X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.

### 2025-05-20 paper-cardinalR.Rmd
Fixed the mathematical formula issue
Added fig.pos latex library

@@
   \usepackage{array}
+  \usepackage{float}
 output: 
@@
 $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable.
-
-<!-- $$ -->
-<!-- X_{2} = \frac{hc}{X_{1}} + \text{non_fac} \cdot \sin(X_{1}) -->
-<!-- $$ -->
-
-:non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
+$X_{2} = \frac{hc}{X_{1}} + \text{nonfac} \cdot \sin(X_{1})$: non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
 $X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.

### 2025-06-19 paper-cardinalR.Rmd
Add conclusion

@@
 
-- Branching: These functions create a controlled environment for testing how effectively various algorithms preserve branching topology and continuity in their low-dimensional embeddings.
-
-- Scurve with a hole allowing for evaluation of how well algorithms handle incomplete manifolds or missing local structure.
-
-- clusteredsphere: This structure allows for cluster separation on curved manifolds in high-dimensional space and can be used to test the ability of NLDR methods and clustering algorithms to detect spherical clusters of different sizes and separations.
-
-- Mobius: The core geometric structure is a Mobius strip—a classic one-sided surface with a half-twist—useful for evaluating how well methods capture non-orientability and twisted manifolds.
-
-- Grided cube: This function is useful for assessing how algorithms preserve uniformly spaced data in high-dimensional spaces.
-
-- Pyrholes: This structure is useful for testing clustering and NLDR algorithms on non-convex and sparse high-dimensional shapes.
+The `cardinalR` package contributes a flexible and extensible framework for generating high-dimensional data structures with well-defined geometric properties.
+It addresses a growing need in the evaluation of machine learning, clustering, and nonlinear dimensionality reduction (NLDR) methods—offering researchers the ability to design customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
+This makes `cardinalR` a useful complement to existing tools like `geozoo`, `snedata`, and `mlbench`, expanding the scope to higher dimensions and more complex shapes.
+
+The suite of included structures is diverse and targeted.
+For example, **branching shapes** allow investigation of continuity and topological preservation in low-dimensional embeddings.
+The **S-curve with a hole** enables exploration of how methods handle incomplete manifolds.
+**Clustered spheres** assess cluster separation on curved surfaces, while the **Möbius strip** challenges NLDR methods with non-orientable geometry.
+**Gridded cubes** test spatial regularity, and **pyrholes** represent non-convex, sparse shapes useful for testing clustering performance in irregular high-dimensional regions.
+
+These structures are not only valuable for algorithm diagnostics but may also be used for **teaching high-dimensional concepts**, **benchmarking reproducibility**, and **evaluating hyperparameter sensitivity**.
+The package's design encourages users to adjust dimensionality, noise, sample size, and clustering properties, promoting **transparent experimentation** and **comparative model evaluation**.
+
+Looking ahead, future extensions of `cardinalR` could incorporate dynamic shape generation, integration with supervised learning simulations, or visual interfaces to assist users in interactively exploring generated structures.
+Expanding the library to include more biologically inspired or real-world data analogs may also broaden its utility across domains like bioinformatics, forensic science, and spatial data analysis.
+
+<!-- - Branching: These functions create a controlled environment for testing how effectively various algorithms preserve branching topology and continuity in their low-dimensional embeddings. -->
+
+<!-- - Scurve with a hole allowing for evaluation of how well algorithms handle incomplete manifolds or missing local structure. -->
+
+<!-- - clusteredsphere: This structure allows for cluster separation on curved manifolds in high-dimensional space and can be used to test the ability of NLDR methods and clustering algorithms to detect spherical clusters of different sizes and separations.
+-->
+
+<!-- - Mobius: The core geometric structure is a Mobius strip—a classic one-sided surface with a half-twist—useful for evaluating how well methods capture non-orientability and twisted manifolds. -->
+
+<!-- - Grided cube: This function is useful for assessing how algorithms preserve uniformly spaced data in high-dimensional spaces. -->
+
+<!-- - Pyrholes: This structure is useful for testing clustering and NLDR algorithms on non-convex and sparse high-dimensional shapes. -->
 

### 2025-06-19 paper-cardinalR.tex
Add conclusion

@@
 
-\begin{itemize}
-\item
-  Branching: These functions create a controlled environment for testing how effectively various algorithms preserve branching topology and continuity in their low-dimensional embeddings.
-\item
-  Scurve with a hole allowing for evaluation of how well algorithms handle incomplete manifolds or missing local structure.
-\item
-  clusteredsphere: This structure allows for cluster separation on curved manifolds in high-dimensional space and can be used to test the ability of NLDR methods and clustering algorithms to detect spherical clusters of different sizes and separations.
-\item
-  Mobius: The core geometric structure is a Mobius strip---a classic one-sided surface with a half-twist---useful for evaluating how well methods capture non-orientability and twisted manifolds.
-\item
-  Grided cube: This function is useful for assessing how algorithms preserve uniformly spaced data in high-dimensional spaces.
-\item
-  Pyrholes: This structure is useful for testing clustering and NLDR algorithms on non-convex and sparse high-dimensional shapes.
-\end{itemize}
+The \texttt{cardinalR} package contributes a flexible and extensible framework for generating high-dimensional data structures with well-defined geometric properties.
+It addresses a growing need in the evaluation of machine learning, clustering, and nonlinear dimensionality reduction (NLDR) methods---offering researchers the ability to design customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
+This makes \texttt{cardinalR} a useful complement to existing tools like \texttt{geozoo}, \texttt{snedata}, and \texttt{mlbench}, expanding the scope to higher dimensions and more complex shapes.
+
+The suite of included structures is diverse and targeted.
+For example, \textbf{branching shapes} allow investigation of continuity and topological preservation in low-dimensional embeddings.
+The \textbf{S-curve with a hole} enables exploration of how methods handle incomplete manifolds.
+\textbf{Clustered spheres} assess cluster separation on curved surfaces, while the \textbf{Möbius strip} challenges NLDR methods with non-orientable geometry.
+\textbf{Gridded cubes} test spatial regularity, and \textbf{pyrholes} represent non-convex, sparse shapes useful for testing clustering performance in irregular high-dimensional regions.
+
+These structures are not only valuable for algorithm diagnostics but may also be used for \textbf{teaching high-dimensional concepts}, \textbf{benchmarking reproducibility}, and \textbf{evaluating hyperparameter sensitivity}.
+The package's design encourages users to adjust dimensionality, noise, sample size, and clustering properties, promoting \textbf{transparent experimentation} and \textbf{comparative model evaluation}.
+
+Looking ahead, future extensions of \texttt{cardinalR} could incorporate dynamic shape generation, integration with supervised learning simulations, or visual interfaces to assist users in interactively exploring generated structures.
+Expanding the library to include more biologically inspired or real-world data analogs may also broaden its utility across domains like bioinformatics, forensic science, and spatial data analysis.
 

### 2025-06-24 paper-cardinalR.tex
Add subsections

@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_expbranches()}}{gen\_expbranches()}}\label{gen_expbranches}
+
 The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in \(2\text{-}D\), with optional noise dimensions to embed the structure in a \(p\)-dimensional space.
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_linearbranches()}}{gen\_linearbranches()}}\label{gen_linearbranches}
 
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_curvybranches()}}{gen\_curvybranches()}}\label{gen_curvybranches}
+
 The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in a \(p\)-dimensional space.
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_orglinearbranches()}}{gen\_orglinearbranches()}}\label{gen_orglinearbranches}
 
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_orgcurvybranches()}}{gen\_orgcurvybranches()}}\label{gen_orgcurvybranches}
 
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_gridcube()}}{gen\_gridcube()}}\label{gen_gridcube}
+
 The \texttt{gen\_gridcube(n,\ p)} constructs a regular grid using approximately \(n\) points by evenly partitioning each of the \(p\) dimensions.
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_unifcube()}}{gen\_unifcube()}}\label{gen_unifcube}
+
 The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in a higher-dimensional space with additional noise dimensions.
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_cubehole()}}{gen\_cubehole()}}\label{gen_cubehole}
 
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_quadratic()}}{gen\_quadratic()}}\label{gen_quadratic}
+
 The \texttt{gen\_quadratic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a quadratic curve in the first two dimensions of a \(p\)-dimensional space.
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_cubic()}}{gen\_cubic()}}\label{gen_cubic}
 
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_pyrrect()}}{gen\_pyrrect()}}\label{gen_pyrrect}
+
 The \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a rectangular base.
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_pyrtri()}}{gen\_pyrtri()}}\label{gen_pyrtri}
+
 The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section.
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_pyrstar()}}{gen\_pyrstar()}}\label{gen_pyrstar}
+
 The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \$n\$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_pyrholes()}}{gen\_pyrholes()}}\label{gen_pyrholes}
 
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_scurve()}}{gen\_scurve()}}\label{gen_scurve}
+
 To simulate an S-curve structure in a higher-dimensional space, we define the function \texttt{gen\_scurve(n,\ p)}, which generates \texttt{n} observations in \texttt{p} dimensions.
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_scurvehole()}}{gen\_scurvehole()}}\label{gen_scurvehole}
 
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_circle()}}{gen\_circle()}}\label{gen_circle}
+
 The function \texttt{gen\_circle(n,\ p)} generates a \(p\)-dimensional dataset of \(n\) observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale.
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_curvycycle()}}{gen\_curvycycle()}}\label{gen_curvycycle}
+
 The \texttt{gen\_curvycycle(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations.
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_unifsphere()}}{gen\_unifsphere()}}\label{gen_unifsphere}
+
 The \texttt{gen\_unifsphere(n,\ p,\ r)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on the surface of a \(3\text{-}D\) sphere of radius \(r\), with additional Gaussian noise dimensions added when \(p > 3\).
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_gridedsphere()}}{gen\_gridedsphere()}}\label{gen_gridedsphere}
+
 The \texttt{gen\_gridedsphere(n,\ p)} function generates a \(p\)-dimensional dataset of approximately \(n\) observations evenly distributed on the surface of a \(3\text{-}D\) unit sphere, with optional Gaussian noise dimensions when \(p > 3\).
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_clusteredspheres()}}{gen\_clusteredspheres()}}\label{gen_clusteredspheres}
 
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_hemisphere()}}{gen\_hemisphere()}}\label{gen_hemisphere}
 
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_crescent()}}{gen\_crescent()}}\label{gen_crescent}
+
 The \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise.
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_curvycylinder()}}{gen\_curvycylinder()}}\label{gen_curvycylinder}
 
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_sphericalspiral()}}{gen\_sphericalspiral()}}\label{gen_sphericalspiral}
+
 The \texttt{gen\_sphericalspiral(n,\ p,\ spins)} function simulates a dataset of \(n\) observations that form a spiraling path on a spherical surface in the first four dimensions.
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_helicalspiral()}}{gen\_helicalspiral()}}\label{gen_helicalspiral}
+
 The \texttt{gen\_helicalspiral(n,\ p)} function generates a dataset of \(n\) observations forming a helical spiral, embedded in the first four dimensions of a \(p\)-dimensional space.
@@
 
+\paragraph{\texorpdfstring{\texttt{gen\_conicspiral()}}{gen\_conicspiral()}}\label{gen_conicspiral}
+
 The \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of a \(p\)-dimensional space.
@@
 \end{figure}
+
+\paragraph{\texorpdfstring{\texttt{gen\_nonlinear()}}{gen\_nonlinear()}}\label{gen_nonlinear}
 

### 2025-06-24 paper-cardinalR.Rmd
Add subsections
Updated gen_expbranches

@@
   \usepackage{float}
+  \newcommand\pD{$p\text{-}D$}
+  \newcommand\gD{$2\text{-}D$}
 output: 
@@
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 
-A branching structure captures trajectories that diverge or bifurcate from a common origin, resembling processes such as cell differentiation in biology.
+A branching structure captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
@@
 
-The `gen_expbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ exponential branches in $2\text{-}D$, with optional noise dimensions to embed the structure in a $p$-dimensional space.
+#### `gen_expbranches()`
+
+The `gen_expbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ exponential branches in $2\text{-}D$, with optional noise dimensions to embed the structure in $p\text{-}D$.
 These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
 
-Each branch $i$ is constructed using:
-
-$X_1 \sim U(-2, 2)$
-$X_2 = \exp(\pm s_i \cdot X_1) + \epsilon$,
-
-where:
-$\epsilon \sim U(0, 0.1)$ adds local jitter,
-
-$s_i \in [0.5, 2]$ is randomly sampled for each branch.
-
+Each branch $i$ is constructed using $X_1 \sim U(-2, 2)$ and $X_2 = \exp(\pm s_iX_1) + \epsilon$, where: $\epsilon \sim U(0, 0.1)$ adds local jitter, $s_i \in [0.5, 2]$ is randomly sampled for each branch.
 The sign alternates between branches (odd: negative, even: positive exponent).
-This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches: decay pattern (reflected about $x$-axis) and even-numbered branches: growth pattern.
-
-For $p > 2$, Gaussian noise is added to embed the $2\text{-}D$ branches in higher dimensions:
-
-$$
-X_j \sim N(0, 0.1^2), \quad j = 3, \dots, p.
-$$
-
-These represent irrelevant or noisy dimensions.
+This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches: decay pattern (reflected) and even-numbered branches: growth pattern.
+For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
 
@@
 
-The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches in a $p$-dimensional space.
+#### `gen_linearbranches()`
+
+The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches in $p\text{-}D$.
 The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.
@@
 The branches differ in direction and location.
-Branches 1 and 2 are initialized with fixed slopes and intercepts.
-The Branch 1 is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5 \cdot X_1 + \epsilon$.
-The Branch 2 is generated from $X_1 \sim U(-6, 2)$, $X_2 = -0.5 \cdot X_1 + \epsilon$.
-$\epsilon \sim U(0, 0.5)$.
-
-Branches 3 to $k$ are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
-
-$X_1$ values span a short segment ($x_{start}$ to $x_{start} + 1$),
-
-$X_2 = s_i \cdot (X_1 - x_{start}) + y_{start} + \epsilon$ where $s_i$ is a sampled slope from a filtered set and $\epsilon \sim U(0, 0.2)$.
-
+Branches $1$ and $2$ are initialized with fixed slopes and intercepts.
+The Branch $1$ is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$.
+The Branch $2$ is generated from $X_1 \sim U(-6, 2)$, $X_2 = -0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$.
+Branches $3$ to $k$ are added iteratively.
+Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
+$X_1$ values span a short segment ($x_{start}$ to $x_{start} + 1$) and $X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon$ where $s_i$ is a sampled slope from a filtered set and $\epsilon \sim U(0, 0.2)$.
 This setup yields $k$ spatially distinct, noisy linear trajectories.
 
-For $p > 2$, the remaining dimensions are pure noise:
-
-$$
-X_j \sim N(0, 0.05^2)\text{ for }j = 3, \dots, p.
-$$
+For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
 
@@
 
+#### `gen_curvybranches()`
+
 The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in a $p$-dimensional space.
@@
 
+#### `gen_orglinearbranches()`
+
 The `gen_orglinearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches embedded in a $p$-dimensional space.
@@
 ```
+
+#### `gen_orgcurvybranches()`
 
@@
 
+#### `gen_gridcube()`
+
 The `gen_gridcube(n, p)` constructs a regular grid using approximately $n$ points by evenly partitioning each of the $p$ dimensions.
@@
 
+#### `gen_unifcube()`
+
 The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in a higher-dimensional space with additional noise dimensions.
@@
 
+#### `gen_cubehole()`
+
 The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p$-dimensional space.
@@
 
+#### `gen_quadratic()`
+
 The `gen_quadratic(n, p, range)` function generates a dataset of $n$ points forming a quadratic curve in the first two dimensions of a $p$-dimensional space.
@@
 ```
+
+#### `gen_cubic()`
 
@@
 
+#### `gen_pyrrect()`
+
 The `gen_pyrrect(n, p, h, l_vec, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a rectangular base.
@@
 
+#### `gen_pyrtri()`
+
 The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section.
@@
 
+#### `gen_pyrstar()`
+
 The `gen_pyrstar(n, p, h, rb)` function generates a dataset of \$n\$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
@@
 
+#### `gen_pyrholes()`
+
 The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p$-dimensional simplex using a chaotic attractor-like midpoint algorithm.
@@
 
+#### `gen_scurve()`
+
 To simulate an S-curve structure in a higher-dimensional space, we define the function `gen_scurve(n, p)`, which generates `n` observations in `p` dimensions.
@@
 ```
+
+#### `gen_scurvehole()`
 
@@
 
+#### `gen_circle()`
+
 The function `gen_circle(n, p)` generates a $p$-dimensional dataset of $n$ observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale.
@@
 
+#### `gen_curvycycle()`
+
 The `gen_curvycycle(n, p)` function generates a $p$-dimensional dataset of $n$ observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations.
@@
 
+#### `gen_unifsphere()`
+
 The `gen_unifsphere(n, p, r)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on the surface of a $3\text{-}D$ sphere of radius $r$, with additional Gaussian noise dimensions added when $p > 3$.
@@
 
+#### `gen_gridedsphere()`
+
 The `gen_gridedsphere(n, p)` function generates a $p$-dimensional dataset of approximately $n$ observations evenly distributed on the surface of a $3\text{-}D$ unit sphere, with optional Gaussian noise dimensions when $p > 3$.
@@
 
+#### `gen_clusteredspheres()`
+
 The `gen_clusteredspheres(n, k, p, r, loc)` function generates a synthetic dataset of $n_1 + k \cdot n_2$ observations in $p$-dimensional space, consisting of one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location.
@@
 ```
+
+#### `gen_hemisphere()`
 
@@
 
+#### `gen_crescent()`
+
 The `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise.
@@
 
+#### `gen_curvycylinder()`
+
 The `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$.
@@
 
+#### `gen_sphericalspiral()`
+
 The `gen_sphericalspiral(n, p, spins)` function simulates a dataset of $n$ observations that form a spiraling path on a spherical surface in the first four dimensions.
@@
 
+#### `gen_helicalspiral()`
+
 The `gen_helicalspiral(n, p)` function generates a dataset of $n$ observations forming a helical spiral, embedded in the first four dimensions of a $p$-dimensional space.
@@
 
+#### `gen_conicspiral()`
+
 The `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of a $p$-dimensional space.
@@
 ```
+
+#### `gen_nonlinear()`
 

### 2025-06-24 paper-cardinalR.tex
Updated gen_expbranches

@@
 
-The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in \(2\text{-}D\), with optional noise dimensions to embed the structure in a \(p\)-dimensional space.
+The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in \(2\text{-}D\), with optional noise dimensions to embed the structure in \(p\text{-}D\).
 These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
 
-Each branch \(i\) is constructed using:
-
-\(X_1 \sim U(-2, 2)\)
-\(X_2 = \exp(\pm s_i \cdot X_1) + \epsilon\),
-
-where:
-\(\epsilon \sim U(0, 0.1)\) adds local jitter,
-
-\(s_i \in [0.5, 2]\) is randomly sampled for each branch.
-
+Each branch \(i\) is constructed using \(X_1 \sim U(-2, 2)\) and \(X_2 = \exp(\pm s_iX_1) + \epsilon\), where: \(\epsilon \sim U(0, 0.1)\) adds local jitter, \(s_i \in [0.5, 2]\) is randomly sampled for each branch.
 The sign alternates between branches (odd: negative, even: positive exponent).
-This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches: decay pattern (reflected about \(x\)-axis) and even-numbered branches: growth pattern.
-
-For \(p > 2\), Gaussian noise is added to embed the \(2\text{-}D\) branches in higher dimensions:
-
-\[
-X_j \sim N(0, 0.1^2), \quad j = 3, \dots, p.
-\]
-
-These represent irrelevant or noisy dimensions.
+This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches: decay pattern (reflected) and even-numbered branches: growth pattern.
+For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
 
@@
 
-The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches in a \(p\)-dimensional space.
+The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches in \(p\text{-}D\).
 The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.
@@
 The branches differ in direction and location.
-Branches 1 and 2 are initialized with fixed slopes and intercepts.
-The Branch 1 is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5 \cdot X_1 + \epsilon\).
-The Branch 2 is generated from \(X_1 \sim U(-6, 2)\), \(X_2 = -0.5 \cdot X_1 + \epsilon\).
-\(\epsilon \sim U(0, 0.5)\).
-
-Branches 3 to \(k\) are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
-
-\(X_1\) values span a short segment (\(x_{start}\) to \(x_{start} + 1\)),
-
-\(X_2 = s_i \cdot (X_1 - x_{start}) + y_{start} + \epsilon\) where \(s_i\) is a sampled slope from a filtered set and \(\epsilon \sim U(0, 0.2)\).
-
+Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts.
+The Branch \(1\) is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\).
+The Branch \(2\) is generated from \(X_1 \sim U(-6, 2)\), \(X_2 = -0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\).
+Branches \(3\) to \(k\) are added iteratively.
+Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
+\(X_1\) values span a short segment (\(x_{start}\) to \(x_{start} + 1\)) and \(X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon\) where \(s_i\) is a sampled slope from a filtered set and \(\epsilon \sim U(0, 0.2)\).
 This setup yields \(k\) spatially distinct, noisy linear trajectories.
 
-For \(p > 2\), the remaining dimensions are pure noise:
-
-\[
-X_j \sim N(0, 0.05^2)\text{ for }j = 3, \dots, p.
-\]
+For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
 

### 2025-06-24 paper-cardinalR.R

@@ -83,7 +83,7 @@ main_tb <- tibble(arg = c("n",
 
 ## ----main-tb-pdf, eval=knitr::is_latex_output()-------------------------------
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
   column_spec(2, width = "8cm")
 

### 2025-06-24 paper-cardinalR.Rmd
Updated gen_linearbranches text

@@
 Branches $3$ to $k$ are added iteratively.
-Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
-$X_1$ values span a short segment ($x_{start}$ to $x_{start} + 1$) and $X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon$ where $s_i$ is a sampled slope from a filtered set and $\epsilon \sim U(0, 0.2)$.
-This setup yields $k$ spatially distinct, noisy linear trajectories.
-
+Each additional branch $i$ starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
+The $X_1$ values are defined over a short range, from $x_{start}$ to $x_{start} + 1$.
+The $X_2$ value is calculated using the formula $X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon$, where $s_i$ is a chosen slope from a selected branch, and $\epsilon \sim U(0, 0.2)$.
 For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
@@
 
-The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in a $p$-dimensional space.
-The underlying geometry lies in the first two dimensions, while the remaining $p - 2$ dimensions contain Gaussian noise.
+The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in $p\text{-}D$.
+The underlying geometry lies in the first two dimensions, while the remaining $(p-2)\text{-}D$ dimensions contain Gaussian noise.
 

### 2025-06-24 paper-cardinalR.tex
Updated gen_linearbranches text

@@
 Branches \(3\) to \(k\) are added iteratively.
-Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
-\(X_1\) values span a short segment (\(x_{start}\) to \(x_{start} + 1\)) and \(X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon\) where \(s_i\) is a sampled slope from a filtered set and \(\epsilon \sim U(0, 0.2)\).
-This setup yields \(k\) spatially distinct, noisy linear trajectories.
-
+Each additional branch \(i\) starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
+The \(X_1\) values are defined over a short range, from \(x_{start}\) to \(x_{start} + 1\).
+The \(X_2\) value is calculated using the formula \(X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon\), where \(s_i\) is a chosen slope from a selected branch, and \(\epsilon \sim U(0, 0.2)\).
 For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
@@
 
-The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in a \(p\)-dimensional space.
-The underlying geometry lies in the first two dimensions, while the remaining \(p - 2\) dimensions contain Gaussian noise.
+The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in \(p\text{-}D\).
+The underlying geometry lies in the first two dimensions, while the remaining \((p-2)\text{-}D\) dimensions contain Gaussian noise.
 

### 2025-06-24 paper-cardinalR.Rmd
Updated Gaussian and Linear sections

@@
 The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in $p\text{-}D$.
-The underlying geometry lies in the first two dimensions, while the remaining $(p-2)\text{-}D$ dimensions contain Gaussian noise.
-
-Branch 1 is generated from $X_1 \sim U(0, 1)$, and $X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
+The underlying geometry lies in the first two dimensions, while the remaining $(p-2)$ dimensions contain Gaussian noise.
+
+Branch $1$ is generated from $X_1 \sim U(0, 1)$, and $X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
 This produces a gently upward-curving parabola in the right half-plane.
-
-Branch 2 is generated from $X_{1} \sim U(-1, 0)$, and $X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
+Branch $2$ is generated from $X_{1} \sim U(-1, 0)$, and $X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
 This creates a steeper, leftward-facing curve in the left half-plane.
-
 Branches $3$ to $k$ are added iteratively.
-Each new branch begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X_{1}$:$X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure:
+Each new branch $i$ begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X_{1}$:$X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure:
 
@@
 
-For $p > 2$, the remaining dimensions are pure Gaussian noise:
-
-$$
-X_j \sim N(0, 0.05^2), \quad \text{for } j = 3, \dots, p.
-$$
+For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
 
@@
 
-The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p$-dimensional space, centered at the origin with user-defined covariance structure.
-
-Each point is independently drawn using the multivariate normal distribution with:
-
-$$
-X_i \sim N_p(\boldsymbol{0}, \Sigma)
-$$
-
-where $\Sigma$ is a user-defined $p \times p$ positive-definite matrix.
+The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure.
+Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, \Sigma)$, where $\Sigma$ is a user-defined $p \times p$ positive-definite matrix.
 
@@
 The `gen_longlinear(n, p)` function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise.
-
-Each variable $X_i$ is created as:
-
-$$
-X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i
-$$
-
-where:
-
-$\text{scale}_i \sim U(-10, 10)$: randomly chosen direction and stretch.
-
-$\text{shift}_i \sim U(-300, 300)$: large offset to spread features apart.
-
-$\epsilon \sim N(0, 0.03 \cdot n)$: Gaussian noise scaled to dataset size to introduce mild irregularity.
+Each variable $X_i$ is created as $X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i$, where $\text{scale}_i \sim U(-10, 10)$ is randomly chosen direction and stretch, $\text{shift}_i \sim U(-300, 300)$ is large offset to spread features apart, and $\epsilon \sim N(0, 0.03 \cdot n)$ is Gaussian noise.
 

### 2025-06-24 paper-cardinalR.tex
Updated Gaussian and Linear sections

@@
 The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in \(p\text{-}D\).
-The underlying geometry lies in the first two dimensions, while the remaining \((p-2)\text{-}D\) dimensions contain Gaussian noise.
-
-Branch 1 is generated from \(X_1 \sim U(0, 1)\), and \(X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
+The underlying geometry lies in the first two dimensions, while the remaining \((p-2)\) dimensions contain Gaussian noise.
+
+Branch \(1\) is generated from \(X_1 \sim U(0, 1)\), and \(X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
 This produces a gently upward-curving parabola in the right half-plane.
-
-Branch 2 is generated from \(X_{1} \sim U(-1, 0)\), and \(X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
+Branch \(2\) is generated from \(X_{1} \sim U(-1, 0)\), and \(X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
 This creates a steeper, leftward-facing curve in the left half-plane.
-
 Branches \(3\) to \(k\) are added iteratively.
-Each new branch begins at a randomly chosen point within a restricted horizontal band: \(X_{1} \in [-0.15, 0.15]\) (to ensure connectivity with earlier branches), spans a unit-length interval on \(X_{1}\):\(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), and has the structure:
+Each new branch \(i\) begins at a randomly chosen point within a restricted horizontal band: \(X_{1} \in [-0.15, 0.15]\) (to ensure connectivity with earlier branches), spans a unit-length interval on \(X_{1}\):\(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), and has the structure:
 
@@
 
-For \(p > 2\), the remaining dimensions are pure Gaussian noise:
-
-\[
-X_j \sim N(0, 0.05^2), \quad \text{for } j = 3, \dots, p.
-\]
+For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
 
@@
 
-The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\)-dimensional space, centered at the origin with user-defined covariance structure.
-
-Each point is independently drawn using the multivariate normal distribution with:
-
-\[
-X_i \sim N_p(\boldsymbol{0}, \Sigma)
-\]
-
-where \(\Sigma\) is a user-defined \(p \times p\) positive-definite matrix.
+The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure.
+Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, \Sigma)\), where \(\Sigma\) is a user-defined \(p \times p\) positive-definite matrix.
 
@@
 The \texttt{gen\_longlinear(n,\ p)} function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise.
-
-Each variable \(X_i\) is created as:
-
-\[
-X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i
-\]
-
-where:
-
-\(\text{scale}_i \sim U(-10, 10)\): randomly chosen direction and stretch.
-
-\(\text{shift}_i \sim U(-300, 300)\): large offset to spread features apart.
-
-\(\epsilon \sim N(0, 0.03 \cdot n)\): Gaussian noise scaled to dataset size to introduce mild irregularity.
+Each variable \(X_i\) is created as \(X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i\), where \(\text{scale}_i \sim U(-10, 10)\) is randomly chosen direction and stretch, \(\text{shift}_i \sim U(-300, 300)\) is large offset to spread features apart, and \(\epsilon \sim N(0, 0.03 \cdot n)\) is Gaussian noise.
 

### 2025-06-24 paper-cardinalR.Rmd
Added Mobius structure

@@
 The `gen_longlinear(n, p)` function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise.
-Each variable $X_i$ is created as $X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i$, where $\text{scale}_i \sim U(-10, 10)$ is randomly chosen direction and stretch, $\text{shift}_i \sim U(-300, 300)$ is large offset to spread features apart, and $\epsilon \sim N(0, 0.03 \cdot n)$ is Gaussian noise.
+Each variable $X_i$ is created as $X_i = \text{scale}_i \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i$, where $\text{scale}_i \sim U(-10, 10)$ is randomly chosen direction and stretch, $\text{shift}_i \sim U(-300, 300)$ is large offset to spread features apart, and $\epsilon \sim N(0, 0.03n)$ is Gaussian noise.
 
@@
 
-The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p$-dimensional space.
+The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$.
 This classical non-orientable surface loops back on itself with a half-twist.
@@
 The Mobius strip is generated using `geozoo::mobius(n, p = 3)$points`, which samples $n$ points from a parametrization of the surface defined by angle $\theta$ and strip width $w$.
-The cartesian coordinates are;
-
-$$
-X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
-$$
-$$
-X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
-$$
-$$
-X_3 = \frac{w}{2} \sin(\theta/2).
-$$
-
-This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable 1-sided surface.
-
-For $p > 3$, the function appends $p - 3$ dimensions with small Gaussian noise:
-
-$$
-x_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
-$$
+The cartesian coordinates are $X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = \frac{w}{2} \sin(\theta/2)$.
+This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface.
+For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ Mobius into $p\text{-}D$, where $j = 4, \dots, p$.
 

### 2025-06-24 paper-cardinalR.tex
Added Mobius structure

@@
 The \texttt{gen\_longlinear(n,\ p)} function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise.
-Each variable \(X_i\) is created as \(X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i\), where \(\text{scale}_i \sim U(-10, 10)\) is randomly chosen direction and stretch, \(\text{shift}_i \sim U(-300, 300)\) is large offset to spread features apart, and \(\epsilon \sim N(0, 0.03 \cdot n)\) is Gaussian noise.
+Each variable \(X_i\) is created as \(X_i = \text{scale}_i \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i\), where \(\text{scale}_i \sim U(-10, 10)\) is randomly chosen direction and stretch, \(\text{shift}_i \sim U(-300, 300)\) is large offset to spread features apart, and \(\epsilon \sim N(0, 0.03n)\) is Gaussian noise.
 
@@
 
-The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\)-dimensional space.
+The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\).
 This classical non-orientable surface loops back on itself with a half-twist.
@@
 The Mobius strip is generated using \texttt{geozoo::mobius(n,\ p\ =\ 3)\$points}, which samples \(n\) points from a parametrization of the surface defined by angle \(\theta\) and strip width \(w\).
-The cartesian coordinates are;
-
-\[
-X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
-\]
-\[
-X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
-\]
-\[
-X_3 = \frac{w}{2} \sin(\theta/2).
-\]
-
-This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable 1-sided surface.
-
-For \(p > 3\), the function appends \(p - 3\) dimensions with small Gaussian noise:
-
-\[
-x_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
-\]
+The cartesian coordinates are \(X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta)\), \(X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta)\), and \(X_3 = \frac{w}{2} \sin(\theta/2)\).
+This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable one-sided surface.
+For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) Mobius into \(p\text{-}D\), where \(j = 4, \dots, p\).
 

### 2025-06-24 paper-cardinalR.Rmd
Updated swissroll

@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, p, w)`, where `n` is the number of points, `p` is the total number of dimensions, and `w` is the vertical range in the third dimension.
-
-The first three dimensions form the classic $3\text{-}D$ Swiss roll shape:
-
-$$
-X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)
-$$
-
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, p, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension.
+The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.
+The $X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)$.
 For $p > 3$, the remaining $p - 3$ dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.

### 2025-06-24 paper-cardinalR.tex
Updated swissroll

@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ p,\ w)}, where \texttt{n} is the number of points, \texttt{p} is the total number of dimensions, and \texttt{w} is the vertical range in the third dimension.
-
-The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape:
-
-\[
-X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)
-\]
-
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ p,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension.
+The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.
+The \(X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)\).
 For \(p > 3\), the remaining \(p - 3\) dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.

### 2025-06-24 paper-cardinalR.Rmd
Updated cone data generation section

@@
 
-This function generates `n` points in a `p`-dimensional space, where the last dimension (`X_p`) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
-
-Points along the height axis are drawn from an exponential distribution to increase density near the tip:
-
-$$
-X_p = h_i \sim \text{Exp}(\lambda = 2/h), \quad \text{truncated at } h
-$$
-
-The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter:
-
-$$
-r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) \cdot \frac{h_i}{h}, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1
-$$
+This function generates $n$ points in a $p\text{-}D$, where the last dimension ($X_p$) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
+Points along the height axis are drawn from an exponential distribution to increase density near the tip.
+The $X_p = h_i \sim \text{Exp}(\lambda = 2/h)$ truncated at $h$.
+The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter $r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1$.
 
@@
 
-The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$.
+The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$ structure.
 This classical non-orientable surface loops back on itself with a half-twist.
@@
 The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.
-The $X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)$.
+The $X_1 = t \cos(t)$, $X_2 = t \sin(t)$, $X_3 \sim U(w_1, w_2), \text{where } t \sim U(0, 3\pi)$.
 For $p > 3$, the remaining $p - 3$ dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.

### 2025-06-24 paper-cardinalR.tex
Updated cone data generation section

@@
 
-This function generates \texttt{n} points in a \texttt{p}-dimensional space, where the last dimension (\texttt{X\_p}) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
-
-Points along the height axis are drawn from an exponential distribution to increase density near the tip:
-
-\[
-X_p = h_i \sim \text{Exp}(\lambda = 2/h), \quad \text{truncated at } h
-\]
-
-The effective radius at height \(h_i\) decreases linearly from base to tip, controlled by a shape ratio parameter:
-
-\[
-r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) \cdot \frac{h_i}{h}, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1
-\]
+This function generates \(n\) points in a \(p\text{-}D\), where the last dimension (\(X_p\)) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
+Points along the height axis are drawn from an exponential distribution to increase density near the tip.
+The \(X_p = h_i \sim \text{Exp}(\lambda = 2/h)\) truncated at \(h\).
+The effective radius at height \(h_i\) decreases linearly from base to tip, controlled by a shape ratio parameter \(r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1\).
 
@@
 
-The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\).
+The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\) structure.
 This classical non-orientable surface loops back on itself with a half-twist.
@@
 The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.
-The \(X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)\).
+The \(X_1 = t \cos(t)\), \(X_2 = t \sin(t)\), \(X_3 \sim U(w_1, w_2), \text{where } t \sim U(0, 3\pi)\).
 For \(p > 3\), the remaining \(p - 3\) dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.

### 2025-06-24 paper-cardinalR.Rmd
Added overall sentence for cube data

@@
 
+A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
+These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:cube-tb-html)', '\\@ref(tab:cube-tb-pdf)'))` outlines these functions and their purposes.
+
 ```{r}

### 2025-06-24 paper-cardinalR.tex
Added overall sentence for cube data

@@
 \subsubsection{Cube}\label{cube}
+
+A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
+These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
+Table \ref{tab:cube-tb-pdf} outlines these functions and their purposes.
 

### 2025-06-24 paper-cardinalR.Rmd
Added overall paragraph for polynomial

@@
 
+A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in high-dimensional space.
+These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
+We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:polynomial-tb-html)', '\\@ref(tab:polynomial-tb-pdf)'))` summarizes these functions and their purposes.
+
 ```{r}

### 2025-06-24 paper-cardinalR.tex
Added overall paragraph for polynomial

@@
 \subsubsection{Polynomial}\label{polynomial}
+
+A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in high-dimensional space.
+These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
+We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
+Table \ref{tab:polynomial-tb-pdf} summarizes these functions and their purposes.
 

### 2025-06-24 paper-cardinalR.Rmd
Added pyramid overall paragraph

@@
 
+A pyramid structure represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
+The functions provided allow users to generate pyramids with various base shapes—including rectangular, triangular, and star-shaped bases—with options for sharp or blunted apexes.
+Additionally, we offer functionality for creating pyramids containing structured holes, enabling the study of non-convex and sparse regions.
+These designs are valuable for testing the ability of dimension reduction and clustering algorithms to preserve and detect complex geometric features.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:pyramid-tb-html)', '\\@ref(tab:pyramid-tb-pdf)'))` summarizes these functions.
+
 ```{r}

### 2025-06-24 paper-cardinalR.tex
Added pyramid overall paragraph

@@
 \subsubsection{Pyramid}\label{pyramid}
+
+A pyramid structure represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
+The functions provided allow users to generate pyramids with various base shapes---including rectangular, triangular, and star-shaped bases---with options for sharp or blunted apexes.
+Additionally, we offer functionality for creating pyramids containing structured holes, enabling the study of non-convex and sparse regions.
+These designs are valuable for testing the ability of dimension reduction and clustering algorithms to preserve and detect complex geometric features.
+Table \ref{tab:pyramid-tb-pdf} summarizes these functions.
 

### 2025-06-24 paper-cardinalR.Rmd
Added overall paragraph for S-curev data

@@
 
+An S-curve structure simulates data that lies along a smooth, non-linear manifold, making it a classic example for testing the performance of non-linear dimension reduction methods.
+This package provides functions to generate standard S-curve shapes as well as variants containing holes to introduce missing or incomplete regions.
+These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:scurve-tb-html)', '\\@ref(tab:scurve-tb-pdf)'))` summarizes these functions.
+
 ```{r}

### 2025-06-24 paper-cardinalR.tex
Added overall paragraph for S-curev data

@@
 \subsubsection{S-curve}\label{s-curve}
+
+An S-curve structure simulates data that lies along a smooth, non-linear manifold, making it a classic example for testing the performance of non-linear dimension reduction methods.
+This package provides functions to generate standard S-curve shapes as well as variants containing holes to introduce missing or incomplete regions.
+These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
+Table \ref{tab:scurve-tb-pdf} summarizes these functions.
 

### 2025-06-24 paper-cardinalR.Rmd
Added sphere overall paragraph

@@
 
+Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+The package provides functions to generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+These shapes help test how well methods preserve symmetry, curvature, and cluster separation.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:sphere-tb-html)', '\\@ref(tab:sphere-tb-pdf)'))` summarizes these functions.
+
 ```{r}

### 2025-06-24 paper-cardinalR.tex
Added sphere overall paragraph

@@
 \subsubsection{Sphere}\label{sphere}
+
+Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+The package provides functions to generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+These shapes help test how well methods preserve symmetry, curvature, and cluster separation.
+Table \ref{tab:sphere-tb-pdf} summarizes these functions.
 

### 2025-06-24 paper-cardinalR.Rmd
Added overall paragraph for trigometric data

@@
 
+Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems.
+The package includes functions to generate a wide range of these forms from crescents and curvy cylinders to spherical, helical, and conic spirals, as well as nonlinear hyperbolas.
+These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trigonometric-tb-html)', '\\@ref(tab:trigonometric-tb-pdf)'))` summarizes these functions.
+
 ```{r}

### 2025-06-24 paper-cardinalR.tex
Added overall paragraph for trigometric data

@@
 \subsubsection{Trigonometric}\label{trigonometric}
+
+Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems.
+The package includes functions to generate a wide range of these forms from crescents and curvy cylinders to spherical, helical, and conic spirals, as well as nonlinear hyperbolas.
+These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
+Table \ref{tab:trigonometric-tb-pdf} summarizes these functions.
 

### 2025-06-24 paper-cardinalR.Rmd
Updated cube section

@@
 
-Each variable $X_j$ represents integer grid coordinates (e.g., 1, 2, 3, ...) in dimension $j$:
-
-$$
-X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n.
-$$
-
-This results in a hypercube-shaped lattice in $p$-dimensional space.
+Each variable $X_j$ represents integer grid coordinates (e.g., 1, 2, 3, ...) in dimension $j$: $X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n$.
+This results in a hypercube-shaped lattice in $p\text{-}D$.
 
@@
 
-The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in a higher-dimensional space with additional noise dimensions.
-
-Points are sampled independently from $X_1, X_2, X_3 \sim U(-0.5, 0.5)$ of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.
-
-For $p > 3$, the remaining dimensions are filled with Gaussian noise:
-
-$$
-X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
-$$
+The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in $p\text{-}D$ with additional noise dimensions.
+
+Points are sampled independently from $X_1, X_2, X_3 \sim U(-0.5, 0.5)$ of a cube of side length $1$.
+This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.
+For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ structure into $p\text{-}D$, where $j = 4, \dots, p$.
 
@@
 
-The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p$-dimensional space.
+The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p\text{-}D$.
 
 The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim U(-0.5, 0.5)$.
-
-The center of the cube is $(0, 0, 0)$. A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$:
-
-Remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$.
-
+The center of the cube is $(0, 0, 0)$.
+A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$: Remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$.
 This results in a shell-like point cloud that fills the cube but avoids the spherical center.
-
-For $p > 3$, Gaussian noise is added to the remaining dimensions (via `gen_unifcube()`):
-
-$$
-X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
-$$.
+For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ structure into $p\text{-}D$, where $j = 4, \dots, p$.
 

### 2025-06-24 paper-cardinalR.tex
Updated cube section

@@
 
-Each variable \(X_j\) represents integer grid coordinates (e.g., 1, 2, 3, \ldots) in dimension \(j\):
-
-\[
-X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n.
-\]
-
-This results in a hypercube-shaped lattice in \(p\)-dimensional space.
+Each variable \(X_j\) represents integer grid coordinates (e.g., 1, 2, 3, \ldots) in dimension \(j\): \(X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n\).
+This results in a hypercube-shaped lattice in \(p\text{-}D\).
 
@@
 
-The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in a higher-dimensional space with additional noise dimensions.
-
-Points are sampled independently from \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\) of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in \(3\text{-}D\) space.
-
-For \(p > 3\), the remaining dimensions are filled with Gaussian noise:
-
-\[
-X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
-\]
+The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in \(p\text{-}D\) with additional noise dimensions.
+
+Points are sampled independently from \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\) of a cube of side length \(1\).
+This results in a uniform cloud of points filling the unit cube in \(3\text{-}D\) space.
+For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) structure into \(p\text{-}D\), where \(j = 4, \dots, p\).
 
@@
 
-The \texttt{gen\_cubehole(n,\ p)} function generates a dataset of \(n\) points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in \(p\)-dimensional space.
+The \texttt{gen\_cubehole(n,\ p)} function generates a dataset of \(n\) points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in \(p\text{-}D\).
 
 The initial points are drawn from a uniform cube centered at the origin: \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\).
-
-The center of the cube is \((0, 0, 0)\). A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\):
-
-Remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\).
-
+The center of the cube is \((0, 0, 0)\).
+A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\): Remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\).
 This results in a shell-like point cloud that fills the cube but avoids the spherical center.
-
-For \(p > 3\), Gaussian noise is added to the remaining dimensions (via \texttt{gen\_unifcube()}):
-
-\[
-X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
-\].
+For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) structure into \(p\text{-}D\), where \(j = 4, \dots, p\).
 

### 2025-06-24 paper-cardinalR.Rmd
Updated cube section

@@
 The center of the cube is $(0, 0, 0)$.
-A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$: Remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$.
+A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$: remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$.
 This results in a shell-like point cloud that fills the cube but avoids the spherical center.
@@
 A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-
-This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate measurement noise or intrinsic variability.
-
-For $p > 2$, the dataset includes Gaussian noise in the remaining dimensions via `gen_noisedims()`:
-
-$$
-X_j \sim N(0, 0.1^2), \quad j = 3, \dots, p.
-$$
+This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise.
+For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ structure into $p\text{-}D$, where $j = 3, \dots, p$.
 
@@
 A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-
-When $p > 2$, the remaining dimensions consist of independent Gaussian noise features generated via `gen_noisedims()`:
-
-$$
-X_j \sim N(0, 0.1^2)\$ for \$j = 3, \dots, p.
-$$
+For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ structure into $p\text{-}D$, where $j = 3, \dots, p$.
 

### 2025-06-24 paper-cardinalR.tex
Updated cube section

@@
 The center of the cube is \((0, 0, 0)\).
-A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\): Remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\).
+A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\): remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\).
 This results in a shell-like point cloud that fills the cube but avoids the spherical center.
@@
 A raw polynomial basis of degree 2 is used to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-
-This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate measurement noise or intrinsic variability.
-
-For \(p > 2\), the dataset includes Gaussian noise in the remaining dimensions via \texttt{gen\_noisedims()}:
-
-\[
-X_j \sim N(0, 0.1^2), \quad j = 3, \dots, p.
-\]
+This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise.
+For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) structure into \(p\text{-}D\), where \(j = 3, \dots, p\).
 
@@
 A cubic transformation with vertical jitter is used to define \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-
-When \(p > 2\), the remaining dimensions consist of independent Gaussian noise features generated via \texttt{gen\_noisedims()}:
-
-\[
-X_j \sim N(0, 0.1^2)\$ for \$j = 3, \dots, p.
-\]
+For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) structure into \(p\text{-}D\), where \(j = 3, \dots, p\).
 

### 2025-06-24 paper-cardinalR.Rmd
Updated S-curve data generation functions

@@
 
-The `gen_quadratic(n, p, range)` function generates a dataset of $n$ points forming a quadratic curve in the first two dimensions of a $p$-dimensional space.
+The `gen_quadratic(n, p, range)` function generates a dataset of $n$ points forming a quadratic curve in the first two dimensions.
 This $2\text{-}D$ parabolic structure is embedded within a higher-dimensional space with additive noise in the remaining dimensions.
@@
 
-The `gen_cubic(n, p, range)` function generates a dataset of $n$ points forming a cubic curve in the first two dimensions of a $p$-dimensional space.
+The `gen_cubic(n, p, range)` function generates a dataset of $n$ points forming a cubic curve in the first two dimensions.
 This function creates a more complex curvilinear structure than a simple parabola.
@@
 
-To simulate an S-curve structure in a higher-dimensional space, we define the function `gen_scurve(n, p)`, which generates `n` observations in `p` dimensions.
+To simulate an S-curve structure in a higher-dimensional space, we define the function `gen_scurve(n, p)`, which generates $n$ observations in $p\text{-}D$.
 
@@
 This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure:
-
-$$
-  X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta) \cdot (\cos(\theta) - 1).
-$$
-
+The first three dimensions form the S-curve structure: $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$, $X_3 = \text{sign}(\theta) \cdot (\cos(\theta) - 1)$.
 This configuration creates a horizontally curled shape in $(X_1, X_3)$, with additional band thickness in the $X_2$ direction.
-
-For $p > 3$, additional noise dimensions are appended introducing structured, wavy perturbations.
+For $p > 3$, additional noise dimensions are appended introducing structured, wavy noise.
 
@@
 To simulate a variant of the S-curve structure with a removed region ("hole"), we define the function `gen_scurvehole(n, p)`.
-This function produces `n` observations in a $p$-dimensional space where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
+This function produces $n$ observations in $p\text{-}D$ where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
 A subset of observations near a designated anchor point is excluded to introduce a hole in the manifold.
 
-To simulate missing regions on the manifold, a fixed anchor point $(0, 1, 0, \ldots)$ is defined in $p$-dimensional space.
-All observations within a Euclidean distance of $\sqrt{0.3} \approx 0.5477$ from the anchor are removed:
-
-$$
-\text{remove if} \quad \sum_{j=1}^p (X_j - a_j)^2 \leq 0.3
-$$.
-
+To simulate missing regions on the manifold, a fixed anchor point $(0, 1, 0, \ldots)$ is defined in $p\text{-}D$.
+All observations within a Euclidean distance of $\sqrt{0.3} \approx 0.5477$ from the anchor are removed if $\sum_{j=1}^p (X_j - a_j)^2 \leq 0.3$.
 This exclusion creates a hole in the manifold centered near the middle vertical region of the S-curve.

### 2025-06-24 paper-cardinalR.tex
Updated S-curve data generation functions

@@
 
-The \texttt{gen\_quadratic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a quadratic curve in the first two dimensions of a \(p\)-dimensional space.
+The \texttt{gen\_quadratic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a quadratic curve in the first two dimensions.
 This \(2\text{-}D\) parabolic structure is embedded within a higher-dimensional space with additive noise in the remaining dimensions.
@@
 
-The \texttt{gen\_cubic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a cubic curve in the first two dimensions of a \(p\)-dimensional space.
+The \texttt{gen\_cubic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a cubic curve in the first two dimensions.
 This function creates a more complex curvilinear structure than a simple parabola.
@@
 
-To simulate an S-curve structure in a higher-dimensional space, we define the function \texttt{gen\_scurve(n,\ p)}, which generates \texttt{n} observations in \texttt{p} dimensions.
+To simulate an S-curve structure in a higher-dimensional space, we define the function \texttt{gen\_scurve(n,\ p)}, which generates \(n\) observations in \(p\text{-}D\).
 
@@
 This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure:
-
-\[
-  X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta) \cdot (\cos(\theta) - 1).
-\]
-
+The first three dimensions form the S-curve structure: \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\), \(X_3 = \text{sign}(\theta) \cdot (\cos(\theta) - 1)\).
 This configuration creates a horizontally curled shape in \((X_1, X_3)\), with additional band thickness in the \(X_2\) direction.
-
-For \(p > 3\), additional noise dimensions are appended introducing structured, wavy perturbations.
+For \(p > 3\), additional noise dimensions are appended introducing structured, wavy noise.
 
@@
 To simulate a variant of the S-curve structure with a removed region (``hole''), we define the function \texttt{gen\_scurvehole(n,\ p)}.
-This function produces \texttt{n} observations in a \(p\)-dimensional space where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
+This function produces \(n\) observations in \(p\text{-}D\) where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
 A subset of observations near a designated anchor point is excluded to introduce a hole in the manifold.
 
-To simulate missing regions on the manifold, a fixed anchor point \((0, 1, 0, \ldots)\) is defined in \(p\)-dimensional space.
-All observations within a Euclidean distance of \(\sqrt{0.3} \approx 0.5477\) from the anchor are removed:
-
-\[
-\text{remove if} \quad \sum_{j=1}^p (X_j - a_j)^2 \leq 0.3
-\].
-
+To simulate missing regions on the manifold, a fixed anchor point \((0, 1, 0, \ldots)\) is defined in \(p\text{-}D\).
+All observations within a Euclidean distance of \(\sqrt{0.3} \approx 0.5477\) from the anchor are removed if \(\sum_{j=1}^p (X_j - a_j)^2 \leq 0.3\).
 This exclusion creates a hole in the manifold centered near the middle vertical region of the S-curve.

### 2025-06-24 paper-cardinalR.Rmd
Updated sphere data generation

@@
 
-A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$. Coordinates in the first two dimensions represent a perfect circle on the plane:
-
-$$
-X_1 = \cos(\theta), \quad X_2 = \sin(\theta)
-$$
-
+A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
+Coordinates in the first two dimensions represent a perfect circle on the plane: $X_1 = \cos(\theta), \quad X_2 = \sin(\theta)$.
 For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
@@
 The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2) \cdot \frac{\pi}{2p}$.
-
-Each additional dimension is computed as:
-
-$$
-X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p
-$$
-
-This structure retains a one-dimensional manifold (the circle) but allows it to be embedded non-linearly and decorrelatively in higher dimensions.
-The decreasing scale ensures that higher dimensions add less variation, preserving the dominance of the circular shape while adding complexity.
+Each additional dimension is computed as: $X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p$.
 
@@
 Together, these define a periodic, non-trivial, closed curve in $3\text{-}D$ with internal folds that produce a more complex geometry than a standard circle or ellipse.
-
 For dimensions $X_4$ through $X_p$, additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
-The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from 4 to $p$, which means that the amplitude decreases as the dimension increases.
+The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from $4$ to $p$, which means that the amplitude decreases as the dimension increases.
 Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j \cdot \sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2) \cdot \frac{\pi}{2p}$.
-
-This introduces decorrelated, low-variance sinusoidal variations across dimensions, retaining the cyclical nature while embedding the data in higher-dimensional space with a coherent but subtle structure.
 
@@
 Each observation lies on the surface of a sphere in $3\text{-}D$, constructed by generating $u \sim U(-1, 1)$ which represents the cosine of the polar angle $\phi$ and $\theta \sim U(0, 2\pi)$ which represents the azimuthal angle.
-
-The corresponding Cartesian coordinates are calculated as;
-
-$$
-X_1 = r \cdot \sqrt{1 - u^2} \cdot \cos(\theta),
-$$
-$$
-X_2 = r \cdot \sqrt{1 - u^2} \cdot \sin(\theta),
-$$
-$$
-X_3 = r \cdot u,
-$$
-
-which gives points uniformly distributed on the surface of a $3\text{-}D$ sphere (not within).
-
-For $p > 3$, additional dimensions $X_4$ to $X_p$ are generated as low-variance Gaussian noise:
-
-$$
-X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p.
-$$
-
-This extends the $3\text{-}D$ spherical manifold into $p$-dimensional space with orthogonal Gaussian noise, preserving the spherical structure while allowing for testing in higher-dimensional settings.
+The corresponding Cartesian coordinates are calculated as; $X_1 = r \cdot \sqrt{1 - u^2} \cdot \cos(\theta)$, $X_2 = r \cdot \sqrt{1 - u^2} \cdot \sin(\theta)$, $X_3 = r \cdot u$, which gives points uniformly distributed on the surface of a $3\text{-}D$ sphere (not within).
+For $p > 3$, additional dimensions $X_4$ to $X_p$ are generated as low-variance Gaussian noise: $X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p$.
 
@@
 
-The base structure consists of a $3\text{-}D$ spherical surface created using a regular grid in spherical coordinates, where $\theta \in [0, 2\pi]$ represents the azimuthal angle (longitude) and $\phi \in [0, \pi]$ denotes the polar angle (co-latitude).
+The base structure consists of a $3\text{-}D$ spherical surface created using a regular grid in spherical coordinates, where $\theta \in [0, 2\pi]$ represents the azimuthal angle and $\phi \in [0, \pi]$ denotes the polar angle.
 
@@
 
-Each point on the sphere is computed using the spherical-to-Cartesian transformation:
-
-$$
-X_1 = \sin(\phi) \cdot \cos(\theta),
-$$
-$$
-X_2 = \sin(\phi) \cdot \sin(\theta),
-$$
-$$
-X_3 = \cos(\phi).
-$$
-
-This forms a structured grid of points on the unit sphere, ideal for studying how regular geometric manifolds are transformed under nonlinear dimensionality reduction.
-
-For $p > 3$, the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions:
-
-$$
-X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p.
-$$
-
-These dimensions represent low-variance orthogonal noise, preserving the spherical shape while testing robustness in high-dimensional settings.
+Each point on the sphere is computed using the spherical-to-Cartesian transformation: $X_1 = \sin(\phi) \cdot \cos(\theta)$, $X_2 = \sin(\phi) \cdot \sin(\theta)$, $X_3 = \cos(\phi)$.
+For $p > 3$, the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions: $X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p$.
 
@@
 
-The `gen_clusteredspheres(n, k, p, r, loc)` function generates a synthetic dataset of $n_1 + k \cdot n_2$ observations in $p$-dimensional space, consisting of one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location.
-
-A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p$-dimensional sphere with a radius of $r_1$.
+The `gen_clusteredspheres(n, k, p, r, loc)` function generates a synthetic dataset of $n_1 + kn_2$ observations in $p$-dimensional space, consisting of one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location.
+
+A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
 The sampling is executed using the function `gen_unifsphere(n_1, p, r_1)`, which generates the desired points in the specified dimensional space.
@@
 
-Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for 3D).
-
+Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for $3\text{-}D$).
 Each observation is classified by cluster, with labels such as "big" for the large central sphere and "small_1" to "small_k" for the smaller spheres.
@@
 Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
-
-The coordinates are transformed into $4\text{-}D$ Cartesian space:
-
-$$
-X_1 = \sin(\theta_1) \cdot \cos(\theta_2),
-$$
-$$
-X_2 = \sin(\theta_1) \cdot \sin(\theta_2),
-$$
-$$
-X_3 = \cos(\theta_1) \cdot \cos(\theta_3),
-$$
-$$
-X_4 = \cos(\theta_1) \cdot \sin(\theta_3).
-$$
-
+The coordinates are transformed into $4\text{-}D$ Cartesian space: $X_1 = \sin(\theta_1) \cdot \cos(\theta_2)$, $X_2 = \sin(\theta_1) \cdot \sin(\theta_2)$, $X_3 = \cos(\theta_1) \cdot \cos(\theta_3)$, $X_4 = \cos(\theta_1) \cdot \sin(\theta_3)$.
 This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
-
-For $p > 4$, additional dimensions $X_5$ to $X_p$ are added as low-variance Gaussian noise:
-
-$$
-X_j \sim N(0, 0.05^2),\text{ for }j = 5, \dots, p.
-$$
-
-These dimensions maintain the spherical structure while simulating embedding into higher-dimensional space with small orthogonal perturbations.
+For $p > 4$, additional dimensions $X_5$ to $X_p$ are added as low-variance Gaussian noise: $X_j \sim N(0, 0.05^2),\text{ for }j = 5, \dots, p$.
 

### 2025-06-24 paper-cardinalR.tex
Updated sphere data generation

@@
 
-A latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\). Coordinates in the first two dimensions represent a perfect circle on the plane:
-
-\[
-X_1 = \cos(\theta), \quad X_2 = \sin(\theta)
-\]
-
+A latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
+Coordinates in the first two dimensions represent a perfect circle on the plane: \(X_1 = \cos(\theta), \quad X_2 = \sin(\theta)\).
 For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
@@
 The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula \(\phi_j = (j - 2) \cdot \frac{\pi}{2p}\).
-
-Each additional dimension is computed as:
-
-\[
-X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p
-\]
-
-This structure retains a one-dimensional manifold (the circle) but allows it to be embedded non-linearly and decorrelatively in higher dimensions.
-The decreasing scale ensures that higher dimensions add less variation, preserving the dominance of the circular shape while adding complexity.
+Each additional dimension is computed as: \(X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p\).
 
@@
 Together, these define a periodic, non-trivial, closed curve in \(3\text{-}D\) with internal folds that produce a more complex geometry than a standard circle or ellipse.
-
 For dimensions \(X_4\) through \(X_p\), additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
-The scaling factor is defined as \(\text{scale}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from 4 to \(p\), which means that the amplitude decreases as the dimension increases.
+The scaling factor is defined as \(\text{scale}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from \(4\) to \(p\), which means that the amplitude decreases as the dimension increases.
 Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j \cdot \sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2) \cdot \frac{\pi}{2p}\).
 
-This introduces decorrelated, low-variance sinusoidal variations across dimensions, retaining the cyclical nature while embedding the data in higher-dimensional space with a coherent but subtle structure.
-
 \begin{verbatim}
@@
 Each observation lies on the surface of a sphere in \(3\text{-}D\), constructed by generating \(u \sim U(-1, 1)\) which represents the cosine of the polar angle \(\phi\) and \(\theta \sim U(0, 2\pi)\) which represents the azimuthal angle.
-
-The corresponding Cartesian coordinates are calculated as;
-
-\[
-X_1 = r \cdot \sqrt{1 - u^2} \cdot \cos(\theta),
-\]
-\[
-X_2 = r \cdot \sqrt{1 - u^2} \cdot \sin(\theta),
-\]
-\[
-X_3 = r \cdot u,
-\]
-
-which gives points uniformly distributed on the surface of a \(3\text{-}D\) sphere (not within).
-
-For \(p > 3\), additional dimensions \(X_4\) to \(X_p\) are generated as low-variance Gaussian noise:
-
-\[
-X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p.
-\]
-
-This extends the \(3\text{-}D\) spherical manifold into \(p\)-dimensional space with orthogonal Gaussian noise, preserving the spherical structure while allowing for testing in higher-dimensional settings.
+The corresponding Cartesian coordinates are calculated as; \(X_1 = r \cdot \sqrt{1 - u^2} \cdot \cos(\theta)\), \(X_2 = r \cdot \sqrt{1 - u^2} \cdot \sin(\theta)\), \(X_3 = r \cdot u\), which gives points uniformly distributed on the surface of a \(3\text{-}D\) sphere (not within).
+For \(p > 3\), additional dimensions \(X_4\) to \(X_p\) are generated as low-variance Gaussian noise: \(X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p\).
 
@@
 
-The base structure consists of a \(3\text{-}D\) spherical surface created using a regular grid in spherical coordinates, where \(\theta \in [0, 2\pi]\) represents the azimuthal angle (longitude) and \(\phi \in [0, \pi]\) denotes the polar angle (co-latitude).
+The base structure consists of a \(3\text{-}D\) spherical surface created using a regular grid in spherical coordinates, where \(\theta \in [0, 2\pi]\) represents the azimuthal angle and \(\phi \in [0, \pi]\) denotes the polar angle.
 
@@
 
-Each point on the sphere is computed using the spherical-to-Cartesian transformation:
-
-\[
-X_1 = \sin(\phi) \cdot \cos(\theta),
-\]
-\[
-X_2 = \sin(\phi) \cdot \sin(\theta),
-\]
-\[
-X_3 = \cos(\phi).
-\]
-
-This forms a structured grid of points on the unit sphere, ideal for studying how regular geometric manifolds are transformed under nonlinear dimensionality reduction.
-
-For \(p > 3\), the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions:
-
-\[
-X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p.
-\]
-
-These dimensions represent low-variance orthogonal noise, preserving the spherical shape while testing robustness in high-dimensional settings.
+Each point on the sphere is computed using the spherical-to-Cartesian transformation: \(X_1 = \sin(\phi) \cdot \cos(\theta)\), \(X_2 = \sin(\phi) \cdot \sin(\theta)\), \(X_3 = \cos(\phi)\).
+For \(p > 3\), the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions: \(X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p\).
 
@@
 
-The \texttt{gen\_clusteredspheres(n,\ k,\ p,\ r,\ loc)} function generates a synthetic dataset of \(n_1 + k \cdot n_2\) observations in \(p\)-dimensional space, consisting of one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location.
-
-A large uniform sphere centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\)-dimensional sphere with a radius of \(r_1\).
+The \texttt{gen\_clusteredspheres(n,\ k,\ p,\ r,\ loc)} function generates a synthetic dataset of \(n_1 + kn_2\) observations in \(p\)-dimensional space, consisting of one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location.
+
+A large uniform sphere centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\text{-}D\) sphere with a radius of \(r_1\).
 The sampling is executed using the function \texttt{gen\_unifsphere(n\_1,\ p,\ r\_1)}, which generates the desired points in the specified dimensional space.
@@
 
-Points on spheres are generated using the standard hyperspherical method, which involves sampling \(u \sim U(-1, 1)\) to determine the cosine of the polar angle, and sampling \(\theta \sim U(0, 2\pi)\) to determine the azimuthal angle (for 3D).
-
+Points on spheres are generated using the standard hyperspherical method, which involves sampling \(u \sim U(-1, 1)\) to determine the cosine of the polar angle, and sampling \(\theta \sim U(0, 2\pi)\) to determine the azimuthal angle (for \(3\text{-}D\)).
 Each observation is classified by cluster, with labels such as ``big'' for the large central sphere and ``small\_1'' to ``small\_k'' for the smaller spheres.
@@
 Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
-
-The coordinates are transformed into \(4\text{-}D\) Cartesian space:
-
-\[
-X_1 = \sin(\theta_1) \cdot \cos(\theta_2),
-\]
-\[
-X_2 = \sin(\theta_1) \cdot \sin(\theta_2),
-\]
-\[
-X_3 = \cos(\theta_1) \cdot \cos(\theta_3),
-\]
-\[
-X_4 = \cos(\theta_1) \cdot \sin(\theta_3).
-\]
-
+The coordinates are transformed into \(4\text{-}D\) Cartesian space: \(X_1 = \sin(\theta_1) \cdot \cos(\theta_2)\), \(X_2 = \sin(\theta_1) \cdot \sin(\theta_2)\), \(X_3 = \cos(\theta_1) \cdot \cos(\theta_3)\), \(X_4 = \cos(\theta_1) \cdot \sin(\theta_3)\).
 This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
-
-For \(p > 4\), additional dimensions \(X_5\) to \(X_p\) are added as low-variance Gaussian noise:
-
-\[
-X_j \sim N(0, 0.05^2),\text{ for }j = 5, \dots, p.
-\]
-
-These dimensions maintain the spherical structure while simulating embedding into higher-dimensional space with small orthogonal perturbations.
+For \(p > 4\), additional dimensions \(X_5\) to \(X_p\) are added as low-variance Gaussian noise: \(X_j \sim N(0, 0.05^2),\text{ for }j = 5, \dots, p\).
 

### 2025-06-24 paper-cardinalR.Rmd
Updated Pyramid function definitions

@@
 The `gen_pyrrect(n, p, h, l_vec, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a rectangular base.
-The pyramid is embedded in a $p$-dimensional space, with a tip at height zero and base at height $h$.
+The pyramid is embedded in $p\text{-}D$, with a tip at height zero and base at height $h$.
 The shape tapers linearly from the base dimensions $(l_x, l_y)$ to a smaller rectangular section with side lengths $(2r_t, 2r_t)$ at the tip.
@@
 
-Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final dimension $x_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$:
-
-$$
-x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)
-$$.
-
-Let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$:
-
-$$
-r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}, \quad
-r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}
-$$.
-
-The first three coordinates are then defined as:
-
-$$
-x_1 \sim U(-r_x(z),\ r_x(z)), \quad
-x_2 \sim U(-r_y(z),\ r_y(z)), \quad
-x_3 \sim U(-r_x(z),\ r_x(z))
-$$.
-
-For $p > 3$, the remaining $p - 4$ dimensions (i.e., $x_4$ to $x_{p-1}$) are tapered toward zero with decreasing height and generated as:
-
-$$
-x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
-$$
-
+Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
+The final dimension $X_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$: $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$.
+Let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$: $r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}$, $r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}$.
+The first three coordinates are then defined as: $X_1 \sim U(-r_x(z),\ r_x(z))$, $X_2 \sim U(-r_y(z),\ r_y(z))$, and $X_3 \sim U(-r_x(z),\ r_x(z))$.
+For $p > 3$, the remaining $p - 4$ dimensions (i.e., $X_4$ to $X_{p-1}$) are tapered toward zero with decreasing height and generated as: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1$.
 This tapering ensures that the structure narrows in all directions as it approaches the tip.
@@
 The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section.
-The structure is embedded in a $p$-dimensional space, with the tip located at height 0 and the base at height $h$.
+The structure is embedded in a $p\text{-}D$, with the tip located at height $0$ and the base at height $h$.
 The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
 
-Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points.
-The final coordinate \$x_p\$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:
-
-$$
-x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
-$$
-
-Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$:
-
-$$
-r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
-$$
-
-A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section:
-
-$$
-u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
-$$
-
-The first three coordinates (triangle plane) are then:
-
-$$
-x_1 = r(z)(1 - u - v), \quad
-x_2 = r(z) \cdot u, \quad
-x_3 = r(z) \cdot v
-$$
-
-For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip:
-
-$$
-x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
-$$
-
-The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section.
-The structure is embedded in a $p$-dimensional space, with the tip located at height 0 and the base at height $h$.
-The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
-
-Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final coordinate $x_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:
-
-$$
-x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
-$$
-
-Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$:
-
-$$
-r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
-$$
-
-A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section:
-
-$$
-u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
-$$
-
-The first three coordinates (triangle plane) are then:
-
-$$
-x_1 = r(z)(1 - u - v), \quad
-x_2 = r(z) \cdot u, \quad
-x_3 = r(z) \cdot v
-$$
-
-For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip:
-
-$$
-x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
-$$
-
-The result is a tapered, high-dimensional triangle-based pyramid with more points near the apex and greater width at the base.
+Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
+The final coordinate $X_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$: $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$.
+Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$: $r(z) = r_t + (l - r_t) \cdot \frac{z}{h}$.
+
+A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
+The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z) \cdot u$, and $X_3 = r(z) \cdot v$.
+For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right)$.
 
@@
 
-The `gen_pyrstar(n, p, h, rb)` function generates a dataset of \$n\$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
+The `gen_pyrstar(n, p, h, rb)` function generates a dataset of $n$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
 The pyramid extends vertically from height $z = 0$ (tip) to height $z = h$ (base), embedded in a $p$-dimensional space.
@@
 
-Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final coordinate $x_p$ encodes the height $z$:
-
-$$
-x_p = z \sim U(0, h).
-$$
-
-The radius at height $z$ scales linearly from zero (tip) to the base radius $r_b$:
-
-$$
-r(z) = r_b \cdot \left(1 - \frac{z}{h}\right).
-$$
-
-Each point is placed within a regular hexagon in the plane $(x_1, x_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor:
-
-$$
-\theta \sim \text{Uniform sample from 6 hexagon angles}, \quad
-r_{\text{point}} \sim \sqrt{U(0, 1)}.
-$$
-
-Then, the first two coordinates are:
-
-$$
-x_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta), \quad
-x_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)
-$$
-
-For dimensions $j = 3, \dots, p-1$, values are tapered toward zero near the tip:
-
-$$
-x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
-$$
-
-This results in a smooth, star-like pyramid in $p$-dimensional space, with a broad hexagonal base and narrow tip.
+Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points. The final coordinate $X_p$ encodes the height $z$: $X_p = z \sim U(0, h)$.
+
+The radius at height $z$ scales linearly from zero (tip) to the base radius $r_b$: $r(z) = r_b \cdot \left(1 - \frac{z}{h}\right)$.
+
+Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
+$r_{\text{point}} \sim \sqrt{U(0, 1)}$.
+Then, the first two coordinates are: $X_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta)$, and $X_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)$.
+For dimensions $j = 3, \dots, p-1$, values are tapered toward zero near the tip: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right)$.
 
@@
 
-The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p$-dimensional simplex using a chaotic attractor-like midpoint algorithm.
+The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p\text{-}D$ simplex using a chaotic attractor-like midpoint algorithm.
 The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a "Sierpinski-like pyramid" in high dimensions.
 
-Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution:
-
-$$
-T_0 \sim U(0, 1)^p.
-$$
-
-Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p$-dimensional simplex.
-At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$:
-
-$$
-T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}.
-$$
-
-This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points. The points remain bounded inside the convex hull of the simplex.
-
-The final output is a $n \times p$ matrix where each row represents a point:
-
-$$
-X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}.
-$$
+Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
+The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
+
+Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
+At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
+This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
+The points remain bounded inside the convex hull of the simplex.
+The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
 

### 2025-06-24 paper-cardinalR.tex
Updated Pyramid function definitions

@@
 The \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a rectangular base.
-The pyramid is embedded in a \(p\)-dimensional space, with a tip at height zero and base at height \(h\).
+The pyramid is embedded in \(p\text{-}D\), with a tip at height zero and base at height \(h\).
 The shape tapers linearly from the base dimensions \((l_x, l_y)\) to a smaller rectangular section with side lengths \((2r_t, 2r_t)\) at the tip.
@@
 
-Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The final dimension \(x_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\):
-
-\[
-x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)
-\].
-
-Let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\):
-
-\[
-r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}, \quad
-r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}
-\].
-
-The first three coordinates are then defined as:
-
-\[
-x_1 \sim U(-r_x(z),\ r_x(z)), \quad
-x_2 \sim U(-r_y(z),\ r_y(z)), \quad
-x_3 \sim U(-r_x(z),\ r_x(z))
-\].
-
-For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(x_4\) to \(x_{p-1}\)) are tapered toward zero with decreasing height and generated as:
-
-\[
-x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
-\]
-
+Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
+The final dimension \(X_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\): \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\).
+Let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\): \(r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}\), \(r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}\).
+The first three coordinates are then defined as: \(X_1 \sim U(-r_x(z),\ r_x(z))\), \(X_2 \sim U(-r_y(z),\ r_y(z))\), and \(X_3 \sim U(-r_x(z),\ r_x(z))\).
+For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are tapered toward zero with decreasing height and generated as: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1\).
 This tapering ensures that the structure narrows in all directions as it approaches the tip.
@@
 The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section.
-The structure is embedded in a \(p\)-dimensional space, with the tip located at height 0 and the base at height \(h\).
+The structure is embedded in a \(p\text{-}D\), with the tip located at height \(0\) and the base at height \(h\).
 The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
 
-Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points.
-The final coordinate \$x\_p\$ encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):
-
-\[
-x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
-\]
-
-Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\):
-
-\[
-r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
-\]
-
-A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section:
-
-\[
-u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
-\]
-
-The first three coordinates (triangle plane) are then:
-
-\[
-x_1 = r(z)(1 - u - v), \quad
-x_2 = r(z) \cdot u, \quad
-x_3 = r(z) \cdot v
-\]
-
-For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip:
-
-\[
-x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
-\]
-
-The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section.
-The structure is embedded in a \(p\)-dimensional space, with the tip located at height 0 and the base at height \(h\).
-The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
-
-Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points.
-The final coordinate \(x_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):
-
-\[
-x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
-\]
-
-Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\):
-
-\[
-r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
-\]
-
-A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section:
-
-\[
-u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
-\]
-
-The first three coordinates (triangle plane) are then:
-
-\[
-x_1 = r(z)(1 - u - v), \quad
-x_2 = r(z) \cdot u, \quad
-x_3 = r(z) \cdot v
-\]
-
-For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip:
-
-\[
-x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
-\]
-
-The result is a tapered, high-dimensional triangle-based pyramid with more points near the apex and greater width at the base.
+Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
+The final coordinate \(X_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\): \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\).
+Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\): \(r(z) = r_t + (l - r_t) \cdot \frac{z}{h}\).
+
+A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section: \(u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v\).
+The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z) \cdot u\), and \(X_3 = r(z) \cdot v\).
+For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right)\).
 
@@
 
-The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \$n\$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
+The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
 The pyramid extends vertically from height \(z = 0\) (tip) to height \(z = h\) (base), embedded in a \(p\)-dimensional space.
@@
 
-Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The final coordinate \(x_p\) encodes the height \(z\):
-
-\[
-x_p = z \sim U(0, h).
-\]
-
-The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r_b\):
-
-\[
-r(z) = r_b \cdot \left(1 - \frac{z}{h}\right).
-\]
-
-Each point is placed within a regular hexagon in the plane \((x_1, x_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}\) and a uniformly random radial scaling factor:
-
-\[
-\theta \sim \text{Uniform sample from 6 hexagon angles}, \quad
-r_{\text{point}} \sim \sqrt{U(0, 1)}.
-\]
-
-Then, the first two coordinates are:
-
-\[
-x_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta), \quad
-x_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)
-\]
-
-For dimensions \(j = 3, \dots, p-1\), values are tapered toward zero near the tip:
-
-\[
-x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
-\]
-
-This results in a smooth, star-like pyramid in \(p\)-dimensional space, with a broad hexagonal base and narrow tip.
+Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points. The final coordinate \(X_p\) encodes the height \(z\): \(X_p = z \sim U(0, h)\).
+
+The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r_b\): \(r(z) = r_b \cdot \left(1 - \frac{z}{h}\right)\).
+
+Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
+\(r_{\text{point}} \sim \sqrt{U(0, 1)}\).
+Then, the first two coordinates are: \(X_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta)\), and \(X_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)\).
+For dimensions \(j = 3, \dots, p-1\), values are tapered toward zero near the tip: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right)\).
 
@@
 
-The \texttt{gen\_pyrholes(n,\ p)} function generates \(n\) points embedded in a \(p\)-dimensional simplex using a chaotic attractor-like midpoint algorithm.
+The \texttt{gen\_pyrholes(n,\ p)} function generates \(n\) points embedded in a \(p\text{-}D\) simplex using a chaotic attractor-like midpoint algorithm.
 The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a ``Sierpinski-like pyramid'' in high dimensions.
 
-Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution:
-
-\[
-T_0 \sim U(0, 1)^p.
-\]
-
-Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\)-dimensional simplex.
-At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\):
-
-\[
-T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}.
-\]
-
-This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points. The points remain bounded inside the convex hull of the simplex.
-
-The final output is a \(n \times p\) matrix where each row represents a point:
-
-\[
-X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}.
-\]
+Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
+The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution: \(T_0 \sim U(0, 1)^p\).
+
+Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\text{-}D\) simplex.
+At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\): \(T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}\).
+This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
+The points remain bounded inside the convex hull of the simplex.
+The final output is a \(n \times p\) matrix where each row represents a point: \(X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}\).
 

### 2025-06-24 paper-cardinalR.Rmd
Updated trigometric functions

@@
 
-The first two dimensions define a crescent-shaped curve using a semi-circular arc. Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
-
-The corresponding $2\text{-}D$ coordinates are defined by:
-
-$$
-X_1 = \cos(\theta),
-$$
-$$
-X_2 = \sin(\theta).
-$$
-
-This forms a nonlinear curvilinear structure in the shape of a tilted crescent in the $X_1$-$X_2$ plane.
-
-For $p > 2$, additional dimensions $X_3$ to $X_p$ are generated using `gen_wavydims1()`:
-
-$$
-X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
-$$
-
+The first two dimensions define a crescent-shaped curve using a semi-circular arc.
+Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
+The corresponding $2\text{-}D$ coordinates are defined by: $X_1 = \cos(\theta)$, $X_2 = \sin(\theta)$.
+For $p > 2$, additional dimensions $X_3$ to $X_p$ are generated using `gen_wavydims1()`: $X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}$.
 These noise dimensions are nonlinearly correlated with $\theta$, producing structured, wave-like variation that aligns with the progression along the crescent curve.
@@
 
-The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension:
-
+The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
 Let $\theta \sim U(0, 3\pi)$ represent a random angle on a circular base and $z \sim U(0, h)$ represent the height along the cylinder.
-
-The coordinates are defined as:
-
-* $X_1 = \cos(\theta)$ (Circular base, x-axis),
-* $X_2 = \sin(\theta)$ (Circular base, y-axis),
-* $X_3 = z$ (Linear height),
-* $X_4 = \sin(z)$ (Nonlinear curvy variation along height).
-
+The coordinates are defined as: $X_1 = \cos(\theta)$ (Circular base, x-axis), $X_2 = \sin(\theta)$ (Circular base, y-axis), $X_3 = z$ (Linear height), and $X_4 = \sin(z)$ (Nonlinear curvy variation along height).
 This forms a curvy cylindrical surface in $4\text{-}D$, where the fourth dimension bends periodically along the height axis, resembling a helicoid or twisting wave along the cylinder.
-
-For $p > 4$, dimensions $X_5$ through $X_p$ are generated as:
-
-$$
-X_j \sim N(0, 0.05^2),\text{ for }j = 5, \ldots, p.
-$$
-
-These dimensions represent unstructured, low-variance noise.
+For $p > 4$, dimensions $X_5$ through $X_p$ are generated as: $X_j \sim N(0, 0.05^2),\text{ for }j = 5, \ldots, p$.
 
@@
 
-Cartesian coordinates from spherical conversion:
-
-  * $X_1 = \sin(\phi) \cdot \cos(\theta)$
-  * $X_2 = \sin(\phi) \cdot \sin(\theta)$
-  * $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter.
-  * $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
-
+Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi) \cdot \cos(\theta)$, $X_2 = \sin(\phi) \cdot \sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
 This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
@@
 These dimensions are functions of the first coordinate $X_1$, introducing dependencies that preserve some geometric coherence.
-Each added dimension follows the form:
-
-$$
-X_j = s_j \cdot (-1)^{\lfloor j/2 \rfloor} \cdot X_1^{a_j} + \eta_j,
-$$
-
-where:
-$a_j \in \{2, 3, 4, 5\}$ is a randomly chosen polynomial power,
-$s_j \sim U(0.5, 2)$ is a scale factor,
-$\eta_j \sim U(-\sigma_j, 2\sigma_j)$, with $\sigma_j \sim U(0, 0.05)$, adds mild randomness.
+Each added dimension follows the form: $X_j = s_j \cdot (-1)^{\lfloor j/2 \rfloor} \cdot X_1^{a_j} + \eta_j$, where $a_j \in \{2, 3, 4, 5\}$ is a randomly chosen polynomial power, $s_j \sim U(0.5, 2)$ is a scale factor, $\eta_j \sim U(-\sigma_j, 2\sigma_j)$, with $\sigma_j \sim U(0, 0.05)$, adds mild randomness.
 
@@
 
-The `gen_helicalspiral(n, p)` function generates a dataset of $n$ observations forming a helical spiral, embedded in the first four dimensions of a $p$-dimensional space.
+The `gen_helicalspiral(n, p)` function generates a dataset of $n$ observations forming a helical spiral, embedded in the first four dimensions of $p\text{-}D$.
 Additional dimensions (if $p > 4$) are filled with structured noise to mimic high-dimensional complexity.
 
-The first four coordinates follow a $3\text{-}D$ helix with an additional oscillating component. Let $\theta \in [0, \frac{5\pi}{4}]$ be a sequence of angles controlling rotation around a circle.
-
-Cartesian coordinates:
-  * $X_1 = \cos(\theta)$: circular trajectory along the x-axis.
-
-  * $X_2 = \sin(\theta)$: circular trajectory along the y-axis.
-
-  * $X_3 = 0.05 \cdot \theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix.
-
-  * $X_4 = 0.1 \cdot \sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
-
+The first four coordinates follow a $3\text{-}D$ helix with an additional oscillating component.
+Let $\theta \in [0, \frac{5\pi}{4}]$ be a sequence of angles controlling rotation around a circle.
+Cartesian coordinates; $X_1 = \cos(\theta)$: circular trajectory along the x-axis, $X_2 = \sin(\theta)$: circular trajectory along the y-axis, $X_3 = 0.05 \cdot \theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix, and $X_4 = 0.1 \cdot \sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
 This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
-
-When $p > 4$, the remaining $p - 4$ dimensions are populated using `gen_noisedims()`. These are independent Gaussian noise dimensions:
-
-$$
-X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
-$$
+When $p > 4$, the remaining $p - 4$ dimensions are populated using `gen_noisedims()`.
+These are independent Gaussian noise dimensions: $X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p$.
 
@@
 
-The `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of a $p$-dimensional space.
+The `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$.
 The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a $3\text{-}D$ conic helix.
 
-The shape is defined by parameter $\theta \in [0, 2\pi \cdot \text{spins}]$, controlling the angular progression of the spiral. The mapping to Cartesian space is as follows:
-
-The Archimedean spiral in the horizontal plane is represented by;
-
-$X_1 = \theta \cdot \cos(\theta)$ for radial expansion in x.
-$X_2 = \theta \cdot \sin(\theta)$ for radial expansion in y.
-
-The growth pattern resembles a cone, with the height increasing according to $X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$
-
-Spiral modulation in the fourth dimension is represented by $X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
-
+The shape is defined by parameter $\theta \in [0, 2\pi \cdot \text{spins}]$, controlling the angular progression of the spiral.
+The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta \cdot \cos(\theta)$ for radial expansion in x, and $X_2 = \theta \cdot \sin(\theta)$ for radial expansion in y.
+The growth pattern resembles a cone, with the height increasing according to $X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
 This results in a $3\text{-}D$ spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
-
-For $p > 4$, the dataset includes isotropic Gaussian noise in the remaining dimensions via `gen_noisedims()`:
-
-$$
-X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
-$$
+For $p > 4$, the dataset includes isotropic Gaussian noise in the remaining dimensions via `gen_noisedims()`: $X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p$.
 
@@
 The `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables.
-The key geometry lies in dimensions 1–4, where:
-
-$X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors).
-$X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable.
-$X_{2} = \frac{hc}{X_{1}} + \text{nonfac} \cdot \sin(X_{1})$: non-linear combination of **hyperbolic and sinusoidal transformations**, creating sharp curvature and oscillation.
-$X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-
+
+The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = \frac{hc}{X_{1}} + \text{nonfac} \cdot \sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-
-If $p > 4$, the remaining dimensions are populated with Gaussian noise using `gen_noisedims()`:
-
-$$
-X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
-$$
+If $p > 4$, the remaining dimensions are populated with Gaussian noise using `gen_noisedims()`: $X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p$.
 

### 2025-06-24 paper-cardinalR.tex
Updated trigometric functions

@@
 
-The first two dimensions define a crescent-shaped curve using a semi-circular arc. Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
-
-The corresponding \(2\text{-}D\) coordinates are defined by:
-
-\[
-X_1 = \cos(\theta),
-\]
-\[
-X_2 = \sin(\theta).
-\]
-
-This forms a nonlinear curvilinear structure in the shape of a tilted crescent in the \(X_1\)-\(X_2\) plane.
-
-For \(p > 2\), additional dimensions \(X_3\) to \(X_p\) are generated using \texttt{gen\_wavydims1()}:
-
-\[
-X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
-\]
-
+The first two dimensions define a crescent-shaped curve using a semi-circular arc.
+Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
+The corresponding \(2\text{-}D\) coordinates are defined by: \(X_1 = \cos(\theta)\), \(X_2 = \sin(\theta)\).
+For \(p > 2\), additional dimensions \(X_3\) to \(X_p\) are generated using \texttt{gen\_wavydims1()}: \(X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}\).
 These noise dimensions are nonlinearly correlated with \(\theta\), producing structured, wave-like variation that aligns with the progression along the crescent curve.
@@
 
-The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension:
-
+The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
 Let \(\theta \sim U(0, 3\pi)\) represent a random angle on a circular base and \(z \sim U(0, h)\) represent the height along the cylinder.
-
-The coordinates are defined as:
-
-\begin{itemize}
-\tightlist
-\item
-  \(X_1 = \cos(\theta)\) (Circular base, x-axis),
-\item
-  \(X_2 = \sin(\theta)\) (Circular base, y-axis),
-\item
-  \(X_3 = z\) (Linear height),
-\item
-  \(X_4 = \sin(z)\) (Nonlinear curvy variation along height).
-\end{itemize}
-
+The coordinates are defined as: \(X_1 = \cos(\theta)\) (Circular base, x-axis), \(X_2 = \sin(\theta)\) (Circular base, y-axis), \(X_3 = z\) (Linear height), and \(X_4 = \sin(z)\) (Nonlinear curvy variation along height).
 This forms a curvy cylindrical surface in \(4\text{-}D\), where the fourth dimension bends periodically along the height axis, resembling a helicoid or twisting wave along the cylinder.
-
-For \(p > 4\), dimensions \(X_5\) through \(X_p\) are generated as:
-
-\[
-X_j \sim N(0, 0.05^2),\text{ for }j = 5, \ldots, p.
-\]
-
-These dimensions represent unstructured, low-variance noise.
+For \(p > 4\), dimensions \(X_5\) through \(X_p\) are generated as: \(X_j \sim N(0, 0.05^2),\text{ for }j = 5, \ldots, p\).
 
@@
 
-Cartesian coordinates from spherical conversion:
-
-\begin{itemize}
-\tightlist
-\item
-  \(X_1 = \sin(\phi) \cdot \cos(\theta)\)
-\item
-  \(X_2 = \sin(\phi) \cdot \sin(\theta)\)
-\item
-  \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter.
-\item
-  \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
-\end{itemize}
-
+Cartesian coordinates from spherical conversion: \(X_1 = \sin(\phi) \cdot \cos(\theta)\), \(X_2 = \sin(\phi) \cdot \sin(\theta)\), \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter, and \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
 This generates a spherical spiral curve embedded in \(4\text{-}D\) space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
@@
 These dimensions are functions of the first coordinate \(X_1\), introducing dependencies that preserve some geometric coherence.
-Each added dimension follows the form:
-
-\[
-X_j = s_j \cdot (-1)^{\lfloor j/2 \rfloor} \cdot X_1^{a_j} + \eta_j,
-\]
-
-where:
-\(a_j \in \{2, 3, 4, 5\}\) is a randomly chosen polynomial power,
-\(s_j \sim U(0.5, 2)\) is a scale factor,
-\(\eta_j \sim U(-\sigma_j, 2\sigma_j)\), with \(\sigma_j \sim U(0, 0.05)\), adds mild randomness.
+Each added dimension follows the form: \(X_j = s_j \cdot (-1)^{\lfloor j/2 \rfloor} \cdot X_1^{a_j} + \eta_j\), where \(a_j \in \{2, 3, 4, 5\}\) is a randomly chosen polynomial power, \(s_j \sim U(0.5, 2)\) is a scale factor, \(\eta_j \sim U(-\sigma_j, 2\sigma_j)\), with \(\sigma_j \sim U(0, 0.05)\), adds mild randomness.
 
@@
 
-The \texttt{gen\_helicalspiral(n,\ p)} function generates a dataset of \(n\) observations forming a helical spiral, embedded in the first four dimensions of a \(p\)-dimensional space.
+The \texttt{gen\_helicalspiral(n,\ p)} function generates a dataset of \(n\) observations forming a helical spiral, embedded in the first four dimensions of \(p\text{-}D\).
 Additional dimensions (if \(p > 4\)) are filled with structured noise to mimic high-dimensional complexity.
 
-The first four coordinates follow a \(3\text{-}D\) helix with an additional oscillating component. Let \(\theta \in [0, \frac{5\pi}{4}]\) be a sequence of angles controlling rotation around a circle.
-
-Cartesian coordinates:
-* \(X_1 = \cos(\theta)\): circular trajectory along the x-axis.
-
-\begin{itemize}
-\item
-  \(X_2 = \sin(\theta)\): circular trajectory along the y-axis.
-\item
-  \(X_3 = 0.05 \cdot \theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix.
-\item
-  \(X_4 = 0.1 \cdot \sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
-\end{itemize}
-
+The first four coordinates follow a \(3\text{-}D\) helix with an additional oscillating component.
+Let \(\theta \in [0, \frac{5\pi}{4}]\) be a sequence of angles controlling rotation around a circle.
+Cartesian coordinates; \(X_1 = \cos(\theta)\): circular trajectory along the x-axis, \(X_2 = \sin(\theta)\): circular trajectory along the y-axis, \(X_3 = 0.05 \cdot \theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix, and \(X_4 = 0.1 \cdot \sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
 This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
-
-When \(p > 4\), the remaining \(p - 4\) dimensions are populated using \texttt{gen\_noisedims()}. These are independent Gaussian noise dimensions:
-
-\[
-X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
-\]
+When \(p > 4\), the remaining \(p - 4\) dimensions are populated using \texttt{gen\_noisedims()}.
+These are independent Gaussian noise dimensions: \(X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p\).
 
@@
 
-The \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of a \(p\)-dimensional space.
+The \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of \(p\text{-}D\).
 The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a \(3\text{-}D\) conic helix.
 
-The shape is defined by parameter \(\theta \in [0, 2\pi \cdot \text{spins}]\), controlling the angular progression of the spiral. The mapping to Cartesian space is as follows:
-
-The Archimedean spiral in the horizontal plane is represented by;
-
-\(X_1 = \theta \cdot \cos(\theta)\) for radial expansion in x.
-\(X_2 = \theta \cdot \sin(\theta)\) for radial expansion in y.
-
-The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\)
-
-Spiral modulation in the fourth dimension is represented by \(X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
-
+The shape is defined by parameter \(\theta \in [0, 2\pi \cdot \text{spins}]\), controlling the angular progression of the spiral.
+The Archimedean spiral in the horizontal plane is represented by; \(X_1 = \theta \cdot \cos(\theta)\) for radial expansion in x, and \(X_2 = \theta \cdot \sin(\theta)\) for radial expansion in y.
+The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\) Spiral modulation in the fourth dimension is represented by \(X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
 This results in a \(3\text{-}D\) spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
-
-For \(p > 4\), the dataset includes isotropic Gaussian noise in the remaining dimensions via \texttt{gen\_noisedims()}:
-
-\[
-X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
-\]
+For \(p > 4\), the dataset includes isotropic Gaussian noise in the remaining dimensions via \texttt{gen\_noisedims()}: \(X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p\).
 
@@
 The \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables.
-The key geometry lies in dimensions 1--4, where:
-
-\(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors).
-\(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable.
-\(X_{2} = \frac{hc}{X_{1}} + \text{nonfac} \cdot \sin(X_{1})\): non-linear combination of \textbf{hyperbolic and sinusoidal transformations}, creating sharp curvature and oscillation.
-\(X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-
+
+The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = \frac{hc}{X_{1}} + \text{nonfac} \cdot \sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-
-If \(p > 4\), the remaining dimensions are populated with Gaussian noise using \texttt{gen\_noisedims()}:
-
-\[
-X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p.
-\]
+If \(p > 4\), the remaining dimensions are populated with Gaussian noise using \texttt{gen\_noisedims()}: \(X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p\).
 

### 2025-06-24 paper-cardinalR.Rmd
Updated branching last two sections

@@
 
-Branch $1$ is generated from $X_1 \sim U(0, 1)$, and $X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
+Branch $1$ is generated from $X_1 \sim U(0, 1)$, and $X_2 = 0.1X_1 + X_1^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
 This produces a gently upward-curving parabola in the right half-plane.
-Branch $2$ is generated from $X_{1} \sim U(-1, 0)$, and $X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
+Branch $2$ is generated from $X_{1} \sim U(-1, 0)$, and $X_2 = 0.1X_{1} - 2X_{1}^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
 This creates a steeper, leftward-facing curve in the left half-plane.
@@
 
-The `gen_orglinearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches embedded in a $p$-dimensional space.
+The `gen_orglinearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches embedded in $p\text{-}D$.
 Each branch lies primarily within a distinct $2\text{-}D$ subspace, while the remaining $p - 2$ dimensions contain Gaussian noise.
@@
 If $k > \binom{p}{2}$, additional pairs are sampled with replacement to reach $k$ total branches.
-Each selected pair $(x_{i1}, x_{i2})$ defines the $2\text{-}D$ plane for branch $i$.
-
-For branch $i$, $n_i$ points are generated, where $\sum_{i=1}^k n_i = n$. The structure follows:
-
-$$
-x_{i1} \sim U(0, 2), \quad
-x_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5),
-$$
-
-where $s_i$ is a scale factor controlling the slope. When $k \leq \binom{p}{2}$, $s_i = 1$. When sampling with replacement, $s_i$ is drawn from the set ${1, 1.5, 2, \dots, 8}$ in increments of $0.5$.
-
-For $p > 2$, the remaining dimensions contain independent Gaussian noise:
-
-$$
-x_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
-$$
+Each selected pair $(X_{i1}, X_{i2})$ defines the $2\text{-}D$ plane for branch $i$.
+
+For branch $i$, $n_i$ points are generated, where $\sum_{i=1}^k n_i = n$.
+The structure follows: $X_{i1} \sim U(0, 2)$, and $X_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5)$, where $s_i$ is a scale factor controlling the slope.
+When $k \leq \binom{p}{2}$, $s_i = 1$.
+When sampling with replacement, $s_i$ is drawn from the set ${1, 1.5, 2, \dots, 8}$ in increments of $0.5$.
+For $p > 2$, the remaining dimensions contain independent Gaussian noise: $X_j \sim N(0, 0.1^2)$, for $j \notin \{i1, i2\}$.
 
@@
 Let $\binom{p}{2}$ denote the number of unique $2\text{-}D$ subspace combinations.
-When $k \leq \binom{p}{2}$, $k$ distinct subspace pairs $(x_{i1}, x_{i2})$ are sampled without replacement.
+When $k \leq \binom{p}{2}$, $k$ distinct subspace pairs $(X_{i1}, X_{i2})$ are sampled without replacement.
 Otherwise, combinations are selected with replacement to reach $k$ total branches.
-For each branch $i = 1, \dots, k$, a random scale factor $s_i$ is used to vary the curvature:
-
-* If $k \leq \binom{p}{2}$: $s_i = 1$
-* If $k > \binom{p}{2}$: $s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}$
+For each branch $i = 1, \dots, k$, a random scale factor $s_i$ is used to vary the curvature; If $k \leq \binom{p}{2}$: $s_i = 1$, or If $k > \binom{p}{2}$: $s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}$.
 
 Each branch contains $n_i$ points such that $\sum_{i=1}^k n_i = n$, where the vector $(n_1, \dots, n_k)$ is randomly drawn using the helper function `gen_nsum()` to partition $n$.
-
-Within its assigned subspace $(x_{i1}, x_{i2})$, branch $i$ is defined by:
-
-$$
-x_{i1} \sim U(0, 2), \quad
-x_{i2} = -s_i \cdot x_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)
-$$
-
-This forms a smooth downward-opening parabola in the plane defined by $(x_{i1}, x_{i2})$, with the degree of curvature controlled by $s_i$.
-
-For $p > 2$, all remaining dimensions are independent Gaussian noise:
-
-$$
-x_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
-$$
+Within its assigned subspace $(X_{i1}, X_{i2})$, branch $i$ is defined by: $X_{i1} \sim U(0, 2)$, and $X_{i2} = -s_i \cdot X_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)$.
+This forms a smooth downward-opening parabola in the plane defined by $(X_{i1}, X_{i2})$, with the degree of curvature controlled by $s_i$.
+For $p > 2$, all remaining dimensions are independent Gaussian noise: $X_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}$.
 

### 2025-06-24 paper-cardinalR.tex
Updated branching last two sections

@@
 
-Branch \(1\) is generated from \(X_1 \sim U(0, 1)\), and \(X_2 = 0.1 \cdot X_1 + 1 \cdot X_1^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
+Branch \(1\) is generated from \(X_1 \sim U(0, 1)\), and \(X_2 = 0.1X_1 + X_1^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
 This produces a gently upward-curving parabola in the right half-plane.
-Branch \(2\) is generated from \(X_{1} \sim U(-1, 0)\), and \(X_2 = 0.1 \cdot X_{1} - 2 \cdot X_{1}^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
+Branch \(2\) is generated from \(X_{1} \sim U(-1, 0)\), and \(X_2 = 0.1X_{1} - 2X_{1}^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
 This creates a steeper, leftward-facing curve in the left half-plane.
@@
 
-The \texttt{gen\_orglinearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches embedded in a \(p\)-dimensional space.
+The \texttt{gen\_orglinearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches embedded in \(p\text{-}D\).
 Each branch lies primarily within a distinct \(2\text{-}D\) subspace, while the remaining \(p - 2\) dimensions contain Gaussian noise.
@@
 If \(k > \binom{p}{2}\), additional pairs are sampled with replacement to reach \(k\) total branches.
-Each selected pair \((x_{i1}, x_{i2})\) defines the \(2\text{-}D\) plane for branch \(i\).
-
-For branch \(i\), \(n_i\) points are generated, where \(\sum_{i=1}^k n_i = n\). The structure follows:
-
-\[
-x_{i1} \sim U(0, 2), \quad
-x_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5),
-\]
-
-where \(s_i\) is a scale factor controlling the slope.
+Each selected pair \((X_{i1}, X_{i2})\) defines the \(2\text{-}D\) plane for branch \(i\).
+
+For branch \(i\), \(n_i\) points are generated, where \(\sum_{i=1}^k n_i = n\).
+The structure follows: \(X_{i1} \sim U(0, 2)\), and \(X_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5)\), where \(s_i\) is a scale factor controlling the slope.
 When \(k \leq \binom{p}{2}\), \(s_i = 1\).
 When sampling with replacement, \(s_i\) is drawn from the set \({1, 1.5, 2, \dots, 8}\) in increments of \(0.5\).
-
-For \(p > 2\), the remaining dimensions contain independent Gaussian noise:
-
-\[
-x_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
-\]
+For \(p > 2\), the remaining dimensions contain independent Gaussian noise: \(X_j \sim N(0, 0.1^2)\), for \(j \notin \{i1, i2\}\).
 
@@
 Let \(\binom{p}{2}\) denote the number of unique \(2\text{-}D\) subspace combinations.
-When \(k \leq \binom{p}{2}\), \(k\) distinct subspace pairs \((x_{i1}, x_{i2})\) are sampled without replacement.
+When \(k \leq \binom{p}{2}\), \(k\) distinct subspace pairs \((X_{i1}, X_{i2})\) are sampled without replacement.
 Otherwise, combinations are selected with replacement to reach \(k\) total branches.
-For each branch \(i = 1, \dots, k\), a random scale factor \(s_i\) is used to vary the curvature:
-
-\begin{itemize}
-\tightlist
-\item
-  If \(k \leq \binom{p}{2}\): \(s_i = 1\)
-\item
-  If \(k > \binom{p}{2}\): \(s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}\)
-\end{itemize}
+For each branch \(i = 1, \dots, k\), a random scale factor \(s_i\) is used to vary the curvature; If \(k \leq \binom{p}{2}\): \(s_i = 1\), or If \(k > \binom{p}{2}\): \(s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}\).
 
 Each branch contains \(n_i\) points such that \(\sum_{i=1}^k n_i = n\), where the vector \((n_1, \dots, n_k)\) is randomly drawn using the helper function \texttt{gen\_nsum()} to partition \(n\).
-
-Within its assigned subspace \((x_{i1}, x_{i2})\), branch \(i\) is defined by:
-
-\[
-x_{i1} \sim U(0, 2), \quad
-x_{i2} = -s_i \cdot x_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)
-\]
-
-This forms a smooth downward-opening parabola in the plane defined by \((x_{i1}, x_{i2})\), with the degree of curvature controlled by \(s_i\).
-
-For \(p > 2\), all remaining dimensions are independent Gaussian noise:
-
-\[
-x_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}.
-\]
+Within its assigned subspace \((X_{i1}, X_{i2})\), branch \(i\) is defined by: \(X_{i1} \sim U(0, 2)\), and \(X_{i2} = -s_i \cdot X_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)\).
+This forms a smooth downward-opening parabola in the plane defined by \((X_{i1}, X_{i2})\), with the degree of curvature controlled by \(s_i\).
+For \(p > 2\), all remaining dimensions are independent Gaussian noise: \(X_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}\).
 

### 2025-06-24 paper-cardinalR.Rmd
Updated \frac

@@
 Branches $3$ to $k$ are added iteratively.
-Each new branch $i$ begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X_{1}$:$X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure:
-
-$$
-X_2 = 0.1 \cdot X_1 - s_i \cdot (X_1^2 - x_{\text{start}}) + y_{\text{start}}, \quad s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}.
-$$
-
+Each new branch $i$ begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X_{1}$:$X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure: $X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}$.
 Here, $s_i$ is a sampled scale factor (slope-like term) controlling curvature, and $(x_{\text{start}}, y_{\text{start}})$ is the starting point from the existing structure.
-
 This construction yields $k$ spatially connected, nonlinear branches with varying curvature.
-
 For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
@@
 For branch $i$, $n_i$ points are generated, where $\sum_{i=1}^k n_i = n$.
-The structure follows: $X_{i1} \sim U(0, 2)$, and $X_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5)$, where $s_i$ is a scale factor controlling the slope.
+The structure follows: $X_{i1} \sim U(0, 2)$, and $X_{i2} = -s_iX_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5)$, where $s_i$ is a scale factor controlling the slope.
 When $k \leq \binom{p}{2}$, $s_i = 1$.
@@
 Each branch contains $n_i$ points such that $\sum_{i=1}^k n_i = n$, where the vector $(n_1, \dots, n_k)$ is randomly drawn using the helper function `gen_nsum()` to partition $n$.
-Within its assigned subspace $(X_{i1}, X_{i2})$, branch $i$ is defined by: $X_{i1} \sim U(0, 2)$, and $X_{i2} = -s_i \cdot X_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)$.
+Within its assigned subspace $(X_{i1}, X_{i2})$, branch $i$ is defined by: $X_{i1} \sim U(0, 2)$, and $X_{i2} = -s_iX_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)$.
 This forms a smooth downward-opening parabola in the plane defined by $(X_{i1}, X_{i2})$, with the degree of curvature controlled by $s_i$.
@@
 The Mobius strip is generated using `geozoo::mobius(n, p = 3)$points`, which samples $n$ points from a parametrization of the surface defined by angle $\theta$ and strip width $w$.
-The cartesian coordinates are $X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = \frac{w}{2} \sin(\theta/2)$.
+The cartesian coordinates are $X_1 = \left(1 + w/2 \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + w/2 \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = w/2 \sin(\theta/2)$.
 This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface.
@@
 The final dimension $X_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$: $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$.
-Let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$: $r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}$, $r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}$.
+Let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$: $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
 The first three coordinates are then defined as: $X_1 \sim U(-r_x(z),\ r_x(z))$, $X_2 \sim U(-r_y(z),\ r_y(z))$, and $X_3 \sim U(-r_x(z),\ r_x(z))$.
-For $p > 3$, the remaining $p - 4$ dimensions (i.e., $X_4$ to $X_{p-1}$) are tapered toward zero with decreasing height and generated as: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1$.
+For $p > 3$, the remaining $p - 4$ dimensions (i.e., $X_4$ to $X_{p-1}$) are tapered toward zero with decreasing height and generated as: $X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1$.
 This tapering ensures that the structure narrows in all directions as it approaches the tip.
@@
 The final coordinate $X_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$: $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$.
-Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$: $r(z) = r_t + (l - r_t) \cdot \frac{z}{h}$.
+Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$: $r(z) = r_t + (l - r_t)z/h$.
 
 A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
-The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z) \cdot u$, and $X_3 = r(z) \cdot v$.
-For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right)$.
+The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v$.
+For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip: $X_j \sim U(-0.1, 0.1)\left(1 - z/h\right)$.
 
@@
 
-The radius at height $z$ scales linearly from zero (tip) to the base radius $r_b$: $r(z) = r_b \cdot \left(1 - \frac{z}{h}\right)$.
-
-Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
+The radius at height $z$ scales linearly from zero (tip) to the base radius $r_b$: $r(z) = r_b\left(1 - z/h\right)$.
+
+Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
 $r_{\text{point}} \sim \sqrt{U(0, 1)}$.
-Then, the first two coordinates are: $X_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta)$, and $X_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)$.
-For dimensions $j = 3, \dots, p-1$, values are tapered toward zero near the tip: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right)$.
+Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
+For dimensions $j = 3, \dots, p-1$, values are tapered toward zero near the tip: $X_j \sim U(-0.1, 0.1)\left(1 - z/h \right)$.
 
@@
 Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
-At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
+At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
 This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
@@
 
-The $3\text{-}D$ geometry is constructed by introducing a latent parameter, $\theta \sim U\left(-\frac{3\pi}{2}, \frac{3\pi}{2}\right)$.
+The $3\text{-}D$ geometry is constructed by introducing a latent parameter, $\theta \sim U\left(-3\pi/2, 3\pi/2\right)$.
 This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure: $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$, $X_3 = \text{sign}(\theta) \cdot (\cos(\theta) - 1)$.
+The first three dimensions form the S-curve structure: $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$, $X_3 = \text{sign}(\theta)(\cos(\theta) - 1)$.
 This configuration creates a horizontally curled shape in $(X_1, X_3)$, with additional band thickness in the $X_2$ direction.
@@
 The first component is a scaling factor that decreases with the dimension index, defined as $\text{scale}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$.
-The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2) \cdot \frac{\pi}{2p}$.
-Each additional dimension is computed as: $X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p$.
+The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2)\pi/2p$.
+Each additional dimension is computed as: $X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p$.
 
@@
 The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle." In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin.
-Additionally, $X_3 = \frac{1}{3} \cdot \cos(3\theta)$ introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
+Additionally, $X_3 = 1/3\cos(3\theta)$ introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
 
@@
 The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from $4$ to $p$, which means that the amplitude decreases as the dimension increases.
-Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j \cdot \sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2) \cdot \frac{\pi}{2p}$.
+Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j\sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2)\pi/2p$.
 
@@
 Each observation lies on the surface of a sphere in $3\text{-}D$, constructed by generating $u \sim U(-1, 1)$ which represents the cosine of the polar angle $\phi$ and $\theta \sim U(0, 2\pi)$ which represents the azimuthal angle.
-The corresponding Cartesian coordinates are calculated as; $X_1 = r \cdot \sqrt{1 - u^2} \cdot \cos(\theta)$, $X_2 = r \cdot \sqrt{1 - u^2} \cdot \sin(\theta)$, $X_3 = r \cdot u$, which gives points uniformly distributed on the surface of a $3\text{-}D$ sphere (not within).
+The corresponding Cartesian coordinates are calculated as; $X_1 = r\sqrt{1 - u^2}\cos(\theta)$, $X_2 = r\sqrt{1 - u^2}\sin(\theta)$, $X_3 = ru$, which gives points uniformly distributed on the surface of a $3\text{-}D$ sphere (not within).
 For $p > 3$, additional dimensions $X_4$ to $X_p$ are generated as low-variance Gaussian noise: $X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p$.
@@
 
-Each point on the sphere is computed using the spherical-to-Cartesian transformation: $X_1 = \sin(\phi) \cdot \cos(\theta)$, $X_2 = \sin(\phi) \cdot \sin(\theta)$, $X_3 = \cos(\phi)$.
+Each point on the sphere is computed using the spherical-to-Cartesian transformation: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi)$.
 For $p > 3$, the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions: $X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p$.
@@
 Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into $4\text{-}D$ Cartesian space: $X_1 = \sin(\theta_1) \cdot \cos(\theta_2)$, $X_2 = \sin(\theta_1) \cdot \sin(\theta_2)$, $X_3 = \cos(\theta_1) \cdot \cos(\theta_3)$, $X_4 = \cos(\theta_1) \cdot \sin(\theta_3)$.
+The coordinates are transformed into $4\text{-}D$ Cartesian space: $X_1 = \sin(\theta_1)\cos(\theta_2)$, $X_2 = \sin(\theta_1)\sin(\theta_2)$, $X_3 = \cos(\theta_1)\cos(\theta_3)$, $X_4 = \cos(\theta_1)\sin(\theta_3)$.
 This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
@@
 The corresponding $2\text{-}D$ coordinates are defined by: $X_1 = \cos(\theta)$, $X_2 = \sin(\theta)$.
-For $p > 2$, additional dimensions $X_3$ to $X_p$ are generated using `gen_wavydims1()`: $X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}$.
+For $p > 2$, additional dimensions $X_3$ to $X_p$ are generated using `gen_wavydims1()`: $X_j = s_j\theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}$.
 These noise dimensions are nonlinearly correlated with $\theta$, producing structured, wave-like variation that aligns with the progression along the crescent curve.
@@
 
-Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi) \cdot \cos(\theta)$, $X_2 = \sin(\phi) \cdot \sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
+Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
 This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
@@
 These dimensions are functions of the first coordinate $X_1$, introducing dependencies that preserve some geometric coherence.
-Each added dimension follows the form: $X_j = s_j \cdot (-1)^{\lfloor j/2 \rfloor} \cdot X_1^{a_j} + \eta_j$, where $a_j \in \{2, 3, 4, 5\}$ is a randomly chosen polynomial power, $s_j \sim U(0.5, 2)$ is a scale factor, $\eta_j \sim U(-\sigma_j, 2\sigma_j)$, with $\sigma_j \sim U(0, 0.05)$, adds mild randomness.
+Each added dimension follows the form: $X_j = s_j(-1)^{\lfloor j/2 \rfloor}X_1^{a_j} + \eta_j$, where $a_j \in \{2, 3, 4, 5\}$ is a randomly chosen polynomial power, $s_j \sim U(0.5, 2)$ is a scale factor, $\eta_j \sim U(-\sigma_j, 2\sigma_j)$, with $\sigma_j \sim U(0, 0.05)$, adds mild randomness.
 
@@
 The first four coordinates follow a $3\text{-}D$ helix with an additional oscillating component.
-Let $\theta \in [0, \frac{5\pi}{4}]$ be a sequence of angles controlling rotation around a circle.
-Cartesian coordinates; $X_1 = \cos(\theta)$: circular trajectory along the x-axis, $X_2 = \sin(\theta)$: circular trajectory along the y-axis, $X_3 = 0.05 \cdot \theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix, and $X_4 = 0.1 \cdot \sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
+Let $\theta \in [0, 5\pi/4]$ be a sequence of angles controlling rotation around a circle.
+Cartesian coordinates; $X_1 = \cos(\theta)$: circular trajectory along the x-axis, $X_2 = \sin(\theta)$: circular trajectory along the y-axis, $X_3 = 0.05\theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix, and $X_4 = 0.1\sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
 This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
@@
 
-The shape is defined by parameter $\theta \in [0, 2\pi \cdot \text{spins}]$, controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta \cdot \cos(\theta)$ for radial expansion in x, and $X_2 = \theta \cdot \sin(\theta)$ for radial expansion in y.
-The growth pattern resembles a cone, with the height increasing according to $X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
+The shape is defined by parameter $\theta \in [0, 2\pi\text{spins}]$, controlling the angular progression of the spiral.
+The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta\cos(\theta)$ for radial expansion in x, and $X_2 = \theta\sin(\theta)$ for radial expansion in y.
+The growth pattern resembles a cone, with the height increasing according to $X_3 = 2\theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta\sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
 This results in a $3\text{-}D$ spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
@@
 
-The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = \frac{hc}{X_{1}} + \text{nonfac} \cdot \sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.

### 2025-06-24 paper-cardinalR.tex
Updated \frac

@@
 Branches \(3\) to \(k\) are added iteratively.
-Each new branch \(i\) begins at a randomly chosen point within a restricted horizontal band: \(X_{1} \in [-0.15, 0.15]\) (to ensure connectivity with earlier branches), spans a unit-length interval on \(X_{1}\):\(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), and has the structure:
-
-\[
-X_2 = 0.1 \cdot X_1 - s_i \cdot (X_1^2 - x_{\text{start}}) + y_{\text{start}}, \quad s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}.
-\]
-
+Each new branch \(i\) begins at a randomly chosen point within a restricted horizontal band: \(X_{1} \in [-0.15, 0.15]\) (to ensure connectivity with earlier branches), spans a unit-length interval on \(X_{1}\):\(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), and has the structure: \(X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}\).
 Here, \(s_i\) is a sampled scale factor (slope-like term) controlling curvature, and \((x_{\text{start}}, y_{\text{start}})\) is the starting point from the existing structure.
-
 This construction yields \(k\) spatially connected, nonlinear branches with varying curvature.
-
 For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
@@
 For branch \(i\), \(n_i\) points are generated, where \(\sum_{i=1}^k n_i = n\).
-The structure follows: \(X_{i1} \sim U(0, 2)\), and \(X_{i2} = -s_i \cdot x_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5)\), where \(s_i\) is a scale factor controlling the slope.
+The structure follows: \(X_{i1} \sim U(0, 2)\), and \(X_{i2} = -s_iX_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5)\), where \(s_i\) is a scale factor controlling the slope.
 When \(k \leq \binom{p}{2}\), \(s_i = 1\).
@@
 Each branch contains \(n_i\) points such that \(\sum_{i=1}^k n_i = n\), where the vector \((n_1, \dots, n_k)\) is randomly drawn using the helper function \texttt{gen\_nsum()} to partition \(n\).
-Within its assigned subspace \((X_{i1}, X_{i2})\), branch \(i\) is defined by: \(X_{i1} \sim U(0, 2)\), and \(X_{i2} = -s_i \cdot X_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)\).
+Within its assigned subspace \((X_{i1}, X_{i2})\), branch \(i\) is defined by: \(X_{i1} \sim U(0, 2)\), and \(X_{i2} = -s_iX_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)\).
 This forms a smooth downward-opening parabola in the plane defined by \((X_{i1}, X_{i2})\), with the degree of curvature controlled by \(s_i\).
@@
 The Mobius strip is generated using \texttt{geozoo::mobius(n,\ p\ =\ 3)\$points}, which samples \(n\) points from a parametrization of the surface defined by angle \(\theta\) and strip width \(w\).
-The cartesian coordinates are \(X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta)\), \(X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta)\), and \(X_3 = \frac{w}{2} \sin(\theta/2)\).
+The cartesian coordinates are \(X_1 = \left(1 + w/2 \cos(\theta/2)\right) \cos(\theta)\), \(X_2 = \left(1 + w/2 \cos(\theta/2)\right) \sin(\theta)\), and \(X_3 = w/2 \sin(\theta/2)\).
 This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable one-sided surface.
@@
 The final dimension \(X_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\): \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\).
-Let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\): \(r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}\), \(r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}\).
+Let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\): \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
 The first three coordinates are then defined as: \(X_1 \sim U(-r_x(z),\ r_x(z))\), \(X_2 \sim U(-r_y(z),\ r_y(z))\), and \(X_3 \sim U(-r_x(z),\ r_x(z))\).
-For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are tapered toward zero with decreasing height and generated as: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1\).
+For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are tapered toward zero with decreasing height and generated as: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1\).
 This tapering ensures that the structure narrows in all directions as it approaches the tip.
@@
 The final coordinate \(X_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\): \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\).
-Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\): \(r(z) = r_t + (l - r_t) \cdot \frac{z}{h}\).
+Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\): \(r(z) = r_t + (l - r_t)z/h\).
 
 A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section: \(u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v\).
-The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z) \cdot u\), and \(X_3 = r(z) \cdot v\).
-For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right)\).
+The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z)u\), and \(X_3 = r(z)v\).
+For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h\right)\).
 
@@
 
-The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r_b\): \(r(z) = r_b \cdot \left(1 - \frac{z}{h}\right)\).
-
-Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
+The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r_b\): \(r(z) = r_b\left(1 - z/h\right)\).
+
+Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
 \(r_{\text{point}} \sim \sqrt{U(0, 1)}\).
-Then, the first two coordinates are: \(X_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta)\), and \(X_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)\).
-For dimensions \(j = 3, \dots, p-1\), values are tapered toward zero near the tip: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right)\).
+Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
+For dimensions \(j = 3, \dots, p-1\), values are tapered toward zero near the tip: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h \right)\).
 
@@
 Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\text{-}D\) simplex.
-At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\): \(T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}\).
+At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\): \(T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}\).
 This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
@@
 
-The \(3\text{-}D\) geometry is constructed by introducing a latent parameter, \(\theta \sim U\left(-\frac{3\pi}{2}, \frac{3\pi}{2}\right)\).
+The \(3\text{-}D\) geometry is constructed by introducing a latent parameter, \(\theta \sim U\left(-3\pi/2, 3\pi/2\right)\).
 This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure: \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\), \(X_3 = \text{sign}(\theta) \cdot (\cos(\theta) - 1)\).
+The first three dimensions form the S-curve structure: \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\), \(X_3 = \text{sign}(\theta)(\cos(\theta) - 1)\).
 This configuration creates a horizontally curled shape in \((X_1, X_3)\), with additional band thickness in the \(X_2\) direction.
@@
 The first component is a scaling factor that decreases with the dimension index, defined as \(\text{scale}_j = \sqrt{(0.5)^{j-2}}\) for \(j = 3, \dots, p\).
-The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula \(\phi_j = (j - 2) \cdot \frac{\pi}{2p}\).
-Each additional dimension is computed as: \(X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p\).
+The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula \(\phi_j = (j - 2)\pi/2p\).
+Each additional dimension is computed as: \(X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p\).
 
@@
 The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle.'' In this configuration, \(X_1 = \cos(\theta)\) represents horizontal oscillation, while \(X_2 = \sqrt{3}/3 + \sin(\theta)\) introduces a vertical offset to avoid centering the curve at the origin.
-Additionally, \(X_3 = \frac{1}{3} \cdot \cos(3\theta)\) introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
+Additionally, \(X_3 = 1/3\cos(3\theta)\) introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
 
@@
 The scaling factor is defined as \(\text{scale}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from \(4\) to \(p\), which means that the amplitude decreases as the dimension increases.
-Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j \cdot \sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2) \cdot \frac{\pi}{2p}\).
+Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j\sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2)\pi/2p\).
 
@@
 Each observation lies on the surface of a sphere in \(3\text{-}D\), constructed by generating \(u \sim U(-1, 1)\) which represents the cosine of the polar angle \(\phi\) and \(\theta \sim U(0, 2\pi)\) which represents the azimuthal angle.
-The corresponding Cartesian coordinates are calculated as; \(X_1 = r \cdot \sqrt{1 - u^2} \cdot \cos(\theta)\), \(X_2 = r \cdot \sqrt{1 - u^2} \cdot \sin(\theta)\), \(X_3 = r \cdot u\), which gives points uniformly distributed on the surface of a \(3\text{-}D\) sphere (not within).
+The corresponding Cartesian coordinates are calculated as; \(X_1 = r\sqrt{1 - u^2}\cos(\theta)\), \(X_2 = r\sqrt{1 - u^2}\sin(\theta)\), \(X_3 = ru\), which gives points uniformly distributed on the surface of a \(3\text{-}D\) sphere (not within).
 For \(p > 3\), additional dimensions \(X_4\) to \(X_p\) are generated as low-variance Gaussian noise: \(X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p\).
@@
 
-Each point on the sphere is computed using the spherical-to-Cartesian transformation: \(X_1 = \sin(\phi) \cdot \cos(\theta)\), \(X_2 = \sin(\phi) \cdot \sin(\theta)\), \(X_3 = \cos(\phi)\).
+Each point on the sphere is computed using the spherical-to-Cartesian transformation: \(X_1 = \sin(\phi)\cos(\theta)\), \(X_2 = \sin(\phi)\sin(\theta)\), \(X_3 = \cos(\phi)\).
 For \(p > 3\), the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions: \(X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p\).
@@
 Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into \(4\text{-}D\) Cartesian space: \(X_1 = \sin(\theta_1) \cdot \cos(\theta_2)\), \(X_2 = \sin(\theta_1) \cdot \sin(\theta_2)\), \(X_3 = \cos(\theta_1) \cdot \cos(\theta_3)\), \(X_4 = \cos(\theta_1) \cdot \sin(\theta_3)\).
+The coordinates are transformed into \(4\text{-}D\) Cartesian space: \(X_1 = \sin(\theta_1)\cos(\theta_2)\), \(X_2 = \sin(\theta_1)\sin(\theta_2)\), \(X_3 = \cos(\theta_1)\cos(\theta_3)\), \(X_4 = \cos(\theta_1)\sin(\theta_3)\).
 This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
@@
 The corresponding \(2\text{-}D\) coordinates are defined by: \(X_1 = \cos(\theta)\), \(X_2 = \sin(\theta)\).
-For \(p > 2\), additional dimensions \(X_3\) to \(X_p\) are generated using \texttt{gen\_wavydims1()}: \(X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}\).
+For \(p > 2\), additional dimensions \(X_3\) to \(X_p\) are generated using \texttt{gen\_wavydims1()}: \(X_j = s_j\theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}\).
 These noise dimensions are nonlinearly correlated with \(\theta\), producing structured, wave-like variation that aligns with the progression along the crescent curve.
@@
 
-Cartesian coordinates from spherical conversion: \(X_1 = \sin(\phi) \cdot \cos(\theta)\), \(X_2 = \sin(\phi) \cdot \sin(\theta)\), \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter, and \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
+Cartesian coordinates from spherical conversion: \(X_1 = \sin(\phi)\cos(\theta)\), \(X_2 = \sin(\phi)\sin(\theta)\), \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter, and \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
 This generates a spherical spiral curve embedded in \(4\text{-}D\) space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
@@
 These dimensions are functions of the first coordinate \(X_1\), introducing dependencies that preserve some geometric coherence.
-Each added dimension follows the form: \(X_j = s_j \cdot (-1)^{\lfloor j/2 \rfloor} \cdot X_1^{a_j} + \eta_j\), where \(a_j \in \{2, 3, 4, 5\}\) is a randomly chosen polynomial power, \(s_j \sim U(0.5, 2)\) is a scale factor, \(\eta_j \sim U(-\sigma_j, 2\sigma_j)\), with \(\sigma_j \sim U(0, 0.05)\), adds mild randomness.
+Each added dimension follows the form: \(X_j = s_j(-1)^{\lfloor j/2 \rfloor}X_1^{a_j} + \eta_j\), where \(a_j \in \{2, 3, 4, 5\}\) is a randomly chosen polynomial power, \(s_j \sim U(0.5, 2)\) is a scale factor, \(\eta_j \sim U(-\sigma_j, 2\sigma_j)\), with \(\sigma_j \sim U(0, 0.05)\), adds mild randomness.
 
@@
 The first four coordinates follow a \(3\text{-}D\) helix with an additional oscillating component.
-Let \(\theta \in [0, \frac{5\pi}{4}]\) be a sequence of angles controlling rotation around a circle.
-Cartesian coordinates; \(X_1 = \cos(\theta)\): circular trajectory along the x-axis, \(X_2 = \sin(\theta)\): circular trajectory along the y-axis, \(X_3 = 0.05 \cdot \theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix, and \(X_4 = 0.1 \cdot \sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
+Let \(\theta \in [0, 5\pi/4]\) be a sequence of angles controlling rotation around a circle.
+Cartesian coordinates; \(X_1 = \cos(\theta)\): circular trajectory along the x-axis, \(X_2 = \sin(\theta)\): circular trajectory along the y-axis, \(X_3 = 0.05\theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix, and \(X_4 = 0.1\sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
 This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
@@
 
-The shape is defined by parameter \(\theta \in [0, 2\pi \cdot \text{spins}]\), controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; \(X_1 = \theta \cdot \cos(\theta)\) for radial expansion in x, and \(X_2 = \theta \cdot \sin(\theta)\) for radial expansion in y.
-The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2 \cdot \theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\) Spiral modulation in the fourth dimension is represented by \(X_4 = \theta \cdot \sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
+The shape is defined by parameter \(\theta \in [0, 2\pi\text{spins}]\), controlling the angular progression of the spiral.
+The Archimedean spiral in the horizontal plane is represented by; \(X_1 = \theta\cos(\theta)\) for radial expansion in x, and \(X_2 = \theta\sin(\theta)\) for radial expansion in y.
+The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2\theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\) Spiral modulation in the fourth dimension is represented by \(X_4 = \theta\sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
 This results in a \(3\text{-}D\) spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
@@
 
-The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = \frac{hc}{X_{1}} + \text{nonfac} \cdot \sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi \cdot X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.

### 2025-06-24 paper-cardinalR.Rmd
Updated the equations

@@
      legend.key.height = unit(0.25, 'cm'),
-     legend.key.width = unit(0.25, 'cm')
+     legend.key.width = unit(0.25, 'cm'),
+     plot.margin = margin(0, 0, 0, 0)
    )
@@
 
-```{r, label = "fig-quadratic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-quadratic_proj1 + quadratic_proj2 + quadratic_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_cubic()`
@@
 
-```{r, label = "fig-cubic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cubic` data."}
+```{r, label = "fig-cubic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` and `cubic` data."}
 #| fig-width: 15
-#| fig-height: 5
-
+#| fig-height: 10
+
+quadratic_proj1 + quadratic_proj2 + quadratic_proj3 +
 cubic_proj1 + cubic_proj2 + cubic_proj3 +
@@
 
-```{r, label = "fig-scurve-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-scurve_proj1 + scurve_proj2 + scurve_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_scurvehole()`
@@
 
-```{r, label = "fig-scurvehole-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurvehole` data."}
+```{r, label = "fig-scurve-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` and `scurvehole` data."}
 #| fig-width: 15
-#| fig-height: 5
-
+#| fig-height: 10
+
+scurve_proj1 + scurve_proj2 + scurve_proj3 + 
   scurvehole_proj1 + scurvehole_proj2 + scurvehole_proj3 +

### 2025-06-24 paper-cardinalR.tex
Updated the equations

@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-quadratic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` data.}\label{fig:fig-quadratic-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_cubic()}}{gen\_cubic()}}\label{gen_cubic}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cubic` data.}\label{fig:fig-cubic-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` and `cubic` data.}\label{fig:fig-cubic-proj}
 \end{figure}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` data.}\label{fig:fig-scurve-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_scurvehole()}}{gen\_scurvehole()}}\label{gen_scurvehole}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurvehole-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurvehole` data.}\label{fig:fig-scurvehole-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` and `scurvehole` data.}\label{fig:fig-scurve-proj}
 \end{figure}

### 2025-06-24 paper-cardinalR.Rmd
Put code and acknowlegement together

@@
    theme(
-     #aspect.ratio = 1,
+     aspect.ratio = 1,
      plot.background = element_rect(fill = 'transparent', colour = NA),
@@
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.3, 
+                                              axis_pos_y = -0.3, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.45, 0.73), 
   title = "a1", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
+                            axis_param = list(limits = 0.45,
+                                              axis_scaled = 1.3, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
                                               threshold = 0.01))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.3, 0.43), 
   title = "a2", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.5,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.35, 0.4), 
   title = "a3", 
@@
 
-# Code
-
-The code is available at [https://github.com/JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR), and source material for this paper is available at [https://github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
-
 # Acknowledgements
+
+The package is available at [https://github.com/JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR), and source material for this paper is available at [https://github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
 

### 2025-06-24 paper-cardinalR.tex
Put code and acknowlegement together

@@
 
-\section{Code}\label{code}
-
-The code is available at \url{https://github.com/JayaniLakshika/cardinalR}, and source material for this paper is available at \url{https://github.com/JayaniLakshika/paper-cardinalR}.
-
 \section{Acknowledgements}\label{acknowledgements}
+
+The package is available at \url{https://github.com/JayaniLakshika/cardinalR}, and source material for this paper is available at \url{https://github.com/JayaniLakshika/paper-cardinalR}.
 

### 2025-06-24 paper-cardinalR.Rmd
Try to reduce the number of pages

@@
 
-```{r, label = "fig-expbranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-expbranches_proj1 + expbranches_proj2 + expbranches_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_linearbranches()`
@@
 
-```{r, label = "fig-linearbranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linearbranches` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-linearbranches_proj1 + linearbranches_proj2 + linearbranches_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_curvybranches()`
@@
 
-```{r, label = "fig-curvybranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvybranches` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-curvybranches_proj1 + curvybranches_proj2 + curvybranches_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_orglinearbranches()`
@@
 
-```{r, label = "fig-orglinearbranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_orgcurvybranches()`
@@
 
-```{r, label = "fig-orgcurvybranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orgcurvybranches` data."}
-#| fig-width: 15
-#| fig-height: 5
-
+```{r, label = "fig-orgcurvybranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, `orgcurvybranches` data.", fig.width=15, fig.height=25}
+
+expbranches_proj1 + expbranches_proj2 + expbranches_proj3 +
+linearbranches_proj1 + linearbranches_proj2 + linearbranches_proj3 +
+curvybranches_proj1 + curvybranches_proj2 + curvybranches_proj3 +
+orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
 orgcurvybranches_proj1 + orgcurvybranches_proj2 + orgcurvybranches_proj3 +
@@
 
-```{r, label = "fig-gridcube-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-gridcube_proj1 + gridcube_proj2 + gridcube_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_unifcube()`
@@
 
-```{r, label = "fig-unifcube-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `unifcube` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-unifcube_proj1 + unifcube_proj2 + unifcube_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_cubehole()`
@@
 
-```{r, label = "fig-cubehole-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cubehole` data."}
-#| fig-width: 15
-#| fig-height: 5
-
+```{r, label = "fig-cubehole-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube`, `unifcube`, and `cubehole` data.", fig.width=15, fig.height=15}
+
+gridcube_proj1 + gridcube_proj2 + gridcube_proj3 +
+unifcube_proj1 + unifcube_proj2 + unifcube_proj3 +
 cubehole_proj1 + cubehole_proj2 + cubehole_proj3 +
@@
 
-```{r, label = "fig-cubic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` and `cubic` data."}
-#| fig-width: 15
-#| fig-height: 10
+```{r, label = "fig-cubic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` and `cubic` data.", fig.width=15, fig.height=10}
 
@@
 
-```{r, label = "fig-pyrrect-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-pyrrect_proj1 + pyrrect_proj2 + pyrrect_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_pyrtri()`
@@
 
-```{r, label = "fig-pyrtri-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrtri` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_pyrstar()`
@@
 
-```{r, label = "fig-pyrstar-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrstar` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-pyrstar_proj1 + pyrstar_proj2 + pyrstar_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_pyrholes()`
@@
 
-```{r, label = "fig-pyrholes-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrholes` data."}
-#| fig-width: 15
-#| fig-height: 5
-
+```{r, label = "fig-pyrholes-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data.", fig.width=15, fig.height=20}
+
+pyrrect_proj1 + pyrrect_proj2 + pyrrect_proj3 +
+pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
+pyrstar_proj1 + pyrstar_proj2 + pyrstar_proj3 +
 pyrholes_proj1 + pyrholes_proj2 + pyrholes_proj3 +
@@
 
-```{r, label = "fig-scurve-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` and `scurvehole` data."}
-#| fig-width: 15
-#| fig-height: 10
+```{r, label = "fig-scurve-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` and `scurvehole` data.", fig.width=15, fig.height=10}
 
@@
 
-```{r, label = "fig-circle-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-circle_proj1 + circle_proj2 + circle_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_curvycycle()`
@@
 
-```{r, label = "fig-curvycycle-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycycle` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_unifsphere()`
@@
 
-```{r, label = "fig-unifsphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `unifsphere` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-unifsphere_proj1 + unifsphere_proj2 + unifsphere_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_gridedsphere()`
@@
 
-```{r, label = "fig-gridedsphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridedsphere` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-gridedsphere_proj1 + gridedsphere_proj2 + gridedsphere_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_clusteredspheres()`
@@
 
-```
-
-```{r, label = "fig-clusteredspheres-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `clusteredspheres` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-clusteredspheres_proj1 + clusteredspheres_proj2 + clusteredspheres_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
 ```
@@
 
-```{r, label = "fig-hemisphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `hemisphere` data."}
-#| fig-width: 15
-#| fig-height: 5
-
+```{r, label = "fig-hemisphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle`, `curvycycle`, `unifsphere`, `gridedsphere`, `clusteredspheres`, and `hemisphere` data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
+
+circle_proj1 + circle_proj2 + circle_proj3 +
+curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3 +
+unifsphere_proj1 + unifsphere_proj2 + unifsphere_proj3 +
+gridedsphere_proj1 + gridedsphere_proj2 + gridedsphere_proj3 +
+clusteredspheres_proj1 + clusteredspheres_proj2 + clusteredspheres_proj3 +
 hemisphere_proj1 + hemisphere_proj2 + hemisphere_proj3 +
@@
 
-```{r, label = "fig-crescent-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-crescent_proj1 + crescent_proj2 + crescent_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_curvycylinder()`
@@
 
-```{r, label = "fig-curvycylinder-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_sphericalspiral()`
@@
 
-```
-
-```{r, label = "fig-sphericalspiral-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `sphericalspiral` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
 ```
@@
 
-```{r, label = "fig-helicalspiral-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `helicalspiral` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-helicalspiral_proj1 + helicalspiral_proj2 + helicalspiral_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_conicspiral()`
@@
 
-```{r, label = "fig-conicspiral-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `conicspiral` data."}
-#| fig-width: 15
-#| fig-height: 5
-
-conicspiral_proj1 + conicspiral_proj2 + conicspiral_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### `gen_nonlinear()`
@@
 
-```{r, label = "fig-nonlinear-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `nonlinear` data."}
-#| fig-width: 15
-#| fig-height: 5
-
+```{r, label = "fig-triginometric-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
+
+crescent_proj1 + crescent_proj2 + crescent_proj3 +
+curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
+sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +
+helicalspiral_proj1 + helicalspiral_proj2 + helicalspiral_proj3 +
+conicspiral_proj1 + conicspiral_proj2 + conicspiral_proj3 +
 nonlinear_proj1 + nonlinear_proj2 + nonlinear_proj3 +

### 2025-06-24 paper-cardinalR.tex
Try to reduce the number of pages

@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-expbranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches` data.}\label{fig:fig-expbranches-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_linearbranches()}}{gen\_linearbranches()}}\label{gen_linearbranches}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-linearbranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linearbranches` data.}\label{fig:fig-linearbranches-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_curvybranches()}}{gen\_curvybranches()}}\label{gen_curvybranches}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-curvybranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvybranches` data.}\label{fig:fig-curvybranches-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_orglinearbranches()}}{gen\_orglinearbranches()}}\label{gen_orglinearbranches}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-orglinearbranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `orglinearbranches` data.}\label{fig:fig-orglinearbranches-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_orgcurvybranches()}}{gen\_orgcurvybranches()}}\label{gen_orgcurvybranches}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-orgcurvybranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `orgcurvybranches` data.}\label{fig:fig-orgcurvybranches-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-orgcurvybranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, `orgcurvybranches` data.}\label{fig:fig-orgcurvybranches-proj}
 \end{figure}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-gridcube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube` data.}\label{fig:fig-gridcube-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_unifcube()}}{gen\_unifcube()}}\label{gen_unifcube}
@@
 \end{verbatim}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-unifcube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `unifcube` data.}\label{fig:fig-unifcube-proj}
-\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubehole-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cubehole` data.}\label{fig:fig-cubehole-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubehole-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube`, `unifcube`, and `cubehole` data.}\label{fig:fig-cubehole-proj}
 \end{figure}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrrect-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` data.}\label{fig:fig-pyrrect-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_pyrtri()}}{gen\_pyrtri()}}\label{gen_pyrtri}
@@
 \end{verbatim}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrtri-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrtri` data.}\label{fig:fig-pyrtri-proj}
-\end{figure}
 
@@
 \end{verbatim}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrstar-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrstar` data.}\label{fig:fig-pyrstar-proj}
-\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrholes-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrholes` data.}\label{fig:fig-pyrholes-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrholes-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data.}\label{fig:fig-pyrholes-proj}
 \end{figure}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-circle-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` data.}\label{fig:fig-circle-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_curvycycle()}}{gen\_curvycycle()}}\label{gen_curvycycle}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-curvycycle-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycycle` data.}\label{fig:fig-curvycycle-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_unifsphere()}}{gen\_unifsphere()}}\label{gen_unifsphere}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-unifsphere-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `unifsphere` data.}\label{fig:fig-unifsphere-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_gridedsphere()}}{gen\_gridedsphere()}}\label{gen_gridedsphere}
@@
 \end{verbatim}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-gridedsphere-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridedsphere` data.}\label{fig:fig-gridedsphere-proj}
-\end{figure}
 
@@
 \end{verbatim}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-clusteredspheres-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `clusteredspheres` data.}\label{fig:fig-clusteredspheres-proj}
-\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-hemisphere-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `hemisphere` data.}\label{fig:fig-hemisphere-proj}
+
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/fig-hemisphere-proj-1}
+
+}
+
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle`, `curvycycle`, `unifsphere`, `gridedsphere`, `clusteredspheres`, and `hemisphere` data.}\label{fig:fig-hemisphere-proj}
 \end{figure}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-crescent-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` data.}\label{fig:fig-crescent-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_curvycylinder()}}{gen\_curvycylinder()}}\label{gen_curvycylinder}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-curvycylinder-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycylinder` data.}\label{fig:fig-curvycylinder-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_sphericalspiral()}}{gen\_sphericalspiral()}}\label{gen_sphericalspiral}
@@
 \end{verbatim}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-sphericalspiral-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `sphericalspiral` data.}\label{fig:fig-sphericalspiral-proj}
-\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-helicalspiral-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `helicalspiral` data.}\label{fig:fig-helicalspiral-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_conicspiral()}}{gen\_conicspiral()}}\label{gen_conicspiral}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-conicspiral-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `conicspiral` data.}\label{fig:fig-conicspiral-proj}
-\end{figure}
-
 \paragraph{\texorpdfstring{\texttt{gen\_nonlinear()}}{gen\_nonlinear()}}\label{gen_nonlinear}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nonlinear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `nonlinear` data.}\label{fig:fig-nonlinear-proj}
+
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/fig-triginometric-proj-1}
+
+}
+
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data.}\label{fig:fig-triginometric-proj}
 \end{figure}

### 2025-06-24 paper-cardinalR.Rmd
Set figure option

@@
   warning=FALSE,
-  out.width = "100%")
+  out.width = "100%",
+  fig.pos = "!")
 

### 2025-06-24 paper-cardinalR.Rmd
Updated acknowledgment only with source code

@@
 
-The package is available at [https://github.com/JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR), and source material for this paper is available at [https://github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
+The source material for this paper is available at [https://github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
 

### 2025-07-05 paper-cardinalR.Rmd
Added Trefoil initial description

@@
 
+### Trefoil Knot
+
+The **Trefoil Knot** is a closed, nontrivial 1-dimensional manifold embedded in 3D or 4D space.
+Unlike simple curves like the S-curve or Swiss roll, the trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of nonlinear dimensionality reduction (NLDR) methods to preserve global structure, loops, and embeddings in high-dimensional data.
+
+We generate the trefoil structure in **4D space** as the true model, and provide a **stereographically projected 3D version** for visualization purposes.
+Both versions support the addition of noise dimensions to embed the structure in higher-dimensional ambient spaces, simulating real-world settings.
+
+#### `gen_trefoil4d()`
+
+The function `gen_trefoil4d()` generates a 4-dimensional trefoil knot using two angular parameters, $\theta$ and $\phi$.
+The knot is constructed on the 3-sphere ($S^3 \subset \mathbb{R}^4$), with a band of thickness around the knot path, controlled by `steps`.
+
+The coordinates are given by:
+
+$$
+\begin{aligned}
+x_1 &= \cos(\theta) \cdot \cos(\phi), \\
+x_2 &= \cos(\theta) \cdot \sin(\phi), \\
+x_3 &= \sin(\theta) \cdot \cos(1.5 \cdot \phi), \\
+x_4 &= \sin(\theta) \cdot \sin(1.5 \cdot \phi).
+\end{aligned}
+$$
+
+The number of `theta` and `phi` values is determined by the `steps` and `n` arguments, respectively.
+Additional independent noise dimensions can be added via `gen_noisedims()` to extend the structure into $p > 4$ dimensions.
+
+```{r}
+trefoil4d <- gen_trefoil4d(n = 500, p = 4, steps = 5)
+```
+
+The resulting dataset represents a true model $T = (x_1, x_2, x_3, x_4)$, capturing the full 4D knot structure.
+
+#### `gen_trefoil3d()`
+
+The function `gen_trefoil3d()` generates a **3-dimensional stereographic projection** of the 4D trefoil knot, mapping each point from $\mathbb{R}^4$ to $\mathbb{R}^3$ using the transformation:
+
+$$
+(x_1', x_2', x_3') = \left( \frac{x_1}{1 - x_4}, \frac{x_2}{1 - x_4}, \frac{x_3}{1 - x_4} \right),
+$$
+
+which excludes points where $x_4 = 1$. The projection preserves the knot topology and enables easier visualization and interpretation.
+
+```{r}
+trefoil3d <- gen_trefoil3d(n = 500, p = 4, steps = 5)
+```
+
+As with the 4D version, this function supports optional noise dimensions to embed the projected knot into higher dimensions.
+
 ### Multiple cluster examples

### 2025-07-05 paper-cardinalR.tex
Set figure option
Updated acknowledgment only with source code
Added Trefoil initial description

@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-orgcurvybranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, `orgcurvybranches` data.}\label{fig:fig-orgcurvybranches-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:fig-cone-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubehole-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube`, `unifcube`, and `cubehole` data.}\label{fig:fig-cubehole-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:fig-gau-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-linear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linear` data.}\label{fig:fig-linear-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-mobius-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `mobius` data.}\label{fig:fig-mobius-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` and `cubic` data.}\label{fig:fig-cubic-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrholes-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data.}\label{fig:fig-pyrholes-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` and `scurvehole` data.}\label{fig:fig-scurve-proj}
@@
 
-\begin{figure}
+\begin{figure}[!]
 
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-swissroll-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `swissroll` data.}\label{fig:fig-swissroll-proj}
@@
 \end{figure}
+
+\subsubsection{Trefoil Knot}\label{trefoil-knot}
+
+The \textbf{Trefoil Knot} is a closed, nontrivial 1-dimensional manifold embedded in 3D or 4D space.
+Unlike simple curves like the S-curve or Swiss roll, the trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of nonlinear dimensionality reduction (NLDR) methods to preserve global structure, loops, and embeddings in high-dimensional data.
+
+We generate the trefoil structure in \textbf{4D space} as the true model, and provide a \textbf{stereographically projected 3D version} for visualization purposes.
+Both versions support the addition of noise dimensions to embed the structure in higher-dimensional ambient spaces, simulating real-world settings.
+
+\paragraph{\texorpdfstring{\texttt{gen\_trefoil4d()}}{gen\_trefoil4d()}}\label{gen_trefoil4d}
+
+The function \texttt{gen\_trefoil4d()} generates a 4-dimensional trefoil knot using two angular parameters, \(\theta\) and \(\phi\).
+The knot is constructed on the 3-sphere (\(S^3 \subset \mathbb{R}^4\)), with a band of thickness around the knot path, controlled by \texttt{steps}.
+
+The coordinates are given by:
+
+\[
+\begin{aligned}
+x_1 &= \cos(\theta) \cdot \cos(\phi), \\
+x_2 &= \cos(\theta) \cdot \sin(\phi), \\
+x_3 &= \sin(\theta) \cdot \cos(1.5 \cdot \phi), \\
+x_4 &= \sin(\theta) \cdot \sin(1.5 \cdot \phi).
+\end{aligned}
+\]
+
+The number of \texttt{theta} and \texttt{phi} values is determined by the \texttt{steps} and \texttt{n} arguments, respectively.
+Additional independent noise dimensions can be added via \texttt{gen\_noisedims()} to extend the structure into \(p > 4\) dimensions.
+
+The resulting dataset represents a true model \(T = (x_1, x_2, x_3, x_4)\), capturing the full 4D knot structure.
+
+\paragraph{\texorpdfstring{\texttt{gen\_trefoil3d()}}{gen\_trefoil3d()}}\label{gen_trefoil3d}
+
+The function \texttt{gen\_trefoil3d()} generates a \textbf{3-dimensional stereographic projection} of the 4D trefoil knot, mapping each point from \(\mathbb{R}^4\) to \(\mathbb{R}^3\) using the transformation:
+
+\[
+(x_1', x_2', x_3') = \left( \frac{x_1}{1 - x_4}, \frac{x_2}{1 - x_4}, \frac{x_3}{1 - x_4} \right),
+\]
+
+which excludes points where \(x_4 = 1\). The projection preserves the knot topology and enables easier visualization and interpretation.
+
+As with the 4D version, this function supports optional noise dimensions to embed the projected knot into higher dimensions.
 
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `mobiusgau` data.
@@
 
-\begin{figure}
+\begin{figure}[!]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
@@
 
-The package is available at \url{https://github.com/JayaniLakshika/cardinalR}, and source material for this paper is available at \url{https://github.com/JayaniLakshika/paper-cardinalR}.
+The source material for this paper is available at \url{https://github.com/JayaniLakshika/paper-cardinalR}.
 

### 2025-07-06 paper-cardinalR.Rmd
Updated Treofoil section

--- 2025-06-24 paper-cardinalR.Rmd
@@
 
+### Trefoil knots
+
+<!--https://laustep.github.io/stlahblog/posts/TorusKnot4D.html-->
+
+The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space.
+The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of NLDR methods to preserve global structure, loops, and embeddings in high-dimensional data.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trefoil-tb-html)', '\\@ref(tab:trefoil-tb-pdf)'))` summarizes these functions.
+
+```{r}
+trefoil_tb <- tibble(fun = c("gen_trefoil4d",
+                            "gen_trefoil3d"), 
+                        exp = c("Generate a trefoil in $4\\text{-}D$.",
+                                "Generate a trefoil in $3\\text{-}D$."))
+```
+
+```{r trefoil-tb-html, eval=knitr::is_html_output()}
+trefoil_tb |> 
+  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r trefoil-tb-pdf, eval=knitr::is_latex_output()}
+trefoil_tb |> 
+  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+#### `gen_trefoil4d()`
+
+The function `gen_trefoil4d()` generates a $4\text{-}D$ trefoil knot using two angular parameters, $\theta$ and $\phi$.
+The knot is constructed on the $3$-sphere ($S^3 \subset \mathbb{R}^4$), with a band of thickness around the knot path, controlled by `steps`.
+
+The coordinates of the $4\text{-}D$ trefoil knot are defined as $X_1 = \cos(\theta) \cos(\phi)$, $X_2 = \cos(\theta) \sin(\phi)$, $X_3 = \sin(\theta) \cos(1.5 \phi)$, and $X_4 = \sin(\theta) \sin(1.5 \phi)$, where $\theta$ and $\phi$ are angular parameters that trace out the knot's structure on the $3$-sphere $S^3 \subset \mathbb{R}^4$.
+
+The number of `theta` and `phi` values is determined by the `steps` and `n` arguments, respectively.
+Additional independent noise dimensions can be added via `gen_noisedims()` to extend the structure into $p > 4$ dimensions.
+
+```{r data-trefoil4d, echo=TRUE}
+trefoil4d <- gen_trefoil4d(n = 500, p = 4, steps = 5)
+```
+
+```{r, eval=FALSE}
+langevitour::langevitour(trefoil4d)
+```
+
+```{r trefoil4d-proj1}
+scaled_data <- scale_data_manual(trefoil4d)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+trefoil4d_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil4d-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+trefoil4d_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil4d-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+trefoil4d_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+#### `gen_trefoil3d()`
+
+The function `gen_trefoil3d()` generates a $3\text{-}D$ stereographic projection of the $4\text{-}D$ trefoil knot by mapping each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3$ using the transformation $X_1' = X_1 / (1 - X_4)$, $X_2' = X_2 / (1 - X_4)$, and $X_3' = X_3 / (1 - X_4)$, while excluding points where $X_4 = 1$ to avoid division by zero.
+
+```{r data-trefoil3d, echo=TRUE}
+trefoil3d <- gen_trefoil3d(n = 500, p = 4, steps = 5)
+```
+
+As with the $4\text{-}D$ version, this function supports optional noise dimensions to embed the projected knot into higher dimensions.
+
+```{r, eval=FALSE}
+langevitour::langevitour(trefoil3d)
+```
+
+```{r trefoil3d-proj1}
+scaled_data <- scale_data_manual(trefoil4d)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+trefoil3d_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil3d-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+trefoil3d_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil3d-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+trefoil3d_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-trefoil-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` and `trefoil3d` data.", fig.width=15, fig.height=10}
+
+trefoil4d_proj1 + trefoil4d_proj2 + trefoil4d_proj3 + 
+  trefoil3d_proj1 + trefoil3d_proj2 + trefoil3d_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Multiple cluster examples

### 2025-07-06 paper-cardinalR.tex
Updated Treofoil section

--- 2025-06-24 paper-cardinalR.tex
@@
 \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data.}\label{fig:fig-triginometric-proj}
+\end{figure}
+
+\subsubsection{Trefoil knots}\label{trefoil-knots}
+
+The Trefoil is a closed, nontrivial one-dimensional manifold embedded in \(3\text{-}D\) or \(4\text{-}D\) space.
+The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of NLDR methods to preserve global structure, loops, and embeddings in high-dimensional data.
+Table \ref{tab:trefoil-tb-pdf} summarizes these functions.
+
+\begin{table}
+
+\caption{\label{tab:trefoil-tb-pdf}cardinalR trefoil data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_trefoil4d & Generate a trefoil in \$4\textbackslash{}text\{-\}D\$.\\
+gen\_trefoil3d & Generate a trefoil in \$3\textbackslash{}text\{-\}D\$.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\paragraph{\texorpdfstring{\texttt{gen\_trefoil4d()}}{gen\_trefoil4d()}}\label{gen_trefoil4d}
+
+The function \texttt{gen\_trefoil4d()} generates a \(4\text{-}D\) trefoil knot using two angular parameters, \(\theta\) and \(\phi\).
+The knot is constructed on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)), with a band of thickness around the knot path, controlled by \texttt{steps}.
+
+The coordinates of the \(4\text{-}D\) trefoil knot are defined as \(X_1 = \cos(\theta) \cos(\phi)\), \(X_2 = \cos(\theta) \sin(\phi)\), \(X_3 = \sin(\theta) \cos(1.5 \phi)\), and \(X_4 = \sin(\theta) \sin(1.5 \phi)\), where \(\theta\) and \(\phi\) are angular parameters that trace out the knot's structure on the \(3\)-sphere \(S^3 \subset \mathbb{R}^4\).
+
+The number of \texttt{theta} and \texttt{phi} values is determined by the \texttt{steps} and \texttt{n} arguments, respectively.
+Additional independent noise dimensions can be added via \texttt{gen\_noisedims()} to extend the structure into \(p > 4\) dimensions.
+
+\begin{verbatim}
+trefoil4d <- gen_trefoil4d(n = 500, p = 4, steps = 5)
+\end{verbatim}
+
+\paragraph{\texorpdfstring{\texttt{gen\_trefoil3d()}}{gen\_trefoil3d()}}\label{gen_trefoil3d}
+
+The function \texttt{gen\_trefoil3d()} generates a \(3\text{-}D\) stereographic projection of the \(4\text{-}D\) trefoil knot by mapping each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\) using the transformation \(X_1' = X_1 / (1 - X_4)\), \(X_2' = X_2 / (1 - X_4)\), and \(X_3' = X_3 / (1 - X_4)\), while excluding points where \(X_4 = 1\) to avoid division by zero.
+
+\begin{verbatim}
+trefoil3d <- gen_trefoil3d(n = 500, p = 4, steps = 5)
+\end{verbatim}
+
+As with the \(4\text{-}D\) version, this function supports optional noise dimensions to embed the projected knot into higher dimensions.
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` and `trefoil3d` data.}\label{fig:fig-trefoil-proj}
 \end{figure}

### 2025-07-06 paper-cardinalR.Rmd
Added some multicluster examples

@@
 odd_shapes_tb <- tibble(fun = c("make_mobiusgau",
-                                "make_multigau"), 
+                                "make_multigau",
+                                "make_curvygau",
+                                "make_multigrid",
+                                "make_threeclust01",
+                                "make_fourclust01",
+                                "make_fiveclust01"), 
                         exp = c("",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
                                 ""))
@@
 
-The package includes various supplementary tools in addition to the shape-generating functions mentioned earlier.
+The package includes various supplementary tools in addition to the shape generating functions mentioned earlier.
 These tools allow users to generate noise dimensions with a normal distribution and various wavy patterns, create background noise, randomize the rows of the data, reposition clusters, generate a vector whose product and sum are approximately equal to a target value, rotate structures, and normalize the data.

### 2025-07-06 paper-cardinalR.tex
Added some multicluster examples

@@
 make\_multigau & \\
+make\_curvygau & \\
+make\_threeclust01 & \\
+make\_fourclust01 & \\
+\addlinespace
+make\_fiveclust01 & \\
 \bottomrule

### 2025-07-06 paper-cardinalR.R

@@ -3555,8 +3555,16 @@ trefoil4d_proj1 + trefoil4d_proj2 + trefoil4d_proj3 +
 
 ## -----------------------------------------------------------------------------
 odd_shapes_tb <- tibble(fun = c("make_mobiusgau",
-                                "make_multigau"), 
+                                "make_multigau",
+                                "make_curvygau",
+                                "make_threeclust01",
+                                "make_fourclust01",
+                                "make_fiveclust01"), 
                         exp = c("",
+                                "",
+                                "",
+                                "",
+                                "",
                                 ""))
 
 

### 2025-07-07 paper-cardinalR.Rmd
Added package installation for the paper

@@
 
+```{r install-libraries, include=FALSE, warning=FALSE, echo=FALSE}
+
+options(repos = c(CRAN = "https://cran.rstudio.com")) # Setup mirror
+
+packages_to_check <- c("remotes", "cardinalR", "tidyverse", "kableExtra", "geozoo", "patchwork", "colorspace")
+
+for (pkg in packages_to_check) {
+  if (!requireNamespace(pkg, quietly = TRUE)) {
+    message(paste("Installing package:", pkg))
+    install.packages(pkg)
+  } else {
+    installed_version <- packageVersion(pkg)
+    available_version <- tryCatch({
+      utils::packageDescription(pkg)$Version
+    }, error = function(e) NA) # Handle cases where package info isn't readily available
+
+    if (!is.na(available_version) && installed_version < package_version(available_version)) {
+      message(paste("A newer version of package", pkg, "is available. Updating..."))
+      install.packages(pkg)
+    } else {
+      message(paste("Package", pkg, "is up-to-date (version", installed_version, ")."))
+    }
+  }
+}
+
+```
+
 ```{r load-libraries}

### 2025-08-04 paper-cardinalR.Rmd
Updated installation

@@
 ```r
-devtools::install_github("JayaniLakshika/cardinalR")
+pak::pak("JayaniLakshika/cardinalR")
 ```
@@
 
-```{r, label = "fig-highd-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobiusgau` data.
+```{r, label = "fig-highd-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
 The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green."}
@@
 
-```{r, label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
+```{r, label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
 

### 2025-08-04 paper-cardinalR.tex
Added package installation for the paper
Updated installation

@@
 \begin{verbatim}
-devtools::install_github("JayaniLakshika/cardinalR")
+pak::pak("JayaniLakshika/cardinalR")
 \end{verbatim}
@@
 make\_curvygau & \\
+make\_multigrid & \\
 make\_threeclust01 & \\
+\addlinespace
 make\_fourclust01 & \\
-\addlinespace
 make\_fiveclust01 & \\
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `mobiusgau` data.
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
 The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:fig-highd-proj}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
 \end{figure}

### 2025-08-04 paper-cardinalR.R

@@ -3753,7 +3753,7 @@ five_clusts_proj3 <- plot_proj(
 
 
 
-## ----label = "fig-highd-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobiusgau` data. The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green."----
+## ----label = "fig-highd-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data. The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green."----
 #| fig-width: 15
 #| fig-height: 5
 
@@ -3836,7 +3836,7 @@ nldr6 <- pca_data |>
         legend.text = element_text(size = 10)) 
 
 
-## ----label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."----
+## ----label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."----
 
 nldr1 + nldr2 + nldr3 +
   nldr4 + nldr5 + nldr6 +

### 2025-08-21 paper-cardinalR.Rmd
Initial work on adding the multi cluster table

@@
                                 "make_curvygau",
-                                "make_multigrid",
-                                "make_threeclust01",
-                                "make_fourclust01",
-                                "make_fiveclust01"), 
+                                "make_klink_circles",
+                                "make_chain_circles",
+                                "make_klink_curvycycle",
+                                "make_chain_curvycycle",
+                                "make_gaucircles",
+                                "make_gaucurvycycle",
+                                "make_onegrid",
+                                "make_twogrid_overlap",
+                                "make_twogrid_shift",
+                                "make_shape_para",
+                                "make_three_clust_"), 
                         exp = c("",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
                                 "",

### 2025-08-21 paper-cardinalR.tex
Initial work on adding the multi cluster table

@@
 make\_curvygau & \\
-make\_multigrid & \\
-make\_threeclust01 & \\
+make\_klink\_circles & \\
+make\_chain\_circles & \\
 \addlinespace
-make\_fourclust01 & \\
-make\_fiveclust01 & \\
+make\_klink\_curvycycle & \\
+make\_chain\_curvycycle & \\
+make\_gaucircles & \\
+make\_gaucurvycycle & \\
+make\_onegrid & \\
+\addlinespace
+make\_twogrid\_overlap & \\
+make\_twogrid\_shift & \\
+make\_shape\_para & \\
+make\_three\_clust\_ & \\
 \bottomrule

### 2025-08-21 paper-cardinalR.Rmd
Add the table for multicluster data generation

--- 2025-08-04 paper-cardinalR.Rmd
@@
                                 "make_curvygau",
-                                "make_multigrid",
-                                "make_threeclust01",
-                                "make_fourclust01",
-                                "make_fiveclust01"), 
-                        exp = c("",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                ""))
+                                "make_klink_circles",
+                                "make_chain_circles",
+                                "make_klink_curvycycle",
+                                "make_chain_curvycycle",
+                                "make_gaucircles",
+                                "make_gaucurvycycle",
+                                "make_onegrid",
+                                "make_twogrid_overlap",
+                                "make_twogrid_shift",
+                                "make_shape_para",
+                                "make_three_clust_"), 
+                        exp = c(
+    "Generates a Möbius-like cluster combined with a Gaussian cluster.",
+    "Generates multiple Gaussian clusters in high-dimensional space.",
+    "Generates a curvilinear cluster with a Gaussian cluster.",
+    "Generates K-link circular clusters (non-linear circular patterns).",
+    "Generates chain-like circular clusters connected sequentially.",
+    "Generates K-link curvy cycle clusters (curvilinear loop structures).",
+    "Generates chain-like curvy cycle clusters connected sequentially.",
+    "Generates circular clusters with a Gaussian cluster in the middle.",
+    "Generates curvy circular clusters with a Gaussian cluster in the middle.",
+    "Generates a single grid in two dimensions.",
+    "Generates two overlapping grids.",
+    "Generates two grids shifted relative to each other.",
+    "Generates parallel shaped clusters.",
+    "Generates three clusters with different shapes. (eg:- 01, 02, ..., 20)"))
 ```

### 2025-08-21 paper-cardinalR.tex
Add the table for multicluster data generation

--- 2025-08-04 paper-cardinalR.tex
@@
 \midrule
-make\_mobiusgau & \\
-make\_multigau & \\
-make\_curvygau & \\
-make\_multigrid & \\
-make\_threeclust01 & \\
+make\_mobiusgau & Generates a Möbius-like cluster combined with a Gaussian cluster.\\
+make\_multigau & Generates multiple Gaussian clusters in high-dimensional space.\\
+make\_curvygau & Generates a curvilinear cluster with a Gaussian cluster.\\
+make\_klink\_circles & Generates K-link circular clusters (non-linear circular patterns).\\
+make\_chain\_circles & Generates chain-like circular clusters connected sequentially.\\
 \addlinespace
-make\_fourclust01 & \\
-make\_fiveclust01 & \\
+make\_klink\_curvycycle & Generates K-link curvy cycle clusters (curvilinear loop structures).\\
+make\_chain\_curvycycle & Generates chain-like curvy cycle clusters connected sequentially.\\
+make\_gaucircles & Generates circular clusters with a Gaussian cluster in the middle.\\
+make\_gaucurvycycle & Generates curvy circular clusters with a Gaussian cluster in the middle.\\
+make\_onegrid & Generates a single grid in two dimensions.\\
+\addlinespace
+make\_twogrid\_overlap & Generates two overlapping grids.\\
+make\_twogrid\_shift & Generates two grids shifted relative to each other.\\
+make\_shape\_para & Generates parallel shaped clusters.\\
+make\_three\_clust\_ & Generates three clusters with different shapes. (eg:- 01, 02, ..., 20)\\
 \bottomrule

### 2025-08-21 paper-cardinalR.R

@@ -3595,20 +3595,21 @@ odd_shapes_tb <- tibble(fun = c("make_mobiusgau",
                                 "make_twogrid_shift",
                                 "make_shape_para",
                                 "make_three_clust_"), 
-                        exp = c("",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                ""))
+                        exp = c(
+    "Generates a Möbius-like cluster combined with a Gaussian cluster.",
+    "Generates multiple Gaussian clusters in high-dimensional space.",
+    "Generates a curvilinear cluster with a Gaussian cluster.",
+    "Generates K-link circular clusters (non-linear circular patterns).",
+    "Generates chain-like circular clusters connected sequentially.",
+    "Generates K-link curvy cycle clusters (curvilinear loop structures).",
+    "Generates chain-like curvy cycle clusters connected sequentially.",
+    "Generates circular clusters with a Gaussian cluster in the middle.",
+    "Generates curvy circular clusters with a Gaussian cluster in the middle.",
+    "Generates a single grid in two dimensions.",
+    "Generates two overlapping grids.",
+    "Generates two grids shifted relative to each other.",
+    "Generates parallel shaped clusters.",
+    "Generates three clusters with different shapes. (eg:- 01, 02, ..., 20)"))
 
 
 ## ----odd-shape-tb-html, eval=knitr::is_html_output()--------------------------

### 2025-08-22 paper-cardinalR.Rmd
Updaded grided sphere section

@@
 
-The `gen_gridedsphere(n, p)` function generates a $p$-dimensional dataset of approximately $n$ observations evenly distributed on the surface of a $3\text{-}D$ unit sphere, with optional Gaussian noise dimensions when $p > 3$.
-
-The base structure consists of a $3\text{-}D$ spherical surface created using a regular grid in spherical coordinates, where $\theta \in [0, 2\pi]$ represents the azimuthal angle and $\phi \in [0, \pi]$ denotes the polar angle.
-
-The number of grid steps along each dimension is determined by factoring $n$ into two approximately equal integers via `gen_nproduct(n, p = 2)`.
-
-Each point on the sphere is computed using the spherical-to-Cartesian transformation: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi)$.
-For $p > 3$, the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions: $X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p$.
+The `gen_gridedsphere(n, p)` function generates a $p$-dimensional dataset of approximately $n$ points evenly distributed on the surface of a $(p!-!1)$-sphere (the unit sphere embedded in $\mathbb{R}^p$).
+
+The construction is based on a regular grid in spherical coordinates with $(p-1)$ angular variables. Specifically:
+
+* For dimensions $j = 1, \dots, p-2$, the polar angles are defined on $[0, \pi]$.
+* For the last angle ($j = p-1$), the azimuthal angle is defined on $[0, 2\pi]$.
+
+The number of grid steps along each angular dimension is determined by factoring $n$ into $(p-1)$ approximately equal integers via `gen_nproduct(n, p - 1)`.
+
+Each grid point is then mapped to Cartesian coordinates using the standard hyperspherical-to-Cartesian transformation:
+
+$$
+\begin{aligned}
+X_1 &= \cos(\theta_1), \\
+X_2 &= \sin(\theta_1)\cos(\theta_2), \\
+X_3 &= \sin(\theta_1)\sin(\theta_2)\cos(\theta_3), \\
+&\;\;\vdots \\
+X_{p-1} &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\cos(\theta_{p-1}), \\
+X_p &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\sin(\theta_{p-1}).
+\end{aligned}
+$$
+
+Thus, the output is a deterministic grid of points lying exactly on the unit $(p-1)$-sphere, with no added noise dimensions.
 

### 2025-08-22 paper-cardinalR.tex
Updaded grided sphere section

@@
 
-The \texttt{gen\_gridedsphere(n,\ p)} function generates a \(p\)-dimensional dataset of approximately \(n\) observations evenly distributed on the surface of a \(3\text{-}D\) unit sphere, with optional Gaussian noise dimensions when \(p > 3\).
-
-The base structure consists of a \(3\text{-}D\) spherical surface created using a regular grid in spherical coordinates, where \(\theta \in [0, 2\pi]\) represents the azimuthal angle and \(\phi \in [0, \pi]\) denotes the polar angle.
-
-The number of grid steps along each dimension is determined by factoring \(n\) into two approximately equal integers via \texttt{gen\_nproduct(n,\ p\ =\ 2)}.
-
-Each point on the sphere is computed using the spherical-to-Cartesian transformation: \(X_1 = \sin(\phi)\cos(\theta)\), \(X_2 = \sin(\phi)\sin(\theta)\), \(X_3 = \cos(\phi)\).
-For \(p > 3\), the manifold is embedded in a higher-dimensional space by adding Gaussian noise dimensions: \(X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p\).
+The \texttt{gen\_gridedsphere(n,\ p)} function generates a \(p\)-dimensional dataset of approximately \(n\) points evenly distributed on the surface of a \((p!-!1)\)-sphere (the unit sphere embedded in \(\mathbb{R}^p\)).
+
+The construction is based on a regular grid in spherical coordinates with \((p-1)\) angular variables. Specifically:
+
+\begin{itemize}
+\tightlist
+\item
+  For dimensions \(j = 1, \dots, p-2\), the polar angles are defined on \([0, \pi]\).
+\item
+  For the last angle (\(j = p-1\)), the azimuthal angle is defined on \([0, 2\pi]\).
+\end{itemize}
+
+The number of grid steps along each angular dimension is determined by factoring \(n\) into \((p-1)\) approximately equal integers via \texttt{gen\_nproduct(n,\ p\ -\ 1)}.
+
+Each grid point is then mapped to Cartesian coordinates using the standard hyperspherical-to-Cartesian transformation:
+
+\[
+\begin{aligned}
+X_1 &= \cos(\theta_1), \\
+X_2 &= \sin(\theta_1)\cos(\theta_2), \\
+X_3 &= \sin(\theta_1)\sin(\theta_2)\cos(\theta_3), \\
+&\;\;\vdots \\
+X_{p-1} &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\cos(\theta_{p-1}), \\
+X_p &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\sin(\theta_{p-1}).
+\end{aligned}
+\]
+
+Thus, the output is a deterministic grid of points lying exactly on the unit \((p-1)\)-sphere, with no added noise dimensions.
 

### 2025-08-22 paper-cardinalR.Rmd
Updated the conclusion

@@
 
-The `cardinalR` package contributes a flexible and extensible framework for generating high-dimensional data structures with well-defined geometric properties.
-It addresses a growing need in the evaluation of machine learning, clustering, and nonlinear dimensionality reduction (NLDR) methods—offering researchers the ability to design customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
-This makes `cardinalR` a useful complement to existing tools like `geozoo`, `snedata`, and `mlbench`, expanding the scope to higher dimensions and more complex shapes.
-
-The suite of included structures is diverse and targeted.
-For example, **branching shapes** allow investigation of continuity and topological preservation in low-dimensional embeddings.
-The **S-curve with a hole** enables exploration of how methods handle incomplete manifolds.
-**Clustered spheres** assess cluster separation on curved surfaces, while the **Möbius strip** challenges NLDR methods with non-orientable geometry.
-**Gridded cubes** test spatial regularity, and **pyrholes** represent non-convex, sparse shapes useful for testing clustering performance in irregular high-dimensional regions.
-
-These structures are not only valuable for algorithm diagnostics but may also be used for **teaching high-dimensional concepts**, **benchmarking reproducibility**, and **evaluating hyperparameter sensitivity**.
-The package's design encourages users to adjust dimensionality, noise, sample size, and clustering properties, promoting **transparent experimentation** and **comparative model evaluation**.
-
-Looking ahead, future extensions of `cardinalR` could incorporate dynamic shape generation, integration with supervised learning simulations, or visual interfaces to assist users in interactively exploring generated structures.
-Expanding the library to include more biologically inspired or real-world data analogs may also broaden its utility across domains like bioinformatics, forensic science, and spatial data analysis.
+The `cardinalR` package introduces a flexible framework for generating high-dimensional data structures with well-defined geometric properties.
+It addresses an important need in the evaluation of clustering, machine learning, and dimensionality reduction (DR) methods by enabling the construction of customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
+In this way, `cardinalR` complements existing packages such as `geozoo`, `snedata`, and `mlbench`, while extending the scope to higher dimensions and more complex shapes.
+
+The included structures cover a wide range of diagnostic settings.
+Branching shapes facilitate the study of continuity and topological preservation, the Scurve with a hole allows investigation of incomplete manifolds, and clustered spheres assess separability on curved surfaces.
+The Möbius strip introduces challenges from non-orientable geometry, while gridded cubes and pyrholes test spatial regularity and clustering in sparse, non-convex regions.
+
+These structures are designed to support not only algorithm diagnostics, but also teaching high-dimensional concepts, benchmarking reproducibility, and evaluating hyperparameter sensitivity.
+By allowing users to adjust dimensionality, sample size, noise, and clustering properties, the package promotes transparent experimentation and comparative model evaluation.
+
+Future extensions of `cardinalR` may include biologically inspired or application-driven data structures would further broaden its utility in domains such as bioinformatics, forensic science, and spatial analysis.
 

### 2025-08-22 paper-cardinalR.tex
Updated the conclusion

@@
 
-The \texttt{cardinalR} package contributes a flexible and extensible framework for generating high-dimensional data structures with well-defined geometric properties.
-It addresses a growing need in the evaluation of machine learning, clustering, and nonlinear dimensionality reduction (NLDR) methods---offering researchers the ability to design customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
-This makes \texttt{cardinalR} a useful complement to existing tools like \texttt{geozoo}, \texttt{snedata}, and \texttt{mlbench}, expanding the scope to higher dimensions and more complex shapes.
-
-The suite of included structures is diverse and targeted.
-For example, \textbf{branching shapes} allow investigation of continuity and topological preservation in low-dimensional embeddings.
-The \textbf{S-curve with a hole} enables exploration of how methods handle incomplete manifolds.
-\textbf{Clustered spheres} assess cluster separation on curved surfaces, while the \textbf{Möbius strip} challenges NLDR methods with non-orientable geometry.
-\textbf{Gridded cubes} test spatial regularity, and \textbf{pyrholes} represent non-convex, sparse shapes useful for testing clustering performance in irregular high-dimensional regions.
-
-These structures are not only valuable for algorithm diagnostics but may also be used for \textbf{teaching high-dimensional concepts}, \textbf{benchmarking reproducibility}, and \textbf{evaluating hyperparameter sensitivity}.
-The package's design encourages users to adjust dimensionality, noise, sample size, and clustering properties, promoting \textbf{transparent experimentation} and \textbf{comparative model evaluation}.
-
-Looking ahead, future extensions of \texttt{cardinalR} could incorporate dynamic shape generation, integration with supervised learning simulations, or visual interfaces to assist users in interactively exploring generated structures.
-Expanding the library to include more biologically inspired or real-world data analogs may also broaden its utility across domains like bioinformatics, forensic science, and spatial data analysis.
+The \texttt{cardinalR} package introduces a flexible framework for generating high-dimensional data structures with well-defined geometric properties.
+It addresses an important need in the evaluation of clustering, machine learning, and dimensionality reduction (DR) methods by enabling the construction of customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
+In this way, \texttt{cardinalR} complements existing packages such as \texttt{geozoo}, \texttt{snedata}, and \texttt{mlbench}, while extending the scope to higher dimensions and more complex shapes.
+
+The included structures cover a wide range of diagnostic settings.
+Branching shapes facilitate the study of continuity and topological preservation, the Scurve with a hole allows investigation of incomplete manifolds, and clustered spheres assess separability on curved surfaces.
+The Möbius strip introduces challenges from non-orientable geometry, while gridded cubes and pyrholes test spatial regularity and clustering in sparse, non-convex regions.
+
+These structures are designed to support not only algorithm diagnostics, but also teaching high-dimensional concepts, benchmarking reproducibility, and evaluating hyperparameter sensitivity.
+By allowing users to adjust dimensionality, sample size, noise, and clustering properties, the package promotes transparent experimentation and comparative model evaluation.
+
+Future extensions of \texttt{cardinalR} may include biologically inspired or application-driven data structures would further broaden its utility in domains such as bioinformatics, forensic science, and spatial analysis.
 

### 2025-08-22 paper-cardinalR.Rmd
Updated figure1

@@
 
-A branching structure captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
+A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
@@
   plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
+  title = "b1", 
   cex = 2, 
@@
   plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
+  title = "b2", 
   cex = 2, 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
+  title = "b3", 
   cex = 2, 
@@
   plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
+  title = "c1", 
   cex = 2, 
@@
   plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
+  title = "c2", 
   cex = 2, 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
+  title = "c3", 
   cex = 2, 
@@
   plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
+  title = "d1", 
   cex = 2, 
@@
   plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
+  title = "d2", 
   cex = 2, 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
+  title = "d3", 
   cex = 2, 
@@
   plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
+  title = "e1", 
   cex = 2, 
@@
   plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
+  title = "e2", 
   cex = 2, 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r, label = "fig-orgcurvybranches-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, `orgcurvybranches` data.", fig.width=15, fig.height=25}
+  title = "e3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r branch-proj, fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=15, fig.height=25}
 

### 2025-08-22 paper-cardinalR.tex
Updated figure1

@@
 
-A branching structure captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
+A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-orgcurvybranches-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, `orgcurvybranches` data.}\label{fig:fig-orgcurvybranches-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.}\label{fig:branch-proj}
 \end{figure}

### 2025-08-22 paper-cardinalR.Rmd
Added more detailed figure titles

@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r, label = "fig-cubehole-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube`, `unifcube`, and `cubehole` data.", fig.width=15, fig.height=15}
+  title = "c3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-cube-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=15, fig.height=15}
 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r, label = "fig-cubic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` and `cubic` data.", fig.width=15, fig.height=10}
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-cubic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.", fig.width=15, fig.height=10}
 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r, label = "fig-pyrholes-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data.", fig.width=15, fig.height=20}
+  title = "d3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-pyr-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=15, fig.height=20}
 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r, label = "fig-scurve-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` and `scurvehole` data.", fig.width=15, fig.height=10}
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-scurve-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.", fig.width=15, fig.height=10}
 
@@
 
-The construction is based on a regular grid in spherical coordinates with $(p-1)$ angular variables. Specifically:
-
-* For dimensions $j = 1, \dots, p-2$, the polar angles are defined on $[0, \pi]$.
-* For the last angle ($j = p-1$), the azimuthal angle is defined on $[0, 2\pi]$.
+The construction is based on a regular grid in spherical coordinates with $(p-1)$ angular variables.
+Specifically, for dimensions $j = 1, \dots, p-2$, the polar angles are defined on $[0, \pi]$.
+For the last angle ($j = p-1$), the azimuthal angle is defined on $[0, 2\pi]$.
 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
+  title = "e3", 
   cex = 2, 
@@
   plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
+  title = "f1", 
   cex = 2, 
@@
   plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
+  title = "f2", 
   cex = 2, 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r, label = "fig-hemisphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle`, `curvycycle`, `unifsphere`, `gridedsphere`, `clusteredspheres`, and `hemisphere` data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
+  title = "f3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-sphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
 
@@
   plot_limits = c(-0.45, 0.73), 
-  title = "a1", 
+  title = "f1", 
   cex = 2, 
@@
   plot_limits = c(-0.3, 0.43), 
-  title = "a2", 
+  title = "f2", 
   cex = 2, 
@@
   plot_limits = c(-0.35, 0.4), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r, label = "fig-triginometric-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
+  title = "f3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-triginometric-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r, label = "fig-trefoil-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` and `trefoil3d` data.", fig.width=15, fig.height=10}
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r, label = "fig-trefoil-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=15, fig.height=10}
 

### 2025-08-22 paper-cardinalR.tex
Added more detailed figure titles

@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubehole-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube`, `unifcube`, and `cubehole` data.}\label{fig:fig-cubehole-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.}\label{fig:fig-cube-proj}
 \end{figure}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` and `cubic` data.}\label{fig:fig-cubic-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.}\label{fig:fig-cubic-proj}
 \end{figure}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyrholes-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data.}\label{fig:fig-pyrholes-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.}\label{fig:fig-pyr-proj}
 \end{figure}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` and `scurvehole` data.}\label{fig:fig-scurve-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.}\label{fig:fig-scurve-proj}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/fig-hemisphere-proj-1}
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/fig-sphere-proj-1}
 
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle`, `curvycycle`, `unifsphere`, `gridedsphere`, `clusteredspheres`, and `hemisphere` data.}\label{fig:fig-hemisphere-proj}
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.}\label{fig:fig-sphere-proj}
 \end{figure}
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data.}\label{fig:fig-triginometric-proj}
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.}\label{fig:fig-triginometric-proj}
 \end{figure}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` and `trefoil3d` data.}\label{fig:fig-trefoil-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.}\label{fig:fig-trefoil-proj}
 \end{figure}

### 2025-08-22 paper-cardinalR.R

@@ -861,7 +861,7 @@ unifcube_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
+  title = "b1", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
@@ -888,7 +888,7 @@ unifcube_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
+  title = "b2", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
@@ -915,7 +915,7 @@ unifcube_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
+  title = "b3", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
@@ -951,7 +951,7 @@ cubehole_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
+  title = "c1", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
@@ -978,7 +978,7 @@ cubehole_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
+  title = "c2", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
@@ -1005,14 +1005,14 @@ cubehole_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
+  title = "c3", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
 
 
 
[…]

### 2025-08-22 paper-cardinalR.Rmd
Updated application section with langevitour output

@@
 
-The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in $4\text{-}D$ space (Figure \@ref(fig:fig-highd-proj)).
+The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
 
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(five_clusts[, -5])
+```{r highd-data, eval=knitr::is_html_output(), fig.cap="`langevitour` output showing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster."}
+
+langevitour::langevitour(five_clusts[, -5], group = five_clusts$cluster, levelColors = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e'))
 ```
@@
 
-```{r, label = "fig-highd-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
+```{r highd-proj, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
 The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green."}

### 2025-08-22 paper-cardinalR.tex
Updated application section with langevitour output

@@
 
-The construction is based on a regular grid in spherical coordinates with \((p-1)\) angular variables. Specifically:
-
-\begin{itemize}
-\tightlist
-\item
-  For dimensions \(j = 1, \dots, p-2\), the polar angles are defined on \([0, \pi]\).
-\item
+The construction is based on a regular grid in spherical coordinates with \((p-1)\) angular variables.
+Specifically, for dimensions \(j = 1, \dots, p-2\), the polar angles are defined on \([0, \pi]\).
 For the last angle (\(j = p-1\)), the azimuthal angle is defined on \([0, 2\pi]\).
-\end{itemize}
 
@@
 
-The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in \(4\text{-}D\) space (Figure \ref{fig:fig-highd-proj}).
+The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in \(4\text{-}D\) space (Figure \ref{fig:highd-proj}).
 
@@
 
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
-The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:fig-highd-proj}
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:highd-proj}
 \end{figure}

### 2025-08-22 paper-cardinalR.Rmd
Updated the figure position

@@
 
-```{r, label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
+```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
 

### 2025-08-22 paper-cardinalR.R

@@ -3852,7 +3852,7 @@ nldr6 <- pca_data |>
         legend.text = element_text(size = 10)) 
 
 
-## ----label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."----
+## ----label = "fig-nldr-layouts", fig.pos="H", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."----
 
 nldr1 + nldr2 + nldr3 +
   nldr4 + nldr5 + nldr6 +

### 2025-08-22 paper-cardinalR.Rmd
Updated Trefoil projection

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.62, 0.62), 
   title = "a3", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.62, 0.62), 
   title = "b3", 

### 2025-08-22 paper-cardinalR.Rmd
Added trefoil data

@@
 library(colorspace)
+library(crosstalk)
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(trefoil4d)
+```{r}
+trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(trefoil3d)
+```{r}
+trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-trefoil-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=15, fig.height=10}
+```{r trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` and `trefoil3d` data in $4\\text{-}D$."}
+
+trefoilfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    trefoil4d_lang, trefoil3d_lang
+  ),
+  device = "xs"
+)
+
+class(trefoilfig) <- c(class(trefoilfig), "htmlwidget")
+
+trefoilfig
+```
+
+```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=15, fig.height=10}
 

### 2025-08-22 paper-cardinalR.tex
Updated the figure position
Added trefoil data

@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.}\label{fig:fig-trefoil-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.}\label{fig:trefoil-proj}
 \end{figure}
@@
 
-\begin{figure}[!]
+\begin{figure}[H]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}

### 2025-08-22 paper-cardinalR.R

@@ -51,6 +51,7 @@ library(kableExtra)
 library(geozoo)
 library(patchwork)
 library(colorspace)
+library(crosstalk)
 
 
 ## -----------------------------------------------------------------------------
@@ -3397,8 +3398,8 @@ trefoil_tb |>
 trefoil4d <- gen_trefoil4d(n = 500, p = 4, steps = 5)
 
 
-## ----eval=FALSE---------------------------------------------------------------
-# langevitour::langevitour(trefoil4d)
+## -----------------------------------------------------------------------------
+trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 
 
 ## ----trefoil4d-proj1----------------------------------------------------------
@@ -3487,8 +3488,8 @@ trefoil4d_proj3 <- plot_proj(
 trefoil3d <- gen_trefoil3d(n = 500, p = 4, steps = 5)
 
 
-## ----eval=FALSE---------------------------------------------------------------
-# langevitour::langevitour(trefoil3d)
+## -----------------------------------------------------------------------------
+trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 
 
 ## ----trefoil3d-proj1----------------------------------------------------------
@@ -3573,7 +3574,22 @@ trefoil3d_proj3 <- plot_proj(
 
 
 
-## ----label = "fig-trefoil-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=15, fig.height=10----
+## ----trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` and `trefoil3d` data in $4\\text{-}D$."----
+# 
+# trefoilfig <- bscols(
+#   htmltools::div(
+#     style = "display: grid; grid-template-columns: 1fr 1fr;",
+#     trefoil4d_lang, trefoil3d_lang
+#   ),
+#   device = "xs"
+# )
+# 
+# class(trefoilfig) <- c(class(trefoilfig), "htmlwidget")
+# 
+# trefoilfig
+
+
[…]

### 2025-08-22 paper-cardinalR.Rmd
Updated some data structures

@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(swissroll)
+```{r}
+swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-swissroll-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `swissroll` data."}
+```{r swissroll, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `swissroll` data in $4\\text{-}D$."}
+
+swissrollfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    swissroll_lang
+  ),
+  device = "xs"
+)
+
+class(swissrollfig) <- c(class(swissrollfig), "htmlwidget")
+
+swissrollfig
+```
+
+```{r swissroll-proj, eval=knitr::is_latex_output(), fig.pos="H", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `swissroll` data."}
 #| fig-width: 15

### 2025-08-22 paper-cardinalR.tex
Updated some data structures

@@
 
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-swissroll-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `swissroll` data.}\label{fig:fig-swissroll-proj}
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/swissroll-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `swissroll` data.}\label{fig:swissroll-proj}
 \end{figure}

### 2025-08-22 paper-cardinalR.Rmd
Added Scurve data

@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(scurve)
+```{r}
+scurve_lang <- langevitour::langevitour(scurve, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(scurvehole)
+```{r}
+scurvehole_lang <- langevitour::langevitour(scurvehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-scurve-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.", fig.width=15, fig.height=10}
+```{r scurve, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `scurve` data in $4\\text{-}D$."}
+
+scurvefig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    scurve_lang, scurvehole_lang
+  ),
+  device = "xs"
+)
+
+class(scurvefig) <- c(class(scurvefig), "htmlwidget")
+
+scurvefig
+```
+
+```{r scurve-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.", fig.width=15, fig.height=10}
 

### 2025-08-22 paper-cardinalR.tex
Added Scurve data

@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.}\label{fig:fig-scurve-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.}\label{fig:scurve-proj}
 \end{figure}

### 2025-08-23 paper-cardinalR.Rmd
Added Mobius data

@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(mobius)
+```{r}
+mobius_lang <- langevitour::langevitour(mobius, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-mobius-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."}
+```{r mobius, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `mobius` data in $4\\text{-}D$."}
+
+mobiusfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    mobius_lang
+  ),
+  device = "xs"
+)
+
+class(mobiusfig) <- c(class(mobiusfig), "htmlwidget")
+
+mobiusfig
+```
+
+```{r mobius-proj, fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."}
 #| fig-width: 15
@@
 
-```{r scurve, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `scurve` data in $4\\text{-}D$."}
+```{r scurve, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `scurve` and `scurvehole` data in $4\\text{-}D$."}
 

### 2025-08-23 paper-cardinalR.tex
Added Mobius data

@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-mobius-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `mobius` data.}\label{fig:fig-mobius-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/mobius-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `mobius` data.}\label{fig:mobius-proj}
 \end{figure}

### 2025-08-23 paper-cardinalR.R

@@ -1221,8 +1221,8 @@ linear_proj1 + linear_proj2 + linear_proj3 +
 mobius <- gen_mobius(n = 1000, p = 4)
 
 
-## ----eval=FALSE---------------------------------------------------------------
-# langevitour::langevitour(mobius)
+## -----------------------------------------------------------------------------
+mobius_lang <- langevitour::langevitour(mobius, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 
 
 ## ----mobius-proj1-------------------------------------------------------------
@@ -1307,7 +1307,22 @@ mobius_proj3 <- plot_proj(
 
 
 
-## ----label = "fig-mobius-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."----
+## ----mobius, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `mobius` data in $4\\text{-}D$."----
+# 
+# mobiusfig <- bscols(
+#   htmltools::div(
+#     style = "display: grid; grid-template-columns: 1fr 1fr;",
+#     mobius_lang
+#   ),
+#   device = "xs"
+# )
+# 
+# class(mobiusfig) <- c(class(mobiusfig), "htmlwidget")
+# 
+# mobiusfig
+
+
+## ----mobius-proj, fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."----
 #| fig-width: 15
 #| fig-height: 5
 
@@ -2112,7 +2127,7 @@ scurvehole_proj3 <- plot_proj(
 
 
 
-## ----scurve, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `scurve` data in $4\\text{-}D$."----
+## ----scurve, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `scurve` and `scurvehole` data in $4\\text{-}D$."----
 # 
 # scurvefig <- bscols(
 #   htmltools::div(

### 2025-08-23 paper-cardinalR.Rmd
Added data for some of data structures

@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(cone)
+```{r}
+cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-cone-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."}
+```{r cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $4\\text{-}D$."}
+
+conefig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    cone_lang
+  ),
+  device = "xs"
+)
+
+class(conefig) <- c(class(conefig), "htmlwidget")
+
+conefig
+```
+
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."}
 #| fig-width: 15
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(gau)
+```{r}
+gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-gau-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gau` data."}
+```{r gau, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gau` data in $4\\text{-}D$."}
+
+gaufig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    gau_lang
+  ),
+  device = "xs"
+)
+
+class(gaufig) <- c(class(gaufig), "htmlwidget")
+
+gaufig
+```
+
+```{r gau-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gau` data."}
 #| fig-width: 15
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(linear)
+```{r}
+linear_lang <- langevitour::langevitour(linear, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-linear-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linear` data."}
+```{r linear, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `linear` data in $4\\text{-}D$."}
+
+linearfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    linear_lang
+  ),
+  device = "xs"
+)
+
+class(linearfig) <- c(class(linearfig), "htmlwidget")
+
+linearfig
+```
+
+```{r linear-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linear` data."}
 #| fig-width: 15
@@
 
-```{r mobius-proj, fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."}
+```{r mobius-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."}
 #| fig-width: 15

### 2025-08-23 paper-cardinalR.tex
Added data for some of data structures

@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:fig-cone-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:cone-proj}
 \end{figure}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:fig-gau-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:gau-proj}
 \end{figure}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-linear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linear` data.}\label{fig:fig-linear-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/linear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linear` data.}\label{fig:linear-proj}
 \end{figure}

### 2025-08-23 paper-cardinalR.Rmd
Added branch data

@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(expbranches)
+```{r}
+expbranches_lang <- langevitour::langevitour(expbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(linearbranches)
+```{r}
+linearbranches_lang <- langevitour::langevitour(linearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(curvybranches)
+```{r}
+curvybranches_lang <- langevitour::langevitour(curvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(orglinearbranches)
+```{r}
+orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(orgcurvybranches)
+```{r}
+orgcurvybranches_lang <- langevitour::langevitour(orgcurvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r branch-proj, fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=15, fig.height=25}
+```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, and `orgcurvybranches` data in $4\\text{-}D$."}
+
+branchfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    expbranches_lang, linearbranches_lang, curvybranches_lang, orglinearbranches_lang, orgcurvybranches_lang
+  ),
+  device = "xs"
+)
+
+class(branchfig) <- c(class(branchfig), "htmlwidget")
+
+branchfig
+```
+
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=15, fig.height=25}
 

### 2025-08-23 paper-cardinalR.Rmd
Added langevitour views of the data in HTML view

@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(gridcube)
+```{r}
+gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(unifcube)
+```{r}
+unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(cubehole)
+```{r}
+cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-cube-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=15, fig.height=15}
+```{r cube, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gridcube`, `unifcube`, and `cubehole` data in $4\\text{-}D$."}
+
+branchfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    gridcube_lang, unifcube_lang, cubehole_lang
+  ),
+  device = "xs"
+)
+
+class(branchfig) <- c(class(branchfig), "htmlwidget")
+
+branchfig
+```
+
+```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=15, fig.height=15}
 
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(quadratic)
+```{r}
+quadratic_lang <- langevitour::langevitour(quadratic, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(cubic)
+```{r}
+cubic_lang <- langevitour::langevitour(cubic, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-cubic-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.", fig.width=15, fig.height=10}
+```{r poly, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `quadratic` and `cubic` data in $4\\text{-}D$."}
+
+polyfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    quadratic_lang, cubic_lang
+  ),
+  device = "xs"
+)
+
+class(polyfig) <- c(class(polyfig), "htmlwidget")
+
+polyfig
+```
+
+```{r poly-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.", fig.width=15, fig.height=10}
 
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(pyrrect)
+```{r}
+pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(pyrtri)
+```{r}
+pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(pyrstar)
+```{r}
+pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(pyrholes)
+```{r}
+pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-pyr-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=15, fig.height=20}
+```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$."}
+
+pyrfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    pyrrect_lang, pyrtri_lang, pyrstar_lang, pyrholes_lang
+  ),
+  device = "xs"
+)
+
+class(pyrfig) <- c(class(pyrfig), "htmlwidget")
+
+pyrfig
+```
+
+```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=15, fig.height=20}
 
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(circle)
+```{r}
+circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(curvycycle)
+```{r}
+curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(unifsphere)
+```{r}
+unifsphere_lang <- langevitour::langevitour(unifsphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(gridedsphere)
+```{r}
+gridedsphere_lang <- langevitour::langevitour(gridedsphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(clusteredspheres)
+```{r}
+clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(hemisphere)
+```{r}
+hemisphere_lang <- langevitour::langevitour(hemisphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-sphere-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
+```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, `unifsphere`, `gridedsphere`, `clusteredspheres`, and `hemisphere` data in $4\\text{-}D$."}
+
+spherefig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    circle_lang, curvycycle_lang, unifsphere_lang, gridedsphere_lang, clusteredspheres_lang, hemisphere_lang
+  ),
+  device = "xs"
+)
+
+class(spherefig) <- c(class(spherefig), "htmlwidget")
+
+spherefig
+```
+
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
 
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(crescent)
+```{r}
+crescent_lang <- langevitour::langevitour(crescent, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(curvycylinder)
+```{r}
+curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(sphericalspiral)
+```{r}
+sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(helicalspiral)
+```{r}
+helicalspiral_lang <- langevitour::langevitour(helicalspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(conicspiral)
+```{r}
+conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, eval=FALSE}
-langevitour::langevitour(nonlinear)
+```{r}
+nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 ```
@@
 
-```{r, label = "fig-triginometric-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
+```{r triginometric, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of tthe `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data in $4\\text{-}D$."}
+
+triginometricfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    crescent_lang, curvycylinder_lang, sphericalspiral_lang, helicalspiral_lang, conicspiral_lang, nonlinear_lang
+  ),
+  device = "xs"
+)
+
+class(triginometricfig) <- c(class(triginometricfig), "htmlwidget")
+
+triginometricfig
+```
+
+```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
 

### 2025-08-23 paper-cardinalR.tex
Added langevitour views of the data in HTML view

@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.}\label{fig:fig-cube-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.}\label{fig:cube-proj}
 \end{figure}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cubic-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.}\label{fig:fig-cubic-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/poly-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.}\label{fig:poly-proj}
 \end{figure}
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.}\label{fig:fig-pyr-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.}\label{fig:pyr-proj}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/fig-sphere-proj-1}
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}
 
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.}\label{fig:fig-sphere-proj}
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.}\label{fig:sphere-proj}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/fig-triginometric-proj-1}
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
 
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.}\label{fig:fig-triginometric-proj}
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.}\label{fig:triginometric-proj}
 \end{figure}

### 2025-08-23 paper-cardinalR.R

@@ -777,8 +777,8 @@ cube_tb |>
 gridcube <- gen_gridcube(n = 1000, p = 4)
 
 
-## ----eval=FALSE---------------------------------------------------------------
-# langevitour::langevitour(gridcube)
+## -----------------------------------------------------------------------------
+gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 
 
 ## ----gridcube-proj1-----------------------------------------------------------
@@ -867,8 +867,8 @@ gridcube_proj3 <- plot_proj(
 unifcube <- gen_unifcube(n = 1000, p = 4)
 
 
-## ----eval=FALSE---------------------------------------------------------------
-# langevitour::langevitour(unifcube)
+## -----------------------------------------------------------------------------
+unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 
 
 ## ----unifcube-proj1-----------------------------------------------------------
@@ -957,8 +957,8 @@ unifcube_proj3 <- plot_proj(
 cubehole <- gen_cubehole(n = 1000, p = 4)
 
 
-## ----eval=FALSE---------------------------------------------------------------
-# langevitour::langevitour(cubehole)
+## -----------------------------------------------------------------------------
+cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
 
 
 ## ----cubehole-proj1-----------------------------------------------------------
@@ -1043,7 +1043,22 @@ cubehole_proj3 <- plot_proj(
 
 
 
-## ----label = "fig-cube-proj", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=15, fig.height=15----
+## ----cube, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gridcube`, `unifcube`, and `cubehole` data in $4\\text{-}D$."----
+# 
+# branchfig <- bscols(
+#   htmltools::div(
+#     style = "display: grid; grid-template-columns: 1fr 1fr;",
+#     gridcube_lang, unifcube_lang, cubehole_lang
[…]

### 2025-08-23 paper-cardinalR.Rmd
Updated cone projections

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.1, 0.1), 
+  plot_limits = c(-0.1, 0.12), 
   title = "a2", 

### 2025-08-23 paper-cardinalR.R

@@ -694,7 +694,7 @@ proj_obj2 <- get_projection(projection = projection,
 cone_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.1, 0.1), 
+  plot_limits = c(-0.1, 0.12), 
   title = "a2", 
   cex = 2, 
   axis_text_size = 5,

### 2025-08-24 paper-cardinalR.Rmd
Started updating the projections

@@
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.83, 0.8), 
   title = "a1", 
@@
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
+                                              axis_pos_x = -0.8, 
+                                              axis_pos_y = -0.8, 
                                               threshold = 0.01))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-1, 1), 
   title = "a2", 
@@
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.8, 0.8), 
   title = "a3", 

### 2025-08-24 paper-cardinalR.R

@@ -794,14 +794,14 @@ proj_obj1 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
                                               threshold = 0.022))
 
 gridcube_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.83, 0.8), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 5,
@@ -821,14 +821,14 @@ proj_obj2 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
+                                              axis_pos_x = -0.8, 
+                                              axis_pos_y = -0.8, 
                                               threshold = 0.01))
 
 gridcube_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-1, 1), 
   title = "a2", 
   cex = 2, 
   axis_text_size = 5,
@@ -848,14 +848,14 @@ proj_obj3 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.6,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
                                               threshold = 0.016))
 
 gridcube_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.8, 0.8), 
   title = "a3", 
   cex = 2, 
   axis_text_size = 5,

### 2025-08-24 paper-cardinalR.Rmd
Updated Trefoil projections

@@
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "a1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.65), 
   title = "a2", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.7,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.62, 0.62), 
+  plot_limits = c(-0.7, 0.62), 
   title = "a3", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "b1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.65), 
   title = "b2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.62, 0.62), 
+  plot_limits = c(-0.7, 0.62), 
   title = "b3", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 12 (f)

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.43), 
+  plot_limits = c(-0.3, 0.46), 
   title = "f2", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.5,
+                            axis_param = list(limits = 0.45,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
                                               threshold = 0.016))

### 2025-08-24 paper-cardinalR.R

@@ -3494,7 +3494,7 @@ proj_obj2 <- get_projection(projection = projection,
 nonlinear_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.43), 
+  plot_limits = c(-0.3, 0.46), 
   title = "f2", 
   cex = 2, 
   axis_text_size = 5,
@@ -3512,10 +3512,10 @@ projection <- cbind(
 proj_obj3 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.5,
+                            axis_param = list(limits = 0.45,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
                                               threshold = 0.016))
 
 nonlinear_proj3 <- plot_proj(

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 12 e

@@
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
                                               threshold = 0.01))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.8, 0.5), 
   title = "e2", 
@@
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.6, 0.62), 
   title = "e3", 

### 2025-08-24 paper-cardinalR.R

@@ -3397,14 +3397,14 @@ proj_obj2 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
                                               threshold = 0.01))
 
 conicspiral_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.8, 0.5), 
   title = "e2", 
   cex = 2, 
   axis_text_size = 5,
@@ -3424,14 +3424,14 @@ proj_obj3 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.6,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
                                               threshold = 0.016))
 
 conicspiral_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.6, 0.62), 
   title = "e3", 
   cex = 2, 
   axis_text_size = 5,

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 12 c

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.5, 0.5), 
   title = "c3", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 12 d

@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
+                                              axis_scaled = 1.7, 
                                               axis_pos_x = -0.57, 
                                               axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              threshold = 0.05))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.75, 0.5), 
   title = "d2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.5, 0.67), 
   title = "d3", 

### 2025-08-24 paper-cardinalR.R

@@ -3306,15 +3306,15 @@ proj_obj2 <- get_projection(projection = projection,
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
+                                              axis_scaled = 1.7, 
                                               axis_pos_x = -0.57, 
                                               axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              threshold = 0.05))
 
 helicalspiral_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.75, 0.5), 
   title = "d2", 
   cex = 2, 
   axis_text_size = 5,
@@ -3341,7 +3341,7 @@ proj_obj3 <- get_projection(projection = projection,
 helicalspiral_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.5, 0.67), 
   title = "d3", 
   cex = 2, 
   axis_text_size = 5,

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 12 b

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.65, 0.5), 
   title = "b1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.5, 0.5), 
   title = "b3", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 12 a

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.3,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.15, 
+                                              axis_pos_y = -0.15, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.2, 0.3), 
   title = "a1", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.02))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.25, 0.2), 
   title = "a2", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.75,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
+                                              threshold = 0.03))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.6, 0.7), 
   title = "a3", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 11

@@
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.65, 0.5), 
   title = "a3", 

### 2025-08-24 paper-cardinalR.Rmd
Figure 10 a, b

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.8), 
   title = "a1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.7), 
   title = "a2", 
@@
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.6, 0.53), 
   title = "a3", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.7, 0.7), 
   title = "b1", 
@@
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.7, 0.62), 
   title = "b3", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 10 b

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "b1", 
@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
+                                              axis_scaled = 1.5, 
                                               axis_pos_x = -0.57, 
                                               axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              threshold = 0.04))
 

### 2025-08-24 paper-cardinalR.Rmd
Figure 10 c

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "c1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.7), 
   title = "c2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.65, 0.65), 
   title = "c3", 

### 2025-08-24 paper-cardinalR.R

@@ -2482,14 +2482,14 @@ proj_obj1 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
                                               threshold = 0.022))
 
 unifsphere_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "c1", 
   cex = 2, 
   axis_text_size = 5,
@@ -2516,7 +2516,7 @@ proj_obj2 <- get_projection(projection = projection,
 unifsphere_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.7), 
   title = "c2", 
   cex = 2, 
   axis_text_size = 5,
@@ -2536,14 +2536,14 @@ proj_obj3 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.6,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
                                               threshold = 0.016))
 
 unifsphere_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.65, 0.65), 
   title = "c3", 
   cex = 2, 
   axis_text_size = 5,

### 2025-08-24 paper-cardinalR.Rmd
Added projections

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "d1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.65), 
   title = "d2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.65, 0.62), 
   title = "d3", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "e1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "f1", 

### 2025-08-24 paper-cardinalR.R

@@ -2572,14 +2572,14 @@ proj_obj1 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
                                               threshold = 0.022))
 
 gridedsphere_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.75, 0.7), 
   title = "d1", 
   cex = 2, 
   axis_text_size = 5,
@@ -2606,7 +2606,7 @@ proj_obj2 <- get_projection(projection = projection,
 gridedsphere_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.65), 
   title = "d2", 
   cex = 2, 
   axis_text_size = 5,
@@ -2626,14 +2626,14 @@ proj_obj3 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.6,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
                                               threshold = 0.016))
 
 gridedsphere_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.65, 0.62), 
   title = "d3", 
   cex = 2, 
   axis_text_size = 5,
@@ -2664,14 +2664,14 @@ proj_obj1 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
[…]

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 10

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.75, 0.62), 
   title = "f3", 

### 2025-08-24 paper-cardinalR.Rmd
Added Figure 9

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.6, 0.62), 
   title = "a3", 

### 2025-08-24 paper-cardinalR.Rmd
Update Figure 8

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.65, 0.5), 
   title = "c1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.4), 
   title = "c2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.7, 0.62), 
   title = "c3", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.75), 
   title = "d2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.7, 0.62), 
   title = "d3", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 8

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.65, 0.5), 
   title = "a1", 
@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
+                                              threshold = 0.08))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.6, 0.6), 
   title = "a2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.5, 0.75), 
   title = "a3", 

### 2025-08-24 paper-cardinalR.R

@@ -1677,7 +1677,7 @@ proj_obj1 <- get_projection(projection = projection,
 pyrrect_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.65, 0.5), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 5,
@@ -1696,15 +1696,15 @@ proj_obj2 <- get_projection(projection = projection,
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
+                                              threshold = 0.08))
 
 pyrrect_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.6, 0.6), 
   title = "a2", 
   cex = 2, 
   axis_text_size = 5,
@@ -1731,7 +1731,7 @@ proj_obj3 <- get_projection(projection = projection,
 pyrrect_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.5, 0.75), 
   title = "a3", 
   cex = 2, 
   axis_text_size = 5,

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 8 a

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.3,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.25, 0.3), 
   title = "b1", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.03))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.3, 0.4), 
   title = "b2", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 7

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.83, 0.65), 
   title = "a1", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.65, 0.62), 
   title = "a3", 
@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.05))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.75, 0.7), 
   title = "b2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.65, 0.62), 
   title = "b3", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 4 and 5

@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.7, 0.62), 
   title = "a3", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.8,
+                            axis_param = list(limits = 0.5,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
+                                              axis_pos_x = -0.3, 
+                                              axis_pos_y = -0.3, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
+  plot_limits = c(-0.4, 0.4), 
   title = "a1", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
                                               threshold = 0.01))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.5, 0.4), 
   title = "a2", 
@@
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.85, 
-                                              axis_pos_y = -0.85, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
                                               threshold = 0.02))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 0.9), 
+  plot_limits = c(-0.7, 0.7), 
   title = "a3", 

### 2025-08-24 paper-cardinalR.R

@@ -1141,14 +1141,14 @@ proj_obj3 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.6,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
                                               threshold = 0.016))
 
 gau_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.7, 0.62), 
   title = "a3", 
   cex = 2, 
   axis_text_size = 5,
@@ -1198,16 +1198,16 @@ projection <- cbind(
 proj_obj1 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.8,
+                            axis_param = list(limits = 0.5,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
+                                              axis_pos_x = -0.3, 
+                                              axis_pos_y = -0.3, 
                                               threshold = 0.022))
 
 linear_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
+  plot_limits = c(-0.4, 0.4), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 5,
@@ -1225,16 +1225,16 @@ projection <- cbind(
 proj_obj2 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
                                               threshold = 0.01))
 
 linear_proj2 <- plot_proj(
[…]

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 3 b and c

@@
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.9, 0.8), 
   title = "b1", 
@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.04))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.9, 0.8), 
   title = "b2", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.7,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.8, 0.8), 
   title = "b3", 
@@
 ```{r data-cubehole, echo=TRUE}
-cubehole <- gen_cubehole(n = 1000, p = 4)
+cubehole <- gen_cubehole(n = 3000, p = 4)
 ```
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.9, 0.75), 
   title = "c1", 
@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.73, 
+                                              axis_pos_y = -0.73, 
+                                              threshold = 0.04))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.85, 0.7), 
   title = "c2", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.7,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.9, 0.8), 
   title = "c3", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 3

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.8, 
-                                              axis_pos_y = -0.8, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 0.8,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
+                                              threshold = 0.05))
 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 1 e

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.3, 
+                                              axis_pos_y = -0.3, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.4, 0.5), 
   title = "e1", 

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 1

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.35, 0.7), 
   title = "a1", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 0.2,
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.16, 
+                                              axis_pos_y = -0.16, 
+                                              threshold = 0.02))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.2, 0.2), 
   title = "a2", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                            axis_param = list(limits = 0.2,
+                                              axis_scaled = 1.3, 
+                                              axis_pos_x = -0.16, 
+                                              axis_pos_y = -0.16, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.2, 0.2), 
   title = "a3", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.7), 
   title = "b2", 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.65, 0.65), 
   title = "b3", 
@@
                                               axis_scaled = 1.2, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
+                                              threshold = 0.022))
+
+curvybranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.8, 0.7), 
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvybranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.02))
+
+curvybranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.3), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvybranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
                                               axis_pos_x = -0.5, 
                                               axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-curvybranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvybranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-curvybranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvybranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-0.6, 0.4), 
   title = "c3", 

### 2025-08-24 paper-cardinalR.R

@@ -183,16 +183,16 @@ projection <- cbind(
 proj_obj1 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
                                               threshold = 0.022))
 
 expbranches_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.35, 0.7), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 5,
@@ -210,16 +210,16 @@ projection <- cbind(
 proj_obj2 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 0.2,
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.16, 
+                                              axis_pos_y = -0.16, 
+                                              threshold = 0.02))
 
 expbranches_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.2, 0.2), 
   title = "a2", 
   cex = 2, 
   axis_text_size = 5,
@@ -237,16 +237,16 @@ projection <- cbind(
 proj_obj3 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
[…]

### 2025-08-24 paper-cardinalR.Rmd
Updated Figure 5

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.5,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.4, 0.4), 
+  plot_limits = c(-0.65, 0.5), 
   title = "a1", 
@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.75,
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
                                               threshold = 0.01))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.4), 
+  plot_limits = c(-0.9, 0.6), 
   title = "a2", 

### 2025-08-24 paper-cardinalR.R

@@ -1198,16 +1198,16 @@ projection <- cbind(
 proj_obj1 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.5,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
                                               threshold = 0.022))
 
 linear_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.4, 0.4), 
+  plot_limits = c(-0.65, 0.5), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 5,
@@ -1225,16 +1225,16 @@ projection <- cbind(
 proj_obj2 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.75,
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
                                               threshold = 0.01))
 
 linear_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.4), 
+  plot_limits = c(-0.9, 0.6), 
   title = "a2", 
   cex = 2, 
   axis_text_size = 5,

### 2025-08-24 paper-cardinalR.Rmd
Added main reference

@@
 
-A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
+A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
@@
 
-To simulate a cone-shaped structure in arbitrary dimensions, we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+To simulate a cone-shaped structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cone)"} else { "Figure \\@ref(fig:cone-proj)"})`) in arbitrary dimensions, we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
 
@@
 
-A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+A cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube)"} else { "Figure \\@ref(fig:cube-proj)"})`) represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
 We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
@@
 
-The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure.
+The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:gau)"} else { "Figure \\@ref(fig:gau-proj)"})`).
 Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, \Sigma)$, where $\Sigma$ is a user-defined $p \times p$ positive-definite matrix.
@@
 
-The `gen_longlinear(n, p)` function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise.
+The `gen_longlinear(n, p)` function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:linear)"} else { "Figure \\@ref(fig:linear-proj)"})`).
 Each variable $X_i$ is created as $X_i = \text{scale}_i \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i$, where $\text{scale}_i \sim U(-10, 10)$ is randomly chosen direction and stretch, $\text{shift}_i \sim U(-300, 300)$ is large offset to spread features apart, and $\epsilon \sim N(0, 0.03n)$ is Gaussian noise.
@@
 
-The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$ structure.
+The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$ structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:mobius)"} else { "Figure \\@ref(fig:mobius-proj)"})`).
 This classical non-orientable surface loops back on itself with a half-twist.
@@
 
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in high-dimensional space.
+A polynomial structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly)"} else { "Figure \\@ref(fig:poly-proj)"})`) generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in high-dimensional space.
 These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
@@
 
-A pyramid structure represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
+A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
 The functions provided allow users to generate pyramids with various base shapes—including rectangular, triangular, and star-shaped bases—with options for sharp or blunted apexes.
@@
 
-An S-curve structure simulates data that lies along a smooth, non-linear manifold, making it a classic example for testing the performance of non-linear dimension reduction methods.
+An S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) simulates data that lies along a smooth, non-linear manifold, making it a classic example for testing the performance of non-linear dimension reduction methods.
 This package provides functions to generate standard S-curve shapes as well as variants containing holes to introduce missing or incomplete regions.
@@
 
-Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+Sphere-shaped structures (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere)"} else { "Figure \\@ref(fig:sphere-proj)"})`) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
 The package provides functions to generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, p, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension.
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, p, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:swissroll)"} else { "Figure \\@ref(fig:swissroll-proj)"})`).
 The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.
@@
 
-Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems.
+Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric)"} else { "Figure \\@ref(fig:triginometric-proj)"})`).
 The package includes functions to generate a wide range of these forms from crescents and curvy cylinders to spherical, helical, and conic spirals, as well as nonlinear hyperbolas.
@@
 
-The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space.
+The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil)"} else { "Figure \\@ref(fig:trefoil-proj)"})`).
 The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of NLDR methods to preserve global structure, loops, and embeddings in high-dimensional data.

### 2025-08-24 paper-cardinalR.tex
Updated Figure 3 b and c
Added main reference

@@
 
-To simulate a cone-shaped structure in arbitrary dimensions, we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+To simulate a cone-shaped structure (Figure \ref{fig:cone-proj}) in arbitrary dimensions, we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
 
@@
 
-A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+A cube structure (Figure \ref{fig:cube-proj}) represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
 We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
@@
 \begin{verbatim}
-cubehole <- gen_cubehole(n = 1000, p = 4)
+cubehole <- gen_cubehole(n = 3000, p = 4)
 \end{verbatim}
@@
 
-The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure.
+The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure (Figure \ref{fig:gau-proj}).
 Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, \Sigma)\), where \(\Sigma\) is a user-defined \(p \times p\) positive-definite matrix.
@@
 
-The \texttt{gen\_longlinear(n,\ p)} function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise.
+The \texttt{gen\_longlinear(n,\ p)} function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise (Figure \ref{fig:linear-proj}).
 Each variable \(X_i\) is created as \(X_i = \text{scale}_i \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i\), where \(\text{scale}_i \sim U(-10, 10)\) is randomly chosen direction and stretch, \(\text{shift}_i \sim U(-300, 300)\) is large offset to spread features apart, and \(\epsilon \sim N(0, 0.03n)\) is Gaussian noise.
@@
 
-The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\) structure.
+The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\) structure (Figure \ref{fig:mobius-proj}).
 This classical non-orientable surface loops back on itself with a half-twist.
@@
 
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in high-dimensional space.
+A polynomial structure (Figure \ref{fig:poly-proj}) generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in high-dimensional space.
 These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
@@
 
-A pyramid structure represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
+A pyramid structure (Figure \ref{fig:pyr-proj}) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
 The functions provided allow users to generate pyramids with various base shapes---including rectangular, triangular, and star-shaped bases---with options for sharp or blunted apexes.
@@
 
-An S-curve structure simulates data that lies along a smooth, non-linear manifold, making it a classic example for testing the performance of non-linear dimension reduction methods.
+An S-curve structure (Figure \ref{fig:scurve-proj}) simulates data that lies along a smooth, non-linear manifold, making it a classic example for testing the performance of non-linear dimension reduction methods.
 This package provides functions to generate standard S-curve shapes as well as variants containing holes to introduce missing or incomplete regions.
@@
 
-Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+Sphere-shaped structures (Figure \ref{fig:sphere-proj}) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
 The package provides functions to generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ p,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension.
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ p,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension (Figure \ref{fig:swissroll-proj}).
 The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.
@@
 
-Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems.
+Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (Figure \ref{fig:triginometric-proj}).
 The package includes functions to generate a wide range of these forms from crescents and curvy cylinders to spherical, helical, and conic spirals, as well as nonlinear hyperbolas.
@@
 
-The Trefoil is a closed, nontrivial one-dimensional manifold embedded in \(3\text{-}D\) or \(4\text{-}D\) space.
+The Trefoil is a closed, nontrivial one-dimensional manifold embedded in \(3\text{-}D\) or \(4\text{-}D\) space (Figure \ref{fig:trefoil-proj}).
 The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of NLDR methods to preserve global structure, loops, and embeddings in high-dimensional data.

### 2025-08-24 paper-cardinalR.Rmd
Updated cross reference in sub sections

@@
 
-The `gen_expbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ exponential branches in $2\text{-}D$, with optional noise dimensions to embed the structure in $p\text{-}D$.
+The `gen_expbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ exponential branches in $2\text{-}D$, with optional noise dimensions to embed the structure in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) a"} else { "Figure \\@ref(fig:branch-proj) a"})`).
 These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
@@
 
-The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches in $p\text{-}D$.
+The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) b"} else { "Figure \\@ref(fig:branch-proj) b"})`).
 The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.
@@
 
-The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in $p\text{-}D$.
+The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) c"} else { "Figure \\@ref(fig:branch-proj) c"})`).
 The underlying geometry lies in the first two dimensions, while the remaining $(p-2)$ dimensions contain Gaussian noise.
@@
 
-The `gen_orglinearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches embedded in $p\text{-}D$.
+The `gen_orglinearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches embedded in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) d"} else { "Figure \\@ref(fig:branch-proj) d"})`).
 Each branch lies primarily within a distinct $2\text{-}D$ subspace, while the remaining $p - 2$ dimensions contain Gaussian noise.
@@
 
-The `gen_orgcurvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in a $p$-dimensional space.
+The `gen_orgcurvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in a $p$-dimensional space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) e"} else { "Figure \\@ref(fig:branch-proj) e"})`).
 Each branch is constructed in a unique or repeated $2\text{-}D$ subspace of the $p$-dimensional space, with curvature induced by a second-degree polynomial structure.
@@
 
-The `gen_gridcube(n, p)` constructs a regular grid using approximately $n$ points by evenly partitioning each of the $p$ dimensions.
+The `gen_gridcube(n, p)` constructs a regular grid using approximately $n$ points by evenly partitioning each of the $p$ dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) a"} else { "Figure \\@ref(fig:cube-proj) a"})`).
 Specifically, the number of grid points along each axis is determined by `gen_nproduct(n, p)`, which calculates the nearest balanced integer factors whose product is close to $n$.
@@
 
-The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in $p\text{-}D$ with additional noise dimensions.
+The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in $p\text{-}D$ with additional noise dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) b"} else { "Figure \\@ref(fig:cube-proj) b"})`).
 
@@
 
-The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p\text{-}D$.
+The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) c"} else { "Figure \\@ref(fig:cube-proj) c"})`).
 
@@
 
-The `gen_quadratic(n, p, range)` function generates a dataset of $n$ points forming a quadratic curve in the first two dimensions.
+The `gen_quadratic(n, p, range)` function generates a dataset of $n$ points forming a quadratic curve in the first two dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) a"} else { "Figure \\@ref(fig:poly-proj) a"})`).
 This $2\text{-}D$ parabolic structure is embedded within a higher-dimensional space with additive noise in the remaining dimensions.
@@
 
-The `gen_cubic(n, p, range)` function generates a dataset of $n$ points forming a cubic curve in the first two dimensions.
+The `gen_cubic(n, p, range)` function generates a dataset of $n$ points forming a cubic curve in the first two dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) b"} else { "Figure \\@ref(fig:poly-proj) b"})`).
 This function creates a more complex curvilinear structure than a simple parabola.
@@
 
-The `gen_pyrrect(n, p, h, l_vec, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a rectangular base.
+The `gen_pyrrect(n, p, h, l_vec, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a rectangular base (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`).
 The pyramid is embedded in $p\text{-}D$, with a tip at height zero and base at height $h$.
@@
 
-The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section.
+The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`).
 The structure is embedded in a $p\text{-}D$, with the tip located at height $0$ and the base at height $h$.
@@
 
-The `gen_pyrstar(n, p, h, rb)` function generates a dataset of $n$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
+The `gen_pyrstar(n, p, h, rb)` function generates a dataset of $n$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`).
 The pyramid extends vertically from height $z = 0$ (tip) to height $z = h$ (base), embedded in a $p$-dimensional space.
@@
 
-The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p\text{-}D$ simplex using a chaotic attractor-like midpoint algorithm.
+The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p\text{-}D$ simplex using a chaotic attractor-like midpoint algorithm (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`).
 The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a "Sierpinski-like pyramid" in high dimensions.
@@
 
-To simulate an S-curve structure in a higher-dimensional space, we define the function `gen_scurve(n, p)`, which generates $n$ observations in $p\text{-}D$.
+To simulate an S-curve structure in a higher-dimensional space, we define the function `gen_scurve(n, p)`, which generates $n$ observations in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`).
 
@@
 
-To simulate a variant of the S-curve structure with a removed region ("hole"), we define the function `gen_scurvehole(n, p)`.
+To simulate a variant of the S-curve structure with a removed region ("hole"), we define the function `gen_scurvehole(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) b"} else { "Figure \\@ref(fig:scurve-proj) b"})`).
 This function produces $n$ observations in $p\text{-}D$ where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
@@
 
-The function `gen_circle(n, p)` generates a $p$-dimensional dataset of $n$ observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale.
+The function `gen_circle(n, p)` generates a $p$-dimensional dataset of $n$ observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
 
@@
 
-The `gen_curvycycle(n, p)` function generates a $p$-dimensional dataset of $n$ observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations.
+The `gen_curvycycle(n, p)` function generates a $p$-dimensional dataset of $n$ observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`).
 The structure forms a one-dimensional nonlinear cycle embedded in higher-dimensional space.
@@
 
-The `gen_unifsphere(n, p, r)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on the surface of a $3\text{-}D$ sphere of radius $r$, with additional Gaussian noise dimensions added when $p > 3$.
+The `gen_unifsphere(n, p, r)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on the surface of a $3\text{-}D$ sphere of radius $r$, with additional Gaussian noise dimensions added when $p > 3$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) c"} else { "Figure \\@ref(fig:sphere-proj) c"})`).
 
@@
 
-The `gen_gridedsphere(n, p)` function generates a $p$-dimensional dataset of approximately $n$ points evenly distributed on the surface of a $(p!-!1)$-sphere (the unit sphere embedded in $\mathbb{R}^p$).
+The `gen_gridedsphere(n, p)` function generates a $p$-dimensional dataset of approximately $n$ points evenly distributed on the surface of a $(p!-!1)$-sphere (the unit sphere embedded in $\mathbb{R}^p$) (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
 
@@
 
-The `gen_clusteredspheres(n, k, p, r, loc)` function generates a synthetic dataset of $n_1 + kn_2$ observations in $p$-dimensional space, consisting of one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location.
+The `gen_clusteredspheres(n, k, p, r, loc)` function generates a synthetic dataset of $n_1 + kn_2$ observations in $p$-dimensional space, consisting of one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
 
@@
 
-The `gen_hemisphere(n, p)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on a $4\text{-}D$ hemisphere, optionally extended with Gaussian noise in additional dimensions when $p > 4$.
+The `gen_hemisphere(n, p)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on a $4\text{-}D$ hemisphere, optionally extended with Gaussian noise in additional dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) f"} else { "Figure \\@ref(fig:sphere-proj) f"})`).
 
@@
 
-The `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise.
+The `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) a"} else { "Figure \\@ref(fig:triginometric-proj) a"})`).
 
@@
 
-The `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$.
+The `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
 
@@
 The `gen_sphericalspiral(n, p, spins)` function simulates a dataset of $n$ observations that form a spiraling path on a spherical surface in the first four dimensions.
-When extended beyond $4\text{-}D$, structured nonlinear noise dimensions are added to simulate more realistic high-dimensional manifolds.
+When extended beyond $4\text{-}D$, structured nonlinear noise dimensions are added to simulate more realistic high-dimensional manifolds (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`).
 
@@
 The `gen_helicalspiral(n, p)` function generates a dataset of $n$ observations forming a helical spiral, embedded in the first four dimensions of $p\text{-}D$.
-Additional dimensions (if $p > 4$) are filled with structured noise to mimic high-dimensional complexity.
+Additional dimensions (if $p > 4$) are filled with structured noise to mimic high-dimensional complexity (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) d"} else { "Figure \\@ref(fig:triginometric-proj) d"})`).
 
@@
 
-The `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$.
+The `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) e"} else { "Figure \\@ref(fig:triginometric-proj) e"})`).
 The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a $3\text{-}D$ conic helix.
@@
 
-The `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables.
+The `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
 
@@
 The function `gen_trefoil4d()` generates a $4\text{-}D$ trefoil knot using two angular parameters, $\theta$ and $\phi$.
-The knot is constructed on the $3$-sphere ($S^3 \subset \mathbb{R}^4$), with a band of thickness around the knot path, controlled by `steps`.
+The knot is constructed on the $3$-sphere ($S^3 \subset \mathbb{R}^4$), with a band of thickness around the knot path, controlled by `steps` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
 
@@
 
-The function `gen_trefoil3d()` generates a $3\text{-}D$ stereographic projection of the $4\text{-}D$ trefoil knot by mapping each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3$ using the transformation $X_1' = X_1 / (1 - X_4)$, $X_2' = X_2 / (1 - X_4)$, and $X_3' = X_3 / (1 - X_4)$, while excluding points where $X_4 = 1$ to avoid division by zero.
+The function `gen_trefoil3d()` generates a $3\text{-}D$ stereographic projection of the $4\text{-}D$ trefoil knot by mapping each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3$ using the transformation $X_1' = X_1 / (1 - X_4)$, $X_2' = X_2 / (1 - X_4)$, and $X_3' = X_3 / (1 - X_4)$, while excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
 

### 2025-08-24 paper-cardinalR.tex
Updated cross reference in sub sections

@@
 
-The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in \(2\text{-}D\), with optional noise dimensions to embed the structure in \(p\text{-}D\).
+The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in \(2\text{-}D\), with optional noise dimensions to embed the structure in \(p\text{-}D\) (Figure \ref{fig:branch-proj} a).
 These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
@@
 
-The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches in \(p\text{-}D\).
+The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches in \(p\text{-}D\) (Figure \ref{fig:branch-proj} b).
 The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.
@@
 
-The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in \(p\text{-}D\).
+The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in \(p\text{-}D\) (Figure \ref{fig:branch-proj} c).
 The underlying geometry lies in the first two dimensions, while the remaining \((p-2)\) dimensions contain Gaussian noise.
@@
 
-The \texttt{gen\_orglinearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches embedded in \(p\text{-}D\).
+The \texttt{gen\_orglinearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches embedded in \(p\text{-}D\) (Figure \ref{fig:branch-proj} d).
 Each branch lies primarily within a distinct \(2\text{-}D\) subspace, while the remaining \(p - 2\) dimensions contain Gaussian noise.
@@
 
-The \texttt{gen\_orgcurvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in a \(p\)-dimensional space.
+The \texttt{gen\_orgcurvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in a \(p\)-dimensional space (Figure \ref{fig:branch-proj} e).
 Each branch is constructed in a unique or repeated \(2\text{-}D\) subspace of the \(p\)-dimensional space, with curvature induced by a second-degree polynomial structure.
@@
 
-The \texttt{gen\_gridcube(n,\ p)} constructs a regular grid using approximately \(n\) points by evenly partitioning each of the \(p\) dimensions.
+The \texttt{gen\_gridcube(n,\ p)} constructs a regular grid using approximately \(n\) points by evenly partitioning each of the \(p\) dimensions (Figure \ref{fig:cube-proj} a).
 Specifically, the number of grid points along each axis is determined by \texttt{gen\_nproduct(n,\ p)}, which calculates the nearest balanced integer factors whose product is close to \(n\).
@@
 
-The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in \(p\text{-}D\) with additional noise dimensions.
+The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in \(p\text{-}D\) with additional noise dimensions (Figure \ref{fig:cube-proj} b).
 
@@
 
-The \texttt{gen\_cubehole(n,\ p)} function generates a dataset of \(n\) points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in \(p\text{-}D\).
+The \texttt{gen\_cubehole(n,\ p)} function generates a dataset of \(n\) points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in \(p\text{-}D\) (Figure \ref{fig:cube-proj} c).
 
@@
 
-The \texttt{gen\_quadratic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a quadratic curve in the first two dimensions.
+The \texttt{gen\_quadratic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a quadratic curve in the first two dimensions (Figure \ref{fig:poly-proj} a).
 This \(2\text{-}D\) parabolic structure is embedded within a higher-dimensional space with additive noise in the remaining dimensions.
@@
 
-The \texttt{gen\_cubic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a cubic curve in the first two dimensions.
+The \texttt{gen\_cubic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a cubic curve in the first two dimensions (Figure \ref{fig:poly-proj} b).
 This function creates a more complex curvilinear structure than a simple parabola.
@@
 
-The \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a rectangular base.
+The \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a rectangular base (Figure \ref{fig:pyr-proj} a).
 The pyramid is embedded in \(p\text{-}D\), with a tip at height zero and base at height \(h\).
@@
 
-The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section.
+The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section (Figure \ref{fig:pyr-proj} b).
 The structure is embedded in a \(p\text{-}D\), with the tip located at height \(0\) and the base at height \(h\).
@@
 
-The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-shaped structure with a hexagonal base.
+The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-shaped structure with a hexagonal base (Figure \ref{fig:pyr-proj} c).
 The pyramid extends vertically from height \(z = 0\) (tip) to height \(z = h\) (base), embedded in a \(p\)-dimensional space.
@@
 
-The \texttt{gen\_pyrholes(n,\ p)} function generates \(n\) points embedded in a \(p\text{-}D\) simplex using a chaotic attractor-like midpoint algorithm.
+The \texttt{gen\_pyrholes(n,\ p)} function generates \(n\) points embedded in a \(p\text{-}D\) simplex using a chaotic attractor-like midpoint algorithm (Figure \ref{fig:pyr-proj} d).
 The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a ``Sierpinski-like pyramid'' in high dimensions.
@@
 
-To simulate an S-curve structure in a higher-dimensional space, we define the function \texttt{gen\_scurve(n,\ p)}, which generates \(n\) observations in \(p\text{-}D\).
+To simulate an S-curve structure in a higher-dimensional space, we define the function \texttt{gen\_scurve(n,\ p)}, which generates \(n\) observations in \(p\text{-}D\) (Figure \ref{fig:scurve-proj} a).
 
@@
 
-To simulate a variant of the S-curve structure with a removed region (``hole''), we define the function \texttt{gen\_scurvehole(n,\ p)}.
+To simulate a variant of the S-curve structure with a removed region (``hole''), we define the function \texttt{gen\_scurvehole(n,\ p)} (Figure \ref{fig:scurve-proj} b).
 This function produces \(n\) observations in \(p\text{-}D\) where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
@@
 
-The function \texttt{gen\_circle(n,\ p)} generates a \(p\)-dimensional dataset of \(n\) observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale.
+The function \texttt{gen\_circle(n,\ p)} generates a \(p\)-dimensional dataset of \(n\) observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale (Figure \ref{fig:sphere-proj} a).
 
@@
 
-The \texttt{gen\_curvycycle(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations.
+The \texttt{gen\_curvycycle(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations (Figure \ref{fig:sphere-proj} b).
 The structure forms a one-dimensional nonlinear cycle embedded in higher-dimensional space.
@@
 
-The \texttt{gen\_unifsphere(n,\ p,\ r)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on the surface of a \(3\text{-}D\) sphere of radius \(r\), with additional Gaussian noise dimensions added when \(p > 3\).
+The \texttt{gen\_unifsphere(n,\ p,\ r)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on the surface of a \(3\text{-}D\) sphere of radius \(r\), with additional Gaussian noise dimensions added when \(p > 3\) (Figure \ref{fig:sphere-proj} c).
 
@@
 
-The \texttt{gen\_gridedsphere(n,\ p)} function generates a \(p\)-dimensional dataset of approximately \(n\) points evenly distributed on the surface of a \((p!-!1)\)-sphere (the unit sphere embedded in \(\mathbb{R}^p\)).
+The \texttt{gen\_gridedsphere(n,\ p)} function generates a \(p\)-dimensional dataset of approximately \(n\) points evenly distributed on the surface of a \((p!-!1)\)-sphere (the unit sphere embedded in \(\mathbb{R}^p\)) (Figure \ref{fig:sphere-proj} d).
 
@@
 
-The \texttt{gen\_clusteredspheres(n,\ k,\ p,\ r,\ loc)} function generates a synthetic dataset of \(n_1 + kn_2\) observations in \(p\)-dimensional space, consisting of one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location.
+The \texttt{gen\_clusteredspheres(n,\ k,\ p,\ r,\ loc)} function generates a synthetic dataset of \(n_1 + kn_2\) observations in \(p\)-dimensional space, consisting of one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location (Figure \ref{fig:sphere-proj} e).
 
@@
 
-The \texttt{gen\_hemisphere(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on a \(4\text{-}D\) hemisphere, optionally extended with Gaussian noise in additional dimensions when \(p > 4\).
+The \texttt{gen\_hemisphere(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on a \(4\text{-}D\) hemisphere, optionally extended with Gaussian noise in additional dimensions when \(p > 4\) (Figure \ref{fig:sphere-proj} f).
 
@@
 
-The \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise.
+The \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise (Figure \ref{fig:triginometric-proj} a).
 
@@
 
-The \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\).
+The \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
 
@@
 The \texttt{gen\_sphericalspiral(n,\ p,\ spins)} function simulates a dataset of \(n\) observations that form a spiraling path on a spherical surface in the first four dimensions.
-When extended beyond \(4\text{-}D\), structured nonlinear noise dimensions are added to simulate more realistic high-dimensional manifolds.
+When extended beyond \(4\text{-}D\), structured nonlinear noise dimensions are added to simulate more realistic high-dimensional manifolds (Figure \ref{fig:triginometric-proj} c).
 
@@
 The \texttt{gen\_helicalspiral(n,\ p)} function generates a dataset of \(n\) observations forming a helical spiral, embedded in the first four dimensions of \(p\text{-}D\).
-Additional dimensions (if \(p > 4\)) are filled with structured noise to mimic high-dimensional complexity.
+Additional dimensions (if \(p > 4\)) are filled with structured noise to mimic high-dimensional complexity (Figure \ref{fig:triginometric-proj} d).
 
@@
 
-The \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of \(p\text{-}D\).
+The \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of \(p\text{-}D\) (Figure \ref{fig:triginometric-proj} e).
 The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a \(3\text{-}D\) conic helix.
@@
 
-The \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables.
+The \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
 
@@
 The function \texttt{gen\_trefoil4d()} generates a \(4\text{-}D\) trefoil knot using two angular parameters, \(\theta\) and \(\phi\).
-The knot is constructed on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)), with a band of thickness around the knot path, controlled by \texttt{steps}.
+The knot is constructed on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)), with a band of thickness around the knot path, controlled by \texttt{steps} (Figure \ref{fig:trefoil-proj} a).
 
@@
 
-The function \texttt{gen\_trefoil3d()} generates a \(3\text{-}D\) stereographic projection of the \(4\text{-}D\) trefoil knot by mapping each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\) using the transformation \(X_1' = X_1 / (1 - X_4)\), \(X_2' = X_2 / (1 - X_4)\), and \(X_3' = X_3 / (1 - X_4)\), while excluding points where \(X_4 = 1\) to avoid division by zero.
+The function \texttt{gen\_trefoil3d()} generates a \(3\text{-}D\) stereographic projection of the \(4\text{-}D\) trefoil knot by mapping each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\) using the transformation \(X_1' = X_1 / (1 - X_4)\), \(X_2' = X_2 / (1 - X_4)\), and \(X_3' = X_3 / (1 - X_4)\), while excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
 

### 2025-09-05 paper-cardinalR.Rmd
ML changes up to linear branching section

@@
 
-There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.
-For example, `geozoo` (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
+There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The `geozoo` package (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
 Another useful package is `snedata` (@james2025), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
 Additionally, `mlbench` (@friedrich2024) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
-In the field of single-cell omics, `splatter` (@luke2017) is particularly simulate complex biological data, effectively capturing nuances such as batch effects and differential expression.
-
-There is a valuable opportunity to improve the generation of high-dimensional data structures by integrating geometric principles with advanced noise control and customizable clustering.
+In the field of single-cell omics, `splatter` (@luke2017) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+
+There is an  opportunity to improve the generation of high-dimensional data structures by integrating geometric principles with advanced noise control and customizable clustering.
 The `geozoo` package provides a strong foundation but could be enhanced to support high-dimensional extensions with controlled noise and user-defined parameters for clustering.
@@
 This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes.
-`cardinalR` offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) create complex clustered data arrangements by using fundamental geometric forms, while maintaining their positions, scales, orientations, and sample sizes in arbitrary dimensional spaces.
-By providing these integrated features, `cardinalR` aims to provide researchers to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
+`cardinalR` offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
+By providing these features, `cardinalR` provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
 This bridges the gap between geometric foundations and the flexible generation of complex synthetic data.
@@
 
-A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
+A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology.
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
@@
 The sign alternates between branches (odd: negative, even: positive exponent).
-This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches: decay pattern (reflected) and even-numbered branches: growth pattern.
+This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches represent a decay pattern (reflected) and even-numbered branches represent a growth pattern.
 For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
@@
 Branches $1$ and $2$ are initialized with fixed slopes and intercepts.
-The Branch $1$ is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$.
+Branch $1$ is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$.
 The Branch $2$ is generated from $X_1 \sim U(-6, 2)$, $X_2 = -0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$.

### 2025-09-05 paper-cardinalR.tex
Rendered after Michael's changes

@@
 
-There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.
-For example, \texttt{geozoo} (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
+There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
 Another useful package is \texttt{snedata} (\citet{james2025}), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
 Additionally, \texttt{mlbench} (\citet{friedrich2024}) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
-In the field of single-cell omics, \texttt{splatter} (\citet{luke2017}) is particularly simulate complex biological data, effectively capturing nuances such as batch effects and differential expression.
-
-There is a valuable opportunity to improve the generation of high-dimensional data structures by integrating geometric principles with advanced noise control and customizable clustering.
+In the field of single-cell omics, \texttt{splatter} (\citet{luke2017}) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+
+There is an opportunity to improve the generation of high-dimensional data structures by integrating geometric principles with advanced noise control and customizable clustering.
 The \texttt{geozoo} package provides a strong foundation but could be enhanced to support high-dimensional extensions with controlled noise and user-defined parameters for clustering.
@@
 This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes.
-\texttt{cardinalR} offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) create complex clustered data arrangements by using fundamental geometric forms, while maintaining their positions, scales, orientations, and sample sizes in arbitrary dimensional spaces.
-By providing these integrated features, \texttt{cardinalR} aims to provide researchers to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
+\texttt{cardinalR} offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
+By providing these features, \texttt{cardinalR} provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
 This bridges the gap between geometric foundations and the flexible generation of complex synthetic data.
@@
 
-A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar processes such as cell differentiation in biology.
+A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology.
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
@@
 The sign alternates between branches (odd: negative, even: positive exponent).
-This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches: decay pattern (reflected) and even-numbered branches: growth pattern.
+This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches represent a decay pattern (reflected) and even-numbered branches represent a growth pattern.
 For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
@@
 Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts.
-The Branch \(1\) is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\).
+Branch \(1\) is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\).
 The Branch \(2\) is generated from \(X_1 \sim U(-6, 2)\), \(X_2 = -0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\).

### 2025-09-05 paper-cardinalR.Rmd
Updated Table 1

@@
                           "is_bkg"),
-                        exp = c("A numeric vector representing the number of points in each cluster.",
-                                "A numeric value representing the number of dimensions.",
-                                "A numeric value representing the number of clusters.",
-                                "A numeric matrix representing the locations/centroids of clusters.",
-                                "A numeric vector representing the scaling factors of clusters.",
-                                "A character vector representing the shapes of clusters.",
-                                "A numeric list which contains plane and the corresponding angle along that plane for each cluster.",
-                                "A Boolean value representing the background noise should exist or not."))
+                  type = c("numeric (vector)",
+                           "numeric",
+                           "numeric",
+                           "numeric (matrix)",
+                           "numeric (vector)",
+                           "character (vector)",
+                           "numeric (list)",
+                           "boolean"),
+                        exp = c("Number of points in each cluster.",
+                                "Number of dimensions.",
+                                "Number of clusters.",
+                                "Locations/centroids of clusters.",
+                                "Scaling factors of clusters.",
+                                "Shapes of clusters.",
+                                "Plane and the corresponding angle along that plane for each cluster.",
+                                "Background noise should exist or not."))
 ```
@@
 main_tb |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Explanation")) 
+  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation")) 
 ```
@@
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
+  column_spec(2, width = "4cm") |>
+  column_spec(3, width = "8cm")
 ```

### 2025-09-05 paper-cardinalR.tex
Updated Table 1

@@
 \centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Argument & Explanation\\
-\midrule
-n & A numeric vector representing the number of points in each cluster.\\
-p & A numeric value representing the number of dimensions.\\
-k & A numeric value representing the number of clusters.\\
-loc & A numeric matrix representing the locations/centroids of clusters.\\
-scale & A numeric vector representing the scaling factors of clusters.\\
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Argument & Type & Explanation\\
+\midrule
+n & numeric (vector) & Number of points in each cluster.\\
+p & numeric & Number of dimensions.\\
+k & numeric & Number of clusters.\\
+loc & numeric (matrix) & Locations/centroids of clusters.\\
+scale & numeric (vector) & Scaling factors of clusters.\\
 \addlinespace
-shape & A character vector representing the shapes of clusters.\\
-rotation & A numeric list which contains plane and the corresponding angle along that plane for each cluster.\\
-is\_bkg & A Boolean value representing the background noise should exist or not.\\
+shape & character (vector) & Shapes of clusters.\\
+rotation & numeric (list) & Plane and the corresponding angle along that plane for each cluster.\\
+is\_bkg & boolean & Background noise should exist or not.\\
 \bottomrule

### 2025-09-05 paper-cardinalR.R

@@ -95,27 +95,36 @@ main_tb <- tibble(arg = c("n",
                           "scale",
                           "shape",
                           "rotation",
-                          "is_bkg"), 
-                        exp = c("A numeric vector representing the number of points in each cluster.",
-                                "A numeric value representing the number of dimensions.",
-                                "A numeric value representing the number of clusters.",
-                                "A numeric matrix representing the locations/centroids of clusters.",
-                                "A numeric vector representing the scaling factors of clusters.",
-                                "A character vector representing the shapes of clusters.",
-                                "A numeric list which contains plane and the corresponding angle along that plane for each cluster.",
-                                "A Boolean value representing the background noise should exist or not."))
+                          "is_bkg"),
+                  type = c("numeric (vector)",
+                           "numeric",
+                           "numeric",
+                           "numeric (matrix)",
+                           "numeric (vector)",
+                           "character (vector)",
+                           "numeric (list)",
+                           "boolean"),
+                        exp = c("Number of points in each cluster.",
+                                "Number of dimensions.",
+                                "Number of clusters.",
+                                "Locations/centroids of clusters.",
+                                "Scaling factors of clusters.",
+                                "Shapes of clusters.",
+                                "Plane and the corresponding angle along that plane for each cluster.",
+                                "Background noise should exist or not."))
 
 
 ## ----main-tb-html, eval=knitr::is_html_output()-------------------------------
 # main_tb |>
-#   kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Explanation"))
+#   kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"))
 
 
 ## ----main-tb-pdf, eval=knitr::is_latex_output()-------------------------------
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
+  column_spec(2, width = "4cm") |>
+  column_spec(3, width = "8cm")
 
 
 ## -----------------------------------------------------------------------------

### 2025-09-07 paper-cardinalR.Rmd
Changes up to cubehole

@@
 The underlying geometry lies in the first two dimensions, while the remaining $(p-2)$ dimensions contain Gaussian noise.
-
 Branch $1$ is generated from $X_1 \sim U(0, 1)$, and $X_2 = 0.1X_1 + X_1^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
@@
 Branches $3$ to $k$ are added iteratively.
-Each new branch $i$ begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X_{1}$:$X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure: $X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}$.
+Each new branch $i$ begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ to ensure connectivity with earlier branches, spans a unit-length interval on $X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure: $X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, \; s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}$.
 Here, $s_i$ is a sampled scale factor (slope-like term) controlling curvature, and $(x_{\text{start}}, y_{\text{start}})$ is the starting point from the existing structure.
@@
 The `gen_orglinearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches embedded in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) d"} else { "Figure \\@ref(fig:branch-proj) d"})`).
-Each branch lies primarily within a distinct $2\text{-}D$ subspace, while the remaining $p - 2$ dimensions contain Gaussian noise.
-
-To construct each branch, a unique or repeated pair of dimensions is selected from the $\binom{p}{2}$ possible $2\text{-}D$ combinations.
+Each branch lies within a distinct $2\text{-}D$ subspace, while the remaining $p - 2$ dimensions contain Gaussian noise.
+
+To construct each branch, a unique or repeated pair of dimensions is selected from the $\binom{p}{2}$ possible combinations of dimensions.
 If $k \leq \binom{p}{2}$, combinations are sampled without replacement.
 If $k > \binom{p}{2}$, additional pairs are sampled with replacement to reach $k$ total branches.
-Each selected pair $(X_{i1}, X_{i2})$ defines the $2\text{-}D$ plane for branch $i$.
+Each selected pair $(i_1, i_2)$ defines the $2\text{-}D$ plane for branch $i$.
 
 For branch $i$, $n_i$ points are generated, where $\sum_{i=1}^k n_i = n$.
-The structure follows: $X_{i1} \sim U(0, 2)$, and $X_{i2} = -s_iX_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5)$, where $s_i$ is a scale factor controlling the slope.
+The structure follows: $X_{i_1} \sim U(0, 2)$, and $X_{i_2} = -s_iX_{i_1} + \epsilon, \quad \epsilon \sim U(0, 0.5)$, where $s_i$ is a scale factor controlling the slope.
 When $k \leq \binom{p}{2}$, $s_i = 1$.
 When sampling with replacement, $s_i$ is drawn from the set ${1, 1.5, 2, \dots, 8}$ in increments of $0.5$.
-For $p > 2$, the remaining dimensions contain independent Gaussian noise: $X_j \sim N(0, 0.1^2)$, for $j \notin \{i1, i2\}$.
+For $p > 2$, the remaining dimensions contain independent Gaussian noise: $X_j \sim N(0, 0.1^2)$, for $j \notin \{i_1, i_2\}$.
 
@@
 Let $\binom{p}{2}$ denote the number of unique $2\text{-}D$ subspace combinations.
-When $k \leq \binom{p}{2}$, $k$ distinct subspace pairs $(X_{i1}, X_{i2})$ are sampled without replacement.
+When $k \leq \binom{p}{2}$, $k$ distinct subspace pairs $(i_1, i_2)$ are sampled without replacement.
 Otherwise, combinations are selected with replacement to reach $k$ total branches.
-For each branch $i = 1, \dots, k$, a random scale factor $s_i$ is used to vary the curvature; If $k \leq \binom{p}{2}$: $s_i = 1$, or If $k > \binom{p}{2}$: $s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}$.
+For each branch $i = 1, \dots, k$, a random scale factor $s_i$ is used to vary the curvature; If $k \leq \binom{p}{2}$: $s_i = 1$, or If $k > \binom{p}{2}$: $s_i \in \{1, 1.5, \dots, 8\}$.
 
 Each branch contains $n_i$ points such that $\sum_{i=1}^k n_i = n$, where the vector $(n_1, \dots, n_k)$ is randomly drawn using the helper function `gen_nsum()` to partition $n$.
-Within its assigned subspace $(X_{i1}, X_{i2})$, branch $i$ is defined by: $X_{i1} \sim U(0, 2)$, and $X_{i2} = -s_iX_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)$.
-This forms a smooth downward-opening parabola in the plane defined by $(X_{i1}, X_{i2})$, with the degree of curvature controlled by $s_i$.
-For $p > 2$, all remaining dimensions are independent Gaussian noise: $X_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}$.
+Within its assigned subspace, branch $i$ is defined by: $X_{i_1} \sim U(0, 2)$, and $X_{i_2} = -s_iX_{i_1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)$.
+This forms a smooth downward-opening parabola in the plane defined by $(X_{i_1}, X_{i_2})$, with the degree of curvature controlled by $s_i$.
+For $p > 2$, all remaining dimensions are independent Gaussian noise: $X_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i_1, i_2\}$.
 
@@
 
-To simulate a cone-shaped structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cone)"} else { "Figure \\@ref(fig:cone-proj)"})`) in arbitrary dimensions, we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
-
-This function generates $n$ points in a $p\text{-}D$, where the last dimension ($X_p$) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
-Points along the height axis are drawn from an exponential distribution to increase density near the tip.
-The $X_p = h_i \sim \text{Exp}(\lambda = 2/h)$ truncated at $h$.
+To simulate a cone-shaped structure in arbitrary dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cone)"} else { "Figure \\@ref(fig:cone-proj)"})`), we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+
+This function generates $n$ points in $p\text{-}D$, where the last dimension ($p$) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
+Points along the height axis, $p$, are drawn from an exponential distribution to increase density near the tip.
+That is, $X_p = h_i \sim \text{Exp}(\lambda = 2/h)$ truncated at $h$.
 The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter $r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1$.

### 2025-09-08 paper-cardinalR.Rmd
Stash

--- 2025-09-05 paper-cardinalR.Rmd
@@
 In the field of single-cell omics, `splatter` (@luke2017) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
-
-There is an  opportunity to improve the generation of high-dimensional data structures by integrating geometric principles with advanced noise control and customizable clustering.
-The `geozoo` package provides a strong foundation but could be enhanced to support high-dimensional extensions with controlled noise and user-defined parameters for clustering.
-Similarly, while `snedata` focuses on abstract datasets for dimensionality reduction, adding features for generating high-dimensional data from geometric layouts would enhance its usability.
-The `mlbench` package could also benefit from allowing users to create datasets with specific geometric structures and noise profiles.
-Additionally, although `splatter` specializes in biological data simulation, it could be expanded to offer a broader framework for generating diverse geometric structures across dimensions, enabling detailed control over noise and clustering.
-Addressing these areas could lead to more robust high-dimensional data generation tools.
+While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.
 
@@
 `cardinalR` offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
+By using clearly defined geometric shapes and controllable features such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets for evaluating algorithm performance in high-dimensional settings.
+
 By providing these features, `cardinalR` provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
@@
 The paper is organized as follows.
-In next section, introduces the implementation of `cardinalR` package on CRAN and GitHub, including demonstration of the package's key functions.
-We illustrate how a clustering data structure affect the dimension reductions in **Application** section.
-Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.
+In the next section, we introduce the implementation of the `cardinalR` package on GitHub, including a demonstration of the package's key functions.
+We illustrate how a clustering data structure affects the dimension reductions in the Application section.
+Finally, we give a brief conclusion of the paper and discuss potential opportunities for the use of our data collection.
 

### 2025-09-08 paper-cardinalR.tex
Stash

@@
 
-There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-Another useful package is \texttt{snedata} (\citet{james2025}), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
-Additionally, \texttt{mlbench} (\citet{friedrich2024}) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
-In the field of single-cell omics, \texttt{splatter} (\citet{luke2017}) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+Several R packages address this need from different perspectives.
+For example, \texttt{geozoo} (\citet{barret2016}) provides a wide range of geometric objects, primarily in low dimensions; \texttt{snedata} (\citet{james2025}) focuses on datasets for dimensionality reduction; \texttt{mlbench} (\citet{friedrich2024}) supplies benchmark datasets for classification and regression; and \texttt{splatter} (\citet{luke2017}) specializes in simulating single-cell omics data.
+While each serves its own domain well, none are designed for flexible, high-dimensional data generation with customizable geometry, noise, and clustering.
 
@@
 The paper is organized as follows.
-In next section, introduces the implementation of \texttt{cardinalR} package on CRAN and GitHub, including demonstration of the package's key functions.
-We illustrate how a clustering data structure affect the dimension reductions in \textbf{Application} section.
-Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.
+In the next section, we introduce the implementation of the \texttt{cardinalR} package on GitHub, including a demonstration of the package's key functions.
+We illustrate how a clustering data structure affects the dimension reductions in the Application section.
+Finally, we give a brief conclusion of the paper and discuss potential opportunities for the use of our data collection.
 

### 2025-09-08 paper-cardinalR.tex
Rendered new one

--- 2025-09-05 paper-cardinalR.tex
@@
 In the field of single-cell omics, \texttt{splatter} (\citet{luke2017}) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
-
-There is an opportunity to improve the generation of high-dimensional data structures by integrating geometric principles with advanced noise control and customizable clustering.
-The \texttt{geozoo} package provides a strong foundation but could be enhanced to support high-dimensional extensions with controlled noise and user-defined parameters for clustering.
-Similarly, while \texttt{snedata} focuses on abstract datasets for dimensionality reduction, adding features for generating high-dimensional data from geometric layouts would enhance its usability.
-The \texttt{mlbench} package could also benefit from allowing users to create datasets with specific geometric structures and noise profiles.
-Additionally, although \texttt{splatter} specializes in biological data simulation, it could be expanded to offer a broader framework for generating diverse geometric structures across dimensions, enabling detailed control over noise and clustering.
-Addressing these areas could lead to more robust high-dimensional data generation tools.
+While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.
 
@@
 \texttt{cardinalR} offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
+By using clearly defined geometric shapes and controllable features such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets for evaluating algorithm performance in high-dimensional settings.
+
 By providing these features, \texttt{cardinalR} provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
@@
 The underlying geometry lies in the first two dimensions, while the remaining \((p-2)\) dimensions contain Gaussian noise.
-
 Branch \(1\) is generated from \(X_1 \sim U(0, 1)\), and \(X_2 = 0.1X_1 + X_1^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
@@
 Branches \(3\) to \(k\) are added iteratively.
-Each new branch \(i\) begins at a randomly chosen point within a restricted horizontal band: \(X_{1} \in [-0.15, 0.15]\) (to ensure connectivity with earlier branches), spans a unit-length interval on \(X_{1}\):\(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), and has the structure: \(X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}\).
+Each new branch \(i\) begins at a randomly chosen point within a restricted horizontal band: \(X_{1} \in [-0.15, 0.15]\) to ensure connectivity with earlier branches, spans a unit-length interval on \(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), and has the structure: \(X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, \; s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}\).
 Here, \(s_i\) is a sampled scale factor (slope-like term) controlling curvature, and \((x_{\text{start}}, y_{\text{start}})\) is the starting point from the existing structure.
@@
 The \texttt{gen\_orglinearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches embedded in \(p\text{-}D\) (Figure \ref{fig:branch-proj} d).
-Each branch lies primarily within a distinct \(2\text{-}D\) subspace, while the remaining \(p - 2\) dimensions contain Gaussian noise.
-
-To construct each branch, a unique or repeated pair of dimensions is selected from the \(\binom{p}{2}\) possible \(2\text{-}D\) combinations.
+Each branch lies within a distinct \(2\text{-}D\) subspace, while the remaining \(p - 2\) dimensions contain Gaussian noise.
+
+To construct each branch, a unique or repeated pair of dimensions is selected from the \(\binom{p}{2}\) possible combinations of dimensions.
 If \(k \leq \binom{p}{2}\), combinations are sampled without replacement.
 If \(k > \binom{p}{2}\), additional pairs are sampled with replacement to reach \(k\) total branches.
-Each selected pair \((X_{i1}, X_{i2})\) defines the \(2\text{-}D\) plane for branch \(i\).
+Each selected pair \((i_1, i_2)\) defines the \(2\text{-}D\) plane for branch \(i\).
 
 For branch \(i\), \(n_i\) points are generated, where \(\sum_{i=1}^k n_i = n\).
-The structure follows: \(X_{i1} \sim U(0, 2)\), and \(X_{i2} = -s_iX_{i1} + \epsilon, \quad \epsilon \sim U(0, 0.5)\), where \(s_i\) is a scale factor controlling the slope.
+The structure follows: \(X_{i_1} \sim U(0, 2)\), and \(X_{i_2} = -s_iX_{i_1} + \epsilon, \quad \epsilon \sim U(0, 0.5)\), where \(s_i\) is a scale factor controlling the slope.
 When \(k \leq \binom{p}{2}\), \(s_i = 1\).
 When sampling with replacement, \(s_i\) is drawn from the set \({1, 1.5, 2, \dots, 8}\) in increments of \(0.5\).
-For \(p > 2\), the remaining dimensions contain independent Gaussian noise: \(X_j \sim N(0, 0.1^2)\), for \(j \notin \{i1, i2\}\).
+For \(p > 2\), the remaining dimensions contain independent Gaussian noise: \(X_j \sim N(0, 0.1^2)\), for \(j \notin \{i_1, i_2\}\).
 
@@
 Let \(\binom{p}{2}\) denote the number of unique \(2\text{-}D\) subspace combinations.
-When \(k \leq \binom{p}{2}\), \(k\) distinct subspace pairs \((X_{i1}, X_{i2})\) are sampled without replacement.
+When \(k \leq \binom{p}{2}\), \(k\) distinct subspace pairs \((i_1, i_2)\) are sampled without replacement.
 Otherwise, combinations are selected with replacement to reach \(k\) total branches.
-For each branch \(i = 1, \dots, k\), a random scale factor \(s_i\) is used to vary the curvature; If \(k \leq \binom{p}{2}\): \(s_i = 1\), or If \(k > \binom{p}{2}\): \(s_i \sim \text{Uniform sample from } {1, 1.5, \dots, 8}\).
+For each branch \(i = 1, \dots, k\), a random scale factor \(s_i\) is used to vary the curvature; If \(k \leq \binom{p}{2}\): \(s_i = 1\), or If \(k > \binom{p}{2}\): \(s_i \in \{1, 1.5, \dots, 8\}\).
 
 Each branch contains \(n_i\) points such that \(\sum_{i=1}^k n_i = n\), where the vector \((n_1, \dots, n_k)\) is randomly drawn using the helper function \texttt{gen\_nsum()} to partition \(n\).
-Within its assigned subspace \((X_{i1}, X_{i2})\), branch \(i\) is defined by: \(X_{i1} \sim U(0, 2)\), and \(X_{i2} = -s_iX_{i1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)\).
-This forms a smooth downward-opening parabola in the plane defined by \((X_{i1}, X_{i2})\), with the degree of curvature controlled by \(s_i\).
-For \(p > 2\), all remaining dimensions are independent Gaussian noise: \(X_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i1, i2\}\).
+Within its assigned subspace, branch \(i\) is defined by: \(X_{i_1} \sim U(0, 2)\), and \(X_{i_2} = -s_iX_{i_1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)\).
+This forms a smooth downward-opening parabola in the plane defined by \((X_{i_1}, X_{i_2})\), with the degree of curvature controlled by \(s_i\).
+For \(p > 2\), all remaining dimensions are independent Gaussian noise: \(X_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i_1, i_2\}\).
 
@@
 
-To simulate a cone-shaped structure (Figure \ref{fig:cone-proj}) in arbitrary dimensions, we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
-
-This function generates \(n\) points in a \(p\text{-}D\), where the last dimension (\(X_p\)) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
-Points along the height axis are drawn from an exponential distribution to increase density near the tip.
-The \(X_p = h_i \sim \text{Exp}(\lambda = 2/h)\) truncated at \(h\).
+To simulate a cone-shaped structure in arbitrary dimensions (Figure \ref{fig:cone-proj}), we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+
+This function generates \(n\) points in \(p\text{-}D\), where the last dimension (\(p\)) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
+Points along the height axis, \(p\), are drawn from an exponential distribution to increase density near the tip.
+That is, \(X_p = h_i \sim \text{Exp}(\lambda = 2/h)\) truncated at \(h\).
 The effective radius at height \(h_i\) decreases linearly from base to tip, controlled by a shape ratio parameter \(r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1\).

### 2025-09-08 paper-cardinalR.Rmd
Changes up to pyramid types

@@
 
-The Mobius strip is generated using `geozoo::mobius(n, p = 3)$points`, which samples $n$ points from a parametrization of the surface defined by angle $\theta$ and strip width $w$.
+The Mobius strip is generated using `geozoo::mobius(n, p = 3)$points`, which samples $n$ points from a parameterization of the surface defined by angle $\theta$ and strip width $w$.
 The cartesian coordinates are $X_1 = \left(1 + w/2 \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + w/2 \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = w/2 \sin(\theta/2)$.

### 2025-09-16 paper-cardinalR.Rmd
Intro changes

--- 2025-09-08 paper-cardinalR.Rmd
@@
 
-The Mobius strip is generated using `geozoo::mobius(n, p = 3)$points`, which samples $n$ points from a parametrization of the surface defined by angle $\theta$ and strip width $w$.
+The Mobius strip is generated using the `geozoo::mobius` function with `p = 3`, and `n` defined as above.
+This function samples $n$ points from a parameterization of the surface defined by angle $\theta$ and strip width $w$.
 The cartesian coordinates are $X_1 = \left(1 + w/2 \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + w/2 \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = w/2 \sin(\theta/2)$.

### 2025-09-16 paper-cardinalR.Rmd
Cube section rework

@@
 
+The first is the regular grid of points of $n$ points in $p$ dimensions.
+This is generated using `gen_gridcube(n, p)`.
+The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to $n$.
+In the base cube function, the cube is represented in integer grid coordinates (e.g.
+1, 2, 3, ...) along each ais.
+This results in a $p\text{-}D$ hyercube-shaped lattice structure.
+
+An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
+The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) b"} else { "Figure \\@ref(fig:cube-proj) b"})`).
+
+Finally, we consider a cube that has a central spherical hole.
+This is generated using the `gen_cubehole(n, p)` function.
+The cube is generated as per the uniformly distributed cube, but points inside sphere of radius $0.5$ are removed, resulting in a hollow cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) c"} else { "Figure \\@ref(fig:cube-proj) c"})`).
+
 ```{r}
@@
 
-#### `gen_gridcube()`
-
-The `gen_gridcube(n, p)` constructs a regular grid using approximately $n$ points by evenly partitioning each of the $p$ dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) a"} else { "Figure \\@ref(fig:cube-proj) a"})`).
-Specifically, the number of grid points along each axis is determined by `gen_nproduct(n, p)`, which calculates the nearest balanced integer factors whose product is close to $n$.
-The full grid is generated via `expand_grid()`, giving Cartesian combinations of these evenly spaced positions.
-
-Each variable $X_j$ represents integer grid coordinates (e.g., 1, 2, 3, ...) in dimension $j$: $X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n$.
-This results in a hypercube-shaped lattice in $p\text{-}D$.
-
 ```{r data-gridcube, echo=TRUE}
@@
 
-#### `gen_unifcube()`
-
-The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in $p\text{-}D$ with additional noise dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) b"} else { "Figure \\@ref(fig:cube-proj) b"})`).
-
-Points are sampled independently from $X_1, X_2, X_3 \sim U(-0.5, 0.5)$ of a cube of side length $1$.
-This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.
-For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ structure into $p\text{-}D$, where $j = 4, \dots, p$.
-
 ```{r data-unifcube, echo=TRUE}
@@
 ```
-
-#### `gen_cubehole()`
-
-The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) c"} else { "Figure \\@ref(fig:cube-proj) c"})`).
-
-The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim U(-0.5, 0.5)$.
-The center of the cube is $(0, 0, 0)$.
-A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$: remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$.
-This results in a shell-like point cloud that fills the cube but avoids the spherical center.
-For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ structure into $p\text{-}D$, where $j = 4, \dots, p$.
 

### 2025-09-16 paper-cardinalR.Rmd
Defined the s parameter in Gaussian function

@@
                                 "A numeric value representing the number of dimensions.",
-                                "A numeric value representing the number of clusters."))
+                                "A numeric value representing the number of branches."))
 ```
@@
 The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:gau)"} else { "Figure \\@ref(fig:gau-proj)"})`).
-Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, \Sigma)$, where $\Sigma$ is a user-defined $p \times p$ positive-definite matrix.
+Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
 

### 2025-09-16 paper-cardinalR.tex
Defined the s parameter in Gaussian function

@@
 p & A numeric value representing the number of dimensions.\\
-k & A numeric value representing the number of clusters.\\
+k & A numeric value representing the number of branches.\\
 \bottomrule
@@
 
+The first is the regular grid of points of \(n\) points in \(p\) dimensions.
+This is generated using \texttt{gen\_gridcube(n,\ p)}.
+The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to \(n\).
+In the base cube function, the cube is represented in integer grid coordinates (e.g.~1, 2, 3, \ldots) along each ais.
+This results in a \(p\text{-}D\) hyercube-shaped lattice structure.
+
+An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
+The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube (Figure \ref{fig:cube-proj} b).
+
+Finally, we consider a cube that has a central spherical hole.
+This is generated using the \texttt{gen\_cubehole(n,\ p)} function.
+The cube is generated as per the uniformly distributed cube, but points inside sphere of radius \(0.5\) are removed, resulting in a hollow cube structure (Figure \ref{fig:cube-proj} c).
+
 \begin{table}
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_gridcube()}}{gen\_gridcube()}}\label{gen_gridcube}
-
-The \texttt{gen\_gridcube(n,\ p)} constructs a regular grid using approximately \(n\) points by evenly partitioning each of the \(p\) dimensions (Figure \ref{fig:cube-proj} a).
-Specifically, the number of grid points along each axis is determined by \texttt{gen\_nproduct(n,\ p)}, which calculates the nearest balanced integer factors whose product is close to \(n\).
-The full grid is generated via \texttt{expand\_grid()}, giving Cartesian combinations of these evenly spaced positions.
-
-Each variable \(X_j\) represents integer grid coordinates (e.g., 1, 2, 3, \ldots) in dimension \(j\): \(X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n\).
-This results in a hypercube-shaped lattice in \(p\text{-}D\).
-
 \begin{verbatim}
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_unifcube()}}{gen\_unifcube()}}\label{gen_unifcube}
-
-The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in \(p\text{-}D\) with additional noise dimensions (Figure \ref{fig:cube-proj} b).
-
-Points are sampled independently from \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\) of a cube of side length \(1\).
-This results in a uniform cloud of points filling the unit cube in \(3\text{-}D\) space.
-For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) structure into \(p\text{-}D\), where \(j = 4, \dots, p\).
-
 \begin{verbatim}
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_cubehole()}}{gen\_cubehole()}}\label{gen_cubehole}
-
-The \texttt{gen\_cubehole(n,\ p)} function generates a dataset of \(n\) points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in \(p\text{-}D\) (Figure \ref{fig:cube-proj} c).
-
-The initial points are drawn from a uniform cube centered at the origin: \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\).
-The center of the cube is \((0, 0, 0)\).
-A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\): remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\).
-This results in a shell-like point cloud that fills the cube but avoids the spherical center.
-For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) structure into \(p\text{-}D\), where \(j = 4, \dots, p\).
-
 \begin{verbatim}
@@
 The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure (Figure \ref{fig:gau-proj}).
-Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, \Sigma)\), where \(\Sigma\) is a user-defined \(p \times p\) positive-definite matrix.
+Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, s)\), where \(s\) is a user-defined \(p \times p\) positive-definite matrix.
 
@@
 
-The Mobius strip is generated using \texttt{geozoo::mobius(n,\ p\ =\ 3)\$points}, which samples \(n\) points from a parametrization of the surface defined by angle \(\theta\) and strip width \(w\).
+The Mobius strip is generated using the \texttt{geozoo::mobius} function with \texttt{p\ =\ 3}, and \texttt{n} defined as above.
+This function samples \(n\) points from a parameterization of the surface defined by angle \(\theta\) and strip width \(w\).
 The cartesian coordinates are \(X_1 = \left(1 + w/2 \cos(\theta/2)\right) \cos(\theta)\), \(X_2 = \left(1 + w/2 \cos(\theta/2)\right) \sin(\theta)\), and \(X_3 = w/2 \sin(\theta/2)\).

### 2025-09-16 paper-cardinalR.R

@@ -158,7 +158,7 @@ arg_branching_tb <- tibble(arg = c("n",
                           "k"), 
                         exp = c("A numeric value representing the number of points.",
                                 "A numeric value representing the number of dimensions.",
-                                "A numeric value representing the number of clusters."))
+                                "A numeric value representing the number of branches."))
 
 
 ## ----arg-branching-tb-html, eval=knitr::is_html_output()----------------------

### 2025-09-16 paper-cardinalR.Rmd
Updated with Thiyanga's comments

@@
 `cardinalR` offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable features such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets for evaluating algorithm performance in high-dimensional settings.
-
-By providing these features, `cardinalR` provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
+By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets for evaluating algorithm performance in high-dimensional settings.
+
+By providing these aspects, `cardinalR` provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
 This bridges the gap between geometric foundations and the flexible generation of complex synthetic data.

### 2025-09-16 paper-cardinalR.tex
Updated with Thiyanga's comments

@@
 \texttt{cardinalR} offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable features such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets for evaluating algorithm performance in high-dimensional settings.
-
-By providing these features, \texttt{cardinalR} provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
+By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets for evaluating algorithm performance in high-dimensional settings.
+
+By providing these aspects, \texttt{cardinalR} provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
 This bridges the gap between geometric foundations and the flexible generation of complex synthetic data.

### 2025-09-16 paper-cardinalR.Rmd
Added r_hole

@@
 Finally, we consider a cube that has a central spherical hole.
-This is generated using the `gen_cubehole(n, p)` function.
-The cube is generated as per the uniformly distributed cube, but points inside sphere of radius $0.5$ are removed, resulting in a hollow cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) c"} else { "Figure \\@ref(fig:cube-proj) c"})`).
+This is generated using the `gen_cubehole(n, p, r_hole)` function.
+The cube is generated as per the uniformly distributed cube, but points inside sphere of radius (`r_hole`) are removed, resulting in a hollow cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) c"} else { "Figure \\@ref(fig:cube-proj) c"})`).
 
@@
 ```{r data-cubehole, echo=TRUE}
-cubehole <- gen_cubehole(n = 3000, p = 4)
+cubehole <- gen_cubehole(n = 3000, p = 4, r_hole = 0.5)
 ```

### 2025-09-16 paper-cardinalR.tex
Added r_hole

@@
 Finally, we consider a cube that has a central spherical hole.
-This is generated using the \texttt{gen\_cubehole(n,\ p)} function.
-The cube is generated as per the uniformly distributed cube, but points inside sphere of radius \(0.5\) are removed, resulting in a hollow cube structure (Figure \ref{fig:cube-proj} c).
+This is generated using the \texttt{gen\_cubehole(n,\ p,\ r\_hole)} function.
+The cube is generated as per the uniformly distributed cube, but points inside sphere of radius (\texttt{r\_hole}) are removed, resulting in a hollow cube structure (Figure \ref{fig:cube-proj} c).
 
@@
 \begin{verbatim}
-cubehole <- gen_cubehole(n = 3000, p = 4)
+cubehole <- gen_cubehole(n = 3000, p = 4, r_hole = 0.5)
 \end{verbatim}

### 2025-09-16 paper-cardinalR.Rmd
Updated pyramid table info

@@
                              "gen_pyrholes"), 
-                        exp = c("Generate a pyramid with a rectangular base, with the option of a sharp or blunted apex.",
-                                "Generate a pyramid with a triangular base, with the option of a sharp or blunted apex.",
-                                "Generate a pyramid with a star-shape base, with the option of a sharp or blunted apex.",
-                                "Generate a pyramid with triangular pyramid shaped holes."))
+                        exp = c("Rectangular-base pyramid, with a sharp or blunted apex.",
+          "Triangular-base pyramid, with a sharp or blunted apex.",
+          "Star-shaped base pyramid, with a sharp or blunted apex.",
+          "Pyramid containing triangular pyramid-shaped holes."))
 ```

### 2025-09-16 paper-cardinalR.tex
Updated pyramid table info

@@
 \midrule
-gen\_pyrrect & Generate a pyramid with a rectangular base, with the option of a sharp or blunted apex.\\
-gen\_pyrtri & Generate a pyramid with a triangular base, with the option of a sharp or blunted apex.\\
-gen\_pyrstar & Generate a pyramid with a star-shape base, with the option of a sharp or blunted apex.\\
-gen\_pyrholes & Generate a pyramid with triangular pyramid shaped holes.\\
+gen\_pyrrect & Rectangular-base pyramid, with a sharp or blunted apex.\\
+gen\_pyrtri & Triangular-base pyramid, with a sharp or blunted apex.\\
+gen\_pyrstar & Star-shaped base pyramid, with a sharp or blunted apex.\\
+gen\_pyrholes & Pyramid containing triangular pyramid-shaped holes.\\
 \bottomrule

### 2025-09-16 paper-cardinalR.Rmd
Added details regarding to the spread

@@
 
-The `gen_longlinear(n, p)` function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:linear)"} else { "Figure \\@ref(fig:linear-proj)"})`).
-Each variable $X_i$ is created as $X_i = \text{scale}_i \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i$, where $\text{scale}_i \sim U(-10, 10)$ is randomly chosen direction and stretch, $\text{shift}_i \sim U(-300, 300)$ is large offset to spread features apart, and $\epsilon \sim N(0, 0.03n)$ is Gaussian noise.
+The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise.
+Each variable is formed as $X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i$, where $\text{scale}\_i \sim U(-10, 10)$ determines the orientation of the line in each dimension, $\text{shift}\_i \sim U(-300, 300)$ offsets the line to separate dimensions, and $\epsilon \sim N(0, (0.03n)^2)$ introduces Gaussian noise.
 
@@
 ```{r trefoil-tb-pdf, eval=knitr::is_latex_output()}
+
 trefoil_tb |> 

### 2025-09-16 paper-cardinalR.tex
Added details regarding to the spread

@@
 
-The \texttt{gen\_longlinear(n,\ p)} function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise (Figure \ref{fig:linear-proj}).
-Each variable \(X_i\) is created as \(X_i = \text{scale}_i \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i\), where \(\text{scale}_i \sim U(-10, 10)\) is randomly chosen direction and stretch, \(\text{shift}_i \sim U(-300, 300)\) is large offset to spread features apart, and \(\epsilon \sim N(0, 0.03n)\) is Gaussian noise.
+The \texttt{gen\_longlinear(n,\ p)} function generates a high-dimensional dataset representing a long linear structure with noise.
+Each variable is formed as \(X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i\), where \(\text{scale}\_i \sim U(-10, 10)\) determines the orientation of the line in each dimension, \(\text{shift}\_i \sim U(-300, 300)\) offsets the line to separate dimensions, and \(\epsilon \sim N(0, (0.03n)^2)\) introduces Gaussian noise.
 

### 2025-09-16 paper-cardinalR.Rmd
Updated the operations ordering of Mobius fuctions
Added descriptions for theta and w in Mobius

@@
 The Mobius strip is generated using the `geozoo::mobius` function with `p = 3`, and `n` defined as above.
-This function samples $n$ points from a parameterization of the surface defined by angle $\theta$ and strip width $w$.
-The cartesian coordinates are $X_1 = \left(1 + w/2 \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + w/2 \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = w/2 \sin(\theta/2)$.
+Each point is sampled from a parameterization defined by an angle $\theta \sim U(0, 2\pi)$, which specifies the position along the circular loop, and a width $w \sim U(-0.4, 0.4)$, which offsets the point across the strip’s band.
+The cartesian coordinates are $X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = (w/2)\sin(\theta/2)$.
 This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface.

### 2025-09-16 paper-cardinalR.tex
Updated the operations ordering of Mobius fuctions
Added descriptions for theta and w in Mobius

@@
 The Mobius strip is generated using the \texttt{geozoo::mobius} function with \texttt{p\ =\ 3}, and \texttt{n} defined as above.
-This function samples \(n\) points from a parameterization of the surface defined by angle \(\theta\) and strip width \(w\).
-The cartesian coordinates are \(X_1 = \left(1 + w/2 \cos(\theta/2)\right) \cos(\theta)\), \(X_2 = \left(1 + w/2 \cos(\theta/2)\right) \sin(\theta)\), and \(X_3 = w/2 \sin(\theta/2)\).
+Each point is sampled from a parameterization defined by an angle \(\theta \sim U(0, 2\pi)\), which specifies the position along the circular loop, and a width \(w \sim U(-0.4, 0.4)\), which offsets the point across the strip's band.
+The cartesian coordinates are \(X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta)\), \(X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta)\), and \(X_3 = (w/2)\sin(\theta/2)\).
 This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable one-sided surface.

### 2025-09-16 paper-cardinalR.Rmd
Updated the Quadratic text with "randomly sampling inputs" than Uniform
Updated the cone generation

@@
 
-This function generates $n$ points in $p\text{-}D$, where the last dimension ($p$) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
-Points along the height axis, $p$, are drawn from an exponential distribution to increase density near the tip.
-That is, $X_p = h_i \sim \text{Exp}(\lambda = 2/h)$ truncated at $h$.
-The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter $r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1$.
-
-For each point, a direction is sampled uniformly from a $(p-1)$-dimensional hypersphere using generalized spherical coordinates (angles).
-The radial coordinates are scaled by $r_i$, which ensures a conical taper.
-Specifically, when $p = 3$, this scaling results in a classic $3\text{-}D$ cone.
-However, for $p > 3$, the introduction of additional angular components allows for a smooth extension into higher dimensions, preserving the conical shape while accommodating the complexities of multi-dimensional geometry.
+This function generates $n$ points in $p-D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip.
+Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, capped at the cone height $h$, producing a higher density of points near the tip.
+At each height $X_p$, the radius of the cross-section decreases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio}$ and $r_{\text{max}} = 1$.
+
+For each point, a direction in the first $p-1$ dimensions is sampled uniformly on a $(p-1)$-dimensional hypersphere using generalized spherical coordinates.
+The radial coordinates are scaled by the height-dependent radius $r$, producing the conical taper.
+In three dimensions ($p = 3$), this results in a classical 3-D cone, while for $p > 3$, additional angular components allow a smooth extension into higher dimensions, preserving the conical shape within multi-dimensional space.
 
@@
 
-The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation.
+The curve is constructed by randomly sampling inputs and applying a second-degree polynomial transformation.
 Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the independent variable.

### 2025-09-16 paper-cardinalR.tex
Updated the Quadratic text with "randomly sampling inputs" than Uniform
Updated the cone generation

@@
 
-This function generates \(n\) points in \(p\text{-}D\), where the last dimension (\(p\)) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.
-Points along the height axis, \(p\), are drawn from an exponential distribution to increase density near the tip.
-That is, \(X_p = h_i \sim \text{Exp}(\lambda = 2/h)\) truncated at \(h\).
-The effective radius at height \(h_i\) decreases linearly from base to tip, controlled by a shape ratio parameter \(r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1\).
-
-For each point, a direction is sampled uniformly from a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates (angles).
-The radial coordinates are scaled by \(r_i\), which ensures a conical taper.
-Specifically, when \(p = 3\), this scaling results in a classic \(3\text{-}D\) cone.
-However, for \(p > 3\), the introduction of additional angular components allows for a smooth extension into higher dimensions, preserving the conical shape while accommodating the complexities of multi-dimensional geometry.
+This function generates \(n\) points in \(p-D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip.
+Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), capped at the cone height \(h\), producing a higher density of points near the tip.
+At each height \(X_p\), the radius of the cross-section decreases linearly from base to tip according to \(r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h\), where \(r_{\text{min}} = \text{ratio}\) and \(r_{\text{max}} = 1\).
+
+For each point, a direction in the first \(p-1\) dimensions is sampled uniformly on a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates.
+The radial coordinates are scaled by the height-dependent radius \(r\), producing the conical taper.
+In three dimensions (\(p = 3\)), this results in a classical 3-D cone, while for \(p > 3\), additional angular components allow a smooth extension into higher dimensions, preserving the conical shape within multi-dimensional space.
 
@@
 
-The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation.
+The curve is constructed by randomly sampling inputs and applying a second-degree polynomial transformation.
 Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the independent variable.

### 2025-09-16 paper-cardinalR.Rmd
Updated cube description with normalise text

@@
 The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to $n$.
-In the base cube function, the cube is represented in integer grid coordinates (e.g.
-1, 2, 3, ...) along each ais.
-This results in a $p\text{-}D$ hyercube-shaped lattice structure.
+Each dimension is then normalized to lie in the interval $[0,1]$, so that the resulting structure forms a true $p\text{-}D$ hypercube.
+This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
 

### 2025-09-16 paper-cardinalR.tex
Updated cube description with normalise text

@@
 The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to \(n\).
-In the base cube function, the cube is represented in integer grid coordinates (e.g.~1, 2, 3, \ldots) along each ais.
-This results in a \(p\text{-}D\) hyercube-shaped lattice structure.
+Each dimension is then normalized to lie in the interval \([0,1]\), so that the resulting structure forms a true \(p\text{-}D\) hypercube.
+This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
 

### 2025-09-16 paper-cardinalR.Rmd
Updated gen_linearbranches text by adding slope and intercept info

@@
 
-Each branch is a segment of a line with added jitter to simulate measurement noise.
-The branches differ in direction and location.
-Branches $1$ and $2$ are initialized with fixed slopes and intercepts.
-Branch $1$ is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$.
-The Branch $2$ is generated from $X_1 \sim U(-6, 2)$, $X_2 = -0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$.
+The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points in $p-D$ forming $k$ linear branches.
+Each branch is a line segment with added jitter to simulate noise.
+Branches $1$ and $2$ are initialized with fixed slopes and intercepts: branch $1$ has slope $0.5$ and intercept $0$, with $X_1 \sim U(-2, 8)$ and $X_2 = 0.5 X_1 + \epsilon$, $\epsilon \sim U(0, 0.5)$; branch $2$ has slope $-0.5$ and intercept $0$, with $X_1 \sim U(-6, 2)$ and $X_2 = -0.5 X_1 + \epsilon$, $\epsilon \sim U(0, 0.5)$.
+
 Branches $3$ to $k$ are added iteratively.
-Each additional branch $i$ starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
-The $X_1$ values are defined over a short range, from $x_{start}$ to $x_{start} + 1$.
-The $X_2$ value is calculated using the formula $X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon$, where $s_i$ is a chosen slope from a selected branch, and $\epsilon \sim U(0, 0.2)$.
-For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
+Each new branch starts at a location outside predefined exclusion zones to avoid overlap with the first two branches.
+Specifically, $X_1$ starting points are sampled outside the intervals $[-8, -7]$, $[-2, 2]$, and $[7, 8]$, and $X_2$ starting points are sampled outside $[7, 8]$.
+For each new branch $i$, $X_1$ values are drawn over a short interval $[x_{\text{start}}, x_{\text{start}} + 1]$, and $X_2$ values are calculated as $X_2 = s_i (X_1 - x_{\text{start}}) + y_{\text{start}} + \epsilon$, where $s_i$ is a randomly chosen slope from a filtered sequence and $\epsilon \sim U(0, 0.2)$.
+
+For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, for $j = 3, \dots, p$.
 

### 2025-09-16 paper-cardinalR.tex
Updated gen_linearbranches text by adding slope and intercept info

@@
 
-Each branch is a segment of a line with added jitter to simulate measurement noise.
-The branches differ in direction and location.
-Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts.
-Branch \(1\) is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\).
-The Branch \(2\) is generated from \(X_1 \sim U(-6, 2)\), \(X_2 = -0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\).
+The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points in \(p-D\) forming \(k\) linear branches.
+Each branch is a line segment with added jitter to simulate noise.
+Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts: branch \(1\) has slope \(0.5\) and intercept \(0\), with \(X_1 \sim U(-2, 8)\) and \(X_2 = 0.5 X_1 + \epsilon\), \(\epsilon \sim U(0, 0.5)\); branch \(2\) has slope \(-0.5\) and intercept \(0\), with \(X_1 \sim U(-6, 2)\) and \(X_2 = -0.5 X_1 + \epsilon\), \(\epsilon \sim U(0, 0.5)\).
+
 Branches \(3\) to \(k\) are added iteratively.
-Each additional branch \(i\) starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.
-The \(X_1\) values are defined over a short range, from \(x_{start}\) to \(x_{start} + 1\).
-The \(X_2\) value is calculated using the formula \(X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon\), where \(s_i\) is a chosen slope from a selected branch, and \(\epsilon \sim U(0, 0.2)\).
-For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
+Each new branch starts at a location outside predefined exclusion zones to avoid overlap with the first two branches.
+Specifically, \(X_1\) starting points are sampled outside the intervals \([-8, -7]\), \([-2, 2]\), and \([7, 8]\), and \(X_2\) starting points are sampled outside \([7, 8]\).
+For each new branch \(i\), \(X_1\) values are drawn over a short interval \([x_{\text{start}}, x_{\text{start}} + 1]\), and \(X_2\) values are calculated as \(X_2 = s_i (X_1 - x_{\text{start}}) + y_{\text{start}} + \epsilon\), where \(s_i\) is a randomly chosen slope from a filtered sequence and \(\epsilon \sim U(0, 0.2)\).
+
+For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), for \(j = 3, \dots, p\).
 

### 2025-09-16 paper-cardinalR.Rmd
Added how clusters can be map with branching data

@@
 A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology.
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points, and number of branches.
+Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:branching-tb-html)', '\\@ref(tab:branching-tb-pdf)'))` outlines these functions.
@@
                                 "A numeric value representing the number of dimensions.",
-                                "A numeric value representing the number of branches."))
+                                "A numeric value representing the number of clusters."))
 ```
@@
 Each branch $i$ is constructed using $X_1 \sim U(-2, 2)$ and $X_2 = \exp(\pm s_iX_1) + \epsilon$, where: $\epsilon \sim U(0, 0.1)$ adds local jitter, $s_i \in [0.5, 2]$ is randomly sampled for each branch.
-The sign alternates between branches (odd: negative, even: positive exponent).
-This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches represent a decay pattern (reflected) and even-numbered branches represent a growth pattern.
+The sign of the exponent alternates between branches: odd-numbered branches have a negative exponent (decaying pattern), and even-numbered branches have a positive exponent (growing pattern), producing mirror-symmetric curves with varying steepness.
 For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
@@
 
-The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points in $p-D$ forming $k$ linear branches.
+The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points in $p\text{-}D$ forming $k$ linear branches.
 Each branch is a line segment with added jitter to simulate noise.
@@
 For each new branch $i$, $X_1$ values are drawn over a short interval $[x_{\text{start}}, x_{\text{start}} + 1]$, and $X_2$ values are calculated as $X_2 = s_i (X_1 - x_{\text{start}}) + y_{\text{start}} + \epsilon$, where $s_i$ is a randomly chosen slope from a filtered sequence and $\epsilon \sim U(0, 0.2)$.
-
 For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, for $j = 3, \dots, p$.
@@
 
-This function generates $n$ points in $p-D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip.
+This function generates $n$ points in $p\text{-}D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip.
 Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, capped at the cone height $h$, producing a higher density of points near the tip.

### 2025-09-16 paper-cardinalR.tex
Added how clusters can be map with branching data

@@
 A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology.
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometry, number of points, and the number of branches.
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points, and number of branches.
+Although these functions can generate multiple branches, they do not produce a formal \emph{multicluster} dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
 Table \ref{tab:branching-tb-pdf} outlines these functions.
@@
 p & A numeric value representing the number of dimensions.\\
-k & A numeric value representing the number of branches.\\
+k & A numeric value representing the number of clusters.\\
 \bottomrule
@@
 Each branch \(i\) is constructed using \(X_1 \sim U(-2, 2)\) and \(X_2 = \exp(\pm s_iX_1) + \epsilon\), where: \(\epsilon \sim U(0, 0.1)\) adds local jitter, \(s_i \in [0.5, 2]\) is randomly sampled for each branch.
-The sign alternates between branches (odd: negative, even: positive exponent).
-This creates mirror-symmetric branches with different steepness and curvature where odd-numbered branches represent a decay pattern (reflected) and even-numbered branches represent a growth pattern.
+The sign of the exponent alternates between branches: odd-numbered branches have a negative exponent (decaying pattern), and even-numbered branches have a positive exponent (growing pattern), producing mirror-symmetric curves with varying steepness.
 For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
@@
 
-The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points in \(p-D\) forming \(k\) linear branches.
+The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points in \(p\text{-}D\) forming \(k\) linear branches.
 Each branch is a line segment with added jitter to simulate noise.
@@
 For each new branch \(i\), \(X_1\) values are drawn over a short interval \([x_{\text{start}}, x_{\text{start}} + 1]\), and \(X_2\) values are calculated as \(X_2 = s_i (X_1 - x_{\text{start}}) + y_{\text{start}} + \epsilon\), where \(s_i\) is a randomly chosen slope from a filtered sequence and \(\epsilon \sim U(0, 0.2)\).
-
 For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), for \(j = 3, \dots, p\).
@@
 
-This function generates \(n\) points in \(p-D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip.
+This function generates \(n\) points in \(p\text{-}D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip.
 Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), capped at the cone height \(h\), producing a higher density of points near the tip.

### 2025-09-16 paper-cardinalR.R

@@ -158,7 +158,7 @@ arg_branching_tb <- tibble(arg = c("n",
                           "k"), 
                         exp = c("A numeric value representing the number of points.",
                                 "A numeric value representing the number of dimensions.",
-                                "A numeric value representing the number of branches."))
+                                "A numeric value representing the number of clusters."))
 
 
 ## ----arg-branching-tb-html, eval=knitr::is_html_output()----------------------

### 2025-09-17 paper-cardinalR.Rmd
Rework pyramid section

@@
 A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
-The functions provided allow users to generate pyramids with various base shapes—including rectangular, triangular, and star-shaped bases—with options for sharp or blunted apexes.
-Additionally, we offer functionality for creating pyramids containing structured holes, enabling the study of non-convex and sparse regions.
-These designs are valuable for testing the ability of dimension reduction and clustering algorithms to preserve and detect complex geometric features.
+The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
+Additionally, similar to the cube generator, it is possible to create a pyramids containing structured holes, enabling the study of non-convex and sparse regions.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:pyramid-tb-html)', '\\@ref(tab:pyramid-tb-pdf)'))` summarizes these functions.
+
+Let $X_1, \dots, X_p$ denote the coordinates of the generated points.
+For the rectangular and triangular based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
+That is, $$X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$$.
+This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
+That is, $X_p = z \sim U(0, h)$.
+
+The remaining dimensions are based on the specific pyramid shape.
+For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`), let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$.
+That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
+The first three coordinates are then defined as: $X_1 \sim U(-r_x(z),\ r_x(z))$, $X_2 \sim U(-r_y(z),\ r_y(z))$, and $X_3 \sim U(-r_x(z),\ r_x(z))$.
+
+For the triangular based pyramid, `gen_pyrtri(n, p, h, l, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`), let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$.
+That is, $r(z) = r_t + (l-r_t)z/h$.
+A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
+The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v$.
+
+For the star based pyramid, `gen_pyrstar(n, p, h, rb)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`), let the radius at height $z$, $r(z)$, be such that the radius scales linearly from zero (tip) to the base radius $r_b$.
+That is, $r(z) = r_b\left(1 - z/h\right)$.
+
+Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
+$r_{\text{point}} \sim \sqrt{U(0, 1)}$. Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
+
+For all pyramid shapes, if $p > 3$, the remaining $p - 4$ dimensions (i.e., $X_4$ to $X_{p-1}$) are tapered toward zero with decreasing height and generated as: $X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1$.
+This tapering ensures that the structure narrows in all directions as it approaches the tip.
 
@@
 
-#### `gen_pyrrect()`
-
-The `gen_pyrrect(n, p, h, l_vec, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a rectangular base (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`).
-The pyramid is embedded in $p\text{-}D$, with a tip at height zero and base at height $h$.
-The shape tapers linearly from the base dimensions $(l_x, l_y)$ to a smaller rectangular section with side lengths $(2r_t, 2r_t)$ at the tip.
-Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.
-
-Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
-The final dimension $X_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$: $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$.
-Let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$: $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
-The first three coordinates are then defined as: $X_1 \sim U(-r_x(z),\ r_x(z))$, $X_2 \sim U(-r_y(z),\ r_y(z))$, and $X_3 \sim U(-r_x(z),\ r_x(z))$.
-For $p > 3$, the remaining $p - 4$ dimensions (i.e., $X_4$ to $X_{p-1}$) are tapered toward zero with decreasing height and generated as: $X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1$.
-This tapering ensures that the structure narrows in all directions as it approaches the tip.
-
 ```{r data-pyrrect, echo=TRUE}
@@
 
-#### `gen_pyrtri()`
-
-The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`).
-The structure is embedded in a $p\text{-}D$, with the tip located at height $0$ and the base at height $h$.
-The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
-
-Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
-The final coordinate $X_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$: $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$.
-Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$: $r(z) = r_t + (l - r_t)z/h$.
-
-A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
-The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v$.
-For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip: $X_j \sim U(-0.1, 0.1)\left(1 - z/h\right)$.
-
 ```{r data-pyrtri, echo=TRUE}
@@
 ```
-
-#### `gen_pyrstar()`
-
-The `gen_pyrstar(n, p, h, rb)` function generates a dataset of $n$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`).
-The pyramid extends vertically from height $z = 0$ (tip) to height $z = h$ (base), embedded in a $p$-dimensional space.
-The distribution concentrates more points near the base, and the hexagonal spread tapers toward the tip.
-
-Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points. The final coordinate $X_p$ encodes the height $z$: $X_p = z \sim U(0, h)$.
-
-The radius at height $z$ scales linearly from zero (tip) to the base radius $r_b$: $r(z) = r_b\left(1 - z/h\right)$.
-
-Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
-$r_{\text{point}} \sim \sqrt{U(0, 1)}$.
-Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
-For dimensions $j = 3, \dots, p-1$, values are tapered toward zero near the tip: $X_j \sim U(-0.1, 0.1)\left(1 - z/h \right)$.
 

### 2025-09-17 paper-cardinalR.Rmd
Updated gen_curvybranches

@@
 The underlying geometry lies in the first two dimensions, while the remaining $(p-2)$ dimensions contain Gaussian noise.
+
 Branch $1$ is generated from $X_1 \sim U(0, 1)$, and $X_2 = 0.1X_1 + X_1^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
@@
 This creates a steeper, leftward-facing curve in the left half-plane.
-Branches $3$ to $k$ are added iteratively.
-Each new branch $i$ begins at a randomly chosen point within a restricted horizontal band: $X_{1} \in [-0.15, 0.15]$ to ensure connectivity with earlier branches, spans a unit-length interval on $X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, and has the structure: $X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, \; s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}$.
-Here, $s_i$ is a sampled scale factor (slope-like term) controlling curvature, and $(x_{\text{start}}, y_{\text{start}})$ is the starting point from the existing structure.
+Branches $3$ to $k$ are then added iteratively.
+Each new branch $i$ begins at a randomly chosen starting point $(x_{\text{start}}, y_{\text{start}})$ from the existing structure, where the horizontal coordinate $x_{\text{start}}$ is restricted to $[-0.15, 0.15]$ to ensure connectivity with the initial branches.
+From this starting point, the branch extends over a unit-length interval $X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, with structure $X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, ; s_i \in {-2, -1.5, -1, -0.5, 0, 0.5, 1.5}$.
+Here, each branch $i$ is assigned a single scale factor $s_i$, sampled independently from the set, which controls the curvature of that branch.
+The parameter $(x_{\text{start}}, y_{\text{start}})$ is the starting point chosen from the existing structure.
 This construction yields $k$ spatially connected, nonlinear branches with varying curvature.

### 2025-09-17 paper-cardinalR.tex
Updated gen_curvybranches

@@
 The underlying geometry lies in the first two dimensions, while the remaining \((p-2)\) dimensions contain Gaussian noise.
+
 Branch \(1\) is generated from \(X_1 \sim U(0, 1)\), and \(X_2 = 0.1X_1 + X_1^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
@@
 This creates a steeper, leftward-facing curve in the left half-plane.
-Branches \(3\) to \(k\) are added iteratively.
-Each new branch \(i\) begins at a randomly chosen point within a restricted horizontal band: \(X_{1} \in [-0.15, 0.15]\) to ensure connectivity with earlier branches, spans a unit-length interval on \(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), and has the structure: \(X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, \; s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1.5\}\).
-Here, \(s_i\) is a sampled scale factor (slope-like term) controlling curvature, and \((x_{\text{start}}, y_{\text{start}})\) is the starting point from the existing structure.
+Branches \(3\) to \(k\) are then added iteratively.
+Each new branch \(i\) begins at a randomly chosen starting point \((x_{\text{start}}, y_{\text{start}})\) from the existing structure, where the horizontal coordinate \(x_{\text{start}}\) is restricted to \([-0.15, 0.15]\) to ensure connectivity with the initial branches.
+From this starting point, the branch extends over a unit-length interval \(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), with structure \(X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, ; s_i \in {-2, -1.5, -1, -0.5, 0, 0.5, 1.5}\).
+Here, each branch \(i\) is assigned a single scale factor \(s_i\), sampled independently from the set, which controls the curvature of that branch.
+The parameter \((x_{\text{start}}, y_{\text{start}})\) is the starting point chosen from the existing structure.
 This construction yields \(k\) spatially connected, nonlinear branches with varying curvature.
@@
 A pyramid structure (Figure \ref{fig:pyr-proj}) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
-The functions provided allow users to generate pyramids with various base shapes---including rectangular, triangular, and star-shaped bases---with options for sharp or blunted apexes.
-Additionally, we offer functionality for creating pyramids containing structured holes, enabling the study of non-convex and sparse regions.
-These designs are valuable for testing the ability of dimension reduction and clustering algorithms to preserve and detect complex geometric features.
+The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
+Additionally, similar to the cube generator, it is possible to create a pyramids containing structured holes, enabling the study of non-convex and sparse regions.
 Table \ref{tab:pyramid-tb-pdf} summarizes these functions.
+
+Let \(X_1, \dots, X_p\) denote the coordinates of the generated points.
+For the rectangular and triangular based pyramid generator functions, the final dimension, \(X_p\), encodes the height of each point and is drawn from an exponential distribution capped at the maximum height \(h\).
+That is, \[X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\].
+This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
+That is, \(X_p = z \sim U(0, h)\).
+
+The remaining dimensions are based on the specific pyramid shape.
+For the rectangular based pyramid, \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} (Figure \ref{fig:pyr-proj} a), let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\).
+That is, \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
+The first three coordinates are then defined as: \(X_1 \sim U(-r_x(z),\ r_x(z))\), \(X_2 \sim U(-r_y(z),\ r_y(z))\), and \(X_3 \sim U(-r_x(z),\ r_x(z))\).
+
+For the triangular based pyramid, \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} (Figure \ref{fig:pyr-proj} b), let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\).
+That is, \(r(z) = r_t + (l-r_t)z/h\).
+A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section: \(u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v\).
+The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z)u\), and \(X_3 = r(z)v\).
+
+For the star based pyramid, \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} (Figure \ref{fig:pyr-proj} c), let the radius at height \(z\), \(r(z)\), be such that the radius scales linearly from zero (tip) to the base radius \(r_b\).
+That is, \(r(z) = r_b\left(1 - z/h\right)\).
+
+Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
+\(r_{\text{point}} \sim \sqrt{U(0, 1)}\). Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
+
+For all pyramid shapes, if \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are tapered toward zero with decreasing height and generated as: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1\).
+This tapering ensures that the structure narrows in all directions as it approaches the tip.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_pyrrect()}}{gen\_pyrrect()}}\label{gen_pyrrect}
-
-The \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a rectangular base (Figure \ref{fig:pyr-proj} a).
-The pyramid is embedded in \(p\text{-}D\), with a tip at height zero and base at height \(h\).
-The shape tapers linearly from the base dimensions \((l_x, l_y)\) to a smaller rectangular section with side lengths \((2r_t, 2r_t)\) at the tip.
-Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.
-
-Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
-The final dimension \(X_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\): \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\).
-Let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\): \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
-The first three coordinates are then defined as: \(X_1 \sim U(-r_x(z),\ r_x(z))\), \(X_2 \sim U(-r_y(z),\ r_y(z))\), and \(X_3 \sim U(-r_x(z),\ r_x(z))\).
-For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are tapered toward zero with decreasing height and generated as: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1\).
-This tapering ensures that the structure narrows in all directions as it approaches the tip.
-
 \begin{verbatim}
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_pyrtri()}}{gen\_pyrtri()}}\label{gen_pyrtri}
-
-The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section (Figure \ref{fig:pyr-proj} b).
-The structure is embedded in a \(p\text{-}D\), with the tip located at height \(0\) and the base at height \(h\).
-The triangle expands linearly in size from tip to base, with more points concentrated near the tip.
-
-Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
-The final coordinate \(X_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\): \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\).
-Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\): \(r(z) = r_t + (l - r_t)z/h\).
-
-A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section: \(u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v\).
-The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z)u\), and \(X_3 = r(z)v\).
-For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h\right)\).
-
 \begin{verbatim}
@@
 \end{verbatim}
-
-\paragraph{\texorpdfstring{\texttt{gen\_pyrstar()}}{gen\_pyrstar()}}\label{gen_pyrstar}
-
-The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-shaped structure with a hexagonal base (Figure \ref{fig:pyr-proj} c).
-The pyramid extends vertically from height \(z = 0\) (tip) to height \(z = h\) (base), embedded in a \(p\)-dimensional space.
-The distribution concentrates more points near the base, and the hexagonal spread tapers toward the tip.
-
-Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points. The final coordinate \(X_p\) encodes the height \(z\): \(X_p = z \sim U(0, h)\).
-
-The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r_b\): \(r(z) = r_b\left(1 - z/h\right)\).
-
-Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
-\(r_{\text{point}} \sim \sqrt{U(0, 1)}\).
-Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
-For dimensions \(j = 3, \dots, p-1\), values are tapered toward zero near the tip: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h \right)\).
 

### 2025-09-17 paper-cardinalR.Rmd
Updated gen_multicluster text

@@
 
-The main function of the package is `gen_multicluster()`.
-This function generates clusters of various shapes, allowing users to specify the number of points in each cluster, as well as their locations, scaling, and rotations across specific dimensions.
-Additionally, users can add background noise into the generated data by using the `is_bkg` option.
+The main function of the package is `gen_multicluster()`, which generates datasets consisting of multiple clusters with user-specified characteristics.
+Users can control the number of clusters (`k`), the number of points in each cluster (`n`), and the dimensionality of the space (`p`).
+Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube), be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (`loc`).
+The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures.
+An optional argument, `is_bkg`, adds background noise drawn from a multivariate normal distribution centered on the dataset’s overall mean with standard deviations matching the observed spread.
 

### 2025-09-17 paper-cardinalR.tex
Updated gen_multicluster text

@@
 
-The main function of the package is \texttt{gen\_multicluster()}.
-This function generates clusters of various shapes, allowing users to specify the number of points in each cluster, as well as their locations, scaling, and rotations across specific dimensions.
-Additionally, users can add background noise into the generated data by using the \texttt{is\_bkg} option.
+The main function of the package is \texttt{gen\_multicluster()}, which generates datasets consisting of multiple clusters with user-specified characteristics.
+Users can control the number of clusters (\texttt{k}), the number of points in each cluster (\texttt{n}), and the dimensionality of the space (\texttt{p}).
+Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube), be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (\texttt{loc}).
+The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures.
+An optional argument, \texttt{is\_bkg}, adds background noise drawn from a multivariate normal distribution centered on the dataset's overall mean with standard deviations matching the observed spread.
 

### 2025-09-17 paper-cardinalR.Rmd
Added reference for branching data

@@
 
-A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology.
+A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (@trapnell2014).
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points, and number of branches.

### 2025-09-17 paper-cardinalR.tex
Added reference for branching data

@@
 
-A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology.
+A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (\citet{trapnell2014}).
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points, and number of branches.

### 2025-09-19 paper-cardinalR.tex
Need to add a section for All noise in one
Updated mobius section with general noise dimensions

@@
 This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable one-sided surface.
-For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) Mobius into \(p\text{-}D\), where \(j = 4, \dots, p\).
+For \(p > 3\), additional noise dimension are added to embed the \(3\text{-}D\) Mobius into \(p\text{-}D\).
 
@@
 \end{figure}
+
+\subsubsection{Generate noise dimensions}\label{generate-noise-dimensions}
 

### 2025-09-19 paper-cardinalR.Rmd
Need to add a section for All noise in one
Updated mobius section with general noise dimensions
Updated pyramid one with noise dimensions

@@
 This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface.
-For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ Mobius into $p\text{-}D$, where $j = 4, \dots, p$.
+For $p > 3$, additional noise dimension are added to embed the $3\text{-}D$ Mobius into $p\text{-}D$.
 
@@
 
-For all pyramid shapes, if $p > 3$, the remaining $p - 4$ dimensions (i.e., $X_4$ to $X_{p-1}$) are tapered toward zero with decreasing height and generated as: $X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1$.
-This tapering ensures that the structure narrows in all directions as it approaches the tip.
+For all pyramid shapes, if $p > 3$, the remaining $p - 3$ dimensions (i.e., $X_4$ to $X_{p-1}$) are additional noise.
 
@@
 ```
+
+### Generate noise dimensions
 

### 2025-09-19 paper-cardinalR.tex
Updated pyramid one with noise dimensions

@@
 
-For all pyramid shapes, if \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are tapered toward zero with decreasing height and generated as: \(X_j \sim U(-0.1, 0.1)\left(1 - z/h\right), \quad \text{for } j = 4, \dots, p-1\).
-This tapering ensures that the structure narrows in all directions as it approaches the tip.
+For all pyramid shapes, if \(p > 3\), the remaining \(p - 3\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are additional noise.
 

### 2025-09-19 paper-cardinalR.Rmd
Updated the introduction sentence

@@
 
-By providing these aspects, `cardinalR` provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
-This bridges the gap between geometric foundations and the flexible generation of complex synthetic data.
+By providing these aspects, `cardinalR` provides researchers with a method to generate more explainable and challenging high-dimensional clustering synthetic datasets focused to the specific needs of evaluating algorithms.
 

### 2025-09-19 paper-cardinalR.tex
Updated the introduction sentence

@@
 
-By providing these aspects, \texttt{cardinalR} provides researchers with a method to generate more explainable and challenging synthetic datasets focused to the specific needs of evaluating algorithms in high-dimensions.
-This bridges the gap between geometric foundations and the flexible generation of complex synthetic data.
+By providing these aspects, \texttt{cardinalR} provides researchers with a method to generate more explainable and challenging high-dimensional clustering synthetic datasets focused to the specific needs of evaluating algorithms.
 

### 2025-09-19 paper-cardinalR.tex
more clear about the p

@@
 The main function of the package is \texttt{gen\_multicluster()}, which generates datasets consisting of multiple clusters with user-specified characteristics.
-Users can control the number of clusters (\texttt{k}), the number of points in each cluster (\texttt{n}), and the dimensionality of the space (\texttt{p}).
+Users can control the number of clusters (\texttt{k}), the number of points in each cluster (\texttt{n}), and the dimensionality of the space (\texttt{p}) for all the clusters.
 Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube), be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (\texttt{loc}).

### 2025-09-19 paper-cardinalR.Rmd
more clear about the p
Remove 0.3 value

@@
 The main function of the package is `gen_multicluster()`, which generates datasets consisting of multiple clusters with user-specified characteristics.
-Users can control the number of clusters (`k`), the number of points in each cluster (`n`), and the dimensionality of the space (`p`).
+Users can control the number of clusters (`k`), the number of points in each cluster (`n`), and the dimensionality of the space (`p`) for all the clusters.
 Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube), be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (`loc`).
@@
 
+### Shape generators
+
 ### Branching
@@
 To simulate missing regions on the manifold, a fixed anchor point $(0, 1, 0, \ldots)$ is defined in $p\text{-}D$.
-All observations within a Euclidean distance of $\sqrt{0.3} \approx 0.5477$ from the anchor are removed if $\sum_{j=1}^p (X_j - a_j)^2 \leq 0.3$.
+All observations within a Euclidean distance of $\sqrt{0.3}$ from the anchor are removed if $\sum_{j=1}^p (X_j - a_j)^2 \leq 0.3$.
 This exclusion creates a hole in the manifold centered near the middle vertical region of the S-curve.

### 2025-09-19 paper-cardinalR.Rmd
Updated cone additional noise dimensions

@@
 The radial coordinates are scaled by the height-dependent radius $r$, producing the conical taper.
-In three dimensions ($p = 3$), this results in a classical 3-D cone, while for $p > 3$, additional angular components allow a smooth extension into higher dimensions, preserving the conical shape within multi-dimensional space.
+In three dimensions ($p = 3$), this results in a classical $3\text{-}D$ cone, while for $p > 3$, additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
 
@@
 The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
-Additionally, similar to the cube generator, it is possible to create a pyramids containing structured holes, enabling the study of non-convex and sparse regions.
+Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:pyramid-tb-html)', '\\@ref(tab:pyramid-tb-pdf)'))` summarizes these functions.
@@
 For the rectangular and triangular based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
-That is, $$X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$$.
-This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+That is, $$X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).$$ This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
 For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
@@
 That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
-The first three coordinates are then defined as: $X_1 \sim U(-r_x(z),\ r_x(z))$, $X_2 \sim U(-r_y(z),\ r_y(z))$, and $X_3 \sim U(-r_x(z),\ r_x(z))$.
+The first three coordinates are then defined as: $X_1 \sim U(-r_x(z),\ r_x(z))$, $X_2 \sim U(-r_y(z),\ r_y(z))$, and $X_3 \sim U(-r_x(z),\ r_x(z)).$
 
@@
 A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
-The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v$.
+The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v.$
 
@@
 
-#### `gen_pyrholes()`
-
-The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p\text{-}D$ simplex using a chaotic attractor-like midpoint algorithm (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`).
+#### `gen_pyrfrac()`
+
+The `gen_pyrfrac(n, p)` function generates $n$ points embedded in a $p\text{-}D$ simplex using a chaotic attractor-like midpoint algorithm (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`).
 The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a "Sierpinski-like pyramid" in high dimensions.
@@
 ```{r data-pyrholes, echo=TRUE}
-pyrholes <- gen_pyrholes(n = 1000, p = 4)
+pyrholes <- gen_pyrfrac(n = 1000, p = 4)
 ```

### 2025-09-19 paper-cardinalR.tex
Remove 0.3 value
Updated cone additional noise dimensions

@@
 
+\subsubsection{Shape generators}\label{shape-generators}
+
 \subsubsection{Branching}\label{branching}
@@
 The radial coordinates are scaled by the height-dependent radius \(r\), producing the conical taper.
-In three dimensions (\(p = 3\)), this results in a classical 3-D cone, while for \(p > 3\), additional angular components allow a smooth extension into higher dimensions, preserving the conical shape within multi-dimensional space.
+In three dimensions (\(p = 3\)), this results in a classical \(3\text{-}D\) cone, while for \(p > 3\), additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
 
@@
 The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
-Additionally, similar to the cube generator, it is possible to create a pyramids containing structured holes, enabling the study of non-convex and sparse regions.
+Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
 Table \ref{tab:pyramid-tb-pdf} summarizes these functions.
@@
 For the rectangular and triangular based pyramid generator functions, the final dimension, \(X_p\), encodes the height of each point and is drawn from an exponential distribution capped at the maximum height \(h\).
-That is, \[X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\].
-This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+That is, \[X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).\] This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
 For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
@@
 That is, \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
-The first three coordinates are then defined as: \(X_1 \sim U(-r_x(z),\ r_x(z))\), \(X_2 \sim U(-r_y(z),\ r_y(z))\), and \(X_3 \sim U(-r_x(z),\ r_x(z))\).
+The first three coordinates are then defined as: \(X_1 \sim U(-r_x(z),\ r_x(z))\), \(X_2 \sim U(-r_y(z),\ r_y(z))\), and \(X_3 \sim U(-r_x(z),\ r_x(z)).\)
 
@@
 A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section: \(u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v\).
-The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z)u\), and \(X_3 = r(z)v\).
+The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z)u\), and \(X_3 = r(z)v.\)
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_pyrholes()}}{gen\_pyrholes()}}\label{gen_pyrholes}
-
-The \texttt{gen\_pyrholes(n,\ p)} function generates \(n\) points embedded in a \(p\text{-}D\) simplex using a chaotic attractor-like midpoint algorithm (Figure \ref{fig:pyr-proj} d).
+\paragraph{\texorpdfstring{\texttt{gen\_pyrfrac()}}{gen\_pyrfrac()}}\label{gen_pyrfrac}
+
+The \texttt{gen\_pyrfrac(n,\ p)} function generates \(n\) points embedded in a \(p\text{-}D\) simplex using a chaotic attractor-like midpoint algorithm (Figure \ref{fig:pyr-proj} d).
 The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a ``Sierpinski-like pyramid'' in high dimensions.
@@
 \begin{verbatim}
-pyrholes <- gen_pyrholes(n = 1000, p = 4)
+pyrholes <- gen_pyrfrac(n = 1000, p = 4)
 \end{verbatim}
@@
 To simulate missing regions on the manifold, a fixed anchor point \((0, 1, 0, \ldots)\) is defined in \(p\text{-}D\).
-All observations within a Euclidean distance of \(\sqrt{0.3} \approx 0.5477\) from the anchor are removed if \(\sum_{j=1}^p (X_j - a_j)^2 \leq 0.3\).
+All observations within a Euclidean distance of \(\sqrt{0.3}\) from the anchor are removed if \(\sum_{j=1}^p (X_j - a_j)^2 \leq 0.3\).
 This exclusion creates a hole in the manifold centered near the middle vertical region of the S-curve.

### 2025-09-19 paper-cardinalR.R

@@ -1929,7 +1929,7 @@ pyrstar_proj3 <- plot_proj(
 
 
 ## ----data-pyrholes, echo=TRUE-------------------------------------------------
-pyrholes <- gen_pyrholes(n = 1000, p = 4)
+pyrholes <- gen_pyrfrac(n = 1000, p = 4)
 
 
 ## -----------------------------------------------------------------------------

### 2025-09-19 paper-cardinalR.Rmd
Updated Swissroll noise
Updated pyramid section

@@
 
-For all pyramid shapes, if $p > 3$, the remaining $p - 3$ dimensions (i.e., $X_4$ to $X_{p-1}$) are additional noise.
+For all the above pyramid shapes, if $p > 3$, the remaining $p - 3$ dimensions (i.e., $X_4$ to $X_{p-1}$) are additional noise.
+
+Finally, for the Sierpinski-like pyramid (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`), let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
+The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
+Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
+At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
+This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
+The points remain bounded inside the convex hull of the simplex.
+The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
 
@@
                              "gen_pyrstar",
-                             "gen_pyrholes"), 
+                             "gen_pyrfrac"), 
                         exp = c("Rectangular-base pyramid, with a sharp or blunted apex.",
@@
 ```
-
-#### `gen_pyrfrac()`
-
-The `gen_pyrfrac(n, p)` function generates $n$ points embedded in a $p\text{-}D$ simplex using a chaotic attractor-like midpoint algorithm (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`).
-The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a "Sierpinski-like pyramid" in high dimensions.
-
-Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
-The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
-
-Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
-At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
-This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
-The points remain bounded inside the convex hull of the simplex.
-The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
 
@@
 The $X_1 = t \cos(t)$, $X_2 = t \sin(t)$, $X_3 \sim U(w_1, w_2), \text{where } t \sim U(0, 3\pi)$.
-For $p > 3$, the remaining $p - 3$ dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.
+For $p > 3$, the remaining $p - 3$ dimensions are filled with noise to simulate high-dimensional complexity.
 

### 2025-09-19 paper-cardinalR.tex
Updated Swissroll noise
Updated pyramid section

@@
 
-For all pyramid shapes, if \(p > 3\), the remaining \(p - 3\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are additional noise.
-
-\begin{table}
-
-\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_pyrrect & Rectangular-base pyramid, with a sharp or blunted apex.\\
-gen\_pyrtri & Triangular-base pyramid, with a sharp or blunted apex.\\
-gen\_pyrstar & Star-shaped base pyramid, with a sharp or blunted apex.\\
-gen\_pyrholes & Pyramid containing triangular pyramid-shaped holes.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{verbatim}
-pyrrect <- gen_pyrrect(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{verbatim}
-pyrtri <- gen_pyrtri(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{verbatim}
-pyrstar <- gen_pyrstar(n = 1000, p = 4)
-\end{verbatim}
-
-\paragraph{\texorpdfstring{\texttt{gen\_pyrfrac()}}{gen\_pyrfrac()}}\label{gen_pyrfrac}
-
-The \texttt{gen\_pyrfrac(n,\ p)} function generates \(n\) points embedded in a \(p\text{-}D\) simplex using a chaotic attractor-like midpoint algorithm (Figure \ref{fig:pyr-proj} d).
-The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a ``Sierpinski-like pyramid'' in high dimensions.
-
-Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
+For all the above pyramid shapes, if \(p > 3\), the remaining \(p - 3\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are additional noise.
+
+Finally, for the Sierpinski-like pyramid (Figure \ref{fig:pyr-proj} d), let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
 The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution: \(T_0 \sim U(0, 1)^p\).
-
 Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\text{-}D\) simplex.
@@
 The final output is a \(n \times p\) matrix where each row represents a point: \(X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}\).
+
+\begin{table}
+
+\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_pyrrect & Rectangular-base pyramid, with a sharp or blunted apex.\\
+gen\_pyrtri & Triangular-base pyramid, with a sharp or blunted apex.\\
+gen\_pyrstar & Star-shaped base pyramid, with a sharp or blunted apex.\\
+gen\_pyrfrac & Pyramid containing triangular pyramid-shaped holes.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{verbatim}
+pyrtri <- gen_pyrtri(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{verbatim}
+pyrstar <- gen_pyrstar(n = 1000, p = 4)
+\end{verbatim}
 
@@
 The \(X_1 = t \cos(t)\), \(X_2 = t \sin(t)\), \(X_3 \sim U(w_1, w_2), \text{where } t \sim U(0, 3\pi)\).
-For \(p > 3\), the remaining \(p - 3\) dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.
+For \(p > 3\), the remaining \(p - 3\) dimensions are filled with noise to simulate high-dimensional complexity.
 

### 2025-09-19 paper-cardinalR.R

@@ -1639,7 +1639,7 @@ cubic_proj1 + cubic_proj2 + cubic_proj3 +
 pyramid_tb <- tibble(fun = c("gen_pyrrect",
                              "gen_pyrtri",
                              "gen_pyrstar",
-                             "gen_pyrholes"), 
+                             "gen_pyrfrac"), 
                         exp = c("Rectangular-base pyramid, with a sharp or blunted apex.",
           "Triangular-base pyramid, with a sharp or blunted apex.",
           "Star-shaped base pyramid, with a sharp or blunted apex.",

### 2025-09-19 paper-cardinalR.Rmd
Added overview paragraph for shape generators

@@
 
+The shape generators form the foundation of the package, providing a collection of functions to create synthetic data structures based on simple, well-defined geometric structures.
+These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures.
+If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
+Users can specify how these noise dimensions are generated (e.g., Gaussian, uniform), offering control over the embedding process.
+All shape generators allow the user to define the number of points and dimensions, and most include additional arguments to customize specific characteristics of the structure.
+
 ### Branching

### 2025-09-19 paper-cardinalR.tex
Added overview paragraph for shape generators

@@
 \subsubsection{Shape generators}\label{shape-generators}
+
+The shape generators form the foundation of the package, providing a collection of functions to create synthetic data structures based on simple, well-defined geometric structures.
+These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures.
+If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
+Users can specify how these noise dimensions are generated (e.g., Gaussian, uniform), offering control over the embedding process.
+All shape generators allow the user to define the number of points and dimensions, and most include additional arguments to customize specific characteristics of the structure.
 

### 2025-09-19 paper-cardinalR.Rmd
Added function for gen_pyrfrac

@@
 
-### Cone
+#### Cone
 
@@
 
-### Cube
+#### Cube
 
@@
 
-### Gaussian
+#### Gaussian
 
@@
 
-### Linear
+#### Linear
 
@@
 
-### Mobius
+#### Mobius
 
@@
 
-### Pyramid
+#### Pyramid
 
@@
 
-Finally, for the Sierpinski-like pyramid (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`), let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
+Finally, for the Sierpinski-like pyramid, `gen_pyrfrac(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`), let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
 The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
@@
 
-### Swiss Roll
+#### Swiss Roll
 

### 2025-09-19 paper-cardinalR.tex
Added function for gen_pyrfrac

@@
 
-\subsubsection{Cone}\label{cone}
+\paragraph{Cone}\label{cone}
 
@@
 
-\subsubsection{Cube}\label{cube}
+\paragraph{Cube}\label{cube}
 
@@
 
-\subsubsection{Gaussian}\label{gaussian}
+\paragraph{Gaussian}\label{gaussian}
 
@@
 
-\subsubsection{Linear}\label{linear}
+\paragraph{Linear}\label{linear}
 
@@
 
-\subsubsection{Mobius}\label{mobius}
+\paragraph{Mobius}\label{mobius}
 
@@
 
-\subsubsection{Pyramid}\label{pyramid}
+\paragraph{Pyramid}\label{pyramid}
 
@@
 
-Finally, for the Sierpinski-like pyramid (Figure \ref{fig:pyr-proj} d), let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
+Finally, for the Sierpinski-like pyramid, \texttt{gen\_pyrfrac(n,\ p)} (Figure \ref{fig:pyr-proj} d), let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
 The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution: \(T_0 \sim U(0, 1)^p\).
@@
 
-\subsubsection{Swiss Roll}\label{swiss-roll}
+\paragraph{Swiss Roll}\label{swiss-roll}
 

### 2025-09-19 paper-cardinalR.Rmd
Updated S-curve section

@@
 
-An S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) simulates data that lies along a smooth, non-linear manifold, making it a classic example for testing the performance of non-linear dimension reduction methods.
-This package provides functions to generate standard S-curve shapes as well as variants containing holes to introduce missing or incomplete regions.
+An S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) simulates data that lies along a smooth, non-linear manifold.
+The functions generate both the standard S-curve shape and, similar to the cube generator, an S-curve variant with structured hole that introduce missing or incomplete region.
 These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
@@
 
-#### `gen_scurve()`
-
-To simulate an S-curve structure in a higher-dimensional space, we define the function `gen_scurve(n, p)`, which generates $n$ observations in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`).
-
-The $3\text{-}D$ geometry is constructed by introducing a latent parameter, $\theta \sim U\left(-3\pi/2, 3\pi/2\right)$.
+For the S-curve structure, `gen_scurve(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`), the $3\text{-}D$ geometry is constructed by introducing a latent parameter, $\theta \sim U\left(-3\pi/2, 3\pi/2\right)$.
 This parameter controls the curvature of the manifold.
@@
 This configuration creates a horizontally curled shape in $(X_1, X_3)$, with additional band thickness in the $X_2$ direction.
-For $p > 3$, additional noise dimensions are appended introducing structured, wavy noise.
+For $p > 3$, additional noise dimensions are appended to embed the structure in higher dimensions.
 
@@
 
-#### `gen_scurvehole()`
-
-To simulate a variant of the S-curve structure with a removed region ("hole"), we define the function `gen_scurvehole(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) b"} else { "Figure \\@ref(fig:scurve-proj) b"})`).
-This function produces $n$ observations in $p\text{-}D$ where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
-A subset of observations near a designated anchor point is excluded to introduce a hole in the manifold.
-
-To simulate missing regions on the manifold, a fixed anchor point $(0, 1, 0, \ldots)$ is defined in $p\text{-}D$.
-All observations within a Euclidean distance of $\sqrt{0.3}$ from the anchor are removed if $\sum_{j=1}^p (X_j - a_j)^2 \leq 0.3$.
-This exclusion creates a hole in the manifold centered near the middle vertical region of the S-curve.
+To introduce missing or incomplete regions on the manifold, `gen_scurvehole(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) b"} else { "Figure \\@ref(fig:scurve-proj) b"})`) removes a subset of points near a designated anchor, creating a hole in the S-curve.
+This produces a hollowed manifold while preserving the overall curvature and structure of the S-curve.
 

### 2025-09-19 paper-cardinalR.tex
Updated S-curve section

@@
 
-An S-curve structure (Figure \ref{fig:scurve-proj}) simulates data that lies along a smooth, non-linear manifold, making it a classic example for testing the performance of non-linear dimension reduction methods.
-This package provides functions to generate standard S-curve shapes as well as variants containing holes to introduce missing or incomplete regions.
+An S-curve structure (Figure \ref{fig:scurve-proj}) simulates data that lies along a smooth, non-linear manifold.
+The functions generate both the standard S-curve shape and, similar to the cube generator, an S-curve variant with structured hole that introduce missing or incomplete region.
 These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_scurve()}}{gen\_scurve()}}\label{gen_scurve}
-
-To simulate an S-curve structure in a higher-dimensional space, we define the function \texttt{gen\_scurve(n,\ p)}, which generates \(n\) observations in \(p\text{-}D\) (Figure \ref{fig:scurve-proj} a).
-
-The \(3\text{-}D\) geometry is constructed by introducing a latent parameter, \(\theta \sim U\left(-3\pi/2, 3\pi/2\right)\).
+For the S-curve structure, \texttt{gen\_scurve(n,\ p)} (Figure \ref{fig:scurve-proj} a), the \(3\text{-}D\) geometry is constructed by introducing a latent parameter, \(\theta \sim U\left(-3\pi/2, 3\pi/2\right)\).
 This parameter controls the curvature of the manifold.
@@
 This configuration creates a horizontally curled shape in \((X_1, X_3)\), with additional band thickness in the \(X_2\) direction.
-For \(p > 3\), additional noise dimensions are appended introducing structured, wavy noise.
+For \(p > 3\), additional noise dimensions are appended to embed the structure in higher dimensions.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_scurvehole()}}{gen\_scurvehole()}}\label{gen_scurvehole}
-
-To simulate a variant of the S-curve structure with a removed region (``hole''), we define the function \texttt{gen\_scurvehole(n,\ p)} (Figure \ref{fig:scurve-proj} b).
-This function produces \(n\) observations in \(p\text{-}D\) where the first three dimensions describe the S-curve manifold, and remaining dimensions add low-variance Gaussian noise.
-A subset of observations near a designated anchor point is excluded to introduce a hole in the manifold.
-
-To simulate missing regions on the manifold, a fixed anchor point \((0, 1, 0, \ldots)\) is defined in \(p\text{-}D\).
-All observations within a Euclidean distance of \(\sqrt{0.3}\) from the anchor are removed if \(\sum_{j=1}^p (X_j - a_j)^2 \leq 0.3\).
-This exclusion creates a hole in the manifold centered near the middle vertical region of the S-curve.
+To introduce missing or incomplete regions on the manifold, \texttt{gen\_scurvehole(n,\ p)} (Figure \ref{fig:scurve-proj} b) removes a subset of points near a designated anchor, creating a hole in the S-curve.
+This produces a hollowed manifold while preserving the overall curvature and structure of the S-curve.
 

### 2025-09-19 paper-cardinalR.Rmd
Updated pyramid section and added fixed radius value

@@
 
+```{r data-pyrrect, echo=TRUE}
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+```
+
+```{r}
+pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r pyrrect-proj1}
+scaled_data <- scale_data_manual(pyrrect)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrrect_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrrect-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
+                                              threshold = 0.08))
+
+pyrrect_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrrect-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrrect_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.75), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
 For the triangular based pyramid, `gen_pyrtri(n, p, h, l, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`), let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$.
@@
 
+```{r data-pyrtri, echo=TRUE}
+pyrtri <- gen_pyrtri(n = 1000, p = 4)
+```
+
+```{r}
+pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r pyrtri-proj1}
+scaled_data <- scale_data_manual(pyrtri)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.022))
+
+pyrtri_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.3), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrtri-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.03))
+
+pyrtri_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.4), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrtri-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrtri_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
 For the star based pyramid, `gen_pyrstar(n, p, h, rb)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`), let the radius at height $z$, $r(z)$, be such that the radius scales linearly from zero (tip) to the base radius $r_b$.
@@
 $r_{\text{point}} \sim \sqrt{U(0, 1)}$. Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
+
+```{r data-pyrstar, echo=TRUE}
+pyrstar <- gen_pyrstar(n = 1000, p = 4)
+```
+
+```{r}
+pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r pyrstar-proj1}
+scaled_data <- scale_data_manual(pyrstar)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrstar_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrstar-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrstar_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.4), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrstar-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+pyrstar_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "c3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
 
@@
 The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
+
+```{r data-pyrholes, echo=TRUE}
+pyrholes <- gen_pyrfrac(n = 1000, p = 4)
+```
+
+```{r}
+pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r pyrholes-proj1}
+scaled_data <- scale_data_manual(pyrholes)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrholes_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "d1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrholes-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrholes_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.75), 
+  title = "d2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrholes-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+pyrholes_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "d3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
 
@@
 
-```{r data-pyrrect, echo=TRUE}
-pyrrect <- gen_pyrrect(n = 1000, p = 4)
-```
-
-```{r}
-pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r pyrrect-proj1}
-scaled_data <- scale_data_manual(pyrrect)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrrect_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrrect-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.7, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.08))
-
-pyrrect_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrrect-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-pyrrect_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.75), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r data-pyrtri, echo=TRUE}
-pyrtri <- gen_pyrtri(n = 1000, p = 4)
-```
-
-```{r}
-pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r pyrtri-proj1}
-scaled_data <- scale_data_manual(pyrtri)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.022))
-
-pyrtri_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.3), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrtri-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.03))
-
-pyrtri_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.4), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrtri-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-pyrtri_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r data-pyrstar, echo=TRUE}
-pyrstar <- gen_pyrstar(n = 1000, p = 4)
-```
-
-```{r}
-pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r pyrstar-proj1}
-scaled_data <- scale_data_manual(pyrstar)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrstar_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrstar-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-pyrstar_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.4), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrstar-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-pyrstar_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r data-pyrholes, echo=TRUE}
-pyrholes <- gen_pyrfrac(n = 1000, p = 4)
-```
-
-```{r}
-pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r pyrholes-proj1}
-scaled_data <- scale_data_manual(pyrholes)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrholes_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrholes-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-pyrholes_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.75), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrholes-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-pyrholes_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
 ```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$."}
@@
 An S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) simulates data that lies along a smooth, non-linear manifold.
-The functions generate both the standard S-curve shape and, similar to the cube generator, an S-curve variant with structured hole that introduce missing or incomplete region.
+The functions generate both the standard S-curve shape and, a S-curve variant with structured hole that introduce missing or incomplete region.
 These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
@@
 
-To introduce missing or incomplete regions on the manifold, `gen_scurvehole(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) b"} else { "Figure \\@ref(fig:scurve-proj) b"})`) removes a subset of points near a designated anchor, creating a hole in the S-curve.
-This produces a hollowed manifold while preserving the overall curvature and structure of the S-curve.
+To introduce missing or incomplete regions on the manifold, `gen_scurvehole(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) b"} else { "Figure \\@ref(fig:scurve-proj) b"})`) removes points in a localized region centered around the middle vertical section of the S-curve.
+Following a similar approach as `gen_cubehole(n, p, r_hole)`, all observations within a fixed radius ($\sqrt{0.3}$) of the anchor point are excluded, creating a hole in the manifold while preserving the overall S-curve structure.
 

### 2025-09-19 paper-cardinalR.tex
Updated pyramid section and added fixed radius value

@@
 
+\begin{verbatim}
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+\end{verbatim}
+
 For the triangular based pyramid, \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} (Figure \ref{fig:pyr-proj} b), let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\).
@@
 
+\begin{verbatim}
+pyrtri <- gen_pyrtri(n = 1000, p = 4)
+\end{verbatim}
+
 For the star based pyramid, \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} (Figure \ref{fig:pyr-proj} c), let the radius at height \(z\), \(r(z)\), be such that the radius scales linearly from zero (tip) to the base radius \(r_b\).
@@
 \(r_{\text{point}} \sim \sqrt{U(0, 1)}\). Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
+
+\begin{verbatim}
+pyrstar <- gen_pyrstar(n = 1000, p = 4)
+\end{verbatim}
 
@@
 
+\begin{verbatim}
+pyrholes <- gen_pyrfrac(n = 1000, p = 4)
+\end{verbatim}
+
 \begin{table}
@@
 
-\begin{verbatim}
-pyrrect <- gen_pyrrect(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{verbatim}
-pyrtri <- gen_pyrtri(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{verbatim}
-pyrstar <- gen_pyrstar(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{verbatim}
-pyrholes <- gen_pyrfrac(n = 1000, p = 4)
-\end{verbatim}
-
 \begin{figure}[!]
@@
 An S-curve structure (Figure \ref{fig:scurve-proj}) simulates data that lies along a smooth, non-linear manifold.
-The functions generate both the standard S-curve shape and, similar to the cube generator, an S-curve variant with structured hole that introduce missing or incomplete region.
+The functions generate both the standard S-curve shape and, a S-curve variant with structured hole that introduce missing or incomplete region.
 These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
@@
 
-To introduce missing or incomplete regions on the manifold, \texttt{gen\_scurvehole(n,\ p)} (Figure \ref{fig:scurve-proj} b) removes a subset of points near a designated anchor, creating a hole in the S-curve.
-This produces a hollowed manifold while preserving the overall curvature and structure of the S-curve.
+To introduce missing or incomplete regions on the manifold, \texttt{gen\_scurvehole(n,\ p)} (Figure \ref{fig:scurve-proj} b) removes points in a localized region centered around the middle vertical section of the S-curve.
+Following a similar approach as \texttt{gen\_cubehole(n,\ p,\ r\_hole)}, all observations within a fixed radius (\(\sqrt{0.3}\)) of the anchor point are excluded, creating a hole in the manifold while preserving the overall S-curve structure.
 

### 2025-09-19 paper-cardinalR.Rmd
Add cube functions
Updated polynomial section

@@
 
+```{r data-gridcube, echo=TRUE}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+```
+
+```{r}
+gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r gridcube-proj1}
+scaled_data <- scale_data_manual(gridcube)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
+                                              threshold = 0.022))
+
+gridcube_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.83, 0.8), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridcube-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.8,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
+                                              threshold = 0.05))
+
+gridcube_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-1, 1), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridcube-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
+                                              threshold = 0.016))
+
+gridcube_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.8, 0.8), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
 An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
 The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) b"} else { "Figure \\@ref(fig:cube-proj) b"})`).
+
+```{r data-unifcube, echo=TRUE}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+```
+
+```{r}
+unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r unifcube-proj1}
+scaled_data <- scale_data_manual(unifcube)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.022))
+
+unifcube_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.9, 0.8), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifcube-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.04))
+
+unifcube_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.9, 0.8), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifcube-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
+                                              threshold = 0.016))
+
+unifcube_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.8, 0.8), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
 
@@
 The cube is generated as per the uniformly distributed cube, but points inside sphere of radius (`r_hole`) are removed, resulting in a hollow cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) c"} else { "Figure \\@ref(fig:cube-proj) c"})`).
+
+```{r data-cubehole, echo=TRUE}
+cubehole <- gen_cubehole(n = 3000, p = 4, r_hole = 0.5)
+```
+
+```{r}
+cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r cubehole-proj1}
+scaled_data <- scale_data_manual(cubehole)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.022))
+
+cubehole_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.9, 0.75), 
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cubehole-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.73, 
+                                              axis_pos_y = -0.73, 
+                                              threshold = 0.04))
+
+cubehole_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.85, 0.7), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cubehole-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.016))
+
+cubehole_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.9, 0.8), 
+  title = "c3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
 
@@
   column_spec(2, width = "8cm")
-```
-
-```{r data-gridcube, echo=TRUE}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-```
-
-```{r}
-gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r gridcube-proj1}
-scaled_data <- scale_data_manual(gridcube)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.7, 
-                                              axis_pos_y = -0.7, 
-                                              threshold = 0.022))
-
-gridcube_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.83, 0.8), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridcube-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.8,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.85, 
-                                              axis_pos_y = -0.85, 
-                                              threshold = 0.05))
-
-gridcube_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 1), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridcube-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.7, 
-                                              axis_pos_y = -0.7, 
-                                              threshold = 0.016))
-
-gridcube_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r data-unifcube, echo=TRUE}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-```
-
-```{r}
-unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r unifcube-proj1}
-scaled_data <- scale_data_manual(unifcube)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.022))
-
-unifcube_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.8), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifcube-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.04))
-
-unifcube_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.8), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifcube-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.016))
-
-unifcube_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r data-cubehole, echo=TRUE}
-cubehole <- gen_cubehole(n = 3000, p = 4, r_hole = 0.5)
-```
-
-```{r}
-cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r cubehole-proj1}
-scaled_data <- scale_data_manual(cubehole)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.022))
-
-cubehole_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.75), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cubehole-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.73, 
-                                              axis_pos_y = -0.73, 
-                                              threshold = 0.04))
-
-cubehole_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.85, 0.7), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cubehole-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.016))
-
-cubehole_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.8), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
 ```
@@
 
-### Polynomial
+#### Polynomial
 
@@
 
-#### `gen_quadratic()`
-
-The `gen_quadratic(n, p, range)` function generates a dataset of $n$ points forming a quadratic curve in the first two dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) a"} else { "Figure \\@ref(fig:poly-proj) a"})`).
-This $2\text{-}D$ parabolic structure is embedded within a higher-dimensional space with additive noise in the remaining dimensions.
-
-The curve is constructed by randomly sampling inputs and applying a second-degree polynomial transformation.
-Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the independent variable.
-A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise.
-For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ structure into $p\text{-}D$, where $j = 3, \dots, p$.
+The first is the quadratic curve of $n$ points in $p$ dimensions.
+This is generated using `gen_quadratic(n, p, range)`.
+The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree 2 is applied to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
+This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
+For $p > 2$, the quadratic structure is embedded into higher dimensions by appending additional noise coordinates  (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) a"} else { "Figure \\@ref(fig:poly-proj) a"})`).
 
@@
 
-#### `gen_cubic()`
-
-The `gen_cubic(n, p, range)` function generates a dataset of $n$ points forming a cubic curve in the first two dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) b"} else { "Figure \\@ref(fig:poly-proj) b"})`).
-This function creates a more complex curvilinear structure than a simple parabola.
-
-The shape is generated using a third-degree raw polynomial basis expansion.
-Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the base input.
-A cubic transformation with vertical jitter is used to define $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ structure into $p\text{-}D$, where $j = 3, \dots, p$.
+The second is the cubic curve of $n$ points in $p$ dimensions.
+This is generated using `gen_cubic(n, p, range)`.
+The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
+This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
+For $p > 2$, the cubic curve is embedded into higher dimensions by appending additional noise coordinates (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) b"} else { "Figure \\@ref(fig:poly-proj) b"})`).
 
@@
 
-### S-curve
+#### S-curve
 

### 2025-09-19 paper-cardinalR.tex
Add cube functions
Updated polynomial section

@@
 
+\begin{verbatim}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+\end{verbatim}
+
 An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
 The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube (Figure \ref{fig:cube-proj} b).
+
+\begin{verbatim}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+\end{verbatim}
 
@@
 
+\begin{verbatim}
+cubehole <- gen_cubehole(n = 3000, p = 4, r_hole = 0.5)
+\end{verbatim}
+
 \begin{table}
@@
 \end{table}
-
-\begin{verbatim}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{verbatim}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{verbatim}
-cubehole <- gen_cubehole(n = 3000, p = 4, r_hole = 0.5)
-\end{verbatim}
 
@@
 
-\subsubsection{Polynomial}\label{polynomial}
+\paragraph{Polynomial}\label{polynomial}
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_quadratic()}}{gen\_quadratic()}}\label{gen_quadratic}
-
-The \texttt{gen\_quadratic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a quadratic curve in the first two dimensions (Figure \ref{fig:poly-proj} a).
-This \(2\text{-}D\) parabolic structure is embedded within a higher-dimensional space with additive noise in the remaining dimensions.
-
-The curve is constructed by randomly sampling inputs and applying a second-degree polynomial transformation.
-Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the independent variable.
-A raw polynomial basis of degree 2 is used to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise.
-For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) structure into \(p\text{-}D\), where \(j = 3, \dots, p\).
+The first is the quadratic curve of \(n\) points in \(p\) dimensions.
+This is generated using \texttt{gen\_quadratic(n,\ p,\ range)}.
+The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree 2 is applied to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
+This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
+For \(p > 2\), the quadratic structure is embedded into higher dimensions by appending additional noise coordinates (Figure \ref{fig:poly-proj} a).
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_cubic()}}{gen\_cubic()}}\label{gen_cubic}
-
-The \texttt{gen\_cubic(n,\ p,\ range)} function generates a dataset of \(n\) points forming a cubic curve in the first two dimensions (Figure \ref{fig:poly-proj} b).
-This function creates a more complex curvilinear structure than a simple parabola.
-
-The shape is generated using a third-degree raw polynomial basis expansion.
-Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the base input.
-A cubic transformation with vertical jitter is used to define \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) structure into \(p\text{-}D\), where \(j = 3, \dots, p\).
+The second is the cubic curve of \(n\) points in \(p\) dimensions.
+This is generated using \texttt{gen\_cubic(n,\ p,\ range)}.
+The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree \(3\) is applied to construct \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
+This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
+For \(p > 2\), the cubic curve is embedded into higher dimensions by appending additional noise coordinates (Figure \ref{fig:poly-proj} b).
 
@@
 
-\subsubsection{S-curve}\label{s-curve}
+\paragraph{S-curve}\label{s-curve}
 

### 2025-09-19 paper-cardinalR.Rmd
Updated Trefoil section

@@
 To introduce missing or incomplete regions on the manifold, `gen_scurvehole(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) b"} else { "Figure \\@ref(fig:scurve-proj) b"})`) removes points in a localized region centered around the middle vertical section of the S-curve.
-Following a similar approach as `gen_cubehole(n, p, r_hole)`, all observations within a fixed radius ($\sqrt{0.3}$) of the anchor point are excluded, creating a hole in the manifold while preserving the overall S-curve structure.
+Following a similar approach as `gen_cubehole()`, all observations within a fixed radius ($\sqrt{0.3}$) of the anchor point are excluded, creating a hole in the manifold while preserving the overall S-curve structure.
 
@@
 The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil)"} else { "Figure \\@ref(fig:trefoil-proj)"})`).
-The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of NLDR methods to preserve global structure, loops, and embeddings in high-dimensional data.
+The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trefoil-tb-html)', '\\@ref(tab:trefoil-tb-pdf)'))` summarizes these functions.
@@
 
-#### `gen_trefoil4d()`
-
-The function `gen_trefoil4d()` generates a $4\text{-}D$ trefoil knot using two angular parameters, $\theta$ and $\phi$.
-The knot is constructed on the $3$-sphere ($S^3 \subset \mathbb{R}^4$), with a band of thickness around the knot path, controlled by `steps` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
-
-The coordinates of the $4\text{-}D$ trefoil knot are defined as $X_1 = \cos(\theta) \cos(\phi)$, $X_2 = \cos(\theta) \sin(\phi)$, $X_3 = \sin(\theta) \cos(1.5 \phi)$, and $X_4 = \sin(\theta) \sin(1.5 \phi)$, where $\theta$ and $\phi$ are angular parameters that trace out the knot's structure on the $3$-sphere $S^3 \subset \mathbb{R}^4$.
-
-The number of `theta` and `phi` values is determined by the `steps` and `n` arguments, respectively.
-Additional independent noise dimensions can be added via `gen_noisedims()` to extend the structure into $p > 4$ dimensions.
+For the $4\text{-}D$ trefoil knot, the function `gen_trefoil4d(n, p, steps)` generates the structure on the $3$-sphere ($S^3 \subset \mathbb{R}^4$) using two angular parameters, $\theta$ and $\phi$.
+A band of thickness around the knot path is controlled by the `steps` argument, while the number of $\theta$ and $\phi$ values is determined by the `steps` and `n` arguments, respectively (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
+The coordinates are defined as $X_1 = \cos(\theta) \cos(\phi)$, $X_2 = \cos(\theta) \sin(\phi)$, $X_3 = \sin(\theta) \cos(1.5 \phi)$, and $X_4 = \sin(\theta) \sin(1.5 \phi)$, where $\theta$ and $\phi$ trace the knot’s path.
+For $p > 4$, the quadratic structure is embedded into higher dimensions by appending additional noise dimensions.
 
@@
 
-#### `gen_trefoil3d()`
-
-The function `gen_trefoil3d()` generates a $3\text{-}D$ stereographic projection of the $4\text{-}D$ trefoil knot by mapping each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3$ using the transformation $X_1' = X_1 / (1 - X_4)$, $X_2' = X_2 / (1 - X_4)$, and $X_3' = X_3 / (1 - X_4)$, while excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
+For the $3\text{-}D$ stereographic projection, `gen_trefoil3d(n, p, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3$ using $X_1' = X_1 / (1 - X_4)$, $X_2' = X_2 / (1 - X_4)$, and $X_3' = X_3 / (1 - X_4)$, excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
+As with the $4\text{-}D$ case, optional noise dimensions can be added to embed the knot into higher-dimensional spaces.
 
@@
 ```
-
-As with the $4\text{-}D$ version, this function supports optional noise dimensions to embed the projected knot into higher dimensions.
 

### 2025-09-19 paper-cardinalR.tex
Updated Trefoil section

@@
 To introduce missing or incomplete regions on the manifold, \texttt{gen\_scurvehole(n,\ p)} (Figure \ref{fig:scurve-proj} b) removes points in a localized region centered around the middle vertical section of the S-curve.
-Following a similar approach as \texttt{gen\_cubehole(n,\ p,\ r\_hole)}, all observations within a fixed radius (\(\sqrt{0.3}\)) of the anchor point are excluded, creating a hole in the manifold while preserving the overall S-curve structure.
+Following a similar approach as \texttt{gen\_cubehole()}, all observations within a fixed radius (\(\sqrt{0.3}\)) of the anchor point are excluded, creating a hole in the manifold while preserving the overall S-curve structure.
 
@@
 The Trefoil is a closed, nontrivial one-dimensional manifold embedded in \(3\text{-}D\) or \(4\text{-}D\) space (Figure \ref{fig:trefoil-proj}).
-The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of NLDR methods to preserve global structure, loops, and embeddings in high-dimensional data.
+The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
 Table \ref{tab:trefoil-tb-pdf} summarizes these functions.
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_trefoil4d()}}{gen\_trefoil4d()}}\label{gen_trefoil4d}
-
-The function \texttt{gen\_trefoil4d()} generates a \(4\text{-}D\) trefoil knot using two angular parameters, \(\theta\) and \(\phi\).
-The knot is constructed on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)), with a band of thickness around the knot path, controlled by \texttt{steps} (Figure \ref{fig:trefoil-proj} a).
-
-The coordinates of the \(4\text{-}D\) trefoil knot are defined as \(X_1 = \cos(\theta) \cos(\phi)\), \(X_2 = \cos(\theta) \sin(\phi)\), \(X_3 = \sin(\theta) \cos(1.5 \phi)\), and \(X_4 = \sin(\theta) \sin(1.5 \phi)\), where \(\theta\) and \(\phi\) are angular parameters that trace out the knot's structure on the \(3\)-sphere \(S^3 \subset \mathbb{R}^4\).
-
-The number of \texttt{theta} and \texttt{phi} values is determined by the \texttt{steps} and \texttt{n} arguments, respectively.
-Additional independent noise dimensions can be added via \texttt{gen\_noisedims()} to extend the structure into \(p > 4\) dimensions.
+For the \(4\text{-}D\) trefoil knot, the function \texttt{gen\_trefoil4d(n,\ p,\ steps)} generates the structure on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)) using two angular parameters, \(\theta\) and \(\phi\).
+A band of thickness around the knot path is controlled by the \texttt{steps} argument, while the number of \(\theta\) and \(\phi\) values is determined by the \texttt{steps} and \texttt{n} arguments, respectively (Figure \ref{fig:trefoil-proj} a).
+The coordinates are defined as \(X_1 = \cos(\theta) \cos(\phi)\), \(X_2 = \cos(\theta) \sin(\phi)\), \(X_3 = \sin(\theta) \cos(1.5 \phi)\), and \(X_4 = \sin(\theta) \sin(1.5 \phi)\), where \(\theta\) and \(\phi\) trace the knot's path.
+For \(p > 4\), the quadratic structure is embedded into higher dimensions by appending additional noise dimensions.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_trefoil3d()}}{gen\_trefoil3d()}}\label{gen_trefoil3d}
-
-The function \texttt{gen\_trefoil3d()} generates a \(3\text{-}D\) stereographic projection of the \(4\text{-}D\) trefoil knot by mapping each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\) using the transformation \(X_1' = X_1 / (1 - X_4)\), \(X_2' = X_2 / (1 - X_4)\), and \(X_3' = X_3 / (1 - X_4)\), while excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
+For the \(3\text{-}D\) stereographic projection, \texttt{gen\_trefoil3d(n,\ p,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\) using \(X_1' = X_1 / (1 - X_4)\), \(X_2' = X_2 / (1 - X_4)\), and \(X_3' = X_3 / (1 - X_4)\), excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
+As with the \(4\text{-}D\) case, optional noise dimensions can be added to embed the knot into higher-dimensional spaces.
 
@@
 \end{verbatim}
-
-As with the \(4\text{-}D\) version, this function supports optional noise dimensions to embed the projected knot into higher dimensions.
 

### 2025-09-19 paper-cardinalR.Rmd
Add summary sentence for noise

@@
 The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree 2 is applied to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
-For $p > 2$, the quadratic structure is embedded into higher dimensions by appending additional noise coordinates  (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) a"} else { "Figure \\@ref(fig:poly-proj) a"})`).
+This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) a"} else { "Figure \\@ref(fig:poly-proj) a"})`).
 
@@
 The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
-For $p > 2$, the cubic curve is embedded into higher dimensions by appending additional noise coordinates (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) b"} else { "Figure \\@ref(fig:poly-proj) b"})`).
+This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) b"} else { "Figure \\@ref(fig:poly-proj) b"})`).
 
@@
 ```
+
+For all polynomial shapes, if $p > 2$, the structure is embedded into higher dimensions by appending additional noise dimensions.
 
@@
 The corresponding $2\text{-}D$ coordinates are defined by: $X_1 = \cos(\theta)$, $X_2 = \sin(\theta)$.
-For $p > 2$, additional dimensions $X_3$ to $X_p$ are generated using `gen_wavydims1()`: $X_j = s_j\theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}$.
-These noise dimensions are nonlinearly correlated with $\theta$, producing structured, wave-like variation that aligns with the progression along the crescent curve.
 
@@
 This forms a curvy cylindrical surface in $4\text{-}D$, where the fourth dimension bends periodically along the height axis, resembling a helicoid or twisting wave along the cylinder.
-For $p > 4$, dimensions $X_5$ through $X_p$ are generated as: $X_j \sim N(0, 0.05^2),\text{ for }j = 5, \ldots, p$.
 
@@
 
-If $p > 4$, the function appends structured, non-linear noise via `gen_wavydims2()`.
-These dimensions are functions of the first coordinate $X_1$, introducing dependencies that preserve some geometric coherence.
-Each added dimension follows the form: $X_j = s_j(-1)^{\lfloor j/2 \rfloor}X_1^{a_j} + \eta_j$, where $a_j \in \{2, 3, 4, 5\}$ is a randomly chosen polynomial power, $s_j \sim U(0.5, 2)$ is a scale factor, $\eta_j \sim U(-\sigma_j, 2\sigma_j)$, with $\sigma_j \sim U(0, 0.05)$, adds mild randomness.
-
 ```{r data-sphericalspiral, echo=TRUE}
@@
 This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
-When $p > 4$, the remaining $p - 4$ dimensions are populated using `gen_noisedims()`.
-These are independent Gaussian noise dimensions: $X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p$.
 
@@
 This results in a $3\text{-}D$ spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
-For $p > 4$, the dataset includes isotropic Gaussian noise in the remaining dimensions via `gen_noisedims()`: $X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p$.
 
@@
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-If $p > 4$, the remaining dimensions are populated with Gaussian noise using `gen_noisedims()`: $X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p$.
 
@@
 
+For trigonometric shapes, additional noise dimensions are appended to embed the structure into higher-dimensional space: for the crescent-shaped curve this applies when $p > 2$ (adding $X_3$ to $X_p$), while for all other trigonometric shapes it applies when $p > 4$ (adding $X_5$ to $X_p$).
+
 ### Trefoil knots
@@
 ### Generate noise dimensions
+
+generate noise dimensions with a normal distribution and various wavy patterns,
 
@@
 The package includes various supplementary tools in addition to the shape generating functions mentioned earlier.
-These tools allow users to generate noise dimensions with a normal distribution and various wavy patterns, create background noise, randomize the rows of the data, reposition clusters, generate a vector whose product and sum are approximately equal to a target value, rotate structures, and normalize the data.
+These tools allow users to create background noise, randomize the rows of the data, relocate clusters, generate a vector whose product and sum are approximately equal to a target value, rotate structures, and normalize the data.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:add-tb-html)', '\\@ref(tab:add-tb-pdf)'))` details these functions.

### 2025-09-19 paper-cardinalR.tex
Add summary sentence for noise

@@
 The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree 2 is applied to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
-For \(p > 2\), the quadratic structure is embedded into higher dimensions by appending additional noise coordinates (Figure \ref{fig:poly-proj} a).
+This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term (Figure \ref{fig:poly-proj} a).
 
@@
 The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree \(3\) is applied to construct \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
-For \(p > 2\), the cubic curve is embedded into higher dimensions by appending additional noise coordinates (Figure \ref{fig:poly-proj} b).
+This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points (Figure \ref{fig:poly-proj} b).
 
@@
 \end{figure}
+
+For all polynomial shapes, if \(p > 2\), the structure is embedded into higher dimensions by appending additional noise dimensions.
 
@@
 The corresponding \(2\text{-}D\) coordinates are defined by: \(X_1 = \cos(\theta)\), \(X_2 = \sin(\theta)\).
-For \(p > 2\), additional dimensions \(X_3\) to \(X_p\) are generated using \texttt{gen\_wavydims1()}: \(X_j = s_j\theta + \varepsilon_j,\text{ where }\varepsilon_j \sim N(0, 0.5^2),\text{ and }s_j\text{ is a random scale}\).
-These noise dimensions are nonlinearly correlated with \(\theta\), producing structured, wave-like variation that aligns with the progression along the crescent curve.
 
@@
 This forms a curvy cylindrical surface in \(4\text{-}D\), where the fourth dimension bends periodically along the height axis, resembling a helicoid or twisting wave along the cylinder.
-For \(p > 4\), dimensions \(X_5\) through \(X_p\) are generated as: \(X_j \sim N(0, 0.05^2),\text{ for }j = 5, \ldots, p\).
 
@@
 This generates a spherical spiral curve embedded in \(4\text{-}D\) space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
-
-If \(p > 4\), the function appends structured, non-linear noise via \texttt{gen\_wavydims2()}.
-These dimensions are functions of the first coordinate \(X_1\), introducing dependencies that preserve some geometric coherence.
-Each added dimension follows the form: \(X_j = s_j(-1)^{\lfloor j/2 \rfloor}X_1^{a_j} + \eta_j\), where \(a_j \in \{2, 3, 4, 5\}\) is a randomly chosen polynomial power, \(s_j \sim U(0.5, 2)\) is a scale factor, \(\eta_j \sim U(-\sigma_j, 2\sigma_j)\), with \(\sigma_j \sim U(0, 0.05)\), adds mild randomness.
 
@@
 This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
-When \(p > 4\), the remaining \(p - 4\) dimensions are populated using \texttt{gen\_noisedims()}.
-These are independent Gaussian noise dimensions: \(X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p\).
 
@@
 This results in a \(3\text{-}D\) spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
-For \(p > 4\), the dataset includes isotropic Gaussian noise in the remaining dimensions via \texttt{gen\_noisedims()}: \(X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p\).
 
@@
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-If \(p > 4\), the remaining dimensions are populated with Gaussian noise using \texttt{gen\_noisedims()}: \(X_j \sim N(0, 0.05^2), \quad j = 5, \dots, p\).
 
@@
 \end{figure}
+
+For trigonometric shapes, additional noise dimensions are appended to embed the structure into higher-dimensional space: for the crescent-shaped curve this applies when \(p > 2\) (adding \(X_3\) to \(X_p\)), while for all other trigonometric shapes it applies when \(p > 4\) (adding \(X_5\) to \(X_p\)).
 
@@
 \subsubsection{Generate noise dimensions}\label{generate-noise-dimensions}
+
+generate noise dimensions with a normal distribution and various wavy patterns,
 
@@
 The package includes various supplementary tools in addition to the shape generating functions mentioned earlier.
-These tools allow users to generate noise dimensions with a normal distribution and various wavy patterns, create background noise, randomize the rows of the data, reposition clusters, generate a vector whose product and sum are approximately equal to a target value, rotate structures, and normalize the data.
+These tools allow users to create background noise, randomize the rows of the data, relocate clusters, generate a vector whose product and sum are approximately equal to a target value, rotate structures, and normalize the data.
 Table \ref{tab:add-tb-pdf} details these functions.

### 2025-09-19 paper-cardinalR.Rmd
Updated trigometric section

@@
 Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric)"} else { "Figure \\@ref(fig:triginometric-proj)"})`).
-The package includes functions to generate a wide range of these forms from crescents and curvy cylinders to spherical, helical, and conic spirals, as well as nonlinear hyperbolas.
+The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when $p > 2$ (adding $X_3$ through $X_p$), and for all other trigonometric shapes when $p > 4$ (adding $X_5$ through $X_p$).
 These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
@@
 
-#### `gen_crescent()`
-
-The `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) a"} else { "Figure \\@ref(fig:triginometric-proj) a"})`).
-
-The first two dimensions define a crescent-shaped curve using a semi-circular arc.
+First, the `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) a"} else { "Figure \\@ref(fig:triginometric-proj) a"})`).
 Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
@@
 
-#### `gen_curvycylinder()`
-
-The `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
-
+Second, the `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
 The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
@@
 The coordinates are defined as: $X_1 = \cos(\theta)$ (Circular base, x-axis), $X_2 = \sin(\theta)$ (Circular base, y-axis), $X_3 = z$ (Linear height), and $X_4 = \sin(z)$ (Nonlinear curvy variation along height).
-This forms a curvy cylindrical surface in $4\text{-}D$, where the fourth dimension bends periodically along the height axis, resembling a helicoid or twisting wave along the cylinder.
 
@@
 
-#### `gen_sphericalspiral()`
-
-The `gen_sphericalspiral(n, p, spins)` function simulates a dataset of $n$ observations that form a spiraling path on a spherical surface in the first four dimensions.
-When extended beyond $4\text{-}D$, structured nonlinear noise dimensions are added to simulate more realistic high-dimensional manifolds (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`).
-
-The first three dimensions represent points on a unit sphere.
-Let $\theta \in [0, 2\pi \times \text{spins}]$ be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
-
+For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$ be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
 Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
@@
 
-#### `gen_helicalspiral()`
-
-The `gen_helicalspiral(n, p)` function generates a dataset of $n$ observations forming a helical spiral, embedded in the first four dimensions of $p\text{-}D$.
-Additional dimensions (if $p > 4$) are filled with structured noise to mimic high-dimensional complexity (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) d"} else { "Figure \\@ref(fig:triginometric-proj) d"})`).
-
-The first four coordinates follow a $3\text{-}D$ helix with an additional oscillating component.
-Let $\theta \in [0, 5\pi/4]$ be a sequence of angles controlling rotation around a circle.
+For a helical spiral in four dimensions, `gen_helicalspiral(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) d"} else { "Figure \\@ref(fig:triginometric-proj) d"})`), let $\theta \in [0, 5\pi/4]$ be a sequence of angles controlling rotation around a circle.
 Cartesian coordinates; $X_1 = \cos(\theta)$: circular trajectory along the x-axis, $X_2 = \sin(\theta)$: circular trajectory along the y-axis, $X_3 = 0.05\theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix, and $X_4 = 0.1\sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
-This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
 
@@
 
-#### `gen_conicspiral()`
-
-The `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) e"} else { "Figure \\@ref(fig:triginometric-proj) e"})`).
+Similarly, the `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) e"} else { "Figure \\@ref(fig:triginometric-proj) e"})`).
 The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a $3\text{-}D$ conic helix.
-
 The shape is defined by parameter $\theta \in [0, 2\pi\text{spins}]$, controlling the angular progression of the spiral.
@@
 The growth pattern resembles a cone, with the height increasing according to $X_3 = 2\theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta\sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
-This results in a $3\text{-}D$ spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
 
@@
 
-#### `gen_nonlinear()`
-
-The `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
-
+Finally, the `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
 The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
@@
 ```
-
-For trigonometric shapes, additional noise dimensions are appended to embed the structure into higher-dimensional space: for the crescent-shaped curve this applies when $p > 2$ (adding $X_3$ to $X_p$), while for all other trigonometric shapes it applies when $p > 4$ (adding $X_5$ to $X_p$).
 

### 2025-09-19 paper-cardinalR.tex
Updated trigometric section

@@
 Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (Figure \ref{fig:triginometric-proj}).
-The package includes functions to generate a wide range of these forms from crescents and curvy cylinders to spherical, helical, and conic spirals, as well as nonlinear hyperbolas.
+The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when \(p > 2\) (adding \(X_3\) through \(X_p\)), and for all other trigonometric shapes when \(p > 4\) (adding \(X_5\) through \(X_p\)).
 These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_crescent()}}{gen\_crescent()}}\label{gen_crescent}
-
-The \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise (Figure \ref{fig:triginometric-proj} a).
-
-The first two dimensions define a crescent-shaped curve using a semi-circular arc.
+First, the \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise (Figure \ref{fig:triginometric-proj} a).
 Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_curvycylinder()}}{gen\_curvycylinder()}}\label{gen_curvycylinder}
-
-The \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
-
+Second, the \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
 The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
@@
 The coordinates are defined as: \(X_1 = \cos(\theta)\) (Circular base, x-axis), \(X_2 = \sin(\theta)\) (Circular base, y-axis), \(X_3 = z\) (Linear height), and \(X_4 = \sin(z)\) (Nonlinear curvy variation along height).
-This forms a curvy cylindrical surface in \(4\text{-}D\), where the fourth dimension bends periodically along the height axis, resembling a helicoid or twisting wave along the cylinder.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_sphericalspiral()}}{gen\_sphericalspiral()}}\label{gen_sphericalspiral}
-
-The \texttt{gen\_sphericalspiral(n,\ p,\ spins)} function simulates a dataset of \(n\) observations that form a spiraling path on a spherical surface in the first four dimensions.
-When extended beyond \(4\text{-}D\), structured nonlinear noise dimensions are added to simulate more realistic high-dimensional manifolds (Figure \ref{fig:triginometric-proj} c).
-
-The first three dimensions represent points on a unit sphere.
-Let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\)be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
-
+For a spiraling path on a spherical surface in the first four dimensions, \texttt{gen\_sphericalspiral(n,\ p,\ spins)} (Figure \ref{fig:triginometric-proj} c), let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\)be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
 Cartesian coordinates from spherical conversion: \(X_1 = \sin(\phi)\cos(\theta)\), \(X_2 = \sin(\phi)\sin(\theta)\), \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter, and \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_helicalspiral()}}{gen\_helicalspiral()}}\label{gen_helicalspiral}
-
-The \texttt{gen\_helicalspiral(n,\ p)} function generates a dataset of \(n\) observations forming a helical spiral, embedded in the first four dimensions of \(p\text{-}D\).
-Additional dimensions (if \(p > 4\)) are filled with structured noise to mimic high-dimensional complexity (Figure \ref{fig:triginometric-proj} d).
-
-The first four coordinates follow a \(3\text{-}D\) helix with an additional oscillating component.
-Let \(\theta \in [0, 5\pi/4]\) be a sequence of angles controlling rotation around a circle.
+For a helical spiral in four dimensions, \texttt{gen\_helicalspiral(n,\ p)} (Figure \ref{fig:triginometric-proj} d), let \(\theta \in [0, 5\pi/4]\) be a sequence of angles controlling rotation around a circle.
 Cartesian coordinates; \(X_1 = \cos(\theta)\): circular trajectory along the x-axis, \(X_2 = \sin(\theta)\): circular trajectory along the y-axis, \(X_3 = 0.05\theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix, and \(X_4 = 0.1\sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
-This results in a helical spiral winding upward along the third axis, with gentle sinusoidal fluctuation in the fourth dimension.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_conicspiral()}}{gen\_conicspiral()}}\label{gen_conicspiral}
-
-The \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of \(p\text{-}D\) (Figure \ref{fig:triginometric-proj} e).
+Similarly, the \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of \(p\text{-}D\) (Figure \ref{fig:triginometric-proj} e).
 The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a \(3\text{-}D\) conic helix.
-
 The shape is defined by parameter \(\theta \in [0, 2\pi\text{spins}]\), controlling the angular progression of the spiral.
@@
 The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2\theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\) Spiral modulation in the fourth dimension is represented by \(X_4 = \theta\sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
-This results in a \(3\text{-}D\) spiral surface expanding upward and outward like a cone, with an oscillatory fourth dimension capturing spiral irregularities.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_nonlinear()}}{gen\_nonlinear()}}\label{gen_nonlinear}
-
-The \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
-
+Finally, the \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
 The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
@@
 \end{figure}
-
-For trigonometric shapes, additional noise dimensions are appended to embed the structure into higher-dimensional space: for the crescent-shaped curve this applies when \(p > 2\) (adding \(X_3\) to \(X_p\)), while for all other trigonometric shapes it applies when \(p > 4\) (adding \(X_5\) to \(X_p\)).
 

### 2025-09-19 paper-cardinalR.Rmd
Updated

@@
 The package provides functions to generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
-These shapes help test how well methods preserve symmetry, curvature, and cluster separation.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:sphere-tb-html)', '\\@ref(tab:sphere-tb-pdf)'))` summarizes these functions.
@@
 A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
-The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle." In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin.
+The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle".
+In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin.
 Additionally, $X_3 = 1/3\cos(3\theta)$ introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
@@
 The corresponding Cartesian coordinates are calculated as; $X_1 = r\sqrt{1 - u^2}\cos(\theta)$, $X_2 = r\sqrt{1 - u^2}\sin(\theta)$, $X_3 = ru$, which gives points uniformly distributed on the surface of a $3\text{-}D$ sphere (not within).
-For $p > 3$, additional dimensions $X_4$ to $X_p$ are generated as low-variance Gaussian noise: $X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p$.
+For $p > 3$, additional noise dimensions $X_4$ to $X_p$ are generated.
 
@@
 This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
-For $p > 4$, additional dimensions $X_5$ to $X_p$ are added as low-variance Gaussian noise: $X_j \sim N(0, 0.05^2),\text{ for }j = 5, \dots, p$.
+For $p > 4$, additional noise dimensions are added to embed the structure in higher dimensions.
 
@@
 
-### Trigonometric
+#### Trigonometric
 
@@
 
-### Trefoil knots
+#### Trefoil knots
 

### 2025-09-19 paper-cardinalR.tex
Updated

@@
 The package provides functions to generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
-These shapes help test how well methods preserve symmetry, curvature, and cluster separation.
 Table \ref{tab:sphere-tb-pdf} summarizes these functions.
@@
 A latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
-The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle.'' In this configuration, \(X_1 = \cos(\theta)\) represents horizontal oscillation, while \(X_2 = \sqrt{3}/3 + \sin(\theta)\) introduces a vertical offset to avoid centering the curve at the origin.
+The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle''.
+In this configuration, \(X_1 = \cos(\theta)\) represents horizontal oscillation, while \(X_2 = \sqrt{3}/3 + \sin(\theta)\) introduces a vertical offset to avoid centering the curve at the origin.
 Additionally, \(X_3 = 1/3\cos(3\theta)\) introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
@@
 The corresponding Cartesian coordinates are calculated as; \(X_1 = r\sqrt{1 - u^2}\cos(\theta)\), \(X_2 = r\sqrt{1 - u^2}\sin(\theta)\), \(X_3 = ru\), which gives points uniformly distributed on the surface of a \(3\text{-}D\) sphere (not within).
-For \(p > 3\), additional dimensions \(X_4\) to \(X_p\) are generated as low-variance Gaussian noise: \(X_j \sim N(0, 0.05^2),\text{ for }j = 4, \dots, p\).
+For \(p > 3\), additional noise dimensions \(X_4\) to \(X_p\) are generated.
 
@@
 This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
-For \(p > 4\), additional dimensions \(X_5\) to \(X_p\) are added as low-variance Gaussian noise: \(X_j \sim N(0, 0.05^2),\text{ for }j = 5, \dots, p\).
+For \(p > 4\), additional noise dimensions are added to embed the structure in higher dimensions.
 

### 2025-09-20 paper-cardinalR.Rmd
Updated sphere section

@@
 
-### Sphere
+#### Sphere
 
 Sphere-shaped structures (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere)"} else { "Figure \\@ref(fig:sphere-proj)"})`) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
-The package provides functions to generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+The functions generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
+For the circle and curvy cycle, this occurs when $p > 2$ and $p > 3$, respectively, while for spheres and hemispheres it occurs when $p > 3$ or $p > 4$.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:sphere-tb-html)', '\\@ref(tab:sphere-tb-pdf)'))` summarizes these functions.
@@
 
-#### `gen_circle()`
-
-The function `gen_circle(n, p)` generates a $p$-dimensional dataset of $n$ observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
-
-A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
+The simplest case, `gen_circle(n, p)` creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
+Let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
 Coordinates in the first two dimensions represent a perfect circle on the plane: $X_1 = \cos(\theta), \quad X_2 = \sin(\theta)$.
@@
 
-#### `gen_curvycycle()`
-
-The `gen_curvycycle(n, p)` function generates a $p$-dimensional dataset of $n$ observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`).
-The structure forms a one-dimensional nonlinear cycle embedded in higher-dimensional space.
-
-A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
+For the one-dimensional nonlinear cycle embedded in $p\text{-}D$ space, `gen_curvycycle(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`), let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
 The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle".
@@
 
-#### `gen_unifsphere()`
-
-The `gen_unifsphere(n, p, r)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on the surface of a $3\text{-}D$ sphere of radius $r$, with additional Gaussian noise dimensions added when $p > 3$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) c"} else { "Figure \\@ref(fig:sphere-proj) c"})`).
-
-Each observation lies on the surface of a sphere in $3\text{-}D$, constructed by generating $u \sim U(-1, 1)$ which represents the cosine of the polar angle $\phi$ and $\theta \sim U(0, 2\pi)$ which represents the azimuthal angle.
-The corresponding Cartesian coordinates are calculated as; $X_1 = r\sqrt{1 - u^2}\cos(\theta)$, $X_2 = r\sqrt{1 - u^2}\sin(\theta)$, $X_3 = ru$, which gives points uniformly distributed on the surface of a $3\text{-}D$ sphere (not within).
-For $p > 3$, additional noise dimensions $X_4$ to $X_p$ are generated.
+Building on simple circular structures, the `gen_unifsphere(n, p, r)` function function extends the idea to three dimensions by generating $n$ observations approximately uniformly distributed on the surface of a sphere of radius $r$, with optional high-dimensional noise when $p > 3$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) c"} else { "Figure \\@ref(fig:sphere-proj) c"})`).
+Each observation is computed from spherical coordinates, with $u \sim U(-1, 1)$ representing $\cos(\phi)$ and $\theta \sim U(0, 2\pi)$ the azimuthal angle.
+Cartesian coordinates are then defined as $X_1 = r\sqrt{1 - u^2}\cos(\theta)$, $X_2 = r\sqrt{1 - u^2}\sin(\theta)$, and $X_3 = ru$, ensuring uniform distribution on the surface (not within) of the sphere.
+For $p > 3$, additional dimensions $X_4$ through $X_p$ are appended as structured noise.
 
@@
 
-#### `gen_gridedsphere()`
-
-The `gen_gridedsphere(n, p)` function generates a $p$-dimensional dataset of approximately $n$ points evenly distributed on the surface of a $(p!-!1)$-sphere (the unit sphere embedded in $\mathbb{R}^p$) (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
-
-The construction is based on a regular grid in spherical coordinates with $(p-1)$ angular variables.
-Specifically, for dimensions $j = 1, \dots, p-2$, the polar angles are defined on $[0, \pi]$.
-For the last angle ($j = p-1$), the azimuthal angle is defined on $[0, 2\pi]$.
-
-The number of grid steps along each angular dimension is determined by factoring $n$ into $(p-1)$ approximately equal integers via `gen_nproduct(n, p - 1)`.
-
-Each grid point is then mapped to Cartesian coordinates using the standard hyperspherical-to-Cartesian transformation:
+In addition, the `gen_gridedsphere(n, p)` function constructs a $p\text{-}D$ dataset consisting of approximately $n$ points that are evenly distributed on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
+The method relies on forming a regular grid in spherical coordinates, parameterized by $(p-1)$ angular variables: for dimensions $j = 1, \dots, p-2$ the polar angles are drawn from $[0, \pi]$, while the final angle ($j = p-1$) represents the azimuth and is drawn from $[0, 2\pi]$.
+The number of grid steps along each angular dimension is chosen by decomposing $n$ into $(p-1)$ approximately equal integer factors using the helper function `gen_nproduct(n, p - 1)`.
+
+Each grid point is subsequently mapped into Cartesian space via the standard hyperspherical-to-Cartesian transformation,
 
@@
 
-Thus, the output is a deterministic grid of points lying exactly on the unit $(p-1)$-sphere, with no added noise dimensions.
+The result is a deterministic grid of points lying exactly on the surface of the unit $(p-1)$-sphere, without any additional noise dimensions.
 
@@
 
-#### `gen_clusteredspheres()`
-
-The `gen_clusteredspheres(n, k, p, r, loc)` function generates a synthetic dataset of $n_1 + kn_2$ observations in $p$-dimensional space, consisting of one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
-
+For more heterogeneous structures, the `gen_clusteredspheres(n, k, p, r, loc)` function generates one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
 A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
 The sampling is executed using the function `gen_unifsphere(n_1, p, r_1)`, which generates the desired points in the specified dimensional space.
-
 In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
 These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
-
 Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for $3\text{-}D$).
@@
 
-#### `gen_hemisphere()`
-
-The `gen_hemisphere(n, p)` function generates a $p$-dimensional dataset of $n$ observations distributed approximately uniformly on a $4\text{-}D$ hemisphere, optionally extended with Gaussian noise in additional dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) f"} else { "Figure \\@ref(fig:sphere-proj) f"})`).
-
-Each observation is situated on a restricted $4\text{-}D$ spherical surface, defined by spherical coordinates.
-The azimuthal angle $\theta_1 \sim U(0, \pi)$ in the $(x_1, x_2)$ plane, while the elevation angle $\theta_2 \sim U(0, \pi)$ in the $(x_2, x_3)$ plane.
+Finally, the `gen_hemisphere(n, p)` function restricts sampling to a hemisphere of a $4\text{-}D$ sphere (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) f"} else { "Figure \\@ref(fig:sphere-proj) f"})`).
+Using spherical coordinates, the azimuthal angle $\theta_1 \sim U(0, \pi)$ in the $(x_1, x_2)$ plane, while the elevation angle $\theta_2 \sim U(0, \pi)$ in the $(x_2, x_3)$ plane.
 Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.

### 2025-09-20 paper-cardinalR.tex
Updated sphere section

@@
 
-\subsubsection{Sphere}\label{sphere}
+\paragraph{Sphere}\label{sphere}
 
 Sphere-shaped structures (Figure \ref{fig:sphere-proj}) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
-The package provides functions to generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+The functions generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
+For the circle and curvy cycle, this occurs when \(p > 2\) and \(p > 3\), respectively, while for spheres and hemispheres it occurs when \(p > 3\) or \(p > 4\).
 Table \ref{tab:sphere-tb-pdf} summarizes these functions.
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_circle()}}{gen\_circle()}}\label{gen_circle}
-
-The function \texttt{gen\_circle(n,\ p)} generates a \(p\)-dimensional dataset of \(n\) observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale (Figure \ref{fig:sphere-proj} a).
-
-A latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
+The simplest case, \texttt{gen\_circle(n,\ p)} creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (Figure \ref{fig:sphere-proj} a).
+Let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
 Coordinates in the first two dimensions represent a perfect circle on the plane: \(X_1 = \cos(\theta), \quad X_2 = \sin(\theta)\).
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_curvycycle()}}{gen\_curvycycle()}}\label{gen_curvycycle}
-
-The \texttt{gen\_curvycycle(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations (Figure \ref{fig:sphere-proj} b).
-The structure forms a one-dimensional nonlinear cycle embedded in higher-dimensional space.
-
-A latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
+For the one-dimensional nonlinear cycle embedded in \(p\text{-}D\) space, \texttt{gen\_curvycycle(n,\ p)} (Figure \ref{fig:sphere-proj} b), let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
 The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle''.
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_unifsphere()}}{gen\_unifsphere()}}\label{gen_unifsphere}
-
-The \texttt{gen\_unifsphere(n,\ p,\ r)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on the surface of a \(3\text{-}D\) sphere of radius \(r\), with additional Gaussian noise dimensions added when \(p > 3\) (Figure \ref{fig:sphere-proj} c).
-
-Each observation lies on the surface of a sphere in \(3\text{-}D\), constructed by generating \(u \sim U(-1, 1)\) which represents the cosine of the polar angle \(\phi\) and \(\theta \sim U(0, 2\pi)\) which represents the azimuthal angle.
-The corresponding Cartesian coordinates are calculated as; \(X_1 = r\sqrt{1 - u^2}\cos(\theta)\), \(X_2 = r\sqrt{1 - u^2}\sin(\theta)\), \(X_3 = ru\), which gives points uniformly distributed on the surface of a \(3\text{-}D\) sphere (not within).
-For \(p > 3\), additional noise dimensions \(X_4\) to \(X_p\) are generated.
+Building on simple circular structures, the \texttt{gen\_unifsphere(n,\ p,\ r)} function function extends the idea to three dimensions by generating \(n\) observations approximately uniformly distributed on the surface of a sphere of radius \(r\), with optional high-dimensional noise when \(p > 3\) (Figure \ref{fig:sphere-proj} c).
+Each observation is computed from spherical coordinates, with \(u \sim U(-1, 1)\) representing \(\cos(\phi)\) and \(\theta \sim U(0, 2\pi)\) the azimuthal angle.
+Cartesian coordinates are then defined as \(X_1 = r\sqrt{1 - u^2}\cos(\theta)\), \(X_2 = r\sqrt{1 - u^2}\sin(\theta)\), and \(X_3 = ru\), ensuring uniform distribution on the surface (not within) of the sphere.
+For \(p > 3\), additional dimensions \(X_4\) through \(X_p\) are appended as structured noise.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_gridedsphere()}}{gen\_gridedsphere()}}\label{gen_gridedsphere}
+In addition, the \texttt{gen\_gridedsphere(n,\ p)} function constructs a \(p\text{-}D\) dataset consisting of approximately \(n\) points that are evenly distributed on the surface of the unit \((p-1)\)-sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
+The method relies on forming a regular grid in spherical coordinates, parameterized by \((p-1)\) angular variables: for dimensions \(j = 1, \dots, p-2\) the polar angles are drawn from \([0, \pi]\), while the final angle (\(j = p-1\)) represents the azimuth and is drawn from \([0, 2\pi]\).
+The number of grid steps along each angular dimension is chosen by decomposing \(n\) into \((p-1)\) approximately equal integer factors using the helper function \texttt{gen\_nproduct(n,\ p\ -\ 1)}.
 
-The \texttt{gen\_gridedsphere(n,\ p)} function generates a \(p\)-dimensional dataset of approximately \(n\) points evenly distributed on the surface of a \((p!-!1)\)-sphere (the unit sphere embedded in \(\mathbb{R}^p\)) (Figure \ref{fig:sphere-proj} d).
-
-The construction is based on a regular grid in spherical coordinates with \((p-1)\) angular variables.
-Specifically, for dimensions \(j = 1, \dots, p-2\), the polar angles are defined on \([0, \pi]\).
-For the last angle (\(j = p-1\)), the azimuthal angle is defined on \([0, 2\pi]\).
-
-The number of grid steps along each angular dimension is determined by factoring \(n\) into \((p-1)\) approximately equal integers via \texttt{gen\_nproduct(n,\ p\ -\ 1)}.
-
-Each grid point is then mapped to Cartesian coordinates using the standard hyperspherical-to-Cartesian transformation:
+Each grid point is subsequently mapped into Cartesian space via the standard hyperspherical-to-Cartesian transformation,
 
@@
 
-Thus, the output is a deterministic grid of points lying exactly on the unit \((p-1)\)-sphere, with no added noise dimensions.
+The result is a deterministic grid of points lying exactly on the surface of the unit \((p-1)\)-sphere, without any additional noise dimensions.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_clusteredspheres()}}{gen\_clusteredspheres()}}\label{gen_clusteredspheres}
-
-The \texttt{gen\_clusteredspheres(n,\ k,\ p,\ r,\ loc)} function generates a synthetic dataset of \(n_1 + kn_2\) observations in \(p\)-dimensional space, consisting of one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location (Figure \ref{fig:sphere-proj} e).
-
+For more heterogeneous structures, the \texttt{gen\_clusteredspheres(n,\ k,\ p,\ r,\ loc)} function generates one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location (Figure \ref{fig:sphere-proj} e).
 A large uniform sphere centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\text{-}D\) sphere with a radius of \(r_1\).
 The sampling is executed using the function \texttt{gen\_unifsphere(n\_1,\ p,\ r\_1)}, which generates the desired points in the specified dimensional space.
-
 In generation of \(k\) smaller uniform spheres, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
 These spheres are positioned at distinct random locations in \(p\)-space, with the center of each sphere being drawn from a normal distribution \(N(0, \texttt{loc}^2 I_p)\).
-
 Points on spheres are generated using the standard hyperspherical method, which involves sampling \(u \sim U(-1, 1)\) to determine the cosine of the polar angle, and sampling \(\theta \sim U(0, 2\pi)\) to determine the azimuthal angle (for \(3\text{-}D\)).
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_hemisphere()}}{gen\_hemisphere()}}\label{gen_hemisphere}
-
-The \texttt{gen\_hemisphere(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations distributed approximately uniformly on a \(4\text{-}D\) hemisphere, optionally extended with Gaussian noise in additional dimensions when \(p > 4\) (Figure \ref{fig:sphere-proj} f).
-
-Each observation is situated on a restricted \(4\text{-}D\) spherical surface, defined by spherical coordinates.
-The azimuthal angle \(\theta_1 \sim U(0, \pi)\) in the \((x_1, x_2)\) plane, while the elevation angle \(\theta_2 \sim U(0, \pi)\) in the \((x_2, x_3)\) plane.
+Finally, the \texttt{gen\_hemisphere(n,\ p)} function restricts sampling to a hemisphere of a \(4\text{-}D\) sphere (Figure \ref{fig:sphere-proj} f).
+Using spherical coordinates, the azimuthal angle \(\theta_1 \sim U(0, \pi)\) in the \((x_1, x_2)\) plane, while the elevation angle \(\theta_2 \sim U(0, \pi)\) in the \((x_2, x_3)\) plane.
 Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
@@
 
-\subsubsection{Trigonometric}\label{trigonometric}
+\paragraph{Trigonometric}\label{trigonometric}
 
@@
 
-\subsubsection{Trefoil knots}\label{trefoil-knots}
+\paragraph{Trefoil knots}\label{trefoil-knots}
 

### 2025-09-20 paper-cardinalR.Rmd
Added noise dimensions section

@@
 
-generate noise dimensions with a normal distribution and various wavy patterns,
+High-dimensional data structures often benefit from the addition of auxiliary noise dimensions, which can be used to assess the robustness of dimensionality reduction and clustering algorithms.
+The functions in this section provide flexible ways to generate random noise dimensions, ranging from purely random Gaussian variables to more structured, wavy patterns that mimic non-linear distortions in high-dimensional space.
+These functions can be applied independently or combined with other geometric structures to create complex simulated datasets.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:noise-tb-html)', '\\@ref(tab:noise-tb-pdf)'))` details these functions.
+
+```{r}
+noise_fun_tb <- tibble(fun = c("gen_noisedims",
+                            "gen_wavydims1",
+                            "gen_wavydims2",
+                            "gen_wavydims3"), 
+                        exp = c("Gaussian noise dimensions with optional mean and standard deviation.",
+                                "Wavy noise dimensions based on a user-specified theta sequence with added jitter.",
+                                "Wavy noise dimensions using polynomial transformations of an existing dimension vector.",
+                                "Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset."))
+```
+
+```{r noise-tb-html, eval=knitr::is_html_output()}
+noise_fun_tb |> 
+  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r noise-tb-pdf, eval=knitr::is_latex_output()}
+noise_fun_tb |> 
+  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+The `gen_noisedims(n, p, m, s)` function generates \$p\$ independent Gaussian noise dimensions,
+
+$$
+X_j \sim N(m_j, s_j^2), \quad j = 1, \dots, p,
+$$
+
+with odd-numbered dimensions multiplied by $-1$ to introduce sign alternation, enhancing variability and decorrelation.
+
+For scenarios where noise should follow a smooth wavy pattern, `gen_wavydims1(n, p, \theta)` generates dimensions as
+
+$$
+X_j = \alpha_j \theta + \varepsilon_j, \quad \varepsilon_j \sim N(0, \sigma^2), \quad j = 1, \dots, p,
+$$
+
+where each dimension is scaled by a different factor $\alpha_j$, producing structured noise that oscillates along the latent parameter $\theta$, mimicking trends or trajectories observed in real-world data.
+
+The `gen_wavydims2(n, p, x_1)` function extends this approach by applying a non-linear transformation to an existing dimension vector $x_1$:
+
+$$
+X_j = \beta_j \, (-1)^{\lfloor j/2 \rfloor} \, x_1^{k_j} + \varepsilon_j, \quad j = 1, \dots, p,
+$$
+
+where $k_j$ is a randomly chosen polynomial power, $\beta_j$ is a scaling factor, and $\varepsilon_j$ is small uniform noise.
+
+Finally, `gen_wavydims3(n, p, \text{data})` generates noise for datasets with multiple correlated dimensions.
+The first three dimensions are small perturbations of the original coordinates $(X_1, X_2, X_3)$, while higher dimensions are constructed via non-linear combinations, including polynomial and trigonometric transformations, e.g.,
+
+$$
+X_j = f_j(X_1, X_2, X_3) + \varepsilon_j, \quad j > 3,
+$$
+
+producing high-dimensional noise that preserves some geometric correlation with the base structure while introducing additional complexity.
 
@@
 ```{r}
-add_fun_tb <- tibble(fun = c("gen_noisedims",
-                            "gen_bkgnoise",
+add_fun_tb <- tibble(fun = c("gen_bkgnoise",
                             "randomize_rows",
@@
                             "gen_nsum",
-                            "gen_wavydims1",
-                            "gen_wavydims2",
-                            "gen_wavydims3",
                             "gen_rotation",
                             "normalize_data"), 
-                        exp = c("Generates additional noise dimensions.",
-                                "Adds background noise.",
+                        exp = c("Adds background noise.",
                                 "Randomizes the rows.",
@@
                                 "Generates a vector of positive integers whose summation is approximately equal to a target value.", 
-                                "Generates random noise dimensions with wavy pattern generated with theta.",
-                                "Generates random noise dimensions with wavy pattern generated with power functions.",
-                                "Generates random noise dimensions with wavy pattern generated with power and sine functions.",
                                 "Generates rotations.", 

### 2025-09-20 paper-cardinalR.tex
Added noise dimensions section

@@
 
-generate noise dimensions with a normal distribution and various wavy patterns,
+High-dimensional data structures often benefit from the addition of auxiliary noise dimensions, which can be used to assess the robustness of dimensionality reduction and clustering algorithms.
+The functions in this section provide flexible ways to generate random noise dimensions, ranging from purely random Gaussian variables to more structured, wavy patterns that mimic non-linear distortions in high-dimensional space.
+These functions can be applied independently or combined with other geometric structures to create complex simulated datasets.
+Table \ref{tab:noise-tb-pdf} details these functions.
+
+\begin{table}
+
+\caption{\label{tab:noise-tb-pdf}cardinalR noise dimensions generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_noisedims & Gaussian noise dimensions with optional mean and standard deviation.\\
+gen\_wavydims1 & Wavy noise dimensions based on a user-specified theta sequence with added jitter.\\
+gen\_wavydims2 & Wavy noise dimensions using polynomial transformations of an existing dimension vector.\\
+gen\_wavydims3 & Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The \texttt{gen\_noisedims(n,\ p,\ m,\ s)} function generates \$p\$ independent Gaussian noise dimensions,
+
+\[
+X_j \sim N(m_j, s_j^2), \quad j = 1, \dots, p,
+\]
+
+with odd-numbered dimensions multiplied by \(-1\) to introduce sign alternation, enhancing variability and decorrelation.
+
+For scenarios where noise should follow a smooth wavy pattern, \texttt{gen\_wavydims1(n,\ p,\ \textbackslash{}theta)} generates dimensions as
+
+\[
+X_j = \alpha_j \theta + \varepsilon_j, \quad \varepsilon_j \sim N(0, \sigma^2), \quad j = 1, \dots, p,
+\]
+
+where each dimension is scaled by a different factor \(\alpha_j\), producing structured noise that oscillates along the latent parameter \(\theta\), mimicking trends or trajectories observed in real-world data.
+
+The \texttt{gen\_wavydims2(n,\ p,\ x\_1)} function extends this approach by applying a non-linear transformation to an existing dimension vector \(x_1\):
+
+\[
+X_j = \beta_j \, (-1)^{\lfloor j/2 \rfloor} \, x_1^{k_j} + \varepsilon_j, \quad j = 1, \dots, p,
+\]
+
+where \(k_j\) is a randomly chosen polynomial power, \(\beta_j\) is a scaling factor, and \(\varepsilon_j\) is small uniform noise.
+
+Finally, \texttt{gen\_wavydims3(n,\ p,\ \textbackslash{}text\{data\})} generates noise for datasets with multiple correlated dimensions.
+The first three dimensions are small perturbations of the original coordinates \((X_1, X_2, X_3)\), while higher dimensions are constructed via non-linear combinations, including polynomial and trigonometric transformations, e.g.,
+
+\[
+X_j = f_j(X_1, X_2, X_3) + \varepsilon_j, \quad j > 3,
+\]
+
+producing high-dimensional noise that preserves some geometric correlation with the base structure while introducing additional complexity.
 
@@
 \midrule
-gen\_noisedims & Generates additional noise dimensions.\\
 gen\_bkgnoise & Adds background noise.\\
@@
 gen\_nproduct & Generates a vector of positive integers whose product is approximately equal to a target value.\\
+gen\_nsum & Generates a vector of positive integers whose summation is approximately equal to a target value.\\
 \addlinespace
-gen\_nsum & Generates a vector of positive integers whose summation is approximately equal to a target value.\\
-gen\_wavydims1 & Generates random noise dimensions with wavy pattern generated with theta.\\
-gen\_wavydims2 & Generates random noise dimensions with wavy pattern generated with power functions.\\
-gen\_wavydims3 & Generates random noise dimensions with wavy pattern generated with power and sine functions.\\
 gen\_rotation & Generates rotations.\\
-\addlinespace
 normalize\_data & Normalizes data.\\

### 2025-09-20 paper-cardinalR.Rmd
Updated branching function table

@@
 A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (@trapnell2014).
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points, and number of branches.
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (`n`), and number of branches (`k`).
 Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
@@
                              "gen_orgcurvybranches"), 
-                      exp = c("Generate a structure with exponential shaped branches.",
-                               "Generate a structure with linear shaped branches.", 
-                               "Generate a structure with curvy shaped branches.",
-                               "Generate a structure with linear shaped branches originated in one point.",
-                               "Generate a structure with curvy shaped branches originated in one point."))
+                      exp = c("Exponential shaped branches.",
+                               "Linear shaped branches.", 
+                               "Curvy shaped branches.",
+                               "Linear shaped branches originated in one point.",
+                               "Curvy shaped branches originated in one point."))
 ```

### 2025-09-20 paper-cardinalR.tex
Updated branching function table

@@
 A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (\citet{trapnell2014}).
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points, and number of branches.
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (\texttt{n}), and number of branches (\texttt{k}).
 Although these functions can generate multiple branches, they do not produce a formal \emph{multicluster} dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.

### 2025-09-20 paper-cardinalR.R

@@ -133,11 +133,11 @@ branch_tb <- tibble(fun = c("gen_expbranches",
                              "gen_curvybranches", 
                              "gen_orglinearbranches", 
                              "gen_orgcurvybranches"), 
-                      exp = c("Generate a structure with exponential shaped branches.",
-                               "Generate a structure with linear shaped branches.", 
-                               "Generate a structure with curvy shaped branches.",
-                               "Generate a structure with linear shaped branches originated in one point.",
-                               "Generate a structure with curvy shaped branches originated in one point."))
+                      exp = c("Exponential shaped branches.",
+                               "Linear shaped branches.", 
+                               "Curvy shaped branches.",
+                               "Linear shaped branches originated in one point.",
+                               "Curvy shaped branches originated in one point."))
 
 
 ## ----branching-tb-html, eval=knitr::is_html_output()--------------------------

### 2025-09-20 paper-cardinalR.Rmd
Updated branching section

@@
 
-### Branching
+#### Branching
 
@@
 
-#### `gen_expbranches()`
-
-The `gen_expbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ exponential branches in $2\text{-}D$, with optional noise dimensions to embed the structure in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) a"} else { "Figure \\@ref(fig:branch-proj) a"})`).
-These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
-
-Each branch $i$ is constructed using $X_1 \sim U(-2, 2)$ and $X_2 = \exp(\pm s_iX_1) + \epsilon$, where: $\epsilon \sim U(0, 0.1)$ adds local jitter, $s_i \in [0.5, 2]$ is randomly sampled for each branch.
-The sign of the exponent alternates between branches: odd-numbered branches have a negative exponent (decaying pattern), and even-numbered branches have a positive exponent (growing pattern), producing mirror-symmetric curves with varying steepness.
-For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
+The simplest structures are approximately linear branches, generated by the `gen_linearbranches(n, p, k)` function (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) b"} else { "Figure \\@ref(fig:branch-proj) b"})`).
+These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability.
+Mathematically, each branch $i$ is defined as
+
+$$
+X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
+$$
+
+where $(x_{\text{start},i}, y_{\text{start},i})$ is the starting point of branch $i$, $\delta$ controls local jitter, and $s_i$ is the slope, initialized as
+
+$$
+s_i =
+\begin{cases}
+0.5 & i = 1, \\
+-0.5 & i = 2, \\
+\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
+\end{cases}
+$$
+
+Branches $1$ and $2$ are initialized with fixed slopes and intercepts, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
+
+```{r data-linearbranches, echo=TRUE}
+linearbranches <- gen_linearbranches(n = 1000, p = 4, k = 4)
+```
+
+```{r}
+linearbranches_lang <- langevitour::langevitour(linearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r linearbranches-proj1}
+scaled_data <- scale_data_manual(linearbranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+linearbranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linearbranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+linearbranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.7), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linearbranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
+                                              threshold = 0.016))
+
+linearbranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.65), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+To introduce curvature, the `gen_curvybranches(n, p, k)` function generates $k$ curvilinear branches in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) c"} else { "Figure \\@ref(fig:branch-proj) c"})`).
+Branches $1$ and $2$ are simple parabolas defined as
+
+$$
+\begin{aligned}
+\text{Branch 1: } & X_1 \sim U(0,1), \quad X_2 = 0.1 X_1 + X_1^2 + \epsilon, \\
+\text{Branch 2: } & X_1 \sim U(-1,0), \quad X_2 = 0.1 X_1 - 2 X_1^2 + \epsilon, \quad \epsilon \sim U(0, \delta),
+\end{aligned}
+$$
+
+where $\delta$ controls local jitter.
+Additional branches are attached iteratively to existing structures.
+Each new branch $i$ starts at a selected point $(x_{\text{start},i}, y_{\text{start},i})$ from the current structure and extends according to
+
+$$
+X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
+$$
+
+where $s_i$ is a scale factor controlling the curvature of branch $i$.
+For the first few initial branches, $s_i$ can be fixed (e.g., $s_1 = 1, s_2 = 2$), while for subsequent branches it is sampled from a predefined set, such as $s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}$, to create variability in curvature.
+
+```{r data-curvybranches, echo=TRUE}
+curvybranches <- gen_curvybranches(n = 1000, p = 4, k = 4)
+```
+
+```{r}
+curvybranches_lang <- langevitour::langevitour(curvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+```
+
+```{r curvybranches-proj1}
+scaled_data <- scale_data_manual(curvybranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
+                                              threshold = 0.022))
+
+curvybranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.8, 0.7), 
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvybranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.02))
+
+curvybranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.3), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvybranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.016))
+
+curvybranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.4), 
+  title = "c3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+The `gen_expbranches(n, p, k)` function creates $k$ exponential branches in $2\text{-}D$, radiating from a central region (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) a"} else { "Figure \\@ref(fig:branch-proj) a"})`).
+Each branch $i$ is defined as
+
+$$
+X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
+$$
+
+where $\sigma_i = (-1)^{i+1}$ alternates the sign of the exponent to produce mirror-symmetric branches.
+The parameter $s_i$ controls the steepness of branch $i$, and $\delta$ introduces small local jitter.
 
@@
 
-#### `gen_linearbranches()`
-
-The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) b"} else { "Figure \\@ref(fig:branch-proj) b"})`).
-The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.
-
-The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points in $p\text{-}D$ forming $k$ linear branches.
-Each branch is a line segment with added jitter to simulate noise.
-Branches $1$ and $2$ are initialized with fixed slopes and intercepts: branch $1$ has slope $0.5$ and intercept $0$, with $X_1 \sim U(-2, 8)$ and $X_2 = 0.5 X_1 + \epsilon$, $\epsilon \sim U(0, 0.5)$; branch $2$ has slope $-0.5$ and intercept $0$, with $X_1 \sim U(-6, 2)$ and $X_2 = -0.5 X_1 + \epsilon$, $\epsilon \sim U(0, 0.5)$.
-
-Branches $3$ to $k$ are added iteratively.
-Each new branch starts at a location outside predefined exclusion zones to avoid overlap with the first two branches.
-Specifically, $X_1$ starting points are sampled outside the intervals $[-8, -7]$, $[-2, 2]$, and $[7, 8]$, and $X_2$ starting points are sampled outside $[7, 8]$.
-For each new branch $i$, $X_1$ values are drawn over a short interval $[x_{\text{start}}, x_{\text{start}} + 1]$, and $X_2$ values are calculated as $X_2 = s_i (X_1 - x_{\text{start}}) + y_{\text{start}} + \epsilon$, where $s_i$ is a randomly chosen slope from a filtered sequence and $\epsilon \sim U(0, 0.2)$.
-For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, for $j = 3, \dots, p$.
-
-```{r data-linearbranches, echo=TRUE}
-linearbranches <- gen_linearbranches(n = 1000, p = 4, k = 4)
-```
-
-```{r}
-linearbranches_lang <- langevitour::langevitour(linearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r linearbranches-proj1}
-scaled_data <- scale_data_manual(linearbranches)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-linearbranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linearbranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-linearbranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linearbranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-linearbranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.65), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-#### `gen_curvybranches()`
-
-The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) c"} else { "Figure \\@ref(fig:branch-proj) c"})`).
-The underlying geometry lies in the first two dimensions, while the remaining $(p-2)$ dimensions contain Gaussian noise.
-
-Branch $1$ is generated from $X_1 \sim U(0, 1)$, and $X_2 = 0.1X_1 + X_1^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
-This produces a gently upward-curving parabola in the right half-plane.
-Branch $2$ is generated from $X_{1} \sim U(-1, 0)$, and $X_2 = 0.1X_{1} - 2X_{1}^2 + \epsilon$, where $\epsilon \sim U(0, 0.05)$.
-This creates a steeper, leftward-facing curve in the left half-plane.
-Branches $3$ to $k$ are then added iteratively.
-Each new branch $i$ begins at a randomly chosen starting point $(x_{\text{start}}, y_{\text{start}})$ from the existing structure, where the horizontal coordinate $x_{\text{start}}$ is restricted to $[-0.15, 0.15]$ to ensure connectivity with the initial branches.
-From this starting point, the branch extends over a unit-length interval $X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]$, with structure $X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, ; s_i \in {-2, -1.5, -1, -0.5, 0, 0.5, 1.5}$.
-Here, each branch $i$ is assigned a single scale factor $s_i$, sampled independently from the set, which controls the curvature of that branch.
-The parameter $(x_{\text{start}}, y_{\text{start}})$ is the starting point chosen from the existing structure.
-This construction yields $k$ spatially connected, nonlinear branches with varying curvature.
-For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.
-
-```{r data-curvybranches, echo=TRUE}
-curvybranches <- gen_curvybranches(n = 1000, p = 4, k = 4)
-```
-
-```{r}
-curvybranches_lang <- langevitour::langevitour(curvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r curvybranches-proj1}
-scaled_data <- scale_data_manual(curvybranches)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.022))
-
-curvybranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.7), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvybranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.02))
-
-curvybranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.3), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvybranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-curvybranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.4), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-#### `gen_orglinearbranches()`
-
-The `gen_orglinearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches embedded in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) d"} else { "Figure \\@ref(fig:branch-proj) d"})`).
-Each branch lies within a distinct $2\text{-}D$ subspace, while the remaining $p - 2$ dimensions contain Gaussian noise.
-
-To construct each branch, a unique or repeated pair of dimensions is selected from the $\binom{p}{2}$ possible combinations of dimensions.
-If $k \leq \binom{p}{2}$, combinations are sampled without replacement.
-If $k > \binom{p}{2}$, additional pairs are sampled with replacement to reach $k$ total branches.
-Each selected pair $(i_1, i_2)$ defines the $2\text{-}D$ plane for branch $i$.
-
-For branch $i$, $n_i$ points are generated, where $\sum_{i=1}^k n_i = n$.
-The structure follows: $X_{i_1} \sim U(0, 2)$, and $X_{i_2} = -s_iX_{i_1} + \epsilon, \quad \epsilon \sim U(0, 0.5)$, where $s_i$ is a scale factor controlling the slope.
-When $k \leq \binom{p}{2}$, $s_i = 1$.
-When sampling with replacement, $s_i$ is drawn from the set ${1, 1.5, 2, \dots, 8}$ in increments of $0.5$.
-For $p > 2$, the remaining dimensions contain independent Gaussian noise: $X_j \sim N(0, 0.1^2)$, for $j \notin \{i_1, i_2\}$.
+High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` and `gen_orgcurvybranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) d-e"} else { "Figure \\@ref(fig:branch-proj) d-e"})`).
+Each branch is embedded in a unique or repeated $2\text{-}D$ subspace of the $p\text{-}D$ space.
+Linear branches follow
+
+$$
+X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = s_i X_{i_1} + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+$$
+
+while curvilinear branches include a quadratic term
+
+$$
+X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = - s_i X_{i_1}^2 + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+$$
+
+where $a_i, b_i$ define the range of the first coordinate for branch $i$, and $\epsilon$ is Gaussian noise added to introduce variability.
+The scale factor $s_i$ controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first $\binom{p}{2}$ branches, $s_i = 1$; for additional branches when $k > \binom{p}{2}$, $s_i$ is randomly drawn from the set $\{1, 1.5, 2, \dots, 8\}$.
 
@@
 
-#### `gen_orgcurvybranches()`
-
-The `gen_orgcurvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in a $p$-dimensional space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) e"} else { "Figure \\@ref(fig:branch-proj) e"})`).
-Each branch is constructed in a unique or repeated $2\text{-}D$ subspace of the $p$-dimensional space, with curvature induced by a second-degree polynomial structure.
-The remaining $p - 2$ dimensions contain Gaussian noise.
-
-Let $\binom{p}{2}$ denote the number of unique $2\text{-}D$ subspace combinations.
-When $k \leq \binom{p}{2}$, $k$ distinct subspace pairs $(i_1, i_2)$ are sampled without replacement.
-Otherwise, combinations are selected with replacement to reach $k$ total branches.
-For each branch $i = 1, \dots, k$, a random scale factor $s_i$ is used to vary the curvature; If $k \leq \binom{p}{2}$: $s_i = 1$, or If $k > \binom{p}{2}$: $s_i \in \{1, 1.5, \dots, 8\}$.
-
-Each branch contains $n_i$ points such that $\sum_{i=1}^k n_i = n$, where the vector $(n_1, \dots, n_k)$ is randomly drawn using the helper function `gen_nsum()` to partition $n$.
-Within its assigned subspace, branch $i$ is defined by: $X_{i_1} \sim U(0, 2)$, and $X_{i_2} = -s_iX_{i_1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)$.
-This forms a smooth downward-opening parabola in the plane defined by $(X_{i_1}, X_{i_2})$, with the degree of curvature controlled by $s_i$.
-For $p > 2$, all remaining dimensions are independent Gaussian noise: $X_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i_1, i_2\}$.
-
 ```{r data-orgcurvybranches, echo=TRUE}
@@
 ```
+
+For all branch-generating functions, when the target dimensionality $p > 2$, additional dimensions are filled with independent noise, allowing the $2\text{-}D$ or $2\text{-}D$-subspace structures to be naturally embedded in higher-dimensional space while preserving the intended geometric patterns.
 

### 2025-09-20 paper-cardinalR.tex
Updated branching section

@@
 
-\subsubsection{Branching}\label{branching}
+\paragraph{Branching}\label{branching}
 
@@
 \midrule
-gen\_expbranches & Generate a structure with exponential shaped branches.\\
-gen\_linearbranches & Generate a structure with linear shaped branches.\\
-gen\_curvybranches & Generate a structure with curvy shaped branches.\\
-gen\_orglinearbranches & Generate a structure with linear shaped branches originated in one point.\\
-gen\_orgcurvybranches & Generate a structure with curvy shaped branches originated in one point.\\
+gen\_expbranches & Exponential shaped branches.\\
+gen\_linearbranches & Linear shaped branches.\\
+gen\_curvybranches & Curvy shaped branches.\\
+gen\_orglinearbranches & Linear shaped branches originated in one point.\\
+gen\_orgcurvybranches & Curvy shaped branches originated in one point.\\
 \bottomrule
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_expbranches()}}{gen\_expbranches()}}\label{gen_expbranches}
+The simplest structures are approximately linear branches, generated by the \texttt{gen\_linearbranches(n,\ p,\ k)} function (Figure \ref{fig:branch-proj} b).
+These consist of \(k\) short line segments in the first two dimensions, with added jitter to simulate variability.
+Mathematically, each branch \(i\) is defined as
 
-The \texttt{gen\_expbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) exponential branches in \(2\text{-}D\), with optional noise dimensions to embed the structure in \(p\text{-}D\) (Figure \ref{fig:branch-proj} a).
-These branches grow in opposite directions, producing a radiating curvilinear structure from a central region.
+\[
+X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
+\]
 
-Each branch \(i\) is constructed using \(X_1 \sim U(-2, 2)\) and \(X_2 = \exp(\pm s_iX_1) + \epsilon\), where: \(\epsilon \sim U(0, 0.1)\) adds local jitter, \(s_i \in [0.5, 2]\) is randomly sampled for each branch.
-The sign of the exponent alternates between branches: odd-numbered branches have a negative exponent (decaying pattern), and even-numbered branches have a positive exponent (growing pattern), producing mirror-symmetric curves with varying steepness.
-For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
+where \((x_{\text{start},i}, y_{\text{start},i})\) is the starting point of branch \(i\), \(\delta\) controls local jitter, and \(s_i\) is the slope, initialized as
+
+\[
+s_i =
+\begin{cases}
+0.5 & i = 1, \\
+-0.5 & i = 2, \\
+\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
+\end{cases}
+\]
+
+Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
+
+\begin{verbatim}
+linearbranches <- gen_linearbranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+To introduce curvature, the \texttt{gen\_curvybranches(n,\ p,\ k)} function generates \(k\) curvilinear branches in \(p\text{-}D\) (Figure \ref{fig:branch-proj} c).
+Branches \(1\) and \(2\) are simple parabolas defined as
+
+\[
+\begin{aligned}
+\text{Branch 1: } & X_1 \sim U(0,1), \quad X_2 = 0.1 X_1 + X_1^2 + \epsilon, \\
+\text{Branch 2: } & X_1 \sim U(-1,0), \quad X_2 = 0.1 X_1 - 2 X_1^2 + \epsilon, \quad \epsilon \sim U(0, \delta),
+\end{aligned}
+\]
+
+where \(\delta\) controls local jitter.
+Additional branches are attached iteratively to existing structures.
+Each new branch \(i\) starts at a selected point \((x_{\text{start},i}, y_{\text{start},i})\) from the current structure and extends according to
+
+\[
+X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
+\]
+
+where \(s_i\) is a scale factor controlling the curvature of branch \(i\).
+For the first few initial branches, \(s_i\) can be fixed (e.g., \(s_1 = 1, s_2 = 2\)), while for subsequent branches it is sampled from a predefined set, such as \(s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}\), to create variability in curvature.
+
+\begin{verbatim}
+curvybranches <- gen_curvybranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+The \texttt{gen\_expbranches(n,\ p,\ k)} function creates \(k\) exponential branches in \(2\text{-}D\), radiating from a central region (Figure \ref{fig:branch-proj} a).
+Each branch \(i\) is defined as
+
+\[
+X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
+\]
+
+where \(\sigma_i = (-1)^{i+1}\) alternates the sign of the exponent to produce mirror-symmetric branches.
+The parameter \(s_i\) controls the steepness of branch \(i\), and \(\delta\) introduces small local jitter.
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_linearbranches()}}{gen\_linearbranches()}}\label{gen_linearbranches}
+High-dimensional generalizations are provided by \texttt{gen\_orglinearbranches(n,\ p,\ k)} and \texttt{gen\_orgcurvybranches(n,\ p,\ k)} (Figure \ref{fig:branch-proj} d-e).
+Each branch is embedded in a unique or repeated \(2\text{-}D\) subspace of the \(p\text{-}D\) space.
+Linear branches follow
 
-The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches in \(p\text{-}D\) (Figure \ref{fig:branch-proj} b).
-The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.
+\[
+X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = s_i X_{i_1} + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+\]
 
-The \texttt{gen\_linearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points in \(p\text{-}D\) forming \(k\) linear branches.
-Each branch is a line segment with added jitter to simulate noise.
-Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts: branch \(1\) has slope \(0.5\) and intercept \(0\), with \(X_1 \sim U(-2, 8)\) and \(X_2 = 0.5 X_1 + \epsilon\), \(\epsilon \sim U(0, 0.5)\); branch \(2\) has slope \(-0.5\) and intercept \(0\), with \(X_1 \sim U(-6, 2)\) and \(X_2 = -0.5 X_1 + \epsilon\), \(\epsilon \sim U(0, 0.5)\).
+while curvilinear branches include a quadratic term
 
-Branches \(3\) to \(k\) are added iteratively.
-Each new branch starts at a location outside predefined exclusion zones to avoid overlap with the first two branches.
-Specifically, \(X_1\) starting points are sampled outside the intervals \([-8, -7]\), \([-2, 2]\), and \([7, 8]\), and \(X_2\) starting points are sampled outside \([7, 8]\).
-For each new branch \(i\), \(X_1\) values are drawn over a short interval \([x_{\text{start}}, x_{\text{start}} + 1]\), and \(X_2\) values are calculated as \(X_2 = s_i (X_1 - x_{\text{start}}) + y_{\text{start}} + \epsilon\), where \(s_i\) is a randomly chosen slope from a filtered sequence and \(\epsilon \sim U(0, 0.2)\).
-For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), for \(j = 3, \dots, p\).
+\[
+X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = - s_i X_{i_1}^2 + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+\]
 
-\begin{verbatim}
-linearbranches <- gen_linearbranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
-
-\paragraph{\texorpdfstring{\texttt{gen\_curvybranches()}}{gen\_curvybranches()}}\label{gen_curvybranches}
-
-The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in \(p\text{-}D\) (Figure \ref{fig:branch-proj} c).
-The underlying geometry lies in the first two dimensions, while the remaining \((p-2)\) dimensions contain Gaussian noise.
-
-Branch \(1\) is generated from \(X_1 \sim U(0, 1)\), and \(X_2 = 0.1X_1 + X_1^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
-This produces a gently upward-curving parabola in the right half-plane.
-Branch \(2\) is generated from \(X_{1} \sim U(-1, 0)\), and \(X_2 = 0.1X_{1} - 2X_{1}^2 + \epsilon\), where \(\epsilon \sim U(0, 0.05)\).
-This creates a steeper, leftward-facing curve in the left half-plane.
-Branches \(3\) to \(k\) are then added iteratively.
-Each new branch \(i\) begins at a randomly chosen starting point \((x_{\text{start}}, y_{\text{start}})\) from the existing structure, where the horizontal coordinate \(x_{\text{start}}\) is restricted to \([-0.15, 0.15]\) to ensure connectivity with the initial branches.
-From this starting point, the branch extends over a unit-length interval \(X_1 \in [x_{\text{start}}, x_{\text{start}} + 1]\), with structure \(X_2 = 0.1X_1 - s_i(X_1^2 - x_{\text{start}}) + y_{\text{start}}, ; s_i \in {-2, -1.5, -1, -0.5, 0, 0.5, 1.5}\).
-Here, each branch \(i\) is assigned a single scale factor \(s_i\), sampled independently from the set, which controls the curvature of that branch.
-The parameter \((x_{\text{start}}, y_{\text{start}})\) is the starting point chosen from the existing structure.
-This construction yields \(k\) spatially connected, nonlinear branches with varying curvature.
-For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).
-
-\begin{verbatim}
-curvybranches <- gen_curvybranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
-
-\paragraph{\texorpdfstring{\texttt{gen\_orglinearbranches()}}{gen\_orglinearbranches()}}\label{gen_orglinearbranches}
-
-The \texttt{gen\_orglinearbranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) approximately linear branches embedded in \(p\text{-}D\) (Figure \ref{fig:branch-proj} d).
-Each branch lies within a distinct \(2\text{-}D\) subspace, while the remaining \(p - 2\) dimensions contain Gaussian noise.
-
-To construct each branch, a unique or repeated pair of dimensions is selected from the \(\binom{p}{2}\) possible combinations of dimensions.
-If \(k \leq \binom{p}{2}\), combinations are sampled without replacement.
-If \(k > \binom{p}{2}\), additional pairs are sampled with replacement to reach \(k\) total branches.
-Each selected pair \((i_1, i_2)\) defines the \(2\text{-}D\) plane for branch \(i\).
-
-For branch \(i\), \(n_i\) points are generated, where \(\sum_{i=1}^k n_i = n\).
-The structure follows: \(X_{i_1} \sim U(0, 2)\), and \(X_{i_2} = -s_iX_{i_1} + \epsilon, \quad \epsilon \sim U(0, 0.5)\), where \(s_i\) is a scale factor controlling the slope.
-When \(k \leq \binom{p}{2}\), \(s_i = 1\).
-When sampling with replacement, \(s_i\) is drawn from the set \({1, 1.5, 2, \dots, 8}\) in increments of \(0.5\).
-For \(p > 2\), the remaining dimensions contain independent Gaussian noise: \(X_j \sim N(0, 0.1^2)\), for \(j \notin \{i_1, i_2\}\).
+where \(a_i, b_i\) define the range of the first coordinate for branch \(i\), and \(\epsilon\) is Gaussian noise added to introduce variability.
+The scale factor \(s_i\) controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first \(\binom{p}{2}\) branches, \(s_i = 1\); for additional branches when \(k > \binom{p}{2}\), \(s_i\) is randomly drawn from the set \(\{1, 1.5, 2, \dots, 8\}\).
 
@@
 
-\paragraph{\texorpdfstring{\texttt{gen\_orgcurvybranches()}}{gen\_orgcurvybranches()}}\label{gen_orgcurvybranches}
-
-The \texttt{gen\_orgcurvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in a \(p\)-dimensional space (Figure \ref{fig:branch-proj} e).
-Each branch is constructed in a unique or repeated \(2\text{-}D\) subspace of the \(p\)-dimensional space, with curvature induced by a second-degree polynomial structure.
-The remaining \(p - 2\) dimensions contain Gaussian noise.
-
-Let \(\binom{p}{2}\) denote the number of unique \(2\text{-}D\) subspace combinations.
-When \(k \leq \binom{p}{2}\), \(k\) distinct subspace pairs \((i_1, i_2)\) are sampled without replacement.
-Otherwise, combinations are selected with replacement to reach \(k\) total branches.
-For each branch \(i = 1, \dots, k\), a random scale factor \(s_i\) is used to vary the curvature; If \(k \leq \binom{p}{2}\): \(s_i = 1\), or If \(k > \binom{p}{2}\): \(s_i \in \{1, 1.5, \dots, 8\}\).
-
-Each branch contains \(n_i\) points such that \(\sum_{i=1}^k n_i = n\), where the vector \((n_1, \dots, n_k)\) is randomly drawn using the helper function \texttt{gen\_nsum()} to partition \(n\).
-Within its assigned subspace, branch \(i\) is defined by: \(X_{i_1} \sim U(0, 2)\), and \(X_{i_2} = -s_iX_{i_1}^2 + \epsilon, \quad \epsilon \sim U(0, 0.5)\).
-This forms a smooth downward-opening parabola in the plane defined by \((X_{i_1}, X_{i_2})\), with the degree of curvature controlled by \(s_i\).
-For \(p > 2\), all remaining dimensions are independent Gaussian noise: \(X_j \sim N(0, 0.1^2), \quad \text{for } j \notin \{i_1, i_2\}\).
-
 \begin{verbatim}
@@
 \end{verbatim}
+
+For all branch-generating functions, when the target dimensionality \(p > 2\), additional dimensions are filled with independent noise, allowing the \(2\text{-}D\) or \(2\text{-}D\)-subspace structures to be naturally embedded in higher-dimensional space while preserving the intended geometric patterns.
 

### 2025-09-20 paper-cardinalR.Rmd
Updated branching section

@@
                           "gen_cubehole"), 
-                      exp = c("Generate a cube with specified grid points along each axes.",
-                              "Generate a cube with uniform points.", 
-                              "Generate a cube with a hole."))
+                      exp = c("Cube with specified grid points along each axes.",
+                              "Cube with uniform points.", 
+                              "Cube with a hole."))
 ```
@@
                                 "gen_cubic"), 
-                        exp = c("Generate a quadratic pattern.",
-                                "Generate a cubic pattern."))
+                        exp = c("Quadratic pattern.",
+                                "Cubic pattern."))
 ```
@@
                             "gen_scurvehole"), 
-                        exp = c("Generate a S-curve.",
-                                "Generate a S-curve with a hole."))
+                        exp = c("S-curve.",
+                                "S-curve with a hole."))
 ```
@@
                             "gen_hemisphere"), 
-                        exp = c("Generate a circle.",
-                                "Generate a curvy cell cycle.",
-                                "Generate a uniform sphere.",
-                                "Generate a grided sphere.",
-                                "Generate multiple small spheres within a big sphere.",
-                                "Generate a hemisphere."))
+                        exp = c("Circle.",
+                                "Curvy cell cycle.",
+                                "Uniform sphere.",
+                                "Grided sphere.",
+                                "Multiple small spheres within a big sphere.",
+                                "Hemisphere."))
 ```
@@
                             "gen_nonlinear"), 
-                        exp = c("Generate a crescent pattern.",
-                                "Generate a curvy cylinder.",
-                                "Generate a spherical spiral.",
-                                "Generate a helical spiral.",
-                                "Generate a conic spiral.",
-                                "Generate a nonlinear hyperbola."))
+                        exp = c("Crescent pattern.",
+                                "Curvy cylinder.",
+                                "Spherical spiral.",
+                                "Helical spiral.",
+                                "Conic spiral.",
+                                "Nonlinear hyperbola."))
 ```
@@
                             "gen_trefoil3d"), 
-                        exp = c("Generate a trefoil in $4\\text{-}D$.",
-                                "Generate a trefoil in $3\\text{-}D$."))
+                        exp = c("Trefoil in $4\\text{-}D$.",
+                                "Trefoil in $3\\text{-}D$."))
 ```
@@
                         exp = c(
-    "Generates a Möbius-like cluster combined with a Gaussian cluster.",
-    "Generates multiple Gaussian clusters in high-dimensional space.",
-    "Generates a curvilinear cluster with a Gaussian cluster.",
-    "Generates K-link circular clusters (non-linear circular patterns).",
-    "Generates chain-like circular clusters connected sequentially.",
-    "Generates K-link curvy cycle clusters (curvilinear loop structures).",
-    "Generates chain-like curvy cycle clusters connected sequentially.",
-    "Generates circular clusters with a Gaussian cluster in the middle.",
-    "Generates curvy circular clusters with a Gaussian cluster in the middle.",
-    "Generates a single grid in two dimensions.",
-    "Generates two overlapping grids.",
-    "Generates two grids shifted relative to each other.",
-    "Generates parallel shaped clusters.",
-    "Generates three clusters with different shapes. (eg:- 01, 02, ..., 20)"))
+    "Möbius-like cluster combined with a Gaussian cluster.",
+    "Multiple Gaussian clusters in high-dimensional space.",
+    "Curvilinear cluster with a Gaussian cluster.",
+    "K-link circular clusters (non-linear circular patterns).",
+    "Chain-like circular clusters connected sequentially.",
+    "K-link curvy cycle clusters (curvilinear loop structures).",
+    "Chain-like curvy cycle clusters connected sequentially.",
+    "Circular clusters with a Gaussian cluster in the middle.",
+    "Curvy circular clusters with a Gaussian cluster in the middle.",
+    "Single grid in two dimensions.",
+    "Two overlapping grids.",
+    "Two grids shifted relative to each other.",
+    "Parallel shaped clusters.",
+    "Three clusters with different shapes. (eg:- 01, 02, ..., 20)"))
 ```

### 2025-09-20 paper-cardinalR.tex
Updated branching section

@@
 \midrule
-gen\_gridcube & Generate a cube with specified grid points along each axes.\\
-gen\_unifcube & Generate a cube with uniform points.\\
-gen\_cubehole & Generate a cube with a hole.\\
+gen\_gridcube & Cube with specified grid points along each axes.\\
+gen\_unifcube & Cube with uniform points.\\
+gen\_cubehole & Cube with a hole.\\
 \bottomrule
@@
 \midrule
-gen\_quadratic & Generate a quadratic pattern.\\
-gen\_cubic & Generate a cubic pattern.\\
+gen\_quadratic & Quadratic pattern.\\
+gen\_cubic & Cubic pattern.\\
 \bottomrule
@@
 \midrule
-gen\_scurve & Generate a S-curve.\\
-gen\_scurvehole & Generate a S-curve with a hole.\\
+gen\_scurve & S-curve.\\
+gen\_scurvehole & S-curve with a hole.\\
 \bottomrule
@@
 \midrule
-gen\_circle & Generate a circle.\\
-gen\_curvycycle & Generate a curvy cell cycle.\\
-gen\_unifsphere & Generate a uniform sphere.\\
-gen\_gridedsphere & Generate a grided sphere.\\
-gen\_clusteredspheres & Generate multiple small spheres within a big sphere.\\
+gen\_circle & Circle.\\
+gen\_curvycycle & Curvy cell cycle.\\
+gen\_unifsphere & Uniform sphere.\\
+gen\_gridedsphere & Grided sphere.\\
+gen\_clusteredspheres & Multiple small spheres within a big sphere.\\
 \addlinespace
-gen\_hemisphere & Generate a hemisphere.\\
+gen\_hemisphere & Hemisphere.\\
 \bottomrule
@@
 \midrule
-gen\_crescent & Generate a crescent pattern.\\
-gen\_curvycylinder & Generate a curvy cylinder.\\
-gen\_sphericalspiral & Generate a spherical spiral.\\
-gen\_helicalspiral & Generate a helical spiral.\\
-gen\_conicspiral & Generate a conic spiral.\\
+gen\_crescent & Crescent pattern.\\
+gen\_curvycylinder & Curvy cylinder.\\
+gen\_sphericalspiral & Spherical spiral.\\
+gen\_helicalspiral & Helical spiral.\\
+gen\_conicspiral & Conic spiral.\\
 \addlinespace
-gen\_nonlinear & Generate a nonlinear hyperbola.\\
+gen\_nonlinear & Nonlinear hyperbola.\\
 \bottomrule
@@
 \midrule
-gen\_trefoil4d & Generate a trefoil in \$4\textbackslash{}text\{-\}D\$.\\
-gen\_trefoil3d & Generate a trefoil in \$3\textbackslash{}text\{-\}D\$.\\
+gen\_trefoil4d & Trefoil in \$4\textbackslash{}text\{-\}D\$.\\
+gen\_trefoil3d & Trefoil in \$3\textbackslash{}text\{-\}D\$.\\
 \bottomrule
@@
 \midrule
-make\_mobiusgau & Generates a Möbius-like cluster combined with a Gaussian cluster.\\
-make\_multigau & Generates multiple Gaussian clusters in high-dimensional space.\\
-make\_curvygau & Generates a curvilinear cluster with a Gaussian cluster.\\
-make\_klink\_circles & Generates K-link circular clusters (non-linear circular patterns).\\
-make\_chain\_circles & Generates chain-like circular clusters connected sequentially.\\
+make\_mobiusgau & Möbius-like cluster combined with a Gaussian cluster.\\
+make\_multigau & Multiple Gaussian clusters in high-dimensional space.\\
+make\_curvygau & Curvilinear cluster with a Gaussian cluster.\\
+make\_klink\_circles & K-link circular clusters (non-linear circular patterns).\\
+make\_chain\_circles & Chain-like circular clusters connected sequentially.\\
 \addlinespace
-make\_klink\_curvycycle & Generates K-link curvy cycle clusters (curvilinear loop structures).\\
-make\_chain\_curvycycle & Generates chain-like curvy cycle clusters connected sequentially.\\
-make\_gaucircles & Generates circular clusters with a Gaussian cluster in the middle.\\
-make\_gaucurvycycle & Generates curvy circular clusters with a Gaussian cluster in the middle.\\
-make\_onegrid & Generates a single grid in two dimensions.\\
+make\_klink\_curvycycle & K-link curvy cycle clusters (curvilinear loop structures).\\
+make\_chain\_curvycycle & Chain-like curvy cycle clusters connected sequentially.\\
+make\_gaucircles & Circular clusters with a Gaussian cluster in the middle.\\
+make\_gaucurvycycle & Curvy circular clusters with a Gaussian cluster in the middle.\\
+make\_onegrid & Single grid in two dimensions.\\
 \addlinespace
-make\_twogrid\_overlap & Generates two overlapping grids.\\
-make\_twogrid\_shift & Generates two grids shifted relative to each other.\\
-make\_shape\_para & Generates parallel shaped clusters.\\
-make\_three\_clust\_ & Generates three clusters with different shapes. (eg:- 01, 02, ..., 20)\\
+make\_twogrid\_overlap & Two overlapping grids.\\
+make\_twogrid\_shift & Two grids shifted relative to each other.\\
+make\_shape\_para & Parallel shaped clusters.\\
+make\_three\_clust\_ & Three clusters with different shapes. (eg:- 01, 02, ..., 20)\\
 \bottomrule

### 2025-09-20 paper-cardinalR.R

@@ -1035,9 +1035,9 @@ cubehole_proj3 <- plot_proj(
 cube_tb <- tibble(fun = c("gen_gridcube",
                           "gen_unifcube",
                           "gen_cubehole"), 
-                      exp = c("Generate a cube with specified grid points along each axes.",
-                              "Generate a cube with uniform points.", 
-                              "Generate a cube with a hole."))
+                      exp = c("Cube with specified grid points along each axes.",
+                              "Cube with uniform points.", 
+                              "Cube with a hole."))
 
 
 ## ----cube-tb-html, eval=knitr::is_html_output()-------------------------------
@@ -1417,8 +1417,8 @@ mobius_proj1 + mobius_proj2 + mobius_proj3 +
 ## -----------------------------------------------------------------------------
 polynomial_tb <- tibble(fun = c("gen_quadratic",
                                 "gen_cubic"), 
-                        exp = c("Generate a quadratic pattern.",
-                                "Generate a cubic pattern."))
+                        exp = c("Quadratic pattern.",
+                                "Cubic pattern."))
 
 
 ## ----polynomial-tb-html, eval=knitr::is_html_output()-------------------------
@@ -2045,8 +2045,8 @@ pyrholes_proj1 + pyrholes_proj2 + pyrholes_proj3 +
 ## -----------------------------------------------------------------------------
 scurve_tb <- tibble(fun = c("gen_scurve",
                             "gen_scurvehole"), 
-                        exp = c("Generate a S-curve.",
-                                "Generate a S-curve with a hole."))
+                        exp = c("S-curve.",
+                                "S-curve with a hole."))
 
 
 ## ----scurve-tb-html, eval=knitr::is_html_output()-----------------------------
@@ -2270,12 +2270,12 @@ sphere_tb <- tibble(fun = c("gen_circle",
                             "gen_gridedsphere",
                             "gen_clusteredspheres",
                             "gen_hemisphere"), 
-                        exp = c("Generate a circle.",
-                                "Generate a curvy cell cycle.",
-                                "Generate a uniform sphere.",
-                                "Generate a grided sphere.",
-                                "Generate multiple small spheres within a big sphere.",
-                                "Generate a hemisphere."))
+                        exp = c("Circle.",
+                                "Curvy cell cycle.",
+                                "Uniform sphere.",
[…]

### 2025-09-20 paper-cardinalR.Rmd
Updated multicluster section

@@
 Users can control the number of clusters (`k`), the number of points in each cluster (`n`), and the dimensionality of the space (`p`) for all the clusters.
-Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube), be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (`loc`).
+Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (`shape`), can be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (`loc`).
 The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures.
 An optional argument, `is_bkg`, adds background noise drawn from a multivariate normal distribution centered on the dataset’s overall mean with standard deviations matching the observed spread.
+Extra arguments (`...`) can be passed to cluster generators, allowing further control over per-cluster characteristics such as noise structure.
 
@@
 If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
-Users can specify how these noise dimensions are generated (e.g., Gaussian, uniform), offering control over the embedding process.
-All shape generators allow the user to define the number of points and dimensions, and most include additional arguments to customize specific characteristics of the structure.
+Users can specify how these noise dimensions are generated (e.g., Gaussian, wavy) (`noise_fun`), offering control over the embedding process.
+All shape generators allow the user to define the number of points (`n`) and dimensions (`p`), and most include additional arguments to customize specific characteristics of the structure.
 

### 2025-09-20 paper-cardinalR.tex
Updated multicluster section

@@
 Users can control the number of clusters (\texttt{k}), the number of points in each cluster (\texttt{n}), and the dimensionality of the space (\texttt{p}) for all the clusters.
-Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube), be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (\texttt{loc}).
+Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (\texttt{shape}), can be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (\texttt{loc}).
 The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures.
 An optional argument, \texttt{is\_bkg}, adds background noise drawn from a multivariate normal distribution centered on the dataset's overall mean with standard deviations matching the observed spread.
+Extra arguments (\texttt{...}) can be passed to cluster generators, allowing further control over per-cluster characteristics such as noise structure.
 
@@
 If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
-Users can specify how these noise dimensions are generated (e.g., Gaussian, uniform), offering control over the embedding process.
-All shape generators allow the user to define the number of points and dimensions, and most include additional arguments to customize specific characteristics of the structure.
+Users can specify how these noise dimensions are generated (e.g., Gaussian, wavy) (\texttt{noise\_fun}), offering control over the embedding process.
+All shape generators allow the user to define the number of points (\texttt{n}) and dimensions (\texttt{p}), and most include additional arguments to customize specific characteristics of the structure.
 

### 2025-09-20 paper-cardinalR.Rmd
Added equations in new line

@@
 Each point is sampled from a parameterization defined by an angle $\theta \sim U(0, 2\pi)$, which specifies the position along the circular loop, and a width $w \sim U(-0.4, 0.4)$, which offsets the point across the strip’s band.
-The cartesian coordinates are $X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = (w/2)\sin(\theta/2)$.
-This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface.
+The cartesian coordinates are $$X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta),  \quad X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta),\text{ and }X_3 = (w/2)\sin(\theta/2).$$ This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface.
 For $p > 3$, additional noise dimension are added to embed the $3\text{-}D$ Mobius into $p\text{-}D$.
@@
 That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
-The first three coordinates are then defined as: $X_1 \sim U(-r_x(z),\ r_x(z))$, $X_2 \sim U(-r_y(z),\ r_y(z))$, and $X_3 \sim U(-r_x(z),\ r_x(z)).$
+The first three coordinates are then defined as: $$X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).$$
 
@@
 This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure: $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$, $X_3 = \text{sign}(\theta)(\cos(\theta) - 1)$.
+The first three dimensions form the S-curve structure: $$X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1)$$.
 This configuration creates a horizontally curled shape in $(X_1, X_3)$, with additional band thickness in the $X_2$ direction.
@@
 Let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
-Coordinates in the first two dimensions represent a perfect circle on the plane: $X_1 = \cos(\theta), \quad X_2 = \sin(\theta)$.
+Coordinates in the first two dimensions represent a perfect circle on the plane: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta)$$.
 For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
@@
 Each observation is computed from spherical coordinates, with $u \sim U(-1, 1)$ representing $\cos(\phi)$ and $\theta \sim U(0, 2\pi)$ the azimuthal angle.
-Cartesian coordinates are then defined as $X_1 = r\sqrt{1 - u^2}\cos(\theta)$, $X_2 = r\sqrt{1 - u^2}\sin(\theta)$, and $X_3 = ru$, ensuring uniform distribution on the surface (not within) of the sphere.
+Cartesian coordinates are then defined as $$X_1 = r\sqrt{1 - u^2}\cos(\theta), \quad X_2 = r\sqrt{1 - u^2}\sin(\theta),\text{ and }X_3 = ru,$$ ensuring uniform distribution on the surface (not within) of the sphere.
 For $p > 3$, additional dimensions $X_4$ through $X_p$ are appended as structured noise.
@@
 Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into $4\text{-}D$ Cartesian space: $X_1 = \sin(\theta_1)\cos(\theta_2)$, $X_2 = \sin(\theta_1)\sin(\theta_2)$, $X_3 = \cos(\theta_1)\cos(\theta_3)$, $X_4 = \cos(\theta_1)\sin(\theta_3)$.
-This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
+The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
 For $p > 4$, additional noise dimensions are added to embed the structure in higher dimensions.
@@
 The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.
-The $X_1 = t \cos(t)$, $X_2 = t \sin(t)$, $X_3 \sim U(w_1, w_2), \text{where } t \sim U(0, 3\pi)$.
-For $p > 3$, the remaining $p - 3$ dimensions are filled with noise to simulate high-dimensional complexity.
+The $$X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \text{ where } t \sim U(0, 3\pi).$$ For $p > 3$, the remaining $p - 3$ dimensions are filled with noise to simulate high-dimensional complexity.
 
@@
 Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
-The corresponding $2\text{-}D$ coordinates are defined by: $X_1 = \cos(\theta)$, $X_2 = \sin(\theta)$.
+The corresponding $2\text{-}D$ coordinates are defined by: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$
 
@@
 A band of thickness around the knot path is controlled by the `steps` argument, while the number of $\theta$ and $\phi$ values is determined by the `steps` and `n` arguments, respectively (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
-The coordinates are defined as $X_1 = \cos(\theta) \cos(\phi)$, $X_2 = \cos(\theta) \sin(\phi)$, $X_3 = \sin(\theta) \cos(1.5 \phi)$, and $X_4 = \sin(\theta) \sin(1.5 \phi)$, where $\theta$ and $\phi$ trace the knot’s path.
+The coordinates are defined as $$X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi)$$, where $\theta$ and $\phi$ trace the knot’s path.
 For $p > 4$, the quadratic structure is embedded into higher dimensions by appending additional noise dimensions.
@@
 
-For the $3\text{-}D$ stereographic projection, `gen_trefoil3d(n, p, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3$ using $X_1' = X_1 / (1 - X_4)$, $X_2' = X_2 / (1 - X_4)$, and $X_3' = X_3 / (1 - X_4)$, excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
+For the $3\text{-}D$ stereographic projection, `gen_trefoil3d(n, p, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $$(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),$$ excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
 As with the $4\text{-}D$ case, optional noise dimensions can be added to embed the knot into higher-dimensional spaces.

### 2025-09-20 paper-cardinalR.tex
Added equations in new line

@@
 Each point is sampled from a parameterization defined by an angle \(\theta \sim U(0, 2\pi)\), which specifies the position along the circular loop, and a width \(w \sim U(-0.4, 0.4)\), which offsets the point across the strip's band.
-The cartesian coordinates are \(X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta)\), \(X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta)\), and \(X_3 = (w/2)\sin(\theta/2)\).
-This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable one-sided surface.
+The cartesian coordinates are \[X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta),  \quad X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta),\text{ and }X_3 = (w/2)\sin(\theta/2).\] This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable one-sided surface.
 For \(p > 3\), additional noise dimension are added to embed the \(3\text{-}D\) Mobius into \(p\text{-}D\).
@@
 That is, \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
-The first three coordinates are then defined as: \(X_1 \sim U(-r_x(z),\ r_x(z))\), \(X_2 \sim U(-r_y(z),\ r_y(z))\), and \(X_3 \sim U(-r_x(z),\ r_x(z)).\)
+The first three coordinates are then defined as: \[X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).\]
 
@@
 This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure: \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\), \(X_3 = \text{sign}(\theta)(\cos(\theta) - 1)\).
+The first three dimensions form the S-curve structure: \[X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1)\].
 This configuration creates a horizontally curled shape in \((X_1, X_3)\), with additional band thickness in the \(X_2\) direction.
@@
 Let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
-Coordinates in the first two dimensions represent a perfect circle on the plane: \(X_1 = \cos(\theta), \quad X_2 = \sin(\theta)\).
+Coordinates in the first two dimensions represent a perfect circle on the plane: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta)\].
 For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
@@
 Each observation is computed from spherical coordinates, with \(u \sim U(-1, 1)\) representing \(\cos(\phi)\) and \(\theta \sim U(0, 2\pi)\) the azimuthal angle.
-Cartesian coordinates are then defined as \(X_1 = r\sqrt{1 - u^2}\cos(\theta)\), \(X_2 = r\sqrt{1 - u^2}\sin(\theta)\), and \(X_3 = ru\), ensuring uniform distribution on the surface (not within) of the sphere.
+Cartesian coordinates are then defined as \[X_1 = r\sqrt{1 - u^2}\cos(\theta), \quad X_2 = r\sqrt{1 - u^2}\sin(\theta),\text{ and }X_3 = ru,\] ensuring uniform distribution on the surface (not within) of the sphere.
 For \(p > 3\), additional dimensions \(X_4\) through \(X_p\) are appended as structured noise.
@@
 Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into \(4\text{-}D\) Cartesian space: \(X_1 = \sin(\theta_1)\cos(\theta_2)\), \(X_2 = \sin(\theta_1)\sin(\theta_2)\), \(X_3 = \cos(\theta_1)\cos(\theta_3)\), \(X_4 = \cos(\theta_1)\sin(\theta_3)\).
-This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
+The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
 For \(p > 4\), additional noise dimensions are added to embed the structure in higher dimensions.
@@
 The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.
-The \(X_1 = t \cos(t)\), \(X_2 = t \sin(t)\), \(X_3 \sim U(w_1, w_2), \text{where } t \sim U(0, 3\pi)\).
-For \(p > 3\), the remaining \(p - 3\) dimensions are filled with noise to simulate high-dimensional complexity.
+The \[X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \text{ where } t \sim U(0, 3\pi).\] For \(p > 3\), the remaining \(p - 3\) dimensions are filled with noise to simulate high-dimensional complexity.
 
@@
 Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
-The corresponding \(2\text{-}D\) coordinates are defined by: \(X_1 = \cos(\theta)\), \(X_2 = \sin(\theta)\).
+The corresponding \(2\text{-}D\) coordinates are defined by: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta).\]
 
@@
 A band of thickness around the knot path is controlled by the \texttt{steps} argument, while the number of \(\theta\) and \(\phi\) values is determined by the \texttt{steps} and \texttt{n} arguments, respectively (Figure \ref{fig:trefoil-proj} a).
-The coordinates are defined as \(X_1 = \cos(\theta) \cos(\phi)\), \(X_2 = \cos(\theta) \sin(\phi)\), \(X_3 = \sin(\theta) \cos(1.5 \phi)\), and \(X_4 = \sin(\theta) \sin(1.5 \phi)\), where \(\theta\) and \(\phi\) trace the knot's path.
+The coordinates are defined as \[X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi)\], where \(\theta\) and \(\phi\) trace the knot's path.
 For \(p > 4\), the quadratic structure is embedded into higher dimensions by appending additional noise dimensions.
@@
 
-For the \(3\text{-}D\) stereographic projection, \texttt{gen\_trefoil3d(n,\ p,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\) using \(X_1' = X_1 / (1 - X_4)\), \(X_2' = X_2 / (1 - X_4)\), and \(X_3' = X_3 / (1 - X_4)\), excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
+For the \(3\text{-}D\) stereographic projection, \texttt{gen\_trefoil3d(n,\ p,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \[(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),\] excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
 As with the \(4\text{-}D\) case, optional noise dimensions can be added to embed the knot into higher-dimensional spaces.

### 2025-09-23 paper-cardinalR.Rmd
Updated this with new org functions
Updated Di's comment on abstract first sentence

@@
 abstract: >
-  A high-dimensional dataset is where each observation is described by many features, or dimensions.
+  A high-dimensional dataset is one where each observation is described by many features, or dimensions, with associations between them.
 Such a dataset might contain various types of structures that have complex geometric properties, such as nonlinear manifolds, clusters, or sparse distributions.
@@
 Each branch is embedded in a unique or repeated $2\text{-}D$ subspace of the $p\text{-}D$ space.
+When `allow_share = TRUE`, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible $\binom{p}{2}$ combinations are exhausted, after which additional branches may repeat subspaces.
 Linear branches follow

### 2025-09-23 paper-cardinalR.tex
Updated this with new org functions
Updated Di's comment on abstract first sentence

@@
 \abstract{%
-A high-dimensional dataset is where each observation is described by many features, or dimensions.
+A high-dimensional dataset is one where each observation is described by many features, or dimensions, with associations between them.
 Such a dataset might contain various types of structures that have complex geometric properties, such as nonlinear manifolds, clusters, or sparse distributions.
@@
 Each branch is embedded in a unique or repeated \(2\text{-}D\) subspace of the \(p\text{-}D\) space.
+When \texttt{allow\_share\ =\ TRUE}, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible \(\binom{p}{2}\) combinations are exhausted, after which additional branches may repeat subspaces.
 Linear branches follow

### 2025-09-23 paper-cardinalR.Rmd
Updated where we can see the characteristics

@@
   A high-dimensional dataset is one where each observation is described by many features, or dimensions, with associations between them.
-Such a dataset might contain various types of structures that have complex geometric properties, such as nonlinear manifolds, clusters, or sparse distributions.
+These datasets contain nonlinear manifolds in image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining.
 We can generate data containing a variety of structures using mathematical functions and statistical distributions.

### 2025-09-23 paper-cardinalR.tex
Updated where we can see the characteristics

@@
 A high-dimensional dataset is one where each observation is described by many features, or dimensions, with associations between them.
-Such a dataset might contain various types of structures that have complex geometric properties, such as nonlinear manifolds, clusters, or sparse distributions.
+These datasets contain nonlinear manifolds in image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining.
 We can generate data containing a variety of structures using mathematical functions and statistical distributions.

### 2025-09-23 paper-cardinalR.Rmd
Remove abstract three sentences
Updated abstract last sentence

@@
 We can generate data containing a variety of structures using mathematical functions and statistical distributions.
-Sampling from a multivariate normal distribution will generate data in an elliptical shape.
-Using a trigonometric function we can generate a spiral.
-A torus function can create a donut shape.
 High-dimensional data structures are useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
 Their controlled complexity allows researchers to understand challenges posed in data analysis and helps to develop robust analytical methods across diverse scientific fields like bioinformatics, machine learning, and forensic science.
-Functions to generate a large variety of structures in high dimensions are organized into the R package `cardinalR`, along with some already generated examples.
+Functions to generate a large variety of structures in high dimensions are organized into the R package `cardinalR`, along with some already generated examples, adding to the existing toolset of benchmark datasets for evaluating algorithms.
 draft: true

### 2025-09-23 paper-cardinalR.tex
Remove abstract three sentences
Updated abstract last sentence

@@
 We can generate data containing a variety of structures using mathematical functions and statistical distributions.
-Sampling from a multivariate normal distribution will generate data in an elliptical shape.
-Using a trigonometric function we can generate a spiral.
-A torus function can create a donut shape.
 High-dimensional data structures are useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
 Their controlled complexity allows researchers to understand challenges posed in data analysis and helps to develop robust analytical methods across diverse scientific fields like bioinformatics, machine learning, and forensic science.
-Functions to generate a large variety of structures in high dimensions are organized into the R package \texttt{cardinalR}, along with some already generated examples.
+Functions to generate a large variety of structures in high dimensions are organized into the R package \texttt{cardinalR}, along with some already generated examples, adding to the existing toolset of benchmark datasets for evaluating algorithms.
 }

### 2025-09-23 paper-cardinalR.Rmd
Updated the word essential

@@
 
-Generating synthetic datasets with clearly defined geometric properties is essential for evaluating and benchmarking algorithms in various fields, such as machine learning, data mining, and computational biology.
+Generating synthetic datasets with clearly defined geometric properties is useful for evaluating and benchmarking algorithms in various fields, such as machine learning, data mining, and computational biology.
 Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.

### 2025-09-23 paper-cardinalR.Rmd
Updated the word "Another useful"

@@
 There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The `geozoo` package (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-Another useful package is `snedata` (@james2025), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
+The package is `snedata` (@james2025), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
 Additionally, `mlbench` (@friedrich2024) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.

### 2025-09-23 paper-cardinalR.tex
Updated the word "Another useful"

@@
 There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-Another useful package is \texttt{snedata} (\citet{james2025}), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
+The package is \texttt{snedata} (\citet{james2025}), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
 Additionally, \texttt{mlbench} (\citet{friedrich2024}) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.

### 2025-09-23 paper-cardinalR.Rmd
Put mlbench into new setence

@@
 The package is `snedata` (@james2025), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
-Additionally, `mlbench` (@friedrich2024) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
-In the field of single-cell omics, `splatter` (@luke2017) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+Additionally, `splatter` (@luke2017) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+In contrast, `mlbench` (@friedrich2024) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+
 While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.

### 2025-09-23 paper-cardinalR.tex
Put mlbench into new setence

@@
 The package is \texttt{snedata} (\citet{james2025}), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
-Additionally, \texttt{mlbench} (\citet{friedrich2024}) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
-In the field of single-cell omics, \texttt{splatter} (\citet{luke2017}) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+Additionally, \texttt{splatter} (\citet{luke2017}) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+In contrast, \texttt{mlbench} (\citet{friedrich2024}) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+
 While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.

### 2025-09-23 paper-cardinalR.Rmd
Combine first two paragraphs of introduction
Did all the comments for introduction

@@
 Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
-
 There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The `geozoo` package (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
@@
 While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.
-
 To address this gap, this paper introduces the `cardinalR` R package.
@@
 `cardinalR` offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets for evaluating algorithm performance in high-dimensional settings.
-
-By providing these aspects, `cardinalR` provides researchers with a method to generate more explainable and challenging high-dimensional clustering synthetic datasets focused to the specific needs of evaluating algorithms.
+By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimensionality reduction (NLDR) methods, clustering algorithms, and visualization techniques.
+Moreover, these datasets can serve as benchmark examples for exploring how different algorithmic choices affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
 

### 2025-09-23 paper-cardinalR.tex
Combine first two paragraphs of introduction
Did all the comments for introduction

@@
 Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
-
 There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
@@
 While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.
-
 To address this gap, this paper introduces the \texttt{cardinalR} R package.
@@
 \texttt{cardinalR} offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets for evaluating algorithm performance in high-dimensional settings.
-
-By providing these aspects, \texttt{cardinalR} provides researchers with a method to generate more explainable and challenging high-dimensional clustering synthetic datasets focused to the specific needs of evaluating algorithms.
+By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimensionality reduction (NLDR) methods, clustering algorithms, and visualization techniques.
+Moreover, these datasets can serve as benchmark examples for exploring how different algorithmic choices affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
 

### 2025-09-23 paper-cardinalR.Rmd
Updated abstract with all the comments
Removed the installation section

@@
 These datasets contain nonlinear manifolds in image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining.
-We can generate data containing a variety of structures using mathematical functions and statistical distributions.
+Data with a variety of structures can be generated using mathematical functions and statistical distributions to create test datasets.
 High-dimensional data structures are useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
@@
 
-## Installation
-
-The development version can be installed from GitHub:
-
-```r
-pak::pak("JayaniLakshika/cardinalR")
-```
+The `cardinalR` R package is available on GitHub at [JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR).
 

### 2025-09-23 paper-cardinalR.Rmd
Remove trefoil summary table

@@
 The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trefoil-tb-html)', '\\@ref(tab:trefoil-tb-pdf)'))` summarizes these functions.
-
-```{r}
-trefoil_tb <- tibble(fun = c("gen_trefoil4d",
-                            "gen_trefoil3d"), 
-                        exp = c("Trefoil in $4\\text{-}D$.",
-                                "Trefoil in $3\\text{-}D$."))
-```
-
-```{r trefoil-tb-html, eval=knitr::is_html_output()}
-trefoil_tb |> 
-  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r trefoil-tb-pdf, eval=knitr::is_latex_output()}
-
-trefoil_tb |> 
-  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
 

### 2025-09-23 paper-cardinalR.tex
Updated abstract with all the comments
Remove trefoil summary table

@@
 
-\subsection{Installation}\label{installation}
-
-The development version can be installed from GitHub:
-
-\begin{verbatim}
-pak::pak("JayaniLakshika/cardinalR")
-\end{verbatim}
+The \texttt{cardinalR} R package is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.
 
@@
 The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
-Table \ref{tab:trefoil-tb-pdf} summarizes these functions.
-
-\begin{table}
-
-\caption{\label{tab:trefoil-tb-pdf}cardinalR trefoil data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_trefoil4d & Trefoil in \$4\textbackslash{}text\{-\}D\$.\\
-gen\_trefoil3d & Trefoil in \$3\textbackslash{}text\{-\}D\$.\\
-\bottomrule
-\end{tabular}
-\end{table}
 

### 2025-09-23 paper-cardinalR.Rmd
Removed all the tables

@@
 
-The main arguments of the `gen_multicluster()` function are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:main-tb-html)', '\\@ref(tab:main-tb-pdf)'))`.
-
-```{r}
-main_tb <- tibble(arg = c("n",
-                          "p",
-                          "k",
-                          "loc",
-                          "scale",
-                          "shape",
-                          "rotation",
-                          "is_bkg"),
-                  type = c("numeric (vector)",
-                           "numeric",
-                           "numeric",
-                           "numeric (matrix)",
-                           "numeric (vector)",
-                           "character (vector)",
-                           "numeric (list)",
-                           "boolean"),
-                        exp = c("Number of points in each cluster.",
-                                "Number of dimensions.",
-                                "Number of clusters.",
-                                "Locations/centroids of clusters.",
-                                "Scaling factors of clusters.",
-                                "Shapes of clusters.",
-                                "Plane and the corresponding angle along that plane for each cluster.",
-                                "Background noise should exist or not."))
-```
-
-```{r main-tb-html, eval=knitr::is_html_output()}
-main_tb |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation")) 
-```
-
-```{r main-tb-pdf, eval=knitr::is_latex_output()}
-main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "4cm") |>
-  column_spec(3, width = "8cm")
-```
-
 ### Shape generators
@@
 Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:branching-tb-html)', '\\@ref(tab:branching-tb-pdf)'))` outlines these functions.
-The main arguments of the functions described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:arg-branching-tb-html)', '\\@ref(tab:arg-branching-tb-pdf)'))`.
-
-```{r}
-branch_tb <- tibble(fun = c("gen_expbranches",
-                             "gen_linearbranches",
-                             "gen_curvybranches", 
-                             "gen_orglinearbranches", 
-                             "gen_orgcurvybranches"), 
-                      exp = c("Exponential shaped branches.",
-                               "Linear shaped branches.", 
-                               "Curvy shaped branches.",
-                               "Linear shaped branches originated in one point.",
-                               "Curvy shaped branches originated in one point."))
-```
-
-```{r branching-tb-html, eval=knitr::is_html_output()}
-branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r branching-tb-pdf, eval=knitr::is_latex_output()}
-branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-```{r}
-arg_branching_tb <- tibble(arg = c("n",
-                          "p",
-                          "k"), 
-                        exp = c("A numeric value representing the number of points.",
-                                "A numeric value representing the number of dimensions.",
-                                "A numeric value representing the number of clusters."))
-```
-
-```{r arg-branching-tb-html, eval=knitr::is_html_output()}
-arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Explanation")) 
-```
-
-```{r arg-branching-tb-pdf, eval=knitr::is_latex_output()}
-arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
 
@@
 These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:cube-tb-html)', '\\@ref(tab:cube-tb-pdf)'))` outlines these functions and their purposes.
 
@@
 
-```
-
-```{r}
-cube_tb <- tibble(fun = c("gen_gridcube",
-                          "gen_unifcube",
-                          "gen_cubehole"), 
-                      exp = c("Cube with specified grid points along each axes.",
-                              "Cube with uniform points.", 
-                              "Cube with a hole."))
-```
-
-```{r cube-tb-html, eval=knitr::is_html_output()}
-cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r cube-tb-pdf, eval=knitr::is_latex_output()}
-cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
 ```
@@
 We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:polynomial-tb-html)', '\\@ref(tab:polynomial-tb-pdf)'))` summarizes these functions and their purposes.
-
-```{r}
-polynomial_tb <- tibble(fun = c("gen_quadratic",
-                                "gen_cubic"), 
-                        exp = c("Quadratic pattern.",
-                                "Cubic pattern."))
-```
-
-```{r polynomial-tb-html, eval=knitr::is_html_output()}
-polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r polynomial-tb-pdf, eval=knitr::is_latex_output()}
-polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
 
@@
 Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:pyramid-tb-html)', '\\@ref(tab:pyramid-tb-pdf)'))` summarizes these functions.
 
@@
 
-```{r}
-pyramid_tb <- tibble(fun = c("gen_pyrrect",
-                             "gen_pyrtri",
-                             "gen_pyrstar",
-                             "gen_pyrfrac"), 
-                        exp = c("Rectangular-base pyramid, with a sharp or blunted apex.",
-          "Triangular-base pyramid, with a sharp or blunted apex.",
-          "Star-shaped base pyramid, with a sharp or blunted apex.",
-          "Pyramid containing triangular pyramid-shaped holes."))
-```
-
-```{r pyramid-tb-html, eval=knitr::is_html_output()}
-pyramid_tb |> 
-  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r pyramid-tb-pdf, eval=knitr::is_latex_output()}
-pyramid_tb |> 
-  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
 ```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$."}
@@
 These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:scurve-tb-html)', '\\@ref(tab:scurve-tb-pdf)'))` summarizes these functions.
-
-```{r}
-scurve_tb <- tibble(fun = c("gen_scurve",
-                            "gen_scurvehole"), 
-                        exp = c("S-curve.",
-                                "S-curve with a hole."))
-```
-
-```{r scurve-tb-html, eval=knitr::is_html_output()}
-scurve_tb |> 
-  kable(caption = "cardinalR S-curve data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r scurve-tb-pdf, eval=knitr::is_latex_output()}
-scurve_tb |> 
-  kable(caption = "cardinalR S-curve data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
 
@@
 For the circle and curvy cycle, this occurs when $p > 2$ and $p > 3$, respectively, while for spheres and hemispheres it occurs when $p > 3$ or $p > 4$.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:sphere-tb-html)', '\\@ref(tab:sphere-tb-pdf)'))` summarizes these functions.
-
-```{r}
-sphere_tb <- tibble(fun = c("gen_circle",
-                            "gen_curvycycle",
-                            "gen_unifsphere",
-                            "gen_gridedsphere",
-                            "gen_clusteredspheres",
-                            "gen_hemisphere"), 
-                        exp = c("Circle.",
-                                "Curvy cell cycle.",
-                                "Uniform sphere.",
-                                "Grided sphere.",
-                                "Multiple small spheres within a big sphere.",
-                                "Hemisphere."))
-```
-
-```{r sphere-tb-html, eval=knitr::is_html_output()}
-sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r sphere-tb-pdf, eval=knitr::is_latex_output()}
-sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
 
@@
 These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trigonometric-tb-html)', '\\@ref(tab:trigonometric-tb-pdf)'))` summarizes these functions.
-
-```{r}
-trigonometric_tb <- tibble(fun = c("gen_crescent",
-                            "gen_curvycylinder",
-                            "gen_sphericalspiral",
-                            "gen_helicalspiral",
-                            "gen_conicspiral",
-                            "gen_nonlinear"), 
-                        exp = c("Crescent pattern.",
-                                "Curvy cylinder.",
-                                "Spherical spiral.",
-                                "Helical spiral.",
-                                "Conic spiral.",
-                                "Nonlinear hyperbola."))
-```
-
-```{r trigonometric-tb-html, eval=knitr::is_html_output()}
-trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r trigonometric-tb-pdf, eval=knitr::is_latex_output()}
-trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
 
@@
 
-```{r}
-noise_fun_tb <- tibble(fun = c("gen_noisedims",
-                            "gen_wavydims1",
-                            "gen_wavydims2",
-                            "gen_wavydims3"), 
-                        exp = c("Gaussian noise dimensions with optional mean and standard deviation.",
-                                "Wavy noise dimensions based on a user-specified theta sequence with added jitter.",
-                                "Wavy noise dimensions using polynomial transformations of an existing dimension vector.",
-                                "Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset."))
-```
-
-```{r noise-tb-html, eval=knitr::is_html_output()}
-noise_fun_tb |> 
-  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r noise-tb-pdf, eval=knitr::is_latex_output()}
-noise_fun_tb |> 
-  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-The `gen_noisedims(n, p, m, s)` function generates \$p\$ independent Gaussian noise dimensions,
+The `gen_noisedims(n, p, m, s)` function generates $p$ independent Gaussian noise dimensions,
 

### 2025-09-23 paper-cardinalR.tex
Removed all the tables

@@
 
-The main arguments of the \texttt{gen\_multicluster()} function are shown in Table \ref{tab:main-tb-pdf}.
-
-\begin{table}
-
-\caption{\label{tab:main-tb-pdf}The main arguments for gen\_multicluster().}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Argument & Type & Explanation\\
-\midrule
-n & numeric (vector) & Number of points in each cluster.\\
-p & numeric & Number of dimensions.\\
-k & numeric & Number of clusters.\\
-loc & numeric (matrix) & Locations/centroids of clusters.\\
-scale & numeric (vector) & Scaling factors of clusters.\\
-\addlinespace
-shape & character (vector) & Shapes of clusters.\\
-rotation & numeric (list) & Plane and the corresponding angle along that plane for each cluster.\\
-is\_bkg & boolean & Background noise should exist or not.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \subsubsection{Shape generators}\label{shape-generators}
@@
 Although these functions can generate multiple branches, they do not produce a formal \emph{multicluster} dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
-Table \ref{tab:branching-tb-pdf} outlines these functions.
-The main arguments of the functions described in Table \ref{tab:arg-branching-tb-pdf}.
-
-\begin{table}
-
-\caption{\label{tab:branching-tb-pdf}cardinalR branching data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_expbranches & Exponential shaped branches.\\
-gen\_linearbranches & Linear shaped branches.\\
-gen\_curvybranches & Curvy shaped branches.\\
-gen\_orglinearbranches & Linear shaped branches originated in one point.\\
-gen\_orgcurvybranches & Curvy shaped branches originated in one point.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:arg-branching-tb-pdf}The main arguments for branching shape generators.}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Argument & Explanation\\
-\midrule
-n & A numeric value representing the number of points.\\
-p & A numeric value representing the number of dimensions.\\
-k & A numeric value representing the number of clusters.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
-Table \ref{tab:cube-tb-pdf} outlines these functions and their purposes.
 
@@
 \end{verbatim}
-
-\begin{table}
-
-\caption{\label{tab:cube-tb-pdf}cardinalR cube data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_gridcube & Cube with specified grid points along each axes.\\
-gen\_unifcube & Cube with uniform points.\\
-gen\_cubehole & Cube with a hole.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-Table \ref{tab:polynomial-tb-pdf} summarizes these functions and their purposes.
-
-\begin{table}
-
-\caption{\label{tab:polynomial-tb-pdf}cardinalR polynomial data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_quadratic & Quadratic pattern.\\
-gen\_cubic & Cubic pattern.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
-Table \ref{tab:pyramid-tb-pdf} summarizes these functions.
 
@@
 
-\begin{table}
-
-\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_pyrrect & Rectangular-base pyramid, with a sharp or blunted apex.\\
-gen\_pyrtri & Triangular-base pyramid, with a sharp or blunted apex.\\
-gen\_pyrstar & Star-shaped base pyramid, with a sharp or blunted apex.\\
-gen\_pyrfrac & Pyramid containing triangular pyramid-shaped holes.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \begin{figure}[!]
@@
 These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
-Table \ref{tab:scurve-tb-pdf} summarizes these functions.
-
-\begin{table}
-
-\caption{\label{tab:scurve-tb-pdf}cardinalR S-curve data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_scurve & S-curve.\\
-gen\_scurvehole & S-curve with a hole.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 For the circle and curvy cycle, this occurs when \(p > 2\) and \(p > 3\), respectively, while for spheres and hemispheres it occurs when \(p > 3\) or \(p > 4\).
-Table \ref{tab:sphere-tb-pdf} summarizes these functions.
-
-\begin{table}
-
-\caption{\label{tab:sphere-tb-pdf}cardinalR sphere data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_circle & Circle.\\
-gen\_curvycycle & Curvy cell cycle.\\
-gen\_unifsphere & Uniform sphere.\\
-gen\_gridedsphere & Grided sphere.\\
-gen\_clusteredspheres & Multiple small spheres within a big sphere.\\
-\addlinespace
-gen\_hemisphere & Hemisphere.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
-Table \ref{tab:trigonometric-tb-pdf} summarizes these functions.
-
-\begin{table}
-
-\caption{\label{tab:trigonometric-tb-pdf}cardinalR trigonometric data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_crescent & Crescent pattern.\\
-gen\_curvycylinder & Curvy cylinder.\\
-gen\_sphericalspiral & Spherical spiral.\\
-gen\_helicalspiral & Helical spiral.\\
-gen\_conicspiral & Conic spiral.\\
-\addlinespace
-gen\_nonlinear & Nonlinear hyperbola.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 
-\begin{table}
-
-\caption{\label{tab:noise-tb-pdf}cardinalR noise dimensions generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_noisedims & Gaussian noise dimensions with optional mean and standard deviation.\\
-gen\_wavydims1 & Wavy noise dimensions based on a user-specified theta sequence with added jitter.\\
-gen\_wavydims2 & Wavy noise dimensions using polynomial transformations of an existing dimension vector.\\
-gen\_wavydims3 & Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-The \texttt{gen\_noisedims(n,\ p,\ m,\ s)} function generates \$p\$ independent Gaussian noise dimensions,
+The \texttt{gen\_noisedims(n,\ p,\ m,\ s)} function generates \(p\) independent Gaussian noise dimensions,
 

### 2025-09-23 paper-cardinalR.R

@@ -87,92 +87,6 @@ theme_set(theme_linedraw() +
 source("scripts/additional_functions.R")
 
 
-## -----------------------------------------------------------------------------
-main_tb <- tibble(arg = c("n",
-                          "p",
-                          "k",
-                          "loc",
-                          "scale",
-                          "shape",
-                          "rotation",
-                          "is_bkg"),
-                  type = c("numeric (vector)",
-                           "numeric",
-                           "numeric",
-                           "numeric (matrix)",
-                           "numeric (vector)",
-                           "character (vector)",
-                           "numeric (list)",
-                           "boolean"),
-                        exp = c("Number of points in each cluster.",
-                                "Number of dimensions.",
-                                "Number of clusters.",
-                                "Locations/centroids of clusters.",
-                                "Scaling factors of clusters.",
-                                "Shapes of clusters.",
-                                "Plane and the corresponding angle along that plane for each cluster.",
-                                "Background noise should exist or not."))
-
-
-## ----main-tb-html, eval=knitr::is_html_output()-------------------------------
-# main_tb |>
-#   kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"))
-
-
-## ----main-tb-pdf, eval=knitr::is_latex_output()-------------------------------
-main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "4cm") |>
-  column_spec(3, width = "8cm")
-
-
-## -----------------------------------------------------------------------------
-branch_tb <- tibble(fun = c("gen_expbranches",
-                             "gen_linearbranches",
-                             "gen_curvybranches", 
-                             "gen_orglinearbranches", 
-                             "gen_orgcurvybranches"), 
-                      exp = c("Exponential shaped branches.",
-                               "Linear shaped branches.", 
[…]

### 2025-09-23 paper-cardinalR.Rmd
Updated Figure width and height

@@
 
-```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=15, fig.height=25}
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=12, fig.height=20}
 
@@
 ```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."}
-#| fig-width: 15
-#| fig-height: 5
+#| fig-width: 12
+#| fig-height: 4
 
@@
 
-```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=15, fig.height=15}
+```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=12, fig.height=12}
 
@@
 ```{r gau-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gau` data."}
-#| fig-width: 15
-#| fig-height: 5
+#| fig-width: 12
+#| fig-height: 4
 
@@
 ```{r linear-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linear` data."}
-#| fig-width: 15
-#| fig-height: 5
+#| fig-width: 12
+#| fig-height: 4
 
@@
 ```{r mobius-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."}
-#| fig-width: 15
-#| fig-height: 5
+#| fig-width: 12
+#| fig-height: 4
 
@@
 
-```{r poly-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.", fig.width=15, fig.height=10}
+```{r poly-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.", fig.width=12, fig.height=8}
 
@@
 
-```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=15, fig.height=20}
+```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=12, fig.height=16}
 
@@
 
-```{r scurve-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.", fig.width=15, fig.height=10}
+```{r scurve-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.", fig.width=12, fig.height=8}
 
@@
 
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=28, out.width="80%", fig.align='center'}
 
@@
 ```{r swissroll-proj, eval=knitr::is_latex_output(), fig.pos="H", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `swissroll` data."}
-#| fig-width: 15
-#| fig-height: 5
+#| fig-width: 12
+#| fig-height: 4
 
@@
 
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=15, fig.height=35, out.width="80%", fig.align='center'}
+```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=28, out.width="80%", fig.align='center'}
 
@@
 
-```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=15, fig.height=10}
+```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=12, fig.height=8}
 
@@
 The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green."}
-#| fig-width: 15
-#| fig-height: 5
+#| fig-width: 12
+#| fig-height: 4
 

### 2025-09-23 paper-cardinalR.R

@@ -552,7 +552,7 @@ orgcurvybranches_proj3 <- plot_proj(
 # branchfig
 
 
-## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=15, fig.height=25----
+## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=12, fig.height=20----
 
 expbranches_proj1 + expbranches_proj2 + expbranches_proj3 +
 linearbranches_proj1 + linearbranches_proj2 + linearbranches_proj3 +
@@ -668,8 +668,8 @@ cone_proj3 <- plot_proj(
 
 
 ## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."----
-#| fig-width: 15
-#| fig-height: 5
+#| fig-width: 12
+#| fig-height: 4
 
 cone_proj1 + cone_proj2 + cone_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
@@ -960,7 +960,7 @@ cubehole_proj3 <- plot_proj(
 # branchfig
 
 
-## ----cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=15, fig.height=15----
+## ----cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=12, fig.height=12----
 
 gridcube_proj1 + gridcube_proj2 + gridcube_proj3 +
[…]

### 2025-09-23 paper-cardinalR.Rmd
Updated HTML and PDF Figures

--- 2025-09-23 paper-cardinalR.Rmd
@@
 ```{r}
-linearbranches_lang <- langevitour::langevitour(linearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+linearbranches_lang <- langevitour::langevitour(linearbranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-curvybranches_lang <- langevitour::langevitour(curvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+curvybranches_lang <- langevitour::langevitour(curvybranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-expbranches_lang <- langevitour::langevitour(expbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+expbranches_lang <- langevitour::langevitour(expbranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-orgcurvybranches_lang <- langevitour::langevitour(orgcurvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+orgcurvybranches_lang <- langevitour::langevitour(orgcurvybranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 
-```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=15, fig.height=15}
+```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=12, fig.height=15}
 
@@
 ```{r}
-gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-linear_lang <- langevitour::langevitour(linear, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+linear_lang <- langevitour::langevitour(linear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-mobius_lang <- langevitour::langevitour(mobius, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+mobius_lang <- langevitour::langevitour(mobius, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-quadratic_lang <- langevitour::langevitour(quadratic, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+quadratic_lang <- langevitour::langevitour(quadratic, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-cubic_lang <- langevitour::langevitour(cubic, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+cubic_lang <- langevitour::langevitour(cubic, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-scurve_lang <- langevitour::langevitour(scurve, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+scurve_lang <- langevitour::langevitour(scurve, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-scurvehole_lang <- langevitour::langevitour(scurvehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+scurvehole_lang <- langevitour::langevitour(scurvehole, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-unifsphere_lang <- langevitour::langevitour(unifsphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+unifsphere_lang <- langevitour::langevitour(unifsphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-gridedsphere_lang <- langevitour::langevitour(gridedsphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+gridedsphere_lang <- langevitour::langevitour(gridedsphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-hemisphere_lang <- langevitour::langevitour(hemisphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+hemisphere_lang <- langevitour::langevitour(hemisphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-crescent_lang <- langevitour::langevitour(crescent, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+crescent_lang <- langevitour::langevitour(crescent, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-helicalspiral_lang <- langevitour::langevitour(helicalspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+helicalspiral_lang <- langevitour::langevitour(helicalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trefoil-tb-html)', '\\@ref(tab:trefoil-tb-pdf)'))` summarizes these functions.
+
+```{r}
+trefoil_tb <- tibble(fun = c("gen_trefoil4d",
+                            "gen_trefoil3d"), 
+                        exp = c("Trefoil in $4\\text{-}D$.",
+                                "Trefoil in $3\\text{-}D$."))
+```
+
+```{r trefoil-tb-html, eval=knitr::is_html_output()}
+trefoil_tb |> 
+  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r trefoil-tb-pdf, eval=knitr::is_latex_output()}
+
+trefoil_tb |> 
+  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
 
@@
 ```{r}
-trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```
@@
 ```{r}
-trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
+trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
 ```

### 2025-09-23 paper-cardinalR.R

@@ -87,12 +87,98 @@ theme_set(theme_linedraw() +
 source("scripts/additional_functions.R")
 
 
+## -----------------------------------------------------------------------------
+main_tb <- tibble(arg = c("n",
+                          "p",
+                          "k",
+                          "loc",
+                          "scale",
+                          "shape",
+                          "rotation",
+                          "is_bkg"),
+                  type = c("numeric (vector)",
+                           "numeric",
+                           "numeric",
+                           "numeric (matrix)",
+                           "numeric (vector)",
+                           "character (vector)",
+                           "numeric (list)",
+                           "boolean"),
+                        exp = c("Number of points in each cluster.",
+                                "Number of dimensions.",
+                                "Number of clusters.",
+                                "Locations/centroids of clusters.",
+                                "Scaling factors of clusters.",
+                                "Shapes of clusters.",
+                                "Plane and the corresponding angle along that plane for each cluster.",
+                                "Background noise should exist or not."))
+
+
+## ----main-tb-html, eval=knitr::is_html_output()-------------------------------
+# main_tb |>
+#   kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"))
+
+
+## ----main-tb-pdf, eval=knitr::is_latex_output()-------------------------------
+main_tb |> 
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "4cm") |>
+  column_spec(3, width = "8cm")
+
+
+## -----------------------------------------------------------------------------
+branch_tb <- tibble(fun = c("gen_expbranches",
+                             "gen_linearbranches",
+                             "gen_curvybranches", 
+                             "gen_orglinearbranches", 
+                             "gen_orgcurvybranches"), 
+                      exp = c("Exponential shaped branches.",
+                               "Linear shaped branches.", 
[…]

### 2025-10-02 paper-cardinalR.Rmd
Added links for residuals data
Start updating the paper from the beginning

--- 2025-09-23 paper-cardinalR.Rmd
@@
 In contrast, `mlbench` (@friedrich2024) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+The `surreal` package (@james2024) implements the "Residual (Sur)Realism" algorithm (@leonard2007) to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
+Meanwhile, the `DHARMa` package (@florian2024) adopts a simulation-based approach to create scaled quantile residuals for generalized linear (mixed) models and related frameworks, supporting model diagnostics through intuitive residuals, plots, and tests for common misspecification issues.
+
+<!--https://cran.r-project.org/web/packages/surreal/index.html-->
+<!--https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html-->
 
@@
 The main function of the package is `gen_multicluster()`, which generates datasets consisting of multiple clusters with user-specified characteristics.
-Users can control the number of clusters (`k`), the number of points in each cluster (`n`), and the dimensionality of the space (`p`) for all the clusters.
-Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (`shape`), can be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (`loc`).
+Users can control the number of clusters (`k`), and the number of points in each cluster (`n`).
+Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (`shape`), can be scaled to adjust its spread, rotated using custom rotation matrices (`rotation`), and positioned at defined centroids (`loc`).
 The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures.
+
+To maintain consistency across generators, the function identifies the arguments required by each chosen generator function and supplies only those arguments that are valid for that specific generator.
+This design enables the combination of cluster types with differing parameter requirements within the same dataset.
+When clusters are generated with fewer dimensions than others, the function augments the lower-dimensional clusters with additional Gaussian noise variables so that all clusters are represented in the same dimensional space.
+These noise dimensions are drawn independently from normal distributions
+
+$$
+X \sim \mathcal{N}(m, s^2),
+$$
+
+where the mean ($m$) is set to the average of the cluster coordinates and the standard deviation ($s$) defaults to $0.2$.
+
 An optional argument, `is_bkg`, adds background noise drawn from a multivariate normal distribution centered on the dataset’s overall mean with standard deviations matching the observed spread.
-Extra arguments (`...`) can be passed to cluster generators, allowing further control over per-cluster characteristics such as noise structure.
+Extra arguments (`...`) can be passed to cluster generators, allowing further control over per-cluster characteristics like radius of the sphere.
+
+The main arguments of the `gen_multicluster()` function are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:main-tb-html)', '\\@ref(tab:main-tb-pdf)'))`.
+
+```{r}
+main_tb <- tibble(arg = c("n",
+                          "k",
+                          "loc",
+                          "scale",
+                          "shape",
+                          "rotation",
+                          "is_bkg"),
+                  type = c("numeric (vector)",
+                           "numeric",
+                           "numeric (matrix)",
+                           "numeric (vector)",
+                           "character (vector)",
+                           "numeric (list)",
+                           "boolean"),
+                        exp = c("Number of points in each cluster.",
+                                "Number of clusters.",
+                                "Locations/centroids of clusters.",
+                                "Scaling factors of clusters.",
+                                "Shapes of clusters.",
+                                "Rotation matrices, one per cluster.",
+                                "Background noise should exist or not."))
+```
+
+```{r main-tb-html, eval=knitr::is_html_output()}
+main_tb |> 
+  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation")) 
+```
+
+```{r main-tb-pdf, eval=knitr::is_latex_output()}
+main_tb |> 
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
+  column_spec(1, width = "2cm") |>
+  column_spec(2, width = "3cm") |>
+  column_spec(3, width = "8cm")
+```
 
@@
 
-#### Branching
-
-A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (@trapnell2014).
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (`n`), and number of branches (`k`).
-Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
-
-The simplest structures are approximately linear branches, generated by the `gen_linearbranches(n, p, k)` function (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) b"} else { "Figure \\@ref(fig:branch-proj) b"})`).
-These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability.
-Mathematically, each branch $i$ is defined as
-
-$$
-X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
-$$
-
-where $(x_{\text{start},i}, y_{\text{start},i})$ is the starting point of branch $i$, $\delta$ controls local jitter, and $s_i$ is the slope, initialized as
-
-$$
-s_i =
-\begin{cases}
-0.5 & i = 1, \\
--0.5 & i = 2, \\
-\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
-\end{cases}
-$$
-
-Branches $1$ and $2$ are initialized with fixed slopes and intercepts, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
-
-```{r data-linearbranches, echo=TRUE}
-linearbranches <- gen_linearbranches(n = 1000, p = 4, k = 4)
-```
-
-```{r}
-linearbranches_lang <- langevitour::langevitour(linearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r linearbranches-proj1}
-scaled_data <- scale_data_manual(linearbranches)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-linearbranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linearbranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-linearbranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linearbranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-linearbranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.65), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-To introduce curvature, the `gen_curvybranches(n, p, k)` function generates $k$ curvilinear branches in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) c"} else { "Figure \\@ref(fig:branch-proj) c"})`).
-Branches $1$ and $2$ are simple parabolas defined as
-
-$$
-\begin{aligned}
-\text{Branch 1: } & X_1 \sim U(0,1), \quad X_2 = 0.1 X_1 + X_1^2 + \epsilon, \\
-\text{Branch 2: } & X_1 \sim U(-1,0), \quad X_2 = 0.1 X_1 - 2 X_1^2 + \epsilon, \quad \epsilon \sim U(0, \delta),
-\end{aligned}
-$$
-
-where $\delta$ controls local jitter.
-Additional branches are attached iteratively to existing structures.
-Each new branch $i$ starts at a selected point $(x_{\text{start},i}, y_{\text{start},i})$ from the current structure and extends according to
-
-$$
-X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
-$$
-
-where $s_i$ is a scale factor controlling the curvature of branch $i$.
-For the first few initial branches, $s_i$ can be fixed (e.g., $s_1 = 1, s_2 = 2$), while for subsequent branches it is sampled from a predefined set, such as $s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}$, to create variability in curvature.
-
-```{r data-curvybranches, echo=TRUE}
-curvybranches <- gen_curvybranches(n = 1000, p = 4, k = 4)
-```
-
-```{r}
-curvybranches_lang <- langevitour::langevitour(curvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r curvybranches-proj1}
-scaled_data <- scale_data_manual(curvybranches)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.022))
-
-curvybranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.7), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvybranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.02))
-
-curvybranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.3), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvybranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-curvybranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.4), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-The `gen_expbranches(n, p, k)` function creates $k$ exponential branches in $2\text{-}D$, radiating from a central region (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) a"} else { "Figure \\@ref(fig:branch-proj) a"})`).
-Each branch $i$ is defined as
-
-$$
-X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
-$$
-
-where $\sigma_i = (-1)^{i+1}$ alternates the sign of the exponent to produce mirror-symmetric branches.
-The parameter $s_i$ controls the steepness of branch $i$, and $\delta$ introduces small local jitter.
-
-```{r data-expbranches, echo=TRUE}
-expbranches <- gen_expbranches(n = 1000, p = 4, k = 4)
-```
-
-```{r}
-expbranches_lang <- langevitour::langevitour(expbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r expbranches-proj1}
-scaled_data <- scale_data_manual(expbranches)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.022))
-
-expbranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.35, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r expbranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.2,
-                                              axis_scaled = 1.7, 
-                                              axis_pos_x = -0.16, 
-                                              axis_pos_y = -0.16, 
-                                              threshold = 0.02))
-
-expbranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.2, 0.2), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r expbranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.2,
-                                              axis_scaled = 1.3, 
-                                              axis_pos_x = -0.16, 
-                                              axis_pos_y = -0.16, 
-                                              threshold = 0.016))
-
-expbranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.2, 0.2), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` and `gen_orgcurvybranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) d-e"} else { "Figure \\@ref(fig:branch-proj) d-e"})`).
-Each branch is embedded in a unique or repeated $2\text{-}D$ subspace of the $p\text{-}D$ space.
-When `allow_share = TRUE`, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible $\binom{p}{2}$ combinations are exhausted, after which additional branches may repeat subspaces.
-Linear branches follow
-
-$$
-X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = s_i X_{i_1} + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
-$$
-
-while curvilinear branches include a quadratic term
-
-$$
-X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = - s_i X_{i_1}^2 + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
-$$
-
-where $a_i, b_i$ define the range of the first coordinate for branch $i$, and $\epsilon$ is Gaussian noise added to introduce variability.
-The scale factor $s_i$ controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first $\binom{p}{2}$ branches, $s_i = 1$; for additional branches when $k > \binom{p}{2}$, $s_i$ is randomly drawn from the set $\{1, 1.5, 2, \dots, 8\}$.
-
-```{r data-orglinearbranches, echo=TRUE}
-orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
-```
-
-```{r}
-orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r orglinearbranches-proj1}
-scaled_data <- scale_data_manual(orglinearbranches)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-orglinearbranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r orglinearbranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-orglinearbranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r orglinearbranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-orglinearbranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r data-orgcurvybranches, echo=TRUE}
-orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
-```
-
-```{r}
-orgcurvybranches_lang <- langevitour::langevitour(orgcurvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r orgcurvybranches-proj1}
-scaled_data <- scale_data_manual(orgcurvybranches)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
-                                              threshold = 0.022))
-
-orgcurvybranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.4, 0.5), 
-  title = "e1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r orgcurvybranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-orgcurvybranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "e2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r orgcurvybranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-orgcurvybranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
-  title = "e3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For all branch-generating functions, when the target dimensionality $p > 2$, additional dimensions are filled with independent noise, allowing the $2\text{-}D$ or $2\text{-}D$-subspace structures to be naturally embedded in higher-dimensional space while preserving the intended geometric patterns.
-
-```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, and `orgcurvybranches` data in $4\\text{-}D$."}
-
-branchfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    expbranches_lang, linearbranches_lang, curvybranches_lang, orglinearbranches_lang, orgcurvybranches_lang
-  ),
-  device = "xs"
-)
-
-class(branchfig) <- c(class(branchfig), "htmlwidget")
-
-branchfig
-```
-
-```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=12, fig.height=20}
-
-expbranches_proj1 + expbranches_proj2 + expbranches_proj3 +
-linearbranches_proj1 + linearbranches_proj2 + linearbranches_proj3 +
-curvybranches_proj1 + curvybranches_proj2 + curvybranches_proj3 +
-orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
-orgcurvybranches_proj1 + orgcurvybranches_proj2 + orgcurvybranches_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Cone
-
-To simulate a cone-shaped structure in arbitrary dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cone)"} else { "Figure \\@ref(fig:cone-proj)"})`), we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
-
-This function generates $n$ points in $p\text{-}D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip.
-Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, capped at the cone height $h$, producing a higher density of points near the tip.
-At each height $X_p$, the radius of the cross-section decreases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio}$ and $r_{\text{max}} = 1$.
-
-For each point, a direction in the first $p-1$ dimensions is sampled uniformly on a $(p-1)$-dimensional hypersphere using generalized spherical coordinates.
-The radial coordinates are scaled by the height-dependent radius $r$, producing the conical taper.
-In three dimensions ($p = 3$), this results in a classical $3\text{-}D$ cone, while for $p > 3$, additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
-
-```{r data-bluntedcone, echo=TRUE}
-cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
-```
-
-```{r}
-cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r cone-proj1}
-scaled_data <- scale_data_manual(cone)
-
-## First projection
-projection <- cbind(
-  c(0.22417,-0.02943,-0.02870,0.08418),
-  c(-0.02715,0.18019,0.05053,0.15254))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.25,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.022))
-
-cone_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.25), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cone-proj2}
-
-## Second projection
-projection <- cbind(
-    c(0.15591,0.17907,0.03198,0.04496),
-    c(-0.15376,0.10435,0.15766,0.00547))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.1,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.08, 
-                                              axis_pos_y = -0.08, 
-                                              threshold = 0.01))
-
-cone_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.1, 0.12), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cone-proj3}
-
-## Third projection
-projection <- cbind(
-    c(-0.13548,0.13832,0.05991,-0.13544),
-    c(0.08479,-0.10719,-0.01572,-0.20123))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.25,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.016))
-
-cone_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.2), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $4\\text{-}D$."}
-
-conefig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    cone_lang
-  ),
-  device = "xs"
-)
-
-class(conefig) <- c(class(conefig), "htmlwidget")
-
-conefig
-```
-
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-cone_proj1 + cone_proj2 + cone_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Cube
-
-A cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube)"} else { "Figure \\@ref(fig:cube-proj)"})`) represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
-We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
-These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
-
-The first is the regular grid of points of $n$ points in $p$ dimensions.
-This is generated using `gen_gridcube(n, p)`.
-The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to $n$.
-Each dimension is then normalized to lie in the interval $[0,1]$, so that the resulting structure forms a true $p\text{-}D$ hypercube.
-This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
-
-```{r data-gridcube, echo=TRUE}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-```
-
-```{r}
-gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r gridcube-proj1}
-scaled_data <- scale_data_manual(gridcube)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.7, 
-                                              axis_pos_y = -0.7, 
-                                              threshold = 0.022))
-
-gridcube_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.83, 0.8), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridcube-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.8,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.85, 
-                                              axis_pos_y = -0.85, 
-                                              threshold = 0.05))
-
-gridcube_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 1), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridcube-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.7, 
-                                              axis_pos_y = -0.7, 
-                                              threshold = 0.016))
-
-gridcube_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
-The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) b"} else { "Figure \\@ref(fig:cube-proj) b"})`).
-
-```{r data-unifcube, echo=TRUE}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-```
-
-```{r}
-unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r unifcube-proj1}
-scaled_data <- scale_data_manual(unifcube)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.022))
-
-unifcube_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.8), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifcube-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.04))
-
-unifcube_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.8), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifcube-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.016))
-
-unifcube_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Finally, we consider a cube that has a central spherical hole.
-This is generated using the `gen_cubehole(n, p, r_hole)` function.
-The cube is generated as per the uniformly distributed cube, but points inside sphere of radius (`r_hole`) are removed, resulting in a hollow cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) c"} else { "Figure \\@ref(fig:cube-proj) c"})`).
-
-```{r data-cubehole, echo=TRUE}
-cubehole <- gen_cubehole(n = 3000, p = 4, r_hole = 0.5)
-```
-
-```{r}
-cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r cubehole-proj1}
-scaled_data <- scale_data_manual(cubehole)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.022))
-
-cubehole_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.75), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cubehole-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.73, 
-                                              axis_pos_y = -0.73, 
-                                              threshold = 0.04))
-
-cubehole_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.85, 0.7), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cubehole-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.016))
-
-cubehole_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.8), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cube, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gridcube`, `unifcube`, and `cubehole` data in $4\\text{-}D$."}
-
-branchfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    gridcube_lang, unifcube_lang, cubehole_lang
-  ),
-  device = "xs"
-)
-
-class(branchfig) <- c(class(branchfig), "htmlwidget")
-
-branchfig
-```
-
-```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=12, fig.height=12}
-
-gridcube_proj1 + gridcube_proj2 + gridcube_proj3 +
-unifcube_proj1 + unifcube_proj2 + unifcube_proj3 +
-cubehole_proj1 + cubehole_proj2 + cubehole_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Gaussian
-
-The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:gau)"} else { "Figure \\@ref(fig:gau-proj)"})`).
-Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
-
-```{r data-gau, echo=TRUE}
-gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
-```
-
-```{r}
-gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r gau-proj1}
-scaled_data <- scale_data_manual(gau)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-gau_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-gau_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-gau_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gau` data in $4\\text{-}D$."}
-
-gaufig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    gau_lang
-  ),
-  device = "xs"
-)
-
-class(gaufig) <- c(class(gaufig), "htmlwidget")
-
-gaufig
-```
-
-```{r gau-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gau` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-gau_proj1 + gau_proj2 + gau_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Linear
-
-The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise.
-Each variable is formed as $X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i$, where $\text{scale}\_i \sim U(-10, 10)$ determines the orientation of the line in each dimension, $\text{shift}\_i \sim U(-300, 300)$ offsets the line to separate dimensions, and $\epsilon \sim N(0, (0.03n)^2)$ introduces Gaussian noise.
-
-```{r data-linear, echo=TRUE}
-linear <- gen_longlinear(n = 1000, p = 4)
-```
-
-```{r}
-linear_lang <- langevitour::langevitour(linear, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r linear-proj1}
-scaled_data <- scale_data_manual(linear)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.022))
-
-linear_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linear-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.75,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.01))
-
-linear_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linear-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.9,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.02))
-
-linear_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linear, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `linear` data in $4\\text{-}D$."}
-
-linearfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    linear_lang
-  ),
-  device = "xs"
-)
-
-class(linearfig) <- c(class(linearfig), "htmlwidget")
-
-linearfig
-```
-
-```{r linear-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linear` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-linear_proj1 + linear_proj2 + linear_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Mobius
-
-The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$ structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:mobius)"} else { "Figure \\@ref(fig:mobius-proj)"})`).
-This classical non-orientable surface loops back on itself with a half-twist.
-
-The Mobius strip is generated using the `geozoo::mobius` function with `p = 3`, and `n` defined as above.
-Each point is sampled from a parameterization defined by an angle $\theta \sim U(0, 2\pi)$, which specifies the position along the circular loop, and a width $w \sim U(-0.4, 0.4)$, which offsets the point across the strip’s band.
-The cartesian coordinates are $$X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta),  \quad X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta),\text{ and }X_3 = (w/2)\sin(\theta/2).$$ This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface.
-For $p > 3$, additional noise dimension are added to embed the $3\text{-}D$ Mobius into $p\text{-}D$.
-
-```{r data-mobius, echo=TRUE}
-mobius <- gen_mobius(n = 1000, p = 4)
-```
-
-```{r}
-mobius_lang <- langevitour::langevitour(mobius, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r mobius-proj1}
-scaled_data <- scale_data_manual(mobius)
-
-## First projection
-projection <- cbind(
-    c(0.53506,0.00763,0.00097,-0.10901),
-    c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-mobius_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r mobius-proj2}
-
-## Second projection
-projection <- cbind(
-    c(-0.06614,0.15134,-0.44278,0.27367),
-    c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-mobius_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r mobius-proj3}
-
-## Third projection
-projection <- cbind(
-    c(0.10704,0.04822,-0.03416,0.53224),
-    c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-mobius_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r mobius, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `mobius` data in $4\\text{-}D$."}
-
-mobiusfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    mobius_lang
-  ),
-  device = "xs"
-)
-
-class(mobiusfig) <- c(class(mobiusfig), "htmlwidget")
-
-mobiusfig
-```
-
-```{r mobius-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `mobius` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-mobius_proj1 + mobius_proj2 + mobius_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Polynomial
-
-A polynomial structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly)"} else { "Figure \\@ref(fig:poly-proj)"})`) generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in high-dimensional space.
-These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
-We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-
-The first is the quadratic curve of $n$ points in $p$ dimensions.
-This is generated using `gen_quadratic(n, p, range)`.
-The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree 2 is applied to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) a"} else { "Figure \\@ref(fig:poly-proj) a"})`).
-
-```{r data-quadratic, echo=TRUE}
-quadratic <- gen_quadratic(n = 1000, p = 4)
-```
-
-```{r}
-quadratic_lang <- langevitour::langevitour(quadratic, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r quadratic-proj1}
-scaled_data <- scale_data_manual(quadratic)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.7, 
-                                              axis_pos_y = -0.7, 
-                                              threshold = 0.022))
-
-quadratic_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.83, 0.65), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r quadratic-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-quadratic_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r quadratic-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-quadratic_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-The second is the cubic curve of $n$ points in $p$ dimensions.
-This is generated using `gen_cubic(n, p, range)`.
-The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:poly) b"} else { "Figure \\@ref(fig:poly-proj) b"})`).
-
-```{r data-cubic, echo=TRUE}
-cubic <- gen_cubic(n = 1000, p = 4)
-```
-
-```{r}
-cubic_lang <- langevitour::langevitour(cubic, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r cubic-proj1}
-scaled_data <- scale_data_manual(cubic)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-cubic_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cubic-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.7, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.05))
-
-cubic_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cubic-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-cubic_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.62), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r poly, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `quadratic` and `cubic` data in $4\\text{-}D$."}
-
-polyfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    quadratic_lang, cubic_lang
-  ),
-  device = "xs"
-)
-
-class(polyfig) <- c(class(polyfig), "htmlwidget")
-
-polyfig
-```
-
-```{r poly-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.", fig.width=12, fig.height=8}
-
-quadratic_proj1 + quadratic_proj2 + quadratic_proj3 +
-cubic_proj1 + cubic_proj2 + cubic_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-For all polynomial shapes, if $p > 2$, the structure is embedded into higher dimensions by appending additional noise dimensions.
-
-#### Pyramid
-
-A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
-The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
-Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
-
-Let $X_1, \dots, X_p$ denote the coordinates of the generated points.
-For the rectangular and triangular based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
-That is, $$X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).$$ This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
-For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
-That is, $X_p = z \sim U(0, h)$.
-
-The remaining dimensions are based on the specific pyramid shape.
-For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`), let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$.
-That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
-The first three coordinates are then defined as: $$X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).$$
-
-```{r data-pyrrect, echo=TRUE}
-pyrrect <- gen_pyrrect(n = 1000, p = 4)
-```
-
-```{r}
-pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r pyrrect-proj1}
-scaled_data <- scale_data_manual(pyrrect)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrrect_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrrect-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.7, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.08))
-
-pyrrect_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrrect-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-pyrrect_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.75), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the triangular based pyramid, `gen_pyrtri(n, p, h, l, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`), let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$.
-That is, $r(z) = r_t + (l-r_t)z/h$.
-A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
-The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v.$
-
-```{r data-pyrtri, echo=TRUE}
-pyrtri <- gen_pyrtri(n = 1000, p = 4)
-```
-
-```{r}
-pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r pyrtri-proj1}
-scaled_data <- scale_data_manual(pyrtri)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.022))
-
-pyrtri_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.3), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrtri-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.03))
-
-pyrtri_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.4), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrtri-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-pyrtri_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the star based pyramid, `gen_pyrstar(n, p, h, rb)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`), let the radius at height $z$, $r(z)$, be such that the radius scales linearly from zero (tip) to the base radius $r_b$.
-That is, $r(z) = r_b\left(1 - z/h\right)$.
-
-Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
-$r_{\text{point}} \sim \sqrt{U(0, 1)}$. Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
-
-```{r data-pyrstar, echo=TRUE}
-pyrstar <- gen_pyrstar(n = 1000, p = 4)
-```
-
-```{r}
-pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r pyrstar-proj1}
-scaled_data <- scale_data_manual(pyrstar)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrstar_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrstar-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-pyrstar_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.4), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrstar-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-pyrstar_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For all the above pyramid shapes, if $p > 3$, the remaining $p - 3$ dimensions (i.e., $X_4$ to $X_{p-1}$) are additional noise.
-
-Finally, for the Sierpinski-like pyramid, `gen_pyrfrac(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`), let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
-The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
-Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
-At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
-This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
-The points remain bounded inside the convex hull of the simplex.
-The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
-
-```{r data-pyrholes, echo=TRUE}
-pyrholes <- gen_pyrfrac(n = 1000, p = 4)
-```
-
-```{r}
-pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r pyrholes-proj1}
-scaled_data <- scale_data_manual(pyrholes)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrholes_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrholes-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-pyrholes_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.75), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrholes-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-pyrholes_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$."}
-
-pyrfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    pyrrect_lang, pyrtri_lang, pyrstar_lang, pyrholes_lang
-  ),
-  device = "xs"
-)
-
-class(pyrfig) <- c(class(pyrfig), "htmlwidget")
-
-pyrfig
-```
-
-```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=12, fig.height=16}
-
-pyrrect_proj1 + pyrrect_proj2 + pyrrect_proj3 +
-pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
-pyrstar_proj1 + pyrstar_proj2 + pyrstar_proj3 +
-pyrholes_proj1 + pyrholes_proj2 + pyrholes_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### S-curve
-
-An S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) simulates data that lies along a smooth, non-linear manifold.
-The functions generate both the standard S-curve shape and, a S-curve variant with structured hole that introduce missing or incomplete region.
-These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
-
-For the S-curve structure, `gen_scurve(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`), the $3\text{-}D$ geometry is constructed by introducing a latent parameter, $\theta \sim U\left(-3\pi/2, 3\pi/2\right)$.
-This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure: $$X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1)$$.
-This configuration creates a horizontally curled shape in $(X_1, X_3)$, with additional band thickness in the $X_2$ direction.
-For $p > 3$, additional noise dimensions are appended to embed the structure in higher dimensions.
-
-```{r data-scurve, echo=TRUE}
-scurve <- gen_scurve(n = 1000, p = 4)
-```
-
-```{r}
-scurve_lang <- langevitour::langevitour(scurve, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r scurve-proj1}
-scaled_data <- scale_data_manual(scurve)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-scurve_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r scurve-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-scurve_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r scurve-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-scurve_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-To introduce missing or incomplete regions on the manifold, `gen_scurvehole(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) b"} else { "Figure \\@ref(fig:scurve-proj) b"})`) removes points in a localized region centered around the middle vertical section of the S-curve.
-Following a similar approach as `gen_cubehole()`, all observations within a fixed radius ($\sqrt{0.3}$) of the anchor point are excluded, creating a hole in the manifold while preserving the overall S-curve structure.
-
-```{r data-scurvehole, echo=TRUE}
-scurvehole <- gen_scurvehole(n = 1000, p = 4)
-```
-
-```{r}
-scurvehole_lang <- langevitour::langevitour(scurvehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r scurvehole-proj1}
-scaled_data <- scale_data_manual(scurvehole)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-scurvehole_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r scurvehole-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-scurvehole_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r scurvehole-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-scurvehole_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r scurve, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `scurve` and `scurvehole` data in $4\\text{-}D$."}
-
-scurvefig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    scurve_lang, scurvehole_lang
-  ),
-  device = "xs"
-)
-
-class(scurvefig) <- c(class(scurvefig), "htmlwidget")
-
-scurvefig
-```
-
-```{r scurve-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.", fig.width=12, fig.height=8}
-
-scurve_proj1 + scurve_proj2 + scurve_proj3 + 
-  scurvehole_proj1 + scurvehole_proj2 + scurvehole_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Sphere
-
-Sphere-shaped structures (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere)"} else { "Figure \\@ref(fig:sphere-proj)"})`) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
-The functions generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
-The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
-For the circle and curvy cycle, this occurs when $p > 2$ and $p > 3$, respectively, while for spheres and hemispheres it occurs when $p > 3$ or $p > 4$.
-
-The simplest case, `gen_circle(n, p)` creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
-Let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
-Coordinates in the first two dimensions represent a perfect circle on the plane: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta)$$.
-For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
-The first component is a scaling factor that decreases with the dimension index, defined as $\text{scale}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$.
-The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2)\pi/2p$.
-Each additional dimension is computed as: $X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p$.
-
-```{r data-circle, echo=TRUE}
-circle <- gen_circle(n = 1000, p = 4)
-```
-
-```{r}
-circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r circle-proj1}
-scaled_data <- scale_data_manual(circle)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-circle_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.8), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r circle-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-circle_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r circle-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-circle_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.53), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the one-dimensional nonlinear cycle embedded in $p\text{-}D$ space, `gen_curvycycle(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`), let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
-The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle".
-In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin.
-Additionally, $X_3 = 1/3\cos(3\theta)$ introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
-
-Together, these define a periodic, non-trivial, closed curve in $3\text{-}D$ with internal folds that produce a more complex geometry than a standard circle or ellipse.
-For dimensions $X_4$ through $X_p$, additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
-The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from $4$ to $p$, which means that the amplitude decreases as the dimension increases.
-Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j\sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2)\pi/2p$.
-
-```{r data-curvycycle, echo=TRUE}
-curvycycle <- gen_curvycycle(n = 1000, p = 4)
-```
-
-```{r}
-curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r curvycycle-proj1}
-scaled_data <- scale_data_manual(curvycycle)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-curvycycle_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycycle-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.04))
-
-curvycycle_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycycle-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-curvycycle_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Building on simple circular structures, the `gen_unifsphere(n, p, r)` function function extends the idea to three dimensions by generating $n$ observations approximately uniformly distributed on the surface of a sphere of radius $r$, with optional high-dimensional noise when $p > 3$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) c"} else { "Figure \\@ref(fig:sphere-proj) c"})`).
-Each observation is computed from spherical coordinates, with $u \sim U(-1, 1)$ representing $\cos(\phi)$ and $\theta \sim U(0, 2\pi)$ the azimuthal angle.
-Cartesian coordinates are then defined as $$X_1 = r\sqrt{1 - u^2}\cos(\theta), \quad X_2 = r\sqrt{1 - u^2}\sin(\theta),\text{ and }X_3 = ru,$$ ensuring uniform distribution on the surface (not within) of the sphere.
-For $p > 3$, additional dimensions $X_4$ through $X_p$ are appended as structured noise.
-
-```{r data-unifsphere, echo=TRUE}
-unifsphere <- gen_unifsphere(n = 1000, p = 4)
-```
-
-```{r}
-unifsphere_lang <- langevitour::langevitour(unifsphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r unifsphere-proj1}
-scaled_data <- scale_data_manual(unifsphere)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-unifsphere_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifsphere-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-unifsphere_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifsphere-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-unifsphere_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.65), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-In addition, the `gen_gridedsphere(n, p)` function constructs a $p\text{-}D$ dataset consisting of approximately $n$ points that are evenly distributed on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
-The method relies on forming a regular grid in spherical coordinates, parameterized by $(p-1)$ angular variables: for dimensions $j = 1, \dots, p-2$ the polar angles are drawn from $[0, \pi]$, while the final angle ($j = p-1$) represents the azimuth and is drawn from $[0, 2\pi]$.
-The number of grid steps along each angular dimension is chosen by decomposing $n$ into $(p-1)$ approximately equal integer factors using the helper function `gen_nproduct(n, p - 1)`.
-
-Each grid point is subsequently mapped into Cartesian space via the standard hyperspherical-to-Cartesian transformation,
-
-$$
-\begin{aligned}
-X_1 &= \cos(\theta_1), \\
-X_2 &= \sin(\theta_1)\cos(\theta_2), \\
-X_3 &= \sin(\theta_1)\sin(\theta_2)\cos(\theta_3), \\
-&\;\;\vdots \\
-X_{p-1} &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\cos(\theta_{p-1}), \\
-X_p &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\sin(\theta_{p-1}).
-\end{aligned}
-$$
-
-The result is a deterministic grid of points lying exactly on the surface of the unit $(p-1)$-sphere, without any additional noise dimensions.
-
-```{r data-gridedsphere, echo=TRUE}
-gridedsphere <- gen_gridedsphere(n = 1000, p = 4)
-```
-
-```{r}
-gridedsphere_lang <- langevitour::langevitour(gridedsphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r gridedsphere-proj1}
-scaled_data <- scale_data_manual(gridedsphere)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-gridedsphere_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridedsphere-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-gridedsphere_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridedsphere-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-gridedsphere_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.62), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For more heterogeneous structures, the `gen_clusteredspheres(n, k, p, r, loc)` function generates one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
-A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
-The sampling is executed using the function `gen_unifsphere(n_1, p, r_1)`, which generates the desired points in the specified dimensional space.
-In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
-These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
-Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for $3\text{-}D$).
-Each observation is classified by cluster, with labels such as "big" for the large central sphere and "small_1" to "small_k" for the smaller spheres.
-
-```{r data-clusteredspheres, echo=TRUE}
-clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, p = 4, r = c(15, 3),
-                                         loc = 10 / sqrt(3)) |>
-  dplyr::select(-cluster)
-```
-
-```{r}
-clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r clusteredspheres-proj1}
-scaled_data <- scale_data_manual(clusteredspheres)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-clusteredspheres_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "e1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r clusteredspheres-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-clusteredspheres_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "e2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r clusteredspheres-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-clusteredspheres_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.62), 
-  title = "e3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Finally, the `gen_hemisphere(n, p)` function restricts sampling to a hemisphere of a $4\text{-}D$ sphere (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) f"} else { "Figure \\@ref(fig:sphere-proj) f"})`).
-Using spherical coordinates, the azimuthal angle $\theta_1 \sim U(0, \pi)$ in the $(x_1, x_2)$ plane, while the elevation angle $\theta_2 \sim U(0, \pi)$ in the $(x_2, x_3)$ plane.
-Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
-For $p > 4$, additional noise dimensions are added to embed the structure in higher dimensions.
-
-```{r data-hemisphere, echo=TRUE}
-hemisphere <- gen_hemisphere(n = 1000, p = 4)
-```
-
-```{r}
-hemisphere_lang <- langevitour::langevitour(hemisphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r hemisphere-proj1}
-scaled_data <- scale_data_manual(hemisphere)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-hemisphere_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "f1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r hemisphere-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-hemisphere_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "f2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r hemisphere-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-hemisphere_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.62), 
-  title = "f3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, `unifsphere`, `gridedsphere`, `clusteredspheres`, and `hemisphere` data in $4\\text{-}D$."}
-
-spherefig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    circle_lang, curvycycle_lang, unifsphere_lang, gridedsphere_lang, clusteredspheres_lang, hemisphere_lang
-  ),
-  device = "xs"
-)
-
-class(spherefig) <- c(class(spherefig), "htmlwidget")
-
-spherefig
-```
-
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=28, out.width="80%", fig.align='center'}
-
-circle_proj1 + circle_proj2 + circle_proj3 +
-curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3 +
-unifsphere_proj1 + unifsphere_proj2 + unifsphere_proj3 +
-gridedsphere_proj1 + gridedsphere_proj2 + gridedsphere_proj3 +
-clusteredspheres_proj1 + clusteredspheres_proj2 + clusteredspheres_proj3 +
-hemisphere_proj1 + hemisphere_proj2 + hemisphere_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Swiss Roll
-
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, p, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:swissroll)"} else { "Figure \\@ref(fig:swissroll-proj)"})`).
-The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.
-The $$X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \text{ where } t \sim U(0, 3\pi).$$ For $p > 3$, the remaining $p - 3$ dimensions are filled with noise to simulate high-dimensional complexity.
-
-```{r data-swissroll, echo=TRUE}
-swissroll <- gen_swissroll(n = 1000, p = 4, w = c(-1, 1))
-```
-
-```{r}
-swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r swissroll-proj1}
-scaled_data <- scale_data_manual(swissroll)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-swissroll_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r swissroll-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-swissroll_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r swissroll-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-swissroll_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r swissroll, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `swissroll` data in $4\\text{-}D$."}
-
-swissrollfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    swissroll_lang
-  ),
-  device = "xs"
-)
-
-class(swissrollfig) <- c(class(swissrollfig), "htmlwidget")
-
-swissrollfig
-```
-
-```{r swissroll-proj, eval=knitr::is_latex_output(), fig.pos="H", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `swissroll` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-swissroll_proj1 + swissroll_proj2 + swissroll_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Trigonometric
-
-Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric)"} else { "Figure \\@ref(fig:triginometric-proj)"})`).
-The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when $p > 2$ (adding $X_3$ through $X_p$), and for all other trigonometric shapes when $p > 4$ (adding $X_5$ through $X_p$).
-These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
-
-First, the `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) a"} else { "Figure \\@ref(fig:triginometric-proj) a"})`).
-Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
-The corresponding $2\text{-}D$ coordinates are defined by: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$
-
-```{r data-crescent, echo=TRUE}
-crescent <- gen_crescent(n = 1000, p = 4)
-```
-
-```{r}
-crescent_lang <- langevitour::langevitour(crescent, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r crescent-proj1}
-scaled_data <- scale_data_manual(crescent)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.15, 
-                                              axis_pos_y = -0.15, 
-                                              threshold = 0.022))
-
-crescent_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.2, 0.3), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r crescent-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.02))
-
-crescent_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.2), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r crescent-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.75,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.03))
-
-crescent_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.7), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Second, the `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
-The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
-Let $\theta \sim U(0, 3\pi)$ represent a random angle on a circular base and $z \sim U(0, h)$ represent the height along the cylinder.
-The coordinates are defined as: $X_1 = \cos(\theta)$ (Circular base, x-axis), $X_2 = \sin(\theta)$ (Circular base, y-axis), $X_3 = z$ (Linear height), and $X_4 = \sin(z)$ (Nonlinear curvy variation along height).
-
-```{r data-curvycylinder, echo=TRUE}
-curvycylinder <- gen_curvycylinder(n = 1000, p = 4, h = 10)
-```
-
-```{r}
-curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r curvycylinder-proj1}
-scaled_data <- scale_data_manual(curvycylinder)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-curvycylinder_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-curvycylinder_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-curvycylinder_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$ be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
-Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
-This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
-
-```{r data-sphericalspiral, echo=TRUE}
-sphericalspiral <- gen_sphericalspiral(n = 1000, p = 4, spins = 1)
-```
-
-```{r}
-sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r sphericalspiral-proj1}
-scaled_data <- scale_data_manual(sphericalspiral)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-sphericalspiral_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphericalspiral-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-sphericalspiral_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphericalspiral-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-sphericalspiral_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For a helical spiral in four dimensions, `gen_helicalspiral(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) d"} else { "Figure \\@ref(fig:triginometric-proj) d"})`), let $\theta \in [0, 5\pi/4]$ be a sequence of angles controlling rotation around a circle.
-Cartesian coordinates; $X_1 = \cos(\theta)$: circular trajectory along the x-axis, $X_2 = \sin(\theta)$: circular trajectory along the y-axis, $X_3 = 0.05\theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix, and $X_4 = 0.1\sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
-
-```{r data-helicalspiral, echo=TRUE}
-helicalspiral <- gen_helicalspiral(n = 1000, p = 4)
-```
-
-```{r}
-helicalspiral_lang <- langevitour::langevitour(helicalspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r helicalspiral-proj1}
-scaled_data <- scale_data_manual(helicalspiral)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-helicalspiral_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r helicalspiral-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.7, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.05))
-
-helicalspiral_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.5), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r helicalspiral-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-helicalspiral_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.67), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Similarly, the `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) e"} else { "Figure \\@ref(fig:triginometric-proj) e"})`).
-The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a $3\text{-}D$ conic helix.
-The shape is defined by parameter $\theta \in [0, 2\pi\text{spins}]$, controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta\cos(\theta)$ for radial expansion in x, and $X_2 = \theta\sin(\theta)$ for radial expansion in y.
-The growth pattern resembles a cone, with the height increasing according to $X_3 = 2\theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta\sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
-
-```{r data-conicspiral, echo=TRUE}
-conicspiral <- gen_conicspiral(n = 1000, p = 4, spins = 1)
-```
-
-```{r}
-conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r conicspiral-proj1}
-scaled_data <- scale_data_manual(conicspiral)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-conicspiral_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "e1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r conicspiral-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.01))
-
-conicspiral_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.5), 
-  title = "e2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r conicspiral-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.016))
-
-conicspiral_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.62), 
-  title = "e3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Finally, the `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
-The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-
-```{r data-nonlinear, echo=TRUE}
-nonlinear <- gen_nonlinear(n = 1000, p = 4, hc = 1, non_fac = 0.5)
-```
-
-```{r}
-nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r nonlinear-proj1}
-scaled_data <- scale_data_manual(nonlinear)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
-                                              threshold = 0.022))
-
-nonlinear_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.45, 0.73), 
-  title = "f1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r nonlinear-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.45,
-                                              axis_scaled = 1.3, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.01))
-
-nonlinear_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.46), 
-  title = "f2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r nonlinear-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.45,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.016))
-
-nonlinear_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.35, 0.4), 
-  title = "f3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r triginometric, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of tthe `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data in $4\\text{-}D$."}
-
-triginometricfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    crescent_lang, curvycylinder_lang, sphericalspiral_lang, helicalspiral_lang, conicspiral_lang, nonlinear_lang
-  ),
-  device = "xs"
-)
-
-class(triginometricfig) <- c(class(triginometricfig), "htmlwidget")
-
-triginometricfig
-```
-
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=28, out.width="80%", fig.align='center'}
-
-crescent_proj1 + crescent_proj2 + crescent_proj3 +
-curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
-sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +
-helicalspiral_proj1 + helicalspiral_proj2 + helicalspiral_proj3 +
-conicspiral_proj1 + conicspiral_proj2 + conicspiral_proj3 +
-nonlinear_proj1 + nonlinear_proj2 + nonlinear_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Trefoil knots
-
-<!--https://laustep.github.io/stlahblog/posts/TorusKnot4D.html-->
-
-The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil)"} else { "Figure \\@ref(fig:trefoil-proj)"})`).
-The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
-
-For the $4\text{-}D$ trefoil knot, the function `gen_trefoil4d(n, p, steps)` generates the structure on the $3$-sphere ($S^3 \subset \mathbb{R}^4$) using two angular parameters, $\theta$ and $\phi$.
-A band of thickness around the knot path is controlled by the `steps` argument, while the number of $\theta$ and $\phi$ values is determined by the `steps` and `n` arguments, respectively (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
-The coordinates are defined as $$X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi)$$, where $\theta$ and $\phi$ trace the knot’s path.
-For $p > 4$, the quadratic structure is embedded into higher dimensions by appending additional noise dimensions.
-
-```{r data-trefoil4d, echo=TRUE}
-trefoil4d <- gen_trefoil4d(n = 500, p = 4, steps = 5)
-```
-
-```{r}
-trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r trefoil4d-proj1}
-scaled_data <- scale_data_manual(trefoil4d)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-trefoil4d_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil4d-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-trefoil4d_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil4d-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-trefoil4d_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the $3\text{-}D$ stereographic projection, `gen_trefoil3d(n, p, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $$(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),$$ excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
-As with the $4\text{-}D$ case, optional noise dimensions can be added to embed the knot into higher-dimensional spaces.
-
-```{r data-trefoil3d, echo=TRUE}
-trefoil3d <- gen_trefoil3d(n = 500, p = 4, steps = 5)
-```
-
-```{r}
-trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
-```
-
-```{r trefoil3d-proj1}
-scaled_data <- scale_data_manual(trefoil4d)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-trefoil3d_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil3d-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-trefoil3d_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil3d-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-trefoil3d_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` and `trefoil3d` data in $4\\text{-}D$."}
-
-trefoilfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    trefoil4d_lang, trefoil3d_lang
-  ),
-  device = "xs"
-)
-
-class(trefoilfig) <- c(class(trefoilfig), "htmlwidget")
-
-trefoilfig
-```
-
-```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=12, fig.height=8}
-
-trefoil4d_proj1 + trefoil4d_proj2 + trefoil4d_proj3 + 
-  trefoil3d_proj1 + trefoil3d_proj2 + trefoil3d_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
+### Wrappers
 
@@
 
+```{r}
+noise_fun_tb <- tibble(fun = c("gen_noisedims",
+                            "gen_wavydims1",
+                            "gen_wavydims2",
+                            "gen_wavydims3"), 
+                        exp = c("Gaussian noise dimensions with optional mean and standard deviation.",
+                                "Wavy noise dimensions based on a user-specified theta sequence with added jitter.",
+                                "Wavy noise dimensions using polynomial transformations of an existing dimension vector.",
+                                "Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset."))
+```
+
+```{r noise-tb-html, eval=knitr::is_html_output()}
+noise_fun_tb |> 
+  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r noise-tb-pdf, eval=knitr::is_latex_output()}
+noise_fun_tb |> 
+  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 The `gen_noisedims(n, p, m, s)` function generates $p$ independent Gaussian noise dimensions,
@@
 
-For scenarios where noise should follow a smooth wavy pattern, `gen_wavydims1(n, p, \theta)` generates dimensions as
+For scenarios where noise should follow a smooth wavy pattern, `gen_wavydims1(n, p, theta)` generates dimensions as
 
@@
 
-Finally, `gen_wavydims3(n, p, \text{data})` generates noise for datasets with multiple correlated dimensions.
+Finally, `gen_wavydims3(n, p, data)` generates noise for datasets with multiple correlated dimensions.
 The first three dimensions are small perturbations of the original coordinates $(X_1, X_2, X_3)$, while higher dimensions are constructed via non-linear combinations, including polynomial and trigonometric transformations, e.g.,
@@
                         exp = c(
-    "Möbius-like cluster combined with a Gaussian cluster.",
+    "Möbius-like cluster combined with a Gaussian.",
     "Multiple Gaussian clusters in high-dimensional space.",
@@
   kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
+  column_spec(1, width = "3.5cm") |>
+  column_spec(2, width = "8.5cm")
 ```
@@
 The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
+Each cluster was chosen to reflect distinct geometric and topological properties, allowing us to test how well DR methods preserve both local and global structures.
+The *helical spiral* cluster is designed to evaluate methods on elongated, twisting structures that challenge linear embeddings such as PCA and require preservation of curvilinear continuity.
+The *hemisphere* provides a curved surface with partial coverage of a $3\text{-}D$ manifold, useful for testing neighborhood preservation and unfolding in algorithms like UMAP and tSNE.
+The *uniform cube* represents isotropic, uniformly distributed data and serves as a control cluster with simple geometric structure to assess baseline embedding fidelity.
+The *cone* introduces variable density along one axis, mimicking structures where point density changes with geometry, helping evaluate how well algorithms maintain relative distances in non-uniform distributions.
+Finally, the *Gaussian* cluster is a standard multivariate normal distribution, included to assess algorithm performance on simple, spherical, high-density clusters.
+Together, these clusters create a challenging synthetic dataset suitable for benchmarking and exploring the strengths and weaknesses of different dimensionality reduction techniques.
 
@@
 ## To generate data
-five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), p = 4, k = 5,
+five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                        loc = positions,
@@
 ```
+
+The five clusters have different geometric structures and each contain different number of points.
+Specifically, the helical spiral cluster includes $2250$ points and was generated with a scale parameter of $0.4$.
+The hemisphere cluster consists of $1500$ points with a scale parameter of $0.35$.
+The uniform cube-shaped cluster contains $750$ points and uses a scale parameter of $0.3$.
+The blunted cone cluster includes $1250$ points, generated with a scale parameter of $1$.
+Finally, the Gaussian-shaped cluster contains $1750$ points and was generated with a scale parameter of $0.3$.
 
@@
 
-The five clusters have different geometric structures and each contain different number of points.
-Specifically, the helical spiral cluster includes $2250$ points and was generated with a scale parameter of $0.4$.
-The hemisphere cluster consists of $1500$ points with a scale parameter of $0.35$.
-The uniform cube-shaped cluster contains $750$ points and uses a scale parameter of $0.3$.
-The blunted cone cluster includes $1250$ points, generated with a scale parameter of $1$.
-Finally, the Gaussian-shaped cluster contains $1750$ points and was generated with a scale parameter of $0.3$.
-
 ```{r layouts}
@@
 
-```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
-
-nldr1 + nldr2 + nldr3 +
-  nldr4 + nldr5 + nldr6 +
-  plot_layout(ncol = 3, guides = "collect")
-```
-
 UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure (Figure \@ref(fig:fig-nldr-layouts)).
@@
 
+```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
+
+nldr1 + nldr2 + nldr3 +
+  nldr4 + nldr5 + nldr6 +
+  plot_layout(ncol = 3, guides = "collect")
+```
+
 # Conclusion
@@
 The `cardinalR` package introduces a flexible framework for generating high-dimensional data structures with well-defined geometric properties.
-It addresses an important need in the evaluation of clustering, machine learning, and dimensionality reduction (DR) methods by enabling the construction of customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
+It addresses an important need in the evaluation of clustering, machine learning, and DR methods by enabling the construction of customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
 In this way, `cardinalR` complements existing packages such as `geozoo`, `snedata`, and `mlbench`, while extending the scope to higher dimensions and more complex shapes.
+
+The motivation for developing this package originated from the need to design a perception–misperception experiment, aimed at investigating how well NLDR methods preserve inter-cluster structure.
+To conduct this study, we required simulated datasets with carefully controlled geometric and clustering properties.
+While some existing packages provided useful starting points, none fully supported the creation of flexible, high-dimensional data with the specific structural variations needed for our experiment.
+Developing these generators for research purposes gradually led to the design of `cardinalR` as a general-purpose package, so that other researchers can benefit from the same tools for simulation, benchmarking, and teaching.
 
@@
 
-The source material for this paper is available at [https://github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
-
+The source material for this paper is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
+
+This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.
 These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].
-This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.

### 2025-10-02 paper-cardinalR.tex
Updated HTML and PDF Figures
Start updating the paper from the beginning

@@
 In contrast, \texttt{mlbench} (\citet{friedrich2024}) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+The \texttt{surreal} package (\citet{james2024}) implements the ``Residual (Sur)Realism'' algorithm (\citet{leonard2007}) to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
+Meanwhile, the \texttt{DHARMa} package (\citet{florian2024}) adopts a simulation-based approach to create scaled quantile residuals for generalized linear (mixed) models and related frameworks, supporting model diagnostics through intuitive residuals, plots, and tests for common misspecification issues.
 
@@
 The main function of the package is \texttt{gen\_multicluster()}, which generates datasets consisting of multiple clusters with user-specified characteristics.
-Users can control the number of clusters (\texttt{k}), the number of points in each cluster (\texttt{n}), and the dimensionality of the space (\texttt{p}) for all the clusters.
-Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (\texttt{shape}), can be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (\texttt{loc}).
+Users can control the number of clusters (\texttt{k}), and the number of points in each cluster (\texttt{n}).
+Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (\texttt{shape}), can be scaled to adjust its spread, rotated using custom rotation matrices (\texttt{rotation}), and positioned at defined centroids (\texttt{loc}).
 The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures.
+
+To maintain consistency across generators, the function identifies the arguments required by each chosen generator function and supplies only those arguments that are valid for that specific generator.
+This design enables the combination of cluster types with differing parameter requirements within the same dataset.
+When clusters are generated with fewer dimensions than others, the function augments the lower-dimensional clusters with additional Gaussian noise variables so that all clusters are represented in the same dimensional space.
+These noise dimensions are drawn independently from normal distributions
+
+\[
+X \sim \mathcal{N}(m, s^2),
+\]
+
+where the mean (\(m\)) is set to the average of the cluster coordinates and the standard deviation (\(s\)) defaults to \(0.2\).
+
 An optional argument, \texttt{is\_bkg}, adds background noise drawn from a multivariate normal distribution centered on the dataset's overall mean with standard deviations matching the observed spread.
-Extra arguments (\texttt{...}) can be passed to cluster generators, allowing further control over per-cluster characteristics such as noise structure.
+Extra arguments (\texttt{...}) can be passed to cluster generators, allowing further control over per-cluster characteristics like radius of the sphere.
+
+The main arguments of the \texttt{gen\_multicluster()} function are shown in Table \ref{tab:main-tb-pdf}.
+
+\begin{table}
+
+\caption{\label{tab:main-tb-pdf}The main arguments for gen\_multicluster().}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Argument & Type & Explanation\\
+\midrule
+n & numeric (vector) & Number of points in each cluster.\\
+k & numeric & Number of clusters.\\
+loc & numeric (matrix) & Locations/centroids of clusters.\\
+scale & numeric (vector) & Scaling factors of clusters.\\
+shape & character (vector) & Shapes of clusters.\\
+\addlinespace
+rotation & numeric (list) & Rotation matrices, one per cluster.\\
+is\_bkg & boolean & Background noise should exist or not.\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-\paragraph{Branching}\label{branching}
-
-A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (\citet{trapnell2014}).
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (\texttt{n}), and number of branches (\texttt{k}).
-Although these functions can generate multiple branches, they do not produce a formal \emph{multicluster} dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
-
-The simplest structures are approximately linear branches, generated by the \texttt{gen\_linearbranches(n,\ p,\ k)} function (Figure \ref{fig:branch-proj} b).
-These consist of \(k\) short line segments in the first two dimensions, with added jitter to simulate variability.
-Mathematically, each branch \(i\) is defined as
-
-\[
-X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
-\]
-
-where \((x_{\text{start},i}, y_{\text{start},i})\) is the starting point of branch \(i\), \(\delta\) controls local jitter, and \(s_i\) is the slope, initialized as
-
-\[
-s_i =
-\begin{cases}
-0.5 & i = 1, \\
--0.5 & i = 2, \\
-\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
-\end{cases}
-\]
-
-Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
-
-\begin{verbatim}
-linearbranches <- gen_linearbranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
-
-To introduce curvature, the \texttt{gen\_curvybranches(n,\ p,\ k)} function generates \(k\) curvilinear branches in \(p\text{-}D\) (Figure \ref{fig:branch-proj} c).
-Branches \(1\) and \(2\) are simple parabolas defined as
-
-\[
-\begin{aligned}
-\text{Branch 1: } & X_1 \sim U(0,1), \quad X_2 = 0.1 X_1 + X_1^2 + \epsilon, \\
-\text{Branch 2: } & X_1 \sim U(-1,0), \quad X_2 = 0.1 X_1 - 2 X_1^2 + \epsilon, \quad \epsilon \sim U(0, \delta),
-\end{aligned}
-\]
-
-where \(\delta\) controls local jitter.
-Additional branches are attached iteratively to existing structures.
-Each new branch \(i\) starts at a selected point \((x_{\text{start},i}, y_{\text{start},i})\) from the current structure and extends according to
-
-\[
-X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
-\]
-
-where \(s_i\) is a scale factor controlling the curvature of branch \(i\).
-For the first few initial branches, \(s_i\) can be fixed (e.g., \(s_1 = 1, s_2 = 2\)), while for subsequent branches it is sampled from a predefined set, such as \(s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}\), to create variability in curvature.
-
-\begin{verbatim}
-curvybranches <- gen_curvybranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
-
-The \texttt{gen\_expbranches(n,\ p,\ k)} function creates \(k\) exponential branches in \(2\text{-}D\), radiating from a central region (Figure \ref{fig:branch-proj} a).
-Each branch \(i\) is defined as
-
-\[
-X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
-\]
-
-where \(\sigma_i = (-1)^{i+1}\) alternates the sign of the exponent to produce mirror-symmetric branches.
-The parameter \(s_i\) controls the steepness of branch \(i\), and \(\delta\) introduces small local jitter.
-
-\begin{verbatim}
-expbranches <- gen_expbranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
-
-High-dimensional generalizations are provided by \texttt{gen\_orglinearbranches(n,\ p,\ k)} and \texttt{gen\_orgcurvybranches(n,\ p,\ k)} (Figure \ref{fig:branch-proj} d-e).
-Each branch is embedded in a unique or repeated \(2\text{-}D\) subspace of the \(p\text{-}D\) space.
-When \texttt{allow\_share\ =\ TRUE}, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible \(\binom{p}{2}\) combinations are exhausted, after which additional branches may repeat subspaces.
-Linear branches follow
-
-\[
-X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = s_i X_{i_1} + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
-\]
-
-while curvilinear branches include a quadratic term
-
-\[
-X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = - s_i X_{i_1}^2 + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
-\]
-
-where \(a_i, b_i\) define the range of the first coordinate for branch \(i\), and \(\epsilon\) is Gaussian noise added to introduce variability.
-The scale factor \(s_i\) controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first \(\binom{p}{2}\) branches, \(s_i = 1\); for additional branches when \(k > \binom{p}{2}\), \(s_i\) is randomly drawn from the set \(\{1, 1.5, 2, \dots, 8\}\).
-
-\begin{verbatim}
-orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
-
-\begin{verbatim}
-orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
-
-For all branch-generating functions, when the target dimensionality \(p > 2\), additional dimensions are filled with independent noise, allowing the \(2\text{-}D\) or \(2\text{-}D\)-subspace structures to be naturally embedded in higher-dimensional space while preserving the intended geometric patterns.
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.}\label{fig:branch-proj}
-\end{figure}
-
-\paragraph{Cone}\label{cone}
-
-To simulate a cone-shaped structure in arbitrary dimensions (Figure \ref{fig:cone-proj}), we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
-
-This function generates \(n\) points in \(p\text{-}D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip.
-Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), capped at the cone height \(h\), producing a higher density of points near the tip.
-At each height \(X_p\), the radius of the cross-section decreases linearly from base to tip according to \(r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h\), where \(r_{\text{min}} = \text{ratio}\) and \(r_{\text{max}} = 1\).
-
-For each point, a direction in the first \(p-1\) dimensions is sampled uniformly on a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates.
-The radial coordinates are scaled by the height-dependent radius \(r\), producing the conical taper.
-In three dimensions (\(p = 3\)), this results in a classical \(3\text{-}D\) cone, while for \(p > 3\), additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
-
-\begin{verbatim}
-cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:cone-proj}
-\end{figure}
-
-\paragraph{Cube}\label{cube}
-
-A cube structure (Figure \ref{fig:cube-proj}) represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
-We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
-These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
-
-The first is the regular grid of points of \(n\) points in \(p\) dimensions.
-This is generated using \texttt{gen\_gridcube(n,\ p)}.
-The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to \(n\).
-Each dimension is then normalized to lie in the interval \([0,1]\), so that the resulting structure forms a true \(p\text{-}D\) hypercube.
-This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
-
-\begin{verbatim}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-\end{verbatim}
-
-An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
-The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube (Figure \ref{fig:cube-proj} b).
-
-\begin{verbatim}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-\end{verbatim}
-
-Finally, we consider a cube that has a central spherical hole.
-This is generated using the \texttt{gen\_cubehole(n,\ p,\ r\_hole)} function.
-The cube is generated as per the uniformly distributed cube, but points inside sphere of radius (\texttt{r\_hole}) are removed, resulting in a hollow cube structure (Figure \ref{fig:cube-proj} c).
-
-\begin{verbatim}
-cubehole <- gen_cubehole(n = 3000, p = 4, r_hole = 0.5)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.}\label{fig:cube-proj}
-\end{figure}
-
-\paragraph{Gaussian}\label{gaussian}
-
-The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure (Figure \ref{fig:gau-proj}).
-Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, s)\), where \(s\) is a user-defined \(p \times p\) positive-definite matrix.
-
-\begin{verbatim}
-gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:gau-proj}
-\end{figure}
-
-\paragraph{Linear}\label{linear}
-
-The \texttt{gen\_longlinear(n,\ p)} function generates a high-dimensional dataset representing a long linear structure with noise.
-Each variable is formed as \(X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i\), where \(\text{scale}\_i \sim U(-10, 10)\) determines the orientation of the line in each dimension, \(\text{shift}\_i \sim U(-300, 300)\) offsets the line to separate dimensions, and \(\epsilon \sim N(0, (0.03n)^2)\) introduces Gaussian noise.
-
-\begin{verbatim}
-linear <- gen_longlinear(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/linear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linear` data.}\label{fig:linear-proj}
-\end{figure}
-
-\paragraph{Mobius}\label{mobius}
-
-The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\) structure (Figure \ref{fig:mobius-proj}).
-This classical non-orientable surface loops back on itself with a half-twist.
-
-The Mobius strip is generated using the \texttt{geozoo::mobius} function with \texttt{p\ =\ 3}, and \texttt{n} defined as above.
-Each point is sampled from a parameterization defined by an angle \(\theta \sim U(0, 2\pi)\), which specifies the position along the circular loop, and a width \(w \sim U(-0.4, 0.4)\), which offsets the point across the strip's band.
-The cartesian coordinates are \[X_1 = \left(1 + (w/2) \cos(\theta/2)\right) \cos(\theta),  \quad X_2 = \left(1 + (w/2) \cos(\theta/2)\right) \sin(\theta),\text{ and }X_3 = (w/2)\sin(\theta/2).\] This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable one-sided surface.
-For \(p > 3\), additional noise dimension are added to embed the \(3\text{-}D\) Mobius into \(p\text{-}D\).
-
-\begin{verbatim}
-mobius <- gen_mobius(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/mobius-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `mobius` data.}\label{fig:mobius-proj}
-\end{figure}
-
-\paragraph{Polynomial}\label{polynomial}
-
-A polynomial structure (Figure \ref{fig:poly-proj}) generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in high-dimensional space.
-These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
-We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-
-The first is the quadratic curve of \(n\) points in \(p\) dimensions.
-This is generated using \texttt{gen\_quadratic(n,\ p,\ range)}.
-The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree 2 is applied to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term (Figure \ref{fig:poly-proj} a).
-
-\begin{verbatim}
-quadratic <- gen_quadratic(n = 1000, p = 4)
-\end{verbatim}
-
-The second is the cubic curve of \(n\) points in \(p\) dimensions.
-This is generated using \texttt{gen\_cubic(n,\ p,\ range)}.
-The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree \(3\) is applied to construct \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points (Figure \ref{fig:poly-proj} b).
-
-\begin{verbatim}
-cubic <- gen_cubic(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/poly-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `quadratic` (a1-a3) and `cubic` (b1-b3) data.}\label{fig:poly-proj}
-\end{figure}
-
-For all polynomial shapes, if \(p > 2\), the structure is embedded into higher dimensions by appending additional noise dimensions.
-
-\paragraph{Pyramid}\label{pyramid}
-
-A pyramid structure (Figure \ref{fig:pyr-proj}) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
-The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
-Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
-
-Let \(X_1, \dots, X_p\) denote the coordinates of the generated points.
-For the rectangular and triangular based pyramid generator functions, the final dimension, \(X_p\), encodes the height of each point and is drawn from an exponential distribution capped at the maximum height \(h\).
-That is, \[X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).\] This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
-For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
-That is, \(X_p = z \sim U(0, h)\).
-
-The remaining dimensions are based on the specific pyramid shape.
-For the rectangular based pyramid, \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} (Figure \ref{fig:pyr-proj} a), let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\).
-That is, \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
-The first three coordinates are then defined as: \[X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).\]
-
-\begin{verbatim}
-pyrrect <- gen_pyrrect(n = 1000, p = 4)
-\end{verbatim}
-
-For the triangular based pyramid, \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} (Figure \ref{fig:pyr-proj} b), let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\).
-That is, \(r(z) = r_t + (l-r_t)z/h\).
-A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section: \(u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v\).
-The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z)u\), and \(X_3 = r(z)v.\)
-
-\begin{verbatim}
-pyrtri <- gen_pyrtri(n = 1000, p = 4)
-\end{verbatim}
-
-For the star based pyramid, \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} (Figure \ref{fig:pyr-proj} c), let the radius at height \(z\), \(r(z)\), be such that the radius scales linearly from zero (tip) to the base radius \(r_b\).
-That is, \(r(z) = r_b\left(1 - z/h\right)\).
-
-Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
-\(r_{\text{point}} \sim \sqrt{U(0, 1)}\). Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
-
-\begin{verbatim}
-pyrstar <- gen_pyrstar(n = 1000, p = 4)
-\end{verbatim}
-
-For all the above pyramid shapes, if \(p > 3\), the remaining \(p - 3\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are additional noise.
-
-Finally, for the Sierpinski-like pyramid, \texttt{gen\_pyrfrac(n,\ p)} (Figure \ref{fig:pyr-proj} d), let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
-The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution: \(T_0 \sim U(0, 1)^p\).
-Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\text{-}D\) simplex.
-At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\): \(T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}\).
-This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
-The points remain bounded inside the convex hull of the simplex.
-The final output is a \(n \times p\) matrix where each row represents a point: \(X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}\).
-
-\begin{verbatim}
-pyrholes <- gen_pyrfrac(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.}\label{fig:pyr-proj}
-\end{figure}
-
-\paragraph{S-curve}\label{s-curve}
-
-An S-curve structure (Figure \ref{fig:scurve-proj}) simulates data that lies along a smooth, non-linear manifold.
-The functions generate both the standard S-curve shape and, a S-curve variant with structured hole that introduce missing or incomplete region.
-These variations are useful for evaluating how well algorithms capture non-linear geometry and handle incomplete manifolds in high-dimensional data.
-
-For the S-curve structure, \texttt{gen\_scurve(n,\ p)} (Figure \ref{fig:scurve-proj} a), the \(3\text{-}D\) geometry is constructed by introducing a latent parameter, \(\theta \sim U\left(-3\pi/2, 3\pi/2\right)\).
-This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure: \[X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1)\].
-This configuration creates a horizontally curled shape in \((X_1, X_3)\), with additional band thickness in the \(X_2\) direction.
-For \(p > 3\), additional noise dimensions are appended to embed the structure in higher dimensions.
-
-\begin{verbatim}
-scurve <- gen_scurve(n = 1000, p = 4)
-\end{verbatim}
-
-To introduce missing or incomplete regions on the manifold, \texttt{gen\_scurvehole(n,\ p)} (Figure \ref{fig:scurve-proj} b) removes points in a localized region centered around the middle vertical section of the S-curve.
-Following a similar approach as \texttt{gen\_cubehole()}, all observations within a fixed radius (\(\sqrt{0.3}\)) of the anchor point are excluded, creating a hole in the manifold while preserving the overall S-curve structure.
-
-\begin{verbatim}
-scurvehole <- gen_scurvehole(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.}\label{fig:scurve-proj}
-\end{figure}
-
-\paragraph{Sphere}\label{sphere}
-
-Sphere-shaped structures (Figure \ref{fig:sphere-proj}) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
-The functions generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
-The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
-For the circle and curvy cycle, this occurs when \(p > 2\) and \(p > 3\), respectively, while for spheres and hemispheres it occurs when \(p > 3\) or \(p > 4\).
-
-The simplest case, \texttt{gen\_circle(n,\ p)} creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (Figure \ref{fig:sphere-proj} a).
-Let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
-Coordinates in the first two dimensions represent a perfect circle on the plane: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta)\].
-For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
-The first component is a scaling factor that decreases with the dimension index, defined as \(\text{scale}_j = \sqrt{(0.5)^{j-2}}\) for \(j = 3, \dots, p\).
-The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula \(\phi_j = (j - 2)\pi/2p\).
-Each additional dimension is computed as: \(X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p\).
-
-\begin{verbatim}
-circle <- gen_circle(n = 1000, p = 4)
-\end{verbatim}
-
-For the one-dimensional nonlinear cycle embedded in \(p\text{-}D\) space, \texttt{gen\_curvycycle(n,\ p)} (Figure \ref{fig:sphere-proj} b), let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
-The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle''.
-In this configuration, \(X_1 = \cos(\theta)\) represents horizontal oscillation, while \(X_2 = \sqrt{3}/3 + \sin(\theta)\) introduces a vertical offset to avoid centering the curve at the origin.
-Additionally, \(X_3 = 1/3\cos(3\theta)\) introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
-
-Together, these define a periodic, non-trivial, closed curve in \(3\text{-}D\) with internal folds that produce a more complex geometry than a standard circle or ellipse.
-For dimensions \(X_4\) through \(X_p\), additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
-The scaling factor is defined as \(\text{scale}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from \(4\) to \(p\), which means that the amplitude decreases as the dimension increases.
-Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j\sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2)\pi/2p\).
-
-\begin{verbatim}
-curvycycle <- gen_curvycycle(n = 1000, p = 4)
-\end{verbatim}
-
-Building on simple circular structures, the \texttt{gen\_unifsphere(n,\ p,\ r)} function function extends the idea to three dimensions by generating \(n\) observations approximately uniformly distributed on the surface of a sphere of radius \(r\), with optional high-dimensional noise when \(p > 3\) (Figure \ref{fig:sphere-proj} c).
-Each observation is computed from spherical coordinates, with \(u \sim U(-1, 1)\) representing \(\cos(\phi)\) and \(\theta \sim U(0, 2\pi)\) the azimuthal angle.
-Cartesian coordinates are then defined as \[X_1 = r\sqrt{1 - u^2}\cos(\theta), \quad X_2 = r\sqrt{1 - u^2}\sin(\theta),\text{ and }X_3 = ru,\] ensuring uniform distribution on the surface (not within) of the sphere.
-For \(p > 3\), additional dimensions \(X_4\) through \(X_p\) are appended as structured noise.
-
-\begin{verbatim}
-unifsphere <- gen_unifsphere(n = 1000, p = 4)
-\end{verbatim}
-
-In addition, the \texttt{gen\_gridedsphere(n,\ p)} function constructs a \(p\text{-}D\) dataset consisting of approximately \(n\) points that are evenly distributed on the surface of the unit \((p-1)\)-sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
-The method relies on forming a regular grid in spherical coordinates, parameterized by \((p-1)\) angular variables: for dimensions \(j = 1, \dots, p-2\) the polar angles are drawn from \([0, \pi]\), while the final angle (\(j = p-1\)) represents the azimuth and is drawn from \([0, 2\pi]\).
-The number of grid steps along each angular dimension is chosen by decomposing \(n\) into \((p-1)\) approximately equal integer factors using the helper function \texttt{gen\_nproduct(n,\ p\ -\ 1)}.
-
-Each grid point is subsequently mapped into Cartesian space via the standard hyperspherical-to-Cartesian transformation,
-
-\[
-\begin{aligned}
-X_1 &= \cos(\theta_1), \\
-X_2 &= \sin(\theta_1)\cos(\theta_2), \\
-X_3 &= \sin(\theta_1)\sin(\theta_2)\cos(\theta_3), \\
-&\;\;\vdots \\
-X_{p-1} &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\cos(\theta_{p-1}), \\
-X_p &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\sin(\theta_{p-1}).
-\end{aligned}
-\]
-
-The result is a deterministic grid of points lying exactly on the surface of the unit \((p-1)\)-sphere, without any additional noise dimensions.
-
-\begin{verbatim}
-gridedsphere <- gen_gridedsphere(n = 1000, p = 4)
-\end{verbatim}
-
-For more heterogeneous structures, the \texttt{gen\_clusteredspheres(n,\ k,\ p,\ r,\ loc)} function generates one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location (Figure \ref{fig:sphere-proj} e).
-A large uniform sphere centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\text{-}D\) sphere with a radius of \(r_1\).
-The sampling is executed using the function \texttt{gen\_unifsphere(n\_1,\ p,\ r\_1)}, which generates the desired points in the specified dimensional space.
-In generation of \(k\) smaller uniform spheres, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
-These spheres are positioned at distinct random locations in \(p\)-space, with the center of each sphere being drawn from a normal distribution \(N(0, \texttt{loc}^2 I_p)\).
-Points on spheres are generated using the standard hyperspherical method, which involves sampling \(u \sim U(-1, 1)\) to determine the cosine of the polar angle, and sampling \(\theta \sim U(0, 2\pi)\) to determine the azimuthal angle (for \(3\text{-}D\)).
-Each observation is classified by cluster, with labels such as ``big'' for the large central sphere and ``small\_1'' to ``small\_k'' for the smaller spheres.
-
-\begin{verbatim}
-clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, p = 4, r = c(15, 3),
-                                         loc = 10 / sqrt(3)) |>
-  dplyr::select(-cluster)
-\end{verbatim}
-
-Finally, the \texttt{gen\_hemisphere(n,\ p)} function restricts sampling to a hemisphere of a \(4\text{-}D\) sphere (Figure \ref{fig:sphere-proj} f).
-Using spherical coordinates, the azimuthal angle \(\theta_1 \sim U(0, \pi)\) in the \((x_1, x_2)\) plane, while the elevation angle \(\theta_2 \sim U(0, \pi)\) in the \((x_2, x_3)\) plane.
-Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
-For \(p > 4\), additional noise dimensions are added to embed the structure in higher dimensions.
-
-\begin{verbatim}
-hemisphere <- gen_hemisphere(n = 1000, p = 4)
-\end{verbatim}
-
-\begin{figure}[!]
-
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}
-
-}
-
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.}\label{fig:sphere-proj}
-\end{figure}
-
-\paragraph{Swiss Roll}\label{swiss-roll}
-
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ p,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension (Figure \ref{fig:swissroll-proj}).
-The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.
-The \[X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \text{ where } t \sim U(0, 3\pi).\] For \(p > 3\), the remaining \(p - 3\) dimensions are filled with noise to simulate high-dimensional complexity.
-
-\begin{verbatim}
-swissroll <- gen_swissroll(n = 1000, p = 4, w = c(-1, 1))
-\end{verbatim}
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/swissroll-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `swissroll` data.}\label{fig:swissroll-proj}
-\end{figure}
-
-\paragraph{Trigonometric}\label{trigonometric}
-
-Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (Figure \ref{fig:triginometric-proj}).
-The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when \(p > 2\) (adding \(X_3\) through \(X_p\)), and for all other trigonometric shapes when \(p > 4\) (adding \(X_5\) through \(X_p\)).
-These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
-
-First, the \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise (Figure \ref{fig:triginometric-proj} a).
-Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
-The corresponding \(2\text{-}D\) coordinates are defined by: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta).\]
-
-\begin{verbatim}
-crescent <- gen_crescent(n = 1000, p = 4)
-\end{verbatim}
-
-Second, the \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
-The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
-Let \(\theta \sim U(0, 3\pi)\) represent a random angle on a circular base and \(z \sim U(0, h)\) represent the height along the cylinder.
-The coordinates are defined as: \(X_1 = \cos(\theta)\) (Circular base, x-axis), \(X_2 = \sin(\theta)\) (Circular base, y-axis), \(X_3 = z\) (Linear height), and \(X_4 = \sin(z)\) (Nonlinear curvy variation along height).
-
-\begin{verbatim}
-curvycylinder <- gen_curvycylinder(n = 1000, p = 4, h = 10)
-\end{verbatim}
-
-For a spiraling path on a spherical surface in the first four dimensions, \texttt{gen\_sphericalspiral(n,\ p,\ spins)} (Figure \ref{fig:triginometric-proj} c), let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\)be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
-Cartesian coordinates from spherical conversion: \(X_1 = \sin(\phi)\cos(\theta)\), \(X_2 = \sin(\phi)\sin(\theta)\), \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter, and \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
-This generates a spherical spiral curve embedded in \(4\text{-}D\) space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
-
-\begin{verbatim}
-sphericalspiral <- gen_sphericalspiral(n = 1000, p = 4, spins = 1)
-\end{verbatim}
-
-For a helical spiral in four dimensions, \texttt{gen\_helicalspiral(n,\ p)} (Figure \ref{fig:triginometric-proj} d), let \(\theta \in [0, 5\pi/4]\) be a sequence of angles controlling rotation around a circle.
-Cartesian coordinates; \(X_1 = \cos(\theta)\): circular trajectory along the x-axis, \(X_2 = \sin(\theta)\): circular trajectory along the y-axis, \(X_3 = 0.05\theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix, and \(X_4 = 0.1\sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
-
-\begin{verbatim}
-helicalspiral <- gen_helicalspiral(n = 1000, p = 4)
-\end{verbatim}
-
-Similarly, the \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of \(p\text{-}D\) (Figure \ref{fig:triginometric-proj} e).
-The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a \(3\text{-}D\) conic helix.
-The shape is defined by parameter \(\theta \in [0, 2\pi\text{spins}]\), controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; \(X_1 = \theta\cos(\theta)\) for radial expansion in x, and \(X_2 = \theta\sin(\theta)\) for radial expansion in y.
-The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2\theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\) Spiral modulation in the fourth dimension is represented by \(X_4 = \theta\sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
-
-\begin{verbatim}
-conicspiral <- gen_conicspiral(n = 1000, p = 4, spins = 1)
-\end{verbatim}
-
-Finally, the \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
-The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-
-\begin{verbatim}
-nonlinear <- gen_nonlinear(n = 1000, p = 4, hc = 1, non_fac = 0.5)
-\end{verbatim}
-
-\begin{figure}[!]
-
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
-
-}
-
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.}\label{fig:triginometric-proj}
-\end{figure}
-
-\paragraph{Trefoil knots}\label{trefoil-knots}
-
-The Trefoil is a closed, nontrivial one-dimensional manifold embedded in \(3\text{-}D\) or \(4\text{-}D\) space (Figure \ref{fig:trefoil-proj}).
-The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
-
-For the \(4\text{-}D\) trefoil knot, the function \texttt{gen\_trefoil4d(n,\ p,\ steps)} generates the structure on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)) using two angular parameters, \(\theta\) and \(\phi\).
-A band of thickness around the knot path is controlled by the \texttt{steps} argument, while the number of \(\theta\) and \(\phi\) values is determined by the \texttt{steps} and \texttt{n} arguments, respectively (Figure \ref{fig:trefoil-proj} a).
-The coordinates are defined as \[X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi)\], where \(\theta\) and \(\phi\) trace the knot's path.
-For \(p > 4\), the quadratic structure is embedded into higher dimensions by appending additional noise dimensions.
-
-\begin{verbatim}
-trefoil4d <- gen_trefoil4d(n = 500, p = 4, steps = 5)
-\end{verbatim}
-
-For the \(3\text{-}D\) stereographic projection, \texttt{gen\_trefoil3d(n,\ p,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \[(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),\] excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
-As with the \(4\text{-}D\) case, optional noise dimensions can be added to embed the knot into higher-dimensional spaces.
-
-\begin{verbatim}
-trefoil3d <- gen_trefoil3d(n = 500, p = 4, steps = 5)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.}\label{fig:trefoil-proj}
-\end{figure}
+\subsubsection{Wrappers}\label{wrappers}
 
@@
 Table \ref{tab:noise-tb-pdf} details these functions.
+
+\begin{table}
+
+\caption{\label{tab:noise-tb-pdf}cardinalR noise dimensions generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_noisedims & Gaussian noise dimensions with optional mean and standard deviation.\\
+gen\_wavydims1 & Wavy noise dimensions based on a user-specified theta sequence with added jitter.\\
+gen\_wavydims2 & Wavy noise dimensions using polynomial transformations of an existing dimension vector.\\
+gen\_wavydims3 & Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-For scenarios where noise should follow a smooth wavy pattern, \texttt{gen\_wavydims1(n,\ p,\ \textbackslash{}theta)} generates dimensions as
+For scenarios where noise should follow a smooth wavy pattern, \texttt{gen\_wavydims1(n,\ p,\ theta)} generates dimensions as
 
@@
 
-Finally, \texttt{gen\_wavydims3(n,\ p,\ \textbackslash{}text\{data\})} generates noise for datasets with multiple correlated dimensions.
+Finally, \texttt{gen\_wavydims3(n,\ p,\ data)} generates noise for datasets with multiple correlated dimensions.
 The first three dimensions are small perturbations of the original coordinates \((X_1, X_2, X_3)\), while higher dimensions are constructed via non-linear combinations, including polynomial and trigonometric transformations, e.g.,
@@
 \centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\raggedright\arraybackslash}p{8.5cm}}
 \toprule
@@
 \midrule
-make\_mobiusgau & Möbius-like cluster combined with a Gaussian cluster.\\
+make\_mobiusgau & Möbius-like cluster combined with a Gaussian.\\
 make\_multigau & Multiple Gaussian clusters in high-dimensional space.\\
@@
 The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in \(4\text{-}D\) space (Figure \ref{fig:highd-proj}).
+Each cluster was chosen to reflect distinct geometric and topological properties, allowing us to test how well DR methods preserve both local and global structures.
+The \emph{helical spiral} cluster is designed to evaluate methods on elongated, twisting structures that challenge linear embeddings such as PCA and require preservation of curvilinear continuity.
+The \emph{hemisphere} provides a curved surface with partial coverage of a \(3\text{-}D\) manifold, useful for testing neighborhood preservation and unfolding in algorithms like UMAP and tSNE.
+The \emph{uniform cube} represents isotropic, uniformly distributed data and serves as a control cluster with simple geometric structure to assess baseline embedding fidelity.
+The \emph{cone} introduces variable density along one axis, mimicking structures where point density changes with geometry, helping evaluate how well algorithms maintain relative distances in non-uniform distributions.
+Finally, the \emph{Gaussian} cluster is a standard multivariate normal distribution, included to assess algorithm performance on simple, spherical, high-density clusters.
+Together, these clusters create a challenging synthetic dataset suitable for benchmarking and exploring the strengths and weaknesses of different dimensionality reduction techniques.
 
@@
 \end{verbatim}
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
-The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:highd-proj}
-\end{figure}
 
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:highd-proj}
 \end{figure}
@@
 
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
+\end{figure}
+
 \section{Conclusion}\label{conclusion}
@@
 The \texttt{cardinalR} package introduces a flexible framework for generating high-dimensional data structures with well-defined geometric properties.
-It addresses an important need in the evaluation of clustering, machine learning, and dimensionality reduction (DR) methods by enabling the construction of customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
+It addresses an important need in the evaluation of clustering, machine learning, and DR methods by enabling the construction of customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
 In this way, \texttt{cardinalR} complements existing packages such as \texttt{geozoo}, \texttt{snedata}, and \texttt{mlbench}, while extending the scope to higher dimensions and more complex shapes.
+
+The motivation for developing this package originated from the need to design a perception--misperception experiment, aimed at investigating how well NLDR methods preserve inter-cluster structure.
+To conduct this study, we required simulated datasets with carefully controlled geometric and clustering properties.
+While some existing packages provided useful starting points, none fully supported the creation of flexible, high-dimensional data with the specific structural variations needed for our experiment.
+Developing these generators for research purposes gradually led to the design of \texttt{cardinalR} as a general-purpose package, so that other researchers can benefit from the same tools for simulation, benchmarking, and teaching.
 
@@
 
-The source material for this paper is available at \url{https://github.com/JayaniLakshika/paper-cardinalR}.
+The source material for this paper is available at \href{https://github.com/JayaniLakshika/paper-cardinalR}{github.com/JayaniLakshika/paper-cardinalR}.
 
+This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.
 These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{tibble} \citep{kirill2023}, \texttt{gtools} \citep{gregory2023}, \texttt{dplyr} \citep{hadley2023}, \texttt{stats} \citep{core2025}, \texttt{tidyr} \citep{hadley2024}, \texttt{purrr} \citep{hadley2025}, \texttt{mvtnorm} \citep{alan2009}, \texttt{geozoo} \citep{barret2016}, and \texttt{MASS} \citep{venables2002}.
-This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.
 

### 2025-10-02 paper-cardinalR.R

@@ -89,7 +89,6 @@ source("scripts/additional_functions.R")
 
 ## -----------------------------------------------------------------------------
 main_tb <- tibble(arg = c("n",
-                          "p",
                           "k",
                           "loc",
                           "scale",
@@ -98,19 +97,17 @@ main_tb <- tibble(arg = c("n",
                           "is_bkg"),
                   type = c("numeric (vector)",
                            "numeric",
-                           "numeric",
                            "numeric (matrix)",
                            "numeric (vector)",
                            "character (vector)",
                            "numeric (list)",
                            "boolean"),
                         exp = c("Number of points in each cluster.",
-                                "Number of dimensions.",
                                 "Number of clusters.",
                                 "Locations/centroids of clusters.",
                                 "Scaling factors of clusters.",
                                 "Shapes of clusters.",
-                                "Plane and the corresponding angle along that plane for each cluster.",
+                                "Rotation matrices, one per cluster.",
                                 "Background noise should exist or not."))
 
 
@@ -122,3668 +119,9 @@ main_tb <- tibble(arg = c("n",
 ## ----main-tb-pdf, eval=knitr::is_latex_output()-------------------------------
 main_tb |> 
   kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "4cm") |>
+  column_spec(1, width = "2cm") |>
+  column_spec(2, width = "3cm") |>
   column_spec(3, width = "8cm")
-
-
-## -----------------------------------------------------------------------------
-branch_tb <- tibble(fun = c("gen_expbranches",
-                             "gen_linearbranches",
-                             "gen_curvybranches", 
-                             "gen_orglinearbranches", 
-                             "gen_orgcurvybranches"), 
-                      exp = c("Exponential shaped branches.",
-                               "Linear shaped branches.", 
-                               "Curvy shaped branches.",
-                               "Linear shaped branches originated in one point.",
[…]

### 2025-10-02 paper-cardinalR.Rmd
Added the section that doesn't need to change

@@
 All shape generators allow the user to define the number of points (`n`) and dimensions (`p`), and most include additional arguments to customize specific characteristics of the structure.
+
+#### Cone
+
+To simulate a cone-shaped structure in arbitrary dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cone)"} else { "Figure \\@ref(fig:cone-proj)"})`), we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+
+This function generates $n$ points in $p\text{-}D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip.
+Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, capped at the cone height $h$, producing a higher density of points near the tip.
+At each height $X_p$, the radius of the cross-section decreases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio}$ and $r_{\text{max}} = 1$.
+
+For each point, a direction in the first $p-1$ dimensions is sampled uniformly on a $(p-1)$-dimensional hypersphere using generalized spherical coordinates.
+The radial coordinates are scaled by the height-dependent radius $r$, producing the conical taper.
+In three dimensions ($p = 3$), this results in a classical $3\text{-}D$ cone, while for $p > 3$, additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
+
+```{r data-bluntedcone, echo=TRUE}
+cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
+```
+
+```{r}
+cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r cone-proj1}
+scaled_data <- scale_data_manual(cone)
+
+## First projection
+projection <- cbind(
+  c(0.22417,-0.02943,-0.02870,0.08418),
+  c(-0.02715,0.18019,0.05053,0.15254))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.25,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.022))
+
+cone_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.25), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cone-proj2}
+
+## Second projection
+projection <- cbind(
+    c(0.15591,0.17907,0.03198,0.04496),
+    c(-0.15376,0.10435,0.15766,0.00547))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.1,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.08, 
+                                              axis_pos_y = -0.08, 
+                                              threshold = 0.01))
+
+cone_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.1, 0.12), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cone-proj3}
+
+## Third projection
+projection <- cbind(
+    c(-0.13548,0.13832,0.05991,-0.13544),
+    c(0.08479,-0.10719,-0.01572,-0.20123))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.25,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.016))
+
+cone_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.2), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $4\\text{-}D$."}
+
+conefig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    cone_lang
+  ),
+  device = "xs"
+)
+
+class(conefig) <- c(class(conefig), "htmlwidget")
+
+conefig
+```
+
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."}
+#| fig-width: 12
+#| fig-height: 4
+
+cone_proj1 + cone_proj2 + cone_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+#### Linear
+
+The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise.
+Each variable is formed as $X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i$, where $\text{scale}\_i \sim U(-10, 10)$ determines the orientation of the line in each dimension, $\text{shift}\_i \sim U(-300, 300)$ offsets the line to separate dimensions, and $\epsilon \sim N(0, (0.03n)^2)$ introduces Gaussian noise.
+
+```{r data-linear, echo=TRUE}
+linear <- gen_longlinear(n = 1000, p = 4)
+```
+
+```{r}
+linear_lang <- langevitour::langevitour(linear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r linear-proj1}
+scaled_data <- scale_data_manual(linear)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
+                                              threshold = 0.022))
+
+linear_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linear-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.75,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.01))
+
+linear_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.9, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linear-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.9,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
+                                              threshold = 0.02))
+
+linear_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.7), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r linear, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `linear` data in $4\\text{-}D$."}
+
+linearfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    linear_lang
+  ),
+  device = "xs"
+)
+
+class(linearfig) <- c(class(linearfig), "htmlwidget")
+
+linearfig
+```
+
+```{r linear-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linear` data."}
+#| fig-width: 12
+#| fig-height: 4
+
+linear_proj1 + linear_proj2 + linear_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+#### Gaussian
+
+The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:gau)"} else { "Figure \\@ref(fig:gau-proj)"})`).
+Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
+
+```{r data-gau, echo=TRUE}
+gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
+```
+
+```{r}
+gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r gau-proj1}
+scaled_data <- scale_data_manual(gau)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+gau_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gau-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+gau_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gau-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+gau_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gau, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gau` data in $4\\text{-}D$."}
+
+gaufig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    gau_lang
+  ),
+  device = "xs"
+)
+
+class(gaufig) <- c(class(gaufig), "htmlwidget")
+
+gaufig
+```
+
+```{r gau-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gau` data."}
+#| fig-width: 12
+#| fig-height: 4
+
+gau_proj1 + gau_proj2 + gau_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
 

### 2025-10-02 paper-cardinalR.tex
Added the section that doesn't need to change

@@
 All shape generators allow the user to define the number of points (\texttt{n}) and dimensions (\texttt{p}), and most include additional arguments to customize specific characteristics of the structure.
+
+\paragraph{Cone}\label{cone}
+
+To simulate a cone-shaped structure in arbitrary dimensions (Figure \ref{fig:cone-proj}), we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+
+This function generates \(n\) points in \(p\text{-}D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip.
+Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), capped at the cone height \(h\), producing a higher density of points near the tip.
+At each height \(X_p\), the radius of the cross-section decreases linearly from base to tip according to \(r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h\), where \(r_{\text{min}} = \text{ratio}\) and \(r_{\text{max}} = 1\).
+
+For each point, a direction in the first \(p-1\) dimensions is sampled uniformly on a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates.
+The radial coordinates are scaled by the height-dependent radius \(r\), producing the conical taper.
+In three dimensions (\(p = 3\)), this results in a classical \(3\text{-}D\) cone, while for \(p > 3\), additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
+
+\begin{verbatim}
+cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
+\end{verbatim}
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:cone-proj}
+\end{figure}
+
+\paragraph{Linear}\label{linear}
+
+The \texttt{gen\_longlinear(n,\ p)} function generates a high-dimensional dataset representing a long linear structure with noise.
+Each variable is formed as \(X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i\), where \(\text{scale}\_i \sim U(-10, 10)\) determines the orientation of the line in each dimension, \(\text{shift}\_i \sim U(-300, 300)\) offsets the line to separate dimensions, and \(\epsilon \sim N(0, (0.03n)^2)\) introduces Gaussian noise.
+
+\begin{verbatim}
+linear <- gen_longlinear(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/linear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linear` data.}\label{fig:linear-proj}
+\end{figure}
+
+\paragraph{Gaussian}\label{gaussian}
+
+The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure (Figure \ref{fig:gau-proj}).
+Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, s)\), where \(s\) is a user-defined \(p \times p\) positive-definite matrix.
+
+\begin{verbatim}
+gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
+\end{verbatim}
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:gau-proj}
+\end{figure}
 

### 2025-10-02 paper-cardinalR.Rmd
Updated trefoil

@@
 
+#### Pyramid
+
+A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
+The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
+Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:pyramid-tb-html)', '\\@ref(tab:pyramid-tb-pdf)'))` summarizes these functions.
+
+Let $X_1, \dots, X_p$ denote the coordinates of the generated points.
+For the rectangular and triangular based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
+That is, $$X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).$$ This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
+That is, $X_p = z \sim U(0, h)$.
+
+The remaining dimensions are based on the specific pyramid shape.
+For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`), let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$.
+That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
+The first three coordinates are then defined as: $$X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).$$
+
+```{r data-pyrrect, echo=TRUE}
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+```
+
+```{r}
+pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r pyrrect-proj1}
+scaled_data <- scale_data_manual(pyrrect)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrrect_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrrect-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
+                                              threshold = 0.08))
+
+pyrrect_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrrect-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrrect_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.75), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For the triangular based pyramid, `gen_pyrtri(n, p, h, l, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`), let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$.
+That is, $r(z) = r_t + (l-r_t)z/h$.
+A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
+The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v.$
+
+```{r data-pyrtri, echo=TRUE}
+pyrtri <- gen_pyrtri(n = 1000, p = 4)
+```
+
+```{r}
+pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r pyrtri-proj1}
+scaled_data <- scale_data_manual(pyrtri)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.022))
+
+pyrtri_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.3), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrtri-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.03))
+
+pyrtri_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.4), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrtri-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrtri_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For the star based pyramid, `gen_pyrstar(n, p, h, rb)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`), let the radius at height $z$, $r(z)$, be such that the radius scales linearly from zero (tip) to the base radius $r_b$.
+That is, $r(z) = r_b\left(1 - z/h\right)$.
+
+Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
+$r_{\text{point}} \sim \sqrt{U(0, 1)}$. Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
+
+```{r data-pyrstar, echo=TRUE}
+pyrstar <- gen_pyrstar(n = 1000, p = 4)
+```
+
+```{r}
+pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r pyrstar-proj1}
+scaled_data <- scale_data_manual(pyrstar)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrstar_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrstar-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrstar_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.4), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrstar-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+pyrstar_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "c3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For all the above pyramid shapes, if $p > 3$, the remaining $p - 3$ dimensions (i.e., $X_4$ to $X_{p-1}$) are additional noise.
+
+Finally, for the Sierpinski-like pyramid, `gen_pyrfrac(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`), let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
+The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
+Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
+At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
+This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
+The points remain bounded inside the convex hull of the simplex.
+The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
+
+```{r data-pyrholes, echo=TRUE}
+pyrholes <- gen_pyrfrac(n = 1000, p = 4)
+```
+
+```{r}
+pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r pyrholes-proj1}
+scaled_data <- scale_data_manual(pyrholes)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrholes_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "d1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrholes-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrholes_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.75), 
+  title = "d2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrholes-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+pyrholes_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "d3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r}
+pyramid_tb <- tibble(fun = c("gen_pyrrect",
+                             "gen_pyrtri",
+                             "gen_pyrstar",
+                             "gen_pyrfrac"), 
+                        exp = c("Rectangular-base pyramid, with a sharp or blunted apex.",
+          "Triangular-base pyramid, with a sharp or blunted apex.",
+          "Star-shaped base pyramid, with a sharp or blunted apex.",
+          "Pyramid containing triangular pyramid-shaped holes."))
+```
+
+```{r pyramid-tb-html, eval=knitr::is_html_output()}
+pyramid_tb |> 
+  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r pyramid-tb-pdf, eval=knitr::is_latex_output()}
+pyramid_tb |> 
+  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$."}
+
+pyrfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    pyrrect_lang, pyrtri_lang, pyrstar_lang, pyrholes_lang
+  ),
+  device = "xs"
+)
+
+class(pyrfig) <- c(class(pyrfig), "htmlwidget")
+
+pyrfig
+```
+
+```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=12, fig.height=16}
+
+pyrrect_proj1 + pyrrect_proj2 + pyrrect_proj3 +
+pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
+pyrstar_proj1 + pyrstar_proj2 + pyrstar_proj3 +
+pyrholes_proj1 + pyrholes_proj2 + pyrholes_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+#### Trefoil knots
+
+<!--https://laustep.github.io/stlahblog/posts/TorusKnot4D.html-->
+
+The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil)"} else { "Figure \\@ref(fig:trefoil-proj)"})`).
+The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trefoil-tb-html)', '\\@ref(tab:trefoil-tb-pdf)'))` summarizes these functions.
+
+```{r}
+trefoil_tb <- tibble(fun = c("gen_trefoil4d",
+                            "gen_trefoil3d"), 
+                        exp = c("Trefoil in $4\\text{-}D$.",
+                                "Trefoil in $3\\text{-}D$."))
+```
+
+```{r trefoil-tb-html, eval=knitr::is_html_output()}
+trefoil_tb |> 
+  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r trefoil-tb-pdf, eval=knitr::is_latex_output()}
+
+trefoil_tb |> 
+  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+For the $4\text{-}D$ trefoil knot, the function `gen_trefoil4d(n, steps)` generates the structure on the $3$-sphere ($S^3 \subset \mathbb{R}^4$) using two angular parameters, $\theta$ and $\phi$.
+A band of thickness around the knot path is controlled by the `steps` argument, while the number of $\theta$ and $\phi$ values is determined by the `steps` and `n` arguments, respectively (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
+The coordinates are defined as $$X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),$$ where $\theta$ and $\phi$ trace the knot’s path.
+
+```{r data-trefoil4d, echo=TRUE}
+trefoil4d <- gen_trefoil4d(n = 500, steps = 5)
+```
+
+```{r}
+trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r trefoil4d-proj1}
+scaled_data <- scale_data_manual(trefoil4d)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.022))
+
+trefoil4d_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil4d-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+trefoil4d_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.65), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil4d-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
+                                              threshold = 0.016))
+
+trefoil4d_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For the $3\text{-}D$ stereographic projection, `gen_trefoil3d(n, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $$(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),$$ excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
+
+```{r data-trefoil3d, echo=TRUE}
+trefoil3d <- gen_trefoil3d(n = 500, steps = 5)
+```
+
+```{r}
+trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r trefoil3d-proj1}
+scaled_data <- scale_data_manual(trefoil3d)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097),
+  c(0.00896,0.52496,0.12597))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.022))
+
+trefoil3d_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.7), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil3d-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278),
+  c(-0.50493,0.14388,0.14615))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+trefoil3d_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.65), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil3d-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416),
+  c(0.50329,0.07199,-0.16064))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
+                                              threshold = 0.016))
+
+trefoil3d_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` and `trefoil3d` data in $4\\text{-}D$."}
+
+trefoilfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    trefoil4d_lang, trefoil3d_lang
+  ),
+  device = "xs"
+)
+
+class(trefoilfig) <- c(class(trefoilfig), "htmlwidget")
+
+trefoilfig
+```
+
+```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=12, fig.height=8}
+
+trefoil4d_proj1 + trefoil4d_proj2 + trefoil4d_proj3 + 
+  trefoil3d_proj1 + trefoil3d_proj2 + trefoil3d_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 ### Wrappers

### 2025-10-02 paper-cardinalR.tex
Updated trefoil

@@
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:gau-proj}
+\end{figure}
+
+\paragraph{Pyramid}\label{pyramid}
+
+A pyramid structure (Figure \ref{fig:pyr-proj}) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
+The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
+Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
+Table \ref{tab:pyramid-tb-pdf} summarizes these functions.
+
+Let \(X_1, \dots, X_p\) denote the coordinates of the generated points.
+For the rectangular and triangular based pyramid generator functions, the final dimension, \(X_p\), encodes the height of each point and is drawn from an exponential distribution capped at the maximum height \(h\).
+That is, \[X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).\] This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
+That is, \(X_p = z \sim U(0, h)\).
+
+The remaining dimensions are based on the specific pyramid shape.
+For the rectangular based pyramid, \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} (Figure \ref{fig:pyr-proj} a), let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\).
+That is, \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
+The first three coordinates are then defined as: \[X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).\]
+
+\begin{verbatim}
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+\end{verbatim}
+
+For the triangular based pyramid, \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} (Figure \ref{fig:pyr-proj} b), let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\).
+That is, \(r(z) = r_t + (l-r_t)z/h\).
+A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section: \(u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v\).
+The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z)u\), and \(X_3 = r(z)v.\)
+
+\begin{verbatim}
+pyrtri <- gen_pyrtri(n = 1000, p = 4)
+\end{verbatim}
+
+For the star based pyramid, \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} (Figure \ref{fig:pyr-proj} c), let the radius at height \(z\), \(r(z)\), be such that the radius scales linearly from zero (tip) to the base radius \(r_b\).
+That is, \(r(z) = r_b\left(1 - z/h\right)\).
+
+Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
+\(r_{\text{point}} \sim \sqrt{U(0, 1)}\). Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
+
+\begin{verbatim}
+pyrstar <- gen_pyrstar(n = 1000, p = 4)
+\end{verbatim}
+
+For all the above pyramid shapes, if \(p > 3\), the remaining \(p - 3\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are additional noise.
+
+Finally, for the Sierpinski-like pyramid, \texttt{gen\_pyrfrac(n,\ p)} (Figure \ref{fig:pyr-proj} d), let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points.
+The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution: \(T_0 \sim U(0, 1)^p\).
+Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\text{-}D\) simplex.
+At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\): \(T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}\).
+This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
+The points remain bounded inside the convex hull of the simplex.
+The final output is a \(n \times p\) matrix where each row represents a point: \(X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}\).
+
+\begin{verbatim}
+pyrholes <- gen_pyrfrac(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{table}
+
+\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_pyrrect & Rectangular-base pyramid, with a sharp or blunted apex.\\
+gen\_pyrtri & Triangular-base pyramid, with a sharp or blunted apex.\\
+gen\_pyrstar & Star-shaped base pyramid, with a sharp or blunted apex.\\
+gen\_pyrfrac & Pyramid containing triangular pyramid-shaped holes.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.}\label{fig:pyr-proj}
+\end{figure}
+
+\paragraph{Trefoil knots}\label{trefoil-knots}
+
+The Trefoil is a closed, nontrivial one-dimensional manifold embedded in \(3\text{-}D\) or \(4\text{-}D\) space (Figure \ref{fig:trefoil-proj}).
+The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
+Table \ref{tab:trefoil-tb-pdf} summarizes these functions.
+
+\begin{table}
+
+\caption{\label{tab:trefoil-tb-pdf}cardinalR trefoil data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_trefoil4d & Trefoil in \$4\textbackslash{}text\{-\}D\$.\\
+gen\_trefoil3d & Trefoil in \$3\textbackslash{}text\{-\}D\$.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+For the \(4\text{-}D\) trefoil knot, the function \texttt{gen\_trefoil4d(n,\ steps)} generates the structure on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)) using two angular parameters, \(\theta\) and \(\phi\).
+A band of thickness around the knot path is controlled by the \texttt{steps} argument, while the number of \(\theta\) and \(\phi\) values is determined by the \texttt{steps} and \texttt{n} arguments, respectively (Figure \ref{fig:trefoil-proj} a).
+The coordinates are defined as \[X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),\] where \(\theta\) and \(\phi\) trace the knot's path.
+
+\begin{verbatim}
+trefoil4d <- gen_trefoil4d(n = 500, steps = 5)
+\end{verbatim}
+
+For the \(3\text{-}D\) stereographic projection, \texttt{gen\_trefoil3d(n,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \[(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),\] excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
+
+\begin{verbatim}
+trefoil3d <- gen_trefoil3d(n = 500, steps = 5)
+\end{verbatim}
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.}\label{fig:trefoil-proj}
 \end{figure}

### 2025-10-02 paper-cardinalR.R

@@ -463,6 +463,635 @@ gau_proj1 + gau_proj2 + gau_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
 
 
+## ----data-pyrrect, echo=TRUE--------------------------------------------------
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+
+
+## -----------------------------------------------------------------------------
+pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+
+
+## ----pyrrect-proj1------------------------------------------------------------
+scaled_data <- scale_data_manual(pyrrect)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrrect_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+
+
+## ----pyrrect-proj2------------------------------------------------------------
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
+                                              threshold = 0.08))
+
+pyrrect_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.6), 
+  title = "a2", 
+  cex = 2, 
[…]

### 2025-10-02 paper-cardinalR.Rmd
Added Swissroll section

@@
 
+#### Polynomial
+
+A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in \gD{} space.
+To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
+These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
+We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:polynomial-tb-html)', '\\@ref(tab:polynomial-tb-pdf)'))` summarizes these functions and their purposes.
+
+```{r}
+polynomial_tb <- tibble(fun = c("gen_quadratic",
+                                "gen_cubic"), 
+                        exp = c("Quadratic pattern.",
+                                "Cubic pattern."))
+```
+
+```{r polynomial-tb-html, eval=knitr::is_html_output()}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r polynomial-tb-pdf, eval=knitr::is_latex_output()}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+The first is the quadratic curve of $n$ points in two dimensions.
+This is generated using `gen_quadratic(n, range)`.
+The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree 2 is applied to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
+This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
+
+```{r data-quadratic, echo=TRUE}
+quadratic <- gen_quadratic(n = 1000)
+```
+
+The second is the cubic curve of $n$ points in two dimensions.
+This is generated using `gen_cubic(n, range)`.
+The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
+This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
+
+```{r data-cubic, echo=TRUE}
+cubic <- gen_cubic(n = 1000)
+```
+
+#### Swiss Roll
+
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:swissroll)"} else { "Figure \\@ref(fig:swissroll-proj)"})`).
+The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.
+The $$X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \text{ where } t \sim U(0, 3\pi).$$
+
+```{r data-swissroll, echo=TRUE}
+swissroll <- gen_swissroll(n = 1000, w = c(-1, 1))
+```
+
+```{r}
+swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r swissroll-proj1}
+scaled_data <- scale_data_manual(swissroll)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097),
+  c(0.00896,0.52496,0.12597))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+swissroll_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r swissroll-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278),
+  c(-0.50493,0.14388,0.14615))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+swissroll_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.65), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r swissroll-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416),
+  c(0.50329,0.07199,-0.16064))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.55, 
+                                              axis_pos_y = -0.55, 
+                                              threshold = 0.016))
+
+swissroll_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r swissroll, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `swissroll` data in $3\\text{-}D$."}
+
+swissrollfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    swissroll_lang
+  ),
+  device = "xs"
+)
+
+class(swissrollfig) <- c(class(swissrollfig), "htmlwidget")
+
+swissrollfig
+```
+
+```{r swissroll-proj, eval=knitr::is_latex_output(), fig.pos="H", fig.cap="Three $2\\text{-}D$ projections from $3\\text{-}D$, for the `swissroll` data."}
+#| fig-width: 12
+#| fig-height: 4
+
+swissroll_proj1 + swissroll_proj2 + swissroll_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 #### Trefoil knots

### 2025-10-02 paper-cardinalR.tex
Added Swissroll section

@@
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.}\label{fig:pyr-proj}
+\end{figure}
+
+\paragraph{Polynomial}\label{polynomial}
+
+A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in $2\text{-}D${} space.
+To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
+These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
+We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
+Table \ref{tab:polynomial-tb-pdf} summarizes these functions and their purposes.
+
+\begin{table}
+
+\caption{\label{tab:polynomial-tb-pdf}cardinalR polynomial data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_quadratic & Quadratic pattern.\\
+gen\_cubic & Cubic pattern.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The first is the quadratic curve of \(n\) points in two dimensions.
+This is generated using \texttt{gen\_quadratic(n,\ range)}.
+The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree 2 is applied to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
+This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
+
+\begin{verbatim}
+quadratic <- gen_quadratic(n = 1000)
+\end{verbatim}
+
+The second is the cubic curve of \(n\) points in two dimensions.
+This is generated using \texttt{gen\_cubic(n,\ range)}.
+The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree \(3\) is applied to construct \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
+This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
+
+\begin{verbatim}
+cubic <- gen_cubic(n = 1000)
+\end{verbatim}
+
+\paragraph{Swiss Roll}\label{swiss-roll}
+
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension (Figure \ref{fig:swissroll-proj}).
+The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.
+The \[X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \text{ where } t \sim U(0, 3\pi).\]
+
+\begin{verbatim}
+swissroll <- gen_swissroll(n = 1000, w = c(-1, 1))
+\end{verbatim}
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/swissroll-proj-1} \caption{Three $2\text{-}D$ projections from $3\text{-}D$, for the `swissroll` data.}\label{fig:swissroll-proj}
 \end{figure}

### 2025-10-02 paper-cardinalR.R

@@ -871,6 +871,146 @@ pyrholes_proj1 + pyrholes_proj2 + pyrholes_proj3 +
 
 
 ## -----------------------------------------------------------------------------
+polynomial_tb <- tibble(fun = c("gen_quadratic",
+                                "gen_cubic"), 
+                        exp = c("Quadratic pattern.",
+                                "Cubic pattern."))
+
+
+## ----polynomial-tb-html, eval=knitr::is_html_output()-------------------------
+# polynomial_tb |>
+#   kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation"))
+
+
+## ----polynomial-tb-pdf, eval=knitr::is_latex_output()-------------------------
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+
+
+## ----data-quadratic, echo=TRUE------------------------------------------------
+quadratic <- gen_quadratic(n = 1000)
+
+
+## ----data-cubic, echo=TRUE----------------------------------------------------
+cubic <- gen_cubic(n = 1000)
+
+
+## ----data-swissroll, echo=TRUE------------------------------------------------
+swissroll <- gen_swissroll(n = 1000, w = c(-1, 1))
+
+
+## -----------------------------------------------------------------------------
+swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+
+
+## ----swissroll-proj1----------------------------------------------------------
+scaled_data <- scale_data_manual(swissroll)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097),
+  c(0.00896,0.52496,0.12597))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+swissroll_proj1 <- plot_proj(
[…]

### 2025-10-02 paper-cardinalR.Rmd
Added S-curve

@@
 
+#### Cube
+
+A cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube)"} else { "Figure \\@ref(fig:cube-proj)"})`) represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
+These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:cube-tb-html)', '\\@ref(tab:cube-tb-pdf)'))` outlines these functions and their purposes.
+
+The first is the regular grid of points of $n$ points in $p$ dimensions.
+This is generated using `gen_gridcube(n, p)`.
+The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to $n$.
+Each dimension is then normalized to lie in the interval $[0,1]$, so that the resulting structure forms a true $p\text{-}D$ hypercube.
+This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
+
+```{r data-gridcube, echo=TRUE}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+```
+
+```{r}
+gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r gridcube-proj1}
+scaled_data <- scale_data_manual(gridcube)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
+                                              threshold = 0.022))
+
+gridcube_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.83, 0.8), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridcube-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.8,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
+                                              threshold = 0.05))
+
+gridcube_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-1, 1), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridcube-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
+                                              threshold = 0.016))
+
+gridcube_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.8, 0.8), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
+The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) b"} else { "Figure \\@ref(fig:cube-proj) b"})`).
+
+```{r data-unifcube, echo=TRUE}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+```
+
+```{r}
+unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r unifcube-proj1}
+scaled_data <- scale_data_manual(unifcube)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.022))
+
+unifcube_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.9, 0.8), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifcube-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.04))
+
+unifcube_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.9, 0.8), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifcube-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
+                                              threshold = 0.016))
+
+unifcube_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.8, 0.8), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r}
+cube_tb <- tibble(fun = c("gen_gridcube",
+                          "gen_unifcube"), 
+                      exp = c("Cube with specified grid points along each axes.",
+                              "Cube with uniform points."))
+```
+
+```{r cube-tb-html, eval=knitr::is_html_output()}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r cube-tb-pdf, eval=knitr::is_latex_output()}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+```{r cube, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gridcube`, `unifcube`, and `cubehole` data in $4\\text{-}D$."}
+
+cubefig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    gridcube_lang, unifcube_lang
+  ),
+  device = "xs"
+)
+
+class(cubefig) <- c(class(cubefig), "htmlwidget")
+
+cubefig
+```
+
+```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=12, fig.height=8}
+
+gridcube_proj1 + gridcube_proj2 + gridcube_proj3 +
+unifcube_proj1 + unifcube_proj2 + unifcube_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 #### Cone
@@
 cubic <- gen_cubic(n = 1000)
+```
+
+#### S-curve
+
+An S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) simulates data that lies along a smooth, non-linear manifold.
+
+For the S-curve structure, `gen_scurve(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`), the $3\text{-}D$ geometry is constructed by introducing a latent parameter, $\theta \sim U\left(-3\pi/2, 3\pi/2\right)$.
+This parameter controls the curvature of the manifold.
+The first three dimensions form the S-curve structure: $$X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1)$$.
+This configuration creates a horizontally curled shape in $(X_1, X_3)$, with additional band thickness in the $X_2$ direction.
+
+```{r data-scurve, echo=TRUE}
+scurve <- gen_scurve(n = 1000)
+```
+
+```{r}
+scurve_lang <- langevitour::langevitour(scurve, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r scurve-proj1}
+scaled_data <- scale_data_manual(scurve)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097),
+  c(0.00896,0.52496,0.12597))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+scurve_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r scurve-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278),
+  c(-0.50493,0.14388,0.14615))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+scurve_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r scurve-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416),
+  c(0.50329,0.07199,-0.16064))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.016))
+
+scurve_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.62), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r scurve, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `scurve` and `scurvehole` data in $4\\text{-}D$."}
+
+scurvefig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    scurve_lang
+  ),
+  device = "xs"
+)
+
+class(scurvefig) <- c(class(scurvefig), "htmlwidget")
+
+scurvefig
+```
+
+```{r scurve-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.", fig.width=12, fig.height=4}
+
+scurve_proj1 + scurve_proj2 + scurve_proj3 + 
+  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-10-02 paper-cardinalR.tex
Added S-curve

@@
 All shape generators allow the user to define the number of points (\texttt{n}) and dimensions (\texttt{p}), and most include additional arguments to customize specific characteristics of the structure.
+
+\paragraph{Cube}\label{cube}
+
+A cube structure (Figure \ref{fig:cube-proj}) represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
+These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
+Table \ref{tab:cube-tb-pdf} outlines these functions and their purposes.
+
+The first is the regular grid of points of \(n\) points in \(p\) dimensions.
+This is generated using \texttt{gen\_gridcube(n,\ p)}.
+The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to \(n\).
+Each dimension is then normalized to lie in the interval \([0,1]\), so that the resulting structure forms a true \(p\text{-}D\) hypercube.
+This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
+
+\begin{verbatim}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+\end{verbatim}
+
+An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
+The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube (Figure \ref{fig:cube-proj} b).
+
+\begin{verbatim}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+\end{verbatim}
+
+\begin{table}
+
+\caption{\label{tab:cube-tb-pdf}cardinalR cube data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_gridcube & Cube with specified grid points along each axes.\\
+gen\_unifcube & Cube with uniform points.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.}\label{fig:cube-proj}
+\end{figure}
 
@@
 \end{verbatim}
+
+\paragraph{S-curve}\label{s-curve}
+
+An S-curve structure (Figure \ref{fig:scurve-proj}) simulates data that lies along a smooth, non-linear manifold.
+
+For the S-curve structure, \texttt{gen\_scurve(n,\ p)} (Figure \ref{fig:scurve-proj} a), the \(3\text{-}D\) geometry is constructed by introducing a latent parameter, \(\theta \sim U\left(-3\pi/2, 3\pi/2\right)\).
+This parameter controls the curvature of the manifold.
+The first three dimensions form the S-curve structure: \[X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1)\].
+This configuration creates a horizontally curled shape in \((X_1, X_3)\), with additional band thickness in the \(X_2\) direction.
+
+\begin{verbatim}
+scurve <- gen_scurve(n = 1000)
+\end{verbatim}
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.}\label{fig:scurve-proj}
+\end{figure}
 

### 2025-10-02 paper-cardinalR.R

@@ -124,6 +124,227 @@ main_tb |>
   column_spec(3, width = "8cm")
 
 
+## ----data-gridcube, echo=TRUE-------------------------------------------------
+gridcube <- gen_gridcube(n = 1000, p = 4)
+
+
+## -----------------------------------------------------------------------------
+gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+
+
+## ----gridcube-proj1-----------------------------------------------------------
+scaled_data <- scale_data_manual(gridcube)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.7, 
+                                              axis_pos_y = -0.7, 
+                                              threshold = 0.022))
+
+gridcube_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.83, 0.8), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+
+
+## ----gridcube-proj2-----------------------------------------------------------
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.8,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
+                                              threshold = 0.05))
+
+gridcube_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-1, 1), 
+  title = "a2", 
+  cex = 2, 
[…]

### 2025-10-03 paper-cardinalR.Rmd
Added all the structues except Mobius

@@
 
+#### Branching
+
+A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (@trapnell2014).
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (`n`), and number of branches (`k`).
+Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:branching-tb-html)', '\\@ref(tab:branching-tb-pdf)'))` outlines these functions.
+The main arguments of the functions described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:arg-branching-tb-html)', '\\@ref(tab:arg-branching-tb-pdf)'))`.
+
+```{r}
+branch_tb <- tibble(fun = c("gen_expbranches",
+                             "gen_linearbranches",
+                             "gen_curvybranches", 
+                             "gen_orglinearbranches", 
+                             "gen_orgcurvybranches"), 
+                      exp = c("Exponential shaped branches.",
+                               "Linear shaped branches.", 
+                               "Curvy shaped branches.",
+                               "Linear shaped branches originated in one point.",
+                               "Curvy shaped branches originated in one point."))
+```
+
+```{r branching-tb-html, eval=knitr::is_html_output()}
+branch_tb |> 
+  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r branching-tb-pdf, eval=knitr::is_latex_output()}
+branch_tb |> 
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+```{r}
+arg_branching_tb <- tibble(arg = c("n",
+                          "p",
+                          "k"), 
+                        exp = c("A numeric value representing the number of points.",
+                                "A numeric value representing the number of dimensions.",
+                                "A numeric value representing the number of clusters."))
+```
+
+```{r arg-branching-tb-html, eval=knitr::is_html_output()}
+arg_branching_tb |> 
+  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Explanation")) 
+```
+
+```{r arg-branching-tb-pdf, eval=knitr::is_latex_output()}
+arg_branching_tb |> 
+  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+The simplest structures are approximately linear branches, generated by the `gen_linearbranches(n, p, k)` function (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) b"} else { "Figure \\@ref(fig:branch-proj) b"})`).
+These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability.
+Mathematically, each branch $i$ is defined as
+
+$$
+X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
+$$
+
+where $(x_{\text{start},i}, y_{\text{start},i})$ is the starting point of branch $i$, $\delta$ controls local jitter, and $s_i$ is the slope, initialized as
+
+$$
+s_i =
+\begin{cases}
+0.5 & i = 1, \\
+-0.5 & i = 2, \\
+\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
+\end{cases}
+$$
+
+Branches $1$ and $2$ are initialized with fixed slopes and intercepts, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
+
+```{r data-linearbranches, echo=TRUE}
+linearbranches <- gen_linearbranches(n = 1000, k = 4)
+```
+
+To introduce curvature, the `gen_curvybranches(n, p, k)` function generates $k$ curvilinear branches in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) c"} else { "Figure \\@ref(fig:branch-proj) c"})`).
+Branches $1$ and $2$ are simple parabolas defined as
+
+$$
+\begin{aligned}
+\text{Branch 1: } & X_1 \sim U(0,1), \quad X_2 = 0.1 X_1 + X_1^2 + \epsilon, \\
+\text{Branch 2: } & X_1 \sim U(-1,0), \quad X_2 = 0.1 X_1 - 2 X_1^2 + \epsilon, \quad \epsilon \sim U(0, \delta),
+\end{aligned}
+$$
+
+where $\delta$ controls local jitter.
+Additional branches are attached iteratively to existing structures.
+Each new branch $i$ starts at a selected point $(x_{\text{start},i}, y_{\text{start},i})$ from the current structure and extends according to
+
+$$
+X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
+$$
+
+where $s_i$ is a scale factor controlling the curvature of branch $i$.
+For the first few initial branches, $s_i$ can be fixed (e.g., $s_1 = 1, s_2 = 2$), while for subsequent branches it is sampled from a predefined set, such as $s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}$, to create variability in curvature.
+
+```{r data-curvybranches, echo=TRUE}
+curvybranches <- gen_curvybranches(n = 1000, k = 4)
+```
+
+The `gen_expbranches(n, p, k)` function creates $k$ exponential branches in $2\text{-}D$, radiating from a central region (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) a"} else { "Figure \\@ref(fig:branch-proj) a"})`).
+Each branch $i$ is defined as
+
+$$
+X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
+$$
+
+where $\sigma_i = (-1)^{i+1}$ alternates the sign of the exponent to produce mirror-symmetric branches.
+The parameter $s_i$ controls the steepness of branch $i$, and $\delta$ introduces small local jitter.
+
+```{r data-expbranches, echo=TRUE}
+expbranches <- gen_expbranches(n = 1000, k = 4)
+```
+
+High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` and `gen_orgcurvybranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) d-e"} else { "Figure \\@ref(fig:branch-proj) d-e"})`).
+Each branch is embedded in a unique or repeated $2\text{-}D$ subspace of the $p\text{-}D$ space.
+When `allow_share = TRUE`, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible $\binom{p}{2}$ combinations are exhausted, after which additional branches may repeat subspaces.
+Linear branches follow
+
+$$
+X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = s_i X_{i_1} + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+$$
+
+while curvilinear branches include a quadratic term
+
+$$
+X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = - s_i X_{i_1}^2 + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+$$
+
+where $a_i, b_i$ define the range of the first coordinate for branch $i$, and $\epsilon$ is Gaussian noise added to introduce variability.
+The scale factor $s_i$ controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first $\binom{p}{2}$ branches, $s_i = 1$; for additional branches when $k > \binom{p}{2}$, $s_i$ is randomly drawn from the set $\{1, 1.5, 2, \dots, 8\}$.
+
+```{r data-orglinearbranches, echo=TRUE}
+orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
+```
+
+```{r}
+orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r orglinearbranches-proj1}
+scaled_data <- scale_data_manual(orglinearbranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+orglinearbranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "d1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r orglinearbranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+orglinearbranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "d2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r orglinearbranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+orglinearbranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "d3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r data-orgcurvybranches, echo=TRUE}
+orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
+```
+
+```{r}
+orgcurvybranches_lang <- langevitour::langevitour(orgcurvybranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r orgcurvybranches-proj1}
+scaled_data <- scale_data_manual(orgcurvybranches)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.3, 
+                                              axis_pos_y = -0.3, 
+                                              threshold = 0.022))
+
+orgcurvybranches_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.4, 0.5), 
+  title = "e1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r orgcurvybranches-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+orgcurvybranches_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "e2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r orgcurvybranches-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+orgcurvybranches_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "e3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For all branch-generating functions, when the target dimensionality $p > 2$, additional dimensions are filled with independent noise, allowing the $2\text{-}D$ or $2\text{-}D$-subspace structures to be naturally embedded in higher-dimensional space while preserving the intended geometric patterns.
+
+```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, and `orgcurvybranches` data in $4\\text{-}D$."}
+
+branchfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    orglinearbranches_lang, orgcurvybranches_lang
+  ),
+  device = "xs"
+)
+
+class(branchfig) <- c(class(branchfig), "htmlwidget")
+
+branchfig
+```
+
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=12, fig.height=8}
+
+orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
+orgcurvybranches_proj1 + orgcurvybranches_proj2 + orgcurvybranches_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 #### Cube
@@
 
+#### Sphere
+
+Sphere-shaped structures (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere)"} else { "Figure \\@ref(fig:sphere-proj)"})`) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+The functions generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:sphere-tb-html)', '\\@ref(tab:sphere-tb-pdf)'))` summarizes these functions.
+
+```{r}
+sphere_tb <- tibble(fun = c("gen_circle",
+                            "gen_curvycycle",
+                            "gen_unifsphere",
+                            "gen_gridedsphere",
+                            "gen_clusteredspheres",
+                            "gen_hemisphere"), 
+                        exp = c("Circle.",
+                                "Curvy cell cycle.",
+                                "Uniform sphere.",
+                                "Grided sphere.",
+                                "Multiple small spheres within a big sphere.",
+                                "Hemisphere."))
+```
+
+```{r sphere-tb-html, eval=knitr::is_html_output()}
+sphere_tb |> 
+  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r sphere-tb-pdf, eval=knitr::is_latex_output()}
+sphere_tb |> 
+  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+The simplest case, `gen_circle(n)` creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
+Let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
+Coordinates in the first two dimensions represent a perfect circle on the plane: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta)$$.
+For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
+The first component is a scaling factor that decreases with the dimension index, defined as $\text{scale}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$.
+The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2)\pi/2p$.
+Each additional dimension is computed as: $X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p$.
+
+```{r data-circle, echo=TRUE}
+circle <- gen_circle(n = 1000, p = 4)
+```
+
+```{r}
+circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r circle-proj1}
+scaled_data <- scale_data_manual(circle)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.022))
+
+circle_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.8), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r circle-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+circle_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.7), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r circle-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.016))
+
+circle_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.53), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For the one-dimensional nonlinear cycle embedded in $p\text{-}D$ space, `gen_curvycycle(n)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`), let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
+The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle".
+In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin.
+Additionally, $X_3 = 1/3\cos(3\theta)$ introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
+
+Together, these define a periodic, non-trivial, closed curve in $3\text{-}D$ with internal folds that produce a more complex geometry than a standard circle or ellipse.
+For dimensions $X_4$ through $X_p$, additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
+The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from $4$ to $p$, which means that the amplitude decreases as the dimension increases.
+Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j\sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2)\pi/2p$.
+
+```{r data-curvycycle, echo=TRUE}
+curvycycle <- gen_curvycycle(n = 1000, p = 4)
+```
+
+```{r}
+curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r curvycycle-proj1}
+scaled_data <- scale_data_manual(curvycycle)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.022))
+
+curvycycle_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.7), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycycle-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.04))
+
+curvycycle_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycycle-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+curvycycle_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+Building on simple circular structures, the `gen_unifsphere(n, r)` function function extends the idea to three dimensions by generating $n$ observations approximately uniformly distributed on the surface of a sphere of radius $r$, with optional high-dimensional noise when $p > 3$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) c"} else { "Figure \\@ref(fig:sphere-proj) c"})`).
+Each observation is computed from spherical coordinates, with $u \sim U(-1, 1)$ representing $\cos(\phi)$ and $\theta \sim U(0, 2\pi)$ the azimuthal angle.
+Cartesian coordinates are then defined as $$X_1 = r\sqrt{1 - u^2}\cos(\theta), \quad X_2 = r\sqrt{1 - u^2}\sin(\theta),\text{ and }X_3 = ru,$$ ensuring uniform distribution on the surface (not within) of the sphere.
+
+```{r data-unifsphere, echo=TRUE}
+unifsphere <- gen_unifsphere(n = 1000, r = 1)
+```
+
+```{r}
+unifsphere_lang <- langevitour::langevitour(unifsphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r unifsphere-proj1}
+scaled_data <- scale_data_manual(unifsphere)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097),
+  c(0.00896,0.52496,0.12597))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.022))
+
+unifsphere_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.7), 
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifsphere-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278),
+  c(-0.50493,0.14388,0.14615))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+unifsphere_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.7), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r unifsphere-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416),
+  c(0.50329,0.07199,-0.16064))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.016))
+
+unifsphere_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.65), 
+  title = "c3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+In addition, the `gen_gridedsphere(n)` function constructs a $p\text{-}D$ dataset consisting of approximately $n$ points that are evenly distributed on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
+The method relies on forming a regular grid in spherical coordinates, parameterized by $(p-1)$ angular variables: for dimensions $j = 1, \dots, p-2$ the polar angles are drawn from $[0, \pi]$, while the final angle ($j = p-1$) represents the azimuth and is drawn from $[0, 2\pi]$.
+The number of grid steps along each angular dimension is chosen by decomposing $n$ into $(p-1)$ approximately equal integer factors using the helper function `gen_nproduct(n, p - 1)`.
+
+Each grid point is subsequently mapped into Cartesian space via the standard hyperspherical-to-Cartesian transformation,
+
+$$
+\begin{aligned}
+X_1 &= \cos(\theta_1), \\
+X_2 &= \sin(\theta_1)\cos(\theta_2), \\
+X_3 &= \sin(\theta_1)\sin(\theta_2)\cos(\theta_3), \\
+&\;\;\vdots \\
+X_{p-1} &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\cos(\theta_{p-1}), \\
+X_p &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\sin(\theta_{p-1}).
+\end{aligned}
+$$
+
+The result is a deterministic grid of points lying exactly on the surface of the unit $(p-1)$-sphere, without any additional noise dimensions.
+
+```{r data-gridedsphere, echo=TRUE}
+gridedsphere <- gen_gridedsphere(n = 1000, p = 4)
+```
+
+```{r}
+gridedsphere_lang <- langevitour::langevitour(gridedsphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r gridedsphere-proj1}
+scaled_data <- scale_data_manual(gridedsphere)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.022))
+
+gridedsphere_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.7), 
+  title = "d1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridedsphere-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+gridedsphere_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.65), 
+  title = "d2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r gridedsphere-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.016))
+
+gridedsphere_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.62), 
+  title = "d3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For more heterogeneous structures, the `gen_clusteredspheres(n, k, r, loc)` function generates one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
+A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
+The sampling is executed using the function `gen_unifsphere(n_1, p, r_1)`, which generates the desired points in the specified dimensional space.
+In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
+These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
+Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for $3\text{-}D$).
+Each observation is classified by cluster, with labels such as "big" for the large central sphere and "small_1" to "small_k" for the smaller spheres.
+
+```{r data-clusteredspheres, echo=TRUE}
+clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, r = c(15, 3),
+                                         loc = 10 / sqrt(3)) |>
+  dplyr::select(-cluster)
+```
+
+```{r}
+clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r clusteredspheres-proj1}
+scaled_data <- scale_data_manual(clusteredspheres)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097),
+  c(0.00896,0.52496,0.12597))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.022))
+
+clusteredspheres_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.7), 
+  title = "e1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r clusteredspheres-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278),
+  c(-0.50493,0.14388,0.14615))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+clusteredspheres_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "e2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r clusteredspheres-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416),
+  c(0.50329,0.07199,-0.16064))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.016))
+
+clusteredspheres_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.62), 
+  title = "e3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+Finally, the `gen_hemisphere(n, p)` function restricts sampling to a hemisphere of a $4\text{-}D$ sphere (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) f"} else { "Figure \\@ref(fig:sphere-proj) f"})`).
+Using spherical coordinates, the azimuthal angle $\theta_1 \sim U(0, \pi)$ in the $(x_1, x_2)$ plane, while the elevation angle $\theta_2 \sim U(0, \pi)$ in the $(x_2, x_3)$ plane.
+Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
+The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
+
+```{r data-hemisphere, echo=TRUE}
+hemisphere <- gen_hemisphere(n = 1000)
+```
+
+```{r}
+hemisphere_lang <- langevitour::langevitour(hemisphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r hemisphere-proj1}
+scaled_data <- scale_data_manual(hemisphere)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.022))
+
+hemisphere_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.7), 
+  title = "f1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r hemisphere-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+hemisphere_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "f2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r hemisphere-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+hemisphere_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.75, 0.62), 
+  title = "f3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, `unifsphere`, `gridedsphere`, `clusteredspheres`, and `hemisphere` data in $4\\text{-}D$."}
+
+spherefig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    circle_lang, curvycycle_lang, unifsphere_lang, gridedsphere_lang, clusteredspheres_lang, hemisphere_lang
+  ),
+  device = "xs"
+)
+
+class(spherefig) <- c(class(spherefig), "htmlwidget")
+
+spherefig
+```
+
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=28, out.width="80%", fig.align='center'}
+
+circle_proj1 + circle_proj2 + circle_proj3 +
+curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3 +
+unifsphere_proj1 + unifsphere_proj2 + unifsphere_proj3 +
+gridedsphere_proj1 + gridedsphere_proj2 + gridedsphere_proj3 +
+clusteredspheres_proj1 + clusteredspheres_proj2 + clusteredspheres_proj3 +
+hemisphere_proj1 + hemisphere_proj2 + hemisphere_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
 #### Swiss Roll
@@
   trefoil3d_proj1 + trefoil3d_proj2 + trefoil3d_proj3 +
+  plot_layout(ncol = 3, guides = "collect") 
+```
+
+#### Trigonometric
+
+Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric)"} else { "Figure \\@ref(fig:triginometric-proj)"})`).
+The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when $p > 2$ (adding $X_3$ through $X_p$), and for all other trigonometric shapes when $p > 4$ (adding $X_5$ through $X_p$).
+These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trigonometric-tb-html)', '\\@ref(tab:trigonometric-tb-pdf)'))` summarizes these functions.
+
+```{r}
+trigonometric_tb <- tibble(fun = c("gen_crescent",
+                            "gen_curvycylinder",
+                            "gen_sphericalspiral",
+                            "gen_helicalspiral",
+                            "gen_conicspiral",
+                            "gen_nonlinear"), 
+                        exp = c("Crescent pattern.",
+                                "Curvy cylinder.",
+                                "Spherical spiral.",
+                                "Helical spiral.",
+                                "Conic spiral.",
+                                "Nonlinear hyperbola."))
+```
+
+```{r trigonometric-tb-html, eval=knitr::is_html_output()}
+trigonometric_tb |> 
+  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r trigonometric-tb-pdf, eval=knitr::is_latex_output()}
+trigonometric_tb |> 
+  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+First, the `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) a"} else { "Figure \\@ref(fig:triginometric-proj) a"})`).
+Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
+The corresponding $2\text{-}D$ coordinates are defined by: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$
+
+```{r data-crescent, echo=TRUE}
+crescent <- gen_crescent(n = 1000)
+```
+
+Second, the `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
+The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
+Let $\theta \sim U(0, 3\pi)$ represent a random angle on a circular base and $z \sim U(0, h)$ represent the height along the cylinder.
+The coordinates are defined as: $X_1 = \cos(\theta)$ (Circular base, x-axis), $X_2 = \sin(\theta)$ (Circular base, y-axis), $X_3 = z$ (Linear height), and $X_4 = \sin(z)$ (Nonlinear curvy variation along height).
+
+```{r data-curvycylinder, echo=TRUE}
+curvycylinder <- gen_curvycylinder(n = 1000, h = 10)
+```
+
+```{r}
+curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r curvycylinder-proj1}
+scaled_data <- scale_data_manual(curvycylinder)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+curvycylinder_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycylinder-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+curvycylinder_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycylinder-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+curvycylinder_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.5), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$ be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
+Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
+This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
+
+```{r data-sphericalspiral, echo=TRUE}
+sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
+```
+
+```{r}
+sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r sphericalspiral-proj1}
+scaled_data <- scale_data_manual(sphericalspiral)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+sphericalspiral_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r sphericalspiral-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+sphericalspiral_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.6), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r sphericalspiral-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+sphericalspiral_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.5), 
+  title = "c3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For a helical spiral in four dimensions, `gen_helicalspiral(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) d"} else { "Figure \\@ref(fig:triginometric-proj) d"})`), let $\theta \in [0, 5\pi/4]$ be a sequence of angles controlling rotation around a circle.
+Cartesian coordinates; $X_1 = \cos(\theta)$: circular trajectory along the x-axis, $X_2 = \sin(\theta)$: circular trajectory along the y-axis, $X_3 = 0.05\theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix, and $X_4 = 0.1\sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
+
+```{r data-helicalspiral, echo=TRUE}
+helicalspiral <- gen_helicalspiral(n = 1000)
+```
+
+Similarly, the `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) e"} else { "Figure \\@ref(fig:triginometric-proj) e"})`).
+The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a $3\text{-}D$ conic helix.
+The shape is defined by parameter $\theta \in [0, 2\pi\text{spins}]$, controlling the angular progression of the spiral.
+The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta\cos(\theta)$ for radial expansion in x, and $X_2 = \theta\sin(\theta)$ for radial expansion in y.
+The growth pattern resembles a cone, with the height increasing according to $X_3 = 2\theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta\sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
+
+```{r data-conicspiral, echo=TRUE}
+conicspiral <- gen_conicspiral(n = 1000, spins = 1)
+```
+
+```{r}
+conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r conicspiral-proj1}
+scaled_data <- scale_data_manual(conicspiral)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+conicspiral_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "e1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r conicspiral-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
+                                              threshold = 0.01))
+
+conicspiral_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.8, 0.5), 
+  title = "e2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r conicspiral-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
+                                              threshold = 0.016))
+
+conicspiral_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.62), 
+  title = "e3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+Finally, the `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
+The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
+
+```{r data-nonlinear, echo=TRUE}
+nonlinear <- gen_nonlinear(n = 1000, hc = 1, non_fac = 0.5)
+```
+
+```{r}
+nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r nonlinear-proj1}
+scaled_data <- scale_data_manual(nonlinear)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.3, 
+                                              axis_pos_y = -0.3, 
+                                              threshold = 0.022))
+
+nonlinear_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.45, 0.73), 
+  title = "f1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r nonlinear-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.45,
+                                              axis_scaled = 1.3, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.01))
+
+nonlinear_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.46), 
+  title = "f2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r nonlinear-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.45,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.016))
+
+nonlinear_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.35, 0.4), 
+  title = "f3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r triginometric, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of tthe `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data in $4\\text{-}D$."}
+
+triginometricfig <- bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    curvycylinder_lang, sphericalspiral_lang, conicspiral_lang, nonlinear_lang
+  ),
+  device = "xs"
+)
+
+class(triginometricfig) <- c(class(triginometricfig), "htmlwidget")
+
+triginometricfig
+```
+
+```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=16, out.width="80%", fig.align='center'}
+
+curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
+sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +
+conicspiral_proj1 + conicspiral_proj2 + conicspiral_proj3 +
+nonlinear_proj1 + nonlinear_proj2 + nonlinear_proj3 +
   plot_layout(ncol = 3, guides = "collect") 

### 2025-10-03 paper-cardinalR.tex
Added all the structues except Mobius

@@
 All shape generators allow the user to define the number of points (\texttt{n}) and dimensions (\texttt{p}), and most include additional arguments to customize specific characteristics of the structure.
+
+\paragraph{Branching}\label{branching}
+
+A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (\citet{trapnell2014}).
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (\texttt{n}), and number of branches (\texttt{k}).
+Although these functions can generate multiple branches, they do not produce a formal \emph{multicluster} dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
+Table \ref{tab:branching-tb-pdf} outlines these functions.
+The main arguments of the functions described in Table \ref{tab:arg-branching-tb-pdf}.
+
+\begin{table}
+
+\caption{\label{tab:branching-tb-pdf}cardinalR branching data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_expbranches & Exponential shaped branches.\\
+gen\_linearbranches & Linear shaped branches.\\
+gen\_curvybranches & Curvy shaped branches.\\
+gen\_orglinearbranches & Linear shaped branches originated in one point.\\
+gen\_orgcurvybranches & Curvy shaped branches originated in one point.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:arg-branching-tb-pdf}The main arguments for branching shape generators.}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Argument & Explanation\\
+\midrule
+n & A numeric value representing the number of points.\\
+p & A numeric value representing the number of dimensions.\\
+k & A numeric value representing the number of clusters.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The simplest structures are approximately linear branches, generated by the \texttt{gen\_linearbranches(n,\ p,\ k)} function (Figure \ref{fig:branch-proj} b).
+These consist of \(k\) short line segments in the first two dimensions, with added jitter to simulate variability.
+Mathematically, each branch \(i\) is defined as
+
+\[
+X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
+\]
+
+where \((x_{\text{start},i}, y_{\text{start},i})\) is the starting point of branch \(i\), \(\delta\) controls local jitter, and \(s_i\) is the slope, initialized as
+
+\[
+s_i =
+\begin{cases}
+0.5 & i = 1, \\
+-0.5 & i = 2, \\
+\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
+\end{cases}
+\]
+
+Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
+
+\begin{verbatim}
+linearbranches <- gen_linearbranches(n = 1000, k = 4)
+\end{verbatim}
+
+To introduce curvature, the \texttt{gen\_curvybranches(n,\ p,\ k)} function generates \(k\) curvilinear branches in \(p\text{-}D\) (Figure \ref{fig:branch-proj} c).
+Branches \(1\) and \(2\) are simple parabolas defined as
+
+\[
+\begin{aligned}
+\text{Branch 1: } & X_1 \sim U(0,1), \quad X_2 = 0.1 X_1 + X_1^2 + \epsilon, \\
+\text{Branch 2: } & X_1 \sim U(-1,0), \quad X_2 = 0.1 X_1 - 2 X_1^2 + \epsilon, \quad \epsilon \sim U(0, \delta),
+\end{aligned}
+\]
+
+where \(\delta\) controls local jitter.
+Additional branches are attached iteratively to existing structures.
+Each new branch \(i\) starts at a selected point \((x_{\text{start},i}, y_{\text{start},i})\) from the current structure and extends according to
+
+\[
+X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
+\]
+
+where \(s_i\) is a scale factor controlling the curvature of branch \(i\).
+For the first few initial branches, \(s_i\) can be fixed (e.g., \(s_1 = 1, s_2 = 2\)), while for subsequent branches it is sampled from a predefined set, such as \(s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}\), to create variability in curvature.
+
+\begin{verbatim}
+curvybranches <- gen_curvybranches(n = 1000, k = 4)
+\end{verbatim}
+
+The \texttt{gen\_expbranches(n,\ p,\ k)} function creates \(k\) exponential branches in \(2\text{-}D\), radiating from a central region (Figure \ref{fig:branch-proj} a).
+Each branch \(i\) is defined as
+
+\[
+X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
+\]
+
+where \(\sigma_i = (-1)^{i+1}\) alternates the sign of the exponent to produce mirror-symmetric branches.
+The parameter \(s_i\) controls the steepness of branch \(i\), and \(\delta\) introduces small local jitter.
+
+\begin{verbatim}
+expbranches <- gen_expbranches(n = 1000, k = 4)
+\end{verbatim}
+
+High-dimensional generalizations are provided by \texttt{gen\_orglinearbranches(n,\ p,\ k)} and \texttt{gen\_orgcurvybranches(n,\ p,\ k)} (Figure \ref{fig:branch-proj} d-e).
+Each branch is embedded in a unique or repeated \(2\text{-}D\) subspace of the \(p\text{-}D\) space.
+When \texttt{allow\_share\ =\ TRUE}, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible \(\binom{p}{2}\) combinations are exhausted, after which additional branches may repeat subspaces.
+Linear branches follow
+
+\[
+X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = s_i X_{i_1} + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+\]
+
+while curvilinear branches include a quadratic term
+
+\[
+X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = - s_i X_{i_1}^2 + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+\]
+
+where \(a_i, b_i\) define the range of the first coordinate for branch \(i\), and \(\epsilon\) is Gaussian noise added to introduce variability.
+The scale factor \(s_i\) controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first \(\binom{p}{2}\) branches, \(s_i = 1\); for additional branches when \(k > \binom{p}{2}\), \(s_i\) is randomly drawn from the set \(\{1, 1.5, 2, \dots, 8\}\).
+
+\begin{verbatim}
+orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+\begin{verbatim}
+orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
+\end{verbatim}
+
+For all branch-generating functions, when the target dimensionality \(p > 2\), additional dimensions are filled with independent noise, allowing the \(2\text{-}D\) or \(2\text{-}D\)-subspace structures to be naturally embedded in higher-dimensional space while preserving the intended geometric patterns.
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.}\label{fig:branch-proj}
+\end{figure}
 
@@
 
+\paragraph{Sphere}\label{sphere}
+
+Sphere-shaped structures (Figure \ref{fig:sphere-proj}) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+The functions generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
+Table \ref{tab:sphere-tb-pdf} summarizes these functions.
+
+\begin{table}
+
+\caption{\label{tab:sphere-tb-pdf}cardinalR sphere data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_circle & Circle.\\
+gen\_curvycycle & Curvy cell cycle.\\
+gen\_unifsphere & Uniform sphere.\\
+gen\_gridedsphere & Grided sphere.\\
+gen\_clusteredspheres & Multiple small spheres within a big sphere.\\
+\addlinespace
+gen\_hemisphere & Hemisphere.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The simplest case, \texttt{gen\_circle(n)} creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (Figure \ref{fig:sphere-proj} a).
+Let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
+Coordinates in the first two dimensions represent a perfect circle on the plane: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta)\].
+For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
+The first component is a scaling factor that decreases with the dimension index, defined as \(\text{scale}_j = \sqrt{(0.5)^{j-2}}\) for \(j = 3, \dots, p\).
+The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula \(\phi_j = (j - 2)\pi/2p\).
+Each additional dimension is computed as: \(X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p\).
+
+\begin{verbatim}
+circle <- gen_circle(n = 1000, p = 4)
+\end{verbatim}
+
+For the one-dimensional nonlinear cycle embedded in \(p\text{-}D\) space, \texttt{gen\_curvycycle(n)} (Figure \ref{fig:sphere-proj} b), let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
+The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle''.
+In this configuration, \(X_1 = \cos(\theta)\) represents horizontal oscillation, while \(X_2 = \sqrt{3}/3 + \sin(\theta)\) introduces a vertical offset to avoid centering the curve at the origin.
+Additionally, \(X_3 = 1/3\cos(3\theta)\) introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
+
+Together, these define a periodic, non-trivial, closed curve in \(3\text{-}D\) with internal folds that produce a more complex geometry than a standard circle or ellipse.
+For dimensions \(X_4\) through \(X_p\), additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
+The scaling factor is defined as \(\text{scale}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from \(4\) to \(p\), which means that the amplitude decreases as the dimension increases.
+Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j\sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2)\pi/2p\).
+
+\begin{verbatim}
+curvycycle <- gen_curvycycle(n = 1000, p = 4)
+\end{verbatim}
+
+Building on simple circular structures, the \texttt{gen\_unifsphere(n,\ r)} function function extends the idea to three dimensions by generating \(n\) observations approximately uniformly distributed on the surface of a sphere of radius \(r\), with optional high-dimensional noise when \(p > 3\) (Figure \ref{fig:sphere-proj} c).
+Each observation is computed from spherical coordinates, with \(u \sim U(-1, 1)\) representing \(\cos(\phi)\) and \(\theta \sim U(0, 2\pi)\) the azimuthal angle.
+Cartesian coordinates are then defined as \[X_1 = r\sqrt{1 - u^2}\cos(\theta), \quad X_2 = r\sqrt{1 - u^2}\sin(\theta),\text{ and }X_3 = ru,\] ensuring uniform distribution on the surface (not within) of the sphere.
+
+\begin{verbatim}
+unifsphere <- gen_unifsphere(n = 1000, r = 1)
+\end{verbatim}
+
+In addition, the \texttt{gen\_gridedsphere(n)} function constructs a \(p\text{-}D\) dataset consisting of approximately \(n\) points that are evenly distributed on the surface of the unit \((p-1)\)-sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
+The method relies on forming a regular grid in spherical coordinates, parameterized by \((p-1)\) angular variables: for dimensions \(j = 1, \dots, p-2\) the polar angles are drawn from \([0, \pi]\), while the final angle (\(j = p-1\)) represents the azimuth and is drawn from \([0, 2\pi]\).
+The number of grid steps along each angular dimension is chosen by decomposing \(n\) into \((p-1)\) approximately equal integer factors using the helper function \texttt{gen\_nproduct(n,\ p\ -\ 1)}.
+
+Each grid point is subsequently mapped into Cartesian space via the standard hyperspherical-to-Cartesian transformation,
+
+\[
+\begin{aligned}
+X_1 &= \cos(\theta_1), \\
+X_2 &= \sin(\theta_1)\cos(\theta_2), \\
+X_3 &= \sin(\theta_1)\sin(\theta_2)\cos(\theta_3), \\
+&\;\;\vdots \\
+X_{p-1} &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\cos(\theta_{p-1}), \\
+X_p &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\sin(\theta_{p-1}).
+\end{aligned}
+\]
+
+The result is a deterministic grid of points lying exactly on the surface of the unit \((p-1)\)-sphere, without any additional noise dimensions.
+
+\begin{verbatim}
+gridedsphere <- gen_gridedsphere(n = 1000, p = 4)
+\end{verbatim}
+
+For more heterogeneous structures, the \texttt{gen\_clusteredspheres(n,\ k,\ r,\ loc)} function generates one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location (Figure \ref{fig:sphere-proj} e).
+A large uniform sphere centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\text{-}D\) sphere with a radius of \(r_1\).
+The sampling is executed using the function \texttt{gen\_unifsphere(n\_1,\ p,\ r\_1)}, which generates the desired points in the specified dimensional space.
+In generation of \(k\) smaller uniform spheres, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
+These spheres are positioned at distinct random locations in \(p\)-space, with the center of each sphere being drawn from a normal distribution \(N(0, \texttt{loc}^2 I_p)\).
+Points on spheres are generated using the standard hyperspherical method, which involves sampling \(u \sim U(-1, 1)\) to determine the cosine of the polar angle, and sampling \(\theta \sim U(0, 2\pi)\) to determine the azimuthal angle (for \(3\text{-}D\)).
+Each observation is classified by cluster, with labels such as ``big'' for the large central sphere and ``small\_1'' to ``small\_k'' for the smaller spheres.
+
+\begin{verbatim}
+clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, r = c(15, 3),
+                                         loc = 10 / sqrt(3)) |>
+  dplyr::select(-cluster)
+\end{verbatim}
+
+Finally, the \texttt{gen\_hemisphere(n,\ p)} function restricts sampling to a hemisphere of a \(4\text{-}D\) sphere (Figure \ref{fig:sphere-proj} f).
+Using spherical coordinates, the azimuthal angle \(\theta_1 \sim U(0, \pi)\) in the \((x_1, x_2)\) plane, while the elevation angle \(\theta_2 \sim U(0, \pi)\) in the \((x_2, x_3)\) plane.
+Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
+The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
+
+\begin{verbatim}
+hemisphere <- gen_hemisphere(n = 1000)
+\end{verbatim}
+
+\begin{figure}[!]
+
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}
+
+}
+
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.}\label{fig:sphere-proj}
+\end{figure}
+
 \paragraph{Swiss Roll}\label{swiss-roll}
@@
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.}\label{fig:trefoil-proj}
+\end{figure}
+
+\paragraph{Trigonometric}\label{trigonometric}
+
+Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (Figure \ref{fig:triginometric-proj}).
+The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when \(p > 2\) (adding \(X_3\) through \(X_p\)), and for all other trigonometric shapes when \(p > 4\) (adding \(X_5\) through \(X_p\)).
+These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
+Table \ref{tab:trigonometric-tb-pdf} summarizes these functions.
+
+\begin{table}
+
+\caption{\label{tab:trigonometric-tb-pdf}cardinalR trigonometric data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_crescent & Crescent pattern.\\
+gen\_curvycylinder & Curvy cylinder.\\
+gen\_sphericalspiral & Spherical spiral.\\
+gen\_helicalspiral & Helical spiral.\\
+gen\_conicspiral & Conic spiral.\\
+\addlinespace
+gen\_nonlinear & Nonlinear hyperbola.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+First, the \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise (Figure \ref{fig:triginometric-proj} a).
+Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
+The corresponding \(2\text{-}D\) coordinates are defined by: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta).\]
+
+\begin{verbatim}
+crescent <- gen_crescent(n = 1000)
+\end{verbatim}
+
+Second, the \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
+The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
+Let \(\theta \sim U(0, 3\pi)\) represent a random angle on a circular base and \(z \sim U(0, h)\) represent the height along the cylinder.
+The coordinates are defined as: \(X_1 = \cos(\theta)\) (Circular base, x-axis), \(X_2 = \sin(\theta)\) (Circular base, y-axis), \(X_3 = z\) (Linear height), and \(X_4 = \sin(z)\) (Nonlinear curvy variation along height).
+
+\begin{verbatim}
+curvycylinder <- gen_curvycylinder(n = 1000, h = 10)
+\end{verbatim}
+
+For a spiraling path on a spherical surface in the first four dimensions, \texttt{gen\_sphericalspiral(n,\ p,\ spins)} (Figure \ref{fig:triginometric-proj} c), let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\)be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
+Cartesian coordinates from spherical conversion: \(X_1 = \sin(\phi)\cos(\theta)\), \(X_2 = \sin(\phi)\sin(\theta)\), \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter, and \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
+This generates a spherical spiral curve embedded in \(4\text{-}D\) space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
+
+\begin{verbatim}
+sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
+\end{verbatim}
+
+For a helical spiral in four dimensions, \texttt{gen\_helicalspiral(n,\ p)} (Figure \ref{fig:triginometric-proj} d), let \(\theta \in [0, 5\pi/4]\) be a sequence of angles controlling rotation around a circle.
+Cartesian coordinates; \(X_1 = \cos(\theta)\): circular trajectory along the x-axis, \(X_2 = \sin(\theta)\): circular trajectory along the y-axis, \(X_3 = 0.05\theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix, and \(X_4 = 0.1\sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
+
+\begin{verbatim}
+helicalspiral <- gen_helicalspiral(n = 1000)
+\end{verbatim}
+
+Similarly, the \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of \(p\text{-}D\) (Figure \ref{fig:triginometric-proj} e).
+The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a \(3\text{-}D\) conic helix.
+The shape is defined by parameter \(\theta \in [0, 2\pi\text{spins}]\), controlling the angular progression of the spiral.
+The Archimedean spiral in the horizontal plane is represented by; \(X_1 = \theta\cos(\theta)\) for radial expansion in x, and \(X_2 = \theta\sin(\theta)\) for radial expansion in y.
+The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2\theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\) Spiral modulation in the fourth dimension is represented by \(X_4 = \theta\sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
+
+\begin{verbatim}
+conicspiral <- gen_conicspiral(n = 1000, spins = 1)
+\end{verbatim}
+
+Finally, the \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
+The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
+
+\begin{verbatim}
+nonlinear <- gen_nonlinear(n = 1000, hc = 1, non_fac = 0.5)
+\end{verbatim}
+
+\begin{figure}[!]
+
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
+
+}
+
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.}\label{fig:triginometric-proj}
 \end{figure}

### 2025-10-03 paper-cardinalR.Rmd
Approximately added all the content that I want to have (still need to add an application where this shape or something similar might be seen)

@@
 These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures.
-If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
+If a shape is not inherently defined in more than two dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
 Users can specify how these noise dimensions are generated (e.g., Gaussian, wavy) (`noise_fun`), offering control over the embedding process.
-All shape generators allow the user to define the number of points (`n`) and dimensions (`p`), and most include additional arguments to customize specific characteristics of the structure.
+All shape generators allow the user to define the number of points (`n`), and most include additional arguments to customize specific characteristics of the structure.
 
@@
   plot_layout(ncol = 3, guides = "collect") 
+```
+
+#### Möbius
+
+The `gen_mobius()` function is a **wrapper** around `geozoo::mobius()`, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
+The function returns a tibble with $n$ sampled points forming the surface of a Möbius strip.
+
+<!-- The Möbius strip is a classical non-orientable surface, looping back on itself with a half-twist.
+Its construction is based on a parameterization with angle $\theta \sim U(0, 2\pi)$ (position along the loop) and width $w \sim U(-0.4, 0.4)$ (offset across the band).
+The Cartesian coordinates are -->
+
+<!-- $$ -->
+<!-- X_1 = \Big(1 + \tfrac{w}{2} \cos(\tfrac{\theta}{2})\Big) \cos(\theta), \quad -->
+<!-- X_2 = \Big(1 + \tfrac{w}{2} \cos(\tfrac{\theta}{2})\Big) \sin(\theta), \quad -->
+<!-- X_3 = \tfrac{w}{2}\sin(\tfrac{\theta}{2}), -->
+<!-- $$ -->
+
+<!-- mapping a two-dimensional band with a half-twist into $\mathbb{R}^3$ as a one-sided, non-orientable surface. -->
+
+<!-- While `gen_mobius()` directly produces a 3-D structure, the package framework allows embedding this strip in higher-dimensional spaces by appending additional Gaussian noise variables.
+This enables the Möbius strip to be used in the same comparative setting as other high-dimensional synthetic structures.
+-->
+
+```{r data-mobius, echo=TRUE}
+mobius <- gen_mobius(n = 1000)
 ```
@@
 
+In contrast, the `gen_hollowsphere(n, p)` function, a wrapper around `geozoo::sphere.hollow()`, generates $n$ points uniformly distributed **only on the surface** of the $(p-1)$-dimensional sphere embedded in $\mathbb{R}^p$.
+This results in a hollow shell-like structure with no interior points.
+
+For example, when $p=3$, `gen_unifsphere()` produces a solid ball in 3D space, whereas `gen_hollowsphere()` produces only the spherical boundary.
+These paired structures allow controlled experiments to investigate how algorithms behave when data is concentrated throughout the full volume versus constrained to the boundary.
+
+```{r data-hollowsphere, echo=TRUE}
+hollowsphere <- gen_hollowsphere(n = 1000, p = 4)
+```
+
 In addition, the `gen_gridedsphere(n)` function constructs a $p\text{-}D$ dataset consisting of approximately $n$ points that are evenly distributed on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
@@
 
-#### Trigonometric
-
-Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric)"} else { "Figure \\@ref(fig:triginometric-proj)"})`).
-The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when $p > 2$ (adding $X_3$ through $X_p$), and for all other trigonometric shapes when $p > 4$ (adding $X_5$ through $X_p$).
-These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trigonometric-tb-html)', '\\@ref(tab:trigonometric-tb-pdf)'))` summarizes these functions.
-
-```{r}
-trigonometric_tb <- tibble(fun = c("gen_crescent",
-                            "gen_curvycylinder",
-                            "gen_sphericalspiral",
-                            "gen_helicalspiral",
-                            "gen_conicspiral",
-                            "gen_nonlinear"), 
-                        exp = c("Crescent pattern.",
-                                "Curvy cylinder.",
-                                "Spherical spiral.",
-                                "Helical spiral.",
-                                "Conic spiral.",
-                                "Nonlinear hyperbola."))
-```
-
-```{r trigonometric-tb-html, eval=knitr::is_html_output()}
-trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r trigonometric-tb-pdf, eval=knitr::is_latex_output()}
-trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-First, the `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) a"} else { "Figure \\@ref(fig:triginometric-proj) a"})`).
-Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
-The corresponding $2\text{-}D$ coordinates are defined by: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$
-
-```{r data-crescent, echo=TRUE}
-crescent <- gen_crescent(n = 1000)
-```
-
-Second, the `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
-The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
-Let $\theta \sim U(0, 3\pi)$ represent a random angle on a circular base and $z \sim U(0, h)$ represent the height along the cylinder.
-The coordinates are defined as: $X_1 = \cos(\theta)$ (Circular base, x-axis), $X_2 = \sin(\theta)$ (Circular base, y-axis), $X_3 = z$ (Linear height), and $X_4 = \sin(z)$ (Nonlinear curvy variation along height).
-
-```{r data-curvycylinder, echo=TRUE}
-curvycylinder <- gen_curvycylinder(n = 1000, h = 10)
-```
-
-```{r}
-curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r curvycylinder-proj1}
-scaled_data <- scale_data_manual(curvycylinder)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-curvycylinder_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-curvycylinder_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-curvycylinder_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$ be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
-Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
-This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
-
-```{r data-sphericalspiral, echo=TRUE}
-sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
-```
-
-```{r}
-sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r sphericalspiral-proj1}
-scaled_data <- scale_data_manual(sphericalspiral)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-sphericalspiral_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphericalspiral-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-sphericalspiral_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphericalspiral-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-sphericalspiral_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For a helical spiral in four dimensions, `gen_helicalspiral(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) d"} else { "Figure \\@ref(fig:triginometric-proj) d"})`), let $\theta \in [0, 5\pi/4]$ be a sequence of angles controlling rotation around a circle.
-Cartesian coordinates; $X_1 = \cos(\theta)$: circular trajectory along the x-axis, $X_2 = \sin(\theta)$: circular trajectory along the y-axis, $X_3 = 0.05\theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix, and $X_4 = 0.1\sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
-
-```{r data-helicalspiral, echo=TRUE}
-helicalspiral <- gen_helicalspiral(n = 1000)
-```
-
-Similarly, the `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) e"} else { "Figure \\@ref(fig:triginometric-proj) e"})`).
-The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a $3\text{-}D$ conic helix.
-The shape is defined by parameter $\theta \in [0, 2\pi\text{spins}]$, controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta\cos(\theta)$ for radial expansion in x, and $X_2 = \theta\sin(\theta)$ for radial expansion in y.
-The growth pattern resembles a cone, with the height increasing according to $X_3 = 2\theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta\sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
-
-```{r data-conicspiral, echo=TRUE}
-conicspiral <- gen_conicspiral(n = 1000, spins = 1)
-```
-
-```{r}
-conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r conicspiral-proj1}
-scaled_data <- scale_data_manual(conicspiral)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-conicspiral_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "e1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r conicspiral-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.01))
-
-conicspiral_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.5), 
-  title = "e2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r conicspiral-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.016))
-
-conicspiral_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.62), 
-  title = "e3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Finally, the `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
-The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-
-```{r data-nonlinear, echo=TRUE}
-nonlinear <- gen_nonlinear(n = 1000, hc = 1, non_fac = 0.5)
-```
-
-```{r}
-nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r nonlinear-proj1}
-scaled_data <- scale_data_manual(nonlinear)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
-                                              threshold = 0.022))
-
-nonlinear_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.45, 0.73), 
-  title = "f1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r nonlinear-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.45,
-                                              axis_scaled = 1.3, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.01))
-
-nonlinear_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.46), 
-  title = "f2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r nonlinear-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.45,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.016))
-
-nonlinear_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.35, 0.4), 
-  title = "f3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r triginometric, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of tthe `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data in $4\\text{-}D$."}
-
-triginometricfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    curvycylinder_lang, sphericalspiral_lang, conicspiral_lang, nonlinear_lang
-  ),
-  device = "xs"
-)
-
-class(triginometricfig) <- c(class(triginometricfig), "htmlwidget")
-
-triginometricfig
-```
-
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=16, out.width="80%", fig.align='center'}
-
-curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
-sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +
-conicspiral_proj1 + conicspiral_proj2 + conicspiral_proj3 +
-nonlinear_proj1 + nonlinear_proj2 + nonlinear_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-### Wrappers
+### Generate a Spherical or Hyperspherical Hole Within a Structure
+
+The package provides functionality for generating datasets with **spherical hole** (in 2D/3D) or, more generally, **hyperspherical hole** (in higher dimensions).
+These structures are valuable for evaluating how dimension reduction methods and clustering algorithms handle incomplete manifolds or missing regions of the data space.
+A hyperspherical hole introduces topological complexity: the structure remains continuous but contains excluded regions (voids) that algorithms must correctly represent in lower-dimensional embeddings.
+
+The core function `gen_hole()` removes points from a dataset that fall within a user-specified hypersphere.
+Formally, given data points ($x \in \mathbb{R}^p$), a center ($a \in \mathbb{R}^p$), and radius ($r > 0$), only points satisfying
+
+$$
+||x - a||_2 > r
+$$
+
+are retained.
+The anchor point (a) can either be user-specified or default to the dataset mean, and radius (r) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
+Because it operates generically on any dataset, spherical or hyperspherical holes can be embedded in a wide range of geometric structures.
+
+Two specialized wrappers illustrate this idea.
+The function `gen_scurvehole()` generates an S-curve with a spherical hole by applying `gen_hole()` to the output of `gen_scurve()`.
+This structure has been used in prior diagnostic studies of NLDR methods [@wang2021], since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
+The second wrapper, `gen_unifcubehole()`, generates uniformly sampled cube data with a hyperspherical hole.
+By embedding a hyperspherical void inside a convex high-dimensional structure, this creates non-convex regions that challenge algorithms in terms of separability and neighborhood preservation.
 

### 2025-10-03 paper-cardinalR.tex
Approximately added all the content that I want to have (still need to add an application where this shape or something similar might be seen)

@@
 These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures.
-If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
+If a shape is not inherently defined in more than two dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
 Users can specify how these noise dimensions are generated (e.g., Gaussian, wavy) (\texttt{noise\_fun}), offering control over the embedding process.
-All shape generators allow the user to define the number of points (\texttt{n}) and dimensions (\texttt{p}), and most include additional arguments to customize specific characteristics of the structure.
+All shape generators allow the user to define the number of points (\texttt{n}), and most include additional arguments to customize specific characteristics of the structure.
 
@@
 
+\paragraph{Möbius}\label{muxf6bius}
+
+The \texttt{gen\_mobius()} function is a \textbf{wrapper} around \texttt{geozoo::mobius()}, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
+The function returns a tibble with \(n\) sampled points forming the surface of a Möbius strip.
+
+\begin{verbatim}
+mobius <- gen_mobius(n = 1000)
+\end{verbatim}
+
 \paragraph{Pyramid}\label{pyramid}
@@
 
+In contrast, the \texttt{gen\_hollowsphere(n,\ p)} function, a wrapper around \texttt{geozoo::sphere.hollow()}, generates \(n\) points uniformly distributed \textbf{only on the surface} of the \((p-1)\)-dimensional sphere embedded in \(\mathbb{R}^p\).
+This results in a hollow shell-like structure with no interior points.
+
+For example, when \(p=3\), \texttt{gen\_unifsphere()} produces a solid ball in 3D space, whereas \texttt{gen\_hollowsphere()} produces only the spherical boundary.
+These paired structures allow controlled experiments to investigate how algorithms behave when data is concentrated throughout the full volume versus constrained to the boundary.
+
+\begin{verbatim}
+hollowsphere <- gen_hollowsphere(n = 1000, p = 4)
+\end{verbatim}
+
 In addition, the \texttt{gen\_gridedsphere(n)} function constructs a \(p\text{-}D\) dataset consisting of approximately \(n\) points that are evenly distributed on the surface of the unit \((p-1)\)-sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
@@
 
-\paragraph{Trigonometric}\label{trigonometric}
+\subsubsection{Generate a Spherical or Hyperspherical Hole Within a Structure}\label{generate-a-spherical-or-hyperspherical-hole-within-a-structure}
 
-Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (Figure \ref{fig:triginometric-proj}).
-The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when \(p > 2\) (adding \(X_3\) through \(X_p\)), and for all other trigonometric shapes when \(p > 4\) (adding \(X_5\) through \(X_p\)).
-These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
-Table \ref{tab:trigonometric-tb-pdf} summarizes these functions.
+The package provides functionality for generating datasets with \textbf{spherical hole} (in 2D/3D) or, more generally, \textbf{hyperspherical hole} (in higher dimensions).
+These structures are valuable for evaluating how dimension reduction methods and clustering algorithms handle incomplete manifolds or missing regions of the data space.
+A hyperspherical hole introduces topological complexity: the structure remains continuous but contains excluded regions (voids) that algorithms must correctly represent in lower-dimensional embeddings.
 
-\begin{table}
+The core function \texttt{gen\_hole()} removes points from a dataset that fall within a user-specified hypersphere.
+Formally, given data points (\(x \in \mathbb{R}^p\)), a center (\(a \in \mathbb{R}^p\)), and radius (\(r > 0\)), only points satisfying
 
-\caption{\label{tab:trigonometric-tb-pdf}cardinalR trigonometric data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_crescent & Crescent pattern.\\
-gen\_curvycylinder & Curvy cylinder.\\
-gen\_sphericalspiral & Spherical spiral.\\
-gen\_helicalspiral & Helical spiral.\\
-gen\_conicspiral & Conic spiral.\\
-\addlinespace
-gen\_nonlinear & Nonlinear hyperbola.\\
-\bottomrule
-\end{tabular}
-\end{table}
+\[
+||x - a||_2 > r
+\]
 
-First, the \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise (Figure \ref{fig:triginometric-proj} a).
-Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
-The corresponding \(2\text{-}D\) coordinates are defined by: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta).\]
+are retained.
+The anchor point (a) can either be user-specified or default to the dataset mean, and radius (r) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
+Because it operates generically on any dataset, spherical or hyperspherical holes can be embedded in a wide range of geometric structures.
 
-\begin{verbatim}
-crescent <- gen_crescent(n = 1000)
-\end{verbatim}
-
-Second, the \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
-The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
-Let \(\theta \sim U(0, 3\pi)\) represent a random angle on a circular base and \(z \sim U(0, h)\) represent the height along the cylinder.
-The coordinates are defined as: \(X_1 = \cos(\theta)\) (Circular base, x-axis), \(X_2 = \sin(\theta)\) (Circular base, y-axis), \(X_3 = z\) (Linear height), and \(X_4 = \sin(z)\) (Nonlinear curvy variation along height).
-
-\begin{verbatim}
-curvycylinder <- gen_curvycylinder(n = 1000, h = 10)
-\end{verbatim}
-
-For a spiraling path on a spherical surface in the first four dimensions, \texttt{gen\_sphericalspiral(n,\ p,\ spins)} (Figure \ref{fig:triginometric-proj} c), let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\)be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
-Cartesian coordinates from spherical conversion: \(X_1 = \sin(\phi)\cos(\theta)\), \(X_2 = \sin(\phi)\sin(\theta)\), \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter, and \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
-This generates a spherical spiral curve embedded in \(4\text{-}D\) space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
-
-\begin{verbatim}
-sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
-\end{verbatim}
-
-For a helical spiral in four dimensions, \texttt{gen\_helicalspiral(n,\ p)} (Figure \ref{fig:triginometric-proj} d), let \(\theta \in [0, 5\pi/4]\) be a sequence of angles controlling rotation around a circle.
-Cartesian coordinates; \(X_1 = \cos(\theta)\): circular trajectory along the x-axis, \(X_2 = \sin(\theta)\): circular trajectory along the y-axis, \(X_3 = 0.05\theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix, and \(X_4 = 0.1\sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
-
-\begin{verbatim}
-helicalspiral <- gen_helicalspiral(n = 1000)
-\end{verbatim}
-
-Similarly, the \texttt{gen\_conicspiral(n,\ p,\ spins)} function generates a dataset of \(n\) points forming a conical spiral in the first four dimensions of \(p\text{-}D\) (Figure \ref{fig:triginometric-proj} e).
-The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a \(3\text{-}D\) conic helix.
-The shape is defined by parameter \(\theta \in [0, 2\pi\text{spins}]\), controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; \(X_1 = \theta\cos(\theta)\) for radial expansion in x, and \(X_2 = \theta\sin(\theta)\) for radial expansion in y.
-The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2\theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\) Spiral modulation in the fourth dimension is represented by \(X_4 = \theta\sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
-
-\begin{verbatim}
-conicspiral <- gen_conicspiral(n = 1000, spins = 1)
-\end{verbatim}
-
-Finally, the \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
-The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-
-\begin{verbatim}
-nonlinear <- gen_nonlinear(n = 1000, hc = 1, non_fac = 0.5)
-\end{verbatim}
-
-\begin{figure}[!]
-
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
-
-}
-
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.}\label{fig:triginometric-proj}
-\end{figure}
-
-\subsubsection{Wrappers}\label{wrappers}
+Two specialized wrappers illustrate this idea.
+The function \texttt{gen\_scurvehole()} generates an S-curve with a spherical hole by applying \texttt{gen\_hole()} to the output of \texttt{gen\_scurve()}.
+This structure has been used in prior diagnostic studies of NLDR methods \citep{wang2021}, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
+The second wrapper, \texttt{gen\_unifcubehole()}, generates uniformly sampled cube data with a hyperspherical hole.
+By embedding a hyperspherical void inside a convex high-dimensional structure, this creates non-convex regions that challenge algorithms in terms of separability and neighborhood preservation.
 

### 2025-10-03 paper-cardinalR.Rmd
Removed Gaussian images

@@
 
-The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:gau)"} else { "Figure \\@ref(fig:gau-proj)"})`).
+The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure.
 Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
@@
 gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
-```
-
-```{r}
-gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r gau-proj1}
-scaled_data <- scale_data_manual(gau)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-gau_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-gau_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-gau_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gau` data in $4\\text{-}D$."}
-
-gaufig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    gau_lang
-  ),
-  device = "xs"
-)
-
-class(gaufig) <- c(class(gaufig), "htmlwidget")
-
-gaufig
-```
-
-```{r gau-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gau` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-gau_proj1 + gau_proj2 + gau_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-10-03 paper-cardinalR.tex
Removed Gaussian images

@@
 
-The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure (Figure \ref{fig:gau-proj}).
+The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure.
 Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, s)\), where \(s\) is a user-defined \(p \times p\) positive-definite matrix.
@@
 \end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:gau-proj}
-\end{figure}
 

### 2025-10-03 paper-cardinalR.Rmd
Added Trigometric section (forgot earlier :))

--- 2025-10-03 paper-cardinalR.Rmd
@@
 main_tb |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation")) 
+  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "") 
 ```
@@
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H", linesep = "")  |>
   column_spec(1, width = "2cm") |>
@@
 
-```{r}
-gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r gau-proj1}
-scaled_data <- scale_data_manual(gau)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-gau_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-gau_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-gau_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gau, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gau` data in $4\\text{-}D$."}
-
-gaufig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    gau_lang
-  ),
-  device = "xs"
-)
-
-class(gaufig) <- c(class(gaufig), "htmlwidget")
-
-gaufig
-```
-
-```{r gau-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gau` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-gau_proj1 + gau_proj2 + gau_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
+#### Möbius
+
+The `gen_mobius()` function is a **wrapper** around `geozoo::mobius()`, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
+The function returns a tibble with $n$ sampled points forming the surface of a Möbius strip.
+
+<!-- The Möbius strip is a classical non-orientable surface, looping back on itself with a half-twist.
+Its construction is based on a parameterization with angle $\theta \sim U(0, 2\pi)$ (position along the loop) and width $w \sim U(-0.4, 0.4)$ (offset across the band).
+The Cartesian coordinates are -->
+
+<!-- $$ -->
+<!-- X_1 = \Big(1 + \tfrac{w}{2} \cos(\tfrac{\theta}{2})\Big) \cos(\theta), \quad -->
+<!-- X_2 = \Big(1 + \tfrac{w}{2} \cos(\tfrac{\theta}{2})\Big) \sin(\theta), \quad -->
+<!-- X_3 = \tfrac{w}{2}\sin(\tfrac{\theta}{2}), -->
+<!-- $$ -->
+
+<!-- mapping a two-dimensional band with a half-twist into $\mathbb{R}^3$ as a one-sided, non-orientable surface. -->
+
+<!-- While `gen_mobius()` directly produces a 3-D structure, the package framework allows embedding this strip in higher-dimensional spaces by appending additional Gaussian noise variables.
+This enables the Möbius strip to be used in the same comparative setting as other high-dimensional synthetic structures.
+-->
+
+```{r data-mobius, echo=TRUE}
+mobius <- gen_mobius(n = 1000)
 ```
@@
 sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation"), linesep = "") 
 ```
@@
 sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "")  |>
   column_spec(1, width = "4cm") |>
@@
 
+```
+
+In contrast, the `gen_hollowsphere(n, p)` function, a wrapper around `geozoo::sphere.hollow()`, generates $n$ points uniformly distributed **only on the surface** of the $(p-1)$-dimensional sphere embedded in $\mathbb{R}^p$.
+This results in a hollow shell-like structure with no interior points.
+
+For example, when $p=3$, `gen_unifsphere()` produces a solid ball in 3D space, whereas `gen_hollowsphere()` produces only the spherical boundary.
+These paired structures allow controlled experiments to investigate how algorithms behave when data is concentrated throughout the full volume versus constrained to the boundary.
+
+```{r data-hollowsphere, echo=TRUE}
+hollowsphere <- gen_hollowsphere(n = 1000, p = 4)
 ```
@@
 
-### Wrappers
+### Generate a Spherical or Hyperspherical Hole Within a Structure
+
+The package provides functionality for generating datasets with **spherical hole** (in 2D/3D) or, more generally, **hyperspherical hole** (in higher dimensions).
+These structures are valuable for evaluating how dimension reduction methods and clustering algorithms handle incomplete manifolds or missing regions of the data space.
+A hyperspherical hole introduces topological complexity: the structure remains continuous but contains excluded regions (voids) that algorithms must correctly represent in lower-dimensional embeddings.
+
+The core function `gen_hole()` removes points from a dataset that fall within a user-specified hypersphere.
+Formally, given data points ($x \in \mathbb{R}^p$), a center ($a \in \mathbb{R}^p$), and radius ($r > 0$), only points satisfying
+
+$$
+||x - a||_2 > r
+$$
+
+are retained.
+The anchor point (a) can either be user-specified or default to the dataset mean, and radius (r) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
+Because it operates generically on any dataset, spherical or hyperspherical holes can be embedded in a wide range of geometric structures.
+
+Two specialized wrappers illustrate this idea.
+The function `gen_scurvehole()` generates an S-curve with a spherical hole by applying `gen_hole()` to the output of `gen_scurve()`.
+This structure has been used in prior diagnostic studies of NLDR methods [@wang2021], since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
+The second wrapper, `gen_unifcubehole()`, generates uniformly sampled cube data with a hyperspherical hole.
+By embedding a hyperspherical void inside a convex high-dimensional structure, this creates non-convex regions that challenge algorithms in terms of separability and neighborhood preservation.
 
@@
 odd_shapes_tb |> 
-  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation"), linesep = "") 
 ```
@@
 odd_shapes_tb |> 
-  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "")  |>
   column_spec(1, width = "3.5cm") |>
@@
 add_fun_tb |> 
-  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation"), linesep = "") 
 ```
@@
 add_fun_tb |> 
-  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "")  |>
   column_spec(1, width = "4cm") |>

### 2025-10-03 paper-cardinalR.tex
Added Trigometric section (forgot earlier :))

--- 2025-10-03 paper-cardinalR.tex
@@
 shape & character (vector) & Shapes of clusters.\\
-\addlinespace
 rotation & numeric (list) & Rotation matrices, one per cluster.\\
@@
 
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/gau-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gau` data.}\label{fig:gau-proj}
-\end{figure}
+\paragraph{Möbius}\label{muxf6bius}
+
+The \texttt{gen\_mobius()} function is a \textbf{wrapper} around \texttt{geozoo::mobius()}, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
+The function returns a tibble with \(n\) sampled points forming the surface of a Möbius strip.
+
+\begin{verbatim}
+mobius <- gen_mobius(n = 1000)
+\end{verbatim}
 
@@
 gen\_clusteredspheres & Multiple small spheres within a big sphere.\\
-\addlinespace
 gen\_hemisphere & Hemisphere.\\
@@
 unifsphere <- gen_unifsphere(n = 1000, r = 1)
+\end{verbatim}
+
+In contrast, the \texttt{gen\_hollowsphere(n,\ p)} function, a wrapper around \texttt{geozoo::sphere.hollow()}, generates \(n\) points uniformly distributed \textbf{only on the surface} of the \((p-1)\)-dimensional sphere embedded in \(\mathbb{R}^p\).
+This results in a hollow shell-like structure with no interior points.
+
+For example, when \(p=3\), \texttt{gen\_unifsphere()} produces a solid ball in 3D space, whereas \texttt{gen\_hollowsphere()} produces only the spherical boundary.
+These paired structures allow controlled experiments to investigate how algorithms behave when data is concentrated throughout the full volume versus constrained to the boundary.
+
+\begin{verbatim}
+hollowsphere <- gen_hollowsphere(n = 1000, p = 4)
 \end{verbatim}
@@
 
-\subsubsection{Wrappers}\label{wrappers}
+\subsubsection{Generate a Spherical or Hyperspherical Hole Within a Structure}\label{generate-a-spherical-or-hyperspherical-hole-within-a-structure}
+
+The package provides functionality for generating datasets with \textbf{spherical hole} (in 2D/3D) or, more generally, \textbf{hyperspherical hole} (in higher dimensions).
+These structures are valuable for evaluating how dimension reduction methods and clustering algorithms handle incomplete manifolds or missing regions of the data space.
+A hyperspherical hole introduces topological complexity: the structure remains continuous but contains excluded regions (voids) that algorithms must correctly represent in lower-dimensional embeddings.
+
+The core function \texttt{gen\_hole()} removes points from a dataset that fall within a user-specified hypersphere.
+Formally, given data points (\(x \in \mathbb{R}^p\)), a center (\(a \in \mathbb{R}^p\)), and radius (\(r > 0\)), only points satisfying
+
+\[
+||x - a||_2 > r
+\]
+
+are retained.
+The anchor point (a) can either be user-specified or default to the dataset mean, and radius (r) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
+Because it operates generically on any dataset, spherical or hyperspherical holes can be embedded in a wide range of geometric structures.
+
+Two specialized wrappers illustrate this idea.
+The function \texttt{gen\_scurvehole()} generates an S-curve with a spherical hole by applying \texttt{gen\_hole()} to the output of \texttt{gen\_scurve()}.
+This structure has been used in prior diagnostic studies of NLDR methods \citep{wang2021}, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
+The second wrapper, \texttt{gen\_unifcubehole()}, generates uniformly sampled cube data with a hyperspherical hole.
+By embedding a hyperspherical void inside a convex high-dimensional structure, this creates non-convex regions that challenge algorithms in terms of separability and neighborhood preservation.
 
@@
 make\_chain\_circles & Chain-like circular clusters connected sequentially.\\
-\addlinespace
 make\_klink\_curvycycle & K-link curvy cycle clusters (curvilinear loop structures).\\
@@
 make\_onegrid & Single grid in two dimensions.\\
-\addlinespace
 make\_twogrid\_overlap & Two overlapping grids.\\
@@
 gen\_nsum & Generates a vector of positive integers whose summation is approximately equal to a target value.\\
-\addlinespace
 gen\_rotation & Generates rotations.\\

### 2025-10-03 paper-cardinalR.Rmd
Changed the order

@@
 
+#### Gaussian
+
+The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure.
+Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
+
+```{r data-gau, echo=TRUE}
+gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
+```
+
 #### Linear
@@
 
-#### Gaussian
-
-The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure.
-Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
-
-```{r data-gau, echo=TRUE}
-gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
-```
-
 #### Möbius
@@
 mobius <- gen_mobius(n = 1000)
+```
+
+#### Polynomial
+
+A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in \gD{} space.
+To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
+These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
+We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:polynomial-tb-html)', '\\@ref(tab:polynomial-tb-pdf)'))` summarizes these functions and their purposes.
+
+```{r}
+polynomial_tb <- tibble(fun = c("gen_quadratic",
+                                "gen_cubic"), 
+                        exp = c("Quadratic pattern.",
+                                "Cubic pattern."))
+```
+
+```{r polynomial-tb-html, eval=knitr::is_html_output()}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation")) 
+```
+
+```{r polynomial-tb-pdf, eval=knitr::is_latex_output()}
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+The first is the quadratic curve of $n$ points in two dimensions.
+This is generated using `gen_quadratic(n, range)`.
+The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree 2 is applied to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
+This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
+
+```{r data-quadratic, echo=TRUE}
+quadratic <- gen_quadratic(n = 1000)
+```
+
+The second is the cubic curve of $n$ points in two dimensions.
+This is generated using `gen_cubic(n, range)`.
+The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
+This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
+
+```{r data-cubic, echo=TRUE}
+cubic <- gen_cubic(n = 1000)
 ```
@@
   plot_layout(ncol = 3, guides = "collect") 
-```
-
-#### Polynomial
-
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in \gD{} space.
-To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
-These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
-We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:polynomial-tb-html)', '\\@ref(tab:polynomial-tb-pdf)'))` summarizes these functions and their purposes.
-
-```{r}
-polynomial_tb <- tibble(fun = c("gen_quadratic",
-                                "gen_cubic"), 
-                        exp = c("Quadratic pattern.",
-                                "Cubic pattern."))
-```
-
-```{r polynomial-tb-html, eval=knitr::is_html_output()}
-polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation")) 
-```
-
-```{r polynomial-tb-pdf, eval=knitr::is_latex_output()}
-polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-The first is the quadratic curve of $n$ points in two dimensions.
-This is generated using `gen_quadratic(n, range)`.
-The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree 2 is applied to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
-
-```{r data-quadratic, echo=TRUE}
-quadratic <- gen_quadratic(n = 1000)
-```
-
-The second is the cubic curve of $n$ points in two dimensions.
-This is generated using `gen_cubic(n, range)`.
-The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
-
-```{r data-cubic, echo=TRUE}
-cubic <- gen_cubic(n = 1000)
 ```
@@
 trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation"), linesep = "") 
 ```
@@
 trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "")  |>
   column_spec(1, width = "4cm") |>

### 2025-10-03 paper-cardinalR.tex
Changed the order

@@
 
+\paragraph{Gaussian}\label{gaussian}
+
+The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure.
+Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, s)\), where \(s\) is a user-defined \(p \times p\) positive-definite matrix.
+
+\begin{verbatim}
+gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
+\end{verbatim}
+
 \paragraph{Linear}\label{linear}
@@
 
-\paragraph{Gaussian}\label{gaussian}
-
-The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure.
-Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, s)\), where \(s\) is a user-defined \(p \times p\) positive-definite matrix.
-
-\begin{verbatim}
-gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
-\end{verbatim}
-
 \paragraph{Möbius}\label{muxf6bius}
@@
 mobius <- gen_mobius(n = 1000)
+\end{verbatim}
+
+\paragraph{Polynomial}\label{polynomial}
+
+A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in $2\text{-}D${} space.
+To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
+These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
+We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
+Table \ref{tab:polynomial-tb-pdf} summarizes these functions and their purposes.
+
+\begin{table}
+
+\caption{\label{tab:polynomial-tb-pdf}cardinalR polynomial data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_quadratic & Quadratic pattern.\\
+gen\_cubic & Cubic pattern.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The first is the quadratic curve of \(n\) points in two dimensions.
+This is generated using \texttt{gen\_quadratic(n,\ range)}.
+The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree 2 is applied to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
+This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
+
+\begin{verbatim}
+quadratic <- gen_quadratic(n = 1000)
+\end{verbatim}
+
+The second is the cubic curve of \(n\) points in two dimensions.
+This is generated using \texttt{gen\_cubic(n,\ range)}.
+The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree \(3\) is applied to construct \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
+This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
+
+\begin{verbatim}
+cubic <- gen_cubic(n = 1000)
 \end{verbatim}
@@
 \end{figure}
-
-\paragraph{Polynomial}\label{polynomial}
-
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in $2\text{-}D${} space.
-To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
-These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
-We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-Table \ref{tab:polynomial-tb-pdf} summarizes these functions and their purposes.
-
-\begin{table}
-
-\caption{\label{tab:polynomial-tb-pdf}cardinalR polynomial data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_quadratic & Quadratic pattern.\\
-gen\_cubic & Cubic pattern.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-The first is the quadratic curve of \(n\) points in two dimensions.
-This is generated using \texttt{gen\_quadratic(n,\ range)}.
-The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree 2 is applied to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
-
-\begin{verbatim}
-quadratic <- gen_quadratic(n = 1000)
-\end{verbatim}
-
-The second is the cubic curve of \(n\) points in two dimensions.
-This is generated using \texttt{gen\_cubic(n,\ range)}.
-The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree \(3\) is applied to construct \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
-This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
-
-\begin{verbatim}
-cubic <- gen_cubic(n = 1000)
-\end{verbatim}
 
@@
 gen\_conicspiral & Conic spiral.\\
-\addlinespace
 gen\_nonlinear & Nonlinear hyperbola.\\

### 2025-10-03 paper-cardinalR.R

@@ -718,6 +718,10 @@ cone_proj1 + cone_proj2 + cone_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
 
 
+## ----data-gau, echo=TRUE------------------------------------------------------
+gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
+
+
 ## ----data-linear, echo=TRUE---------------------------------------------------
 linear <- gen_longlinear(n = 1000, p = 4)
 
@@ -831,12 +835,35 @@ linear_proj1 + linear_proj2 + linear_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
 
 
-## ----data-gau, echo=TRUE------------------------------------------------------
-gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
-
-
 ## ----data-mobius, echo=TRUE---------------------------------------------------
 mobius <- gen_mobius(n = 1000)
+
+
+## -----------------------------------------------------------------------------
+polynomial_tb <- tibble(fun = c("gen_quadratic",
+                                "gen_cubic"), 
+                        exp = c("Quadratic pattern.",
+                                "Cubic pattern."))
+
+
+## ----polynomial-tb-html, eval=knitr::is_html_output()-------------------------
+# polynomial_tb |>
+#   kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation"))
+
+
+## ----polynomial-tb-pdf, eval=knitr::is_latex_output()-------------------------
+polynomial_tb |> 
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+
+
+## ----data-quadratic, echo=TRUE------------------------------------------------
+quadratic <- gen_quadratic(n = 1000)
+
+
+## ----data-cubic, echo=TRUE----------------------------------------------------
+cubic <- gen_cubic(n = 1000)
 
 
 ## ----data-pyrrect, echo=TRUE--------------------------------------------------
@@ -1244,33 +1271,6 @@ pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
 pyrstar_proj1 + pyrstar_proj2 + pyrstar_proj3 +
 pyrholes_proj1 + pyrholes_proj2 + pyrholes_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
-
-
-## -----------------------------------------------------------------------------
[…]

### 2025-10-03 paper-cardinalR.Rmd
Updated Trefoil table
Removed adding noise dimnsions text

@@
 
-For all branch-generating functions, when the target dimensionality $p > 2$, additional dimensions are filled with independent noise, allowing the $2\text{-}D$ or $2\text{-}D$-subspace structures to be naturally embedded in higher-dimensional space while preserving the intended geometric patterns.
-
 ```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, and `orgcurvybranches` data in $4\\text{-}D$."}
@@
                             "gen_trefoil3d"), 
-                        exp = c("Trefoil in $4\\text{-}D$.",
-                                "Trefoil in $3\\text{-}D$."))
+                        exp = c("Trefoil in 4-D.",
+                                "Trefoil in 3-D."))
 ```

### 2025-10-03 paper-cardinalR.tex
Updated Trefoil table
Removed adding noise dimnsions text

@@
 \end{verbatim}
-
-For all branch-generating functions, when the target dimensionality \(p > 2\), additional dimensions are filled with independent noise, allowing the \(2\text{-}D\) or \(2\text{-}D\)-subspace structures to be naturally embedded in higher-dimensional space while preserving the intended geometric patterns.
 
@@
 \midrule
-gen\_trefoil4d & Trefoil in \$4\textbackslash{}text\{-\}D\$.\\
-gen\_trefoil3d & Trefoil in \$3\textbackslash{}text\{-\}D\$.\\
+gen\_trefoil4d & Trefoil in 4-D.\\
+gen\_trefoil3d & Trefoil in 3-D.\\
 \bottomrule

### 2025-10-03 paper-cardinalR.Rmd
Updated the positions of the tables

@@
 main_tb |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "") 
+  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", table.pos = "H") 
 ```
@@
 branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
 ```
@@
 branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Explanation")) 
+  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Explanation"), table.pos = "H") 
 ```
@@
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
+  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
 ```
@@
 cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
 ```
@@
 polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 pyramid_tb |> 
-  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
 ```
@@
 pyramid_tb |> 
-  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation"), linesep = "") 
+  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "H") 
 ```
@@
 sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "")  |>
+  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 trefoil_tb |> 
-  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
 ```
@@
 trefoil_tb |> 
-  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation"), linesep = "") 
+  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "H") 
 ```
@@
 trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "")  |>
+  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 
-### Generate a Spherical or Hyperspherical Hole Within a Structure
+### Generate a spherical or hyperspherical hole within a structure
 
@@
 noise_fun_tb |> 
-  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation")) 
+  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
 ```
@@
 noise_fun_tb |> 
-  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
@@
 odd_shapes_tb |> 
-  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation"), linesep = "") 
+  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "H") 
 ```
@@
 odd_shapes_tb |> 
-  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "")  |>
+  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "H")  |>
   column_spec(1, width = "3.5cm") |>
@@
 add_fun_tb |> 
-  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation"), linesep = "") 
+  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "H") 
 ```
@@
 add_fun_tb |> 
-  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "")  |>
+  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "H")  |>
   column_spec(1, width = "4cm") |>

### 2025-10-03 paper-cardinalR.R

@@ -113,7 +113,7 @@ main_tb <- tibble(arg = c("n",
 
 ## ----main-tb-html, eval=knitr::is_html_output()-------------------------------
 # main_tb |>
-#   kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "")
+#   kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", table.pos = "H")
 
 
 ## ----main-tb-pdf, eval=knitr::is_latex_output()-------------------------------
@@ -139,12 +139,12 @@ branch_tb <- tibble(fun = c("gen_expbranches",
 
 ## ----branching-tb-html, eval=knitr::is_html_output()--------------------------
 # branch_tb |>
-#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"))
+#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), table.pos = "H")
 
 
 ## ----branching-tb-pdf, eval=knitr::is_latex_output()--------------------------
 branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
   column_spec(2, width = "8cm")
 
@@ -160,12 +160,12 @@ arg_branching_tb <- tibble(arg = c("n",
 
 ## ----arg-branching-tb-html, eval=knitr::is_html_output()----------------------
 # arg_branching_tb |>
-#   kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Explanation"))
+#   kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Explanation"), table.pos = "H")
 
 
 ## ----arg-branching-tb-pdf, eval=knitr::is_latex_output()----------------------
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T)  |>
[…]

### 2025-10-05 paper-cardinalR.Rmd
Updated Table 12

@@
 
-An S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) simulates data that lies along a smooth, non-linear manifold.
-
-For the S-curve structure, `gen_scurve(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`), the $3\text{-}D$ geometry is constructed by introducing a latent parameter, $\theta \sim U\left(-3\pi/2, 3\pi/2\right)$.
-This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure: $$X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1)$$.
-This configuration creates a horizontally curled shape in $(X_1, X_3)$, with additional band thickness in the $X_2$ direction.
+The S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) represents a smooth, non-linear manifold in 3D space.
+Using `gen_scurve(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`), the geometry is defined by a latent parameter $\theta \sim U(-3\pi/2, 3\pi/2)$, which controls the curvature of the manifold.
+The first three coordinates form the S-shape:
+
+$$
+X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1),
+$$
+
+creating a horizontally curled pattern in $(X_1, X_3)$ with additional thickness along $X_2$.
 
@@
 The function `gen_scurvehole()` generates an S-curve with a spherical hole by applying `gen_hole()` to the output of `gen_scurve()`.
-This structure has been used in prior diagnostic studies of NLDR methods [@wang2021], since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
+This structure has been used in prior diagnostic studies of NLDR methods, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
 The second wrapper, `gen_unifcubehole()`, generates uniformly sampled cube data with a hyperspherical hole.
@@
     "Circular clusters with a Gaussian cluster in the middle.",
-    "Curvy circular clusters with a Gaussian cluster in the middle.",
+    "Curvy circular clusters with a Gaussian in the middle.",
     "Single grid in two dimensions.",

### 2025-10-05 paper-cardinalR.tex
Updated the positions of the tables
Updated Table 12

@@
 
-An S-curve structure (Figure \ref{fig:scurve-proj}) simulates data that lies along a smooth, non-linear manifold.
+The S-curve structure (Figure \ref{fig:scurve-proj}) represents a smooth, non-linear manifold in 3D space.
+Using \texttt{gen\_scurve(n,\ p)} (Figure \ref{fig:scurve-proj} a), the geometry is defined by a latent parameter \(\theta \sim U(-3\pi/2, 3\pi/2)\), which controls the curvature of the manifold.
+The first three coordinates form the S-shape:
 
-For the S-curve structure, \texttt{gen\_scurve(n,\ p)} (Figure \ref{fig:scurve-proj} a), the \(3\text{-}D\) geometry is constructed by introducing a latent parameter, \(\theta \sim U\left(-3\pi/2, 3\pi/2\right)\).
-This parameter controls the curvature of the manifold.
-The first three dimensions form the S-curve structure: \[X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1)\].
-This configuration creates a horizontally curled shape in \((X_1, X_3)\), with additional band thickness in the \(X_2\) direction.
+\[
+X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1),
+\]
+
+creating a horizontally curled pattern in \((X_1, X_3)\) with additional thickness along \(X_2\).
 
@@
 
-\subsubsection{Generate a Spherical or Hyperspherical Hole Within a Structure}\label{generate-a-spherical-or-hyperspherical-hole-within-a-structure}
+\subsubsection{Generate a spherical or hyperspherical hole within a structure}\label{generate-a-spherical-or-hyperspherical-hole-within-a-structure}
 
@@
 The function \texttt{gen\_scurvehole()} generates an S-curve with a spherical hole by applying \texttt{gen\_hole()} to the output of \texttt{gen\_scurve()}.
-This structure has been used in prior diagnostic studies of NLDR methods \citep{wang2021}, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
+This structure has been used in prior diagnostic studies of NLDR methods, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
 The second wrapper, \texttt{gen\_unifcubehole()}, generates uniformly sampled cube data with a hyperspherical hole.
@@
 make\_gaucircles & Circular clusters with a Gaussian cluster in the middle.\\
-make\_gaucurvycycle & Curvy circular clusters with a Gaussian cluster in the middle.\\
+make\_gaucurvycycle & Curvy circular clusters with a Gaussian in the middle.\\
 make\_onegrid & Single grid in two dimensions.\\

### 2025-10-05 paper-cardinalR.Rmd
Updated Shape generator overview paragraph

@@
 
-The shape generators form the foundation of the package, providing a collection of functions to create synthetic data structures based on simple, well-defined geometric structures.
-These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures.
-If a shape is not inherently defined in more than two dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
-Users can specify how these noise dimensions are generated (e.g., Gaussian, wavy) (`noise_fun`), offering control over the embedding process.
-All shape generators allow the user to define the number of points (`n`), and most include additional arguments to customize specific characteristics of the structure.
+The shape generators form the foundation of the package, providing functions to create synthetic datasets from simple, well-defined geometric forms such as cones, pyramids, spheres, grids, and branching structures.
+Each generator includes the parameter `n`, which specifies the number of points to generate.
+Some functions, such as `gen_unifcube()`, also take the dimension `p`, while others include arguments specific to the geometry (e.g., radius for spheres (`r`), width for bands (`w`)).
+If higher-dimensional data are required, additional noise dimensions can be appended after data generation using any noise generator function.
+This flexibility allows users to construct both low- and high-dimensional datasets from the same underlying structures.
 

### 2025-10-05 paper-cardinalR.tex
Updated Shape generator overview paragraph

@@
 
-The shape generators form the foundation of the package, providing a collection of functions to create synthetic data structures based on simple, well-defined geometric structures.
-These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures.
-If a shape is not inherently defined in more than two dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space.
-Users can specify how these noise dimensions are generated (e.g., Gaussian, wavy) (\texttt{noise\_fun}), offering control over the embedding process.
-All shape generators allow the user to define the number of points (\texttt{n}), and most include additional arguments to customize specific characteristics of the structure.
+The shape generators form the foundation of the package, providing functions to create synthetic datasets from simple, well-defined geometric forms such as cones, pyramids, spheres, grids, and branching structures.
+Each generator includes the parameter \texttt{n}, which specifies the number of points to generate.
+Some functions, such as \texttt{gen\_unifcube()}, also take the dimension \texttt{p}, while others include arguments specific to the geometry (e.g., radius for spheres (\texttt{r}), width for bands (\texttt{w})).
+If higher-dimensional data are required, additional noise dimensions can be appended after data generation using any noise generator function.
+This flexibility allows users to construct both low- and high-dimensional datasets from the same underlying structures.
 

### 2025-10-05 paper-cardinalR.Rmd
Remove linear projections

@@
 
-```{r}
-linear_lang <- langevitour::langevitour(linear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r linear-proj1}
-scaled_data <- scale_data_manual(linear)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.022))
-
-linear_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linear-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.75,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.01))
-
-linear_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linear-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.9,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.02))
-
-linear_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r linear, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `linear` data in $4\\text{-}D$."}
-
-linearfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    linear_lang
-  ),
-  device = "xs"
-)
-
-class(linearfig) <- c(class(linearfig), "htmlwidget")
-
-linearfig
-```
-
-```{r linear-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `linear` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-linear_proj1 + linear_proj2 + linear_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
 #### Möbius

### 2025-10-05 paper-cardinalR.Rmd
Updated Table 6

@@
                              "gen_pyrfrac"), 
-                        exp = c("Rectangular-base pyramid, with a sharp or blunted apex.",
-          "Triangular-base pyramid, with a sharp or blunted apex.",
-          "Star-shaped base pyramid, with a sharp or blunted apex.",
-          "Pyramid containing triangular pyramid-shaped holes."))
+                        exp = c("Rectangular-base, with a sharp or blunted apex.",
+          "Triangular-base, with a sharp or blunted apex.",
+          "Star-shaped base, with a sharp or blunted apex.",
+          "Pyramid with triangular pyramid-shaped holes."))
 ```

### 2025-10-05 paper-cardinalR.tex
Remove linear projections
Updated Table 6

@@
 
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/linear-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `linear` data.}\label{fig:linear-proj}
-\end{figure}
-
 \paragraph{Möbius}\label{muxf6bius}
@@
 \midrule
-gen\_pyrrect & Rectangular-base pyramid, with a sharp or blunted apex.\\
-gen\_pyrtri & Triangular-base pyramid, with a sharp or blunted apex.\\
-gen\_pyrstar & Star-shaped base pyramid, with a sharp or blunted apex.\\
-gen\_pyrfrac & Pyramid containing triangular pyramid-shaped holes.\\
+gen\_pyrrect & Rectangular-base, with a sharp or blunted apex.\\
+gen\_pyrtri & Triangular-base, with a sharp or blunted apex.\\
+gen\_pyrstar & Star-shaped base, with a sharp or blunted apex.\\
+gen\_pyrfrac & Pyramid with triangular pyramid-shaped holes.\\
 \bottomrule

### 2025-10-05 paper-cardinalR.Rmd
Removed Cube projections

@@
 
-A cube structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube)"} else { "Figure \\@ref(fig:cube-proj)"})`) represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
 We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
@@
 
-```{r}
-gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r gridcube-proj1}
-scaled_data <- scale_data_manual(gridcube)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.7, 
-                                              axis_pos_y = -0.7, 
-                                              threshold = 0.022))
-
-gridcube_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.83, 0.8), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridcube-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.8,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.85, 
-                                              axis_pos_y = -0.85, 
-                                              threshold = 0.05))
-
-gridcube_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 1), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridcube-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.7, 
-                                              axis_pos_y = -0.7, 
-                                              threshold = 0.016))
-
-gridcube_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
 An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
@@
 unifcube <- gen_unifcube(n = 1000, p = 4)
-```
-
-```{r}
-unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r unifcube-proj1}
-scaled_data <- scale_data_manual(unifcube)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.022))
-
-unifcube_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.8), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifcube-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.04))
-
-unifcube_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.9, 0.8), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifcube-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.016))
-
-unifcube_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
 ```
@@
   column_spec(2, width = "8cm")
-```
-
-```{r cube, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `gridcube`, `unifcube`, and `cubehole` data in $4\\text{-}D$."}
-
-cubefig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    gridcube_lang, unifcube_lang
-  ),
-  device = "xs"
-)
-
-class(cubefig) <- c(class(cubefig), "htmlwidget")
-
-cubefig
-```
-
-```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=12, fig.height=8}
-
-gridcube_proj1 + gridcube_proj2 + gridcube_proj3 +
-unifcube_proj1 + unifcube_proj2 + unifcube_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-10-05 paper-cardinalR.tex
Removed Cube projections

@@
 
-A cube structure (Figure \ref{fig:cube-proj}) represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
 We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
@@
 \end{table}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cube-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.}\label{fig:cube-proj}
-\end{figure}
 

### 2025-10-05 paper-cardinalR.R

@@ -388,180 +388,8 @@ orgcurvybranches_proj1 + orgcurvybranches_proj2 + orgcurvybranches_proj3 +
 gridcube <- gen_gridcube(n = 1000, p = 4)
 
 
-## -----------------------------------------------------------------------------
-gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-
-
-## ----gridcube-proj1-----------------------------------------------------------
-scaled_data <- scale_data_manual(gridcube)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.7, 
-                                              axis_pos_y = -0.7, 
-                                              threshold = 0.022))
-
-gridcube_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.83, 0.8), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-
-
-## ----gridcube-proj2-----------------------------------------------------------
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.8,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.85, 
-                                              axis_pos_y = -0.85, 
-                                              threshold = 0.05))
-
-gridcube_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 1), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
[…]

### 2025-10-05 paper-cardinalR.Rmd
Removed S-curve and Swiss roll projections (these are common benchmark strutures)

@@
 
-The S-curve structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve)"} else { "Figure \\@ref(fig:scurve-proj)"})`) represents a smooth, non-linear manifold in 3D space.
+The S-curve structure represents a smooth, non-linear manifold in 3D space.
 Using `gen_scurve(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`), the geometry is defined by a latent parameter $\theta \sim U(-3\pi/2, 3\pi/2)$, which controls the curvature of the manifold.
@@
 scurve <- gen_scurve(n = 1000)
-```
-
-```{r}
-scurve_lang <- langevitour::langevitour(scurve, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r scurve-proj1}
-scaled_data <- scale_data_manual(scurve)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097),
-  c(0.00896,0.52496,0.12597))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-scurve_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r scurve-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278),
-  c(-0.50493,0.14388,0.14615))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-scurve_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r scurve-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416),
-  c(0.50329,0.07199,-0.16064))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-scurve_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r scurve, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `scurve` and `scurvehole` data in $4\\text{-}D$."}
-
-scurvefig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    scurve_lang
-  ),
-  device = "xs"
-)
-
-class(scurvefig) <- c(class(scurvefig), "htmlwidget")
-
-scurvefig
-```
-
-```{r scurve-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.", fig.width=12, fig.height=4}
-
-scurve_proj1 + scurve_proj2 + scurve_proj3 + 
-  plot_layout(ncol = 3, guides = "collect") 
 ```
@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:swissroll)"} else { "Figure \\@ref(fig:swissroll-proj)"})`).
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension.
 The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.
@@
 swissroll <- gen_swissroll(n = 1000, w = c(-1, 1))
-```
-
-```{r}
-swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r swissroll-proj1}
-scaled_data <- scale_data_manual(swissroll)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097),
-  c(0.00896,0.52496,0.12597))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-swissroll_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r swissroll-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278),
-  c(-0.50493,0.14388,0.14615))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-swissroll_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r swissroll-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416),
-  c(0.50329,0.07199,-0.16064))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-swissroll_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r swissroll, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `swissroll` data in $3\\text{-}D$."}
-
-swissrollfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    swissroll_lang
-  ),
-  device = "xs"
-)
-
-class(swissrollfig) <- c(class(swissrollfig), "htmlwidget")
-
-swissrollfig
-```
-
-```{r swissroll-proj, eval=knitr::is_latex_output(), fig.pos="H", fig.cap="Three $2\\text{-}D$ projections from $3\\text{-}D$, for the `swissroll` data."}
-#| fig-width: 12
-#| fig-height: 4
-
-swissroll_proj1 + swissroll_proj2 + swissroll_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
 ```

### 2025-10-05 paper-cardinalR.tex
Removed S-curve and Swiss roll projections (these are common benchmark strutures)

@@
 
-The S-curve structure (Figure \ref{fig:scurve-proj}) represents a smooth, non-linear manifold in 3D space.
+The S-curve structure represents a smooth, non-linear manifold in 3D space.
 Using \texttt{gen\_scurve(n,\ p)} (Figure \ref{fig:scurve-proj} a), the geometry is defined by a latent parameter \(\theta \sim U(-3\pi/2, 3\pi/2)\), which controls the curvature of the manifold.
@@
 \end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/scurve-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `scurve` (a1-a3) and `scurvehole` (b1-b3) data.}\label{fig:scurve-proj}
-\end{figure}
 
@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension (Figure \ref{fig:swissroll-proj}).
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension.
 The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.
@@
 \end{verbatim}
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/swissroll-proj-1} \caption{Three $2\text{-}D$ projections from $3\text{-}D$, for the `swissroll` data.}\label{fig:swissroll-proj}
-\end{figure}
 

### 2025-10-05 paper-cardinalR.Rmd
Updated Table 3

@@
 arg_branching_tb <- tibble(arg = c("n",
-                          "p",
                           "k"), 
-                        exp = c("A numeric value representing the number of points.",
-                                "A numeric value representing the number of dimensions.",
-                                "A numeric value representing the number of clusters."))
+                          type = c("integer", "integer"),
+                        exp = c("Number of points.",
+                                "Number of clusters."))
 ```
@@
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Explanation"), table.pos = "H") 
+  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), table.pos = "H") 
 ```
@@
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Explanation"), booktabs = T, table.pos = "H")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
+  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
+  column_spec(1, width = "1cm") |>
+  column_spec(1, width = "2cm") |>
+  column_spec(2, width = "3cm")
 ```
@@
   kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
+  column_spec(1, width = "3cm") |>
+  column_spec(2, width = "9cm")
 ```

### 2025-10-05 paper-cardinalR.Rmd
Updated table 10

@@
   kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
-  column_spec(1, width = "3cm") |>
-  column_spec(2, width = "9cm")
+  column_spec(1, width = "2.5cm") |>
+  column_spec(2, width = "10.5cm")
 ```

### 2025-10-05 paper-cardinalR.Rmd
Updated branching section

@@
 
-The simplest structures are approximately linear branches, generated by the `gen_linearbranches(n, p, k)` function (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) b"} else { "Figure \\@ref(fig:branch-proj) b"})`).
+The simplest structures are approximately linear branches in \gD{}, generated by the `gen_linearbranches(n, k)` function.
 These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability.
@@
 
-To introduce curvature, the `gen_curvybranches(n, p, k)` function generates $k$ curvilinear branches in $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) c"} else { "Figure \\@ref(fig:branch-proj) c"})`).
-Branches $1$ and $2$ are simple parabolas defined as
+To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in \gD{}. Branches $1$ and $2$ are simple parabolas defined as
 
@@
 An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
-The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cube) b"} else { "Figure \\@ref(fig:cube-proj) b"})`).
+The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube.
 
@@
 The S-curve structure represents a smooth, non-linear manifold in 3D space.
-Using `gen_scurve(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve) a"} else { "Figure \\@ref(fig:scurve-proj) a"})`), the geometry is defined by a latent parameter $\theta \sim U(-3\pi/2, 3\pi/2)$, which controls the curvature of the manifold.
+Using `gen_scurve(n, p)`, the geometry is defined by a latent parameter $\theta \sim U(-3\pi/2, 3\pi/2)$, which controls the curvature of the manifold.
 The first three coordinates form the S-shape:

### 2025-10-05 paper-cardinalR.tex
Updated Table 3
Updated table 10
Updated branching section

@@
 \centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{3cm}l}
 \toprule
-Argument & Explanation\\
+Argument & Type & Explanation\\
 \midrule
-n & A numeric value representing the number of points.\\
-p & A numeric value representing the number of dimensions.\\
-k & A numeric value representing the number of clusters.\\
+n & integer & Number of points.\\
+k & integer & Number of clusters.\\
 \bottomrule
@@
 
-The simplest structures are approximately linear branches, generated by the \texttt{gen\_linearbranches(n,\ p,\ k)} function (Figure \ref{fig:branch-proj} b).
+The simplest structures are approximately linear branches in $2\text{-}D${}, generated by the \texttt{gen\_linearbranches(n,\ k)} function.
 These consist of \(k\) short line segments in the first two dimensions, with added jitter to simulate variability.
@@
 
-To introduce curvature, the \texttt{gen\_curvybranches(n,\ p,\ k)} function generates \(k\) curvilinear branches in \(p\text{-}D\) (Figure \ref{fig:branch-proj} c).
-Branches \(1\) and \(2\) are simple parabolas defined as
+To introduce curvature, the \texttt{gen\_curvybranches(n,\ k)} function generates \(k\) curvilinear branches in $2\text{-}D${}. Branches \(1\) and \(2\) are simple parabolas defined as
 
@@
 An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
-The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube (Figure \ref{fig:cube-proj} b).
+The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube.
 
@@
 The S-curve structure represents a smooth, non-linear manifold in 3D space.
-Using \texttt{gen\_scurve(n,\ p)} (Figure \ref{fig:scurve-proj} a), the geometry is defined by a latent parameter \(\theta \sim U(-3\pi/2, 3\pi/2)\), which controls the curvature of the manifold.
+Using \texttt{gen\_scurve(n,\ p)}, the geometry is defined by a latent parameter \(\theta \sim U(-3\pi/2, 3\pi/2)\), which controls the curvature of the manifold.
 The first three coordinates form the S-shape:
@@
 \centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{10.5cm}}
 \toprule

### 2025-10-05 paper-cardinalR.Rmd
Add cone before cube in ordering

@@
 
-The `gen_expbranches(n, p, k)` function creates $k$ exponential branches in $2\text{-}D$, radiating from a central region (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) a"} else { "Figure \\@ref(fig:branch-proj) a"})`).
-Each branch $i$ is defined as
+The `gen_expbranches(n, k)` function creates $k$ exponential branches in \gD{}, radiating from a central region. Each branch $i$ is defined as
 
@@
 
-#### Cube
-
-A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
-We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
-These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:cube-tb-html)', '\\@ref(tab:cube-tb-pdf)'))` outlines these functions and their purposes.
-
-The first is the regular grid of points of $n$ points in $p$ dimensions.
-This is generated using `gen_gridcube(n, p)`.
-The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to $n$.
-Each dimension is then normalized to lie in the interval $[0,1]$, so that the resulting structure forms a true $p\text{-}D$ hypercube.
-This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
-
-```{r data-gridcube, echo=TRUE}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-```
-
-An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
-The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube.
-
-```{r data-unifcube, echo=TRUE}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-```
-
-```{r}
-cube_tb <- tibble(fun = c("gen_gridcube",
-                          "gen_unifcube"), 
-                      exp = c("Cube with specified grid points along each axes.",
-                              "Cube with uniform points."))
-```
-
-```{r cube-tb-html, eval=knitr::is_html_output()}
-cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
-```
-
-```{r cube-tb-pdf, eval=knitr::is_latex_output()}
-cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
 #### Cone
@@
   plot_layout(ncol = 3, guides = "collect") 
+```
+
+#### Cube
+
+A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
+We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
+These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:cube-tb-html)', '\\@ref(tab:cube-tb-pdf)'))` outlines these functions and their purposes.
+
+The first is the regular grid of points of $n$ points in $p$ dimensions.
+This is generated using `gen_gridcube(n, p)`.
+The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to $n$.
+Each dimension is then normalized to lie in the interval $[0,1]$, so that the resulting structure forms a true $p\text{-}D$ hypercube.
+This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
+
+```{r data-gridcube, echo=TRUE}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+```
+
+An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
+The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube.
+
+```{r data-unifcube, echo=TRUE}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+```
+
+```{r}
+cube_tb <- tibble(fun = c("gen_gridcube",
+                          "gen_unifcube"), 
+                      exp = c("Cube with specified grid points along each axes.",
+                              "Cube with uniform points."))
+```
+
+```{r cube-tb-html, eval=knitr::is_html_output()}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
+```
+
+```{r cube-tb-pdf, eval=knitr::is_latex_output()}
+cube_tb |> 
+  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
 ```

### 2025-10-05 paper-cardinalR.tex
Add cone before cube in ordering

@@
 
-The \texttt{gen\_expbranches(n,\ p,\ k)} function creates \(k\) exponential branches in \(2\text{-}D\), radiating from a central region (Figure \ref{fig:branch-proj} a).
-Each branch \(i\) is defined as
+The \texttt{gen\_expbranches(n,\ k)} function creates \(k\) exponential branches in $2\text{-}D${}, radiating from a central region. Each branch \(i\) is defined as
 
@@
 
+\paragraph{Cone}\label{cone}
+
+To simulate a cone-shaped structure in arbitrary dimensions (Figure \ref{fig:cone-proj}), we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+
+This function generates \(n\) points in \(p\text{-}D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip.
+Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), capped at the cone height \(h\), producing a higher density of points near the tip.
+At each height \(X_p\), the radius of the cross-section decreases linearly from base to tip according to \(r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h\), where \(r_{\text{min}} = \text{ratio}\) and \(r_{\text{max}} = 1\).
+
+For each point, a direction in the first \(p-1\) dimensions is sampled uniformly on a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates.
+The radial coordinates are scaled by the height-dependent radius \(r\), producing the conical taper.
+In three dimensions (\(p = 3\)), this results in a classical \(3\text{-}D\) cone, while for \(p > 3\), additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
+
+\begin{verbatim}
+cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
+\end{verbatim}
+
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:cone-proj}
+\end{figure}
+
 \paragraph{Cube}\label{cube}
@@
 \end{table}
-
-\paragraph{Cone}\label{cone}
-
-To simulate a cone-shaped structure in arbitrary dimensions (Figure \ref{fig:cone-proj}), we define a function \texttt{gen\_cone(n,\ p,\ h,\ ratio)}, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
-
-This function generates \(n\) points in \(p\text{-}D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip.
-Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), capped at the cone height \(h\), producing a higher density of points near the tip.
-At each height \(X_p\), the radius of the cross-section decreases linearly from base to tip according to \(r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h\), where \(r_{\text{min}} = \text{ratio}\) and \(r_{\text{max}} = 1\).
-
-For each point, a direction in the first \(p-1\) dimensions is sampled uniformly on a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates.
-The radial coordinates are scaled by the height-dependent radius \(r\), producing the conical taper.
-In three dimensions (\(p = 3\)), this results in a classical \(3\text{-}D\) cone, while for \(p > 3\), additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
-
-\begin{verbatim}
-cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
-\end{verbatim}
-
-\begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:cone-proj}
-\end{figure}
 

### 2025-10-05 paper-cardinalR.R

@@ -384,33 +384,6 @@ orgcurvybranches_proj1 + orgcurvybranches_proj2 + orgcurvybranches_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
 
 
-## ----data-gridcube, echo=TRUE-------------------------------------------------
-gridcube <- gen_gridcube(n = 1000, p = 4)
-
-
-## ----data-unifcube, echo=TRUE-------------------------------------------------
-unifcube <- gen_unifcube(n = 1000, p = 4)
-
-
-## -----------------------------------------------------------------------------
-cube_tb <- tibble(fun = c("gen_gridcube",
-                          "gen_unifcube"), 
-                      exp = c("Cube with specified grid points along each axes.",
-                              "Cube with uniform points."))
-
-
-## ----cube-tb-html, eval=knitr::is_html_output()-------------------------------
-# cube_tb |>
-#   kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), table.pos = "H")
-
-
-## ----cube-tb-pdf, eval=knitr::is_latex_output()-------------------------------
-cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-
-
 ## ----data-bluntedcone, echo=TRUE----------------------------------------------
 cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
 
@@ -522,6 +495,33 @@ cone_proj3 <- plot_proj(
 
 cone_proj1 + cone_proj2 + cone_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
+
+
+## ----data-gridcube, echo=TRUE-------------------------------------------------
+gridcube <- gen_gridcube(n = 1000, p = 4)
+
+
+## ----data-unifcube, echo=TRUE-------------------------------------------------
+unifcube <- gen_unifcube(n = 1000, p = 4)
+
+
+## -----------------------------------------------------------------------------
+cube_tb <- tibble(fun = c("gen_gridcube",
+                          "gen_unifcube"), 
+                      exp = c("Cube with specified grid points along each axes.",
+                              "Cube with uniform points."))
+
+
+## ----cube-tb-html, eval=knitr::is_html_output()-------------------------------
+# cube_tb |>
[…]

### 2025-10-05 paper-cardinalR.Rmd
Removed on of branching projections

@@
 
-High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` and `gen_orgcurvybranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch) d-e"} else { "Figure \\@ref(fig:branch-proj) d-e"})`).
+High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) and `gen_orgcurvybranches(n, p, k)`.
 Each branch is embedded in a unique or repeated $2\text{-}D$ subspace of the $p\text{-}D$ space.
@@
   plot_limits = c(-0.65, 0.7), 
-  title = "d1", 
+  title = "a1", 
   cex = 2, 
@@
   plot_limits = c(-0.7, 0.6), 
-  title = "d2", 
+  title = "a2", 
   cex = 2, 
@@
   plot_limits = c(-0.5, 0.62), 
-  title = "d3", 
+  title = "a3", 
   cex = 2, 
@@
 
-```{r}
-orgcurvybranches_lang <- langevitour::langevitour(orgcurvybranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r orgcurvybranches-proj1}
-scaled_data <- scale_data_manual(orgcurvybranches)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
-                                              threshold = 0.022))
-
-orgcurvybranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.4, 0.5), 
-  title = "e1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r orgcurvybranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-orgcurvybranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "e2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r orgcurvybranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-orgcurvybranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
-  title = "e3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `expbranches`, `linearbranches`, `curvybranches`, `orglinearbranches`, and `orgcurvybranches` data in $4\\text{-}D$."}
+```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `orglinearbranches` data in $4\\text{-}D$."}
 
@@
     style = "display: grid; grid-template-columns: 1fr 1fr;",
-    orglinearbranches_lang, orgcurvybranches_lang
+    orglinearbranches_lang
   ),
@@
 
-```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.", fig.width=12, fig.height=8}
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` (a1-a3) data.", fig.width=12, fig.height=4}
 
 orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
-orgcurvybranches_proj1 + orgcurvybranches_proj2 + orgcurvybranches_proj3 +
   plot_layout(ncol = 3, guides = "collect") 

### 2025-10-05 paper-cardinalR.tex
Removed on of branching projections

@@
 
-High-dimensional generalizations are provided by \texttt{gen\_orglinearbranches(n,\ p,\ k)} and \texttt{gen\_orgcurvybranches(n,\ p,\ k)} (Figure \ref{fig:branch-proj} d-e).
+High-dimensional generalizations are provided by \texttt{gen\_orglinearbranches(n,\ p,\ k)} (Figure \ref{fig:branch-proj}) and \texttt{gen\_orgcurvybranches(n,\ p,\ k)}.
 Each branch is embedded in a unique or repeated \(2\text{-}D\) subspace of the \(p\text{-}D\) space.
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `expbranches` (a1-a3), `linearbranches` (b1-b3), `curvybranches` (c1-c3), `orglinearbranches` (d1-d3), `orgcurvybranches` (e1-e3) data.}\label{fig:branch-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `orglinearbranches` (a1-a3) data.}\label{fig:branch-proj}
 \end{figure}

### 2025-10-06 paper-cardinalR.Rmd
Added wrapper information for cube

@@
 
-A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
-We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
-These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
+A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, and boundary properties in high dimensions.
+We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, and uniform random points.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:cube-tb-html)', '\\@ref(tab:cube-tb-pdf)'))` outlines these functions and their purposes.
-
-The first is the regular grid of points of $n$ points in $p$ dimensions.
-This is generated using `gen_gridcube(n, p)`.
-The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to $n$.
-Each dimension is then normalized to lie in the interval $[0,1]$, so that the resulting structure forms a true $p\text{-}D$ hypercube.
-This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
-
-```{r data-gridcube, echo=TRUE}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-```
-
-An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
-The function `gen_unifcube(n, p)` is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the $p\text{-}D$ cube.
-
-```{r data-unifcube, echo=TRUE}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-```
 
@@
 ```
+
+The function `gen_gridcube(n, p)` is a wrapper around `geozoo::cube.solid.grid()`.
+It generates a regular lattice of points in \pD{}, producing a uniform hypercube grid.
+Each axis contains equally spaced coordinates, resulting in a well-defined geometric structure.
+
+```{r data-gridcube, echo=TRUE}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+```
+
+By contrast, `gen_unifcube(n, p)` wraps `geozoo::cube.solid.random()`, producing uniformly distributed points within a $p\text{-}D$ cube.
+To avoid including the cube’s vertices, these points are removed after generation.
+This results in a hypercube filled with random, but evenly distributed, samples rather than structured lattice points.
+
+```{r data-unifcube, echo=TRUE}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+```
+
+Such cube-based structures are commonly used as benchmarks in Monte Carlo sampling, computational geometry, and density estimation, where assessing how algorithms behave under uniform or grid-like distributions is critical [@devroye1986; @niederreiter1992].
 

### 2025-10-06 paper-cardinalR.tex
Added wrapper information for cube

@@
 
-A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, spacing, and boundary properties in high dimensions.
-We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, uniform random points, and cubes with missing regions or holes.
-These structures are valuable for testing the ability of algorithms to maintain uniform spacing or to detect gaps in the data.
+A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, and boundary properties in high dimensions.
+We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, and uniform random points.
 Table \ref{tab:cube-tb-pdf} outlines these functions and their purposes.
-
-The first is the regular grid of points of \(n\) points in \(p\) dimensions.
-This is generated using \texttt{gen\_gridcube(n,\ p)}.
-The number of grid points along each axis is determined by finding the nearest integer factors whose product is close to \(n\).
-Each dimension is then normalized to lie in the interval \([0,1]\), so that the resulting structure forms a true \(p\text{-}D\) hypercube.
-This produces a lattice of evenly spaced points along all axes, providing a uniform and interpretable high-dimensional grid.
-
-\begin{verbatim}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-\end{verbatim}
-
-An extension to the regular grid of points is to consider the points being uniformly distributed along each axis, as opposed to evenly spaced.
-The function \texttt{gen\_unifcube(n,\ p)} is identical to the regular grid of points, except instead of points being placed in integer grid coordinates, they are placed at a uniformly distributed point inside the \(p\text{-}D\) cube.
-
-\begin{verbatim}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-\end{verbatim}
 
@@
 \end{table}
+
+The function \texttt{gen\_gridcube(n,\ p)} is a wrapper around \texttt{geozoo::cube.solid.grid()}.
+It generates a regular lattice of points in $p\text{-}D${}, producing a uniform hypercube grid.
+Each axis contains equally spaced coordinates, resulting in a well-defined geometric structure.
+
+\begin{verbatim}
+gridcube <- gen_gridcube(n = 1000, p = 4)
+\end{verbatim}
+
+By contrast, \texttt{gen\_unifcube(n,\ p)} wraps \texttt{geozoo::cube.solid.random()}, producing uniformly distributed points within a \(p\text{-}D\) cube.
+To avoid including the cube's vertices, these points are removed after generation.
+This results in a hypercube filled with random, but evenly distributed, samples rather than structured lattice points.
+
+\begin{verbatim}
+unifcube <- gen_unifcube(n = 1000, p = 4)
+\end{verbatim}
+
+Such cube-based structures are commonly used as benchmarks in Monte Carlo sampling, computational geometry, and density estimation, where assessing how algorithms behave under uniform or grid-like distributions is critical \citep{devroye1986, niederreiter1992}.
 

### 2025-10-06 paper-cardinalR.R

@@ -410,14 +410,6 @@ cone_proj1 + cone_proj2 + cone_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
 
 
-## ----data-gridcube, echo=TRUE-------------------------------------------------
-gridcube <- gen_gridcube(n = 1000, p = 4)
-
-
-## ----data-unifcube, echo=TRUE-------------------------------------------------
-unifcube <- gen_unifcube(n = 1000, p = 4)
-
-
 ## -----------------------------------------------------------------------------
 cube_tb <- tibble(fun = c("gen_gridcube",
                           "gen_unifcube"), 
@@ -435,6 +427,14 @@ cube_tb |>
   kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "4cm") |>
   column_spec(2, width = "8cm")
+
+
+## ----data-gridcube, echo=TRUE-------------------------------------------------
+gridcube <- gen_gridcube(n = 1000, p = 4)
+
+
+## ----data-unifcube, echo=TRUE-------------------------------------------------
+unifcube <- gen_unifcube(n = 1000, p = 4)
 
 
 ## ----data-gau, echo=TRUE------------------------------------------------------

### 2025-10-06 paper-cardinalR.Rmd
Added application for linear

@@
 
+Gaussian clouds are common benchmark structures in statistics and machine learning, used in clustering, classification, and anomaly detection, with applications in image segmentation, speech recognition, and forensic analysis [@geoffrey2000].
+
 #### Linear
@@
 
+This structure appears in \pD{} data when variation is driven by a single factor, such as time-course or sensor measurements, providing a useful test case for trajectory and regression methods [@trapnell2014].
+
 #### Möbius
 
-The `gen_mobius()` function is a **wrapper** around `geozoo::mobius()`, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
+The `gen_mobius()` function is a wrapper around `geozoo::mobius()`, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
 The function returns a tibble with $n$ sampled points forming the surface of a Möbius strip.
@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension.
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, w)`, where $n$ is the number of points, and $w$ is the vertical range in the third dimension.
 The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.

### 2025-10-06 paper-cardinalR.tex
Added application for linear

@@
 
+Gaussian clouds are common benchmark structures in statistics and machine learning, used in clustering, classification, and anomaly detection, with applications in image segmentation, speech recognition, and forensic analysis \citep{geoffrey2000}.
+
 \paragraph{Linear}\label{linear}
@@
 
+This structure appears in $p\text{-}D${} data when variation is driven by a single factor, such as time-course or sensor measurements, providing a useful test case for trajectory and regression methods \citep{trapnell2014}.
+
 \paragraph{Möbius}\label{muxf6bius}
 
-The \texttt{gen\_mobius()} function is a \textbf{wrapper} around \texttt{geozoo::mobius()}, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
+The \texttt{gen\_mobius()} function is a wrapper around \texttt{geozoo::mobius()}, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
 The function returns a tibble with \(n\) sampled points forming the surface of a Möbius strip.
@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension.
+To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ w)}, where \(n\) is the number of points, and \(w\) is the vertical range in the third dimension.
 The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.

### 2025-10-06 paper-cardinalR.Rmd
Added Mobius application

@@
 The function returns a tibble with $n$ sampled points forming the surface of a Möbius strip.
-
-<!-- The Möbius strip is a classical non-orientable surface, looping back on itself with a half-twist.
-Its construction is based on a parameterization with angle $\theta \sim U(0, 2\pi)$ (position along the loop) and width $w \sim U(-0.4, 0.4)$ (offset across the band).
-The Cartesian coordinates are -->
-
-<!-- $$ -->
-<!-- X_1 = \Big(1 + \tfrac{w}{2} \cos(\tfrac{\theta}{2})\Big) \cos(\theta), \quad -->
-<!-- X_2 = \Big(1 + \tfrac{w}{2} \cos(\tfrac{\theta}{2})\Big) \sin(\theta), \quad -->
-<!-- X_3 = \tfrac{w}{2}\sin(\tfrac{\theta}{2}), -->
-<!-- $$ -->
-
-<!-- mapping a two-dimensional band with a half-twist into $\mathbb{R}^3$ as a one-sided, non-orientable surface. -->
-
-<!-- While `gen_mobius()` directly produces a 3-D structure, the package framework allows embedding this strip in higher-dimensional spaces by appending additional Gaussian noise variables.
-This enables the Möbius strip to be used in the same comparative setting as other high-dimensional synthetic structures.
--->
-
+e
 ```{r data-mobius, echo=TRUE}
@@
 ```
+
+The Möbius strip structure can model twisted or cyclic surfaces in physics and engineering, such as conveyor belts, molecular structures, or optical systems with non-orientable geometries [@optica2023].
 

### 2025-10-06 paper-cardinalR.Rmd
Updated Pyramid section

@@
 
-A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
+A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in \pD{} space.
 The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
@@
 Let $X_1, \dots, X_p$ denote the coordinates of the generated points.
-For the rectangular and triangular based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
-That is, $$X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).$$ This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+For the rectangular, triangular, and star-shaped based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
+That is, $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).$ This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
 For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
@@
 That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
-The first three coordinates are then defined as: $$X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).$$
+The first three coordinates are then defined as $X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).$
 
@@
 That is, $r(z) = r_b\left(1 - z/h\right)$.
-
-Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
-$r_{\text{point}} \sim \sqrt{U(0, 1)}$. Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
+Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$, $r_{\text{point}} \sim \sqrt{U(0, 1)}$.
+Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
 
@@
 
-For all the above pyramid shapes, if $p > 3$, the remaining $p - 3$ dimensions (i.e., $X_4$ to $X_{p-1}$) are additional noise.
+For rectangular and triangular pyramids, the remaining dimensions $X_4$ to $X_{p-1}$, and for star-based pyramids $X_3$ to $X_{p-1}$, are treated as noise.
 
@@
 Let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
-Coordinates in the first two dimensions represent a perfect circle on the plane: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta)$$.
-For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
+Coordinates in the first two dimensions represent a perfect circle on the plane: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$ For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
 The first component is a scaling factor that decreases with the dimension index, defined as $\text{scale}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$.
@@
 This results in a hollow shell-like structure with no interior points.
-
-For example, when $p=3$, `gen_unifsphere()` produces a solid ball in 3D space, whereas `gen_hollowsphere()` produces only the spherical boundary.
+For example, when $p=3$, `gen_unifsphere()` produces a solid ball in $3\text{-}D$ space, whereas `gen_hollowsphere()` produces only the spherical boundary.
 These paired structures allow controlled experiments to investigate how algorithms behave when data is concentrated throughout the full volume versus constrained to the boundary.

### 2025-10-06 paper-cardinalR.tex
Added Mobius application
Updated Pyramid section

@@
 The function returns a tibble with \(n\) sampled points forming the surface of a Möbius strip.
+e
 
@@
 \end{verbatim}
+
+The Möbius strip structure can model twisted or cyclic surfaces in physics and engineering, such as conveyor belts, molecular structures, or optical systems with non-orientable geometries \citep{optica2023}.
 
@@
 
-A pyramid structure (Figure \ref{fig:pyr-proj}) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in high-dimensional space.
+A pyramid structure (Figure \ref{fig:pyr-proj}) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in $p\text{-}D${} space.
 The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
@@
 Let \(X_1, \dots, X_p\) denote the coordinates of the generated points.
-For the rectangular and triangular based pyramid generator functions, the final dimension, \(X_p\), encodes the height of each point and is drawn from an exponential distribution capped at the maximum height \(h\).
-That is, \[X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).\] This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+For the rectangular, triangular, and star-shaped based pyramid generator functions, the final dimension, \(X_p\), encodes the height of each point and is drawn from an exponential distribution capped at the maximum height \(h\).
+That is, \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).\) This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
 For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
@@
 That is, \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
-The first three coordinates are then defined as: \[X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).\]
+The first three coordinates are then defined as \(X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).\)
 
@@
 That is, \(r(z) = r_b\left(1 - z/h\right)\).
-
-Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
-\(r_{\text{point}} \sim \sqrt{U(0, 1)}\). Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
+Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\), \(r_{\text{point}} \sim \sqrt{U(0, 1)}\).
+Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
 
@@
 
-For all the above pyramid shapes, if \(p > 3\), the remaining \(p - 3\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are additional noise.
+For rectangular and triangular pyramids, the remaining dimensions \(X_4\) to \(X_{p-1}\), and for star-based pyramids \(X_3\) to \(X_{p-1}\), are treated as noise.
 
@@
 Let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
-Coordinates in the first two dimensions represent a perfect circle on the plane: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta)\].
-For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
+Coordinates in the first two dimensions represent a perfect circle on the plane: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta).\] For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
 The first component is a scaling factor that decreases with the dimension index, defined as \(\text{scale}_j = \sqrt{(0.5)^{j-2}}\) for \(j = 3, \dots, p\).
@@
 This results in a hollow shell-like structure with no interior points.
-
-For example, when \(p=3\), \texttt{gen\_unifsphere()} produces a solid ball in 3D space, whereas \texttt{gen\_hollowsphere()} produces only the spherical boundary.
+For example, when \(p=3\), \texttt{gen\_unifsphere()} produces a solid ball in \(3\text{-}D\) space, whereas \texttt{gen\_hollowsphere()} produces only the spherical boundary.
 These paired structures allow controlled experiments to investigate how algorithms behave when data is concentrated throughout the full volume versus constrained to the boundary.

### 2025-10-06 paper-cardinalR.Rmd
Added the pyramid applications
Updated S-curve section

@@
 Extra arguments (`...`) can be passed to cluster generators, allowing further control over per-cluster characteristics like radius of the sphere.
-
 The main arguments of the `gen_multicluster()` function are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:main-tb-html)', '\\@ref(tab:main-tb-pdf)'))`.
@@
 The function returns a tibble with $n$ sampled points forming the surface of a Möbius strip.
-e
+
 ```{r data-mobius, echo=TRUE}
@@
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:pyramid-tb-html)', '\\@ref(tab:pyramid-tb-pdf)'))` summarizes these functions.
-
-Let $X_1, \dots, X_p$ denote the coordinates of the generated points.
-For the rectangular, triangular, and star-shaped based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
-That is, $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).$ This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
-For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
-That is, $X_p = z \sim U(0, h)$.
-
-The remaining dimensions are based on the specific pyramid shape.
-For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`), let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$.
-That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
-The first three coordinates are then defined as $X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).$
-
-```{r data-pyrrect, echo=TRUE}
-pyrrect <- gen_pyrrect(n = 1000, p = 4)
-```
-
-```{r}
-pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r pyrrect-proj1}
-scaled_data <- scale_data_manual(pyrrect)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrrect_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrrect-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.7, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.08))
-
-pyrrect_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrrect-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-pyrrect_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.75), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the triangular based pyramid, `gen_pyrtri(n, p, h, l, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`), let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$.
-That is, $r(z) = r_t + (l-r_t)z/h$.
-A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
-The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v.$
-
-```{r data-pyrtri, echo=TRUE}
-pyrtri <- gen_pyrtri(n = 1000, p = 4)
-```
-
-```{r}
-pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r pyrtri-proj1}
-scaled_data <- scale_data_manual(pyrtri)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.022))
-
-pyrtri_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.3), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrtri-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.03))
-
-pyrtri_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.4), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrtri-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-pyrtri_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the star based pyramid, `gen_pyrstar(n, p, h, rb)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`), let the radius at height $z$, $r(z)$, be such that the radius scales linearly from zero (tip) to the base radius $r_b$.
-That is, $r(z) = r_b\left(1 - z/h\right)$.
-Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$, $r_{\text{point}} \sim \sqrt{U(0, 1)}$.
-Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
-
-```{r data-pyrstar, echo=TRUE}
-pyrstar <- gen_pyrstar(n = 1000, p = 4)
-```
-
-```{r}
-pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r pyrstar-proj1}
-scaled_data <- scale_data_manual(pyrstar)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrstar_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrstar-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-pyrstar_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.4), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrstar-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-pyrstar_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For rectangular and triangular pyramids, the remaining dimensions $X_4$ to $X_{p-1}$, and for star-based pyramids $X_3$ to $X_{p-1}$, are treated as noise.
-
-Finally, for the Sierpinski-like pyramid, `gen_pyrfrac(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`), let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
-The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
-Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
-At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
-This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
-The points remain bounded inside the convex hull of the simplex.
-The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
-
-```{r data-pyrholes, echo=TRUE}
-pyrholes <- gen_pyrfrac(n = 1000, p = 4)
-```
-
-```{r}
-pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r pyrholes-proj1}
-scaled_data <- scale_data_manual(pyrholes)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-pyrholes_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrholes-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-pyrholes_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.75), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrholes-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-pyrholes_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
 
@@
 
+Let $X_1, \dots, X_p$ denote the coordinates of the generated points.
+For the rectangular, triangular, and star-shaped based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
+That is, $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).$ This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
+For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
+That is, $X_p = z \sim U(0, h)$.
+
+The remaining dimensions are based on the specific pyramid shape.
+For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`), let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$.
+That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
+The first three coordinates are then defined as $X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).$
+
+```{r data-pyrrect, echo=TRUE}
+pyrrect <- gen_pyrrect(n = 1000, p = 4)
+```
+
+```{r}
+pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r pyrrect-proj1}
+scaled_data <- scale_data_manual(pyrrect)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrrect_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrrect-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.7, 
+                                              axis_pos_x = -0.45, 
+                                              axis_pos_y = -0.45, 
+                                              threshold = 0.08))
+
+pyrrect_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.6, 0.6), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrrect-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrrect_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.75), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For the triangular based pyramid, `gen_pyrtri(n, p, h, l, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`), let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$.
+That is, $r(z) = r_t + (l-r_t)z/h$.
+A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
+The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v.$
+
+```{r data-pyrtri, echo=TRUE}
+pyrtri <- gen_pyrtri(n = 1000, p = 4)
+```
+
+```{r}
+pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r pyrtri-proj1}
+scaled_data <- scale_data_manual(pyrtri)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.2, 
+                                              axis_pos_y = -0.2, 
+                                              threshold = 0.022))
+
+pyrtri_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.25, 0.3), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrtri-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.03))
+
+pyrtri_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.3, 0.4), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrtri-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.016))
+
+pyrtri_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.62), 
+  title = "b3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For the star based pyramid, `gen_pyrstar(n, p, h, rb)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`), let the radius at height $z$, $r(z)$, be such that the radius scales linearly from zero (tip) to the base radius $r_b$.
+That is, $r(z) = r_b\left(1 - z/h\right)$.
+Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$, $r_{\text{point}} \sim \sqrt{U(0, 1)}$.
+Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
+
+```{r data-pyrstar, echo=TRUE}
+pyrstar <- gen_pyrstar(n = 1000, p = 4)
+```
+
+```{r}
+pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r pyrstar-proj1}
+scaled_data <- scale_data_manual(pyrstar)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrstar_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.5), 
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrstar-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrstar_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.4), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrstar-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+pyrstar_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "c3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For rectangular and triangular pyramids, the remaining dimensions $X_4$ to $X_{p-1}$, and for star-based pyramids $X_3$ to $X_{p-1}$, are treated as noise.
+
+Finally, for the Sierpinski-like pyramid, `gen_pyrfrac(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`), let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
+The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
+Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
+At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
+This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
+The points remain bounded inside the convex hull of the simplex.
+The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
+
+```{r data-pyrholes, echo=TRUE}
+pyrholes <- gen_pyrfrac(n = 1000, p = 4)
+```
+
+```{r}
+pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r pyrholes-proj1}
+scaled_data <- scale_data_manual(pyrholes)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.022))
+
+pyrholes_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.65, 0.7), 
+  title = "d1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrholes-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.7,
+                                              axis_scaled = 3, 
+                                              axis_pos_x = -0.57, 
+                                              axis_pos_y = -0.57, 
+                                              threshold = 0.01))
+
+pyrholes_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.75), 
+  title = "d2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r pyrholes-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.2, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.6,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.016))
+
+pyrholes_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.7, 0.62), 
+  title = "d3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+Pyramid structures mimic tapering or layered geometries seen in architecture, crystals, and fractal-like natural patterns [@kirkby1983].
+
 ```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$."}
@@
 
-The S-curve structure represents a smooth, non-linear manifold in 3D space.
-Using `gen_scurve(n, p)`, the geometry is defined by a latent parameter $\theta \sim U(-3\pi/2, 3\pi/2)$, which controls the curvature of the manifold.
-The first three coordinates form the S-shape:
+The S-curve is a smooth, non-linear manifold in $3\text{-}D$ space. Using `gen_scurve(n)`, it is defined by
 
 $$
-X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1),
+X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1), \quad \theta \sim U(-3\pi/2, 3\pi/2).
 $$
 
-creating a horizontally curled pattern in $(X_1, X_3)$ with additional thickness along $X_2$.
+This follows the `s_curve()` function from snedata [@james2025], itself adapted from *scikit-learn*, but differs by returning a tibble with standardized names (`x1`, `x2`, `x3`), excluding the color variable, and omitting built-in noise (which can be added separately).
+S-curve is commonly used in manifold learning and dimension reduction as benchmarks for unfolding curved structure.
 

### 2025-10-06 paper-cardinalR.tex
Added the pyramid applications
Updated S-curve section

@@
 Extra arguments (\texttt{...}) can be passed to cluster generators, allowing further control over per-cluster characteristics like radius of the sphere.
-
 The main arguments of the \texttt{gen\_multicluster()} function are shown in Table \ref{tab:main-tb-pdf}.
@@
 The function returns a tibble with \(n\) sampled points forming the surface of a Möbius strip.
-e
 
@@
 
+\begin{table}
+
+\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Explanation\\
+\midrule
+gen\_pyrrect & Rectangular-base, with a sharp or blunted apex.\\
+gen\_pyrtri & Triangular-base, with a sharp or blunted apex.\\
+gen\_pyrstar & Star-shaped base, with a sharp or blunted apex.\\
+gen\_pyrfrac & Pyramid with triangular pyramid-shaped holes.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 Let \(X_1, \dots, X_p\) denote the coordinates of the generated points.
@@
 
-\begin{table}
-
-\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-gen\_pyrrect & Rectangular-base, with a sharp or blunted apex.\\
-gen\_pyrtri & Triangular-base, with a sharp or blunted apex.\\
-gen\_pyrstar & Star-shaped base, with a sharp or blunted apex.\\
-gen\_pyrfrac & Pyramid with triangular pyramid-shaped holes.\\
-\bottomrule
-\end{tabular}
-\end{table}
+Pyramid structures mimic tapering or layered geometries seen in architecture, crystals, and fractal-like natural patterns \citep{kirkby1983}.
 
@@
 
-The S-curve structure represents a smooth, non-linear manifold in 3D space.
-Using \texttt{gen\_scurve(n,\ p)}, the geometry is defined by a latent parameter \(\theta \sim U(-3\pi/2, 3\pi/2)\), which controls the curvature of the manifold.
-The first three coordinates form the S-shape:
+The S-curve is a smooth, non-linear manifold in \(3\text{-}D\) space. Using \texttt{gen\_scurve(n)}, it is defined by
 
 \[
-X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1),
+X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1), \quad \theta \sim U(-3\pi/2, 3\pi/2).
 \]
 
-creating a horizontally curled pattern in \((X_1, X_3)\) with additional thickness along \(X_2\).
+This follows the \texttt{s\_curve()} function from snedata \citep{james2025}, itself adapted from \emph{scikit-learn}, but differs by returning a tibble with standardized names (\texttt{x1}, \texttt{x2}, \texttt{x3}), excluding the color variable, and omitting built-in noise (which can be added separately).
+S-curve is commonly used in manifold learning and dimension reduction as benchmarks for unfolding curved structure.
 

### 2025-10-06 paper-cardinalR.Rmd
Updated trefoil bind version

@@
 
-The S-curve is a smooth, non-linear manifold in $3\text{-}D$ space. Using `gen_scurve(n)`, it is defined by
-
-$$
-X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1), \quad \theta \sim U(-3\pi/2, 3\pi/2).
-$$
+The S-curve is a smooth, non-linear manifold in $3\text{-}D$ space.
+Using `gen_scurve(n)`, it is defined by $X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1), \quad \theta \sim U(-3\pi/2, 3\pi/2).$
 
@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, w)`, where $n$ is the number of points, and $w$ is the vertical range in the third dimension.
-The first three dimensions form the classic $3\text{-}D$ Swiss roll shape.
-The $$X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \text{ where } t \sim U(0, 3\pi).$$
+The Swiss roll is a standard nonlinear manifold, representing a \gD{} plane curled into $3\text{-}D$.
+The `gen_swissroll(n, w)` generates points as $X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad t \sim U(0, 3\pi).$
 
@@
 ```
+
+Compared with `sndata::swiss_roll()` [@james2025], this implementation (i) samples $t$ over $[0, 3\pi]$ instead of $[1.5\pi, 4.5\pi]$, (ii) allows a flexible vertical range $w = (w_1, w_2)$ rather than fixing $z \in [0, z_{\max}]$, and (iii) returns a tibble with `x1, x2, x3` instead of adding a color variable.
+
+The Swiss roll is a classic benchmark for manifold learning, illustrating how a curved surface can be “unrolled” into lower dimensions.
+Similar spiral-like forms appear in galaxies, protein folding, and coiled materials [@dimitris2002].
 
@@
 
-For the $3\text{-}D$ stereographic projection, `gen_trefoil3d(n, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $$(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),$$ excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
+For the $3\text{-}D$ stereographic projection, `gen_trefoil3d(n, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),$ excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
 

### 2025-10-06 paper-cardinalR.tex
Updated trefoil bind version

@@
 
-The S-curve is a smooth, non-linear manifold in \(3\text{-}D\) space. Using \texttt{gen\_scurve(n)}, it is defined by
-
-\[
-X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1), \quad \theta \sim U(-3\pi/2, 3\pi/2).
-\]
+The S-curve is a smooth, non-linear manifold in \(3\text{-}D\) space.
+Using \texttt{gen\_scurve(n)}, it is defined by \(X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1), \quad \theta \sim U(-3\pi/2, 3\pi/2).\)
 
@@
 
-To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ w)}, where \(n\) is the number of points, and \(w\) is the vertical range in the third dimension.
-The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape.
-The \[X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \text{ where } t \sim U(0, 3\pi).\]
+The Swiss roll is a standard nonlinear manifold, representing a $2\text{-}D${} plane curled into \(3\text{-}D\).
+The \texttt{gen\_swissroll(n,\ w)} generates points as \(X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad t \sim U(0, 3\pi).\)
 
@@
 \end{verbatim}
+
+Compared with \texttt{sndata::swiss\_roll()} \citep{james2025}, this implementation (i) samples \(t\) over \([0, 3\pi]\) instead of \([1.5\pi, 4.5\pi]\), (ii) allows a flexible vertical range \(w = (w_1, w_2)\) rather than fixing \(z \in [0, z_{\max}]\), and (iii) returns a tibble with \texttt{x1,\ x2,\ x3} instead of adding a color variable.
+
+The Swiss roll is a classic benchmark for manifold learning, illustrating how a curved surface can be ``unrolled'' into lower dimensions.
+Similar spiral-like forms appear in galaxies, protein folding, and coiled materials \citep{dimitris2002}.
 
@@
 
-For the \(3\text{-}D\) stereographic projection, \texttt{gen\_trefoil3d(n,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \[(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),\] excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
+For the \(3\text{-}D\) stereographic projection, \texttt{gen\_trefoil3d(n,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),\) excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
 

### 2025-10-06 paper-cardinalR.Rmd
Added application for Trefoil knots
Added application for cone

@@
 
+Cone-shaped structures appear in particle dispersions, light beams, and tapering processes, where spread decreases along one axis.
+They are also used to benchmark clustering and dimensionality reduction methods [@hadsell2006].
+
 ```{r}
@@
 ```
+
+The trefoil knot appears in molecular biology (DNA/protein knotting), fluid dynamics (knotted vortices), and physics (topological phases), making it a useful benchmark for testing whether dimension reduction preserves global loops and topology [@witten1985; @arsuaga2002].
 

### 2025-10-06 paper-cardinalR.tex
Added application for Trefoil knots
Added application for cone

@@
 \end{verbatim}
+
+Cone-shaped structures appear in particle dispersions, light beams, and tapering processes, where spread decreases along one axis.
+They are also used to benchmark clustering and dimensionality reduction methods \citep{hadsell2006}.
 
@@
 
+The trefoil knot appears in molecular biology (DNA/protein knotting), fluid dynamics (knotted vortices), and physics (topological phases), making it a useful benchmark for testing whether dimension reduction preserves global loops and topology \citep{witten1985, arsuaga2002}.
+
 \begin{figure}[!]

### 2025-10-06 paper-cardinalR.Rmd
Updated figure positions and trigometric section

@@
 
-```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` (a1-a3) data.", fig.width=12, fig.height=4}
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` (a1-a3) data.", fig.width=12, fig.height=4, fig.pos="H"}
 
@@
 
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."}
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data.", fig.pos="H"}
 #| fig-width: 12
@@
 
-```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=12, fig.height=16}
+```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=12, fig.height=16, fig.pos="H"}
 
@@
 
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=28, out.width="80%", fig.align='center'}
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H"}
 
@@
 curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3  +
-unifsphere_proj1 + unifsphere_proj2 + unifsphere_proj3 +
-gridedsphere_proj1 + gridedsphere_proj2 + gridedsphere_proj3 +
 clusteredspheres_proj1 + clusteredspheres_proj2 + clusteredspheres_proj3 +
-hemisphere_proj1 + hemisphere_proj2 + hemisphere_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
@@
 
-```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=12, fig.height=8}
+```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=12, fig.height=8, fig.pos="H"}
 
@@
 Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric)"} else { "Figure \\@ref(fig:triginometric-proj)"})`).
-The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when $p > 2$ (adding $X_3$ through $X_p$), and for all other trigonometric shapes when $p > 4$ (adding $X_5$ through $X_p$).
-These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
+The main geometry is defined by the first few coordinates: crescents ($p=2$), cylinders, spirals, and helices ($p=4$).
+These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features [@calladine1997; @gershenfeld2000].
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trigonometric-tb-html)', '\\@ref(tab:trigonometric-tb-pdf)'))` summarizes these functions.
@@
 
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=16, out.width="80%", fig.align='center'}
+```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=16, out.width="80%", fig.align='center', fig.pos="H"}
 

### 2025-10-06 paper-cardinalR.tex
Updated figure positions and trigometric section

@@
 
-\begin{figure}[!]
+\begin{figure}[H]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `orglinearbranches` (a1-a3) data.}\label{fig:branch-proj}
@@
 
-\begin{figure}[!]
+\begin{figure}[H]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:cone-proj}
@@
 
-\begin{figure}[!]
+\begin{figure}[H]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.}\label{fig:pyr-proj}
@@
 
-\begin{figure}[!]
+\begin{figure}[H]
 
@@
 
-\begin{figure}[!]
+\begin{figure}[H]
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.}\label{fig:trefoil-proj}
@@
 Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (Figure \ref{fig:triginometric-proj}).
-The first few coordinates define the main geometric structure (e.g., crescent, cylinder, spiral, or helix), while for higher-dimensional embeddings, additional noise dimensions are appended: for the crescent shape this occurs when \(p > 2\) (adding \(X_3\) through \(X_p\)), and for all other trigonometric shapes when \(p > 4\) (adding \(X_5\) through \(X_p\)).
-These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features.
+The main geometry is defined by the first few coordinates: crescents (\(p=2\)), cylinders, spirals, and helices (\(p=4\)).
+These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features \citep{calladine1997, gershenfeld2000}.
 Table \ref{tab:trigonometric-tb-pdf} summarizes these functions.

### 2025-10-06 paper-cardinalR.R

@@ -291,7 +291,7 @@ orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 # branchfig
 
 
-## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` (a1-a3) data.", fig.width=12, fig.height=4----
+## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` (a1-a3) data.", fig.width=12, fig.height=4, fig.pos="H"----
 
 orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
@@ -402,7 +402,7 @@ cone_proj3 <- plot_proj(
 # conefig
 
 
-## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data."----
+## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data.", fig.pos="H"----
 #| fig-width: 12
 #| fig-height: 4
 
@@ -874,7 +874,7 @@ pyrholes_proj3 <- plot_proj(
 # pyrfig
 
 
-## ----pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=12, fig.height=16----
+## ----pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=12, fig.height=16, fig.pos="H"----
 
 pyrrect_proj1 + pyrrect_proj2 + pyrrect_proj3 +
 pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
@@ -1475,14 +1475,11 @@ hemisphere_proj3 <- plot_proj(
 # spherefig
[…]

### 2025-10-06 paper-cardinalR.Rmd
Added reference for spheres

@@
 
-Sphere-shaped structures (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere)"} else { "Figure \\@ref(fig:sphere-proj)"})`) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
-The functions generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+The functions generate a variety of spherical forms, including simple circles, uniform and hollow spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
 The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:sphere-tb-html)', '\\@ref(tab:sphere-tb-pdf)'))` summarizes these functions.
+Such spherical or hemispherical structures frequently appear in physical and biological systems, for example in models of celestial bodies, molecular shells, or cell membranes [@tinkham2003; @alberts2014].Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:sphere-tb-html)', '\\@ref(tab:sphere-tb-pdf)'))` summarizes these functions.
 
@@
                             "gen_unifsphere",
+                            "gen_hollowsphere",
                             "gen_gridedsphere",
@@
                                 "Uniform sphere.",
+                                "Hollow sphere.",
                                 "Grided sphere.",
@@
 
-The simplest case, `gen_circle(n)` creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
+The simplest case, `gen_circle(n, p)` creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
 Let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
@@
 
-For the one-dimensional nonlinear cycle embedded in $p\text{-}D$ space, `gen_curvycycle(n)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`), let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
+For the one-dimensional nonlinear cycle embedded in $p\text{-}D$ space, `gen_curvycycle(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`), let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
 The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle".
@@
 
-Building on simple circular structures, the `gen_unifsphere(n, r)` function function extends the idea to three dimensions by generating $n$ observations approximately uniformly distributed on the surface of a sphere of radius $r$, with optional high-dimensional noise when $p > 3$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) c"} else { "Figure \\@ref(fig:sphere-proj) c"})`).
+Building on simple circular structures, the `gen_unifsphere(n, r)` function function extends the idea to three dimensions by generating $n$ observations approximately uniformly distributed on the surface of a sphere of radius $r$..
 Each observation is computed from spherical coordinates, with $u \sim U(-1, 1)$ representing $\cos(\phi)$ and $\theta \sim U(0, 2\pi)$ the azimuthal angle.
@@
 
-```{r}
-unifsphere_lang <- langevitour::langevitour(unifsphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r unifsphere-proj1}
-scaled_data <- scale_data_manual(unifsphere)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097),
-  c(0.00896,0.52496,0.12597))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-unifsphere_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifsphere-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278),
-  c(-0.50493,0.14388,0.14615))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-unifsphere_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r unifsphere-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416),
-  c(0.50329,0.07199,-0.16064))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-unifsphere_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.65), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-In contrast, the `gen_hollowsphere(n, p)` function, a wrapper around `geozoo::sphere.hollow()`, generates $n$ points uniformly distributed **only on the surface** of the $(p-1)$-dimensional sphere embedded in $\mathbb{R}^p$.
+In contrast, the `gen_hollowsphere(n, p)` function, a wrapper around `geozoo::sphere.hollow()`, generates $n$ points uniformly distributed only on the surface of the $(p-1)$-dimensional sphere embedded in $\mathbb{R}^p$.
 This results in a hollow shell-like structure with no interior points.
@@
 
-```{r data-gridedsphere, echo=TRUE}
-gridedsphere <- gen_gridedsphere(n = 1000, p = 4)
-```
-
-```{r}
-gridedsphere_lang <- langevitour::langevitour(gridedsphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r gridedsphere-proj1}
-scaled_data <- scale_data_manual(gridedsphere)
+For more heterogeneous structures, the `gen_clusteredspheres(n, k, r, loc)` function generates one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
+A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
+The sampling is executed using the function `gen_unifsphere(n_1, r_1)`, which generates the desired points in the specified dimensional space.
+In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
+These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
+Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for $3\text{-}D$).
+Each observation is classified by cluster, with labels such as "big" for the large central sphere and "small_1" to "small_k" for the smaller spheres.
+
+```{r data-clusteredspheres, echo=TRUE}
+clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, r = c(15, 3),
+                                         loc = 10 / sqrt(3)) |>
+  dplyr::select(-cluster)
+```
+
+```{r}
+clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r clusteredspheres-proj1}
+scaled_data <- scale_data_manual(clusteredspheres)
 
@@
 projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
+  c(0.53506,0.00763,0.00097),
+  c(0.00896,0.52496,0.12597))
 
@@
 
-gridedsphere_proj1 <- plot_proj(
+clusteredspheres_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
@@
   plot_limits = c(-0.75, 0.7), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridedsphere-proj2}
+  title = "c1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r clusteredspheres-proj2}
 
@@
 projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
+  c(-0.06614,0.15134,-0.44278),
+  c(-0.50493,0.14388,0.14615))
 
@@
 
-gridedsphere_proj2 <- plot_proj(
+clusteredspheres_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r gridedsphere-proj3}
+  plot_limits = c(-0.7, 0.6), 
+  title = "c2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r clusteredspheres-proj3}
 
@@
 projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
+  c(0.10704,0.04822,-0.03416),
+  c(0.50329,0.07199,-0.16064))
 
@@
 
-gridedsphere_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.62), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For more heterogeneous structures, the `gen_clusteredspheres(n, k, r, loc)` function generates one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
-A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
-The sampling is executed using the function `gen_unifsphere(n_1, p, r_1)`, which generates the desired points in the specified dimensional space.
-In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
-These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
-Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for $3\text{-}D$).
-Each observation is classified by cluster, with labels such as "big" for the large central sphere and "small_1" to "small_k" for the smaller spheres.
-
-```{r data-clusteredspheres, echo=TRUE}
-clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, r = c(15, 3),
-                                         loc = 10 / sqrt(3)) |>
-  dplyr::select(-cluster)
-```
-
-```{r}
-clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r clusteredspheres-proj1}
-scaled_data <- scale_data_manual(clusteredspheres)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097),
-  c(0.00896,0.52496,0.12597))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-clusteredspheres_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "e1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r clusteredspheres-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278),
-  c(-0.50493,0.14388,0.14615))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-clusteredspheres_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "e2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r clusteredspheres-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416),
-  c(0.50329,0.07199,-0.16064))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
 clusteredspheres_proj3 <- plot_proj(
@@
   plot_limits = c(-0.6, 0.62), 
-  title = "e3", 
+  title = "c3", 
   cex = 2, 
@@
 
-```{r data-hemisphere, echo=TRUE}
-hemisphere <- gen_hemisphere(n = 1000)
-```
-
-```{r}
-hemisphere_lang <- langevitour::langevitour(hemisphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r hemisphere-proj1}
-scaled_data <- scale_data_manual(hemisphere)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-hemisphere_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "f1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r hemisphere-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-hemisphere_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "f2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r hemisphere-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-hemisphere_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.62), 
-  title = "f3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, `unifsphere`, `gridedsphere`, `clusteredspheres`, and `hemisphere` data in $4\\text{-}D$."}
+```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, and `clusteredspheres` data in $4\\text{-}D$."}
 
@@
     style = "display: grid; grid-template-columns: 1fr 1fr;",
-    circle_lang, curvycycle_lang, unifsphere_lang, gridedsphere_lang, clusteredspheres_lang, hemisphere_lang
+    circle_lang, curvycycle_lang, clusteredspheres_lang
   ),

### 2025-10-06 paper-cardinalR.tex
Added reference for spheres

@@
 
-Sphere-shaped structures (Figure \ref{fig:sphere-proj}) are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
-The functions generate a variety of spherical forms, including simple circles, uniform spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
+Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+The functions generate a variety of spherical forms, including simple circles, uniform and hollow spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
 The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
-Table \ref{tab:sphere-tb-pdf} summarizes these functions.
+Such spherical or hemispherical structures frequently appear in physical and biological systems, for example in models of celestial bodies, molecular shells, or cell membranes \citep{tinkham2003, alberts2014}.Table \ref{tab:sphere-tb-pdf} summarizes these functions.
 
@@
 gen\_unifsphere & Uniform sphere.\\
+gen\_hollowsphere & Hollow sphere.\\
 gen\_gridedsphere & Grided sphere.\\
@@
 
-The simplest case, \texttt{gen\_circle(n)} creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (Figure \ref{fig:sphere-proj} a).
+The simplest case, \texttt{gen\_circle(n,\ p)} creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (Figure \ref{fig:sphere-proj} a).
 Let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
@@
 
-For the one-dimensional nonlinear cycle embedded in \(p\text{-}D\) space, \texttt{gen\_curvycycle(n)} (Figure \ref{fig:sphere-proj} b), let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
+For the one-dimensional nonlinear cycle embedded in \(p\text{-}D\) space, \texttt{gen\_curvycycle(n,\ p)} (Figure \ref{fig:sphere-proj} b), let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
 The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle''.
@@
 
-Building on simple circular structures, the \texttt{gen\_unifsphere(n,\ r)} function function extends the idea to three dimensions by generating \(n\) observations approximately uniformly distributed on the surface of a sphere of radius \(r\), with optional high-dimensional noise when \(p > 3\) (Figure \ref{fig:sphere-proj} c).
+Building on simple circular structures, the \texttt{gen\_unifsphere(n,\ r)} function function extends the idea to three dimensions by generating \(n\) observations approximately uniformly distributed on the surface of a sphere of radius \(r\)..
 Each observation is computed from spherical coordinates, with \(u \sim U(-1, 1)\) representing \(\cos(\phi)\) and \(\theta \sim U(0, 2\pi)\) the azimuthal angle.
@@
 
-In contrast, the \texttt{gen\_hollowsphere(n,\ p)} function, a wrapper around \texttt{geozoo::sphere.hollow()}, generates \(n\) points uniformly distributed \textbf{only on the surface} of the \((p-1)\)-dimensional sphere embedded in \(\mathbb{R}^p\).
+In contrast, the \texttt{gen\_hollowsphere(n,\ p)} function, a wrapper around \texttt{geozoo::sphere.hollow()}, generates \(n\) points uniformly distributed only on the surface of the \((p-1)\)-dimensional sphere embedded in \(\mathbb{R}^p\).
 This results in a hollow shell-like structure with no interior points.
@@
 
-\begin{verbatim}
-gridedsphere <- gen_gridedsphere(n = 1000, p = 4)
-\end{verbatim}
-
 For more heterogeneous structures, the \texttt{gen\_clusteredspheres(n,\ k,\ r,\ loc)} function generates one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location (Figure \ref{fig:sphere-proj} e).
 A large uniform sphere centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\text{-}D\) sphere with a radius of \(r_1\).
-The sampling is executed using the function \texttt{gen\_unifsphere(n\_1,\ p,\ r\_1)}, which generates the desired points in the specified dimensional space.
+The sampling is executed using the function \texttt{gen\_unifsphere(n\_1,\ r\_1)}, which generates the desired points in the specified dimensional space.
 In generation of \(k\) smaller uniform spheres, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
@@
 The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
-
-\begin{verbatim}
-hemisphere <- gen_hemisphere(n = 1000)
-\end{verbatim}
 

### 2025-10-06 paper-cardinalR.Rmd
Updated the gen hole section

@@
 
-The package provides functionality for generating datasets with **spherical hole** (in 2D/3D) or, more generally, **hyperspherical hole** (in higher dimensions).
+The package provides functionality for generating datasets with spherical hole (in $2\text{-}D$/$3\text{-}D$) or, more generally, hyperspherical hole (in higher dimensions).
 These structures are valuable for evaluating how dimension reduction methods and clustering algorithms handle incomplete manifolds or missing regions of the data space.
@@
 
-The core function `gen_hole()` removes points from a dataset that fall within a user-specified hypersphere.
-Formally, given data points ($x \in \mathbb{R}^p$), a center ($a \in \mathbb{R}^p$), and radius ($r > 0$), only points satisfying
-
-$$
-||x - a||_2 > r
-$$
-
-are retained.
-The anchor point (a) can either be user-specified or default to the dataset mean, and radius (r) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
+The core function `gen_hole(df, anchor, r)` removes points from a dataset that fall within a user-specified hypersphere.
+Formally, given data points ($x \in \mathbb{R}^p$), a center ($a \in \mathbb{R}^p$), and radius ($r > 0$), only points satisfying $||x - a||_2 > r$ are retained.
+The anchor point ($a$) can either be user-specified or default to the dataset mean, and radius ($r$) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
 Because it operates generically on any dataset, spherical or hyperspherical holes can be embedded in a wide range of geometric structures.
@@
 Two specialized wrappers illustrate this idea.
-The function `gen_scurvehole()` generates an S-curve with a spherical hole by applying `gen_hole()` to the output of `gen_scurve()`.
+The function `gen_scurvehole(n, r_hole)` generates an S-curve with a spherical hole by applying `gen_hole()` to the output of `gen_scurve()`.
 This structure has been used in prior diagnostic studies of NLDR methods, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
-The second wrapper, `gen_unifcubehole()`, generates uniformly sampled cube data with a hyperspherical hole.
+The second wrapper, `gen_unifcubehole(n, p, r_hole)`, generates uniformly sampled cube data with a hyperspherical hole.
 By embedding a hyperspherical void inside a convex high-dimensional structure, this creates non-convex regions that challenge algorithms in terms of separability and neighborhood preservation.

### 2025-10-06 paper-cardinalR.tex
Updated the gen hole section

@@
 
-The package provides functionality for generating datasets with \textbf{spherical hole} (in 2D/3D) or, more generally, \textbf{hyperspherical hole} (in higher dimensions).
+The package provides functionality for generating datasets with spherical hole (in \(2\text{-}D\)/\(3\text{-}D\)) or, more generally, hyperspherical hole (in higher dimensions).
 These structures are valuable for evaluating how dimension reduction methods and clustering algorithms handle incomplete manifolds or missing regions of the data space.
@@
 
-The core function \texttt{gen\_hole()} removes points from a dataset that fall within a user-specified hypersphere.
-Formally, given data points (\(x \in \mathbb{R}^p\)), a center (\(a \in \mathbb{R}^p\)), and radius (\(r > 0\)), only points satisfying
-
-\[
-||x - a||_2 > r
-\]
-
-are retained.
-The anchor point (a) can either be user-specified or default to the dataset mean, and radius (r) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
+The core function \texttt{gen\_hole(df,\ anchor,\ r)} removes points from a dataset that fall within a user-specified hypersphere.
+Formally, given data points (\(x \in \mathbb{R}^p\)), a center (\(a \in \mathbb{R}^p\)), and radius (\(r > 0\)), only points satisfying \(||x - a||_2 > r\) are retained.
+The anchor point (\(a\)) can either be user-specified or default to the dataset mean, and radius (\(r\)) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
 Because it operates generically on any dataset, spherical or hyperspherical holes can be embedded in a wide range of geometric structures.
@@
 Two specialized wrappers illustrate this idea.
-The function \texttt{gen\_scurvehole()} generates an S-curve with a spherical hole by applying \texttt{gen\_hole()} to the output of \texttt{gen\_scurve()}.
+The function \texttt{gen\_scurvehole(n,\ r\_hole)} generates an S-curve with a spherical hole by applying \texttt{gen\_hole()} to the output of \texttt{gen\_scurve()}.
 This structure has been used in prior diagnostic studies of NLDR methods, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
-The second wrapper, \texttt{gen\_unifcubehole()}, generates uniformly sampled cube data with a hyperspherical hole.
+The second wrapper, \texttt{gen\_unifcubehole(n,\ p,\ r\_hole)}, generates uniformly sampled cube data with a hyperspherical hole.
 By embedding a hyperspherical void inside a convex high-dimensional structure, this creates non-convex regions that challenge algorithms in terms of separability and neighborhood preservation.

### 2025-10-07 paper-cardinalR.tex
Renumbered the sections

@@
 
-\section{Implementation}\label{implementation}
+\section{Usage}\label{usage}
 
@@
 
-\subsection{Usage}\label{usage}
-
-\subsubsection{Main function}\label{main-function}
+\subsection{Main function}\label{main-function}
 
@@
 
-\subsubsection{Shape generators}\label{shape-generators}
+\subsection{Shape generators}\label{shape-generators}
 
@@
 
-\paragraph{Branching}\label{branching}
+\subsubsection{Branching}\label{branching}
 
@@
 
-\paragraph{Cone}\label{cone}
+\subsubsection{Cone}\label{cone}
 
@@
 
-\paragraph{Cube}\label{cube}
+\subsubsection{Cube}\label{cube}
 
@@
 
-\paragraph{Gaussian}\label{gaussian}
+\subsubsection{Gaussian}\label{gaussian}
 
@@
 
-\paragraph{Linear}\label{linear}
+\subsubsection{Linear}\label{linear}
 
@@
 
-\paragraph{Möbius}\label{muxf6bius}
+\subsubsection{Möbius}\label{muxf6bius}
 
@@
 
-\paragraph{Polynomial}\label{polynomial}
+\subsubsection{Polynomial}\label{polynomial}
 
@@
 
-\paragraph{Pyramid}\label{pyramid}
+\subsubsection{Pyramid}\label{pyramid}
 
@@
 
-\paragraph{S-curve}\label{s-curve}
+\subsubsection{S-curve}\label{s-curve}
 
@@
 
-\paragraph{Sphere}\label{sphere}
+\subsubsection{Sphere}\label{sphere}
 
@@
 
-\paragraph{Swiss Roll}\label{swiss-roll}
+\subsubsection{Swiss Roll}\label{swiss-roll}
 
@@
 
-\paragraph{Trefoil knots}\label{trefoil-knots}
+\subsubsection{Trefoil knots}\label{trefoil-knots}
 
@@
 
-\paragraph{Trigonometric}\label{trigonometric}
+\subsubsection{Trigonometric}\label{trigonometric}
 
@@
 
-\subsubsection{Generate a spherical or hyperspherical hole within a structure}\label{generate-a-spherical-or-hyperspherical-hole-within-a-structure}
+\subsection{Generate a spherical or hyperspherical hole within a structure}\label{generate-a-spherical-or-hyperspherical-hole-within-a-structure}
 
@@
 
-\subsubsection{Generate noise dimensions}\label{generate-noise-dimensions}
+\subsection{Generate noise dimensions}\label{generate-noise-dimensions}
 
@@
 
-\subsubsection{Multiple cluster examples}\label{multiple-cluster-examples}
+\subsection{Multiple cluster examples}\label{multiple-cluster-examples}
 
@@
 
-\subsubsection{Additional functions}\label{additional-functions}
+\subsection{Additional functions}\label{additional-functions}
 
@@
 \end{table}
+
+\section{Implementation}\label{implementation}
 

### 2025-10-08 paper-cardinalR.Rmd
Renumbered the sections
Updated branching projection

@@
 
-# Implementation
+# Usage
 
@@
 
-## Usage
-
-### Main function
+## Main function
 
@@
 
-### Shape generators
+## Shape generators
 
@@
 
-#### Branching
+### Branching
 
@@
 projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
+    c(-0.083573,-0.019043,0.006574,-0.035774),
+    c(0.005380,0.068541,-0.030810,-0.054717))
 
 proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                            proj_scale = 1.8, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 0.085,
+                                              axis_scaled = 6, 
+                                              axis_pos_x = -0.055, 
+                                              axis_pos_y = -0.055, 
+                                              threshold = 0.006))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.07, 0.07), 
   title = "a1", 
@@
 
-#### Cone
+### Cone
 
@@
 
-#### Cube
+### Cube
 
@@
 
-#### Gaussian
+### Gaussian
 
@@
 
-#### Linear
+### Linear
 
@@
 
-#### Möbius
+### Möbius
 
@@
 
-#### Polynomial
+### Polynomial
 
@@
 
-#### Pyramid
+### Pyramid
 
@@
 
-#### S-curve
+### S-curve
 
@@
 
-#### Sphere
+### Sphere
 
@@
 
-#### Swiss Roll
+### Swiss Roll
 
@@
 
-#### Trefoil knots
+### Trefoil knots
 
@@
 
-#### Trigonometric
+### Trigonometric
 
@@
 
-### Generate a spherical or hyperspherical hole within a structure
+## Generate a spherical or hyperspherical hole within a structure
 
@@
 
-### Generate noise dimensions
+## Generate noise dimensions
 
@@
 
-### Multiple cluster examples
+## Multiple cluster examples
 
@@
 
-### Additional functions
+## Additional functions
 
@@
 
+# Implementation
+
 # Application

### 2025-10-08 paper-cardinalR.Rmd
Added multicluster code example

@@
 
+The following example demonstrates how to use `gen_multicluster()` to create a $4\text{-}D$ dataset with three clusters of different shapes and orientations:
+
+```{r, echo=TRUE}
+# Define example rotation matrices for 4D space
+rot1 <- gen_rotation(p = 4, planes_angles = list(
+  list(plane = c(1, 2), angle = 60),
+  list(plane = c(3, 4), angle = 90)
+))
+
+rot2 <- gen_rotation(p = 4, planes_angles = list(
+  list(plane = c(1, 3), angle = 30)
+))
+
+rot3 <- gen_rotation(p = 4, planes_angles = list(
+  list(plane = c(2, 4), angle = 45)
+))
+
+# Generate the clustered dataset
+clust_data <- gen_multicluster(
+  n = c(200, 300, 500),
+  k = 3,
+  loc = matrix(c(
+    0, 0, 0, 0,
+    5, 9, 0, 0,
+    3, 4, 10, 7
+  ), nrow = 3, byrow = TRUE),
+  scale = c(3, 1, 2),
+  shape = c("gaussian", "cone", "unifcube"),
+  rotation = list(rot1, rot2, rot3),
+  is_bkg = FALSE
+)
+
+```
+
 ## Shape generators

### 2025-10-08 paper-cardinalR.tex
Added multicluster code example

@@
 \end{table}
+
+The following example demonstrates how to use \texttt{gen\_multicluster()} to create a \(4\text{-}D\) dataset with three clusters of different shapes and orientations:
+
+\begin{verbatim}
+# Define example rotation matrices for 4D space
+rot1 <- gen_rotation(p = 4, planes_angles = list(
+  list(plane = c(1, 2), angle = 60),
+  list(plane = c(3, 4), angle = 90)
+))
+
+rot2 <- gen_rotation(p = 4, planes_angles = list(
+  list(plane = c(1, 3), angle = 30)
+))
+
+rot3 <- gen_rotation(p = 4, planes_angles = list(
+  list(plane = c(2, 4), angle = 45)
+))
+
+# Generate the clustered dataset
+clust_data <- gen_multicluster(
+  n = c(200, 300, 500),
+  k = 3,
+  loc = matrix(c(
+    0, 0, 0, 0,
+    5, 9, 0, 0,
+    3, 4, 10, 7
+  ), nrow = 3, byrow = TRUE),
+  scale = c(3, 1, 2),
+  shape = c("gaussian", "cone", "unifcube"),
+  rotation = list(rot1, rot2, rot3),
+  is_bkg = FALSE
+)
+\end{verbatim}
 

### 2025-10-08 paper-cardinalR.Rmd
Updated application by adding two new sections

@@
                                               axis_scaled = 6, 
-                                              axis_pos_x = -0.055, 
-                                              axis_pos_y = -0.055, 
+                                              axis_pos_x = -0.07, 
+                                              axis_pos_y = -0.07, 
                                               threshold = 0.006))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.07, 0.07), 
+  plot_limits = c(-0.1, 0.1), 
   title = "a1", 
@@
 
-<!-- ## Assessing the performance of dimension reduction on different geometric structures in high-dimensions -->
-
-This section illustrates the use of package by generating a synthetic dataset to evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
-
-The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
-Each cluster was chosen to reflect distinct geometric and topological properties, allowing us to test how well DR methods preserve both local and global structures.
-The *helical spiral* cluster is designed to evaluate methods on elongated, twisting structures that challenge linear embeddings such as PCA and require preservation of curvilinear continuity.
-The *hemisphere* provides a curved surface with partial coverage of a $3\text{-}D$ manifold, useful for testing neighborhood preservation and unfolding in algorithms like UMAP and tSNE.
-The *uniform cube* represents isotropic, uniformly distributed data and serves as a control cluster with simple geometric structure to assess baseline embedding fidelity.
-The *cone* introduces variable density along one axis, mimicking structures where point density changes with geometry, helping evaluate how well algorithms maintain relative distances in non-uniform distributions.
-Finally, the *Gaussian* cluster is a standard multivariate normal distribution, included to assess algorithm performance on simple, spherical, high-density clusters.
-Together, these clusters create a challenging synthetic dataset suitable for benchmarking and exploring the strengths and weaknesses of different dimensionality reduction techniques.
+This section demonstrates how the package can be used to generate complex high-dimensional datasets, apply dimension reduction (DR) techniques, and evaluate clustering performance.
+The example shows how diverse geometric structures can be simulated and analyzed to assess algorithmic behavior.
+
+## Generating high-dimensional clustered data
+
+To illustrate, we generate a dataset with five clusters in $4\text{-}D$, each representing distinct geometric characteristics: a *helical spiral* (elongated and twisted), a *hemisphere* (curved surface), a *uniform cube* (isotropic distribution), a *cone* (density gradient), and a *Gaussian* cluster (compact and spherical) (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
+Each cluster has a unique number of points and scaling factor, representing variation in cluster size and spread across the $4\text{-}D$ space.
 
@@
 
-The five clusters have different geometric structures and each contain different number of points.
-Specifically, the helical spiral cluster includes $2250$ points and was generated with a scale parameter of $0.4$.
-The hemisphere cluster consists of $1500$ points with a scale parameter of $0.35$.
-The uniform cube-shaped cluster contains $750$ points and uses a scale parameter of $0.3$.
-The blunted cone cluster includes $1250$ points, generated with a scale parameter of $1$.
-Finally, the Gaussian-shaped cluster contains $1750$ points and was generated with a scale parameter of $0.3$.
-
 ```{r highd-data, eval=knitr::is_html_output(), fig.cap="`langevitour` output showing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster."}
@@
 ```
+
+## Evaluating dimension reduction (DR) methods
+
+Next, we apply six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 
@@
 
-UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure (Figure \@ref(fig:fig-nldr-layouts)).
+```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
+
+nldr1 + nldr2 + nldr3 +
+  nldr4 + nldr5 + nldr6 +
+  plot_layout(ncol = 3, guides = "collect")
+```
+
+These methods vary in their ability to preserve local and global data structures (Figure \@ref(fig:fig-nldr-layouts)).
+UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure.
 However, PHATE reveals three non-linear clusters, even though two of them do not show non-linearity.
@@
 
-```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
-
-nldr1 + nldr2 + nldr3 +
-  nldr4 + nldr5 + nldr6 +
-  plot_layout(ncol = 3, guides = "collect")
-```
+## Benchmarking clustering algorithms
 

### 2025-10-08 paper-cardinalR.tex
Updated application by adding two new sections

@@
 
-This section illustrates the use of package by generating a synthetic dataset to evaluate the performance of six popular dimension reduction techniques: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
+This section demonstrates how the package can be used to generate complex high-dimensional datasets, apply dimension reduction (DR) techniques, and evaluate clustering performance.
+The example shows how diverse geometric structures can be simulated and analyzed to assess algorithmic behavior.
 
-The following code generates a dataset of five clusters, positioned with equal inter-cluster distances in \(4\text{-}D\) space (Figure \ref{fig:highd-proj}).
-Each cluster was chosen to reflect distinct geometric and topological properties, allowing us to test how well DR methods preserve both local and global structures.
-The \emph{helical spiral} cluster is designed to evaluate methods on elongated, twisting structures that challenge linear embeddings such as PCA and require preservation of curvilinear continuity.
-The \emph{hemisphere} provides a curved surface with partial coverage of a \(3\text{-}D\) manifold, useful for testing neighborhood preservation and unfolding in algorithms like UMAP and tSNE.
-The \emph{uniform cube} represents isotropic, uniformly distributed data and serves as a control cluster with simple geometric structure to assess baseline embedding fidelity.
-The \emph{cone} introduces variable density along one axis, mimicking structures where point density changes with geometry, helping evaluate how well algorithms maintain relative distances in non-uniform distributions.
-Finally, the \emph{Gaussian} cluster is a standard multivariate normal distribution, included to assess algorithm performance on simple, spherical, high-density clusters.
-Together, these clusters create a challenging synthetic dataset suitable for benchmarking and exploring the strengths and weaknesses of different dimensionality reduction techniques.
+\subsection{Generating high-dimensional clustered data}\label{generating-high-dimensional-clustered-data}
+
+To illustrate, we generate a dataset with five clusters in \(4\text{-}D\), each representing distinct geometric characteristics: a \emph{helical spiral} (elongated and twisted), a \emph{hemisphere} (curved surface), a \emph{uniform cube} (isotropic distribution), a \emph{cone} (density gradient), and a \emph{Gaussian} cluster (compact and spherical) (Figure \ref{fig:highd-proj}).
+Each cluster has a unique number of points and scaling factor, representing variation in cluster size and spread across the \(4\text{-}D\) space.
 
@@
 
-The five clusters have different geometric structures and each contain different number of points.
-Specifically, the helical spiral cluster includes \(2250\) points and was generated with a scale parameter of \(0.4\).
-The hemisphere cluster consists of \(1500\) points with a scale parameter of \(0.35\).
-The uniform cube-shaped cluster contains \(750\) points and uses a scale parameter of \(0.3\).
-The blunted cone cluster includes \(1250\) points, generated with a scale parameter of \(1\).
-Finally, the Gaussian-shaped cluster contains \(1750\) points and was generated with a scale parameter of \(0.3\).
-
 \begin{figure}[H]
@@
 
-UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure (Figure \ref{fig:fig-nldr-layouts}).
+\subsection{Evaluating dimension reduction (DR) methods}\label{evaluating-dimension-reduction-dr-methods}
+
+Next, we apply six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
+\end{figure}
+
+These methods vary in their ability to preserve local and global data structures (Figure \ref{fig:fig-nldr-layouts}).
+UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure.
 However, PHATE reveals three non-linear clusters, even though two of them do not show non-linearity.
@@
 
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
-\end{figure}
+\subsection{Benchmarking clustering algorithms}\label{benchmarking-clustering-algorithms}
 

### 2025-10-08 paper-cardinalR.R

@@ -235,14 +235,14 @@ proj_obj1 <- get_projection(projection = projection,
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.085,
                                               axis_scaled = 6, 
-                                              axis_pos_x = -0.055, 
-                                              axis_pos_y = -0.055, 
+                                              axis_pos_x = -0.07, 
+                                              axis_pos_y = -0.07, 
                                               threshold = 0.006))
 
 orglinearbranches_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.07, 0.07), 
+  plot_limits = c(-0.1, 0.1), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 5,

### 2025-10-08 paper-cardinalR.Rmd
Updated the branching projection

@@
                             scaled_data = scaled_data, 
-                            axis_param = list(limits = 0.085,
-                                              axis_scaled = 6, 
-                                              axis_pos_x = -0.07, 
-                                              axis_pos_y = -0.07, 
-                                              threshold = 0.006))
+                            axis_param = list(limits = 0.15,
+                                              axis_scaled = 10, 
+                                              axis_pos_x = -0.12, 
+                                              axis_pos_y = -0.12, 
+                                              threshold = 0.015))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.1, 0.1), 
+  plot_limits = c(-0.15, 0.15), 
   title = "a1", 

### 2025-10-08 paper-cardinalR.Rmd
Initial codes for clustering

@@
 
+```{r, echo=TRUE}
+library(fpc)
+
+# Extract high-dimensional data
+data <- five_clusts[, -5]
+true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))
+
+# Hierarchical clustering
+dist_mat <- dist(data)
+hc_res <- cutree(hclust(dist_mat, method = "ward.D2"), k = 5)
+```
+
+```{r, echo=TRUE}
+# K-means clustering
+kmeans_res <- kmeans(data, centers = 5, nstart = 20)
+```
+
+```{r, echo=TRUE}
+library(mclust)
+
+# Model-based clustering
+mclust_res <- Mclust(data, G = 5)$classification
+```
+
+```{r}
+# Evaluate clustering performance
+comp_kmeans <- cluster.stats(dist_mat, true_labels, kmeans_res$cluster)
+comp_hclust <- cluster.stats(dist_mat, true_labels, hc_res)
+comp_mclust <- cluster.stats(dist_mat, true_labels, mclust_res)
+
+# Summarize key performance metrics
+data.frame(
+  Method = c("K-means", "Hierarchical", "Model-based"),
+  ARI = c(comp_kmeans$corrected.rand,
+          comp_hclust$corrected.rand,
+          comp_mclust$corrected.rand),
+  Entropy = c(comp_kmeans$entropy,
+              comp_hclust$entropy,
+              comp_mclust$entropy)
+)
+```
+
 # Conclusion
@@
 The source material for this paper is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
-
 This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.

### 2025-10-08 paper-cardinalR.tex
Initial codes for clustering

@@
 
+\begin{verbatim}
+library(fpc)
+
+# Extract high-dimensional data
+data <- five_clusts[, -5]
+true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))
+
+# Hierarchical clustering
+dist_mat <- dist(data)
+hc_res <- cutree(hclust(dist_mat, method = "ward.D2"), k = 5)
+\end{verbatim}
+
+\begin{verbatim}
+# K-means clustering
+kmeans_res <- kmeans(data, centers = 5, nstart = 20)
+\end{verbatim}
+
+\begin{verbatim}
+library(mclust)
+
+# Model-based clustering
+mclust_res <- Mclust(data, G = 5)$classification
+\end{verbatim}
+
+\begin{verbatim}
+#>         Method ARI  Entropy
+#> 1      K-means   1 1.551532
+#> 2 Hierarchical   1 1.551532
+#> 3  Model-based   1 1.551532
+\end{verbatim}
+
 \section{Conclusion}\label{conclusion}

### 2025-10-08 paper-cardinalR.Rmd
Try to make a complex dataset

--- 2025-10-08 paper-cardinalR.Rmd
@@
 
+# Implementation
+
 # Usage
@@
 
-# Implementation
-
 # Application
@@
 positions <- geozoo::simplex(p=4)$points
-positions <- positions * 0.8
+positions <- positions * 0.3 #positions * 0.8
 
@@
                        loc = positions,
-                       scale = c(0.4, 0.35, 0.3, 1, 0.3),
+                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
                        shape = c("helicalspiral", "hemisphere", "unifcube", 
@@
 
+```{r}
+library(fpc)
+
+# Extract high-dimensional data
+data <- five_clusts[, -5]
+true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))
+
+# Hierarchical clustering
+dist_mat <- dist(data)
+hc_res <- cutree(hclust(dist_mat, method = "ward.D2"), k = 5)
+```
+
+```{r}
+# K-means clustering
+kmeans_res <- kmeans(data, centers = 5, nstart = 20)
+```
+
+```{r}
+library(mclust)
+
+# Model-based clustering
+mclust_res <- Mclust(data, G = 5, modelNames = "VVV")$classification
+```
+
+```{r}
+# Evaluate clustering performance
+comp_kmeans <- cluster.stats(dist_mat, true_labels, kmeans_res$cluster)
+comp_hclust <- cluster.stats(dist_mat, true_labels, hc_res)
+comp_mclust <- cluster.stats(dist_mat, true_labels, mclust_res)
+
+# Summarize key performance metrics
+data.frame(
+  Method = c("K-means", "Hierarchical", "Model-based"),
+  Avg_Silhouette = c(comp_kmeans$avg.silwidth, comp_hclust$avg.silwidth, comp_mclust$avg.silwidth),
+  Dunn_Index = c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn),
+  Entropy = c(comp_kmeans$entropy, comp_hclust$entropy, comp_mclust$entropy),
+  Corrected_Rand = c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand)
+)
+
+```
+
 # Conclusion

### 2025-10-08 paper-cardinalR.tex
Try to make a complex dataset

--- 2025-10-08 paper-cardinalR.tex
@@
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for the use of our data collection.
+
+\section{Implementation}\label{implementation}
 
@@
 
-\section{Implementation}\label{implementation}
-
 \section{Application}\label{application}
@@
 
+\begin{verbatim}
+#>         Method Avg_Silhouette  Dunn_Index  Entropy Corrected_Rand
+#> 1      K-means       0.227417 0.008968602 1.551532      0.4223504
+#> 2 Hierarchical       0.227417 0.008968602 1.551532      0.5029130
+#> 3  Model-based       0.227417 0.008968602 1.551532      0.7525318
+\end{verbatim}
+
 \section{Conclusion}\label{conclusion}
@@
 The source material for this paper is available at \href{https://github.com/JayaniLakshika/paper-cardinalR}{github.com/JayaniLakshika/paper-cardinalR}.
-
 This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.

### 2025-10-12 paper-cardinalR.Rmd
Added draft image for the implementation

@@
 
-<!--https://cran.r-project.org/web/packages/surreal/index.html-->
-<!--https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html-->
-
 While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.
@@
 
+```{r, out.width = "100%"}
+knitr::include_graphics("figures/cardinalR.pdf")
+```
+
 # Usage
@@
 The `cardinalR` R package is available on GitHub at [JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR).
-
-<!-- ## Web site -->
-
-<!-- More documentation of the package can be found at the web site [https://jayanilakshika.github.io/cardinalR/](https://jayanilakshika.github.io/cardinalR/). -->
 

### 2025-10-13 paper-cardinalR.Rmd
Add intro paragraph for implementation

@@
 
-```{r, out.width = "100%"}
+The `cardinalR` package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes.
+The main function, `gen_multicluster()`, combines these shapes into a single dataset by applying scaling, rotation, and translation through `gen_rotation()`.
+Each generated shape is assigned a unique cluster label.
+This design allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
+Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(fig:workflow)', '\\@ref(fig:workflow)'))` illustrates the workflow of `gen_multicluster()`.
+
+```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow of."}
 knitr::include_graphics("figures/cardinalR.pdf")

### 2025-10-13 paper-cardinalR.tex
Added draft image for the implementation
Add intro paragraph for implementation

@@
 \section{Implementation}\label{implementation}
+
+The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes.
+The main function, \texttt{gen\_multicluster()}, combines these shapes into a single dataset by applying scaling, rotation, and translation through \texttt{gen\_rotation()}.
+Each generated shape is assigned a unique cluster label.
+This design allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
+Figure \ref{fig:workflow} illustrates the workflow of \texttt{gen\_multicluster()}.
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{figures/cardinalR} \caption{Workflow of.}\label{fig:workflow}
+\end{figure}
 

### 2025-10-13 paper-cardinalR.Rmd
Added benchmarking clustering algorithm

@@
 library(crosstalk)
+library(fpc)
+library(mclust)
 ```
@@
 
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=16, out.width="80%", fig.align='center', fig.pos="H"}
+```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}
 
@@
 positions <- geozoo::simplex(p=4)$points
-positions <- positions * 0.3 #positions * 0.8
+positions <- positions * 0.3
 
@@
   theme(aspect.ratio = 1,
-        legend.text = element_text(size = 10))
+        legend.position = "none")
 
@@
   theme(aspect.ratio = 1,
-        legend.text = element_text(size = 10)) 
+        legend.position = "none") 
 ```
@@
   nldr4 + nldr5 + nldr6 +
-  plot_layout(ncol = 3, guides = "collect")
+  plot_layout(ncol = 3)
 ```
@@
 
-```{r}
-library(fpc)
-
+To further evaluate the structure of the generated data, we benchmarked three clustering algorithms: **$k$-means** [Chapter 20 of @boehmke2019], **hierarchical** [@murtagh2012], and **model-based clustering** [@chris2002; @scrucca2023] using the simulated dataset.
+The model-based clustering was performed with the `"VVV"` covariance structure, allowing each cluster to vary in volume, shape, and orientation.
+Cluster validity statistics were computed using the `cluster.stats()` function from the `fpc` package [@christian2024].
+
+```{r}
 # Extract high-dimensional data
@@
 ```{r}
-library(mclust)
-
 # Model-based clustering
@@
 # Summarize key performance metrics
-data.frame(
-  Method = c("K-means", "Hierarchical", "Model-based"),
-  Avg_Silhouette = c(comp_kmeans$avg.silwidth, comp_hclust$avg.silwidth, comp_mclust$avg.silwidth),
-  Dunn_Index = c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn),
-  Entropy = c(comp_kmeans$entropy, comp_hclust$entropy, comp_mclust$entropy),
-  Corrected_Rand = c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand)
+summary_clust <- tibble::tibble(
+  Metric = c("k-means", "Hierarchical", "Model-based"),
+  `wb.ratio` = round(c(comp_kmeans$wb.ratio, comp_hclust$wb.ratio, comp_mclust$wb.ratio), 2),
+  `Dunn Index` = round(c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn), 2),
+  `Corrected rand` = round(c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand), 2),
+  VI = round(c(comp_kmeans$vi, comp_hclust$vi, comp_mclust$vi), 2)
 )
-
-```
+```
+
+```{r summaryclust-tb-html, eval=knitr::is_html_output()}
+summary_clust |> 
+  kable(caption = "cardinalR branching data generation functions", col.names = c("Metric", "wb.ratio", "Dunn_Index", "Corrected_Rand", "VI"), table.pos = "H") 
+```
+
+```{r summaryclust-tb-pdf, eval=knitr::is_latex_output()}
+summary_clust |> 
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Metric", "wb.ratio", "Dunn_Index", "Corrected_Rand", "VI"), booktabs = T, table.pos = "H")  |>
+  column_spec(1, width = "2.5cm") |>
+  column_spec(2, width = "2cm") |>
+  column_spec(3, width = "2cm") |>
+  column_spec(4, width = "2cm") |>
+  column_spec(5, width = "2cm")
+```
+
+Overall, all methods produced similar compactness and separation, as reflected by the *within–between cluster ratios (wb.ratio)* and *Dunn indices*.
+However, the **model-based clustering** achieved the highest *Corrected Rand Index* ($0.75$) and lowest *Variation of Information (VI)* ($0.65$), indicating the best recovery of the true underlying groups.
+In comparison, $k$-means and hierarchical clustering showed moderate agreement with the true labels.
+These findings demonstrate that mixture-based approaches can more effectively capture the heterogeneity of clusters in high-dimensional, non-spherical data.
 

### 2025-10-13 paper-cardinalR.tex
Added benchmarking clustering algorithm

@@
 
-\begin{verbatim}
-#>         Method Avg_Silhouette  Dunn_Index  Entropy Corrected_Rand
-#> 1      K-means       0.227417 0.008968602 1.551532      0.4223504
-#> 2 Hierarchical       0.227417 0.008968602 1.551532      0.5029130
-#> 3  Model-based       0.227417 0.008968602 1.551532      0.7525318
-\end{verbatim}
+To further evaluate the structure of the generated data, we benchmarked three clustering algorithms: \textbf{\(k\)-means} \citep[Chapter 20 of][]{boehmke2019}, \textbf{hierarchical} \citep{murtagh2012}, and \textbf{model-based clustering} \citep{chris2002, scrucca2023} using the simulated dataset.
+The model-based clustering was performed with the \texttt{"VVV"} covariance structure, allowing each cluster to vary in volume, shape, and orientation.
+Cluster validity statistics were computed using the \texttt{cluster.stats()} function from the \texttt{fpc} package \citep{christian2024}.
+
+\begin{table}
+
+\caption{\label{tab:summaryclust-tb-pdf}cardinalR branching data generation functions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}}
+\toprule
+Metric & wb.ratio & Dunn\_Index & Corrected\_Rand & VI\\
+\midrule
+k-means & 0.61 & 0.01 & 0.42 & 1.32\\
+Hierarchical & 0.61 & 0.01 & 0.50 & 1.15\\
+Model-based & 0.61 & 0.01 & 0.75 & 0.65\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Overall, all methods produced similar compactness and separation, as reflected by the \emph{within--between cluster ratios (wb.ratio)} and \emph{Dunn indices}.
+However, the \textbf{model-based clustering} achieved the highest \emph{Corrected Rand Index} (\(0.75\)) and lowest \emph{Variation of Information (VI)} (\(0.65\)), indicating the best recovery of the true underlying groups.
+In comparison, \(k\)-means and hierarchical clustering showed moderate agreement with the true labels.
+These findings demonstrate that mixture-based approaches can more effectively capture the heterogeneity of clusters in high-dimensional, non-spherical data.
 

### 2025-10-13 scripts/five_clusts/01_data_generation.R
Regenerate embeddings with new data

@@ -3,15 +3,16 @@ library(readr)
 set.seed(20240412)
 
 positions <- geozoo::simplex(p=4)$points
-positions <- positions * 0.8
+positions <- positions * 0.3
 
 ## To generate data
-five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), p = 4, k = 5,
-                       loc = positions,
-                       scale = c(0.4, 0.35, 0.3, 1, 0.3),
-                       shape = c("helicalspiral", "hemisphere", "unifcube", "bluntedcone", "gaussian"),
-                       rotation = NULL,
-                       is_bkg = FALSE)
+five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
+                                loc = positions,
+                                scale = c(0.25, 0.35, 0.3, 1, 0.3),
+                                shape = c("helicalspiral", "hemisphere", "unifcube",
+                                          "cone", "gaussian"),
+                                rotation = NULL,
+                                is_bkg = FALSE)
 
 langevitour::langevitour(five_clusts[, -5])
 

### 2025-10-13 paper-cardinalR.Rmd
Need draw the implementation diagram

@@
 
-```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow of."}
+```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow of.", eval=FALSE}
 knitr::include_graphics("figures/cardinalR.pdf")
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.8), 
+  plot_limits = c(-0.8, 0.8), 
   title = "a1", 
@@
 
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H"}
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}
 
@@
 summary_clust |> 
-  kable(caption = "cardinalR branching data generation functions", col.names = c("Metric", "wb.ratio", "Dunn_Index", "Corrected_Rand", "VI"), table.pos = "H") 
+  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", col.names = c("Metric", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), tab.pos = "H")
 ```
@@
 summary_clust |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Metric", "wb.ratio", "Dunn_Index", "Corrected_Rand", "VI"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("Metric", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), booktabs = T, table.pos = "H")  |>
   column_spec(1, width = "2.5cm") |>
@@
   column_spec(3, width = "2cm") |>
-  column_spec(4, width = "2cm") |>
+  column_spec(4, width = "2.5cm") |>
   column_spec(5, width = "2cm")
@@
 Overall, all methods produced similar compactness and separation, as reflected by the *within–between cluster ratios (wb.ratio)* and *Dunn indices*.
-However, the **model-based clustering** achieved the highest *Corrected Rand Index* ($0.75$) and lowest *Variation of Information (VI)* ($0.65$), indicating the best recovery of the true underlying groups.
+However, the **model-based clustering** achieved the highest *Corrected Rand Index* ($0.75$) and lowest *Variation of Information (VI)* ($0.65$), indicating the best recovery of the true underlying groups (Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:summaryclust-tb-html)', '\\@ref(tab:summaryclust-tb-pdf)'))`).
 In comparison, $k$-means and hierarchical clustering showed moderate agreement with the true labels.

### 2025-10-13 paper-cardinalR.tex
Regenerate embeddings with new data
Need draw the implementation diagram

@@
 Figure \ref{fig:workflow} illustrates the workflow of \texttt{gen\_multicluster()}.
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/cardinalR} \caption{Workflow of.}\label{fig:workflow}
-\end{figure}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}
+{\centering \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
+{\centering \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
 
@@
 
-\caption{\label{tab:summaryclust-tb-pdf}cardinalR branching data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}}
-\toprule
-Metric & wb.ratio & Dunn\_Index & Corrected\_Rand & VI\\
+\caption{\label{tab:summaryclust-tb-pdf}Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}}
+\toprule
+Metric & wb.ratio & Dunn Index & Corrected Rand & VI\\
 \midrule
@@
 Overall, all methods produced similar compactness and separation, as reflected by the \emph{within--between cluster ratios (wb.ratio)} and \emph{Dunn indices}.
-However, the \textbf{model-based clustering} achieved the highest \emph{Corrected Rand Index} (\(0.75\)) and lowest \emph{Variation of Information (VI)} (\(0.65\)), indicating the best recovery of the true underlying groups.
+However, the \textbf{model-based clustering} achieved the highest \emph{Corrected Rand Index} (\(0.75\)) and lowest \emph{Variation of Information (VI)} (\(0.65\)), indicating the best recovery of the true underlying groups (Table \ref{tab:summaryclust-tb-pdf}).
 In comparison, \(k\)-means and hierarchical clustering showed moderate agreement with the true labels.

### 2025-10-13 paper-cardinalR.R

@@ -89,8 +89,8 @@ theme_set(theme_linedraw() +
 source("scripts/additional_functions.R")
 
 
-## ----workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow of."--------
-knitr::include_graphics("figures/cardinalR.pdf")
+## ----workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow of.", eval=FALSE----
+# knitr::include_graphics("figures/cardinalR.pdf")
 
 
 ## -----------------------------------------------------------------------------
@@ -982,7 +982,7 @@ proj_obj1 <- get_projection(projection = projection,
 circle_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.8), 
+  plot_limits = c(-0.8, 0.8), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 5,
@@ -1249,7 +1249,7 @@ clusteredspheres_proj3 <- plot_proj(
 # spherefig
 
 
-## ----sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H"----
+## ----sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"----
 
 circle_proj1 + circle_proj2 + circle_proj3 +
 curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3  +
@@ -2234,15 +2234,15 @@ summary_clust <- tibble::tibble(
 
 ## ----summaryclust-tb-html, eval=knitr::is_html_output()-----------------------
 # summary_clust |>
[…]

### 2025-10-13 paper-cardinalR.Rmd
Added initial NLDR summary statistics

@@
 
+```{r}
+summary_dr <- read_rds("data/five_clusts/dr_summary.rds")
+
+summary_dr <- summary_dr |>
+  mutate(across(where(is.numeric), ~ round(.x, 2)))
+```
+
+```{r summarydr-tb-html, eval=knitr::is_html_output()}
+summary_dr |> 
+  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", col.names = c("Method", "svm", "cte", "rte", "global_score"), tab.pos = "H", linesep = "")
+```
+
+```{r summarydr-tb-pdf, eval=knitr::is_latex_output()}
+summary_dr |> 
+  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("Method", "svm", "cte", "rte", "global_score"), booktabs = T, table.pos = "H", linesep = "")  |>
+  column_spec(1, width = "2cm") |>
+  column_spec(2, width = "2cm") |>
+  column_spec(3, width = "2cm") |>
+  column_spec(4, width = "2cm") |>
+  column_spec(5, width = "2cm")
+```
+
 These methods vary in their ability to preserve local and global data structures (Figure \@ref(fig:fig-nldr-layouts)).
@@
   Metric = c("k-means", "Hierarchical", "Model-based"),
-  `wb.ratio` = round(c(comp_kmeans$wb.ratio, comp_hclust$wb.ratio, comp_mclust$wb.ratio), 2),
-  `Dunn Index` = round(c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn), 2),
-  `Corrected rand` = round(c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand), 2),
-  VI = round(c(comp_kmeans$vi, comp_hclust$vi, comp_mclust$vi), 2)
+  `wb.ratio` = c(comp_kmeans$wb.ratio, comp_hclust$wb.ratio, comp_mclust$wb.ratio),
+  `Dunn Index` = c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn),
+  `Corrected rand` = c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand),
+  VI = c(comp_kmeans$vi, comp_hclust$vi, comp_mclust$vi)
 )
+
+summary_clust <- summary_clust |>
+  mutate(across(where(is.numeric), ~ round(.x, 2)))
 ```

### 2025-10-13 paper-cardinalR.tex
Added initial NLDR summary statistics

@@
 
+\begin{table}
+
+\caption{\label{tab:summarydr-tb-pdf}Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}}
+\toprule
+Method & svm & cte & rte & global\_score\\
+\midrule
+tSNE & 0.98 & 0.42 & 0.70 & 0.46\\
+UMAP & 0.98 & 0.62 & 0.73 & 0.58\\
+PHATE & 0.90 & 0.47 & 0.72 & 0.55\\
+TriMAP & 0.93 & 0.58 & 0.75 & 0.59\\
+PaCMAP & 0.99 & 0.24 & 0.72 & 0.54\\
+PCA & 0.78 & 0.53 & 0.83 & 1.00\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 These methods vary in their ability to preserve local and global data structures (Figure \ref{fig:fig-nldr-layouts}).

### 2025-10-14 paper-cardinalR.Rmd
Added implementation figure

@@
 This design allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
-Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(fig:workflow)', '\\@ref(fig:workflow)'))` illustrates the workflow of `gen_multicluster()`.
-
-```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow of.", eval=FALSE}
-knitr::include_graphics("figures/cardinalR.pdf")
+Figure `r '\\@ref(fig:workflow)'` illustrates the workflow of `gen_multicluster()`.
+
+```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow for generating high-dimensional clustered data.
+The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
+Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset."}
+
+knitr::include_graphics("figures/cardinalR_workflow.png")
 ```

### 2025-10-14 paper-cardinalR.tex
Added implementation figure

@@
 Figure \ref{fig:workflow} illustrates the workflow of \texttt{gen\_multicluster()}.
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{figures/cardinalR_workflow} \caption{Workflow for generating high-dimensional clustered data.
+The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
+Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.}\label{fig:workflow}
+\end{figure}
 

### 2025-10-16 paper-cardinalR.Rmd
Rewrite the evaluation of DR section

@@
   is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.text = element_text(size = 10))
+  theme(legend.position = "none")
 
@@
   is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.text = element_text(size = 10))
+  theme(legend.position = "none")
 ```
@@
 five_clusts_proj1 + five_clusts_proj2 + five_clusts_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
+  plot_layout(ncol = 3) 
 ```
@@
 
-Next, we apply six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
+We applied six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 
@@
   ggplot(aes(x = tSNE1,
-             y = tSNE2,
-             color = cluster))+
-  geom_point(alpha=0.1, size=1) +
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour = '#636363') +
   interior_annotation("a", c(0.08, 0.93)) + 
-  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(aspect.ratio = 1,
-        legend.position = "none")
+  theme(aspect.ratio = 1)
 
@@
   ggplot(aes(x = UMAP1,
-             y = UMAP2,
-             color = cluster))+
-  geom_point(alpha=0.1, size=1) +
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour = '#d95f02') +
   interior_annotation("b", c(0.08, 0.93)) + 
-  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(aspect.ratio = 1,
-        legend.position = "none")
+  theme(aspect.ratio = 1)
 
@@
   ggplot(aes(x = PHATE1,
-             y = PHATE2,
-             color = cluster))+
-  geom_point(alpha=0.1, size=1) +
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1, colour = '#7570b3') +
   interior_annotation("c", c(0.08, 0.93)) + 
-  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(aspect.ratio = 1,
-        legend.position = "none")
+  theme(aspect.ratio = 1)
 
@@
   ggplot(aes(x = TriMAP1,
-             y = TriMAP2,
-             color = cluster))+
-  geom_point(alpha=0.1, size=1) +
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1, colour = '#e7298a') +
   interior_annotation("d", c(0.08, 0.93)) + 
-  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(aspect.ratio = 1,
-        legend.position = "none")
+  theme(aspect.ratio = 1)
 
@@
   ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2,
-             color = cluster))+
-  geom_point(alpha=0.1, size=1) +
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1, colour = '#66a61e') +
   interior_annotation("e", c(0.08, 0.93)) + 
-  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(aspect.ratio = 1,
-        legend.position = "none")
+  theme(aspect.ratio = 1)
 
@@
   ggplot(aes(x = pca1,
-             y = pca2,
-             color = cluster))+
-  geom_point(alpha=0.1, size=1) +
+             y = pca2))+
+  geom_point(alpha=0.1, size=1, colour = '#a6761d') +
   interior_annotation("f", c(0.08, 0.93)) + 
-  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(aspect.ratio = 1,
-        legend.position = "none") 
-```
-
-```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
-
-nldr1 + nldr2 + nldr3 +
-  nldr4 + nldr5 + nldr6 +
-  plot_layout(ncol = 3)
-```
-
-```{r}
-summary_dr <- read_rds("data/five_clusts/dr_summary.rds")
-
-summary_dr <- summary_dr |>
-  mutate(across(where(is.numeric), ~ round(.x, 2)))
-```
-
-```{r summarydr-tb-html, eval=knitr::is_html_output()}
-summary_dr |> 
-  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", col.names = c("Method", "svm", "cte", "rte", "global_score"), tab.pos = "H", linesep = "")
-```
-
-```{r summarydr-tb-pdf, eval=knitr::is_latex_output()}
-summary_dr |> 
-  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("Method", "svm", "cte", "rte", "global_score"), booktabs = T, table.pos = "H", linesep = "")  |>
-  column_spec(1, width = "2cm") |>
-  column_spec(2, width = "2cm") |>
-  column_spec(3, width = "2cm") |>
-  column_spec(4, width = "2cm") |>
-  column_spec(5, width = "2cm")
-```
-
-These methods vary in their ability to preserve local and global data structures (Figure \@ref(fig:fig-nldr-layouts)).
-UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure.
-However, PHATE reveals three non-linear clusters, even though two of them do not show non-linearity.
-UMAP, TriMAP, and PaCMAP successfully maintain the local structures of the data.
-In contrast, tSNE divides the non-linear cluster into sub-clusters.
-Also, tSNE fails to preserve the distances between the clusters.
-PCA, on the other hand, preserves the local structures of the clusters, but some clusters are incorrectly merged that should remain distinct.
+  theme(aspect.ratio = 1) 
+```
+
+```{r}
+error_five_clust <- read_rds("data/five_clusts/five_clusts_rmse.rds")
+
+error_five_clust <- error_five_clust |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(b1 >= 5) |>
+  group_by(method, a1) |>
+  filter(RMSE == min(RMSE)) |>
+  ungroup()
+
+error_plot_five_clust <- plot_rmse(error_five_clust) +
+  scale_x_continuous(breaks = sort(unique(error_five_clust$a1))[c(1, 5, 9, 13, 17, 21, 25)]) +
+  scale_color_manual(values=c("tSNE" = '#636363', "UMAP" = '#d95f02', "PHATE" = '#7570b3', "TriMAP" = '#e7298a', "PaCMAP" = '#66a61e', "PCA" = '#a6761d'))
+```
+
+```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using RMSE for varying binwidth ($a_1$).
+Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
+Layout f is universally poor.
+Layouts a and b are universally optimal.
+Layout b shows six well-separated clusters and latout a shows close clusters, thus layout a is the best choice."}
+
+error_plot_five_clust + wrap_plots(nldr1, nldr2, nldr3, 
+                                    nldr4, nldr5, nldr6, ncol = 2)
+```
+
+To assess their performance, we computed the root mean squared error (RMSE) between the observed high-dimensional data and the fitted values, defined as the high-dimensional mappings of the bin centroids [@gamage2025c].
+A lower RMSE indicates that the method better preserves the high-dimensional structure in its low-dimensional embedding.
+
+As shown in Figure \@ref(fig:fig-nldr-layouts), tSNE (Figure \@ref(fig:fig-nldr-layouts) a) achieved the lowest RMSE across bin widths (mostly tiny), indicating high preservation of both local and global structures.
+Its layout displays well-separated clusters with minimal inter-cluster distances, making it the most faithful representation of the underlying data structure.
+UMAP and PaCMAP (Figure \@ref(fig:fig-nldr-layouts) b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \@ref(fig:fig-nldr-layouts) c) show non-linear cluster structures irrespective of the original structure.
+Also, TriMAP (Figure \@ref(fig:fig-nldr-layouts) d) has high RMSE, and show three clusters with small distances.
+PCA (Figure \@ref(fig:fig-nldr-layouts) f) failed to capture the non-linear geometry, leading to the highest RMSE.
 

### 2025-10-16 paper-cardinalR.tex
Rewrite the evaluation of DR section

--- 2025-10-13 paper-cardinalR.tex
@@
 
-Next, we apply six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
+We applied six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
 
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Six different dimension reduction representations of the five clusters data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA.}\label{fig:fig-nldr-layouts}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using RMSE for varying binwidth ($a_1$).
+Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
+Layout f is universally poor.
+Layouts a and b are universally optimal.
+Layout b shows six well-separated clusters and latout a shows close clusters, thus layout a is the best choice.}\label{fig:fig-nldr-layouts}
 \end{figure}
 
-These methods vary in their ability to preserve local and global data structures (Figure \ref{fig:fig-nldr-layouts}).
-UMAP, PHATE, TriMAP, and PaCMAP effectively separate the five clusters and show the preservation of the global structure.
-However, PHATE reveals three non-linear clusters, even though two of them do not show non-linearity.
-UMAP, TriMAP, and PaCMAP successfully maintain the local structures of the data.
-In contrast, tSNE divides the non-linear cluster into sub-clusters.
-Also, tSNE fails to preserve the distances between the clusters.
-PCA, on the other hand, preserves the local structures of the clusters, but some clusters are incorrectly merged that should remain distinct.
+To assess their performance, we computed the root mean squared error (RMSE) between the observed high-dimensional data and the fitted values, defined as the high-dimensional mappings of the bin centroids \citep{gamage2025c}.
+A lower RMSE indicates that the method better preserves the high-dimensional structure in its low-dimensional embedding.
+
+As shown in Figure \ref{fig:fig-nldr-layouts}, tSNE (Figure \ref{fig:fig-nldr-layouts} a) achieved the lowest RMSE across bin widths (mostly tiny), indicating high preservation of both local and global structures.
+Its layout displays well-separated clusters with minimal inter-cluster distances, making it the most faithful representation of the underlying data structure.
+UMAP and PaCMAP (Figure \ref{fig:fig-nldr-layouts} b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \ref{fig:fig-nldr-layouts} c) show non-linear cluster structures irrespective of the original structure.
+Also, TriMAP (Figure \ref{fig:fig-nldr-layouts} d) has high RMSE, and show three clusters with small distances.
+PCA (Figure \ref{fig:fig-nldr-layouts} f) failed to capture the non-linear geometry, leading to the highest RMSE.
 

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 7 title

@@
 
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}
 
@@
   plot_limits = c(-0.65, 0.5), 
-  title = "b1", 
+  title = "a1", 
   cex = 2, 
@@
   plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
+  title = "a2", 
   cex = 2, 
@@
   plot_limits = c(-0.5, 0.5), 
-  title = "b3", 
+  title = "a3", 
   cex = 2, 
@@
   plot_limits = c(-0.65, 0.7), 
-  title = "c1", 
+  title = "b1", 
   cex = 2, 
@@
   plot_limits = c(-0.7, 0.6), 
-  title = "c2", 
+  title = "b2", 
   cex = 2, 
@@
   plot_limits = c(-0.5, 0.5), 
-  title = "c3", 
+  title = "b3", 
   cex = 2, 
@@
   plot_limits = c(-0.65, 0.7), 
-  title = "e1", 
+  title = "c1", 
   cex = 2, 
@@
   plot_limits = c(-0.8, 0.5), 
-  title = "e2", 
+  title = "c2", 
   cex = 2, 
@@
   plot_limits = c(-0.45, 0.73), 
-  title = "f1", 
+  title = "d1", 
   cex = 2, 
@@
   plot_limits = c(-0.3, 0.46), 
-  title = "f2", 
+  title = "d2", 
   cex = 2, 
@@
   plot_limits = c(-0.35, 0.4), 
-  title = "f3", 
+  title = "d3", 
   cex = 2, 
@@
 
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}
+```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}
 

### 2025-10-22 paper-cardinalR.tex
Updated Figure 7 title

@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.}\label{fig:sphere-proj}
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).}\label{fig:sphere-proj}
 \end{figure}
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `crescent` (a1-a3), `curvycylinder` (b1-b3), `sphericalspiral` (c1-c3), `helicalspiral` (d1-d3), `conicspiral` (e1-e3), and `nonlinear` (f1-f3) data.}\label{fig:triginometric-proj}
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.}\label{fig:triginometric-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.R

@@ -1250,7 +1250,7 @@ clusteredspheres_proj3 <- plot_proj(
 # spherefig
 
 
-## ----sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"----
+## ----sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"----
 
 circle_proj1 + circle_proj2 + circle_proj3 +
 curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3  +
@@ -1544,7 +1544,7 @@ curvycylinder_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.65, 0.5), 
-  title = "b1", 
+  title = "a1", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
@@ -1571,7 +1571,7 @@ curvycylinder_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
+  title = "a2", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
@@ -1598,7 +1598,7 @@ curvycylinder_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
   plot_limits = c(-0.5, 0.5), 
-  title = "b3", 
+  title = "a3", 
   cex = 2, 
   axis_text_size = 5,
   is_color = FALSE)
@@ -1634,7 +1634,7 @@ sphericalspiral_proj1 <- plot_proj(
[…]

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 2 title

@@
 
-```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` (a1-a3) data.", fig.width=12, fig.height=4, fig.pos="H"}
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `orglinearbranches` data.
+Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
+These views highlight the dataset’s underlying branching structure and demonstrate how projections can help reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H"}
 

### 2025-10-22 paper-cardinalR.tex
Updated Figure 2 title

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `orglinearbranches` (a1-a3) data.}\label{fig:branch-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `orglinearbranches` data.
+Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
+These views highlight the dataset’s underlying branching structure and demonstrate how projections can help reveal patterns that are otherwise hidden in higher dimensions.}\label{fig:branch-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.R

@@ -330,7 +330,7 @@ orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 # branchfig
 
 
-## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `orglinearbranches` (a1-a3) data.", fig.width=12, fig.height=4, fig.pos="H"----
+## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `orglinearbranches` data. Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles. These views highlight the dataset’s underlying branching structure and demonstrate how projections can help reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H"----
 
 orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
   plot_layout(ncol = 3, guides = "collect") 

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 1 title

@@
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
-These views highlight the dataset’s underlying branching structure and demonstrate how projections can help reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H"}
+These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H"}
 

### 2025-10-22 paper-cardinalR.tex
Updated Figure 1 title

@@
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
-These views highlight the dataset’s underlying branching structure and demonstrate how projections can help reveal patterns that are otherwise hidden in higher dimensions.}\label{fig:branch-proj}
+These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.}\label{fig:branch-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.Rmd
Added Figure 7 title

@@
 
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `cone` data.", fig.pos="H"}
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `cone` data.", fig.pos="H"}
 #| fig-width: 12
@@
                                               axis_pos_y = -0.2, 
-                                              threshold = 0.01))
+                                              threshold = 0.03))
 
@@
 
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}
+```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.
+Each shows variations in curvature, density, and continuity.
+Various geometric forms captured by the trigonometric shape generators, illustrating how different non-linear geometric structures can challenge to test the robustness of algorithms.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}
 

### 2025-10-22 paper-cardinalR.tex
Added Figure 7 title

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `cone` data.}\label{fig:cone-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `cone` data.}\label{fig:cone-proj}
 \end{figure}
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.}\label{fig:triginometric-proj}
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.
+Each shows variations in curvature, density, and continuity.
+Various geometric forms captured by the trigonometric shape generators, illustrating how different non-linear geometric structures can challenge to test the robustness of algorithms.}\label{fig:triginometric-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 4 caption

@@
 
-```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.", fig.width=12, fig.height=16, fig.pos="H"}
+```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.
+The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
+`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
+These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H"}
 

### 2025-10-22 paper-cardinalR.tex
Updated Figure 4 caption

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.}\label{fig:pyr-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.
+The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
+`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
+These projections illustrate a range of pyramid-like geometries that vary in density and structure.}\label{fig:pyr-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 6 caption

@@
 
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).
+The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern resembling a twisted ring.
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}
 
@@
 
-```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.", fig.width=12, fig.height=8, fig.pos="H"}
+```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.
+The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
+In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
+These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H"}
 

### 2025-10-22 paper-cardinalR.tex
Updated Figure 6 caption

@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).}\label{fig:sphere-proj}
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).
+The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern resembling a twisted ring.
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\text{-}D$ projections, highlighting how projection can distort spatial relationships.}\label{fig:sphere-proj}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.}\label{fig:trefoil-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.
+The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
+In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
+These projections illustrate a range of looping structures in high-dimensions.}\label{fig:trefoil-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 3 caption

@@
 
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `cone` data.", fig.pos="H"}
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `cone` data.
+Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
+These projections show how the conical geometry is preserved.", fig.pos="H"}
 #| fig-width: 12

### 2025-10-22 paper-cardinalR.tex
Updated Figure 3 caption

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `cone` data.}\label{fig:cone-proj}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `cone` data.
+Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
+These projections show how the conical geometry is preserved.}\label{fig:cone-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.R

@@ -441,7 +441,7 @@ cone_proj3 <- plot_proj(
 # conefig
 
 
-## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `cone` data.", fig.pos="H"----
+## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `cone` data. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex. These projections show how the conical geometry is preserved.", fig.pos="H"----
 #| fig-width: 12
 #| fig-height: 4
 

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 7 title

@@
 ```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.
-Each shows variations in curvature, density, and continuity.
-Various geometric forms captured by the trigonometric shape generators, illustrating how different non-linear geometric structures can challenge to test the robustness of algorithms.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}
+The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
+The `sphericalspiral` traces a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
+The `conicspiral` expands radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
+The `nonlinear` dataset exhibits a warped 2D surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
+Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}
 

### 2025-10-22 paper-cardinalR.tex
Updated Figure 7 title

@@
 \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.
-Each shows variations in curvature, density, and continuity.
-Various geometric forms captured by the trigonometric shape generators, illustrating how different non-linear geometric structures can challenge to test the robustness of algorithms.}\label{fig:triginometric-proj}
+The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
+The `sphericalspiral` traces a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
+The `conicspiral` expands radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
+The `nonlinear` dataset exhibits a warped 2D surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
+Each shows variations in curvature, density, and continuity.}\label{fig:triginometric-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.R

@@ -1894,7 +1894,7 @@ nonlinear_proj3 <- plot_proj(
 # triginometricfig
 
 
-## ----triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data. Each shows variations in curvature, density, and continuity. Various geometric forms captured by the trigonometric shape generators, illustrating how different non-linear geometric structures can challenge to test the robustness of algorithms.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"----
+## ----triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data. The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature. The `sphericalspiral` traces a spiral path on a spherical surface, combining circular and vertical motion in a helical form. The `conicspiral` expands radially while ascending, forming a conical helix with twisting variations in a non-radial dimension. The `nonlinear` dataset exhibits a warped 2D surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions. Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"----
 
 curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
 sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 5

@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
+                                              axis_scaled = 1, 
                                               axis_pos_x = -0.57, 

### 2025-10-22 paper-cardinalR.R

@@ -1192,7 +1192,7 @@ proj_obj2 <- get_projection(projection = projection,
                             proj_scale = 1.2, 
                             scaled_data = scaled_data, 
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
+                                              axis_scaled = 1, 
                                               axis_pos_x = -0.57, 
                                               axis_pos_y = -0.57, 
                                               threshold = 0.01))
@@ -1200,7 +1200,7 @@ proj_obj2 <- get_projection(projection = projection,
 clusteredspheres_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
+  plot_limits = c(-0.7, 0.7), 
   title = "c2", 
   cex = 2, 
   axis_text_size = 5,

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 5 caption

@@
 ```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).
-The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern resembling a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}
+The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
+These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}
 

### 2025-10-22 paper-cardinalR.tex
Updated Figure 5 caption

@@
 \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).
-The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern resembling a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\text{-}D$ projections, highlighting how projection can distort spatial relationships.}\label{fig:sphere-proj}
+The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\text{-}D$ projections, highlighting how projection can distort spatial relationships.
+These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\text{-}D$, emphasizing the effects of projection on density, continuity, and separation}\label{fig:sphere-proj}
 \end{figure}

### 2025-10-22 paper-cardinalR.R

@@ -1250,7 +1250,7 @@ clusteredspheres_proj3 <- plot_proj(
 # spherefig
 
 
-## ----sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3). The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern resembling a twisted ring. The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"----
+## ----sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3). The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring. The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships. These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"----
 
 circle_proj1 + circle_proj2 + circle_proj3 +
 curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3  +

### 2025-10-22 paper-cardinalR.Rmd
Updated Figure 7 caption

@@
 The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
-The `sphericalspiral` traces a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
-The `conicspiral` expands radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped 2D surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
+The `sphericalspiral` forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
+The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
+The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
 Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}

### 2025-10-22 paper-cardinalR.tex
Updated Figure 7 caption

@@
 The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
-The `sphericalspiral` traces a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
-The `conicspiral` expands radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped 2D surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
+The `sphericalspiral` forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
+The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
+The `nonlinear` dataset exhibits a warped $2\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
 Each shows variations in curvature, density, and continuity.}\label{fig:triginometric-proj}

### 2025-10-22 paper-cardinalR.R

@@ -1894,7 +1894,7 @@ nonlinear_proj3 <- plot_proj(
 # triginometricfig
 
 
-## ----triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data. The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature. The `sphericalspiral` traces a spiral path on a spherical surface, combining circular and vertical motion in a helical form. The `conicspiral` expands radially while ascending, forming a conical helix with twisting variations in a non-radial dimension. The `nonlinear` dataset exhibits a warped 2D surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions. Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"----
+## ----triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data. The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature. The `sphericalspiral` forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form. The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension. The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions. Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"----
 
 curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
 sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +

### 2025-10-22 paper-cardinalR.Rmd
Start updating langevitour titles

@@
 
-```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `orglinearbranches` data in $4\\text{-}D$."}
+```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$.
+The tour views shows how the linear branches appear from multiple viewing angles."}
 
@@
 
-```{r cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $4\\text{-}D$."}
+```{r cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $3\\text{-}D$.
+Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex."}
 
@@
 
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `cone` data.
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ `cone` data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
@@
 
-```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$."}
+```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$.
+The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
+`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape."}
 
@@
 
-```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, and `clusteredspheres` data in $4\\text{-}D$."}
+```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, and `clusteredspheres` data in $4\\text{-}D$.
+The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships."}
 
@@
 
-```{r trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` and `trefoil3d` data in $4\\text{-}D$."}
+```{r trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` and `trefoil3d` data in $4\\text{-}D$.
+The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
+In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space."}
 
@@
 
-```{r triginometric, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of tthe `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data in $4\\text{-}D$."}
+```{r triginometric, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of tthe `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data in $4\\text{-}D$.
+The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
+The `sphericalspiral` forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
+The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
+The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions."}
 

### 2025-10-22 paper-cardinalR.tex
Start updating langevitour titles

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `cone` data.
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ `cone` data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.

### 2025-10-22 paper-cardinalR.R

@@ -315,7 +315,7 @@ orglinearbranches_proj3 <- plot_proj(
 orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 
 
-## ----branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `orglinearbranches` data in $4\\text{-}D$."----
+## ----branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$. The tour views shows how the linear branches appear from multiple viewing angles."----
 # 
 # branchfig <- bscols(
 #   htmltools::div(
@@ -426,7 +426,7 @@ cone_proj3 <- plot_proj(
 
 
 
-## ----cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $4\\text{-}D$."----
+## ----cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $3\\text{-}D$. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex."----
 # 
 # conefig <- bscols(
 #   htmltools::div(
@@ -441,7 +441,7 @@ cone_proj3 <- plot_proj(
 # conefig
 
 
-## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `cone` data. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex. These projections show how the conical geometry is preserved.", fig.pos="H"----
+## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ `cone` data. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex. These projections show how the conical geometry is preserved.", fig.pos="H"----
 #| fig-width: 12
 #| fig-height: 4
 
@@ -898,7 +898,7 @@ pyrholes_proj3 <- plot_proj(
 
 
 
-## ----pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$."----
[…]

### 2025-10-22 paper-cardinalR.Rmd
Added langevitour to the center

@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    style = "display: grid; grid-template-columns: 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
     orglinearbranches_lang
@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    style = "display: grid; grid-template-columns: 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
     cone_lang
@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
     pyrrect_lang, pyrtri_lang, pyrstar_lang, pyrholes_lang
@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
     circle_lang, curvycycle_lang, clusteredspheres_lang
@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
     trefoil4d_lang, trefoil3d_lang
@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
+    style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
     curvycylinder_lang, sphericalspiral_lang, conicspiral_lang, nonlinear_lang

### 2025-10-22 paper-cardinalR.R

@@ -319,7 +319,12 @@ orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 # 
 # branchfig <- bscols(
 #   htmltools::div(
-#     style = "display: grid; grid-template-columns: 1fr 1fr;",
+#     style = "display: grid; grid-template-columns: 1fr;
+#     gap: 0px;
+#     align-items: start;
+#     justify-items: center;
+#     margin: 0;
+#     padding: 0;",
 #     orglinearbranches_lang
 #   ),
 #   device = "xs"
@@ -430,7 +435,12 @@ cone_proj3 <- plot_proj(
 # 
 # conefig <- bscols(
 #   htmltools::div(
-#     style = "display: grid; grid-template-columns: 1fr 1fr;",
+#     style = "display: grid; grid-template-columns: 1fr;
+#     gap: 0px;
+#     align-items: start;
+#     justify-items: center;
+#     margin: 0;
+#     padding: 0;",
 #     cone_lang
 #   ),
 #   device = "xs"
@@ -902,7 +912,12 @@ pyrholes_proj3 <- plot_proj(
 # 
 # pyrfig <- bscols(
 #   htmltools::div(
-#     style = "display: grid; grid-template-columns: 1fr 1fr;",
+#     style = "display: grid; grid-template-columns: 1fr 1fr;
+#     gap: 0px;
+#     align-items: start;
+#     justify-items: center;
+#     margin: 0;
+#     padding: 0;",
 #     pyrrect_lang, pyrtri_lang, pyrstar_lang, pyrholes_lang
 #   ),
 #   device = "xs"
@@ -1239,7 +1254,12 @@ clusteredspheres_proj3 <- plot_proj(
 # 
 # spherefig <- bscols(
 #   htmltools::div(
-#     style = "display: grid; grid-template-columns: 1fr 1fr;",
+#     style = "display: grid; grid-template-columns: 1fr 1fr;
+#     gap: 0px;
+#     align-items: start;
+#     justify-items: center;
+#     margin: 0;
+#     padding: 0;",
 #     circle_lang, curvycycle_lang, clusteredspheres_lang
 #   ),
 #   device = "xs"
@@ -1466,7 +1486,12 @@ trefoil3d_proj3 <- plot_proj(
 # 
 # trefoilfig <- bscols(
 #   htmltools::div(
-#     style = "display: grid; grid-template-columns: 1fr 1fr;",
[…]

### 2025-10-22 paper-cardinalR.Rmd
Changed the original dimensions of the data

@@
 
-```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, and `clusteredspheres` data in $4\\text{-}D$.
+```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle` data in $4\\text{-}D$, and `clusteredspheres` data in $3\\text{-}D$.
 The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
@@
 
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, `circle` (a1-a3), `curvycycle` (b1-b3), and, $3\\text{-}D$ `clusteredspheres` (c1-c3).
 The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
@@
 
-```{r trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` and `trefoil3d` data in $4\\text{-}D$.
+```{r trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` in $4\\text{-}D$ and `trefoil3d` data in $3\\text{-}D$.
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
@@
 
-```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.
+```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, `trefoil4d` (a1-a3) and $3\\text{-}D$ `trefoil3d` (b1-b3) data.
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.

### 2025-10-22 paper-cardinalR.tex
Changed the original dimensions of the data

@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3).
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, `circle` (a1-a3), `curvycycle` (b1-b3), and, $3\text{-}D$ `clusteredspheres` (c1-c3).
 The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `trefoil4d` (a1-a3) and `trefoil3d` (b1-b3) data.
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, `trefoil4d` (a1-a3) and $3\text{-}D$ `trefoil3d` (b1-b3) data.
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.

### 2025-10-22 paper-cardinalR.R

@@ -1250,7 +1250,7 @@ clusteredspheres_proj3 <- plot_proj(
 
 
 
-## ----sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle`, and `clusteredspheres` data in $4\\text{-}D$. The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring. The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships."----
+## ----sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle` data in $4\\text{-}D$, and `clusteredspheres` data in $3\\text{-}D$. The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring. The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships."----
 # 
 # spherefig <- bscols(
 #   htmltools::div(
@@ -1270,7 +1270,7 @@ clusteredspheres_proj3 <- plot_proj(
 # spherefig
 
 
-## ----sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), and, `clusteredspheres` (c1-c3). The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring. The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships. These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"----
[…]

### 2025-10-22 paper-cardinalR.R
Updated Figure 1 title

@@ -315,7 +315,7 @@ orglinearbranches_proj3 <- plot_proj(
 orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 
 
-## ----branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$. The tour views shows how the linear branches appear from multiple viewing angles."----
+## ----branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$. The tour view shows how the linear branches appear from multiple viewing angles."----
 # 
 # branchfig <- bscols(
 #   htmltools::div(

### 2025-11-07 paper-cardinalR.Rmd
Updated Figure 1 title
Updated HBE

@@
 ```{r}
-arg_branching_tb <- tibble(arg = c("n",
-                          "k"), 
+arg_branching_tb <- tibble(arg = c("n", "k"), 
                           type = c("integer", "integer"),
@@
 ```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$.
-The tour views shows how the linear branches appear from multiple viewing angles."}
+The tour view shows how the linear branches appear from multiple viewing angles."}
 
@@
 ```{r}
-error_five_clust <- read_rds("data/five_clusts/five_clusts_rmse.rds")
+error_five_clust <- read_rds("data/five_clusts/five_clusts_hbe.rds")
 
@@
   group_by(method, a1) |>
-  filter(RMSE == min(RMSE)) |>
+  filter(HBE == min(HBE)) |>
   ungroup()
 
-error_plot_five_clust <- plot_rmse(error_five_clust) +
+error_plot_five_clust <- plot_hbe(error_five_clust) +
   scale_x_continuous(breaks = sort(unique(error_five_clust$a1))[c(1, 5, 9, 13, 17, 21, 25)]) +
@@
 
-```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using RMSE for varying binwidth ($a_1$).
+```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
 Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
@@
 
-To assess their performance, we computed the root mean squared error (RMSE) between the observed high-dimensional data and the fitted values, defined as the high-dimensional mappings of the bin centroids [@gamage2025c].
-A lower RMSE indicates that the method better preserves the high-dimensional structure in its low-dimensional embedding.
-
-As shown in Figure \@ref(fig:fig-nldr-layouts), tSNE (Figure \@ref(fig:fig-nldr-layouts) a) achieved the lowest RMSE across bin widths (mostly tiny), indicating high preservation of both local and global structures.
+To assess their performance, we computed the hexbin error (HBE) between the observed high-dimensional data and the fitted values, defined as the high-dimensional mappings of the bin centroids [@gamage2025c].
+A lower HBE indicates that the method better preserves the high-dimensional structure in its low-dimensional embedding.
+
+As shown in Figure \@ref(fig:fig-nldr-layouts), tSNE (Figure \@ref(fig:fig-nldr-layouts) a) achieved the lowest HBE across bin widths (mostly tiny), indicating high preservation of both local and global structures.
 Its layout displays well-separated clusters with minimal inter-cluster distances, making it the most faithful representation of the underlying data structure.
 UMAP and PaCMAP (Figure \@ref(fig:fig-nldr-layouts) b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \@ref(fig:fig-nldr-layouts) c) show non-linear cluster structures irrespective of the original structure.
-Also, TriMAP (Figure \@ref(fig:fig-nldr-layouts) d) has high RMSE, and show three clusters with small distances.
-PCA (Figure \@ref(fig:fig-nldr-layouts) f) failed to capture the non-linear geometry, leading to the highest RMSE.
+Also, TriMAP (Figure \@ref(fig:fig-nldr-layouts) d) has high HBE, and show three clusters with small distances.
+PCA (Figure \@ref(fig:fig-nldr-layouts) f) failed to capture the non-linear geometry, leading to the highest HBE.
 

### 2025-11-07 paper-cardinalR.tex
Updated HBE

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using RMSE for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
 Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
@@
 
-To assess their performance, we computed the root mean squared error (RMSE) between the observed high-dimensional data and the fitted values, defined as the high-dimensional mappings of the bin centroids \citep{gamage2025c}.
-A lower RMSE indicates that the method better preserves the high-dimensional structure in its low-dimensional embedding.
-
-As shown in Figure \ref{fig:fig-nldr-layouts}, tSNE (Figure \ref{fig:fig-nldr-layouts} a) achieved the lowest RMSE across bin widths (mostly tiny), indicating high preservation of both local and global structures.
+To assess their performance, we computed the hexbin error (HBE) between the observed high-dimensional data and the fitted values, defined as the high-dimensional mappings of the bin centroids \citep{gamage2025c}.
+A lower HBE indicates that the method better preserves the high-dimensional structure in its low-dimensional embedding.
+
+As shown in Figure \ref{fig:fig-nldr-layouts}, tSNE (Figure \ref{fig:fig-nldr-layouts} a) achieved the lowest HBE across bin widths (mostly tiny), indicating high preservation of both local and global structures.
 Its layout displays well-separated clusters with minimal inter-cluster distances, making it the most faithful representation of the underlying data structure.
 UMAP and PaCMAP (Figure \ref{fig:fig-nldr-layouts} b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \ref{fig:fig-nldr-layouts} c) show non-linear cluster structures irrespective of the original structure.
-Also, TriMAP (Figure \ref{fig:fig-nldr-layouts} d) has high RMSE, and show three clusters with small distances.
-PCA (Figure \ref{fig:fig-nldr-layouts} f) failed to capture the non-linear geometry, leading to the highest RMSE.
+Also, TriMAP (Figure \ref{fig:fig-nldr-layouts} d) has high HBE, and show three clusters with small distances.
+PCA (Figure \ref{fig:fig-nldr-layouts} f) failed to capture the non-linear geometry, leading to the highest HBE.
 

### 2025-11-07 scripts/additional_functions.R

@@ -182,15 +182,15 @@ gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axi
 }
 
 
-plot_rmse <- function(error_df) {
+plot_hbe <- function(error_df) {
 
   ggplot(error_df,
          aes(x = a1,
-             y = RMSE,
+             y = HBE,
              colour = method)) +
     geom_point(size = 0.8) +
     geom_line(linewidth = 0.3) +
-    ylab("RMSE") +
+    ylab("HBE") +
     xlab(expression(paste("binwidth (", a[1], ")"))) +
     theme_minimal() +
     theme(panel.border = element_rect(fill = 'transparent'),

### 2025-11-07 paper-cardinalR.Rmd
Break into two lines 1570

@@
 A band of thickness around the knot path is controlled by the `steps` argument, while the number of $\theta$ and $\phi$ values is determined by the `steps` and `n` arguments, respectively (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
-The coordinates are defined as $$X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),$$ where $\theta$ and $\phi$ trace the knot’s path.
+The coordinates are defined as $$X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \\\quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),$$ where $\theta$ and $\phi$ trace the knot’s path.
 
@@
                         exp = c("Adds background noise.",
-                                "Randomizes the rows.",
+                                "Randomizes the rows of input data.",
                                 "Relocates the clusters.",

### 2025-11-07 paper-cardinalR.tex
Break into two lines 1570

@@
 A band of thickness around the knot path is controlled by the \texttt{steps} argument, while the number of \(\theta\) and \(\phi\) values is determined by the \texttt{steps} and \texttt{n} arguments, respectively (Figure \ref{fig:trefoil-proj} a).
-The coordinates are defined as \[X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),\] where \(\theta\) and \(\phi\) trace the knot's path.
+The coordinates are defined as \[X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \\\quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),\] where \(\theta\) and \(\phi\) trace the knot's path.
 
@@
 gen\_bkgnoise & Adds background noise.\\
-randomize\_rows & Randomizes the rows.\\
+randomize\_rows & Randomizes the rows of input data.\\
 relocate\_clusters & Relocates the clusters.\\

### 2025-11-07 paper-cardinalR.Rmd
Try to reduce the number of pages

@@
 `Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
-These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H"}
+These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center'}
 
@@
 The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
-These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}
+These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H"}
 
@@
 In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
-These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H"}
+These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center'}
 
@@
 The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
-Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="100%", fig.align='center', fig.pos="H"}
+Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H"}
 

### 2025-11-07 paper-cardinalR.tex
Try to reduce the number of pages

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.
+
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1}
+
+}
+
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.
 The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}
 
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, `trefoil4d` (a1-a3) and $3\text{-}D$ `trefoil3d` (b1-b3) data.
+
+{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
+
+}
+
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, `trefoil4d` (a1-a3) and $3\text{-}D$ `trefoil3d` (b1-b3) data.
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
+{\centering \includegraphics[width=0.7\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
 

### 2025-11-07 paper-cardinalR.R

@@ -927,7 +927,7 @@ pyrholes_proj3 <- plot_proj(
 # pyrfig
 
 
-## ----pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data. The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges. `Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape. These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H"----
+## ----pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data. The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges. `Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape. These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center'----
 
 pyrrect_proj1 + pyrrect_proj2 + pyrrect_proj3 +
 pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
@@ -1269,7 +1269,7 @@ clusteredspheres_proj3 <- plot_proj(
 # spherefig
 
 
[…]

### 2025-11-07 paper-cardinalR.Rmd
Need to work on rot 2, 3 later

@@
 rot1 <- gen_rotation(p = 4, planes_angles = list(
-  list(plane = c(1, 2), angle = 60),
-  list(plane = c(3, 4), angle = 90)
-))
-
-rot2 <- gen_rotation(p = 4, planes_angles = list(
-  list(plane = c(1, 3), angle = 30)
-))
-
-rot3 <- gen_rotation(p = 4, planes_angles = list(
-  list(plane = c(2, 4), angle = 45)
+  list(plane = c(1, 2), angle = 60), list(plane = c(3, 4), angle = 90)
 ))
@@
   shape = c("gaussian", "cone", "unifcube"),
-  rotation = list(rot1, rot2, rot3),
+  rotation = list(rot1, NULL, NULL),
   is_bkg = FALSE

### 2025-11-07 paper-cardinalR.Rmd
Fixed the latex issue

@@
 
-The simplest structures are approximately linear branches in \gD{}, generated by the `gen_linearbranches(n, k)` function.
+The simplest structures are approximately linear branches in $2\text{-}D$, generated by the `gen_linearbranches(n, k)` function.
 These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability.
@@
 
-To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in \gD{}. Branches $1$ and $2$ are simple parabolas defined as
+To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in $2\text{-}D$. Branches $1$ and $2$ are simple parabolas defined as
 
@@
 
-The `gen_expbranches(n, k)` function creates $k$ exponential branches in \gD{}, radiating from a central region. Each branch $i$ is defined as
+The `gen_expbranches(n, k)` function creates $k$ exponential branches in $2\text{-}D$, radiating from a central region. Each branch $i$ is defined as
 
@@
 The function `gen_gridcube(n, p)` is a wrapper around `geozoo::cube.solid.grid()`.
-It generates a regular lattice of points in \pD{}, producing a uniform hypercube grid.
+It generates a regular lattice of points in $p\text{-}D$, producing a uniform hypercube grid.
 Each axis contains equally spaced coordinates, resulting in a well-defined geometric structure.
@@
 
-This structure appears in \pD{} data when variation is driven by a single factor, such as time-course or sensor measurements, providing a useful test case for trajectory and regression methods [@trapnell2014].
+This structure appears in $p\text{-}D$ data when variation is driven by a single factor, such as time-course or sensor measurements, providing a useful test case for trajectory and regression methods [@trapnell2014].
 
@@
 
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in \gD{} space.
+A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in  space.
 To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
@@
 
-A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in \pD{} space.
+A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in $p\text{-}D$ space.
 The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
@@
 
-The Swiss roll is a standard nonlinear manifold, representing a \gD{} plane curled into $3\text{-}D$.
+The Swiss roll is a standard nonlinear manifold, representing a  plane curled into $3\text{-}D$.
 The `gen_swissroll(n, w)` generates points as $X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad t \sim U(0, 3\pi).$

### 2025-11-07 paper-cardinalR.tex
Fixed the latex issue

@@
 
-The simplest structures are approximately linear branches in $2\text{-}D${}, generated by the \texttt{gen\_linearbranches(n,\ k)} function.
+The simplest structures are approximately linear branches in \(2\text{-}D\), generated by the \texttt{gen\_linearbranches(n,\ k)} function.
 These consist of \(k\) short line segments in the first two dimensions, with added jitter to simulate variability.
@@
 
-To introduce curvature, the \texttt{gen\_curvybranches(n,\ k)} function generates \(k\) curvilinear branches in $2\text{-}D${}. Branches \(1\) and \(2\) are simple parabolas defined as
+To introduce curvature, the \texttt{gen\_curvybranches(n,\ k)} function generates \(k\) curvilinear branches in \(2\text{-}D\). Branches \(1\) and \(2\) are simple parabolas defined as
 
@@
 
-The \texttt{gen\_expbranches(n,\ k)} function creates \(k\) exponential branches in $2\text{-}D${}, radiating from a central region. Each branch \(i\) is defined as
+The \texttt{gen\_expbranches(n,\ k)} function creates \(k\) exponential branches in \(2\text{-}D\), radiating from a central region. Each branch \(i\) is defined as
 
@@
 The function \texttt{gen\_gridcube(n,\ p)} is a wrapper around \texttt{geozoo::cube.solid.grid()}.
-It generates a regular lattice of points in $p\text{-}D${}, producing a uniform hypercube grid.
+It generates a regular lattice of points in \(p\text{-}D\), producing a uniform hypercube grid.
 Each axis contains equally spaced coordinates, resulting in a well-defined geometric structure.
@@
 
-This structure appears in $p\text{-}D${} data when variation is driven by a single factor, such as time-course or sensor measurements, providing a useful test case for trajectory and regression methods \citep{trapnell2014}.
+This structure appears in \(p\text{-}D\) data when variation is driven by a single factor, such as time-course or sensor measurements, providing a useful test case for trajectory and regression methods \citep{trapnell2014}.
 
@@
 
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in $2\text{-}D${} space.
+A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in space.
 To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
@@
 
-A pyramid structure (Figure \ref{fig:pyr-proj}) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in $p\text{-}D${} space.
+A pyramid structure (Figure \ref{fig:pyr-proj}) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in \(p\text{-}D\) space.
 The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
@@
 
-The Swiss roll is a standard nonlinear manifold, representing a $2\text{-}D${} plane curled into \(3\text{-}D\).
+The Swiss roll is a standard nonlinear manifold, representing a plane curled into \(3\text{-}D\).
 The \texttt{gen\_swissroll(n,\ w)} generates points as \(X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad t \sim U(0, 3\pi).\)

### 2025-11-08 paper-cardinalR.Rmd
Updated reference and introduction comments

@@
   \usepackage{float}
-  \newcommand\pD{$p\text{-}D$}
-  \newcommand\gD{$2\text{-}D$}
 output: 
@@
 There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The `geozoo` package (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-The package is `snedata` (@james2025), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
+The package is `snedata` (@james2025) provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding (@hinton2002) and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], to explore how well these techniques preserve data structures in lower dimensions.
 Additionally, `splatter` (@luke2017) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
@@
 `cardinalR` offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimensionality reduction (NLDR) methods, clustering algorithms, and visualization techniques.
+By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimension reduction (NLDR) methods, clustering algorithms, and visualization techniques.
 Moreover, these datasets can serve as benchmark examples for exploring how different algorithmic choices affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
@@
 The core function `gen_hole(df, anchor, r)` removes points from a dataset that fall within a user-specified hypersphere.
-Formally, given data points ($x \in \mathbb{R}^p$), a center ($a \in \mathbb{R}^p$), and radius ($r > 0$), only points satisfying $||x - a||_2 > r$ are retained.
+Formally, given data points ($x \in \mathbb{R}^p$), a center ($a \in \mathbb{R}^p$), and radius ($r > 0$), only points satisfying $\sqrt{\sum_{i=1}^n(x_i-a_i)^2} > r$ are retained.
 The anchor point ($a$) can either be user-specified or default to the dataset mean, and radius ($r$) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
@@
 The function `gen_scurvehole(n, r_hole)` generates an S-curve with a spherical hole by applying `gen_hole()` to the output of `gen_scurve()`.
-This structure has been used in prior diagnostic studies of NLDR methods, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
+This structure has been used in prior diagnostic studies of NLDR methods [@tenenbaum2000, @van2007], since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
 The second wrapper, `gen_unifcubehole(n, p, r_hole)`, generates uniformly sampled cube data with a hyperspherical hole.
@@
 
-We applied six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
+We applied six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) [@jolliffe2011], tSNE, uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 

### 2025-11-08 paper-cardinalR.tex
Updated reference and introduction comments

@@
 There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-The package is \texttt{snedata} (\citet{james2025}), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures.
+The package is \texttt{snedata} (\citet{james2025}) provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding (\citet{hinton2002}) and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, to explore how well these techniques preserve data structures in lower dimensions.
 Additionally, \texttt{splatter} (\citet{luke2017}) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
@@
 \texttt{cardinalR} offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimensionality reduction (NLDR) methods, clustering algorithms, and visualization techniques.
+By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimension reduction (NLDR) methods, clustering algorithms, and visualization techniques.
 Moreover, these datasets can serve as benchmark examples for exploring how different algorithmic choices affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
@@
 The core function \texttt{gen\_hole(df,\ anchor,\ r)} removes points from a dataset that fall within a user-specified hypersphere.
-Formally, given data points (\(x \in \mathbb{R}^p\)), a center (\(a \in \mathbb{R}^p\)), and radius (\(r > 0\)), only points satisfying \(||x - a||_2 > r\) are retained.
+Formally, given data points (\(x \in \mathbb{R}^p\)), a center (\(a \in \mathbb{R}^p\)), and radius (\(r > 0\)), only points satisfying \(\sqrt{\sum_{i=1}^n(x_i-a_i)^2} > r\) are retained.
 The anchor point (\(a\)) can either be user-specified or default to the dataset mean, and radius (\(r\)) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
@@
 The function \texttt{gen\_scurvehole(n,\ r\_hole)} generates an S-curve with a spherical hole by applying \texttt{gen\_hole()} to the output of \texttt{gen\_scurve()}.
-This structure has been used in prior diagnostic studies of NLDR methods, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
+This structure has been used in prior diagnostic studies of NLDR methods \citep[\citet{van2007}]{tenenbaum2000}, since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
 The second wrapper, \texttt{gen\_unifcubehole(n,\ p,\ r\_hole)}, generates uniformly sampled cube data with a hyperspherical hole.
@@
 
-We applied six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) \citep{jolliffe2011}, t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
+We applied six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) \citep{jolliffe2011}, tSNE, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}.
 

### 2025-11-08 paper-cardinalR.Rmd
Added conclusion sentence

@@
 
-These structures are designed to support not only algorithm diagnostics, but also teaching high-dimensional concepts, benchmarking reproducibility, and evaluating hyperparameter sensitivity.
+These structures are designed to support not only algorithm diagnostics, but also teaching high-dimensional concepts, benchmarking reproducibility, and evaluating hyper-parameter sensitivity.
 By allowing users to adjust dimensionality, sample size, noise, and clustering properties, the package promotes transparent experimentation and comparative model evaluation.
-
-Future extensions of `cardinalR` may include biologically inspired or application-driven data structures would further broaden its utility in domains such as bioinformatics, forensic science, and spatial analysis.
+Together, these capabilities make `cardinalR` a versatile tool for generating interpretable, high-dimensional datasets that advance research, teaching, and evaluation of data-analytic methods.
+
+Future extensions of `cardinalR` may include biologically inspired or application-driven data structures that would further broaden its utility in domains such as bioinformatics, forensic science, and spatial analysis.
 

### 2025-11-08 paper-cardinalR.tex
Added conclusion sentence

@@
 
-These structures are designed to support not only algorithm diagnostics, but also teaching high-dimensional concepts, benchmarking reproducibility, and evaluating hyperparameter sensitivity.
+These structures are designed to support not only algorithm diagnostics, but also teaching high-dimensional concepts, benchmarking reproducibility, and evaluating hyper-parameter sensitivity.
 By allowing users to adjust dimensionality, sample size, noise, and clustering properties, the package promotes transparent experimentation and comparative model evaluation.
-
-Future extensions of \texttt{cardinalR} may include biologically inspired or application-driven data structures would further broaden its utility in domains such as bioinformatics, forensic science, and spatial analysis.
+Together, these capabilities make \texttt{cardinalR} a versatile tool for generating interpretable, high-dimensional datasets that advance research, teaching, and evaluation of data-analytic methods.
+
+Future extensions of \texttt{cardinalR} may include biologically inspired or application-driven data structures that would further broaden its utility in domains such as bioinformatics, forensic science, and spatial analysis.
 

### 2025-11-08 paper-cardinalR.Rmd
Added what is illustarting

@@
 
-To illustrate, we generate a dataset with five clusters in $4\text{-}D$, each representing distinct geometric characteristics: a *helical spiral* (elongated and twisted), a *hemisphere* (curved surface), a *uniform cube* (isotropic distribution), a *cone* (density gradient), and a *Gaussian* cluster (compact and spherical) (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
+To illustrate how high-dimensional clustered data can be generated using `cardinalR`, we generate a dataset with five clusters in $4\text{-}D$, each representing distinct geometric characteristics: a *helical spiral* (elongated and twisted), a *hemisphere* (curved surface), a *uniform cube* (isotropic distribution), a *cone* (density gradient), and a *Gaussian* cluster (compact and spherical) (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
 Each cluster has a unique number of points and scaling factor, representing variation in cluster size and spread across the $4\text{-}D$ space.

### 2025-11-08 paper-cardinalR.tex
Added what is illustarting

@@
 
-To illustrate, we generate a dataset with five clusters in \(4\text{-}D\), each representing distinct geometric characteristics: a \emph{helical spiral} (elongated and twisted), a \emph{hemisphere} (curved surface), a \emph{uniform cube} (isotropic distribution), a \emph{cone} (density gradient), and a \emph{Gaussian} cluster (compact and spherical) (Figure \ref{fig:highd-proj}).
+To illustrate how high-dimensional clustered data can be generated using \texttt{cardinalR}, we generate a dataset with five clusters in \(4\text{-}D\), each representing distinct geometric characteristics: a \emph{helical spiral} (elongated and twisted), a \emph{hemisphere} (curved surface), a \emph{uniform cube} (isotropic distribution), a \emph{cone} (density gradient), and a \emph{Gaussian} cluster (compact and spherical) (Figure \ref{fig:highd-proj}).
 Each cluster has a unique number of points and scaling factor, representing variation in cluster size and spread across the \(4\text{-}D\) space.

### 2025-11-08 paper-cardinalR.Rmd
Added supplementary section with gen_sum and gen_product

@@
 These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].
+
+# Supplementary numeric generators
+
+Two helper functions, `gen_nproduct()` and `gen_nsum()`, generate numeric vectors of positive integers that approximately satisfy a user-specified target product or sum, respectively.
+
+The function `gen_nsum(n, k)` divides a total sum `n` into `k` positive integers.
+It first assigns an equal base value to each element and then randomly distributes any remainder, ensuring the elements sum exactly to `n`.
+
+```{r, echo=TRUE}
+gen_nsum(n = 100, k = 3)
+```
+
+The function `gen_nproduct(n, p)` aims to produce `p` positive integers whose product is approximately `n`.
+It starts with all elements equal to the rounded $p^{th}$ root of `n` and iteratively adjusts elements up or down in a randomized manner until the product is within a small tolerance of `n`.
+This accommodates the fact that exact integer solutions for a given product are often impossible.
+
+```{r, echo=TRUE}
+gen_nproduct(n = 500, p = 4)
+```

### 2025-11-08 paper-cardinalR.tex
Added supplementary section with gen_sum and gen_product

@@
 
+\section{Supplementary numeric generators}\label{supplementary-numeric-generators}
+
+Two helper functions, \texttt{gen\_nproduct()} and \texttt{gen\_nsum()}, generate numeric vectors of positive integers that approximately satisfy a user-specified target product or sum, respectively.
+
+The function \texttt{gen\_nsum(n,\ k)} divides a total sum \texttt{n} into \texttt{k} positive integers.
+It first assigns an equal base value to each element and then randomly distributes any remainder, ensuring the elements sum exactly to \texttt{n}.
+
+\begin{verbatim}
+gen_nsum(n = 100, k = 3)
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 33 34 33
+\end{verbatim}
+
+The function \texttt{gen\_nproduct(n,\ p)} aims to produce \texttt{p} positive integers whose product is approximately \texttt{n}.
+It starts with all elements equal to the rounded \(p^{th}\) root of \texttt{n} and iteratively adjusts elements up or down in a randomized manner until the product is within a small tolerance of \texttt{n}.
+This accommodates the fact that exact integer solutions for a given product are often impossible.
+
+\begin{verbatim}
+gen_nproduct(n = 500, p = 4)
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 4 5 5 5
+\end{verbatim}
+
 \bibliography{paper-cardinalR.bib}

### 2025-11-08 paper-cardinalR.R

@@ -2267,3 +2267,11 @@ summary_clust |>
   column_spec(4, width = "2.5cm") |>
   column_spec(5, width = "2cm")
 
+
+## ----echo=TRUE----------------------------------------------------------------
+gen_nsum(n = 100, k = 3)
+
+
+## ----echo=TRUE----------------------------------------------------------------
+gen_nproduct(n = 500, p = 4)
+

### 2025-11-08 paper-cardinalR.Rmd
Trying to make the paper shorter

@@
 
-```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow for generating high-dimensional clustered data.
+```{r workflow, out.width = "70%", fig.pos="H", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data.
 The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
@@
 ```{r, echo=TRUE}
-# Define example rotation matrices for 4D space
 rot1 <- gen_rotation(p = 4, planes_angles = list(
@@
 
-# Generate the clustered dataset
 clust_data <- gen_multicluster(
@@
 Each grid point is subsequently mapped into Cartesian space via the standard hyperspherical-to-Cartesian transformation,
+
+<!-- $$ -->
+<!-- X_i = -->
+<!-- \begin{cases} -->
+<!-- \displaystyle \prod_{j=1}^{i-1} \sin(\theta_j) , \cos(\theta_i), & 1 \le i \le p-1, \\ -->
+<!-- \displaystyle \prod_{j=1}^{p-1} \sin(\theta_j), & i = p, -->
+<!-- \end{cases} -->
+<!-- $$ -->
+
+<!-- with the convention that $\prod_{j=1}^{0} \sin(\theta_j) = 1$. -->
 
@@
 
-## To generate data
 five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
@@
 The included structures cover a wide range of diagnostic settings.
-Branching shapes facilitate the study of continuity and topological preservation, the Scurve with a hole allows investigation of incomplete manifolds, and clustered spheres assess separability on curved surfaces.
+Branching shapes facilitate the study of continuity and topological preservation, the S-curve with a hole allows investigation of incomplete manifolds, and clustered spheres assess separability on curved surfaces.
 The Möbius strip introduces challenges from non-orientable geometry, while gridded cubes and pyrholes test spatial regularity and clustering in sparse, non-convex regions.

### 2025-11-08 paper-cardinalR.tex
Trying to make the paper shorter

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/cardinalR_workflow} \caption{Workflow for generating high-dimensional clustered data.
+
+{\centering \includegraphics[width=0.7\linewidth]{figures/cardinalR_workflow}
+
+}
+
+\caption{Workflow for generating high-dimensional clustered data.
 The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
@@
 \begin{verbatim}
-# Define example rotation matrices for 4D space
 rot1 <- gen_rotation(p = 4, planes_angles = list(

### 2025-11-08 paper-cardinalR.Rmd
Started adding fig.alt

@@
 The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
-Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset."}
+Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.", fig.alt="Diagram showing the workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions."}
 
@@
 ```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$.
-The tour view shows how the linear branches appear from multiple viewing angles."}
+The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles."}
 
@@
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
-These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H"}
+These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three 2-D projections of the 4-D orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure."}
 
@@
 ```{r cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $3\\text{-}D$.
-Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex."}
+Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex."}
 
@@
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
-These projections show how the conical geometry is preserved.", fig.pos="H"}
+These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt="Three 2-D projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex."}
 #| fig-width: 12
@@
 The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
-`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape."}
+`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.", fig.alt="Interactive langevitour of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing dense bases, triangular spreads, branching points, and hollow regions."}
 
@@
 `Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
-These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center'}
+These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Nine 2-D projections of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions."}
 
@@
 The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships."}
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.alt="Interactive langevitour of three datasets: circle and curvycycle in 4-D showing closed-loop and wavy twisted-ring structures, and clusteredspheres in 3-D showing multiple compact spherical clusters with slight overlap in projections."}
 
@@
 The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
-These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H"}
+These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Nine 2-D projections of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation."}
 
@@
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
-In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space."}
+In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.", fig.alt="Interactive langevitour of two trefoil knot datasets: trefoil4d in 4-D showing complex twisting and looping patterns, and trefoil3d in 3-D showing a simpler compact knot-like form."}
 
@@
 In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
-These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center'}
+These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Six 2-D projections of trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions."}
 
@@
 The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions."}
+The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.", fig.alt="Interactive langevitour of six datasets in 4-D: crescent, curvycylinder, sphericalspiral, helicalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral paths, conical helices, and warped 2-D surfaces with smooth and sharp variations."}
 
@@
 The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
-Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H"}
+Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Twelve 2-D projections of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity."}
 
@@
 
-```{r highd-data, eval=knitr::is_html_output(), fig.cap="`langevitour` output showing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster."}
+```{r highd-data, eval=knitr::is_html_output(), fig.cap="`langevitour` output showing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in 4-D: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure."}
 
@@
 ```{r highd-proj, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
-The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green."}
-#| fig-width: 12
-#| fig-height: 4
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three 2-D projections of five 4-D synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=4}
 
@@
 Layouts a and b are universally optimal.
-Layout b shows six well-separated clusters and latout a shows close clusters, thus layout a is the best choice."}
+Layout b shows six well-separated clusters and layout a shows close clusters, thus layout a is the best choice.", fig.alt="Composite plot assessing six NLDR layouts of five clusters data using HBE over varying binwidths, showing tSNE, UMAP, PAHTE, TriMAP, PaCMAP, and PCA.
+Layouts a and b are optimal, f is poor, with color-coded lines matching the scatterplots of each layout."}
 

### 2025-11-08 paper-cardinalR.tex
Started adding fig.alt

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/cardinalR_workflow}
+{\centering \includegraphics[width=0.7\linewidth,alt={Diagram showing the workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.}]{figures/cardinalR_workflow}
 
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `orglinearbranches` data.
+\includegraphics[width=1\linewidth,alt={Three 2-D projections of the 4-D orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `orglinearbranches` data.
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ `cone` data.
+\includegraphics[width=1\linewidth,alt={Three 2-D projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ `cone` data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/pyr-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Nine 2-D projections of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.}]{paper-cardinalR_files/figure-latex/pyr-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Nine 2-D projections of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Six 2-D projections of trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.}]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
+{\centering \includegraphics[width=0.7\linewidth,alt={Twelve 2-D projections of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity.}]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
 
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
+\includegraphics[width=1\linewidth,alt={Three 2-D projections of five 4-D synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.}]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
 The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:highd-proj}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth,alt={Composite plot assessing six NLDR layouts of five clusters data using HBE over varying binwidths, showing tSNE, UMAP, PAHTE, TriMAP, PaCMAP, and PCA.
+Layouts a and b are optimal, f is poor, with color-coded lines matching the scatterplots of each layout.}]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
 Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
@@
 Layouts a and b are universally optimal.
-Layout b shows six well-separated clusters and latout a shows close clusters, thus layout a is the best choice.}\label{fig:fig-nldr-layouts}
+Layout b shows six well-separated clusters and layout a shows close clusters, thus layout a is the best choice.}\label{fig:fig-nldr-layouts}
 \end{figure}

### 2025-11-08 paper-cardinalR.Rmd
Added fig.alt

@@
 The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
-Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.", fig.alt="Diagram showing the workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions."}
+Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions."}
 
@@
 `Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
-These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Nine 2-D projections of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions."}
+These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three 2-D projections from each of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions."}
 
@@
 The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
-These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Nine 2-D projections of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation."}
+These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Three 2-D projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation."}
 
@@
 In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
-These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Six 2-D projections of trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions."}
+These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three 2-D projections from each of two trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions."}
 
@@
 The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
-Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Twelve 2-D projections of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity."}
+Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Three 2-D projections from each of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity."}
 

### 2025-11-08 paper-cardinalR.tex
Added fig.alt

@@
 
-{\centering \includegraphics[width=0.7\linewidth,alt={Diagram showing the workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.}]{figures/cardinalR_workflow}
+{\centering \includegraphics[width=0.7\linewidth,alt={Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.}]{figures/cardinalR_workflow}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Nine 2-D projections of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.}]{paper-cardinalR_files/figure-latex/pyr-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Three 2-D projections from each of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.}]{paper-cardinalR_files/figure-latex/pyr-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Nine 2-D projections of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Three 2-D projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Six 2-D projections of trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.}]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Three 2-D projections from each of two trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.}]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth,alt={Twelve 2-D projections of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity.}]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
+{\centering \includegraphics[width=0.7\linewidth,alt={Three 2-D projections from each of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity.}]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
 

### 2025-11-08 paper-cardinalR.Rmd
Fixed the rotation issue

@@
   shape = c("gaussian", "cone", "unifcube"),
-  rotation = list(rot1, NULL, NULL),
+  rotation = rot1,
   is_bkg = FALSE

### 2025-11-08 paper-cardinalR.R

@@ -146,7 +146,7 @@ clust_data <- gen_multicluster(
   ), nrow = 3, byrow = TRUE),
   scale = c(3, 1, 2),
   shape = c("gaussian", "cone", "unifcube"),
-  rotation = list(rot1, NULL, NULL),
+  rotation = rot1,
   is_bkg = FALSE
 )
 

### 2025-11-09 paper-cardinalR.Rmd
Updated figure positions

@@
 
-```{r workflow, out.width = "70%", fig.pos="H", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data.
+```{r workflow, out.width = "70%", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data.
 The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
-Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions."}
+Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.", fig.pos="!ht"}
 
@@
 ```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$.
-The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles."}
+The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles.", fig.pos="!ht"}
 
@@
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
-These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three 2-D projections of the 4-D orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure."}
+These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three 2-D projections of the 4-D orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"}
 
@@
 ```{r cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $3\\text{-}D$.
-Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex."}
+Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex.", fig.pos="!ht"}
 
@@
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
-These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt="Three 2-D projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex."}
-#| fig-width: 12
-#| fig-height: 4
+These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt="Three 2-D projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.", fig.pos="!ht", fig.width=12, fig.height=4}
 
@@
 The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
-`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.", fig.alt="Interactive langevitour of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing dense bases, triangular spreads, branching points, and hollow regions."}
+`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.", fig.alt="Interactive langevitour of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing dense bases, triangular spreads, branching points, and hollow regions.", fig.pos="!ht"}
 
@@
 `Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
-These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three 2-D projections from each of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions."}
+These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three 2-D projections from each of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.", fig.pos="!ht"}
 
@@
 The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.alt="Interactive langevitour of three datasets: circle and curvycycle in 4-D showing closed-loop and wavy twisted-ring structures, and clusteredspheres in 3-D showing multiple compact spherical clusters with slight overlap in projections."}
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.alt="Interactive langevitour of three datasets: circle and curvycycle in 4-D showing closed-loop and wavy twisted-ring structures, and clusteredspheres in 3-D showing multiple compact spherical clusters with slight overlap in projections.", fig.pos="!ht"}
 
@@
 The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
-These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Three 2-D projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation."}
+These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Three 2-D projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.", fig.pos="!ht"}
 
@@
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
-In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.", fig.alt="Interactive langevitour of two trefoil knot datasets: trefoil4d in 4-D showing complex twisting and looping patterns, and trefoil3d in 3-D showing a simpler compact knot-like form."}
+In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.", fig.alt="Interactive langevitour of two trefoil knot datasets: trefoil4d in 4-D showing complex twisting and looping patterns, and trefoil3d in 3-D showing a simpler compact knot-like form.", fig.pos="!ht"}
 
@@
 In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
-These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three 2-D projections from each of two trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions."}
+These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three 2-D projections from each of two trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.", fig.pos="!ht"}
 
@@
 The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.", fig.alt="Interactive langevitour of six datasets in 4-D: crescent, curvycylinder, sphericalspiral, helicalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral paths, conical helices, and warped 2-D surfaces with smooth and sharp variations."}
+The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.", fig.alt="Interactive langevitour of six datasets in 4-D: crescent, curvycylinder, sphericalspiral, helicalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral paths, conical helices, and warped 2-D surfaces with smooth and sharp variations.", fig.pos="!ht"}
 
@@
 The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
-Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Three 2-D projections from each of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity."}
+Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Three 2-D projections from each of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity.", fig.pos="!ht"}
 
@@
 
-```{r highd-data, eval=knitr::is_html_output(), fig.cap="`langevitour` output showing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in 4-D: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure."}
+```{r highd-data, eval=knitr::is_html_output(), fig.cap="`langevitour` output showing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in 4-D: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht"}
 
@@
 
-```{r highd-proj, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
-The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three 2-D projections of five 4-D synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=4}
+```{r highd-proj, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three 2-D projections of five 4-D synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=4, fig.pos="!ht"}
 
@@
 Layout b shows six well-separated clusters and layout a shows close clusters, thus layout a is the best choice.", fig.alt="Composite plot assessing six NLDR layouts of five clusters data using HBE over varying binwidths, showing tSNE, UMAP, PAHTE, TriMAP, PaCMAP, and PCA.
-Layouts a and b are optimal, f is poor, with color-coded lines matching the scatterplots of each layout."}
+Layouts a and b are optimal, f is poor, with color-coded lines matching the scatterplots of each layout.", fig.pos="!ht"}
 

### 2025-11-09 paper-cardinalR.R

@@ -89,7 +89,7 @@ theme_set(theme_linedraw() +
 source("scripts/additional_functions.R")
 
 
-## ----workflow, out.width = "70%", fig.pos="H", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data. The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise). Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions."----
+## ----workflow, out.width = "70%", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data. The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise). Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.", fig.pos="!ht"----
 
 knitr::include_graphics("figures/cardinalR_workflow.png")
 
@@ -303,7 +303,7 @@ orglinearbranches_proj3 <- plot_proj(
 orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 
 
-## ----branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles."----
+## ----branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles.", fig.pos="!ht"----
 # 
 # branchfig <- bscols(
 #   htmltools::div(
[…]

### 2025-11-11 paper-cardinalR.Rmd
Swap usage and implementation sections

@@
 There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The `geozoo` package (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-The package is `snedata` (@james2025) provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding (@hinton2002) and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], to explore how well these techniques preserve data structures in lower dimensions.
+The package `snedata` (@james2025) provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding (@hinton2002) and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], to explore how well these techniques preserve data structures in lower dimensions.
 Additionally, `splatter` (@luke2017) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
@@
 
-# Implementation
-
-The `cardinalR` package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes.
-The main function, `gen_multicluster()`, combines these shapes into a single dataset by applying scaling, rotation, and translation through `gen_rotation()`.
-Each generated shape is assigned a unique cluster label.
-This design allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
+# Usage
+
+The `cardinalR` package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes, which can then be combined into a single dataset including scaling, rotation and translation.
+The package is available on CRAN, and the source is available on GitHub at [JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR).
+
+The main function, `gen_multicluster()`, is an all-in-one function that includes generating individual shapes, handles scaling and rotating of these shapes, and combines the result into a single unified dataset.
+This function and associated workflow allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
 Figure `r '\\@ref(fig:workflow)'` illustrates the workflow of `gen_multicluster()`.
@@
 
-# Usage
-
-The `cardinalR` R package is available on GitHub at [JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR).
-
-## Main function
-
-The main function of the package is `gen_multicluster()`, which generates datasets consisting of multiple clusters with user-specified characteristics.
 Users can control the number of clusters (`k`), and the number of points in each cluster (`n`).
@@
 
+The following is an example of a three shape multiclustered dataset. The first shape is gaussian, the second conical, and the third a cube.
+
+```{r, echo=TRUE}
+clust_data <- gen_multicluster(
+  n = c(200, 300, 500),
+  k = 3,
+  loc = matrix(c(
+    0, 0, 0, 0,
+    5, 9, 0, 0,
+    3, 4, 10, 7
+  ), nrow = 3, byrow = TRUE),
+  scale = c(3, 1, 2),
+  shape = c("gaussian", "cone", "unifcube"),
+  is_bkg = FALSE
+)
+```
+
+Here, the shapes have 200, 300 and 500 points respectively (`n`), are positioned in 4-D space according to a location matrix, `loc`, and stretched according to the `scale`.
+The details of the individual shape generators, and the noise elements are contained in the following sections.
+
+# Implementation
+
+The main function of the package is `gen_multicluster()`, which generates datasets consisting of multiple clusters with user-specified characteristics. 
 To maintain consistency across generators, the function identifies the arguments required by each chosen generator function and supplies only those arguments that are valid for that specific generator.
@@
   column_spec(3, width = "8cm")
-```
-
-The following example demonstrates how to use `gen_multicluster()` to create a $4\text{-}D$ dataset with three clusters of different shapes and orientations:
-
-```{r, echo=TRUE}
-rot1 <- gen_rotation(p = 4, planes_angles = list(
-  list(plane = c(1, 2), angle = 60), list(plane = c(3, 4), angle = 90)
-))
-
-clust_data <- gen_multicluster(
-  n = c(200, 300, 500),
-  k = 3,
-  loc = matrix(c(
-    0, 0, 0, 0,
-    5, 9, 0, 0,
-    3, 4, 10, 7
-  ), nrow = 3, byrow = TRUE),
-  scale = c(3, 1, 2),
-  shape = c("gaussian", "cone", "unifcube"),
-  rotation = rot1,
-  is_bkg = FALSE
-)
-
 ```

### 2025-11-11 paper-cardinalR.Rmd
Removed supplementary section

@@
 These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].
-
-# Supplementary numeric generators
-
-Two helper functions, `gen_nproduct()` and `gen_nsum()`, generate numeric vectors of positive integers that approximately satisfy a user-specified target product or sum, respectively.
-
-The function `gen_nsum(n, k)` divides a total sum `n` into `k` positive integers.
-It first assigns an equal base value to each element and then randomly distributes any remainder, ensuring the elements sum exactly to `n`.
-
-```{r, echo=TRUE}
-gen_nsum(n = 100, k = 3)
-```
-
-The function `gen_nproduct(n, p)` aims to produce `p` positive integers whose product is approximately `n`.
-It starts with all elements equal to the rounded $p^{th}$ root of `n` and iteratively adjusts elements up or down in a randomized manner until the product is within a small tolerance of `n`.
-This accommodates the fact that exact integer solutions for a given product are often impossible.
-
-```{r, echo=TRUE}
-gen_nproduct(n = 500, p = 4)
-```

### 2025-11-11 paper-cardinalR.tex
Updated figure positions
Removed supplementary section

@@
 There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-The package is \texttt{snedata} (\citet{james2025}) provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding (\citet{hinton2002}) and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, to explore how well these techniques preserve data structures in lower dimensions.
+The package \texttt{snedata} (\citet{james2025}) provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding (\citet{hinton2002}) and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, to explore how well these techniques preserve data structures in lower dimensions.
 Additionally, \texttt{splatter} (\citet{luke2017}) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
@@
 
-\section{Implementation}\label{implementation}
-
-The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes.
-The main function, \texttt{gen\_multicluster()}, combines these shapes into a single dataset by applying scaling, rotation, and translation through \texttt{gen\_rotation()}.
-Each generated shape is assigned a unique cluster label.
-This design allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
+\section{Usage}\label{usage}
+
+The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes, which can then be combined into a single dataset including scaling, rotation and translation.
+The package is available on CRAN, and the source is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.
+
+The main function, \texttt{gen\_multicluster()}, is an all-in-one function that includes generating individual shapes, handles scaling and rotating of these shapes, and combines the result into a single unified dataset.
+This function and associated workflow allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
 Figure \ref{fig:workflow} illustrates the workflow of \texttt{gen\_multicluster()}.
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 
@@
 
-\section{Usage}\label{usage}
-
-The \texttt{cardinalR} R package is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.
-
-\subsection{Main function}\label{main-function}
-
-The main function of the package is \texttt{gen\_multicluster()}, which generates datasets consisting of multiple clusters with user-specified characteristics.
 Users can control the number of clusters (\texttt{k}), and the number of points in each cluster (\texttt{n}).
@@
 
-To maintain consistency across generators, the function identifies the arguments required by each chosen generator function and supplies only those arguments that are valid for that specific generator.
-This design enables the combination of cluster types with differing parameter requirements within the same dataset.
-When clusters are generated with fewer dimensions than others, the function augments the lower-dimensional clusters with additional Gaussian noise variables so that all clusters are represented in the same dimensional space.
-These noise dimensions are drawn independently from normal distributions
-
-\[
-X \sim \mathcal{N}(m, s^2),
-\]
-
-where the mean (\(m\)) is set to the average of the cluster coordinates and the standard deviation (\(s\)) defaults to \(0.2\).
-
-An optional argument, \texttt{is\_bkg}, adds background noise drawn from a multivariate normal distribution centered on the dataset's overall mean with standard deviations matching the observed spread.
-Extra arguments (\texttt{...}) can be passed to cluster generators, allowing further control over per-cluster characteristics like radius of the sphere.
-The main arguments of the \texttt{gen\_multicluster()} function are shown in Table \ref{tab:main-tb-pdf}.
-
-\begin{table}
-
-\caption{\label{tab:main-tb-pdf}The main arguments for gen\_multicluster().}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Argument & Type & Explanation\\
-\midrule
-n & numeric (vector) & Number of points in each cluster.\\
-k & numeric & Number of clusters.\\
-loc & numeric (matrix) & Locations/centroids of clusters.\\
-scale & numeric (vector) & Scaling factors of clusters.\\
-shape & character (vector) & Shapes of clusters.\\
-rotation & numeric (list) & Rotation matrices, one per cluster.\\
-is\_bkg & boolean & Background noise should exist or not.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-The following example demonstrates how to use \texttt{gen\_multicluster()} to create a \(4\text{-}D\) dataset with three clusters of different shapes and orientations:
-
-\begin{verbatim}
-rot1 <- gen_rotation(p = 4, planes_angles = list(
-  list(plane = c(1, 2), angle = 60), list(plane = c(3, 4), angle = 90)
-))
-
+The following is an example of a three shape multiclustered dataset. The first shape is gaussian, the second conical, and the third a cube.
+
+\begin{verbatim}
 clust_data <- gen_multicluster(
@@
   shape = c("gaussian", "cone", "unifcube"),
-  rotation = rot1,
   is_bkg = FALSE
@@
 \end{verbatim}
+
+Here, the shapes have 200, 300 and 500 points respectively (\texttt{n}), are positioned in 4-D space according to a location matrix, \texttt{loc}, and stretched according to the \texttt{scale}.
+The details of the individual shape generators, and the noise elements are contained in the following sections.
+
+\section{Implementation}\label{implementation}
+
+The main function of the package is \texttt{gen\_multicluster()}, which generates datasets consisting of multiple clusters with user-specified characteristics.
+To maintain consistency across generators, the function identifies the arguments required by each chosen generator function and supplies only those arguments that are valid for that specific generator.
+This design enables the combination of cluster types with differing parameter requirements within the same dataset.
+When clusters are generated with fewer dimensions than others, the function augments the lower-dimensional clusters with additional Gaussian noise variables so that all clusters are represented in the same dimensional space.
+These noise dimensions are drawn independently from normal distributions
+
+\[
+X \sim \mathcal{N}(m, s^2),
+\]
+
+where the mean (\(m\)) is set to the average of the cluster coordinates and the standard deviation (\(s\)) defaults to \(0.2\).
+
+An optional argument, \texttt{is\_bkg}, adds background noise drawn from a multivariate normal distribution centered on the dataset's overall mean with standard deviations matching the observed spread.
+Extra arguments (\texttt{...}) can be passed to cluster generators, allowing further control over per-cluster characteristics like radius of the sphere.
+The main arguments of the \texttt{gen\_multicluster()} function are shown in Table \ref{tab:main-tb-pdf}.
+
+\begin{table}
+
+\caption{\label{tab:main-tb-pdf}The main arguments for gen\_multicluster().}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Argument & Type & Explanation\\
+\midrule
+n & numeric (vector) & Number of points in each cluster.\\
+k & numeric & Number of clusters.\\
+loc & numeric (matrix) & Locations/centroids of clusters.\\
+scale & numeric (vector) & Scaling factors of clusters.\\
+shape & character (vector) & Shapes of clusters.\\
+rotation & numeric (list) & Rotation matrices, one per cluster.\\
+is\_bkg & boolean & Background noise should exist or not.\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Three 2-D projections of the 4-D orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `orglinearbranches` data.
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Three 2-D projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ `cone` data.
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Three 2-D projections of five 4-D synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.}]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Composite plot assessing six NLDR layouts of five clusters data using HBE over varying binwidths, showing tSNE, UMAP, PAHTE, TriMAP, PaCMAP, and PCA.
@@
 These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{tibble} \citep{kirill2023}, \texttt{gtools} \citep{gregory2023}, \texttt{dplyr} \citep{hadley2023}, \texttt{stats} \citep{core2025}, \texttt{tidyr} \citep{hadley2024}, \texttt{purrr} \citep{hadley2025}, \texttt{mvtnorm} \citep{alan2009}, \texttt{geozoo} \citep{barret2016}, and \texttt{MASS} \citep{venables2002}.
-
-\section{Supplementary numeric generators}\label{supplementary-numeric-generators}
-
-Two helper functions, \texttt{gen\_nproduct()} and \texttt{gen\_nsum()}, generate numeric vectors of positive integers that approximately satisfy a user-specified target product or sum, respectively.
-
-The function \texttt{gen\_nsum(n,\ k)} divides a total sum \texttt{n} into \texttt{k} positive integers.
-It first assigns an equal base value to each element and then randomly distributes any remainder, ensuring the elements sum exactly to \texttt{n}.
-
-\begin{verbatim}
-gen_nsum(n = 100, k = 3)
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 33 34 33
-\end{verbatim}
-
-The function \texttt{gen\_nproduct(n,\ p)} aims to produce \texttt{p} positive integers whose product is approximately \texttt{n}.
-It starts with all elements equal to the rounded \(p^{th}\) root of \texttt{n} and iteratively adjusts elements up or down in a randomized manner until the product is within a small tolerance of \texttt{n}.
-This accommodates the fact that exact integer solutions for a given product are often impossible.
-
-\begin{verbatim}
-gen_nproduct(n = 500, p = 4)
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 4 5 5 5
-\end{verbatim}
 

### 2025-11-11 paper-cardinalR.Rmd
Updated positions of the table

@@
 main_tb |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", table.pos = "H") 
+  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", table.pos = "!ht") 
 ```
@@
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H", linesep = "")  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "!ht", linesep = "")  |>
   column_spec(1, width = "2cm") |>
@@
 branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
+  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
 ```
@@
 branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), table.pos = "H") 
+  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), table.pos = "!ht") 
 ```
@@
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "1cm") |>
@@
 cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
+  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
 ```
@@
 cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
+  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
 ```
@@
 polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 pyramid_tb |> 
-  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
+  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
 ```
@@
 pyramid_tb |> 
-  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "H") 
+  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "!ht") 
 ```
@@
 sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "H")  |>
+  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 trefoil_tb |> 
-  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
+  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
 ```
@@
 trefoil_tb |> 
-  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "H") 
+  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "!ht") 
 ```
@@
 trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "H")  |>
+  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 noise_fun_tb |> 
-  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation"), table.pos = "H") 
+  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
 ```
@@
 noise_fun_tb |> 
-  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "2.5cm") |>
@@
 producing high-dimensional noise that preserves some geometric correlation with the base structure while introducing additional complexity.
+
+## Rotate shape generators
 
@@
 odd_shapes_tb |> 
-  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "H") 
+  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "!ht") 
 ```
@@
 odd_shapes_tb |> 
-  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "H")  |>
+  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "!ht")  |>
   column_spec(1, width = "3.5cm") |>
@@
                             "gen_nsum",
-                            "gen_rotation",
                             "normalize_data"), 
@@
                                 "Generates a vector of positive integers whose summation is approximately equal to a target value.", 
-                                "Generates rotations.", 
                                 "Normalizes data."))
@@
 add_fun_tb |> 
-  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "H") 
+  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "!ht") 
 ```
@@
 add_fun_tb |> 
-  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "H")  |>
+  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 summary_clust |> 
-  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("Metric", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("Metric", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "2.5cm") |>

### 2025-11-11 paper-cardinalR.tex
Updated positions of the table

@@
 producing high-dimensional noise that preserves some geometric correlation with the base structure while introducing additional complexity.
+
+\subsection{Rotate shape generators}\label{rotate-shape-generators}
 
@@
 gen\_nsum & Generates a vector of positive integers whose summation is approximately equal to a target value.\\
-gen\_rotation & Generates rotations.\\
 normalize\_data & Normalizes data.\\

### 2025-11-11 paper-cardinalR.R

@@ -135,12 +135,12 @@ main_tb <- tibble(arg = c("n",
 
 ## ----main-tb-html, eval=knitr::is_html_output()-------------------------------
 # main_tb |>
-#   kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", table.pos = "H")
+#   kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", table.pos = "!ht")
 
 
 ## ----main-tb-pdf, eval=knitr::is_latex_output()-------------------------------
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "H", linesep = "")  |>
+  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "!ht", linesep = "")  |>
   column_spec(1, width = "2cm") |>
   column_spec(2, width = "3cm") |>
   column_spec(3, width = "8cm")
@@ -161,12 +161,12 @@ branch_tb <- tibble(fun = c("gen_expbranches",
 
 ## ----branching-tb-html, eval=knitr::is_html_output()--------------------------
 # branch_tb |>
-#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), table.pos = "H")
+#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht")
 
 
 ## ----branching-tb-pdf, eval=knitr::is_latex_output()--------------------------
 branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "H")  |>
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
   column_spec(2, width = "8cm")
 
@@ -180,12 +180,12 @@ arg_branching_tb <- tibble(arg = c("n", "k"),
 
 ## ----arg-branching-tb-html, eval=knitr::is_html_output()----------------------
 # arg_branching_tb |>
[…]

### 2025-11-11 paper-cardinalR.Rmd
Added link for the webpage of additional functions
Added rotation sub section

@@
 
-## Rotate shape generators
+## Rotating shape generators
+
+In $p\text{-}D$ space, a rotation is an orthogonal transformation that changes the orientation of data while preserving its total variance and pairwise distances.
+The function `gen_rotation()` generates such rotation matrices for any dimension, given a list of rotation planes (axis pairs) and angles.
 
@@
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:add-tb-html)', '\\@ref(tab:add-tb-pdf)'))` details these functions.
+More detailed explanations are available in <https://jayanilakshika.github.io/cardinalR/articles/03additionalfun.html>.
 

### 2025-11-11 paper-cardinalR.tex
Added link for the webpage of additional functions
Added rotation sub section

@@
 
-\subsection{Rotate shape generators}\label{rotate-shape-generators}
+\subsection{Rotating shape generators}\label{rotating-shape-generators}
+
+In \(p\text{-}D\) space, a rotation is an orthogonal transformation that changes the orientation of data while preserving its total variance and pairwise distances.
+The function \texttt{gen\_rotation()} generates such rotation matrices for any dimension, given a list of rotation planes (axis pairs) and angles.
 
@@
 Table \ref{tab:add-tb-pdf} details these functions.
+More detailed explanations are available in \url{https://jayanilakshika.github.io/cardinalR/articles/03additionalfun.html}.
 

### 2025-11-11 paper-cardinalR.Rmd
Updated Figure 1 caption

@@
 
-```{r workflow, out.width = "70%", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data.
-The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
-Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.", fig.pos="!ht"}
+```{r workflow, out.width = "80%", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data.
+The user specifies input parameters such as the number of points ($n$), number of clusters ($k$), cluster locations, shapes, scaling, rotations, and optional background noise.
+Each cluster shape is generated by a shape generator, optionally rotated or scaled, and combined into a single dataset.
+Additional background noise can be added, and each observation is labeled by shape.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.", fig.pos="!ht"}
 

### 2025-11-11 paper-cardinalR.tex
Updated Figure 1 caption

@@
 
-{\centering \includegraphics[width=0.7\linewidth,alt={Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.}]{figures/cardinalR_workflow}
+{\centering \includegraphics[width=0.8\linewidth,alt={Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.}]{figures/cardinalR_workflow}
 
@@
 \caption{Workflow for generating high-dimensional clustered data.
-The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise).
-Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.}\label{fig:workflow}
+The user specifies input parameters such as the number of points ($n$), number of clusters ($k$), cluster locations, shapes, scaling, rotations, and optional background noise.
+Each cluster shape is generated by a shape generator, optionally rotated or scaled, and combined into a single dataset.
+Additional background noise can be added, and each observation is labeled by shape.}\label{fig:workflow}
 \end{figure}

### 2025-11-17 paper-cardinalR.Rmd
revised abstract

@@
 abstract: >
-  A high-dimensional dataset is one where each observation is described by many features, or dimensions, with associations between them.
-These datasets contain nonlinear manifolds in image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining.
-Data with a variety of structures can be generated using mathematical functions and statistical distributions to create test datasets.
-High-dimensional data structures are useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
-Their controlled complexity allows researchers to understand challenges posed in data analysis and helps to develop robust analytical methods across diverse scientific fields like bioinformatics, machine learning, and forensic science.
-Functions to generate a large variety of structures in high dimensions are organized into the R package `cardinalR`, along with some already generated examples, adding to the existing toolset of benchmark datasets for evaluating algorithms.
+  High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association.
+It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold.
+Simulated high-dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
+These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets.
+By controlling the complexity researchers can better understand how different methods work and can be improved.
+A variety of functions to generate a large variety of structures in high dimensions are organized into the R package `cardinalR`, along with several sample sets.
+This enriches the existing toolset of benchmark datasets for evaluating algorithms.
 draft: true

### 2025-11-17 paper-cardinalR.tex
revised abstract

@@
 \abstract{%
-A high-dimensional dataset is one where each observation is described by many features, or dimensions, with associations between them.
-These datasets contain nonlinear manifolds in image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining.
-Data with a variety of structures can be generated using mathematical functions and statistical distributions to create test datasets.
-High-dimensional data structures are useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
-Their controlled complexity allows researchers to understand challenges posed in data analysis and helps to develop robust analytical methods across diverse scientific fields like bioinformatics, machine learning, and forensic science.
-Functions to generate a large variety of structures in high dimensions are organized into the R package \texttt{cardinalR}, along with some already generated examples, adding to the existing toolset of benchmark datasets for evaluating algorithms.
+High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association.
+It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold.
+Simulated high-dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
+These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets.
+By controlling the complexity researchers can better understand how different methods work and can be improved.
+A variety of functions to generate a large variety of structures in high dimensions are organized into the R package \texttt{cardinalR}, along with several sample sets.
+This enriches the existing toolset of benchmark datasets for evaluating algorithms.
 }

### 2025-11-17 paper-cardinalR.Rmd
modification to intro

@@
 It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold.
-Simulated high-dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, clustering, machine learning, and visualization.
+Simulated high-dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
 These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets.
@@
 The `surreal` package (@james2024) implements the "Residual (Sur)Realism" algorithm (@leonard2007) to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
-Meanwhile, the `DHARMa` package (@florian2024) adopts a simulation-based approach to create scaled quantile residuals for generalized linear (mixed) models and related frameworks, supporting model diagnostics through intuitive residuals, plots, and tests for common misspecification issues.
-
-While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.
+<!--Meanwhile, the `DHARMa` package (@florian2024) adopts a simulation-based approach to create scaled quantile residuals for generalized linear (mixed) models and related frameworks, supporting model diagnostics through intuitive residuals, plots, and tests for common misspecification issues.-->
+
+These packages are especially valuable for specific applications, but together they are not sufficiently rish for simulating high-dimensional data for testing dimensionality reduction, supervised and unsupervised learning algorithms.
 To address this gap, this paper introduces the `cardinalR` R package.
-This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes.
-`cardinalR` offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
+This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
 By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimension reduction (NLDR) methods, clustering algorithms, and visualization techniques.

### 2025-11-17 paper-cardinalR.tex
modification to intro

@@
 The \texttt{surreal} package (\citet{james2024}) implements the ``Residual (Sur)Realism'' algorithm (\citet{leonard2007}) to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
-Meanwhile, the \texttt{DHARMa} package (\citet{florian2024}) adopts a simulation-based approach to create scaled quantile residuals for generalized linear (mixed) models and related frameworks, supporting model diagnostics through intuitive residuals, plots, and tests for common misspecification issues.
-
-While these packages are valuable, their scope is often limited to specific applications or low-dimensional structures.
+
+These packages are especially valuable for specific applications, but together they are not sufficiently rish for simulating high-dimensional data for testing dimensionality reduction, supervised and unsupervised learning algorithms.
 To address this gap, this paper introduces the \texttt{cardinalR} R package.
-This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes.
-\texttt{cardinalR} offers important functionalities that extend beyond the capabilities of existing tools, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
+This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
 By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimension reduction (NLDR) methods, clustering algorithms, and visualization techniques.

### 2025-11-17 paper-cardinalR.Rmd
introduction and conclusions modified

@@
 
-These packages are especially valuable for specific applications, but together they are not sufficiently rish for simulating high-dimensional data for testing dimensionality reduction, supervised and unsupervised learning algorithms.
+These packages are especially valuable for specific applications, but together they are not sufficiently rich for simulating high-dimensional data for testing dimensionality reduction, supervised and unsupervised learning algorithms.
 To address this gap, this paper introduces the `cardinalR` R package.
 This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimension reduction (NLDR) methods, clustering algorithms, and visualization techniques.
+By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of NLDR methods, clustering algorithms, and visualization techniques.
 Moreover, these datasets can serve as benchmark examples for exploring how different algorithmic choices affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
+
+The motivation for developing this package originated from our own work in studying nonlinear dimension reduction (NLDR) algorithms.
+We wanted to conduct a visualization experiment to understand perception and misperception of a variety of NLDR methods.
+This required simulated datasets with carefully controlled geometric and clustering properties.
+While some existing packages provided useful starting points, none fully supported the creation of flexible, high-dimensional data with the specific structural variations needed for our experiment.
+Developing these generators for research purposes underlies `cardinalR`, which is now a general-purpose package that should be useful for research and teaching.
 
@@
 
-The motivation for developing this package originated from the need to design a perception–misperception experiment, aimed at investigating how well NLDR methods preserve inter-cluster structure.
-To conduct this study, we required simulated datasets with carefully controlled geometric and clustering properties.
-While some existing packages provided useful starting points, none fully supported the creation of flexible, high-dimensional data with the specific structural variations needed for our experiment.
-Developing these generators for research purposes gradually led to the design of `cardinalR` as a general-purpose package, so that other researchers can benefit from the same tools for simulation, benchmarking, and teaching.
-
 The included structures cover a wide range of diagnostic settings.

### 2025-11-17 paper-cardinalR.tex
introduction and conclusions modified

@@
 
-These packages are especially valuable for specific applications, but together they are not sufficiently rish for simulating high-dimensional data for testing dimensionality reduction, supervised and unsupervised learning algorithms.
+These packages are especially valuable for specific applications, but together they are not sufficiently rich for simulating high-dimensional data for testing dimensionality reduction, supervised and unsupervised learning algorithms.
 To address this gap, this paper introduces the \texttt{cardinalR} R package.
 This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of nonlinear dimension reduction (NLDR) methods, clustering algorithms, and visualization techniques.
+By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of NLDR methods, clustering algorithms, and visualization techniques.
 Moreover, these datasets can serve as benchmark examples for exploring how different algorithmic choices affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
+
+The motivation for developing this package originated from our own work in studying nonlinear dimension reduction (NLDR) algorithms.
+We wanted to conduct a visualization experiment to understand perception and misperception of a variety of NLDR methods.
+This required simulated datasets with carefully controlled geometric and clustering properties.
+While some existing packages provided useful starting points, none fully supported the creation of flexible, high-dimensional data with the specific structural variations needed for our experiment.
+Developing these generators for research purposes underlies \texttt{cardinalR}, which is now a general-purpose package that should be useful for research and teaching.
 
@@
 
-The motivation for developing this package originated from the need to design a perception--misperception experiment, aimed at investigating how well NLDR methods preserve inter-cluster structure.
-To conduct this study, we required simulated datasets with carefully controlled geometric and clustering properties.
-While some existing packages provided useful starting points, none fully supported the creation of flexible, high-dimensional data with the specific structural variations needed for our experiment.
-Developing these generators for research purposes gradually led to the design of \texttt{cardinalR} as a general-purpose package, so that other researchers can benefit from the same tools for simulation, benchmarking, and teaching.
-
 The included structures cover a wide range of diagnostic settings.

### 2025-11-19 paper-cardinalR.Rmd
Added trefoil references

@@
 Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
-There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The `geozoo` package (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-The package `snedata` (@james2025) provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding (@hinton2002) and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], to explore how well these techniques preserve data structures in lower dimensions.
-Additionally, `splatter` (@luke2017) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
-In contrast, `mlbench` (@friedrich2024) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
-The `surreal` package (@james2024) implements the "Residual (Sur)Realism" algorithm (@leonard2007) to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
+There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The `geozoo` package [@barret2016] offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
+The package `snedata` [@james2025] provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding [@hinton2002] and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], to explore how well these techniques preserve data structures in lower dimensions.
+Additionally, `splatter` [@luke2017] is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+In contrast, `mlbench` [@friedrich2024] includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+The `surreal` package [@james2024] implements the "Residual (Sur)Realism" algorithm [@leonard2007] to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
 <!--Meanwhile, the `DHARMa` package (@florian2024) adopts a simulation-based approach to create scaled quantile residuals for generalized linear (mixed) models and related frameworks, supporting model diagnostics through intuitive residuals, plots, and tests for common misspecification issues.-->
@@
 
-A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (@trapnell2014).
+A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology [@trapnell2014].
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (`n`), and number of branches (`k`).
@@
 
-For the $4\text{-}D$ trefoil knot, the function `gen_trefoil4d(n, steps)` generates the structure on the $3$-sphere ($S^3 \subset \mathbb{R}^4$) using two angular parameters, $\theta$ and $\phi$.
+For the $4\text{-}D$ trefoil knot [@laurent2024], the function `gen_trefoil4d(n, steps)` generates the structure on the $3$-sphere ($S^3 \subset \mathbb{R}^4$) using two angular parameters, $\theta$ and $\phi$.
 A band of thickness around the knot path is controlled by the `steps` argument, while the number of $\theta$ and $\phi$ values is determined by the `steps` and `n` arguments, respectively (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
@@
 
-For the $3\text{-}D$ stereographic projection, `gen_trefoil3d(n, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),$ excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
+For the $3\text{-}D$ stereographic projection [@laurent2024], `gen_trefoil3d(n, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),$ excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
 

### 2025-11-19 paper-cardinalR.tex
Added trefoil references

@@
 Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
-There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package (\citet{barret2016}) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-The package \texttt{snedata} (\citet{james2025}) provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding (\citet{hinton2002}) and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, to explore how well these techniques preserve data structures in lower dimensions.
-Additionally, \texttt{splatter} (\citet{luke2017}) is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
-In contrast, \texttt{mlbench} (\citet{friedrich2024}) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
-The \texttt{surreal} package (\citet{james2024}) implements the ``Residual (Sur)Realism'' algorithm (\citet{leonard2007}) to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
+There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package \citep{barret2016} offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
+The package \texttt{snedata} \citep{james2025} provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding \citep{hinton2002} and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, to explore how well these techniques preserve data structures in lower dimensions.
+Additionally, \texttt{splatter} \citep{luke2017} is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+In contrast, \texttt{mlbench} \citep{friedrich2024} includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+The \texttt{surreal} package \citep{james2024} implements the ``Residual (Sur)Realism'' algorithm \citep{leonard2007} to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
 
@@
 
-A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology (\citet{trapnell2014}).
+A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology \citep{trapnell2014}.
 We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (\texttt{n}), and number of branches (\texttt{k}).
@@
 
-For the \(4\text{-}D\) trefoil knot, the function \texttt{gen\_trefoil4d(n,\ steps)} generates the structure on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)) using two angular parameters, \(\theta\) and \(\phi\).
+For the \(4\text{-}D\) trefoil knot \citep{laurent2024}, the function \texttt{gen\_trefoil4d(n,\ steps)} generates the structure on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)) using two angular parameters, \(\theta\) and \(\phi\).
 A band of thickness around the knot path is controlled by the \texttt{steps} argument, while the number of \(\theta\) and \(\phi\) values is determined by the \texttt{steps} and \texttt{n} arguments, respectively (Figure \ref{fig:trefoil-proj} a).
@@
 
-For the \(3\text{-}D\) stereographic projection, \texttt{gen\_trefoil3d(n,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),\) excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
+For the \(3\text{-}D\) stereographic projection \citep{laurent2024}, \texttt{gen\_trefoil3d(n,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),\) excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
 

### 2025-11-19 paper-cardinalR.Rmd
Updated clustering algorithm summary table

@@
 summary_clust |> 
-  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", col.names = c("Metric", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), tab.pos = "H")
+  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", col.names = c("method", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), tab.pos = "H")
 ```
@@
 summary_clust |> 
-  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("Metric", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), booktabs = T, table.pos = "!ht")  |>
+  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("method", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "2.5cm") |>

### 2025-11-19 paper-cardinalR.Rmd
Updated application intro

@@
 
-This section demonstrates how the package can be used to generate complex high-dimensional datasets, apply dimension reduction (DR) techniques, and evaluate clustering performance.
+This section demonstrates how the package can be used to generate complex high-dimensional datasets, evaluate dimension reduction (DR) and clustering methods.
 The example shows how diverse geometric structures can be simulated and analyzed to assess algorithmic behavior.

### 2025-11-20 paper-cardinalR.Rmd
Updated Paul's comment on grided sphere

@@
 Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+Throughout this section, we follow the standard mathematical terminology: a *sphere* refers to the hollow $(p-1)$-dimensional surface in $\mathbb{R}^p$, while a *ball* refers to the filled interior region.
 The functions generate a variety of spherical forms, including simple circles, uniform and hollow spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
@@
                                 "Curvy cell cycle.",
-                                "Uniform sphere.",
+                                "Uniform ball.",
                                 "Hollow sphere.",
@@
 
-In addition, the `gen_gridedsphere(n)` function constructs a $p\text{-}D$ dataset consisting of approximately $n$ points that are evenly distributed on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
-The method relies on forming a regular grid in spherical coordinates, parameterized by $(p-1)$ angular variables: for dimensions $j = 1, \dots, p-2$ the polar angles are drawn from $[0, \pi]$, while the final angle ($j = p-1$) represents the azimuth and is drawn from $[0, 2\pi]$.
-The number of grid steps along each angular dimension is chosen by decomposing $n$ into $(p-1)$ approximately equal integer factors using the helper function `gen_nproduct(n, p - 1)`.
+In addition, the `gen_gridedsphere(n)` function constructs a $p$-dimensional dataset consisting of approximately $n$ points arranged on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
+Rather than sampling points uniformly, this function creates a deterministic grid in spherical coordinates, using $(p-1)$ angular variables: the first $(p-2)$ angles are taken from $[0, \pi]$, and the final angle from $[0, 2\pi]$.
+The number of grid points along each angular dimension is determined by decomposing $n$ into $(p-1)$ approximately equal integer factors via `gen_nproduct(n, p - 1)`.
 
@@
 For more heterogeneous structures, the `gen_clusteredspheres(n, k, r, loc)` function generates one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
-A large uniform sphere centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
+A large Uniform ball centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
 The sampling is executed using the function `gen_unifsphere(n_1, r_1)`, which generates the desired points in the specified dimensional space.
-In generation of $k$ smaller uniform spheres, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
+In generation of $k$ smaller Uniform balls, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
 These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.

### 2025-11-20 paper-cardinalR.tex
Updated clustering algorithm summary table
Updated Paul's comment on grided sphere

@@
 Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
+Throughout this section, we follow the standard mathematical terminology: a \emph{sphere} refers to the hollow \((p-1)\)-dimensional surface in \(\mathbb{R}^p\), while a \emph{ball} refers to the filled interior region.
 The functions generate a variety of spherical forms, including simple circles, uniform and hollow spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
@@
 gen\_curvycycle & Curvy cell cycle.\\
-gen\_unifsphere & Uniform sphere.\\
+gen\_unifsphere & Uniform ball.\\
 gen\_hollowsphere & Hollow sphere.\\
@@
 
-In addition, the \texttt{gen\_gridedsphere(n)} function constructs a \(p\text{-}D\) dataset consisting of approximately \(n\) points that are evenly distributed on the surface of the unit \((p-1)\)-sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
-The method relies on forming a regular grid in spherical coordinates, parameterized by \((p-1)\) angular variables: for dimensions \(j = 1, \dots, p-2\) the polar angles are drawn from \([0, \pi]\), while the final angle (\(j = p-1\)) represents the azimuth and is drawn from \([0, 2\pi]\).
-The number of grid steps along each angular dimension is chosen by decomposing \(n\) into \((p-1)\) approximately equal integer factors using the helper function \texttt{gen\_nproduct(n,\ p\ -\ 1)}.
+In addition, the \texttt{gen\_gridedsphere(n)} function constructs a \(p\)-dimensional dataset consisting of approximately \(n\) points arranged on the surface of the unit \((p-1)\)-sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
+Rather than sampling points uniformly, this function creates a deterministic grid in spherical coordinates, using \((p-1)\) angular variables: the first \((p-2)\) angles are taken from \([0, \pi]\), and the final angle from \([0, 2\pi]\).
+The number of grid points along each angular dimension is determined by decomposing \(n\) into \((p-1)\) approximately equal integer factors via \texttt{gen\_nproduct(n,\ p\ -\ 1)}.
 
@@
 For more heterogeneous structures, the \texttt{gen\_clusteredspheres(n,\ k,\ r,\ loc)} function generates one large sphere of radius \(r_1\) and \(k\) smaller spheres of radius \(r_2\), each centered at a different random location (Figure \ref{fig:sphere-proj} e).
-A large uniform sphere centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\text{-}D\) sphere with a radius of \(r_1\).
+A large Uniform ball centered at the origin is created by sampling \(n_1\) points uniformly on the surface of a \(p\text{-}D\) sphere with a radius of \(r_1\).
 The sampling is executed using the function \texttt{gen\_unifsphere(n\_1,\ r\_1)}, which generates the desired points in the specified dimensional space.
-In generation of \(k\) smaller uniform spheres, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
+In generation of \(k\) smaller Uniform balls, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
 These spheres are positioned at distinct random locations in \(p\)-space, with the center of each sphere being drawn from a normal distribution \(N(0, \texttt{loc}^2 I_p)\).
@@
 \toprule
-Metric & wb.ratio & Dunn Index & Corrected Rand & VI\\
+method & wb.ratio & Dunn Index & Corrected Rand & VI\\
 \midrule

### 2025-11-20 paper-cardinalR.tex
Linked shape generators

@@
 
-The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes, which can then be combined into a single dataset including scaling, rotation and translation.
+The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes (\url{https://jayanilakshika.github.io/cardinalR/reference/index.html}), which can then be combined into a single dataset including scaling, rotation, and translation.
 The package is available on CRAN, and the source is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.
 
-The main function, \texttt{gen\_multicluster()}, is an all-in-one function that includes generating individual shapes, handles scaling and rotating of these shapes, and combines the result into a single unified dataset.
-This function and associated workflow allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
+The main function, \texttt{gen\_multicluster()}, is an all-in-one function that includes generating individual shapes, handling scaling and rotating of these shapes, and combining the result into a single unified dataset.
+This function and associated workflow allow flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
 Figure \ref{fig:workflow} illustrates the workflow of \texttt{gen\_multicluster()}.
@@
 
-The following is an example of a three shape multiclustered dataset. The first shape is gaussian, the second conical, and the third a cube.
+The following is an example of a three-shape multiclustered dataset. The first shape is Gaussian, the second conical, and the third a cube.
 
@@
 
-Here, the shapes have 200, 300 and 500 points respectively (\texttt{n}), are positioned in 4-D space according to a location matrix, \texttt{loc}, and stretched according to the \texttt{scale}.
-The details of the individual shape generators, and the noise elements are contained in the following sections.
+Here, the shapes have \(200\), \(300\), and \(500\) points respectively (\texttt{n}), are positioned in \(4\text{-}D\) space according to a location matrix, \texttt{loc}, and stretched according to the \texttt{scale}.
+The details of the individual shape generators and the noise elements are contained in the following sections.
 

### 2025-11-20 paper-cardinalR.Rmd
Linked shape generators documentation

@@
 
-The `cardinalR` package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes, which can then be combined into a single dataset including scaling, rotation and translation.
+The `cardinalR` package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes (A full list of available shape generators are available at <https://jayanilakshika.github.io/cardinalR/reference/index.html>.), which can then be combined into a single dataset including scaling, rotation, and translation.
 The package is available on CRAN, and the source is available on GitHub at [JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR).
 
-The main function, `gen_multicluster()`, is an all-in-one function that includes generating individual shapes, handles scaling and rotating of these shapes, and combines the result into a single unified dataset.
-This function and associated workflow allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
+The main function, `gen_multicluster()`, is an all-in-one function that includes generating individual shapes, handling scaling and rotating of these shapes, and combining the result into a single unified dataset.
+This function and associated workflow allow flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
 Figure `r '\\@ref(fig:workflow)'` illustrates the workflow of `gen_multicluster()`.
@@
 
-Here, the shapes have 200, 300 and 500 points respectively (`n`), are positioned in 4-D space according to a location matrix, `loc`, and stretched according to the `scale`.
-The details of the individual shape generators, and the noise elements are contained in the following sections.
+Here, the shapes have $200$, $300$, and $500$ points respectively (`n`), are positioned in $4\text{-}D$ space according to a location matrix, `loc`, and stretched according to the `scale`.
+The details of the individual shape generators and the noise elements are contained in the following sections.
 

### 2025-11-20 paper-cardinalR.Rmd
Rewrote longlinear paragraph

@@
 
+<!-- The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise.
+Each variable is formed as $X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i$, where $\text{scale}\_i \sim U(-10, 10)$ determines the orientation of the line in each dimension, $\text{shift}\_i \sim U(-300, 300)$ offsets the line to separate dimensions, and $\epsilon \sim N(0, (0.03n)^2)$ introduces Gaussian noise.
+-->
+
 The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise.
-Each variable is formed as $X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i$, where $\text{scale}\_i \sim U(-10, 10)$ determines the orientation of the line in each dimension, $\text{shift}\_i \sim U(-300, 300)$ offsets the line to separate dimensions, and $\epsilon \sim N(0, (0.03n)^2)$ introduces Gaussian noise.
+Each variable is constructed by first creating a sequence of integers from $0$ to $n-1$, adding Gaussian noise with standard deviation $0.03n$, and then applying a random scaling and shift.
+Specifically, for the $j$-th dimension, a scale factor $\text{scale}_j \sim U(-10, 10)$ determines the orientation and stretching of the line, while a shift factor $\text{shift}_j \sim U(-300, 300)$ is added inside the multiplication to offset the line.
+This results in each variable being of the form $X_j = \text{scale}_j \cdot \big((0,1,\dots,n-1) + \epsilon + \text{shift}_j\big)$, where $\epsilon \sim N(0, (0.03n)^2)$ is independent Gaussian noise.
 

### 2025-11-20 paper-cardinalR.tex
Rewrote longlinear paragraph

@@
 
-The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes (\url{https://jayanilakshika.github.io/cardinalR/reference/index.html}), which can then be combined into a single dataset including scaling, rotation, and translation.
+The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes (A full list of available shape generators are available at \url{https://jayanilakshika.github.io/cardinalR/reference/index.html}.), which can then be combined into a single dataset including scaling, rotation, and translation.
 The package is available on CRAN, and the source is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.
@@
 The \texttt{gen\_longlinear(n,\ p)} function generates a high-dimensional dataset representing a long linear structure with noise.
-Each variable is formed as \(X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i\), where \(\text{scale}\_i \sim U(-10, 10)\) determines the orientation of the line in each dimension, \(\text{shift}\_i \sim U(-300, 300)\) offsets the line to separate dimensions, and \(\epsilon \sim N(0, (0.03n)^2)\) introduces Gaussian noise.
+Each variable is constructed by first creating a sequence of integers from \(0\) to \(n-1\), adding Gaussian noise with standard deviation \(0.03n\), and then applying a random scaling and shift.
+Specifically, for the \(j\)-th dimension, a scale factor \(\text{scale}_j \sim U(-10, 10)\) determines the orientation and stretching of the line, while a shift factor \(\text{shift}_j \sim U(-300, 300)\) is added inside the multiplication to offset the line.
+This results in each variable being of the form \(X_j = \text{scale}_j \cdot \big((0,1,\dots,n-1) + \epsilon + \text{shift}_j\big)\), where \(\epsilon \sim N(0, (0.03n)^2)\) is independent Gaussian noise.
 

### 2025-11-20 paper-cardinalR.Rmd
Try to fix rjtool issue with abstract

@@
 ---
-title: "cardinalR: Generating interesting high-dimensional data structures"
-abstract: >
+title: "cardinalR: Generating Interesting High-Dimensional Data Structures"
+abstract: |
   High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association.
@@
 By controlling the complexity researchers can better understand how different methods work and can be improved.
-A variety of functions to generate a large variety of structures in high dimensions are organized into the R package `cardinalR`, along with several sample sets.
+A variety of functions to generate a large variety of structures in high dimensions are organized into the R package cardinalR, along with several sample sets.
 This enriches the existing toolset of benchmark datasets for evaluating algorithms.
@@
     orcid: 0000-0002-6265-6481
-    email:  \email{jayani.piyadigamage@monash.edu}
+    email: jayani.piyadigamage@monash.edu
   - name: Dianne Cook

### 2025-11-20 paper-cardinalR.tex
Try to fix rjtool issue with abstract

@@
 % !TeX root = RJwrapper.tex
-\title{cardinalR: Generating interesting high-dimensional data structures}
+\title{cardinalR: Generating Interesting High-Dimensional Data Structures}
 
@@
 By controlling the complexity researchers can better understand how different methods work and can be improved.
-A variety of functions to generate a large variety of structures in high dimensions are organized into the R package \texttt{cardinalR}, along with several sample sets.
+A variety of functions to generate a large variety of structures in high dimensions are organized into the R package cardinalR, along with several sample sets.
 This enriches the existing toolset of benchmark datasets for evaluating algorithms.
@@
 \textit{ORCiD: \href{https://orcid.org/0000-0002-6265-6481}{0000-0002-6265-6481}}\\%
-\email{jayani.piyadigamage@monash.edu}%
+\href{mailto:jayani.piyadigamage@monash.edu}{\nolinkurl{jayani.piyadigamage@monash.edu}}%
 }

### 2025-11-25 paper-cardinalR.Rmd
Try to find why rjtolls have the error

@@
 title: "cardinalR: Generating Interesting High-Dimensional Data Structures"
-abstract: |
+abstract: >
   High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association.
 It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold.
-Simulated high-dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
+Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
 These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets.

### 2025-11-25 paper-cardinalR.tex
Try to find why rjtolls have the error

@@
 It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold.
-Simulated high-dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
+Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
 These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets.

### 2025-11-26 paper-cardinalR.Rmd
Added installation from package file

@@
 
-options(repos = c(CRAN = "https://cran.rstudio.com")) # Setup mirror
-
-packages_to_check <- c("remotes", "cardinalR", "tidyverse", "kableExtra", "geozoo", "patchwork", "colorspace")
-
-for (pkg in packages_to_check) {
-  if (!requireNamespace(pkg, quietly = TRUE)) {
-    message(paste("Installing package:", pkg))
-    install.packages(pkg)
-  } else {
-    installed_version <- packageVersion(pkg)
-    available_version <- tryCatch({
-      utils::packageDescription(pkg)$Version
-    }, error = function(e) NA) # Handle cases where package info isn't readily available
-
-    if (!is.na(available_version) && installed_version < package_version(available_version)) {
-      message(paste("A newer version of package", pkg, "is available. Updating..."))
-      install.packages(pkg)
-    } else {
-      message(paste("Package", pkg, "is up-to-date (version", installed_version, ")."))
+# Ensure remotes is available
+if (!requireNamespace("remotes", quietly = TRUE)) {
+  install.packages("remotes")
 }
+
+# Read the package list
+pkgs_raw <- readLines("_Rpackages.txt")
+pkgs_raw <- pkgs_raw[nzchar(pkgs_raw)]  # remove empty lines
+
+# Split into package and version
+parts <- strsplit(pkgs_raw, "==")
+pkgs <- vapply(parts, `[[`, "", 1)
+versions <- vapply(parts, `[[`, "", 2)
+
+# Install packages with exact versions
+for (i in seq_along(pkgs)) {
+  pkg <- pkgs[i]
+  ver <- versions[i]
+
+  if (!requireNamespace(pkg, quietly = TRUE) ||
+      as.character(packageVersion(pkg)) != ver) {
+
+    message(sprintf("Installing %s (%s)", pkg, ver))
+    remotes::install_version(pkg, version = ver, upgrade = "never")
   }

### 2025-11-28 paper-cardinalR.Rmd
Added code for table functions

@@
                                "Linear shaped branches originated in one point.",
-                               "Curvy shaped branches originated in one point."))
+                               "Curvy shaped branches originated in one point.")) |> 
+   mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
 ```
@@
 branch_tb |>
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), table.pos = "!ht") 
+  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), escape = FALSE, table.pos = "!ht") 
 ```
@@
 arg_branching_tb |> 
-  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "!ht")  |>
+  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, escape = FALSE, table.pos = "!ht")  |>
   column_spec(1, width = "1cm") |>
@@
                       exp = c("Cube with specified grid points along each axes.",
-                              "Cube with uniform points."))
+                              "Cube with uniform points.")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |>
+  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
 ```
@@
 cube_tb |> 
-  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |>
+  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
                         exp = c("Quadratic pattern.",
-                                "Cubic pattern."))
+                                "Cubic pattern.")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
 ```
@@
 polynomial_tb |> 
-  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
           "Star-shaped base, with a sharp or blunted apex.",
-          "Pyramid with triangular pyramid-shaped holes."))
+          "Pyramid with triangular pyramid-shaped holes.")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 pyramid_tb |> 
-  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
 ```
@@
 pyramid_tb |> 
-  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, escape = FALSE, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
                                 "Multiple small spheres within a big sphere.",
-                                "Hemisphere."))
+                                "Hemisphere.")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
 ```
@@
 sphere_tb |> 
-  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, linesep = "", table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
                         exp = c("Trefoil in 4-D.",
-                                "Trefoil in 3-D."))
+                                "Trefoil in 3-D.")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 trefoil_tb |> 
-  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
 ```
@@
 trefoil_tb |>
-  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
                                 "Conic spiral.",
-                                "Nonlinear hyperbola."))
+                                "Nonlinear hyperbola.")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
 ```
@@
 trigonometric_tb |> 
-  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, linesep = "", table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
@@
                                 "Wavy noise dimensions using polynomial transformations of an existing dimension vector.",
-                                "Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset."))
+                                "Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 noise_fun_tb |> 
-  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
 ```
@@
 noise_fun_tb |> 
-  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "2.5cm") |>
@@
     "Parallel shaped clusters.",
-    "Three clusters with different shapes. (eg:- 01, 02, ..., 20)"))
+    "Three clusters with different shapes. (eg:- 01, 02, ..., 20)")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 odd_shapes_tb |> 
-  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
 ```
@@
 odd_shapes_tb |> 
-  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, linesep = "", table.pos = "!ht")  |>
   column_spec(1, width = "3.5cm") |>
@@
                                 "Generates a vector of positive integers whose summation is approximately equal to a target value.", 
-                                "Normalizes data."))
+                                "Normalizes data.")) |> 
+  mutate(fun = gsub("_", "\\\\_", fun))
 ```
@@
 add_fun_tb |> 
-  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation"), linesep = "", table.pos = "!ht") 
+  mutate(fun = paste0("<code>", fun, "</code>")) |> 
+  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
 ```
@@
 add_fun_tb |> 
-  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", table.pos = "!ht")  |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", escape = FALSE, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>

### 2025-11-28 paper-cardinalR.tex
Added code for table functions

@@
 \midrule
-gen\_expbranches & Exponential shaped branches.\\
-gen\_linearbranches & Linear shaped branches.\\
-gen\_curvybranches & Curvy shaped branches.\\
-gen\_orglinearbranches & Linear shaped branches originated in one point.\\
-gen\_orgcurvybranches & Curvy shaped branches originated in one point.\\
+\texttt{gen\_expbranches} & Exponential shaped branches.\\
+\texttt{gen\_linearbranches} & Linear shaped branches.\\
+\texttt{gen\_curvybranches} & Curvy shaped branches.\\
+\texttt{gen\_orglinearbranches} & Linear shaped branches originated in one point.\\
+\texttt{gen\_orgcurvybranches} & Curvy shaped branches originated in one point.\\
 \bottomrule
@@
 \midrule
-gen\_gridcube & Cube with specified grid points along each axes.\\
-gen\_unifcube & Cube with uniform points.\\
+\texttt{gen\_gridcube} & Cube with specified grid points along each axes.\\
+\texttt{gen\_unifcube} & Cube with uniform points.\\
 \bottomrule
@@
 \midrule
-gen\_quadratic & Quadratic pattern.\\
-gen\_cubic & Cubic pattern.\\
+\texttt{gen\_quadratic} & Quadratic pattern.\\
+\texttt{gen\_cubic} & Cubic pattern.\\
 \bottomrule
@@
 \midrule
-gen\_pyrrect & Rectangular-base, with a sharp or blunted apex.\\
-gen\_pyrtri & Triangular-base, with a sharp or blunted apex.\\
-gen\_pyrstar & Star-shaped base, with a sharp or blunted apex.\\
-gen\_pyrfrac & Pyramid with triangular pyramid-shaped holes.\\
+\texttt{gen\_pyrrect} & Rectangular-base, with a sharp or blunted apex.\\
+\texttt{gen\_pyrtri} & Triangular-base, with a sharp or blunted apex.\\
+\texttt{gen\_pyrstar} & Star-shaped base, with a sharp or blunted apex.\\
+\texttt{gen\_pyrfrac} & Pyramid with triangular pyramid-shaped holes.\\
 \bottomrule
@@
 \midrule
-gen\_circle & Circle.\\
-gen\_curvycycle & Curvy cell cycle.\\
-gen\_unifsphere & Uniform ball.\\
-gen\_hollowsphere & Hollow sphere.\\
-gen\_gridedsphere & Grided sphere.\\
-gen\_clusteredspheres & Multiple small spheres within a big sphere.\\
-gen\_hemisphere & Hemisphere.\\
+\texttt{gen\_circle} & Circle.\\
+\texttt{gen\_curvycycle} & Curvy cell cycle.\\
+\texttt{gen\_unifsphere} & Uniform ball.\\
+\texttt{gen\_hollowsphere} & Hollow sphere.\\
+\texttt{gen\_gridedsphere} & Grided sphere.\\
+\texttt{gen\_clusteredspheres} & Multiple small spheres within a big sphere.\\
+\texttt{gen\_hemisphere} & Hemisphere.\\
 \bottomrule
@@
 \midrule
-gen\_trefoil4d & Trefoil in 4-D.\\
-gen\_trefoil3d & Trefoil in 3-D.\\
+\texttt{gen\_trefoil4d} & Trefoil in 4-D.\\
+\texttt{gen\_trefoil3d} & Trefoil in 3-D.\\
 \bottomrule
@@
 \midrule
-gen\_crescent & Crescent pattern.\\
-gen\_curvycylinder & Curvy cylinder.\\
-gen\_sphericalspiral & Spherical spiral.\\
-gen\_helicalspiral & Helical spiral.\\
-gen\_conicspiral & Conic spiral.\\
-gen\_nonlinear & Nonlinear hyperbola.\\
+\texttt{gen\_crescent} & Crescent pattern.\\
+\texttt{gen\_curvycylinder} & Curvy cylinder.\\
+\texttt{gen\_sphericalspiral} & Spherical spiral.\\
+\texttt{gen\_helicalspiral} & Helical spiral.\\
+\texttt{gen\_conicspiral} & Conic spiral.\\
+\texttt{gen\_nonlinear} & Nonlinear hyperbola.\\
 \bottomrule
@@
 \midrule
-gen\_noisedims & Gaussian noise dimensions with optional mean and standard deviation.\\
-gen\_wavydims1 & Wavy noise dimensions based on a user-specified theta sequence with added jitter.\\
-gen\_wavydims2 & Wavy noise dimensions using polynomial transformations of an existing dimension vector.\\
-gen\_wavydims3 & Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.\\
+\texttt{gen\_noisedims} & Gaussian noise dimensions with optional mean and standard deviation.\\
+\texttt{gen\_wavydims1} & Wavy noise dimensions based on a user-specified theta sequence with added jitter.\\
+\texttt{gen\_wavydims2} & Wavy noise dimensions using polynomial transformations of an existing dimension vector.\\
+\texttt{gen\_wavydims3} & Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.\\
 \bottomrule
@@
 \midrule
-make\_mobiusgau & Möbius-like cluster combined with a Gaussian.\\
-make\_multigau & Multiple Gaussian clusters in high-dimensional space.\\
-make\_curvygau & Curvilinear cluster with a Gaussian cluster.\\
-make\_klink\_circles & K-link circular clusters (non-linear circular patterns).\\
-make\_chain\_circles & Chain-like circular clusters connected sequentially.\\
-make\_klink\_curvycycle & K-link curvy cycle clusters (curvilinear loop structures).\\
-make\_chain\_curvycycle & Chain-like curvy cycle clusters connected sequentially.\\
-make\_gaucircles & Circular clusters with a Gaussian cluster in the middle.\\
-make\_gaucurvycycle & Curvy circular clusters with a Gaussian in the middle.\\
-make\_onegrid & Single grid in two dimensions.\\
-make\_twogrid\_overlap & Two overlapping grids.\\
-make\_twogrid\_shift & Two grids shifted relative to each other.\\
-make\_shape\_para & Parallel shaped clusters.\\
-make\_three\_clust\_ & Three clusters with different shapes. (eg:- 01, 02, ..., 20)\\
+\texttt{make\_mobiusgau} & Möbius-like cluster combined with a Gaussian.\\
+\texttt{make\_multigau} & Multiple Gaussian clusters in high-dimensional space.\\
+\texttt{make\_curvygau} & Curvilinear cluster with a Gaussian cluster.\\
+\texttt{make\_klink\_circles} & K-link circular clusters (non-linear circular patterns).\\
+\texttt{make\_chain\_circles} & Chain-like circular clusters connected sequentially.\\
+\texttt{make\_klink\_curvycycle} & K-link curvy cycle clusters (curvilinear loop structures).\\
+\texttt{make\_chain\_curvycycle} & Chain-like curvy cycle clusters connected sequentially.\\
+\texttt{make\_gaucircles} & Circular clusters with a Gaussian cluster in the middle.\\
+\texttt{make\_gaucurvycycle} & Curvy circular clusters with a Gaussian in the middle.\\
+\texttt{make\_onegrid} & Single grid in two dimensions.\\
+\texttt{make\_twogrid\_overlap} & Two overlapping grids.\\
+\texttt{make\_twogrid\_shift} & Two grids shifted relative to each other.\\
+\texttt{make\_shape\_para} & Parallel shaped clusters.\\
+\texttt{make\_three\_clust\_} & Three clusters with different shapes. (eg:- 01, 02, ..., 20)\\
 \bottomrule
@@
 \midrule
-gen\_bkgnoise & Adds background noise.\\
-randomize\_rows & Randomizes the rows of input data.\\
-relocate\_clusters & Relocates the clusters.\\
-gen\_nproduct & Generates a vector of positive integers whose product is approximately equal to a target value.\\
-gen\_nsum & Generates a vector of positive integers whose summation is approximately equal to a target value.\\
-normalize\_data & Normalizes data.\\
+\texttt{gen\_bkgnoise} & Adds background noise.\\
+\texttt{randomize\_rows} & Randomizes the rows of input data.\\
+\texttt{relocate\_clusters} & Relocates the clusters.\\
+\texttt{gen\_nproduct} & Generates a vector of positive integers whose product is approximately equal to a target value.\\
+\texttt{gen\_nsum} & Generates a vector of positive integers whose summation is approximately equal to a target value.\\
+\texttt{normalize\_data} & Normalizes data.\\
 \bottomrule

### 2025-11-28 paper-cardinalR.R

@@ -160,17 +160,20 @@ branch_tb <- tibble(fun = c("gen_expbranches",
                                "Linear shaped branches.", 
                                "Curvy shaped branches.",
                                "Linear shaped branches originated in one point.",
-                               "Curvy shaped branches originated in one point."))
+                               "Curvy shaped branches originated in one point.")) |> 
+   mutate(fun = gsub("_", "\\\\_", fun))
 
 
 ## ----branching-tb-html, eval=knitr::is_html_output()--------------------------
 # branch_tb |>
-#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), table.pos = "!ht")
+#   mutate(fun = paste0("<code>", fun, "</code>")) |>
+#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht")
 
 
 ## ----branching-tb-pdf, eval=knitr::is_latex_output()--------------------------
-branch_tb |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, table.pos = "!ht")  |>
+branch_tb |>
+  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
   column_spec(1, width = "4cm") |>
   column_spec(2, width = "8cm")
 
@@ -184,12 +187,12 @@ arg_branching_tb <- tibble(arg = c("n", "k"),
 
 ## ----arg-branching-tb-html, eval=knitr::is_html_output()----------------------
 # arg_branching_tb |>
-#   kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), table.pos = "!ht")
+#   kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), escape = FALSE, table.pos = "!ht")
 
 
 ## ----arg-branching-tb-pdf, eval=knitr::is_latex_output()----------------------
 arg_branching_tb |> 
[…]

### 2025-11-28 paper-cardinalR.Rmd
Updated trefoil table

@@
                             "gen_trefoil3d"), 
-                        exp = c("Trefoil in 4-D.",
-                                "Trefoil in 3-D.")) |> 
+                        exp = c("Trefoil in $4\\text{-}D$.",
+                                "Trefoil in $3\\text{-}D$.")) |> 
   mutate(fun = gsub("_", "\\\\_", fun))
@@
                                 "make_twogrid_shift",
-                                "make_shape_para",
-                                "make_three_clust_"), 
+                                "make_shape_para"), 
                         exp = c(
@@
     "Two grids shifted relative to each other.",
-    "Parallel shaped clusters.",
-    "Three clusters with different shapes. (eg:- 01, 02, ..., 20)")) |> 
+    "Parallel shaped clusters.")) |> 
   mutate(fun = gsub("_", "\\\\_", fun))

### 2025-11-28 paper-cardinalR.R

@@ -1284,8 +1284,8 @@ swissroll <- gen_swissroll(n = 1000, w = c(-1, 1))
 ## -----------------------------------------------------------------------------
 trefoil_tb <- tibble(fun = c("gen_trefoil4d",
                             "gen_trefoil3d"), 
-                        exp = c("Trefoil in 4-D.",
-                                "Trefoil in 3-D.")) |> 
+                        exp = c("Trefoil in $4\\text{-}D$.",
+                                "Trefoil in $3\\text{-}D$.")) |> 
   mutate(fun = gsub("_", "\\\\_", fun))
 
 
@@ -1977,8 +1977,7 @@ odd_shapes_tb <- tibble(fun = c("make_mobiusgau",
                                 "make_onegrid",
                                 "make_twogrid_overlap",
                                 "make_twogrid_shift",
-                                "make_shape_para",
-                                "make_three_clust_"), 
+                                "make_shape_para"), 
                         exp = c(
     "Möbius-like cluster combined with a Gaussian.",
     "Multiple Gaussian clusters in high-dimensional space.",
@@ -1992,8 +1991,7 @@ odd_shapes_tb <- tibble(fun = c("make_mobiusgau",
     "Single grid in two dimensions.",
     "Two overlapping grids.",
     "Two grids shifted relative to each other.",
-    "Parallel shaped clusters.",
-    "Three clusters with different shapes. (eg:- 01, 02, ..., 20)")) |> 
+    "Parallel shaped clusters.")) |> 
   mutate(fun = gsub("_", "\\\\_", fun))
 
 

### 2025-11-28 paper-cardinalR.Rmd
Make main argument table to code

@@
                                 "Rotation matrices, one per cluster.",
-                                "Background noise should exist or not."))
+                                "Background noise should exist or not.")) |> 
+    mutate(arg = gsub("_", "\\\\_", arg))
 ```
@@
 main_tb |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", table.pos = "!ht") 
+  mutate(arg = paste0("<code>", arg, "</code>")) |> 
+  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
 ```
@@
 main_tb |> 
-  kable(caption = "The main arguments for gen\\_multicluster().", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "!ht", linesep = "")  |>
+    mutate(arg = paste0("\\texttt{", arg, "}")) |> 
+  kable(caption = "The main arguments for \\texttt{gen\\_multicluster()}.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "!ht", linesep = "", escape = FALSE)  |>
   column_spec(1, width = "2cm") |>

### 2025-11-28 paper-cardinalR.tex
Updated trefoil table
Make main argument table to code

@@
 
-\caption{\label{tab:main-tb-pdf}The main arguments for gen\_multicluster().}
+\caption{\label{tab:main-tb-pdf}The main arguments for \texttt{gen\_multicluster()}.}
 \centering
@@
 \midrule
-n & numeric (vector) & Number of points in each cluster.\\
-k & numeric & Number of clusters.\\
-loc & numeric (matrix) & Locations/centroids of clusters.\\
-scale & numeric (vector) & Scaling factors of clusters.\\
-shape & character (vector) & Shapes of clusters.\\
-rotation & numeric (list) & Rotation matrices, one per cluster.\\
-is\_bkg & boolean & Background noise should exist or not.\\
+\texttt{n} & numeric (vector) & Number of points in each cluster.\\
+\texttt{k} & numeric & Number of clusters.\\
+\texttt{loc} & numeric (matrix) & Locations/centroids of clusters.\\
+\texttt{scale} & numeric (vector) & Scaling factors of clusters.\\
+\texttt{shape} & character (vector) & Shapes of clusters.\\
+\texttt{rotation} & numeric (list) & Rotation matrices, one per cluster.\\
+\texttt{is\_bkg} & boolean & Background noise should exist or not.\\
 \bottomrule
@@
 \midrule
-\texttt{gen\_trefoil4d} & Trefoil in 4-D.\\
-\texttt{gen\_trefoil3d} & Trefoil in 3-D.\\
+\texttt{gen\_trefoil4d} & Trefoil in $4\text{-}D$.\\
+\texttt{gen\_trefoil3d} & Trefoil in $3\text{-}D$.\\
 \bottomrule
@@
 \texttt{make\_shape\_para} & Parallel shaped clusters.\\
-\texttt{make\_three\_clust\_} & Three clusters with different shapes. (eg:- 01, 02, ..., 20)\\
 \bottomrule

### 2025-11-28 paper-cardinalR.Rmd
Added branch parameters to the table

@@
                              "gen_orgcurvybranches"),
+                    parm = c("n, k",
+                             "n, k",
+                             "n, k",
+                             "n, p, k",
+                             "n, p, k"),
                       exp = c("Exponential shaped branches.",
@@
 branch_tb |> 
+  mutate(parm = paste0("<code>", parm, "</code>")) |>
   mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
+  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, table.pos = "!ht") 
 ```
@@
 branch_tb |>
+  mutate(parm = paste0("\\texttt{", parm, "}")) |>
   mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-```{r}
-arg_branching_tb <- tibble(arg = c("n", "k"), 
-                          type = c("integer", "integer"),
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
+  column_spec(1, width = "3.5cm") |>
+  column_spec(2, width = "1.5cm") |>
+  column_spec(3, width = "8cm")
+```
+
+```{r}
+arg_branching_tb <- tibble(arg = c("n", "k", "p"), 
+                          type = c("integer", "integer", "integer"),
                         exp = c("Number of points.",
-                                "Number of clusters."))
+                                "Number of clusters.",
+                                "Number of dimensions."))
 ```
@@
 arg_branching_tb |> 
+  mutate(arg = paste0("<code>", arg, "</code>")) |>
   kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), escape = FALSE, table.pos = "!ht") 
@@
 arg_branching_tb |> 
+  mutate(arg = paste0("\\texttt{", arg, "}")) |> 
   kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, escape = FALSE, table.pos = "!ht")  |>
@@
 
-```{r data-linearbranches, echo=TRUE}
+```{r data-linearbranches}
 linearbranches <- gen_linearbranches(n = 1000, k = 4)
@@
 
-```{r data-curvybranches, echo=TRUE}
+```{r data-curvybranches}
 curvybranches <- gen_curvybranches(n = 1000, k = 4)
@@
 
-```{r data-expbranches, echo=TRUE}
+```{r data-expbranches}
 expbranches <- gen_expbranches(n = 1000, k = 4)
@@
 
-```{r data-orglinearbranches, echo=TRUE}
+```{r data-orglinearbranches}
 orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
@@
 
-```{r data-orgcurvybranches, echo=TRUE}
+```{r data-orgcurvybranches}
 orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)

### 2025-11-28 paper-cardinalR.tex
Added branch parameters to the table

@@
 \centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-\texttt{gen\_expbranches} & Exponential shaped branches.\\
-\texttt{gen\_linearbranches} & Linear shaped branches.\\
-\texttt{gen\_curvybranches} & Curvy shaped branches.\\
-\texttt{gen\_orglinearbranches} & Linear shaped branches originated in one point.\\
-\texttt{gen\_orgcurvybranches} & Curvy shaped branches originated in one point.\\
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+Function & Arguments & Explanation\\
+\midrule
+\texttt{gen\_expbranches} & \texttt{n, k} & Exponential shaped branches.\\
+\texttt{gen\_linearbranches} & \texttt{n, k} & Linear shaped branches.\\
+\texttt{gen\_curvybranches} & \texttt{n, k} & Curvy shaped branches.\\
+\texttt{gen\_orglinearbranches} & \texttt{n, p, k} & Linear shaped branches originated in one point.\\
+\texttt{gen\_orgcurvybranches} & \texttt{n, p, k} & Curvy shaped branches originated in one point.\\
 \bottomrule
@@
 \midrule
-n & integer & Number of points.\\
-k & integer & Number of clusters.\\
+\texttt{n} & integer & Number of points.\\
+\texttt{k} & integer & Number of clusters.\\
+\texttt{p} & integer & Number of dimensions.\\
 \bottomrule
@@
 
-\begin{verbatim}
-linearbranches <- gen_linearbranches(n = 1000, k = 4)
-\end{verbatim}
-
 To introduce curvature, the \texttt{gen\_curvybranches(n,\ k)} function generates \(k\) curvilinear branches in \(2\text{-}D\). Branches \(1\) and \(2\) are simple parabolas defined as
@@
 
-\begin{verbatim}
-curvybranches <- gen_curvybranches(n = 1000, k = 4)
-\end{verbatim}
-
 The \texttt{gen\_expbranches(n,\ k)} function creates \(k\) exponential branches in \(2\text{-}D\), radiating from a central region. Each branch \(i\) is defined as
@@
 The parameter \(s_i\) controls the steepness of branch \(i\), and \(\delta\) introduces small local jitter.
-
-\begin{verbatim}
-expbranches <- gen_expbranches(n = 1000, k = 4)
-\end{verbatim}
 
@@
 The scale factor \(s_i\) controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first \(\binom{p}{2}\) branches, \(s_i = 1\); for additional branches when \(k > \binom{p}{2}\), \(s_i\) is randomly drawn from the set \(\{1, 1.5, 2, \dots, 8\}\).
-
-\begin{verbatim}
-orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
-
-\begin{verbatim}
-orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
-\end{verbatim}
 

### 2025-12-03 paper-cardinalR.Rmd
Started adding the table with parameters

@@
                              "gen_orglinearbranches", 
-                             "gen_orgcurvybranches"),
+                             "gen_orgcurvybranches",
+                             "gen_cone",
+                             "gen_gridcube",
+                             "gen_unifcube",
+                            "gen_gaussian",
+                            "gen_longlinear",
+                            "gen_mobius",
+                            "gen_quadratic",
+                             "gen_cubic",
+                            "gen_pyrrect",
+                             "gen_pyrtri",
+                             "gen_pyrstar",
+                             "gen_pyrfrac",
+                            "gen_scurve",
+                            "gen_circle",
+                            "gen_curvycycle",
+                            "gen_unifsphere",
+                            "gen_hollowsphere",
+                            "gen_gridedsphere",
+                            "gen_clusteredspheres",
+                            "gen_hemisphere",
+                            "gen_swissroll",
+                            "gen_trefoil4d",
+                            "gen_trefoil3d",
+                             "gen_crescent",
+                            "gen_curvycylinder",
+                            "gen_sphericalspiral",
+                            "gen_helicalspiral",
+                            "gen_conicspiral",
+                            "gen_nonlinear"),
                     parm = c("n, k",
@@
                              "n, p, k",
-                             "n, p, k"),
+                             "n, p, k",
+                             "n, p, h, ratio",
+                             "n, p",
+                             "n, p",
+                             "n, p, s",
+                             "n, p",
+                             "n",
+                             "n",
+                             "n",
+                             "n, p",
+                             "n, p",
+                             "n, p",
+                             "n, p",
+                              "n",
+                             "n, p",
+                             "n, p",
+                             "n, r",
+                             "n, p",
+                             "n, k, r, loc",
+                             "n, w",
+                             "n, steps",
+                             "n, steps",
+                             "n", 
+                             "n, h", 
+                             "n, spins", 
+                             "n", 
+                             "n, spins", 
+                             "n, hc, non_fac",
+                             "n",
+                             "n"),
                       exp = c("Exponential shaped branches.",
@@
                                "Linear shaped branches originated in one point.",
-                               "Curvy shaped branches originated in one point.")) |> 
-   mutate(fun = gsub("_", "\\\\_", fun))
+                               "Curvy shaped branches originated in one point.",
+                               "Cone-shaped structure.",
+                              "Cube with specified grid points along each axes.",
+                              "Cube with uniform points.",
+                              "Multivariate Gaussian cloud.",
+                              "Long linear structure.",
+                              "Möbius strip in $3\text{-}D$.",
+                              "Quadratic pattern in $2\text{-}D$.",
+                                "Cubic pattern in $2\text{-}D$.",
+                              "Rectangular-base, with a sharp or blunted apex.",
+          "Triangular-base, with a sharp or blunted apex.",
+          "Star-shaped base, with a sharp or blunted apex.",
+          "Pyramid with triangular pyramid-shaped holes.",
+          "S-curve in $3\text{-}D$.",
+          "Circle.",
+                                "Curvy cell cycle.",
+                                "Uniform ball.",
+                                "Hollow sphere.",
+                                "Grided sphere.",
+                                "Multiple small spheres within a big sphere.",
+                                "Hemisphere.",
+          "Swissroll structure.",
+          "Trefoil in $4\\text{-}D$.", 
+          "Trefoil in $3\\text{-}D$.",
+          "Crescent pattern.",
+                                "Curvy cylinder.",
+                                "Spherical spiral.",
+                                "Helical spiral.",
+                                "Conic spiral.",
+                                "Nonlinear hyperbola.")) |> 
+   mutate(fun = gsub("_", "\\\\_", fun)) |> 
+   mutate(parm = gsub("_", "\\\\_", parm))
 ```
@@
 
-```{r data-bluntedcone, echo=TRUE}
+```{r data-bluntedcone}
 cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
@@
 
-```{r}
-cube_tb <- tibble(fun = c("gen_gridcube",
-                          "gen_unifcube"), 
-                      exp = c("Cube with specified grid points along each axes.",
-                              "Cube with uniform points.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r cube-tb-html, eval=knitr::is_html_output()}
-cube_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |>
-  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
-```
-
-```{r cube-tb-pdf, eval=knitr::is_latex_output()}
-cube_tb |> 
-  mutate(fun = paste0("\\texttt{", fun, "}")) |>
-  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
 The function `gen_gridcube(n, p)` is a wrapper around `geozoo::cube.solid.grid()`.
@@
 
-```{r data-gridcube, echo=TRUE}
+```{r data-gridcube}
 gridcube <- gen_gridcube(n = 1000, p = 4)
@@
 
-```{r data-unifcube, echo=TRUE}
+```{r data-unifcube}
 unifcube <- gen_unifcube(n = 1000, p = 4)
@@
 
-```{r}
-polynomial_tb <- tibble(fun = c("gen_quadratic",
-                                "gen_cubic"), 
-                        exp = c("Quadratic pattern.",
-                                "Cubic pattern.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r polynomial-tb-html, eval=knitr::is_html_output()}
-polynomial_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
-```
-
-```{r polynomial-tb-pdf, eval=knitr::is_latex_output()}
-polynomial_tb |> 
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
 The first is the quadratic curve of $n$ points in two dimensions.
@@
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:pyramid-tb-html)', '\\@ref(tab:pyramid-tb-pdf)'))` summarizes these functions.
-
-```{r}
-pyramid_tb <- tibble(fun = c("gen_pyrrect",
-                             "gen_pyrtri",
-                             "gen_pyrstar",
-                             "gen_pyrfrac"), 
-                        exp = c("Rectangular-base, with a sharp or blunted apex.",
-          "Triangular-base, with a sharp or blunted apex.",
-          "Star-shaped base, with a sharp or blunted apex.",
-          "Pyramid with triangular pyramid-shaped holes.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r pyramid-tb-html, eval=knitr::is_html_output()}
-pyramid_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
-```
-
-```{r pyramid-tb-pdf, eval=knitr::is_latex_output()}
-pyramid_tb |> 
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, escape = FALSE, table.pos = "!ht")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
 
@@
 
-```{r}
-sphere_tb <- tibble(fun = c("gen_circle",
-                            "gen_curvycycle",
-                            "gen_unifsphere",
-                            "gen_hollowsphere",
-                            "gen_gridedsphere",
-                            "gen_clusteredspheres",
-                            "gen_hemisphere"), 
-                        exp = c("Circle.",
-                                "Curvy cell cycle.",
-                                "Uniform ball.",
-                                "Hollow sphere.",
-                                "Grided sphere.",
-                                "Multiple small spheres within a big sphere.",
-                                "Hemisphere.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r sphere-tb-html, eval=knitr::is_html_output()}
-sphere_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
-```
-
-```{r sphere-tb-pdf, eval=knitr::is_latex_output()}
-sphere_tb |> 
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, linesep = "", table.pos = "!ht")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
 The simplest case, `gen_circle(n, p)` creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
@@
 
-```{r}
-trefoil_tb <- tibble(fun = c("gen_trefoil4d",
-                            "gen_trefoil3d"), 
-                        exp = c("Trefoil in $4\\text{-}D$.",
-                                "Trefoil in $3\\text{-}D$.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r trefoil-tb-html, eval=knitr::is_html_output()}
-trefoil_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
-```
-
-```{r trefoil-tb-pdf, eval=knitr::is_latex_output()}
-
-trefoil_tb |>
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
 For the $4\text{-}D$ trefoil knot [@laurent2024], the function `gen_trefoil4d(n, steps)` generates the structure on the $3$-sphere ($S^3 \subset \mathbb{R}^4$) using two angular parameters, $\theta$ and $\phi$.
@@
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trigonometric-tb-html)', '\\@ref(tab:trigonometric-tb-pdf)'))` summarizes these functions.
-
-```{r}
-trigonometric_tb <- tibble(fun = c("gen_crescent",
-                            "gen_curvycylinder",
-                            "gen_sphericalspiral",
-                            "gen_helicalspiral",
-                            "gen_conicspiral",
-                            "gen_nonlinear"), 
-                        exp = c("Crescent pattern.",
-                                "Curvy cylinder.",
-                                "Spherical spiral.",
-                                "Helical spiral.",
-                                "Conic spiral.",
-                                "Nonlinear hyperbola.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r trigonometric-tb-html, eval=knitr::is_html_output()}
-trigonometric_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
-```
-
-```{r trigonometric-tb-pdf, eval=knitr::is_latex_output()}
-trigonometric_tb |> 
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, linesep = "", table.pos = "!ht")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
 

### 2025-12-03 paper-cardinalR.tex
Added DOI for all
Started adding the table with parameters

@@
 \texttt{gen\_orgcurvybranches} & \texttt{n, p, k} & Curvy shaped branches originated in one point.\\
+\addlinespace
+\texttt{gen\_cone} & \texttt{n, p, h, ratio} & Cone-shaped structure.\\
+\texttt{gen\_gridcube} & \texttt{n, p} & Cube with specified grid points along each axes.\\
+\texttt{gen\_unifcube} & \texttt{n, p} & Cube with uniform points.\\
+\texttt{gen\_gaussian} & \texttt{n, p, s} & Multivariate Gaussian cloud.\\
+\texttt{gen\_longlinear} & \texttt{n, p} & Long linear structure.\\
+\addlinespace
+\texttt{gen\_mobius} & \texttt{n} & Möbius strip in $3  ext{-}D$.\\
+\texttt{gen\_quadratic} & \texttt{n} & Quadratic pattern in $2  ext{-}D$.\\
+\texttt{gen\_cubic} & \texttt{n} & Cubic pattern in $2  ext{-}D$.\\
+\texttt{gen\_pyrrect} & \texttt{n, p} & Rectangular-base, with a sharp or blunted apex.\\
+\texttt{gen\_pyrtri} & \texttt{n, p} & Triangular-base, with a sharp or blunted apex.\\
+\addlinespace
+\texttt{gen\_pyrstar} & \texttt{n, p} & Star-shaped base, with a sharp or blunted apex.\\
+\texttt{gen\_pyrfrac} & \texttt{n, p} & Pyramid with triangular pyramid-shaped holes.\\
+\texttt{gen\_scurve} & \texttt{n} & S-curve in $3   ext{-}D$.\\
+\texttt{gen\_circle} & \texttt{n, p} & Circle.\\
+\texttt{gen\_curvycycle} & \texttt{n, p} & Curvy cell cycle.\\
+\addlinespace
+\texttt{gen\_unifsphere} & \texttt{n, r} & Uniform ball.\\
+\texttt{gen\_hollowsphere} & \texttt{n, p} & Hollow sphere.\\
+\texttt{gen\_gridedsphere} & \texttt{n, k, r, loc} & Grided sphere.\\
+\texttt{gen\_clusteredspheres} & \texttt{n, w} & Multiple small spheres within a big sphere.\\
+\texttt{gen\_hemisphere} & \texttt{n, steps} & Hemisphere.\\
+\addlinespace
+\texttt{gen\_swissroll} & \texttt{n, steps} & Swissroll structure.\\
+\texttt{gen\_trefoil4d} & \texttt{n} & Trefoil in $4\text{-}D$.\\
+\texttt{gen\_trefoil3d} & \texttt{n, h} & Trefoil in $3\text{-}D$.\\
+\texttt{gen\_crescent} & \texttt{n, spins} & Crescent pattern.\\
+\texttt{gen\_curvycylinder} & \texttt{n} & Curvy cylinder.\\
+\addlinespace
+\texttt{gen\_sphericalspiral} & \texttt{n, spins} & Spherical spiral.\\
+\texttt{gen\_helicalspiral} & \texttt{n, hc, non\_fac} & Helical spiral.\\
+\texttt{gen\_conicspiral} & \texttt{n} & Conic spiral.\\
+\texttt{gen\_nonlinear} & \texttt{n} & Nonlinear hyperbola.\\
 \bottomrule
@@
 
-\begin{verbatim}
-cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
-\end{verbatim}
-
 Cone-shaped structures appear in particle dispersions, light beams, and tapering processes, where spread decreases along one axis.
@@
 
-\begin{table}
-
-\caption{\label{tab:cube-tb-pdf}cardinalR cube data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-\texttt{gen\_gridcube} & Cube with specified grid points along each axes.\\
-\texttt{gen\_unifcube} & Cube with uniform points.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 The function \texttt{gen\_gridcube(n,\ p)} is a wrapper around \texttt{geozoo::cube.solid.grid()}.
@@
 
-\begin{verbatim}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-\end{verbatim}
-
 By contrast, \texttt{gen\_unifcube(n,\ p)} wraps \texttt{geozoo::cube.solid.random()}, producing uniformly distributed points within a \(p\text{-}D\) cube.
@@
 This results in a hypercube filled with random, but evenly distributed, samples rather than structured lattice points.
-
-\begin{verbatim}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-\end{verbatim}
 
@@
 
-\begin{table}
-
-\caption{\label{tab:polynomial-tb-pdf}cardinalR polynomial data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-\texttt{gen\_quadratic} & Quadratic pattern.\\
-\texttt{gen\_cubic} & Cubic pattern.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 The first is the quadratic curve of \(n\) points in two dimensions.
@@
 Table \ref{tab:pyramid-tb-pdf} summarizes these functions.
-
-\begin{table}
-
-\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-\texttt{gen\_pyrrect} & Rectangular-base, with a sharp or blunted apex.\\
-\texttt{gen\_pyrtri} & Triangular-base, with a sharp or blunted apex.\\
-\texttt{gen\_pyrstar} & Star-shaped base, with a sharp or blunted apex.\\
-\texttt{gen\_pyrfrac} & Pyramid with triangular pyramid-shaped holes.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 Such spherical or hemispherical structures frequently appear in physical and biological systems, for example in models of celestial bodies, molecular shells, or cell membranes \citep{tinkham2003, alberts2014}.Table \ref{tab:sphere-tb-pdf} summarizes these functions.
-
-\begin{table}
-
-\caption{\label{tab:sphere-tb-pdf}cardinalR sphere data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-\texttt{gen\_circle} & Circle.\\
-\texttt{gen\_curvycycle} & Curvy cell cycle.\\
-\texttt{gen\_unifsphere} & Uniform ball.\\
-\texttt{gen\_hollowsphere} & Hollow sphere.\\
-\texttt{gen\_gridedsphere} & Grided sphere.\\
-\texttt{gen\_clusteredspheres} & Multiple small spheres within a big sphere.\\
-\texttt{gen\_hemisphere} & Hemisphere.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 
-\begin{table}
-
-\caption{\label{tab:trefoil-tb-pdf}cardinalR trefoil data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-\texttt{gen\_trefoil4d} & Trefoil in $4\text{-}D$.\\
-\texttt{gen\_trefoil3d} & Trefoil in $3\text{-}D$.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 For the \(4\text{-}D\) trefoil knot \citep{laurent2024}, the function \texttt{gen\_trefoil4d(n,\ steps)} generates the structure on the \(3\)-sphere (\(S^3 \subset \mathbb{R}^4\)) using two angular parameters, \(\theta\) and \(\phi\).
@@
 Table \ref{tab:trigonometric-tb-pdf} summarizes these functions.
-
-\begin{table}
-
-\caption{\label{tab:trigonometric-tb-pdf}cardinalR trigonometric data generation functions}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-Function & Explanation\\
-\midrule
-\texttt{gen\_crescent} & Crescent pattern.\\
-\texttt{gen\_curvycylinder} & Curvy cylinder.\\
-\texttt{gen\_sphericalspiral} & Spherical spiral.\\
-\texttt{gen\_helicalspiral} & Helical spiral.\\
-\texttt{gen\_conicspiral} & Conic spiral.\\
-\texttt{gen\_nonlinear} & Nonlinear hyperbola.\\
-\bottomrule
-\end{tabular}
-\end{table}
 

### 2025-12-03 paper-cardinalR.R

@@ -158,18 +158,106 @@ branch_tb <- tibble(fun = c("gen_expbranches",
                              "gen_linearbranches",
                              "gen_curvybranches", 
                              "gen_orglinearbranches", 
-                             "gen_orgcurvybranches"),
+                             "gen_orgcurvybranches",
+                             "gen_cone",
+                             "gen_gridcube",
+                             "gen_unifcube",
+                            "gen_gaussian",
+                            "gen_longlinear",
+                            "gen_mobius",
+                            "gen_quadratic",
+                             "gen_cubic",
+                            "gen_pyrrect",
+                             "gen_pyrtri",
+                             "gen_pyrstar",
+                             "gen_pyrfrac",
+                            "gen_scurve",
+                            "gen_circle",
+                            "gen_curvycycle",
+                            "gen_unifsphere",
+                            "gen_hollowsphere",
+                            "gen_gridedsphere",
+                            "gen_clusteredspheres",
+                            "gen_hemisphere",
+                            "gen_swissroll",
+                            "gen_trefoil4d",
+                            "gen_trefoil3d",
+                             "gen_crescent",
+                            "gen_curvycylinder",
+                            "gen_sphericalspiral",
+                            "gen_helicalspiral",
+                            "gen_conicspiral",
+                            "gen_nonlinear"),
                     parm = c("n, k",
                              "n, k",
                              "n, k",
                              "n, p, k",
-                             "n, p, k"),
+                             "n, p, k",
+                             "n, p, h, ratio",
+                             "n, p",
+                             "n, p",
+                             "n, p, s",
+                             "n, p",
+                             "n",
+                             "n",
+                             "n",
+                             "n, p",
+                             "n, p",
+                             "n, p",
+                             "n, p",
+                              "n",
+                             "n, p",
+                             "n, p",
+                             "n, r",
+                             "n, p",
[…]

### 2025-12-03 paper-cardinalR.Rmd
Removed the code chunks

@@
                              "n, p",
+                             "n",
                              "n, k, r, loc",
+                             "n, p",
                              "n, w",
@@
                              "n, spins", 
-                             "n, hc, non_fac",
-                             "n",
-                             "n"),
+                             "n, hc, non_fac"),
                       exp = c("Exponential shaped branches.",
@@
   mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
+  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht", linesep = "")  |>
   column_spec(1, width = "3.5cm") |>
@@
 
-```{r data-gau, echo=TRUE}
+```{r data-gau}
 gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
@@
 
-```{r data-linear, echo=TRUE}
+```{r data-linear}
 linear <- gen_longlinear(n = 1000, p = 4)
@@
 
-```{r data-mobius, echo=TRUE}
+```{r data-mobius}
 mobius <- gen_mobius(n = 1000)
@@
 
-```{r data-quadratic, echo=TRUE}
+```{r data-quadratic}
 quadratic <- gen_quadratic(n = 1000)
@@
 
-```{r data-cubic, echo=TRUE}
+```{r data-cubic}
 cubic <- gen_cubic(n = 1000)
@@
 
-```{r data-pyrrect, echo=TRUE}
+```{r data-pyrrect}
 pyrrect <- gen_pyrrect(n = 1000, p = 4)
@@
 
-```{r data-pyrtri, echo=TRUE}
+```{r data-pyrtri}
 pyrtri <- gen_pyrtri(n = 1000, p = 4)
@@
 
-```{r data-pyrstar, echo=TRUE}
+```{r data-pyrstar}
 pyrstar <- gen_pyrstar(n = 1000, p = 4)
@@
 
-```{r data-pyrholes, echo=TRUE}
+```{r data-pyrholes}
 pyrholes <- gen_pyrfrac(n = 1000, p = 4)
@@
 
-```{r data-scurve, echo=TRUE}
+```{r data-scurve}
 scurve <- gen_scurve(n = 1000)
@@
 
-```{r data-circle, echo=TRUE}
+```{r data-circle}
 circle <- gen_circle(n = 1000, p = 4)
@@
 
-```{r data-curvycycle, echo=TRUE}
+```{r data-curvycycle}
 curvycycle <- gen_curvycycle(n = 1000, p = 4)
@@
 
-```{r data-unifsphere, echo=TRUE}
+```{r data-unifsphere}
 unifsphere <- gen_unifsphere(n = 1000, r = 1)
@@
 
-```{r data-hollowsphere, echo=TRUE}
+```{r data-hollowsphere}
 hollowsphere <- gen_hollowsphere(n = 1000, p = 4)
@@
 
-```{r data-clusteredspheres, echo=TRUE}
+```{r data-clusteredspheres}
 clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, r = c(15, 3),
@@
 
-```{r data-swissroll, echo=TRUE}
+```{r data-swissroll}
 swissroll <- gen_swissroll(n = 1000, w = c(-1, 1))
@@
 
-```{r data-trefoil4d, echo=TRUE}
+```{r data-trefoil4d}
 trefoil4d <- gen_trefoil4d(n = 500, steps = 5)
@@
 
-```{r data-trefoil3d, echo=TRUE}
+```{r data-trefoil3d}
 trefoil3d <- gen_trefoil3d(n = 500, steps = 5)
@@
 
-```{r data-crescent, echo=TRUE}
+```{r data-crescent}
 crescent <- gen_crescent(n = 1000)
@@
 
-```{r data-curvycylinder, echo=TRUE}
+```{r data-curvycylinder}
 curvycylinder <- gen_curvycylinder(n = 1000, h = 10)
@@
 
-```{r data-sphericalspiral, echo=TRUE}
+```{r data-sphericalspiral}
 sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
@@
 
-```{r data-helicalspiral, echo=TRUE}
+```{r data-helicalspiral}
 helicalspiral <- gen_helicalspiral(n = 1000)
@@
 
-```{r data-conicspiral, echo=TRUE}
+```{r data-conicspiral}
 conicspiral <- gen_conicspiral(n = 1000, spins = 1)
@@
 
-```{r data-nonlinear, echo=TRUE}
+```{r data-nonlinear}
 nonlinear <- gen_nonlinear(n = 1000, hc = 1, non_fac = 0.5)

### 2025-12-03 paper-cardinalR.tex
Removed the code chunks

@@
 \texttt{gen\_orgcurvybranches} & \texttt{n, p, k} & Curvy shaped branches originated in one point.\\
-\addlinespace
 \texttt{gen\_cone} & \texttt{n, p, h, ratio} & Cone-shaped structure.\\
@@
 \texttt{gen\_longlinear} & \texttt{n, p} & Long linear structure.\\
-\addlinespace
 \texttt{gen\_mobius} & \texttt{n} & Möbius strip in $3  ext{-}D$.\\
@@
 \texttt{gen\_pyrtri} & \texttt{n, p} & Triangular-base, with a sharp or blunted apex.\\
-\addlinespace
 \texttt{gen\_pyrstar} & \texttt{n, p} & Star-shaped base, with a sharp or blunted apex.\\
@@
 \texttt{gen\_curvycycle} & \texttt{n, p} & Curvy cell cycle.\\
-\addlinespace
 \texttt{gen\_unifsphere} & \texttt{n, r} & Uniform ball.\\
 \texttt{gen\_hollowsphere} & \texttt{n, p} & Hollow sphere.\\
-\texttt{gen\_gridedsphere} & \texttt{n, k, r, loc} & Grided sphere.\\
-\texttt{gen\_clusteredspheres} & \texttt{n, w} & Multiple small spheres within a big sphere.\\
-\texttt{gen\_hemisphere} & \texttt{n, steps} & Hemisphere.\\
-\addlinespace
-\texttt{gen\_swissroll} & \texttt{n, steps} & Swissroll structure.\\
-\texttt{gen\_trefoil4d} & \texttt{n} & Trefoil in $4\text{-}D$.\\
-\texttt{gen\_trefoil3d} & \texttt{n, h} & Trefoil in $3\text{-}D$.\\
-\texttt{gen\_crescent} & \texttt{n, spins} & Crescent pattern.\\
-\texttt{gen\_curvycylinder} & \texttt{n} & Curvy cylinder.\\
-\addlinespace
+\texttt{gen\_gridedsphere} & \texttt{n} & Grided sphere.\\
+\texttt{gen\_clusteredspheres} & \texttt{n, k, r, loc} & Multiple small spheres within a big sphere.\\
+\texttt{gen\_hemisphere} & \texttt{n, p} & Hemisphere.\\
+\texttt{gen\_swissroll} & \texttt{n, w} & Swissroll structure.\\
+\texttt{gen\_trefoil4d} & \texttt{n, steps} & Trefoil in $4\text{-}D$.\\
+\texttt{gen\_trefoil3d} & \texttt{n, steps} & Trefoil in $3\text{-}D$.\\
+\texttt{gen\_crescent} & \texttt{n} & Crescent pattern.\\
+\texttt{gen\_curvycylinder} & \texttt{n, h} & Curvy cylinder.\\
 \texttt{gen\_sphericalspiral} & \texttt{n, spins} & Spherical spiral.\\
-\texttt{gen\_helicalspiral} & \texttt{n, hc, non\_fac} & Helical spiral.\\
-\texttt{gen\_conicspiral} & \texttt{n} & Conic spiral.\\
-\texttt{gen\_nonlinear} & \texttt{n} & Nonlinear hyperbola.\\
+\texttt{gen\_helicalspiral} & \texttt{n} & Helical spiral.\\
+\texttt{gen\_conicspiral} & \texttt{n, spins} & Conic spiral.\\
+\texttt{gen\_nonlinear} & \texttt{n, hc, non\_fac} & Nonlinear hyperbola.\\
 \bottomrule
@@
 
-\begin{verbatim}
-gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
-\end{verbatim}
-
 Gaussian clouds are common benchmark structures in statistics and machine learning, used in clustering, classification, and anomaly detection, with applications in image segmentation, speech recognition, and forensic analysis \citep{geoffrey2000}.
@@
 
-\begin{verbatim}
-linear <- gen_longlinear(n = 1000, p = 4)
-\end{verbatim}
-
 This structure appears in \(p\text{-}D\) data when variation is driven by a single factor, such as time-course or sensor measurements, providing a useful test case for trajectory and regression methods \citep{trapnell2014}.
@@
 The function returns a tibble with \(n\) sampled points forming the surface of a Möbius strip.
-
-\begin{verbatim}
-mobius <- gen_mobius(n = 1000)
-\end{verbatim}
 
@@
 
-\begin{verbatim}
-quadratic <- gen_quadratic(n = 1000)
-\end{verbatim}
-
 The second is the cubic curve of \(n\) points in two dimensions.
@@
 This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
-
-\begin{verbatim}
-cubic <- gen_cubic(n = 1000)
-\end{verbatim}
 
@@
 
-\begin{verbatim}
-pyrrect <- gen_pyrrect(n = 1000, p = 4)
-\end{verbatim}
-
 For the triangular based pyramid, \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} (Figure \ref{fig:pyr-proj} b), let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\).
@@
 
-\begin{verbatim}
-pyrtri <- gen_pyrtri(n = 1000, p = 4)
-\end{verbatim}
-
 For the star based pyramid, \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} (Figure \ref{fig:pyr-proj} c), let the radius at height \(z\), \(r(z)\), be such that the radius scales linearly from zero (tip) to the base radius \(r_b\).
@@
 Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
-
-\begin{verbatim}
-pyrstar <- gen_pyrstar(n = 1000, p = 4)
-\end{verbatim}
 
@@
 The final output is a \(n \times p\) matrix where each row represents a point: \(X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}\).
-
-\begin{verbatim}
-pyrholes <- gen_pyrfrac(n = 1000, p = 4)
-\end{verbatim}
 
@@
 
-\begin{verbatim}
-scurve <- gen_scurve(n = 1000)
-\end{verbatim}
-
 \subsubsection{Sphere}\label{sphere}
@@
 
-\begin{verbatim}
-circle <- gen_circle(n = 1000, p = 4)
-\end{verbatim}
-
 For the one-dimensional nonlinear cycle embedded in \(p\text{-}D\) space, \texttt{gen\_curvycycle(n,\ p)} (Figure \ref{fig:sphere-proj} b), let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
@@
 
-\begin{verbatim}
-curvycycle <- gen_curvycycle(n = 1000, p = 4)
-\end{verbatim}
-
 Building on simple circular structures, the \texttt{gen\_unifsphere(n,\ r)} function function extends the idea to three dimensions by generating \(n\) observations approximately uniformly distributed on the surface of a sphere of radius \(r\)..
@@
 Cartesian coordinates are then defined as \[X_1 = r\sqrt{1 - u^2}\cos(\theta), \quad X_2 = r\sqrt{1 - u^2}\sin(\theta),\text{ and }X_3 = ru,\] ensuring uniform distribution on the surface (not within) of the sphere.
-
-\begin{verbatim}
-unifsphere <- gen_unifsphere(n = 1000, r = 1)
-\end{verbatim}
 
@@
 These paired structures allow controlled experiments to investigate how algorithms behave when data is concentrated throughout the full volume versus constrained to the boundary.
-
-\begin{verbatim}
-hollowsphere <- gen_hollowsphere(n = 1000, p = 4)
-\end{verbatim}
 
@@
 
-\begin{verbatim}
-clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, r = c(15, 3),
-                                         loc = 10 / sqrt(3)) |>
-  dplyr::select(-cluster)
-\end{verbatim}
-
 Finally, the \texttt{gen\_hemisphere(n,\ p)} function restricts sampling to a hemisphere of a \(4\text{-}D\) sphere (Figure \ref{fig:sphere-proj} f).
@@
 
-\begin{verbatim}
-swissroll <- gen_swissroll(n = 1000, w = c(-1, 1))
-\end{verbatim}
-
 Compared with \texttt{sndata::swiss\_roll()} \citep{james2025}, this implementation (i) samples \(t\) over \([0, 3\pi]\) instead of \([1.5\pi, 4.5\pi]\), (ii) allows a flexible vertical range \(w = (w_1, w_2)\) rather than fixing \(z \in [0, z_{\max}]\), and (iii) returns a tibble with \texttt{x1,\ x2,\ x3} instead of adding a color variable.
@@
 
-\begin{verbatim}
-trefoil4d <- gen_trefoil4d(n = 500, steps = 5)
-\end{verbatim}
-
 For the \(3\text{-}D\) stereographic projection \citep{laurent2024}, \texttt{gen\_trefoil3d(n,\ steps)} maps each point \((X_1, X_2, X_3, X_4) \in \mathbb{R}^4\) to \((X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),\) excluding points where \(X_4 = 1\) to avoid division by zero (Figure \ref{fig:trefoil-proj} b).
-
-\begin{verbatim}
-trefoil3d <- gen_trefoil3d(n = 500, steps = 5)
-\end{verbatim}
 
@@
 
-\begin{verbatim}
-crescent <- gen_crescent(n = 1000)
-\end{verbatim}
-
 Second, the \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
@@
 
-\begin{verbatim}
-curvycylinder <- gen_curvycylinder(n = 1000, h = 10)
-\end{verbatim}
-
 For a spiraling path on a spherical surface in the first four dimensions, \texttt{gen\_sphericalspiral(n,\ p,\ spins)} (Figure \ref{fig:triginometric-proj} c), let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\)be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
@@
 
-\begin{verbatim}
-sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
-\end{verbatim}
-
 For a helical spiral in four dimensions, \texttt{gen\_helicalspiral(n,\ p)} (Figure \ref{fig:triginometric-proj} d), let \(\theta \in [0, 5\pi/4]\) be a sequence of angles controlling rotation around a circle.
 Cartesian coordinates; \(X_1 = \cos(\theta)\): circular trajectory along the x-axis, \(X_2 = \sin(\theta)\): circular trajectory along the y-axis, \(X_3 = 0.05\theta + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.5, 0.5)\): linear progression (height) with vertical jitter, simulating a helix, and \(X_4 = 0.1\sin(\theta)\): oscillates with \(\theta\), representing a periodic ``wobble'' along the fourth dimension.
-
-\begin{verbatim}
-helicalspiral <- gen_helicalspiral(n = 1000)
-\end{verbatim}
 
@@
 
-\begin{verbatim}
-conicspiral <- gen_conicspiral(n = 1000, spins = 1)
-\end{verbatim}
-
 Finally, the \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
@@
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-
-\begin{verbatim}
-nonlinear <- gen_nonlinear(n = 1000, hc = 1, non_fac = 0.5)
-\end{verbatim}
 

### 2025-12-03 paper-cardinalR.Rmd
Removed table references

@@
 
-### Branching
-
-A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology [@trapnell2014].
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (`n`), and number of branches (`k`).
-Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:branching-tb-html)', '\\@ref(tab:branching-tb-pdf)'))` outlines these functions.
@@
 
+### Branching
+
+A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology [@trapnell2014].
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (`n`), and number of branches (`k`).
+Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
+
 The simplest structures are approximately linear branches in $2\text{-}D$, generated by the `gen_linearbranches(n, k)` function.
@@
 We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, and uniform random points.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:cube-tb-html)', '\\@ref(tab:cube-tb-pdf)'))` outlines these functions and their purposes.
 
@@
 We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:polynomial-tb-html)', '\\@ref(tab:polynomial-tb-pdf)'))` summarizes these functions and their purposes.
 
@@
 The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
-Such spherical or hemispherical structures frequently appear in physical and biological systems, for example in models of celestial bodies, molecular shells, or cell membranes [@tinkham2003; @alberts2014].Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:sphere-tb-html)', '\\@ref(tab:sphere-tb-pdf)'))` summarizes these functions.
+Such spherical or hemispherical structures frequently appear in physical and biological systems, for example in models of celestial bodies, molecular shells, or cell membranes [@tinkham2003; @alberts2014].
 
@@
 The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trefoil-tb-html)', '\\@ref(tab:trefoil-tb-pdf)'))` summarizes these functions.
 
@@
 These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features [@calladine1997; @gershenfeld2000].
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trigonometric-tb-html)', '\\@ref(tab:trigonometric-tb-pdf)'))` summarizes these functions.
 

### 2025-12-03 paper-cardinalR.tex
Removed table references

@@
 
-\subsubsection{Branching}\label{branching}
-
-A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology \citep{trapnell2014}.
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (\texttt{n}), and number of branches (\texttt{k}).
-Although these functions can generate multiple branches, they do not produce a formal \emph{multicluster} dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
-Table \ref{tab:branching-tb-pdf} outlines these functions.
-The main arguments of the functions described in Table \ref{tab:arg-branching-tb-pdf}.
+Table \ref{tab:branching-tb-pdf} outlines these functions. The main arguments of the functions described in Table \ref{tab:arg-branching-tb-pdf}.
 
@@
 \end{table}
+
+\subsubsection{Branching}\label{branching}
+
+A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology \citep{trapnell2014}.
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (\texttt{n}), and number of branches (\texttt{k}).
+Although these functions can generate multiple branches, they do not produce a formal \emph{multicluster} dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
 
@@
 We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, and uniform random points.
-Table \ref{tab:cube-tb-pdf} outlines these functions and their purposes.
 
@@
 We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-Table \ref{tab:polynomial-tb-pdf} summarizes these functions and their purposes.
 
@@
 The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
-Such spherical or hemispherical structures frequently appear in physical and biological systems, for example in models of celestial bodies, molecular shells, or cell membranes \citep{tinkham2003, alberts2014}.Table \ref{tab:sphere-tb-pdf} summarizes these functions.
+Such spherical or hemispherical structures frequently appear in physical and biological systems, for example in models of celestial bodies, molecular shells, or cell membranes \citep{tinkham2003, alberts2014}.
 
@@
 The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
-Table \ref{tab:trefoil-tb-pdf} summarizes these functions.
 
@@
 These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features \citep{calladine1997, gershenfeld2000}.
-Table \ref{tab:trigonometric-tb-pdf} summarizes these functions.
 

### 2025-12-04 paper-cardinalR.Rmd
set install chunk to not evaluate
figure captions and langevitour cite with new paragraph in intro

@@
 
-```{r install-libraries, include=FALSE, warning=FALSE, echo=FALSE}
+```{r install-libraries, include=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
 
@@
 Developing these generators for research purposes underlies `cardinalR`, which is now a general-purpose package that should be useful for research and teaching.
+
+The example data structures are best viewed using a tour [@As85].
+These show the data as a sequence of low dimensional projections (typically 2D), providing a good sense of the shape in high dimensions.
+The interactive tour plots included in this paper are produced using the software langevitour [@paul2023].
 
@@
 
-```{r branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$.
-The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles.", fig.pos="!ht"}
+```{r branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ output of the `gen_orgcurvybranches()`.
+The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"}
 
@@
 
-```{r cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $3\\text{-}D$.
+```{r cone, eval=knitr::is_html_output(), fig.cap="Viewing the $3\\text{-}D$ output  of `gen_cone()`.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex.", fig.pos="!ht"}
@@
 
-```{r pyr, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data in $4\\text{-}D$.
+```{r pyr, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$  `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data.
 The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
@@
 ```{r}
-circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
 ```
@@
 ```{r}
-curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
 ```
@@
 ```{r}
-clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
 ```
@@
 
-```{r sphere, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `circle`, `curvycycle` data in $4\\text{-}D$, and `clusteredspheres` data in $3\\text{-}D$.
+```{r sphere, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `circle`, `curvycycle` data, and $3\\text{-}D$ `clusteredspheres` data.
 The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.alt="Interactive langevitour of three datasets: circle and curvycycle in 4-D showing closed-loop and wavy twisted-ring structures, and clusteredspheres in 3-D showing multiple compact spherical clusters with slight overlap in projections.", fig.pos="!ht"}
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.alt="Interactive langevitour of three datasets: circle and curvycycle in 4-D showing closed-loop and wavy twisted-ring structures, and clusteredspheres in 3-D showing multiple compact spherical clusters with slight overlap in projections.", fig.pos="!ht", layout="l-body"}
 
@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;
+    style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
     gap: 0px;
@@
     padding: 0;",
-    circle_lang, curvycycle_lang, clusteredspheres_lang
-  ),
+    circle_lang, curvycycle_lang, clusteredspheres_lang),
+  #widths=c(3,3,3),
   device = "xs"
@@
 
-The Swiss roll is a standard nonlinear manifold, representing a plane curled into $3\text{-}D$.
+The Swiss roll is a plane curled into $3\text{-}D$, and is a commonly used example of a nonlinear manifold.
 The `gen_swissroll(n, w)` generates points as $X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad t \sim U(0, 3\pi).$
@@
 
-```{r trefoil, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `trefoil4d` in $4\\text{-}D$ and `trefoil3d` data in $3\\text{-}D$.
+```{r trefoil, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `trefoil4d` and $3\\text{-}D$ `trefoil3d` data.
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
@@
 
-```{r triginometric, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of tthe `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data in $4\\text{-}D$.
+```{r triginometric, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data.
 The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
@@
 
-```{r highd-data, eval=knitr::is_html_output(), fig.cap="`langevitour` output showing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in 4-D: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht"}
+```{r highd-data, eval=knitr::is_html_output(), fig.cap="Viewing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in 4-D: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht"}
 

### 2025-12-04 paper-cardinalR.tex
figure captions and langevitour cite with new paragraph in intro

@@
 Developing these generators for research purposes underlies \texttt{cardinalR}, which is now a general-purpose package that should be useful for research and teaching.
+
+The example data structures are best viewed using a tour \citep{As85}.
+These show the data as a sequence of low dimensional projections (typically 2D), providing a good sense of the shape in high dimensions.
+The interactive tour plots included in this paper are produced using the software langevitour \citep{paul2023}.
 
@@
 
-The Swiss roll is a standard nonlinear manifold, representing a plane curled into \(3\text{-}D\).
+The Swiss roll is a plane curled into \(3\text{-}D\), and is a commonly used example of a nonlinear manifold.
 The \texttt{gen\_swissroll(n,\ w)} generates points as \(X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad t \sim U(0, 3\pi).\)

### 2025-12-04 paper-cardinalR.R

@@ -406,7 +406,7 @@ orglinearbranches_proj3 <- plot_proj(
 orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 
 
-## ----branch, eval=knitr::is_html_output(), fig.cap="`langevitour` output of the `orglinearbranches` data in $4\\text{-}D$. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles.", fig.pos="!ht"----
+## ----branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ output of the `gen_orgcurvybranches()`. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"----
 # 
 # branchfig <- bscols(
 #   htmltools::div(
@@ -522,7 +522,7 @@ cone_proj3 <- plot_proj(
 
 
 
-## ----cone, eval=knitr::is_html_output(), fig.cap="`langevitour` output  of the `cone` data in $3\\text{-}D$. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex.", fig.pos="!ht"----
+## ----cone, eval=knitr::is_html_output(), fig.cap="Viewing the $3\\text{-}D$ output  of `gen_cone()`. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex.", fig.pos="!ht"----
 # 
 # conefig <- bscols(
 #   htmltools::div(
@@ -936,7 +936,7 @@ pyrholes_proj3 <- plot_proj(
 
 
 
[…]

### 2025-12-04 paper-cardinalR.Rmd
Updated mathematical notations

@@
 ```{r branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ output of the `gen_orgcurvybranches()`.
-The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"}
+The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"}
 
@@
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
-These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three 2-D projections of the 4-D orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"}
+These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the $4\\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"}
 
@@
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
-These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt="Three 2-D projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.", fig.pos="!ht", fig.width=12, fig.height=4}
+These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.", fig.pos="!ht", fig.width=12, fig.height=4}
 
@@
 The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
-`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.", fig.alt="Interactive langevitour of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing dense bases, triangular spreads, branching points, and hollow regions.", fig.pos="!ht"}
+`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.", fig.alt="Interactive langevitour of four $4\\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing dense bases, triangular spreads, branching points, and hollow regions.", fig.pos="!ht"}
 
@@
 `Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
-These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three 2-D projections from each of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.", fig.pos="!ht"}
+These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.", fig.pos="!ht"}
 
@@
 The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.alt="Interactive langevitour of three datasets: circle and curvycycle in 4-D showing closed-loop and wavy twisted-ring structures, and clusteredspheres in 3-D showing multiple compact spherical clusters with slight overlap in projections.", fig.pos="!ht", layout="l-body"}
+The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.alt="Interactive langevitour of three datasets: circle and curvycycle in $4\\text{-}D$ showing closed-loop and wavy twisted-ring structures, and clusteredspheres in 3-D showing multiple compact spherical clusters with slight overlap in projections.", fig.pos="!ht", layout="l-body"}
 
@@
 The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
-These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Three 2-D projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.", fig.pos="!ht"}
+These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Three $2\\text{-}D$ projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.", fig.pos="!ht"}
 
@@
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
-In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.", fig.alt="Interactive langevitour of two trefoil knot datasets: trefoil4d in 4-D showing complex twisting and looping patterns, and trefoil3d in 3-D showing a simpler compact knot-like form.", fig.pos="!ht"}
+In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.", fig.alt="Interactive langevitour of two trefoil knot datasets: trefoil4d in $4\\text{-}D$ showing complex twisting and looping patterns, and trefoil3d in 3-D showing a simpler compact knot-like form.", fig.pos="!ht"}
 
@@
 In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
-These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three 2-D projections from each of two trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.", fig.pos="!ht"}
+These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three $2\\text{-}D$ projections from each of two trefoil knot datasets, showing complex $4\\text{-}D$ trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.", fig.pos="!ht"}
 
@@
 The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.", fig.alt="Interactive langevitour of six datasets in 4-D: crescent, curvycylinder, sphericalspiral, helicalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral paths, conical helices, and warped 2-D surfaces with smooth and sharp variations.", fig.pos="!ht"}
+The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.", fig.alt="Interactive langevitour of six datasets in $4\\text{-}D$: crescent, curvycylinder, sphericalspiral, helicalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral paths, conical helices, and warped $2\\text{-}D$ surfaces with smooth and sharp variations.", fig.pos="!ht"}
 
@@
 The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
-Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Three 2-D projections from each of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity.", fig.pos="!ht"}
+Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped $2\\text{-}D$ surfaces, illustrating curvature, density, and continuity.", fig.pos="!ht"}
 
@@
 
-```{r highd-data, eval=knitr::is_html_output(), fig.cap="Viewing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in 4-D: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht"}
+```{r highd-data, eval=knitr::is_html_output(), fig.cap="Viewing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in $4\\text{-}D$: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht"}
 
@@
 ```{r highd-proj, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
-The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three 2-D projections of five 4-D synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=4, fig.pos="!ht"}
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three $2\\text{-}D$ projections of five $4\\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=4, fig.pos="!ht"}
 

### 2025-12-04 paper-cardinalR.tex
Updated mathematical notations

@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three 2-D projections of the 4-D orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `orglinearbranches` data.
+\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the $4\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `orglinearbranches` data.
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three 2-D projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ `cone` data.
+\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ `cone` data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Three 2-D projections from each of four 4-D pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.}]{paper-cardinalR_files/figure-latex/pyr-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Three $2\text{-}D$ projections from each of four $4\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.}]{paper-cardinalR_files/figure-latex/pyr-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Three 2-D projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Three $2\text{-}D$ projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Three 2-D projections from each of two trefoil knot datasets, showing complex 4-D trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.}]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={Three $2\text{-}D$ projections from each of two trefoil knot datasets, showing complex $4\text{-}D$ trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.}]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth,alt={Three 2-D projections from each of four 4-D datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped 2-D surfaces, illustrating curvature, density, and continuity.}]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
+{\centering \includegraphics[width=0.7\linewidth,alt={Three $2\text{-}D$ projections from each of four $4\text{-}D$ datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped $2\text{-}D$ surfaces, illustrating curvature, density, and continuity.}]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
 
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three 2-D projections of five 4-D synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.}]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
+\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of five $4\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.}]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
 The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:highd-proj}

### 2025-12-04 paper-cardinalR.R

@@ -406,7 +406,7 @@ orglinearbranches_proj3 <- plot_proj(
 orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 
 
-## ----branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ output of the `gen_orgcurvybranches()`. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in 4-D, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"----
+## ----branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ output of the `gen_orgcurvybranches()`. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"----
 # 
 # branchfig <- bscols(
 #   htmltools::div(
@@ -426,7 +426,7 @@ orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 # branchfig
 
 
-## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `orglinearbranches` data. Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles. These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three 2-D projections of the 4-D orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"----
[…]

### 2025-12-07 paper-cardinalR.Rmd
revised intro and abstract

@@
 abstract: >
-  High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association.
-It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold.
   Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
-These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets.
-By controlling the complexity researchers can better understand how different methods work and can be improved.
-A variety of functions to generate a large variety of structures in high dimensions are organized into the R package cardinalR, along with several sample sets.
-This enriches the existing toolset of benchmark datasets for evaluating algorithms.
+High-dimensional data is characterized by multiple variables that are dependent or associated in some way, such as linear, nonlinear, clustering or anomalies.
+Here we provide new methods for generating a variety of high-dimensional structures using mathematical functions and statistical distributionsorganized into the R package cardinalR.
+Several example data sets are also provided.
+These will be useful for researchers to better understand how different analytical methods work and can be improved, with a special focus on nonlinear dimension reduction methods.
+This package enriches the existing toolset of benchmark datasets for evaluating algorithms.
 draft: true
@@
 Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
-There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The `geozoo` package [@barret2016] offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-The package `snedata` [@james2025] provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding [@hinton2002] and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], to explore how well these techniques preserve data structures in lower dimensions.
-Additionally, `splatter` [@luke2017] is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
-In contrast, `mlbench` [@friedrich2024] includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.
+The `geozoo` package [@barret2016] provides functions to generate standard high-dimensional data like cubes, spheres and simplexes, along with some prepared datasets.
+The `snedata` package  [@james2025] provides functions for generating common examples used in dimension reduction publications and to download benchmark data sets.
+The `splatter` package [@luke2017] is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+The `mlbench` package [@friedrich2024] provides access to benchmark datasets commonly associated with established classification or regression challenges.
 The `surreal` package [@james2024] implements the "Residual (Sur)Realism" algorithm [@leonard2007] to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
@@
 
-These packages are especially valuable for specific applications, but together they are not sufficiently rich for simulating high-dimensional data for testing dimensionality reduction, supervised and unsupervised learning algorithms.
-To address this gap, this paper introduces the `cardinalR` R package.
-This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; `cardinalR` allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of NLDR methods, clustering algorithms, and visualization techniques.
-Moreover, these datasets can serve as benchmark examples for exploring how different algorithmic choices affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
+The current work implemented in the `cardinalR` R package builds on these approaches.
+It provides functions to generate a more extensive set of high-dimensional data structures, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine the shapes to produce multiple clusters.
+The user can control characteristics such as number of dimensions, shape and sample size.
+It is designed to resource researchers with synthetic datasets to evaluate the performance and interpret the fit of NLDR methods, clustering algorithms, and visualization techniques.
+These datasets can also serve as benchmark examples for exploring how different choices of algorithm parameters affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
 

### 2025-12-07 paper-cardinalR.tex
revised intro and abstract

@@
 \abstract{%
-High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association.
-It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold.
 Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
-These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets.
-By controlling the complexity researchers can better understand how different methods work and can be improved.
-A variety of functions to generate a large variety of structures in high dimensions are organized into the R package cardinalR, along with several sample sets.
-This enriches the existing toolset of benchmark datasets for evaluating algorithms.
+High-dimensional data is characterized by multiple variables that are dependent or associated in some way, such as linear, nonlinear, clustering or anomalies.
+Here we provide new methods for generating a variety of high-dimensional structures using mathematical functions and statistical distributionsorganized into the R package cardinalR.
+Several example data sets are also provided.
+These will be useful for researchers to better understand how different analytical methods work and can be improved, with a special focus on nonlinear dimension reduction methods.
+This package enriches the existing toolset of benchmark datasets for evaluating algorithms.
 }
@@
 Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
-There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.The \texttt{geozoo} package \citep{barret2016} offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces.
-The package \texttt{snedata} \citep{james2025} provides functions for generating simple simulation datasets for use in Stochastic Neighbor Embedding \citep{hinton2002} and related dimension reduction methods, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, to explore how well these techniques preserve data structures in lower dimensions.
-Additionally, \texttt{splatter} \citep{luke2017} is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
-In contrast, \texttt{mlbench} \citep{friedrich2024} includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges.
+There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.
+The \texttt{geozoo} package \citep{barret2016} provides functions to generate standard high-dimensional data like cubes, spheres and simplexes, along with some prepared datasets.
+The \texttt{snedata} package \citep{james2025} provides functions for generating common examples used in dimension reduction publications and to download benchmark data sets.
+The \texttt{splatter} package \citep{luke2017} is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
+The \texttt{mlbench} package \citep{friedrich2024} provides access to benchmark datasets commonly associated with established classification or regression challenges.
 The \texttt{surreal} package \citep{james2024} implements the ``Residual (Sur)Realism'' algorithm \citep{leonard2007} to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
 
-These packages are especially valuable for specific applications, but together they are not sufficiently rich for simulating high-dimensional data for testing dimensionality reduction, supervised and unsupervised learning algorithms.
-To address this gap, this paper introduces the \texttt{cardinalR} R package.
-This package provides a collection of functions designed to generate customizable data structures in any number of dimensions, starting from basic geometric shapes, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine high-dimensional datasets into a single multi-faceted, clustered dataset in a space of arbitrary dimension.
-By using clearly defined geometric shapes and controllable characteristics such as number of dimensions, sample size; \texttt{cardinalR} allows researchers to generate transparent and interpretable synthetic datasets useful for evaluating the performance of NLDR methods, clustering algorithms, and visualization techniques.
-Moreover, these datasets can serve as benchmark examples for exploring how different algorithmic choices affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
+The current work implemented in the \texttt{cardinalR} R package builds on these approaches.
+It provides functions to generate a more extensive set of high-dimensional data structures, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine the shapes to produce multiple clusters.
+The user can control characteristics such as number of dimensions, shape and sample size.
+It is designed to resource researchers with synthetic datasets to evaluate the performance and interpret the fit of NLDR methods, clustering algorithms, and visualization techniques.
+These datasets can also serve as benchmark examples for exploring how different choices of algorithm parameters affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
 

### 2025-12-08 paper-cardinalR.Rmd
Started fixing captions with code in PDF

@@
 
-```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `orglinearbranches` data.
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ data generated using \\texttt{gen\\_orgcurvybranches}.
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;",
-    circle_lang, curvycycle_lang, clusteredspheres_lang),
-  #widths=c(3,3,3),
+    style = "
+      display: grid; 
+      grid-template-columns: repeat(3, 1fr);
+      gap: 0;
+      width: 100%;
+      margin: 0 auto;
+      max-width: 100%;
+    ",
+    circle_lang,
+    curvycycle_lang,
+    clusteredspheres_lang
+  ),
   device = "xs"
@@
 
+htmltools::div(
+  style = "max-width: 100%; width: 100%; margin: 0 auto;",
   spherefig
+)
+
 ```

### 2025-12-08 paper-cardinalR.tex
Started fixing captions with code in PDF

@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the $4\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ `orglinearbranches` data.
+\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the $4\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ data generated using \texttt{gen\_orgcurvybranches}.
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.

### 2025-12-08 paper-cardinalR.R

@@ -426,7 +426,7 @@ orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 # branchfig
 
 
-## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ `orglinearbranches` data. Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles. These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the $4\\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"----
+## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ data generated using \\texttt{gen\\_orgcurvybranches}. Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles. These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the $4\\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"----
 
 orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
   plot_layout(ncol = 3, guides = "collect") 
@@ -1253,20 +1253,28 @@ clusteredspheres_proj3 <- plot_proj(
 # 
 # spherefig <- bscols(
 #   htmltools::div(
-#     style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
-#     gap: 0px;
-#     align-items: start;
-#     justify-items: center;
-#     margin: 0;
-#     padding: 0;",
-#     circle_lang, curvycycle_lang, clusteredspheres_lang),
-#   #widths=c(3,3,3),
+#     style = "
+#       display: grid;
+#       grid-template-columns: repeat(3, 1fr);
+#       gap: 0;
+#       width: 100%;
+#       margin: 0 auto;
+#       max-width: 100%;
+#     ",
+#     circle_lang,
+#     curvycycle_lang,
+#     clusteredspheres_lang
+#   ),
[…]

### 2025-12-08 paper-cardinalR.Rmd
Fixed the width of Figure5

--- 2025-12-07 paper-cardinalR.Rmd
@@
 ```{r}
-circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
+circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
 ```
@@
 ```{r}
-curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
+curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
 ```
@@
 ```{r}
-clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
+clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
 ```
@@
 Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
+The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
 
@@
 
-Compared with `sndata::swiss_roll()` [@james2025], this implementation (i) samples $t$ over $[0, 3\pi]$ instead of $[1.5\pi, 4.5\pi]$, (ii) allows a flexible vertical range $w = (w_1, w_2)$ rather than fixing $z \in [0, z_{\max}]$, and (iii) returns a tibble with `x1, x2, x3` instead of adding a color variable.
+Compared with `snedata::swiss_roll()` [@james2025], this implementation (i) samples $t$ over $[0, 3\pi]$ instead of $[1.5\pi, 4.5\pi]$, (ii) allows a flexible vertical range $w = (w_1, w_2)$ rather than fixing $z \in [0, z_{\max}]$, and (iii) returns a tibble with `x1, x2, x3` instead of adding a color variable.
 

### 2025-12-08 paper-cardinalR.tex
Fixed the width of Figure5

@@
 Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
+The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
 
@@
 
-Compared with \texttt{sndata::swiss\_roll()} \citep{james2025}, this implementation (i) samples \(t\) over \([0, 3\pi]\) instead of \([1.5\pi, 4.5\pi]\), (ii) allows a flexible vertical range \(w = (w_1, w_2)\) rather than fixing \(z \in [0, z_{\max}]\), and (iii) returns a tibble with \texttt{x1,\ x2,\ x3} instead of adding a color variable.
+Compared with \texttt{snedata::swiss\_roll()} \citep{james2025}, this implementation (i) samples \(t\) over \([0, 3\pi]\) instead of \([1.5\pi, 4.5\pi]\), (ii) allows a flexible vertical range \(w = (w_1, w_2)\) rather than fixing \(z \in [0, z_{\max}]\), and (iii) returns a tibble with \texttt{x1,\ x2,\ x3} instead of adding a color variable.
 

### 2025-12-08 paper-cardinalR.R

@@ -974,7 +974,7 @@ circle <- gen_circle(n = 1000, p = 4)
 
 
 ## -----------------------------------------------------------------------------
-circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
+circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
 
 
 ## ----circle-proj1-------------------------------------------------------------
@@ -1064,7 +1064,7 @@ curvycycle <- gen_curvycycle(n = 1000, p = 4)
 
 
 ## -----------------------------------------------------------------------------
-curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
+curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
 
 
 ## ----curvycycle-proj1---------------------------------------------------------
@@ -1164,7 +1164,7 @@ clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, r = c(15, 3),
 
 
 ## -----------------------------------------------------------------------------
-clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "300px", height = "300px")
+clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
 
 
 ## ----clusteredspheres-proj1---------------------------------------------------
@@ -1253,28 +1253,20 @@ clusteredspheres_proj3 <- plot_proj(
 # 
 # spherefig <- bscols(
 #   htmltools::div(
-#     style = "
-#       display: grid;
-#       grid-template-columns: repeat(3, 1fr);
-#       gap: 0;
-#       width: 100%;
-#       margin: 0 auto;
-#       max-width: 100%;
-#     ",
-#     circle_lang,
-#     curvycycle_lang,
-#     clusteredspheres_lang
-#   ),
+#     style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
+#     gap: 0px;
+#     align-items: start;
+#     justify-items: center;
+#     margin: 0;
+#     padding: 0;",
+#     circle_lang, curvycycle_lang, clusteredspheres_lang),
+#   #widths=c(3,3,3),
[…]

### 2025-12-08 paper-cardinalR.Rmd
Fixed the caption issue and figure width

@@
 
-```{r branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ output of the `gen_orgcurvybranches()`.
+```{r branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `orgcurvybranches` data.
 The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"}
@@
 
-```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ data generated using \\texttt{gen\\_orgcurvybranches}.
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ \\texttt{orgcurvybranches} data.
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
@@
 
-```{r cone, eval=knitr::is_html_output(), fig.cap="Viewing the $3\\text{-}D$ output  of `gen_cone()`.
+```{r cone, eval=knitr::is_html_output(), fig.cap="Viewing the $3\\text{-}D$ output  of `cone` data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex.", fig.pos="!ht"}
@@
 
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ `cone` data.
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
@@
 
-```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.
-The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
-`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
+```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the \\texttt{pyrrect} (a1-a3), \\texttt{pyrtri} (b1-b3), \\texttt{pyrstar} (c1-c3), and \\texttt{pyrholes} (d1-d3) data.
+The \\texttt{pyrrrect} structure forms a dense rectangular base tapering to a narrow tip, while \\texttt{pytri} shows a more triangular spread with sharper edges.
+\\texttt{pyrstar} extends into multiple pointed branches radiating from a common core, and \\texttt{pyrholes} reveals hollow or open regions within an otherwise compact shape.
 These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.", fig.pos="!ht"}
@@
 Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
+The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
 
@@
 
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, `circle` (a1-a3), `curvycycle` (b1-b3), and, $3\\text{-}D$ `clusteredspheres` (c1-c3).
-The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
+```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, \\texttt{circle} (a1-a3), \\texttt{curvycycle} (b1-b3), and, $3\\text{-}D$ \\texttt{clusteredspheres} (c1-c3).
+The \\texttt{circle} structure forms a smooth, closed loop, while \\texttt{curvycycle} shows a wavy, continuous pattern forming a twisted ring.
+The \\texttt{clusteredspheres} dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
 These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Three $2\\text{-}D$ projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.", fig.pos="!ht"}
@@
 
-```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, `trefoil4d` (a1-a3) and $3\\text{-}D$ `trefoil3d` (b1-b3) data.
-The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
-In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
+```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, \\texttt{trefoil4d} (a1-a3) and $3\\text{-}D$ \\texttt{trefoil3d} (b1-b3) data.
+The \\texttt{trefoil4d} structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
+In contrast, the \\texttt{trefoil3d} dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
 These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three $2\\text{-}D$ projections from each of two trefoil knot datasets, showing complex $4\\text{-}D$ trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.", fig.pos="!ht"}
@@
 
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.
-The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
-The `sphericalspiral` forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
-The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
+```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the \\texttt{curvycylinder} (a1-a3), \\texttt{sphericalspiral} (b1-b3), \\texttt{conicspiral} (c1-c3), and \\texttt{nonlinear} (d1-d3) data.
+The \\texttt{curvycylinder} shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
+The \\texttt{sphericalspiral} forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
+The \\texttt{conicspiral} spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
+The \\texttt{nonlinear} dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
 Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped $2\\text{-}D$ surfaces, illustrating curvature, density, and continuity.", fig.pos="!ht"}

### 2025-12-08 paper-cardinalR.tex
Fixed the caption issue and figure width

@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the $4\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ data generated using \texttt{gen\_orgcurvybranches}.
+\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the $4\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ \texttt{orgcurvybranches} data.
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ `cone` data.
+\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ \texttt{cone} data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `pyrrect` (a1-a3), `pyrtri` (b1-b3), `pyrstar` (c1-c3), and `pyrholes` (d1-d3) data.
-The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
-`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the \texttt{pyrrect} (a1-a3), \texttt{pyrtri} (b1-b3), \texttt{pyrstar} (c1-c3), and \texttt{pyrholes} (d1-d3) data.
+The \texttt{pyrrrect} structure forms a dense rectangular base tapering to a narrow tip, while \texttt{pytri} shows a more triangular spread with sharper edges.
+\texttt{pyrstar} extends into multiple pointed branches radiating from a common core, and \texttt{pyrholes} reveals hollow or open regions within an otherwise compact shape.
 These projections illustrate a range of pyramid-like geometries that vary in density and structure.}\label{fig:pyr-proj}
@@
 Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
+The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
 
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, `circle` (a1-a3), `curvycycle` (b1-b3), and, $3\text{-}D$ `clusteredspheres` (c1-c3).
-The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\text{-}D$ projections, highlighting how projection can distort spatial relationships.
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, \texttt{circle} (a1-a3), \texttt{curvycycle} (b1-b3), and, $3\text{-}D$ \texttt{clusteredspheres} (c1-c3).
+The \texttt{circle} structure forms a smooth, closed loop, while \texttt{curvycycle} shows a wavy, continuous pattern forming a twisted ring.
+The \texttt{clusteredspheres} dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\text{-}D$ projections, highlighting how projection can distort spatial relationships.
 These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\text{-}D$, emphasizing the effects of projection on density, continuity, and separation}\label{fig:sphere-proj}
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, `trefoil4d` (a1-a3) and $3\text{-}D$ `trefoil3d` (b1-b3) data.
-The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
-In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, \texttt{trefoil4d} (a1-a3) and $3\text{-}D$ \texttt{trefoil3d} (b1-b3) data.
+The \texttt{trefoil4d} structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
+In contrast, the \texttt{trefoil3d} dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
 These projections illustrate a range of looping structures in high-dimensions.}\label{fig:trefoil-proj}
@@
 
-\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the `curvycylinder` (a1-a3), `sphericalspiral` (b1-b3), `conicspiral` (c1-c3), and `nonlinear` (d1-d3) data.
-The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
-The `sphericalspiral` forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
-The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped $2\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
+\caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the \texttt{curvycylinder} (a1-a3), \texttt{sphericalspiral} (b1-b3), \texttt{conicspiral} (c1-c3), and \texttt{nonlinear} (d1-d3) data.
+The \texttt{curvycylinder} shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
+The \texttt{sphericalspiral} forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
+The \texttt{conicspiral} spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
+The \texttt{nonlinear} dataset exhibits a warped $2\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
 Each shows variations in curvature, density, and continuity.}\label{fig:triginometric-proj}

### 2025-12-08 paper-cardinalR.R

@@ -406,7 +406,7 @@ orglinearbranches_proj3 <- plot_proj(
 orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 
 
-## ----branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ output of the `gen_orgcurvybranches()`. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"----
+## ----branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `orgcurvybranches` data. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"----
 # 
 # branchfig <- bscols(
 #   htmltools::div(
@@ -426,7 +426,7 @@ orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 # branchfig
 
 
-## ----branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ data generated using \\texttt{gen\\_orgcurvybranches}. Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles. These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the $4\\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"----
[…]

### 2025-12-08 paper-cardinalR.Rmd
Added table caption

@@
 
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:branching-tb-html)', '\\@ref(tab:branching-tb-pdf)'))` outlines these functions.
-The main arguments of the functions described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:arg-branching-tb-html)', '\\@ref(tab:arg-branching-tb-pdf)'))`.
+Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:shape-tb-html)', '\\@ref(tab:shape-tb-pdf)'))` outlines these functions.
+The main arguments of the functions described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:arg-shape-tb-html)', '\\@ref(tab:arg-shape-tb-pdf)'))`.
 
 ```{r}
-branch_tb <- tibble(fun = c("gen_expbranches",
+shape_tb <- tibble(fun = c("gen_expbranches",
                              "gen_linearbranches",
@@
 
-```{r branching-tb-html, eval=knitr::is_html_output()}
-branch_tb |> 
+```{r shape-tb-html, eval=knitr::is_html_output()}
+shape_tb |> 
   mutate(parm = paste0("<code>", parm, "</code>")) |>
   mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, table.pos = "!ht") 
-```
-
-```{r branching-tb-pdf, eval=knitr::is_latex_output()}
-branch_tb |>
+  kable(caption = "Overview of shape-generation functions, including their required parameters and a brief description of each geometric structure produced.
+The generators cover branching patterns, spheres, spirals, pyramids, Gaussian clouds, and other nonlinear shapes.", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, table.pos = "!ht")
+```
+
+```{r shape-tb-pdf, eval=knitr::is_latex_output()}
+shape_tb |>
   mutate(parm = paste0("\\texttt{", parm, "}")) |>
   mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht", linesep = "")  |>
+  kable(caption = "Overview of shape-generation functions, including their required parameters and a brief description of each geometric structure produced.
+The generators cover branching patterns, spheres, spirals, pyramids, Gaussian clouds, and other nonlinear shapes.", format="latex", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht", linesep = "")  |>
   column_spec(1, width = "3.5cm") |>
@@
 
-```{r arg-branching-tb-html, eval=knitr::is_html_output()}
+```{r arg-shape-tb-html, eval=knitr::is_html_output()}
 arg_branching_tb |> 
@@
 
-```{r arg-branching-tb-pdf, eval=knitr::is_latex_output()}
+```{r arg-shape-tb-pdf, eval=knitr::is_latex_output()}
 arg_branching_tb |> 

### 2025-12-08 paper-cardinalR.tex
Added table caption

@@
 
-Table \ref{tab:branching-tb-pdf} outlines these functions. The main arguments of the functions described in Table \ref{tab:arg-branching-tb-pdf}.
+Table \ref{tab:shape-tb-pdf} outlines these functions. The main arguments of the functions described in Table \ref{tab:arg-shape-tb-pdf}.
 
@@
 
-\caption{\label{tab:branching-tb-pdf}cardinalR branching data generation functions}
+\caption{\label{tab:shape-tb-pdf}Overview of shape-generation functions, including their required parameters and a brief description of each geometric structure produced.
+The generators cover branching patterns, spheres, spirals, pyramids, Gaussian clouds, and other nonlinear shapes.}
 \centering
@@
 
-\caption{\label{tab:arg-branching-tb-pdf}The main arguments for branching shape generators.}
+\caption{\label{tab:arg-shape-tb-pdf}The main arguments for branching shape generators.}
 \centering

### 2025-12-08 paper-cardinalR.R

@@ -154,7 +154,7 @@ main_tb |>
 
 
 ## -----------------------------------------------------------------------------
-branch_tb <- tibble(fun = c("gen_expbranches",
+shape_tb <- tibble(fun = c("gen_expbranches",
                              "gen_linearbranches",
                              "gen_curvybranches", 
                              "gen_orglinearbranches", 
@@ -260,18 +260,18 @@ branch_tb <- tibble(fun = c("gen_expbranches",
    mutate(parm = gsub("_", "\\\\_", parm))
 
 
-## ----branching-tb-html, eval=knitr::is_html_output()--------------------------
-# branch_tb |>
+## ----shape-tb-html, eval=knitr::is_html_output()------------------------------
+# shape_tb |>
 #   mutate(parm = paste0("<code>", parm, "</code>")) |>
 #   mutate(fun = paste0("<code>", fun, "</code>")) |>
-#   kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, table.pos = "!ht")
+#   kable(caption = "Overview of shape-generation functions, including their required parameters and a brief description of each geometric structure produced. The generators cover branching patterns, spheres, spirals, pyramids, Gaussian clouds, and other nonlinear shapes.", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, table.pos = "!ht")
 
 
-## ----branching-tb-pdf, eval=knitr::is_latex_output()--------------------------
-branch_tb |>
+## ----shape-tb-pdf, eval=knitr::is_latex_output()------------------------------
+shape_tb |>
   mutate(parm = paste0("\\texttt{", parm, "}")) |>
   mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht", linesep = "")  |>
+  kable(caption = "Overview of shape-generation functions, including their required parameters and a brief description of each geometric structure produced. The generators cover branching patterns, spheres, spirals, pyramids, Gaussian clouds, and other nonlinear shapes.", format="latex", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht", linesep = "")  |>
[…]

### 2025-12-09 paper-cardinalR.Rmd
revised cluster benchmarking section

--- 2025-12-08 paper-cardinalR.Rmd
@@
 High-dimensional data is characterized by multiple variables that are dependent or associated in some way, such as linear, nonlinear, clustering or anomalies.
-Here we provide new methods for generating a variety of high-dimensional structures using mathematical functions and statistical distributionsorganized into the R package cardinalR.
+Here we provide new methods for generating a variety of high-dimensional structures using mathematical functions and statistical distributions organized into the R package cardinalR.
 Several example data sets are also provided.
@@
 set.seed(20240412)
-```
-
-```{r install-libraries, include=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
-
-# Ensure remotes is available
-if (!requireNamespace("remotes", quietly = TRUE)) {
-  install.packages("remotes")
-}
-
-# Read the package list
-pkgs_raw <- readLines("_Rpackages.txt")
-pkgs_raw <- pkgs_raw[nzchar(pkgs_raw)]  # remove empty lines
-
-# Split into package and version
-parts <- strsplit(pkgs_raw, "==")
-pkgs <- vapply(parts, `[[`, "", 1)
-versions <- vapply(parts, `[[`, "", 2)
-
-# Install packages with exact versions
-for (i in seq_along(pkgs)) {
-  pkg <- pkgs[i]
-  ver <- versions[i]
-
-  if (!requireNamespace(pkg, quietly = TRUE) ||
-      as.character(packageVersion(pkg)) != ver) {
-
-    message(sprintf("Installing %s (%s)", pkg, ver))
-    remotes::install_version(pkg, version = ver, upgrade = "never")
-  }
-}
-
 ```
@@
 # Extract high-dimensional data
-data <- five_clusts[, -5]
+# bad practice!
+#data <- five_clusts[, -5]
 true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))
@@
 # Hierarchical clustering
-dist_mat <- dist(data)
-hc_res <- cutree(hclust(dist_mat, method = "ward.D2"), k = 5)
+dist_mat <- dist(five_clusts[, -5])
+hc_fit <- hclust(dist_mat, method = "ward.D2")
+hc_cl <- NULL
+for (ncl in 2:10) {
+  hcl <- cutree(hc_fit, ncl)
+  hc_cl <- cbind(hc_cl, hcl)
+}
+colnames(hc_cl) <- paste0("cl", 2:10)
+
+x <- cluster.stats(dist_mat,
+  hc_cl[,1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
+                    "wb.ratio", "ch", "sindex")]
+hc_stats <- tibble(method="hc", cl=2) |>
+  bind_cols(as_tibble(x))
+for (i in 3:10) {
+  x <- cluster.stats(dist_mat,
+    hc_cl[,i-1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
+                    "wb.ratio", "ch", "sindex")]
+  hcms <- tibble(method="hc", cl=i) |>
+    bind_cols(as_tibble(x))
+  hc_stats <- hc_stats |>
+    bind_rows(hcms)
+}
 ```
@@
 # K-means clustering
-kmeans_res <- kmeans(data, centers = 5, nstart = 20)
-```
-
-```{r}
+kmeans_cl <- NULL
+for (ncl in 2:10) {
+  kcl <- kmeans(five_clusts[, -5], centers = ncl, nstart = 20)$cluster
+  kmeans_cl <- cbind(kmeans_cl, kcl)
+}
+colnames(kmeans_cl) <- paste0("cl", 2:10)
+
+x <- cluster.stats(dist_mat,
+  kmeans_cl[,1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
+                    "wb.ratio", "ch", "sindex")]
+kmeans_stats <- tibble(method="km", cl=2) |>
+  bind_cols(as_tibble(x))
+for (i in 3:10) {
+  x <- cluster.stats(dist_mat,
+    kmeans_cl[,i-1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
+                    "wb.ratio", "ch", "sindex")]
+  kms <- tibble(method="km", cl=i) |>
+    bind_cols(as_tibble(x))
+  kmeans_stats <- kmeans_stats |>
+    bind_rows(kms)
+}
+```
+
+```{r fig-cluster-stats, fig.width=10, fig.height=3, out.width="100%", layout="l-page"}
+# Examine the cluster stats
+all_stats <- bind_rows(hc_stats, kmeans_stats)
+all_stats |>
+  pivot_longer(within.cluster.ss:sindex, names_to = "stat", values_to = "value") |>
+  ggplot(aes(x=cl, y=value, colour=method)) +
+    geom_line() +
+    facet_wrap(~stat, ncol=3, scales="free_y") +
+    xlab("Number of clusters") +
+    ylab("")
+```
+
+@fig-cluster-stats shows a selection of cluster metrics for 2-10 clusters for each of the methods, $k$-means, hierarchical, ...
+As is typical, the suggestion of best solution varies between cluster statistics.
+There is some agreement that 4-5 clusters is better than other choices (say why), and that $k$-means performs better than others (this might not hold when you add the next methods).
+
+```{r eval=FALSE}
 # Model-based clustering
-mclust_res <- Mclust(data, G = 5, modelNames = "VVV")$classification
-```
-
-```{r}
+mclust_res <- Mclust(five_clusts[, -5], G = 5, modelNames = "VVV")$classification
+```
+
+```{r eval=FALSE}
 # Evaluate clustering performance
+# shouldn't use true labels here
 comp_kmeans <- cluster.stats(dist_mat, true_labels, kmeans_res$cluster)
@@
 
-```{r summaryclust-tb-html, eval=knitr::is_html_output()}
+```{r summaryclust-tb-html,  eval=FALSE}
 summary_clust |> 
@@
 
-```{r summaryclust-tb-pdf, eval=knitr::is_latex_output()}
+```{r summaryclust-tb-pdf,  eval=FALSE}
 summary_clust |> 

### 2025-12-09 paper-cardinalR.tex
revised cluster benchmarking section

--- 2025-12-08 paper-cardinalR.tex
@@
 
-\begin{table}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1}
 
-\caption{\label{tab:summaryclust-tb-pdf}Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}}
-\toprule
-method & wb.ratio & Dunn Index & Corrected Rand & VI\\
-\midrule
-k-means & 0.61 & 0.01 & 0.42 & 1.32\\
-Hierarchical & 0.61 & 0.01 & 0.50 & 1.15\\
-Model-based & 0.61 & 0.01 & 0.75 & 0.65\\
-\bottomrule
-\end{tabular}
-\end{table}
+\citet{fig-cluster-stats} shows a selection of cluster metrics for 2-10 clusters for each of the methods, \(k\)-means, hierarchical, \ldots{} As is typical, the suggestion of best solution varies between cluster statistics.
+There is some agreement that 4-5 clusters is better than other choices (say why), and that \(k\)-means performs better than others (this might not hold when you add the next methods).
 

### 2025-12-09 paper-cardinalR.Rmd
Rendered new version

@@
 
-## Generating high-dimensional clustered data
-
 To illustrate how high-dimensional clustered data can be generated using `cardinalR`, we generate a dataset with five clusters in $4\text{-}D$, each representing distinct geometric characteristics: a *helical spiral* (elongated and twisted), a *hemisphere* (curved surface), a *uniform cube* (isotropic distribution), a *cone* (density gradient), and a *Gaussian* cluster (compact and spherical) (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
@@
 
-```{r fig-cluster-stats, fig.width=10, fig.height=3, out.width="100%", layout="l-page"}
+```{r fig-cluster-stats, fig.width=10, fig.height=3, out.width="100%", layout="l-body"}
 # Examine the cluster stats

### 2025-12-11 paper-cardinalR.Rmd
Added scripts
Fixed the clustering figure

@@
 
+<!--scripts/five_clusts/02_gen_embeddings.R-->
 ```{r layouts}
@@
 
+<!--scripts/five_clusts/03_gen_rmse.R-->
 ```{r}
@@
     xlab("Number of clusters") +
-    ylab("")
+    ylab("") +
+    theme_minimal() +
+    theme(panel.border = element_rect(fill = 'transparent'),
+          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+          axis.ticks.x = element_line(),
+          axis.ticks.y = element_line(),
+          legend.position = "none",
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7),
+          axis.title.x = element_text(size = 7),
+          axis.title.y = element_text(size = 7),
+          plot.margin = margin(0, 0, 0, 0))
 ```

### 2025-12-11 paper-cardinalR.Rmd
Added sample for model-based clusters

--- 2025-12-09 paper-cardinalR.Rmd
@@
 
-```{r fig-cluster-stats, fig.width=10, fig.height=3, out.width="100%", layout="l-body"}
+```{r}
+# Model-based clustering
+modelbased_cl <- NULL
+for (ncl in 2:10) {
+  mcl <- Mclust(five_clusts[, -5], G = ncl, modelNames = "VVV")$classification
+  modelbased_cl <- cbind(modelbased_cl, mcl)
+}
+colnames(modelbased_cl) <- paste0("cl", 2:10)
+
+x <- cluster.stats(dist_mat,
+  modelbased_cl[,1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
+                    "wb.ratio", "ch", "sindex")]
+modelbased_stats <- tibble(method="mb", cl=2) |>
+  bind_cols(as_tibble(x))
+for (i in 3:10) {
+  x <- cluster.stats(dist_mat,
+    modelbased_cl[,i-1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
+                    "wb.ratio", "ch", "sindex")]
+  mbs <- tibble(method="mb", cl=i) |>
+    bind_cols(as_tibble(x))
+  modelbased_stats <- modelbased_stats |>
+    bind_rows(mbs)
+}
+```
+
+```{r, label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body"}
 # Examine the cluster stats
-all_stats <- bind_rows(hc_stats, kmeans_stats)
+all_stats <- bind_rows(hc_stats, kmeans_stats, modelbased_stats)
 all_stats |>
@@
     xlab("Number of clusters") +
-    ylab("")
-```
-
-@fig-cluster-stats shows a selection of cluster metrics for 2-10 clusters for each of the methods, $k$-means, hierarchical, ...
+    ylab("") +
+    theme_minimal() +
+    theme(panel.border = element_rect(fill = 'transparent'),
+          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+          axis.ticks.x = element_line(),
+          axis.ticks.y = element_line(),
+          legend.position = "none",
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7),
+          axis.title.x = element_text(size = 7),
+          axis.title.y = element_text(size = 7),
+          plot.margin = margin(0, 0, 0, 0))
+```
+
+@fig-cluster-stats shows a selection of cluster metrics for $2-10$ clusters for each of the methods, $k$-means, hierarchical, model-based.
 As is typical, the suggestion of best solution varies between cluster statistics.
-There is some agreement that 4-5 clusters is better than other choices (say why), and that $k$-means performs better than others (this might not hold when you add the next methods).
+There is some agreement that $4-5$ clusters is better than other choices (say why), and that $k$-means performs better than others (this might not hold when you add the next methods).
 

### 2025-12-11 paper-cardinalR.tex
Rendered new version
Added sample for model-based clusters

@@
 
-\subsection{Generating high-dimensional clustered data}\label{generating-high-dimensional-clustered-data}
-
 To illustrate how high-dimensional clustered data can be generated using \texttt{cardinalR}, we generate a dataset with five clusters in \(4\text{-}D\), each representing distinct geometric characteristics: a \emph{helical spiral} (elongated and twisted), a \emph{hemisphere} (curved surface), a \emph{uniform cube} (isotropic distribution), a \emph{cone} (density gradient), and a \emph{Gaussian} cluster (compact and spherical) (Figure \ref{fig:highd-proj}).
@@
 
-\citet{fig-cluster-stats} shows a selection of cluster metrics for 2-10 clusters for each of the methods, \(k\)-means, hierarchical, \ldots{} As is typical, the suggestion of best solution varies between cluster statistics.
-There is some agreement that 4-5 clusters is better than other choices (say why), and that \(k\)-means performs better than others (this might not hold when you add the next methods).
+\citet{fig-cluster-stats} shows a selection of cluster metrics for \(2-10\) clusters for each of the methods, \(k\)-means, hierarchical, model-based.
+As is typical, the suggestion of best solution varies between cluster statistics.
+There is some agreement that \(4-5\) clusters is better than other choices (say why), and that \(k\)-means performs better than others (this might not hold when you add the next methods).
 

### 2025-12-12 paper-cardinalR.Rmd
Added the best model-based model

@@
 for (ncl in 2:10) {
-  mcl <- Mclust(five_clusts[, -5], G = ncl, modelNames = "VVV")$classification
+  mcl <- Mclust(five_clusts[, -5], G = ncl, modelNames = "EII")$classification
   modelbased_cl <- cbind(modelbased_cl, mcl)
@@
           axis.ticks.y = element_line(),
-          legend.position = "none",
+          legend.position = "bottom",
           axis.text.x = element_text(size = 7),

### 2025-12-12 paper-cardinalR.R

@@ -2207,7 +2207,7 @@ for (i in 3:10) {
 # Model-based clustering
 modelbased_cl <- NULL
 for (ncl in 2:10) {
-  mcl <- Mclust(five_clusts[, -5], G = ncl, modelNames = "VVV")$classification
+  mcl <- Mclust(five_clusts[, -5], G = ncl, modelNames = "EII")$classification
   modelbased_cl <- cbind(modelbased_cl, mcl)
 }
 colnames(modelbased_cl) <- paste0("cl", 2:10)
@@ -2243,7 +2243,7 @@ all_stats |>
           plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
           axis.ticks.x = element_line(),
           axis.ticks.y = element_line(),
-          legend.position = "none",
+          legend.position = "bottom",
           axis.text.x = element_text(size = 7),
           axis.text.y = element_text(size = 7),
           axis.title.x = element_text(size = 7),

### 2025-12-12 paper-cardinalR.Rmd
Fixed the figure caption

@@
 
-```{r, label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body"}
+```{r, label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body", fig.cap="Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods."}
+
 # Examine the cluster stats
@@
     ylab("") +
+    scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3')) +
     theme_minimal() +
@@
 
-@fig-cluster-stats shows a selection of cluster metrics for $2-10$ clusters for each of the methods, $k$-means, hierarchical, model-based.
+Figure \@ref(fig:fig-cluster-stats) shows a selection of cluster metrics for $2-10$ clusters for each of the methods, $k$-means, hierarchical, model-based.
 As is typical, the suggestion of best solution varies between cluster statistics.
-There is some agreement that $4-5$ clusters is better than other choices (say why), and that $k$-means performs better than others (this might not hold when you add the next methods).
+There is some agreement that $4-5$ clusters is better than other choices (say why), and that $k$-means performs better than others.
 

### 2025-12-12 paper-cardinalR.tex
Fixed the figure caption

@@
 
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1}
+\begin{figure}[!]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.}\label{fig:fig-cluster-stats}
+\end{figure}
 
-\citet{fig-cluster-stats} shows a selection of cluster metrics for \(2-10\) clusters for each of the methods, \(k\)-means, hierarchical, model-based.
+Figure \ref{fig:fig-cluster-stats} shows a selection of cluster metrics for \(2-10\) clusters for each of the methods, \(k\)-means, hierarchical, model-based.
 As is typical, the suggestion of best solution varies between cluster statistics.
-There is some agreement that \(4-5\) clusters is better than other choices (say why), and that \(k\)-means performs better than others (this might not hold when you add the next methods).
+There is some agreement that \(4-5\) clusters is better than other choices (say why), and that \(k\)-means performs better than others.
 

### 2025-12-12 paper-cardinalR.R

@@ -2228,7 +2228,8 @@ for (i in 3:10) {
 }
 
 
-## ----label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body"----
+## ----label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body", fig.cap="Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods."----
+
 # Examine the cluster stats
 all_stats <- bind_rows(hc_stats, kmeans_stats, modelbased_stats)
 all_stats |>
@@ -2238,6 +2239,7 @@ all_stats |>
     facet_wrap(~stat, ncol=3, scales="free_y") +
     xlab("Number of clusters") +
     ylab("") +
+    scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3')) +
     theme_minimal() +
     theme(panel.border = element_rect(fill = 'transparent'),
           plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),

### 2025-12-12 paper-cardinalR.Rmd
Added description of clusters stats

@@
 To further evaluate the structure of the generated data, we benchmarked three clustering algorithms: **$k$-means** [Chapter 20 of @boehmke2019], **hierarchical** [@murtagh2012], and **model-based clustering** [@chris2002; @scrucca2023] using the simulated dataset.
-The model-based clustering was performed with the `"VVV"` covariance structure, allowing each cluster to vary in volume, shape, and orientation.
+Model-based clustering performed the `"EII"` covariance structure.
+Under this parameterization, clusters are spherical with equal volume and equal shape, and no orientation parameter is estimated.
 Cluster validity statistics were computed using the `cluster.stats()` function from the `fpc` package [@christian2024].
@@
 
-Figure \@ref(fig:fig-cluster-stats) shows a selection of cluster metrics for $2-10$ clusters for each of the methods, $k$-means, hierarchical, model-based.
-As is typical, the suggestion of best solution varies between cluster statistics.
-There is some agreement that $4-5$ clusters is better than other choices (say why), and that $k$-means performs better than others.
-
-```{r eval=FALSE}
-# Model-based clustering
-mclust_res <- Mclust(five_clusts[, -5], G = 5, modelNames = "VVV")$classification
-```
-
-```{r eval=FALSE}
-# Evaluate clustering performance
-# shouldn't use true labels here
-comp_kmeans <- cluster.stats(dist_mat, true_labels, kmeans_res$cluster)
-comp_hclust <- cluster.stats(dist_mat, true_labels, hc_res)
-comp_mclust <- cluster.stats(dist_mat, true_labels, mclust_res)
-
-# Summarize key performance metrics
-summary_clust <- tibble::tibble(
-  Metric = c("k-means", "Hierarchical", "Model-based"),
-  `wb.ratio` = c(comp_kmeans$wb.ratio, comp_hclust$wb.ratio, comp_mclust$wb.ratio),
-  `Dunn Index` = c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn),
-  `Corrected rand` = c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand),
-  VI = c(comp_kmeans$vi, comp_hclust$vi, comp_mclust$vi)
-)
-
-summary_clust <- summary_clust |>
-  mutate(across(where(is.numeric), ~ round(.x, 2)))
-```
-
-```{r summaryclust-tb-html,  eval=FALSE}
-summary_clust |> 
-  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", col.names = c("method", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), tab.pos = "H")
-```
-
-```{r summaryclust-tb-pdf,  eval=FALSE}
-summary_clust |> 
-  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("method", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), booktabs = T, table.pos = "!ht")  |>
-  column_spec(1, width = "2.5cm") |>
-  column_spec(2, width = "2cm") |>
-  column_spec(3, width = "2cm") |>
-  column_spec(4, width = "2.5cm") |>
-  column_spec(5, width = "2cm")
-```
-
-Overall, all methods produced similar compactness and separation, as reflected by the *within–between cluster ratios (wb.ratio)* and *Dunn indices*.
-However, the **model-based clustering** achieved the highest *Corrected Rand Index* ($0.75$) and lowest *Variation of Information (VI)* ($0.65$), indicating the best recovery of the true underlying groups (Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:summaryclust-tb-html)', '\\@ref(tab:summaryclust-tb-pdf)'))`).
-In comparison, $k$-means and hierarchical clustering showed moderate agreement with the true labels.
-These findings demonstrate that mixture-based approaches can more effectively capture the heterogeneity of clusters in high-dimensional, non-spherical data.
+Figure \@ref(fig:fig-cluster-stats) shows a selection of cluster metrics for $2-10$ clusters for each of the methods, $k$-means, hierarchical, and model-based.
+As is typical, the suggestion of the best solution varies between cluster statistics.
+Although the metrics differ in their preferences, several show consistent support for a $4–5$ cluster solution.
+The Calinski–Harabasz index increases sharply up to four or five clusters before levelling off, the Dunn and Pearson gamma indices attain local maxima in this range, and both the silhouette width and within-cluster sum of squares show diminishing improvement beyond five clusters.
+Together, these patterns indicate that $4–5$ clusters provide a good balance of separation and compactness, with $k$-means performing slightly better than the other methods across most metrics.
 
@@
 
-<!-- - Branching: These functions create a controlled environment for testing how effectively various algorithms preserve branching topology and continuity in their low-dimensional embeddings. -->
-
-<!-- - Scurve with a hole allowing for evaluation of how well algorithms handle incomplete manifolds or missing local structure. -->
-
-<!-- - clusteredsphere: This structure allows for cluster separation on curved manifolds in high-dimensional space and can be used to test the ability of NLDR methods and clustering algorithms to detect spherical clusters of different sizes and separations.
--->
-
-<!-- - Mobius: The core geometric structure is a Mobius strip—a classic one-sided surface with a half-twist—useful for evaluating how well methods capture non-orientability and twisted manifolds. -->
-
-<!-- - Grided cube: This function is useful for assessing how algorithms preserve uniformly spaced data in high-dimensional spaces. -->
-
-<!-- - Pyrholes: This structure is useful for testing clustering and NLDR algorithms on non-convex and sparse high-dimensional shapes. -->
-
-<!-- The application of our high-dimensional data generation package to evaluate the interplay between dimensionality reduction, nuisance variables, and hierarchical clustering yielded several key insights.
-The ability to generate synthetic datasets with well-defined underlying structures, coupled with the controlled introduction of nuisance variables, provided a valuable platform for assessing the robustness of downstream unsupervised learning techniques.
--->
-
-<!-- Our findings demonstrated that the choice of dimensionality reduction method significantly impacted the ability of hierarchical clustering to recover the true underlying clusters.
-Methods that effectively preserved the global structure of the data, as defined by our generation process, generally led to more accurate and interpretable hierarchical clustering results.
-However, the presence of nuisance variables often confounded the low-dimensional embeddings, making it more challenging for hierarchical clustering to separate truly distinct groups.
-This highlights a critical consideration in real-world data analysis, where unmeasured or latent factors can obscure the signal of interest.
--->
-
-<!-- The hierarchical clustering analysis itself, and the choice of linkage criteria, also played a crucial role.
-Different linkage methods revealed varying degrees of sensitivity to the distorted representations caused by the nuisance variables.
-For instance, methods that prioritize compact clusters might have been more susceptible to being misled by variance introduced by nuisance factors, while others focusing on the distance between clusters might have shown more resilience.
-The dendrograms generated by hierarchical clustering provided a visual means to explore the relationships between samples and the potential influence of nuisance, although determining the optimal number of clusters remained a challenge in the presence of these confounding factors.
--->
-
-<!-- This application underscores the utility of our data generation package as a powerful tool for controlled experimentation in unsupervised learning.
-By providing the ground truth cluster assignments and the ability to systematically manipulate data characteristics like dimensionality, geometric structure, and the presence of nuisance variables, researchers can gain a deeper understanding of the strengths and limitations of various DR and clustering algorithms.
-This controlled environment allows for a more objective evaluation than often possible with real-world datasets where the underlying structure is unknown.
--->
-
-<!-- One limitation of this particular application was the specific type and magnitude of the nuisance variables introduced.
-Future work could explore a wider range of nuisance types (e.g., batch effects, technical noise with specific distributions) and their varying degrees of influence on different DR and clustering methodologies.
-Furthermore, investigating strategies for mitigating the impact of nuisance variables, either during the DR step or within the clustering process itself, would be a valuable extension.
--->
-
-<!-- In conclusion, this example demonstrates the critical role of synthetic data generation in dissecting the complex interactions within unsupervised learning pipelines.
-Our package provides a flexible and controlled means to create such data, enabling researchers to systematically evaluate the performance and robustness of dimensionality reduction and clustering algorithms under well-defined conditions, including the presence of confounding factors.
-This capability contributes to a more rigorous and informed approach to high-dimensional data analysis.
--->
-
 # Acknowledgements

### 2025-12-12 paper-cardinalR.tex
Added description of clusters stats

@@
 To further evaluate the structure of the generated data, we benchmarked three clustering algorithms: \textbf{\(k\)-means} \citep[Chapter 20 of][]{boehmke2019}, \textbf{hierarchical} \citep{murtagh2012}, and \textbf{model-based clustering} \citep{chris2002, scrucca2023} using the simulated dataset.
-The model-based clustering was performed with the \texttt{"VVV"} covariance structure, allowing each cluster to vary in volume, shape, and orientation.
+Model-based clustering performed the \texttt{"EII"} covariance structure.
+Under this parameterization, clusters are spherical with equal volume and equal shape, and no orientation parameter is estimated.
 Cluster validity statistics were computed using the \texttt{cluster.stats()} function from the \texttt{fpc} package \citep{christian2024}.
@@
 
-Figure \ref{fig:fig-cluster-stats} shows a selection of cluster metrics for \(2-10\) clusters for each of the methods, \(k\)-means, hierarchical, model-based.
-As is typical, the suggestion of best solution varies between cluster statistics.
-There is some agreement that \(4-5\) clusters is better than other choices (say why), and that \(k\)-means performs better than others.
-
-Overall, all methods produced similar compactness and separation, as reflected by the \emph{within--between cluster ratios (wb.ratio)} and \emph{Dunn indices}.
-However, the \textbf{model-based clustering} achieved the highest \emph{Corrected Rand Index} (\(0.75\)) and lowest \emph{Variation of Information (VI)} (\(0.65\)), indicating the best recovery of the true underlying groups (Table \ref{tab:summaryclust-tb-pdf}).
-In comparison, \(k\)-means and hierarchical clustering showed moderate agreement with the true labels.
-These findings demonstrate that mixture-based approaches can more effectively capture the heterogeneity of clusters in high-dimensional, non-spherical data.
+Figure \ref{fig:fig-cluster-stats} shows a selection of cluster metrics for \(2-10\) clusters for each of the methods, \(k\)-means, hierarchical, and model-based.
+As is typical, the suggestion of the best solution varies between cluster statistics.
+Although the metrics differ in their preferences, several show consistent support for a \(4–5\) cluster solution.
+The Calinski--Harabasz index increases sharply up to four or five clusters before levelling off, the Dunn and Pearson gamma indices attain local maxima in this range, and both the silhouette width and within-cluster sum of squares show diminishing improvement beyond five clusters.
+Together, these patterns indicate that \(4–5\) clusters provide a good balance of separation and compactness, with \(k\)-means performing slightly better than the other methods across most metrics.
 

### 2025-12-12 paper-cardinalR.Rmd
Put clustering stats in a new script (speed up the paper rendering)

@@
 library(mclust)
+library(htmltools)
 ```
@@
 
-```{r}
-# Extract high-dimensional data
-# bad practice!
-#data <- five_clusts[, -5]
-true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))
-
-# Hierarchical clustering
-dist_mat <- dist(five_clusts[, -5])
-hc_fit <- hclust(dist_mat, method = "ward.D2")
-hc_cl <- NULL
-for (ncl in 2:10) {
-  hcl <- cutree(hc_fit, ncl)
-  hc_cl <- cbind(hc_cl, hcl)
-}
-colnames(hc_cl) <- paste0("cl", 2:10)
-
-x <- cluster.stats(dist_mat,
-  hc_cl[,1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
-hc_stats <- tibble(method="hc", cl=2) |>
-  bind_cols(as_tibble(x))
-for (i in 3:10) {
-  x <- cluster.stats(dist_mat,
-    hc_cl[,i-1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
-  hcms <- tibble(method="hc", cl=i) |>
-    bind_cols(as_tibble(x))
-  hc_stats <- hc_stats |>
-    bind_rows(hcms)
-}
-```
-
-```{r}
-# K-means clustering
-kmeans_cl <- NULL
-for (ncl in 2:10) {
-  kcl <- kmeans(five_clusts[, -5], centers = ncl, nstart = 20)$cluster
-  kmeans_cl <- cbind(kmeans_cl, kcl)
-}
-colnames(kmeans_cl) <- paste0("cl", 2:10)
-
-x <- cluster.stats(dist_mat,
-  kmeans_cl[,1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
-kmeans_stats <- tibble(method="km", cl=2) |>
-  bind_cols(as_tibble(x))
-for (i in 3:10) {
-  x <- cluster.stats(dist_mat,
-    kmeans_cl[,i-1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
-  kms <- tibble(method="km", cl=i) |>
-    bind_cols(as_tibble(x))
-  kmeans_stats <- kmeans_stats |>
-    bind_rows(kms)
-}
-```
-
-```{r}
-# Model-based clustering
-modelbased_cl <- NULL
-for (ncl in 2:10) {
-  mcl <- Mclust(five_clusts[, -5], G = ncl, modelNames = "EII")$classification
-  modelbased_cl <- cbind(modelbased_cl, mcl)
-}
-colnames(modelbased_cl) <- paste0("cl", 2:10)
-
-x <- cluster.stats(dist_mat,
-  modelbased_cl[,1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
-modelbased_stats <- tibble(method="mb", cl=2) |>
-  bind_cols(as_tibble(x))
-for (i in 3:10) {
-  x <- cluster.stats(dist_mat,
-    modelbased_cl[,i-1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
-  mbs <- tibble(method="mb", cl=i) |>
-    bind_cols(as_tibble(x))
-  modelbased_stats <- modelbased_stats |>
-    bind_rows(mbs)
-}
-```
-
-```{r, label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body", fig.cap="Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods."}
+<!--script/05_gen_cluster_stat.R-->
+```{r, label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
+Several indices consistently suggest that $4–5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics."}
 
 # Examine the cluster stats
-all_stats <- bind_rows(hc_stats, kmeans_stats, modelbased_stats)
+all_stats <- read_rds("data/five_clusts/cluster_stat.rds")
 all_stats |>
@@
 As is typical, the suggestion of the best solution varies between cluster statistics.
-Although the metrics differ in their preferences, several show consistent support for a $4–5$ cluster solution.
+Although the metrics differ in their preferences, several show consistent support for a $4-5$ cluster solution.
 The Calinski–Harabasz index increases sharply up to four or five clusters before levelling off, the Dunn and Pearson gamma indices attain local maxima in this range, and both the silhouette width and within-cluster sum of squares show diminishing improvement beyond five clusters.
-Together, these patterns indicate that $4–5$ clusters provide a good balance of separation and compactness, with $k$-means performing slightly better than the other methods across most metrics.
+Together, these patterns indicate that $4-5$ clusters provide a good balance of separation and compactness, with $k$-means performing slightly better than the other methods across most metrics.
+
+```{r}
+## k-means with  four clusters
+kcl4 <- kmeans(five_clusts[, -5], centers = 4, nstart = 20)$cluster
+
+## k-means with  five clusters
+kcl5 <- kmeans(five_clusts[, -5], centers = 5, nstart = 20)$cluster
+```
+
+```{r highd-data-clusters-algo-html, eval=knitr::is_html_output(), fig.cap="Viewing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in $4\\text{-}D$: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht", layout = "l-body"}
+
+four_km_langevitour <- langevitour::langevitour(
+  five_clusts[, -5],
+  group = kcl4,
+  levelColors = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e'),
+  width = "400px",
+  height = "300px"
+)
+
+five_km_langevitour <- langevitour::langevitour(
+  five_clusts[, -5],
+  group = kcl5,
+  levelColors = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e'),
+  width = "400px",
+  height = "300px"
+)
+
+km_results <- browsable(
+  div(
+    style = "
+      display: flex;
+      justify-content: space-between;
+      gap: 10px;
+    ",
+    div(style = "flex: 1;", four_km_langevitour),
+    div(style = "flex: 1;", five_km_langevitour)
+  )
+)
+
+class(km_results) <- c(class(km_results), "htmlwidget")
+
+km_results
+```
+
+```{r five-4clusts-projections1}
+scaled_data <- scale_data_manual(five_clusts[, -5])
+
+## First projection
+projection <- cbind(
+  c(-0.59692,0.46414,0.42436,-0.10048),
+  c(-0.54999,-0.67673,-0.03631,-0.01193))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.05))
+
+proj_obj1[["cluster"]] <- as.character(kcl4)
+
+five_clusts_proj1_1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
+
+```
+
+```{r five-4clusts-projections2}
+## Second projection
+projection <- cbind(
+  c(0.34673,-0.35774,0.66272,0.27298),
+  c(0.24023,0.01170,-0.41886,0.72707))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.07))
+
+proj_obj2[["cluster"]] <- as.character(kcl4)
+
+five_clusts_proj2_1 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1.2), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
+```
+
+```{r five-4clusts-projections3}
+## Third projection
+projection <- cbind(
+  c(0.47388,-0.20984,-0.56560,0.41644),
+  c(-0.51124,-0.60184,0.06463,0.36627))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.05))
+
+proj_obj3[["cluster"]] <- as.character(kcl4)
+
+five_clusts_proj3_1 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1.1), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
+
+```
+
+```{r five-5clusts-projections1}
+scaled_data <- scale_data_manual(five_clusts[, -5])
+
+## First projection
+projection <- cbind(
+  c(-0.59692,0.46414,0.42436,-0.10048),
+  c(-0.54999,-0.67673,-0.03631,-0.01193))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.05))
+
+proj_obj1[["cluster"]] <- as.character(kcl5)
+
+five_clusts_proj1_2 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
+
+```
+
+```{r five-5clusts-projections2}
+## Second projection
+projection <- cbind(
+  c(0.34673,-0.35774,0.66272,0.27298),
+  c(0.24023,0.01170,-0.41886,0.72707))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.07))
+
+proj_obj2[["cluster"]] <- as.character(kcl5)
+
+five_clusts_proj2_2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1.2), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
+```
+
+```{r five-5clusts-projections3}
+## Third projection
+projection <- cbind(
+  c(0.47388,-0.20984,-0.56560,0.41644),
+  c(-0.51124,-0.60184,0.06463,0.36627))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.9, 
+                                              axis_pos_y = -0.9, 
+                                              threshold = 0.05))
+
+proj_obj3[["cluster"]] <- as.character(kcl5)
+
+five_clusts_proj3_2 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.1, 1.1), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
+
+```
+
+```{r highd-proj-clust-algo-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three $2\\text{-}D$ projections of five $4\\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=8, fig.pos="!ht"}
+
+five_clusts_proj1_1 + five_clusts_proj2_1 + five_clusts_proj3_1 +
+  five_clusts_proj1_2 + five_clusts_proj2_2 + five_clusts_proj3_2 +
+  plot_layout(ncol = 3) 
+```
+
+The Langevitour displays of the four- and five-cluster $k$-means solutions further clarify why these partitions do not fully recover the underlying structure.
+In both visualizations, the estimated clusters show noticeable overlap, with colours repeatedly blending as the tour moves—evidence that points from different true groups are assigned to the same cluster.
+The $4$-cluster solution shows large, diffuse regions where several true clusters are merged, while the $5$-cluster solution partially separates some structures but still misclassifies substantial portions of multiple groups.
+These patterns reflect the mismatch between the nonlinear shapes present in the data and the spherical decision boundaries imposed by $k$-means.
+Even though the cluster validity indices favour a $4–5$ cluster solution overall, the dynamic projections reveal that neither solution provides clean separation, highlighting the importance of visual diagnostics when interpreting clustering results.
 

### 2025-12-12 paper-cardinalR.tex
Put clustering stats in a new script (speed up the paper rendering)

@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.}\label{fig:fig-cluster-stats}
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
+Several indices consistently suggest that $4–5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics.}\label{fig:fig-cluster-stats}
 \end{figure}
@@
 As is typical, the suggestion of the best solution varies between cluster statistics.
-Although the metrics differ in their preferences, several show consistent support for a \(4–5\) cluster solution.
+Although the metrics differ in their preferences, several show consistent support for a \(4-5\) cluster solution.
 The Calinski--Harabasz index increases sharply up to four or five clusters before levelling off, the Dunn and Pearson gamma indices attain local maxima in this range, and both the silhouette width and within-cluster sum of squares show diminishing improvement beyond five clusters.
-Together, these patterns indicate that \(4–5\) clusters provide a good balance of separation and compactness, with \(k\)-means performing slightly better than the other methods across most metrics.
+Together, these patterns indicate that \(4-5\) clusters provide a good balance of separation and compactness, with \(k\)-means performing slightly better than the other methods across most metrics.
+
+\begin{figure}[!ht]
+\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of five $4\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.}]{paper-cardinalR_files/figure-latex/highd-proj-clust-algo-pdf-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:highd-proj-clust-algo-pdf}
+\end{figure}
+
+The Langevitour displays of the four- and five-cluster \(k\)-means solutions further clarify why these partitions do not fully recover the underlying structure.
+In both visualizations, the estimated clusters show noticeable overlap, with colours repeatedly blending as the tour moves---evidence that points from different true groups are assigned to the same cluster.
+The \(4\)-cluster solution shows large, diffuse regions where several true clusters are merged, while the \(5\)-cluster solution partially separates some structures but still misclassifies substantial portions of multiple groups.
+These patterns reflect the mismatch between the nonlinear shapes present in the data and the spherical decision boundaries imposed by \(k\)-means.
+Even though the cluster validity indices favour a \(4–5\) cluster solution overall, the dynamic projections reveal that neither solution provides clean separation, highlighting the importance of visual diagnostics when interpreting clustering results.
 

### 2025-12-12 paper-cardinalR.R

@@ -27,6 +27,7 @@ library(colorspace)
 library(crosstalk)
 library(fpc)
 library(mclust)
+library(htmltools)
 
 
 ## -----------------------------------------------------------------------------
@@ -2146,92 +2147,10 @@ error_plot_five_clust + wrap_plots(nldr1, nldr2, nldr3,
                                     nldr4, nldr5, nldr6, ncol = 2)
 
 
-## -----------------------------------------------------------------------------
-# Extract high-dimensional data
-# bad practice!
-#data <- five_clusts[, -5]
-true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))
-
-# Hierarchical clustering
-dist_mat <- dist(five_clusts[, -5])
-hc_fit <- hclust(dist_mat, method = "ward.D2")
-hc_cl <- NULL
-for (ncl in 2:10) {
-  hcl <- cutree(hc_fit, ncl)
-  hc_cl <- cbind(hc_cl, hcl)
-}
-colnames(hc_cl) <- paste0("cl", 2:10)
-
-x <- cluster.stats(dist_mat,
-  hc_cl[,1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
-hc_stats <- tibble(method="hc", cl=2) |>
-  bind_cols(as_tibble(x))
-for (i in 3:10) {
-  x <- cluster.stats(dist_mat,
-    hc_cl[,i-1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
-  hcms <- tibble(method="hc", cl=i) |>
-    bind_cols(as_tibble(x))
-  hc_stats <- hc_stats |>
-    bind_rows(hcms)
-}
-
-
-## -----------------------------------------------------------------------------
-# K-means clustering
-kmeans_cl <- NULL
-for (ncl in 2:10) {
-  kcl <- kmeans(five_clusts[, -5], centers = ncl, nstart = 20)$cluster
-  kmeans_cl <- cbind(kmeans_cl, kcl)
-}
-colnames(kmeans_cl) <- paste0("cl", 2:10)
-
-x <- cluster.stats(dist_mat,
-  kmeans_cl[,1])[c("within.cluster.ss", "pearsongamma", "dunn2", 
-                    "wb.ratio", "ch", "sindex")]
[…]

### 2025-12-12 paper-cardinalR.Rmd
Updated figure titles

@@
 
-```{r highd-data-clusters-algo-html, eval=knitr::is_html_output(), fig.cap="Viewing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in $4\\text{-}D$: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht", layout = "l-body"}
+```{r highd-data-clusters-algo-html, eval=knitr::is_html_output(), fig.cap="Views of the five-cluster synthetic dataset coloured by the $k$-means four- and five-cluster solutions.
+These show how the five true structures: helical spiral, hemisphere, uniform cube, blunted cone, and Gaussian-shaped cluster appear under $k$-means partitions.
+Both the $4$- and $5$-cluster solutions mix colours, revealing substantial misclassification and limited ability of $k$-means to recover the underlying geometric structure.", fig.pos="!ht", layout = "l-body"}
 
@@
   levelColors = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e'),
-  width = "400px",
+  width = "350px",
   height = "300px"
@@
       justify-content: space-between;
-      gap: 10px;
+      gap: 5px;
     ",
@@
 
-```{r highd-proj-clust-algo-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
-The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three $2\\text{-}D$ projections of five $4\\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=8, fig.pos="!ht"}
+```{r highd-proj-clust-algo-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data coloured by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
+The intermixing of colours within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.", fig.width=12, fig.height=8, fig.pos="!ht"}
 
@@
 
-The Langevitour displays of the four- and five-cluster $k$-means solutions further clarify why these partitions do not fully recover the underlying structure.
-In both visualizations, the estimated clusters show noticeable overlap, with colours repeatedly blending as the tour moves—evidence that points from different true groups are assigned to the same cluster.
-The $4$-cluster solution shows large, diffuse regions where several true clusters are merged, while the $5$-cluster solution partially separates some structures but still misclassifies substantial portions of multiple groups.
-These patterns reflect the mismatch between the nonlinear shapes present in the data and the spherical decision boundaries imposed by $k$-means.
-Even though the cluster validity indices favour a $4–5$ cluster solution overall, the dynamic projections reveal that neither solution provides clean separation, highlighting the importance of visual diagnostics when interpreting clustering results.
+`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` shows the four- and five-cluster $k$-means solutions further clarify why these partitions do not fully identify the underlying structure.
+The predicted clusters show noticeable overlap, with colours repeatedly blending as the tour moves; evidence that points from different true clusters are assigned to the same cluster.
+The $4$-cluster solution shows large, diffuse regions where several true clusters are merged, while the $5$-cluster solution partially separates some structures but still misclassifies substantial portions of multiple clusters.
+These patterns reflect the mismatch between the helical spherical shape present in the data and the spherical decision boundaries imposed by $k$-means.
+Even though the cluster statistics favour a $4–5$ cluster solution overall, `r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` reveals that neither solution provides clear separation, highlighting the importance of visual diagnostics when interpreting clustering results.
 

### 2025-12-12 paper-cardinalR.tex
Updated figure titles

--- 2025-12-12 paper-cardinalR.tex
@@
 As is typical, the suggestion of the best solution varies between cluster statistics.
-Although the metrics differ in their preferences, several show consistent support for a \(4–5\) cluster solution.
+Although the metrics differ in their preferences, several show consistent support for a \(4-5\) cluster solution.
 The Calinski--Harabasz index increases sharply up to four or five clusters before levelling off, the Dunn and Pearson gamma indices attain local maxima in this range, and both the silhouette width and within-cluster sum of squares show diminishing improvement beyond five clusters.
-Together, these patterns indicate that \(4–5\) clusters provide a good balance of separation and compactness, with \(k\)-means performing slightly better than the other methods across most metrics.
+Together, these patterns indicate that \(4-5\) clusters provide a good balance of separation and compactness, with \(k\)-means performing slightly better than the other methods across most metrics.
+
+\begin{figure}[!ht]
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/highd-proj-clust-algo-pdf-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data coloured by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
+The intermixing of colours within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.}\label{fig:highd-proj-clust-algo-pdf}
+\end{figure}
+
+Figure \ref{fig:highd-proj-clust-algo-pdf} shows the four- and five-cluster \(k\)-means solutions further clarify why these partitions do not fully identify the underlying structure.
+The predicted clusters show noticeable overlap, with colours repeatedly blending as the tour moves; evidence that points from different true clusters are assigned to the same cluster.
+The \(4\)-cluster solution shows large, diffuse regions where several true clusters are merged, while the \(5\)-cluster solution partially separates some structures but still misclassifies substantial portions of multiple clusters.
+These patterns reflect the mismatch between the helical spherical shape present in the data and the spherical decision boundaries imposed by \(k\)-means.
+Even though the cluster statistics favour a \(4–5\) cluster solution overall, Figure \ref{fig:highd-proj-clust-algo-pdf} reveals that neither solution provides clear separation, highlighting the importance of visual diagnostics when interpreting clustering results.
 

### 2025-12-12 paper-cardinalR.Rmd
Updated additional function link

@@
 This flexibility allows users to construct both low- and high-dimensional datasets from the same underlying structures.
-
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:shape-tb-html)', '\\@ref(tab:shape-tb-pdf)'))` outlines these functions.
@@
 ```{r}
-arg_branching_tb <- tibble(arg = c("n", "k", "p"), 
-                          type = c("integer", "integer", "integer"),
+arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "k", "loc", "w", "steps", "spins", "hc", "non_fac"), 
+                          type = c("integer", "integer", "integer", "",
+                                   "", "", "", "", "", "", "", "", "", ""),
                         exp = c("Number of points.",
                                 "Number of clusters.",
-                                "Number of dimensions."))
+                                "Number of dimensions.",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "",
+                                "")) |> 
+   mutate(arg = gsub("_", "\\\\_", arg))
 ```
@@
 Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
+The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
 
@@
 Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:add-tb-html)', '\\@ref(tab:add-tb-pdf)'))` details these functions.
-More detailed explanations are available in <https://jayanilakshika.github.io/cardinalR/articles/03additionalfun.html>.
+More detailed explanations are available in [jayanilakshika.github.io/cardinalR/articles/03additionalfun](https://jayanilakshika.github.io/cardinalR/articles/03additionalfun.html).
 

### 2025-12-12 paper-cardinalR.tex
Updated additional function link

@@
 This flexibility allows users to construct both low- and high-dimensional datasets from the same underlying structures.
-
-Table \ref{tab:shape-tb-pdf} outlines these functions. The main arguments of the functions described in Table \ref{tab:arg-shape-tb-pdf}.
+Table \ref{tab:shape-tb-pdf} outlines these functions.
+The main arguments of the functions described in Table \ref{tab:arg-shape-tb-pdf}.
 
@@
 \texttt{p} & integer & Number of dimensions.\\
+\texttt{h} &  & \\
+\texttt{ratio} &  & \\
+\addlinespace
+\texttt{s} &  & \\
+\texttt{r} &  & \\
+\texttt{k} &  & \\
+\texttt{loc} &  & \\
+\texttt{w} &  & \\
+\addlinespace
+\texttt{steps} &  & \\
+\texttt{spins} &  & \\
+\texttt{hc} &  & \\
+\texttt{non\_fac} &  & \\
 \bottomrule
@@
 Additionally, \(\theta_3 \sim U(0, \pi/2)\) in the \((x_3, x_4)\) plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
+The coordinates are transformed into \(4\text{-}D\) Cartesian space: \[X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).\] This produces points on one side of a \(4\text{-}D\) unit sphere, effectively generating a \(4\text{-}D\) hemisphere.
 
@@
 Table \ref{tab:add-tb-pdf} details these functions.
-More detailed explanations are available in \url{https://jayanilakshika.github.io/cardinalR/articles/03additionalfun.html}.
+More detailed explanations are available in \href{https://jayanilakshika.github.io/cardinalR/articles/03additionalfun.html}{jayanilakshika.github.io/cardinalR/articles/03additionalfun}.
 

### 2025-12-12 paper-cardinalR.Rmd
Added table caption for the argument table

@@
 ```{r}
-arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "k", "loc", "w", "steps", "spins", "hc", "non_fac"), 
-                          type = c("integer", "integer", "integer", "",
-                                   "", "", "", "", "", "", "", "", "", ""),
+arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac"), 
+                          type = c("integer", "integer", "integer", "real value",
+                                   "real value", "real value", "real value", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value"),
                         exp = c("Number of points.",
@@
                                 "Number of dimensions.",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "",
-                                "")) |> 
+                                "Height.",
+                                "Radius tip to radius base ratio.",
+                                "Variance-covariance matrix.",
+                                "Radius.",
+                                "Number of small spheres.",
+                                "Radius of the big and small spheres.",
+                                "How far the small spheres are placed from each other.",
+                                "Vertical variation",
+                                "Number of steps for the theta parameter.",
+                                "Number of loops of the spiral.",
+                                "Steepness and vertical scaling of the hyperbola.",
+                                "Strength of this sinusoidal effect.")) |> 
    mutate(arg = gsub("_", "\\\\_", arg))
@@
   mutate(arg = paste0("<code>", arg, "</code>")) |>
-  kable(caption = "The main arguments for branching shape generators.", col.names = c("Argument", "Type", "Explanation"), escape = FALSE, table.pos = "!ht") 
+  kable(caption = "Argument definitions for the shape generators.
+The table lists each argument, its data type, and a description of its role in controlling geometric structure, dimensionality, scaling, curvature, spacing, and other features of the simulated high-dimensional datasets.", col.names = c("Argument", "Type (positive)", "Explanation"), escape = FALSE, table.pos = "!ht")
 ```
@@
   mutate(arg = paste0("\\texttt{", arg, "}")) |> 
-  kable(caption = "The main arguments for branching shape generators.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, escape = FALSE, table.pos = "!ht")  |>
+  kable(caption = "Argument definitions for the shape generators.
+The table lists each argument, its data type, and a description of its role in controlling geometric structure, dimensionality, scaling, curvature, spacing, and other features of the simulated high-dimensional datasets.", format="latex", col.names = c("Argument", "Type (positive)", "Explanation"), booktabs = T, escape = FALSE, table.pos = "!ht")  |>
   column_spec(1, width = "1cm") |>

### 2025-12-12 paper-cardinalR.tex
Added table caption for the argument table

@@
 
-\caption{\label{tab:arg-shape-tb-pdf}The main arguments for branching shape generators.}
+\caption{\label{tab:arg-shape-tb-pdf}Argument definitions for the shape generators.
+The table lists each argument, its data type, and a description of its role in controlling geometric structure, dimensionality, scaling, curvature, spacing, and other features of the simulated high-dimensional datasets.}
 \centering
@@
 \toprule
-Argument & Type & Explanation\\
+Argument & Type (positive) & Explanation\\
 \midrule
@@
 \texttt{p} & integer & Number of dimensions.\\
-\texttt{h} &  & \\
-\texttt{ratio} &  & \\
+\texttt{h} & real value & Height.\\
+\texttt{ratio} & real value & Radius tip to radius base ratio.\\
 \addlinespace
-\texttt{s} &  & \\
-\texttt{r} &  & \\
-\texttt{k} &  & \\
-\texttt{loc} &  & \\
-\texttt{w} &  & \\
+\texttt{s} & real value & Variance-covariance matrix.\\
+\texttt{r} & real value & Radius.\\
+\texttt{k\_small} & integer & Number of small spheres.\\
+\texttt{r\_vec} & real values & Radius of the big and small spheres.\\
+\texttt{spe} & real value & How far the small spheres are placed from each other.\\
 \addlinespace
-\texttt{steps} &  & \\
-\texttt{spins} &  & \\
-\texttt{hc} &  & \\
-\texttt{non\_fac} &  & \\
+\texttt{w} & real value & Vertical variation\\
+\texttt{steps} & integer & Number of steps for the theta parameter.\\
+\texttt{spins} & integer & Number of loops of the spiral.\\
+\texttt{hc} & real value & Steepness and vertical scaling of the hyperbola.\\
+\texttt{non\_fac} & real value & Strength of this sinusoidal effect.\\
 \bottomrule

### 2025-12-12 paper-cardinalR.Rmd
Added all arguments

@@
 ```{r}
-arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac"), 
+arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "n_vec", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac"), 
                           type = c("integer", "integer", "integer", "real value",
-                                   "real value", "real value", "real value", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value"),
+                                   "real value", "real value", "real value", "integers", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value"),
                         exp = c("Number of points.",
@@
                                 "Radius.",
+                                "Sample sizes of the big and small spheres",
                                 "Number of small spheres.",

### 2025-12-12 paper-cardinalR.R

@@ -248,9 +248,9 @@ shape_tb |>
 
 
 ## -----------------------------------------------------------------------------
-arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac"), 
+arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "n_vec", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac"), 
                           type = c("integer", "integer", "integer", "real value",
-                                   "real value", "real value", "real value", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value"),
+                                   "real value", "real value", "real value", "integers", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value"),
                         exp = c("Number of points.",
                                 "Number of clusters.",
                                 "Number of dimensions.",
@@ -258,6 +258,7 @@ arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "k_sma
                                 "Radius tip to radius base ratio.",
                                 "Variance-covariance matrix.",
                                 "Radius.",
+                                "Sample sizes of the big and small spheres",
                                 "Number of small spheres.",
                                 "Radius of the big and small spheres.",
                                 "How far the small spheres are placed from each other.",

### 2025-12-12 paper-cardinalR.Rmd
Updated noise dimension section

--- 2025-12-12 paper-cardinalR.Rmd
@@
                                 "Radius.",
+                                "Sample sizes of the big and small spheres",
                                 "Number of small spheres.",
                                 "Radius of the big and small spheres.",
-                                "How far the small spheres are placed from each other.",
+                                "How far apart the small spheres are placed.",
                                 "Vertical variation",
@@
 
-with odd-numbered dimensions multiplied by $-1$ to introduce sign alternation, enhancing variability and decorrelation.
+with odd-numbered dimensions multiplied by $-1$.
+This does not affect independence, since all noise dimensions are generated independently.
+The sign alternation is included only to avoid consistent directional drift and to ensure a symmetric appearance of noise when visualized or combined with other simulated structures.
 

### 2025-12-12 paper-cardinalR.tex
Added all arguments
Updated noise dimension section

@@
 \texttt{r} & real value & Radius.\\
+\texttt{n\_vec} & integers & Sample sizes of the big and small spheres\\
 \texttt{k\_small} & integer & Number of small spheres.\\
 \texttt{r\_vec} & real values & Radius of the big and small spheres.\\
-\texttt{spe} & real value & How far the small spheres are placed from each other.\\
 \addlinespace
+\texttt{spe} & real value & How far apart the small spheres are placed.\\
 \texttt{w} & real value & Vertical variation\\
@@
 \texttt{hc} & real value & Steepness and vertical scaling of the hyperbola.\\
+\addlinespace
 \texttt{non\_fac} & real value & Strength of this sinusoidal effect.\\
@@
 
-with odd-numbered dimensions multiplied by \(-1\) to introduce sign alternation, enhancing variability and decorrelation.
+with odd-numbered dimensions multiplied by \(-1\).
+This does not affect independence, since all noise dimensions are generated independently.
+The sign alternation is included only to avoid consistent directional drift and to ensure a symmetric appearance of noise when visualized or combined with other simulated structures.
 

### 2025-12-13 paper-cardinalR.Rmd
Updated implementation section

@@
 When clusters are generated with fewer dimensions than others, the function augments the lower-dimensional clusters with additional Gaussian noise variables so that all clusters are represented in the same dimensional space.
-These noise dimensions are drawn independently from normal distributions
-
-$$
-X \sim \mathcal{N}(m, s^2),
-$$
-
-where the mean ($m$) is set to the average of the cluster coordinates and the standard deviation ($s$) defaults to $0.2$.
+These noise dimensions are drawn independently from normal distributions $X \sim \mathcal{N}(\mu, \sigma^2)$, where the mean ($\mu$) is set to the average of the cluster coordinates and the standard deviation ($\sigma$) defaults to $0.2$.
 
@@
 These patterns reflect the mismatch between the helical spherical shape present in the data and the spherical decision boundaries imposed by $k$-means.
-Even though the cluster statistics favour a $4–5$ cluster solution overall, `r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` reveals that neither solution provides clear separation, highlighting the importance of visual diagnostics when interpreting clustering results.
+Even though the cluster statistics favour a $4-5$ cluster solution overall, `r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` reveals that neither solution provides clear separation, highlighting the importance of visual diagnostics when interpreting clustering results.
 

### 2025-12-13 paper-cardinalR.tex
Updated implementation section

@@
 When clusters are generated with fewer dimensions than others, the function augments the lower-dimensional clusters with additional Gaussian noise variables so that all clusters are represented in the same dimensional space.
-These noise dimensions are drawn independently from normal distributions
-
-\[
-X \sim \mathcal{N}(m, s^2),
-\]
-
-where the mean (\(m\)) is set to the average of the cluster coordinates and the standard deviation (\(s\)) defaults to \(0.2\).
+These noise dimensions are drawn independently from normal distributions \(X \sim \mathcal{N}(\mu, \sigma^2)\), where the mean (\(\mu\)) is set to the average of the cluster coordinates and the standard deviation (\(\sigma\)) defaults to \(0.2\).
 
@@
 These patterns reflect the mismatch between the helical spherical shape present in the data and the spherical decision boundaries imposed by \(k\)-means.
-Even though the cluster statistics favour a \(4–5\) cluster solution overall, Figure \ref{fig:highd-proj-clust-algo-pdf} reveals that neither solution provides clear separation, highlighting the importance of visual diagnostics when interpreting clustering results.
+Even though the cluster statistics favour a \(4-5\) cluster solution overall, Figure \ref{fig:highd-proj-clust-algo-pdf} reveals that neither solution provides clear separation, highlighting the importance of visual diagnostics when interpreting clustering results.
 

### 2025-12-13 paper-cardinalR.Rmd
Updating branching section

@@
 A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology [@trapnell2014].
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (`n`), and number of branches (`k`).
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, total number of points (`n`) generated across all branches, with points allocated approximately evenly among branches, and number of branches (`k`).
 Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
@@
 
-Branches $1$ and $2$ are initialized with fixed slopes and intercepts, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
+The jitter term is sampled from a one-sided uniform distribution to introduce directional variability without altering branch orientation.
+
+Branches $1$ and $2$ are initialized with fixed slopes and starting points, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
 
@@
 where $s_i$ is a scale factor controlling the curvature of branch $i$.
-For the first few initial branches, $s_i$ can be fixed (e.g., $s_1 = 1, s_2 = 2$), while for subsequent branches it is sampled from a predefined set, such as $s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}$, to create variability in curvature.
+For the first few initial branches, $s_i$ can be fixed (e.g., $s_1 = 1, s_2 = 2$) to establish reference shapes, while for subsequent branches it is sampled from a predefined set, such as $s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}$, to control curvature magnitude and direction.
 
@@
 High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) and `gen_orgcurvybranches(n, p, k)`.
-Each branch is embedded in a unique or repeated $2\text{-}D$ subspace of the $p\text{-}D$ space.
+For branch
+$i$, the active coordinate pair $(i_1, i_2)$ indexes the selected $2\text{-}D$ subspace.
 When `allow_share = TRUE`, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible $\binom{p}{2}$ combinations are exhausted, after which additional branches may repeat subspaces.
@@
 
-where $a_i, b_i$ define the range of the first coordinate for branch $i$, and $\epsilon$ is Gaussian noise added to introduce variability.
+where $a_i, b_i$ define the range of the first coordinate for branch $i$, and $\epsilon$ is Gaussian noise added to introduce smooth variability in $p\text{-}D$ without sharp boundaries.
 The scale factor $s_i$ controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first $\binom{p}{2}$ branches, $s_i = 1$; for additional branches when $k > \binom{p}{2}$, $s_i$ is randomly drawn from the set $\{1, 1.5, 2, \dots, 8\}$.
+
+Across all branching generators, the scale parameter $s_i$ controls the strength of deviation from linearity, determining slope, curvature, or growth rate depending on the branch geometry.
 
@@
 ```{r data-clusteredspheres}
-clusteredspheres <- gen_clusteredspheres(n = c(1000, 100), k = 3, r = c(15, 3),
-                                         loc = 10 / sqrt(3)) |>
+clusteredspheres <- gen_clusteredspheres(n_vec = c(1000, 100), k_small = 3, r_vec = c(15, 3),
+                                         sep = 10 / sqrt(3)) |>
   dplyr::select(-cluster)

### 2025-12-13 paper-cardinalR.tex
Updating branching section

@@
 A branching structure (Figure \ref{fig:branch-proj}) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology \citep{trapnell2014}.
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, numbers of points (\texttt{n}), and number of branches (\texttt{k}).
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, total number of points (\texttt{n}) generated across all branches, with points allocated approximately evenly among branches, and number of branches (\texttt{k}).
 Although these functions can generate multiple branches, they do not produce a formal \emph{multicluster} dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
@@
 
-Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
+The jitter term is sampled from a one-sided uniform distribution to introduce directional variability without altering branch orientation.
+
+Branches \(1\) and \(2\) are initialized with fixed slopes and starting points, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
 
@@
 where \(s_i\) is a scale factor controlling the curvature of branch \(i\).
-For the first few initial branches, \(s_i\) can be fixed (e.g., \(s_1 = 1, s_2 = 2\)), while for subsequent branches it is sampled from a predefined set, such as \(s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}\), to create variability in curvature.
+For the first few initial branches, \(s_i\) can be fixed (e.g., \(s_1 = 1, s_2 = 2\)) to establish reference shapes, while for subsequent branches it is sampled from a predefined set, such as \(s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}\), to control curvature magnitude and direction.
 
@@
 
-High-dimensional generalizations are provided by \texttt{gen\_orglinearbranches(n,\ p,\ k)} (Figure \ref{fig:branch-proj}) and \texttt{gen\_orgcurvybranches(n,\ p,\ k)}.
-Each branch is embedded in a unique or repeated \(2\text{-}D\) subspace of the \(p\text{-}D\) space.
+High-dimensional generalizations are provided by \texttt{gen\_orglinearbranches(n,\ p,\ k)} (Figure \ref{fig:branch-proj}) and \texttt{gen\_orgcurvybranches(n,\ p,\ k)}. For branch
+\(i\), the active coordinate pair \((i_1, i_2)\) indexes the selected \(2\text{-}D\) subspace.
 When \texttt{allow\_share\ =\ TRUE}, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible \(\binom{p}{2}\) combinations are exhausted, after which additional branches may repeat subspaces.
@@
 
-where \(a_i, b_i\) define the range of the first coordinate for branch \(i\), and \(\epsilon\) is Gaussian noise added to introduce variability.
+where \(a_i, b_i\) define the range of the first coordinate for branch \(i\), and \(\epsilon\) is Gaussian noise added to introduce smooth variability in \(p\text{-}D\) without sharp boundaries.
 The scale factor \(s_i\) controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first \(\binom{p}{2}\) branches, \(s_i = 1\); for additional branches when \(k > \binom{p}{2}\), \(s_i\) is randomly drawn from the set \(\{1, 1.5, 2, \dots, 8\}\).
+
+Across all branching generators, the scale parameter \(s_i\) controls the strength of deviation from linearity, determining slope, curvature, or growth rate depending on the branch geometry.
 

### 2025-12-13 paper-cardinalR.Rmd
Updated branching section

@@
 
-To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in $2\text{-}D$. Branches $1$ and $2$ are simple parabolas defined as
+To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in $2\text{-}D$. Each branch follows a quadratic trajectory of the form
 
 $$
-\begin{aligned}
-\text{Branch 1: } & X_1 \sim U(0,1), \quad X_2 = 0.1 X_1 + X_1^2 + \epsilon, \\
-\text{Branch 2: } & X_1 \sim U(-1,0), \quad X_2 = 0.1 X_1 - 2 X_1^2 + \epsilon, \quad \epsilon \sim U(0, \delta),
-\end{aligned}
+X_1 \sim U(a_i, b_i), \quad X_2 = 0.1 X_1 + s_i X_1^2 + \epsilon, \quad \epsilon \sim U(-\delta, \delta),
 $$
-
-where $\delta$ controls local jitter.
+where $(a_i, b_i)$ defines the domain of the branch, $s_i$ controls curvature, and $\delta$ introduces local jitter.
+For the first two branches, the parameters are fixed to establish reference shapes: $(a_1,b_1,s_1) = (0,1,1), \quad (a_2,b_2,s_2) = (-1,0,-2)$.
 Additional branches are attached iteratively to existing structures.
@@
 When `allow_share = TRUE`, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible $\binom{p}{2}$ combinations are exhausted, after which additional branches may repeat subspaces.
-Linear branches follow
+
+In both cases, branch $i$ is generated according to
 
 $$
-X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = s_i X_{i_1} + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+X_{i_1} \sim U(a_i, b_i), \quad
+X_{i_2} = f_i(X_{i_1}) + \epsilon, \quad
+\epsilon \sim N(0, \sigma^2),
 $$
-
-while curvilinear branches include a quadratic term
-
+where $a_i$ and $b_i$ define the domain of the branch and $\epsilon$ introduces smooth variability in the $p\text{-}D$ space. The function $f_i(\cdot)$ determines the branch geometry:
 $$
-X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = - s_i X_{i_1}^2 + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+f_i(x) =
+\begin{cases}
+s_i x, & \text{linear branches}, \\
+-s_i x^2, & \text{curvilinear branches}.
+\end{cases}
 $$
 
-where $a_i, b_i$ define the range of the first coordinate for branch $i$, and $\epsilon$ is Gaussian noise added to introduce smooth variability in $p\text{-}D$ without sharp boundaries.
 The scale factor $s_i$ controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first $\binom{p}{2}$ branches, $s_i = 1$; for additional branches when $k > \binom{p}{2}$, $s_i$ is randomly drawn from the set $\{1, 1.5, 2, \dots, 8\}$.

### 2025-12-13 paper-cardinalR.tex
Updated branching section

@@
 
-To introduce curvature, the \texttt{gen\_curvybranches(n,\ k)} function generates \(k\) curvilinear branches in \(2\text{-}D\). Branches \(1\) and \(2\) are simple parabolas defined as
+To introduce curvature, the \texttt{gen\_curvybranches(n,\ k)} function generates \(k\) curvilinear branches in \(2\text{-}D\). Each branch follows a quadratic trajectory of the form
 
 \[
-\begin{aligned}
-\text{Branch 1: } & X_1 \sim U(0,1), \quad X_2 = 0.1 X_1 + X_1^2 + \epsilon, \\
-\text{Branch 2: } & X_1 \sim U(-1,0), \quad X_2 = 0.1 X_1 - 2 X_1^2 + \epsilon, \quad \epsilon \sim U(0, \delta),
-\end{aligned}
+X_1 \sim U(a_i, b_i), \quad X_2 = 0.1 X_1 + s_i X_1^2 + \epsilon, \quad \epsilon \sim U(-\delta, \delta),
 \]
-
-where \(\delta\) controls local jitter.
+where \((a_i, b_i)\) defines the domain of the branch, \(s_i\) controls curvature, and \(\delta\) introduces local jitter.
+For the first two branches, the parameters are fixed to establish reference shapes: \((a_1,b_1,s_1) = (0,1,1), \quad (a_2,b_2,s_2) = (-1,0,-2)\).
 Additional branches are attached iteratively to existing structures.
@@
 When \texttt{allow\_share\ =\ TRUE}, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible \(\binom{p}{2}\) combinations are exhausted, after which additional branches may repeat subspaces.
-Linear branches follow
+
+In both cases, branch \(i\) is generated according to
 
 \[
-X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = s_i X_{i_1} + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
+X_{i_1} \sim U(a_i, b_i), \quad
+X_{i_2} = f_i(X_{i_1}) + \epsilon, \quad
+\epsilon \sim N(0, \sigma^2),
+\]
+where \(a_i\) and \(b_i\) define the domain of the branch and \(\epsilon\) introduces smooth variability in the \(p\text{-}D\) space. The function \(f_i(\cdot)\) determines the branch geometry:
+\[
+f_i(x) =
+\begin{cases}
+s_i x, & \text{linear branches}, \\
+-s_i x^2, & \text{curvilinear branches}.
+\end{cases}
 \]
 
-while curvilinear branches include a quadratic term
-
-\[
-X_{i_1} \sim U(a_i,b_i), \quad X_{i_2} = - s_i X_{i_1}^2 + \epsilon, \quad \epsilon \sim N(0, \sigma^2),
-\]
-
-where \(a_i, b_i\) define the range of the first coordinate for branch \(i\), and \(\epsilon\) is Gaussian noise added to introduce smooth variability in \(p\text{-}D\) without sharp boundaries.
 The scale factor \(s_i\) controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first \(\binom{p}{2}\) branches, \(s_i = 1\); for additional branches when \(k > \binom{p}{2}\), \(s_i\) is randomly drawn from the set \(\{1, 1.5, 2, \dots, 8\}\).

### 2025-12-13 paper-cardinalR.Rmd
Updated cube and Gaussian sections

@@
 This function generates $n$ points in $p\text{-}D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip.
-Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, capped at the cone height $h$, producing a higher density of points near the tip.
-At each height $X_p$, the radius of the cross-section decreases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio}$ and $r_{\text{max}} = 1$.
+Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, truncated to the interval $[0, h]$, producing a higher density of points near the tip.
+At each height $X_p$, the radius of the cross-section increases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio} \in [0, 1]$ and $r_{\text{max}} = 1$.
 
@@
 It generates a regular lattice of points in $p\text{-}D$, producing a uniform hypercube grid.
-Each axis contains equally spaced coordinates, resulting in a well-defined geometric structure.
+The parameter `n` controls the approximate number of points by determining the grid resolution along each axis.
 
@@
 To avoid including the cube’s vertices, these points are removed after generation.
-This results in a hypercube filled with random, but evenly distributed, samples rather than structured lattice points.
+This results in a hypercube filled with random samples rather than structured lattice points.
 
@@
 
-The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure.
-Each point is independently drawn using the multivariate normal distribution with $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
+The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure. For 
+$i=1,\dots,n$, each observation is independently drawn from a multivariate normal distribution,  $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
 

### 2025-12-13 paper-cardinalR.tex
Updated cube and Gaussian sections

@@
 This function generates \(n\) points in \(p\text{-}D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip.
-Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), capped at the cone height \(h\), producing a higher density of points near the tip.
-At each height \(X_p\), the radius of the cross-section decreases linearly from base to tip according to \(r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h\), where \(r_{\text{min}} = \text{ratio}\) and \(r_{\text{max}} = 1\).
+Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), truncated to the interval \([0, h]\), producing a higher density of points near the tip.
+At each height \(X_p\), the radius of the cross-section increases linearly from base to tip according to \(r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h\), where \(r_{\text{min}} = \text{ratio} \in [0, 1]\) and \(r_{\text{max}} = 1\).
 
@@
 It generates a regular lattice of points in \(p\text{-}D\), producing a uniform hypercube grid.
-Each axis contains equally spaced coordinates, resulting in a well-defined geometric structure.
+The parameter \texttt{n} controls the approximate number of points by determining the grid resolution along each axis.
 
@@
 To avoid including the cube's vertices, these points are removed after generation.
-This results in a hypercube filled with random, but evenly distributed, samples rather than structured lattice points.
+This results in a hypercube filled with random samples rather than structured lattice points.
 
@@
 
-The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure.
-Each point is independently drawn using the multivariate normal distribution with \(X_i \sim N_p(\boldsymbol{0}, s)\), where \(s\) is a user-defined \(p \times p\) positive-definite matrix.
+The \texttt{gen\_gaussian(n,\ p,\ s)} function generates a multivariate Gaussian cloud in \(p\text{-}D\), centered at the origin with user-defined covariance structure. For
+\(i=1,\dots,n\), each observation is independently drawn from a multivariate normal distribution, \(X_i \sim N_p(\boldsymbol{0}, s)\), where \(s\) is a user-defined \(p \times p\) positive-definite matrix.
 

### 2025-12-13 paper-cardinalR.Rmd
Updated rectangular base pyramid

@@
 The remaining dimensions are based on the specific pyramid shape.
-For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`), let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$.
-That is, $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
-The first three coordinates are then defined as $X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).$
+For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`) the base shape is a rectangle whose size shrinks linearly with height.
+Let $l_x$ and $l_y$ denote the half-widths of the rectangular base in the $X_1$ and $X_2$ directions, specified via $l=(l_x,l_y)$, and let $r_t$ denote the half-width at the pyramid tip.
+At height $z\in [0,h]$, the half-widths of the rectangular cross-section are $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
+The first three coordinates are then defined as $X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z))$.
 
@@
 That is, $r(z) = r_b\left(1 - z/h\right)$.
-Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$, $r_{\text{point}} \sim \sqrt{U(0, 1)}$.
+Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim DiscreteUniform\{0, \pi/3, \dots, 5\pi/3\}$, $r_{\text{point}} \sim \sqrt{U(0, 1)}$.
 Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
@@
 The simplest case, `gen_circle(n, p)` creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
-Let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
+Let a latent angle variable $\theta \sim U(0, 2\pi)$.
 Coordinates in the first two dimensions represent a perfect circle on the plane: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$ For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
@@
 
-For the one-dimensional nonlinear cycle embedded in $p\text{-}D$ space, `gen_curvycycle(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`), let a latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$.
-The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle".
+For the one-dimensional nonlinear cycle embedded in $p\text{-}D$ space, `gen_curvycycle(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`), let a latent angle variable $\theta \sim U(0, 2\pi)$.
+The first three dimensions define a non-circular closed curve, referred to as a `"curvy cycle"`.
 In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin.
@@
 The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from $4$ to $p$, which means that the amplitude decreases as the dimension increases.
-Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j\sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2)\pi/2p$.
+Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j\sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2)\pi/(2p)$.
 
@@
 A band of thickness around the knot path is controlled by the `steps` argument, while the number of $\theta$ and $\phi$ values is determined by the `steps` and `n` arguments, respectively (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
-The coordinates are defined as $$X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \\\quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),$$ where $\theta$ and $\phi$ trace the knot’s path.
+The coordinates are defined as $$X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \\\quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),$$ where $\theta$ parameterizes the band thickness and $\phi$ parameterizes the knot trajectory.
 
@@
 First, the `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) a"} else { "Figure \\@ref(fig:triginometric-proj) a"})`).
-Let $\theta \in [\pi/6, 2\pi]$ be a sequence of $n$ evenly spaced angles.
-The corresponding $2\text{-}D$ coordinates are defined by: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$
+Let $\{\theta_i\}_{i=1}^n$  be a sequence of $n$ evenly spaced angles on the interval $[\pi/6, 2\pi]$, defined as $\theta_i = \frac{\pi}{6} + (i-1)\frac{2\pi - \pi/6}{n-1}, \quad i = 1,\dots,n$.
+The corresponding $2\text{-}D$ coordinates are defined by: $$X_{i1} = \cos(\theta_i), \quad X_{i2} = \sin(\theta_i).$$
 
@@
 
-For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$ be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
+For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$  be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$ be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
 Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
@@
 The shape is defined by parameter $\theta \in [0, 2\pi\text{spins}]$, controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta\cos(\theta)$ for radial expansion in x, and $X_2 = \theta\sin(\theta)$ for radial expansion in y.
+The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta\cos(\theta)$ for radial expansion in $x$, and $X_2 = \theta\sin(\theta)$ for radial expansion in $y$.
 The growth pattern resembles a cone, with the height increasing according to $X_3 = 2\theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta\sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
@@
 Finally, the `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
-The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{non\_fac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.

### 2025-12-13 paper-cardinalR.tex
Updated rectangular base pyramid

@@
 The remaining dimensions are based on the specific pyramid shape.
-For the rectangular based pyramid, \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} (Figure \ref{fig:pyr-proj} a), let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\).
-That is, \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
-The first three coordinates are then defined as \(X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z)).\)
+For the rectangular based pyramid, \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} (Figure \ref{fig:pyr-proj} a) the base shape is a rectangle whose size shrinks linearly with height.
+Let \(l_x\) and \(l_y\) denote the half-widths of the rectangular base in the \(X_1\) and \(X_2\) directions, specified via \(l=(l_x,l_y)\), and let \(r_t\) denote the half-width at the pyramid tip.
+At height \(z\in [0,h]\), the half-widths of the rectangular cross-section are \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).
+The first three coordinates are then defined as \(X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z))\).
 
@@
 That is, \(r(z) = r_b\left(1 - z/h\right)\).
-Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\), \(r_{\text{point}} \sim \sqrt{U(0, 1)}\).
+Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}\) and a uniformly random radial scaling factor: \(\theta \sim DiscreteUniform\{0, \pi/3, \dots, 5\pi/3\}\), \(r_{\text{point}} \sim \sqrt{U(0, 1)}\).
 Then, the first two coordinates are: \(X_1 = r(z)r_{\text{point}}\cos(\theta)\), and \(X_2 = r(z)r_{\text{point}}\sin(\theta)\).
@@
 The simplest case, \texttt{gen\_circle(n,\ p)} creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (Figure \ref{fig:sphere-proj} a).
-Let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
+Let a latent angle variable \(\theta \sim U(0, 2\pi)\).
 Coordinates in the first two dimensions represent a perfect circle on the plane: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta).\] For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
@@
 
-For the one-dimensional nonlinear cycle embedded in \(p\text{-}D\) space, \texttt{gen\_curvycycle(n,\ p)} (Figure \ref{fig:sphere-proj} b), let a latent angle variable \(\theta\) is uniformly sampled from the interval \([0, 2\pi]\).
-The first three dimensions define a non-circular closed curve, referred to as a ``curvy cycle''.
+For the one-dimensional nonlinear cycle embedded in \(p\text{-}D\) space, \texttt{gen\_curvycycle(n,\ p)} (Figure \ref{fig:sphere-proj} b), let a latent angle variable \(\theta \sim U(0, 2\pi)\).
+The first three dimensions define a non-circular closed curve, referred to as a \texttt{"curvy\ cycle"}.
 In this configuration, \(X_1 = \cos(\theta)\) represents horizontal oscillation, while \(X_2 = \sqrt{3}/3 + \sin(\theta)\) introduces a vertical offset to avoid centering the curve at the origin.
@@
 The scaling factor is defined as \(\text{scale}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from \(4\) to \(p\), which means that the amplitude decreases as the dimension increases.
-Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j\sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2)\pi/2p\).
+Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j\sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2)\pi/(2p)\).
 
@@
 A band of thickness around the knot path is controlled by the \texttt{steps} argument, while the number of \(\theta\) and \(\phi\) values is determined by the \texttt{steps} and \texttt{n} arguments, respectively (Figure \ref{fig:trefoil-proj} a).
-The coordinates are defined as \[X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \\\quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),\] where \(\theta\) and \(\phi\) trace the knot's path.
+The coordinates are defined as \[X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \\\quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),\] where \(\theta\) parameterizes the band thickness and \(\phi\) parameterizes the knot trajectory.
 
@@
 First, the \texttt{gen\_crescent(n,\ p)} function generates a \(p\)-dimensional dataset of \(n\) observations based on a \(2\text{-}D\) crescent-shaped manifold with optional structured high-dimensional noise (Figure \ref{fig:triginometric-proj} a).
-Let \(\theta \in [\pi/6, 2\pi]\) be a sequence of \(n\) evenly spaced angles.
-The corresponding \(2\text{-}D\) coordinates are defined by: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta).\]
+Let \(\{\theta_i\}_{i=1}^n\) be a sequence of \(n\) evenly spaced angles on the interval \([\pi/6, 2\pi]\), defined as \(\theta_i = \frac{\pi}{6} + (i-1)\frac{2\pi - \pi/6}{n-1}, \quad i = 1,\dots,n\).
+The corresponding \(2\text{-}D\) coordinates are defined by: \[X_{i1} = \cos(\theta_i), \quad X_{i2} = \sin(\theta_i).\]
 
@@
 
-For a spiraling path on a spherical surface in the first four dimensions, \texttt{gen\_sphericalspiral(n,\ p,\ spins)} (Figure \ref{fig:triginometric-proj} c), let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\)be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
+For a spiraling path on a spherical surface in the first four dimensions, \texttt{gen\_sphericalspiral(n,\ p,\ spins)} (Figure \ref{fig:triginometric-proj} c), let \(\theta \in [0, 2\pi \times \text{spins}]\) be the azimuthal angle (longitude), controls the number of spiral turns and the \(\phi \in [0, \pi]\) be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
 Cartesian coordinates from spherical conversion: \(X_1 = \sin(\phi)\cos(\theta)\), \(X_2 = \sin(\phi)\sin(\theta)\), \(X_3 = \cos(\phi) + \varepsilon\), where \(\varepsilon \sim U(-0.5, 0.5)\) introduces vertical jitter, and \(X_4 = \theta / \max(\theta)\): a normalized progression along the spiral path.
@@
 The shape is defined by parameter \(\theta \in [0, 2\pi\text{spins}]\), controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; \(X_1 = \theta\cos(\theta)\) for radial expansion in x, and \(X_2 = \theta\sin(\theta)\) for radial expansion in y.
+The Archimedean spiral in the horizontal plane is represented by; \(X_1 = \theta\cos(\theta)\) for radial expansion in \(x\), and \(X_2 = \theta\sin(\theta)\) for radial expansion in \(y\).
 The growth pattern resembles a cone, with the height increasing according to \(X_3 = 2\theta / \max(\theta) + \varepsilon_3\), with \(\varepsilon_3 \sim U(-0.1, 0.6).\) Spiral modulation in the fourth dimension is represented by \(X_4 = \theta\sin(2\theta) + \varepsilon_4\), with \(\varepsilon_4 \sim U(-0.1, 0.6)\) which simulates a twisting helical component in a non-radial dimension.
@@
 Finally, the \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
-The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{nonfac}\sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{non\_fac}\sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
 These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.

### 2025-12-13 paper-cardinalR.Rmd
Updated the polynomial section

@@
 This is generated using `gen_quadratic(n, range)`.
-The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree 2 is applied to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
+The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and the response is generated as $X_2 = X_1 - X_1^2 + \varepsilon$, where $\varepsilon \sim U(0, 0.5)$.
 This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.

### 2025-12-13 paper-cardinalR.tex
Updated the polynomial section

@@
 This is generated using \texttt{gen\_quadratic(n,\ range)}.
-The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree 2 is applied to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
+The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and the response is generated as \(X_2 = X_1 - X_1^2 + \varepsilon\), where \(\varepsilon \sim U(0, 0.5)\).
 This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.

### 2025-12-13 paper-cardinalR.Rmd
Fix range notation in polynomial functions

@@
 
-The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise.
-Each variable is constructed by first creating a sequence of integers from $0$ to $n-1$, adding Gaussian noise with standard deviation $0.03n$, and then applying a random scaling and shift.
-Specifically, for the $j$-th dimension, a scale factor $\text{scale}_j \sim U(-10, 10)$ determines the orientation and stretching of the line, while a shift factor $\text{shift}_j \sim U(-300, 300)$ is added inside the multiplication to offset the line.
-This results in each variable being of the form $X_j = \text{scale}_j \cdot \big((0,1,\dots,n-1) + \epsilon + \text{shift}_j\big)$, where $\epsilon \sim N(0, (0.03n)^2)$ is independent Gaussian noise.
+The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a single noisy linear trajectory.
+Let $t_i = i - 1, \quad i = 1, \dots, n$, denote a common latent index shared across all dimensions.
+For each dimension $j = 1, \dots, p)$, independent scale and shift parameters are sampled as $a_j \sim U(-10, 10), \qquad b_j \sim U(-300, 300)$.
+Gaussian noise $\varepsilon_{ij} \sim N(0, (0.03n)^2)$ is added independently across observations and dimensions.
+The observed variables are then defined as $X_{ij} = a_j \bigl(t_i + b_j + \varepsilon_{ij}\bigr), \quad i = 1, \dots, n$.cThis construction yields a single elongated linear structure embedded in $p\text{-}D$, with each dimension exhibiting a different orientation, scale, and offset.
 
@@
 This is generated using `gen_quadratic(n, range)`.
-The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and the response is generated as $X_2 = X_1 - X_1^2 + \varepsilon$, where $\varepsilon \sim U(0, 0.5)$.
+Let $range = [a, b]$.
+The independent variable is defined as $X_1 \sim U(a, b)$, and the response is generated as $X_2 = X_1 - X_1^2 + \varepsilon$, where $\varepsilon \sim U(0, 0.5)$.
 This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
@@
 This is generated using `gen_cubic(n, range)`.
-The independent variable is defined as $X_1 \sim U(\text{range}[1], \text{range}[2])$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
+Let $range = [a, b]$.
+The independent variable is defined as $X_1 \sim U(a, b)$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
 This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.

### 2025-12-13 paper-cardinalR.tex
Fix range notation in polynomial functions

@@
 
-The \texttt{gen\_longlinear(n,\ p)} function generates a high-dimensional dataset representing a long linear structure with noise.
-Each variable is constructed by first creating a sequence of integers from \(0\) to \(n-1\), adding Gaussian noise with standard deviation \(0.03n\), and then applying a random scaling and shift.
-Specifically, for the \(j\)-th dimension, a scale factor \(\text{scale}_j \sim U(-10, 10)\) determines the orientation and stretching of the line, while a shift factor \(\text{shift}_j \sim U(-300, 300)\) is added inside the multiplication to offset the line.
-This results in each variable being of the form \(X_j = \text{scale}_j \cdot \big((0,1,\dots,n-1) + \epsilon + \text{shift}_j\big)\), where \(\epsilon \sim N(0, (0.03n)^2)\) is independent Gaussian noise.
+The \texttt{gen\_longlinear(n,\ p)} function generates a high-dimensional dataset representing a single noisy linear trajectory.
+Let \(t_i = i - 1, \quad i = 1, \dots, n\), denote a common latent index shared across all dimensions.
+For each dimension \(j = 1, \dots, p)\), independent scale and shift parameters are sampled as \(a_j \sim U(-10, 10), \qquad b_j \sim U(-300, 300)\).
+Gaussian noise \(\varepsilon_{ij} \sim N(0, (0.03n)^2)\) is added independently across observations and dimensions.
+The observed variables are then defined as \(X_{ij} = a_j \bigl(t_i + b_j + \varepsilon_{ij}\bigr), \quad i = 1, \dots, n\).cThis construction yields a single elongated linear structure embedded in \(p\text{-}D\), with each dimension exhibiting a different orientation, scale, and offset.
 
@@
 This is generated using \texttt{gen\_quadratic(n,\ range)}.
-The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and the response is generated as \(X_2 = X_1 - X_1^2 + \varepsilon\), where \(\varepsilon \sim U(0, 0.5)\).
+Let \(range = [a, b]\).
+The independent variable is defined as \(X_1 \sim U(a, b)\), and the response is generated as \(X_2 = X_1 - X_1^2 + \varepsilon\), where \(\varepsilon \sim U(0, 0.5)\).
 This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
@@
 This is generated using \texttt{gen\_cubic(n,\ range)}.
-The independent variable is defined as \(X_1 \sim U(\text{range}[1], \text{range}[2])\), and a raw polynomial basis of degree \(3\) is applied to construct \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
+Let \(range = [a, b]\).
+The independent variable is defined as \(X_1 \sim U(a, b)\), and a raw polynomial basis of degree \(3\) is applied to construct \(X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\).
 This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.

### 2025-12-15 paper-cardinalR.Rmd
clustering example refined

@@
 <!--script/05_gen_cluster_stat.R-->
-```{r, label = "fig-cluster-stats", fig.width=10, fig.height=3, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
+```{r, label = "fig-cluster-stats", fig.width=8, fig.height=6, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
 Several indices consistently suggest that $4–5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics."}
@@
 Although the metrics differ in their preferences, several show consistent support for a $4-5$ cluster solution.
-The Calinski–Harabasz index increases sharply up to four or five clusters before levelling off, the Dunn and Pearson gamma indices attain local maxima in this range, and both the silhouette width and within-cluster sum of squares show diminishing improvement beyond five clusters.
-Together, these patterns indicate that $4-5$ clusters provide a good balance of separation and compactness, with $k$-means performing slightly better than the other methods across most metrics.
+Pearson gamma (`pearsongamma`) increases sharply up to five clusters before levelling off, Calinski–Harabasz index (`ch`) increases sharply from 4 to 5 clusters and Dunn (`dunn2`) has a maximum at 5 for two methods and at 4 for $k$-means.
+All of these are interpreted as higher is better.
+With the other three, lower is better.
+WB ratio (`wbratio`) and within-cluster sum of squares (`within.cluster.ss`) steadily decline with number of clusters, possibly elbowing around 5 clusters.
+The S-index (`sindex`) is optimised at 4 clusters for $k$-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.
+Overall, $k$-means performs slightly better than the hierarchical and model-based clustering across most metrics and number of clusters.
 
 ```{r}
+set.seed(1214)
+
 ## k-means with  four clusters
@@
 kcl5 <- kmeans(five_clusts[, -5], centers = 5, nstart = 20)$cluster
+
 ```
@@
 
-`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` shows the four- and five-cluster $k$-means solutions further clarify why these partitions do not fully identify the underlying structure.
-The predicted clusters show noticeable overlap, with colours repeatedly blending as the tour moves; evidence that points from different true clusters are assigned to the same cluster.
-The $4$-cluster solution shows large, diffuse regions where several true clusters are merged, while the $5$-cluster solution partially separates some structures but still misclassifies substantial portions of multiple clusters.
-These patterns reflect the mismatch between the helical spherical shape present in the data and the spherical decision boundaries imposed by $k$-means.
-Even though the cluster statistics favour a $4-5$ cluster solution overall, `r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` reveals that neither solution provides clear separation, highlighting the importance of visual diagnostics when interpreting clustering results.
+`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` shows the four- and five-cluster $k$-means solutions, with cluster id used to colour the points.
+Neither solution captures the geometric nature of the true clusters, but they are both reasonable partitions of the data.
+To examine either one, it is best to subset to a single cluster to view in the tour.
+With each solution, the five original shapes are each split by the clustering.
+More than 5 clusters would be needed to better capture the original shapes.
 

### 2025-12-15 paper-cardinalR.tex
clustering example refined

@@
 Although the metrics differ in their preferences, several show consistent support for a \(4-5\) cluster solution.
-The Calinski--Harabasz index increases sharply up to four or five clusters before levelling off, the Dunn and Pearson gamma indices attain local maxima in this range, and both the silhouette width and within-cluster sum of squares show diminishing improvement beyond five clusters.
-Together, these patterns indicate that \(4-5\) clusters provide a good balance of separation and compactness, with \(k\)-means performing slightly better than the other methods across most metrics.
+Pearson gamma (\texttt{pearsongamma}) increases sharply up to five clusters before levelling off, Calinski--Harabasz index (\texttt{ch}) increases sharply from 4 to 5 clusters and Dunn (\texttt{dunn2}) has a maximum at 5 for two methods and at 4 for \(k\)-means.
+All of these are interpreted as higher is better.
+With the other three, lower is better.
+WB ratio (\texttt{wbratio}) and within-cluster sum of squares (\texttt{within.cluster.ss}) steadily decline with number of clusters, possibly elbowing around 5 clusters.
+The S-index (\texttt{sindex}) is optimised at 4 clusters for \(k\)-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.
+Overall, \(k\)-means performs slightly better than the hierarchical and model-based clustering across most metrics and number of clusters.
 
@@
 
-Figure \ref{fig:highd-proj-clust-algo-pdf} shows the four- and five-cluster \(k\)-means solutions further clarify why these partitions do not fully identify the underlying structure.
-The predicted clusters show noticeable overlap, with colours repeatedly blending as the tour moves; evidence that points from different true clusters are assigned to the same cluster.
-The \(4\)-cluster solution shows large, diffuse regions where several true clusters are merged, while the \(5\)-cluster solution partially separates some structures but still misclassifies substantial portions of multiple clusters.
-These patterns reflect the mismatch between the helical spherical shape present in the data and the spherical decision boundaries imposed by \(k\)-means.
-Even though the cluster statistics favour a \(4-5\) cluster solution overall, Figure \ref{fig:highd-proj-clust-algo-pdf} reveals that neither solution provides clear separation, highlighting the importance of visual diagnostics when interpreting clustering results.
+Figure \ref{fig:highd-proj-clust-algo-pdf} shows the four- and five-cluster \(k\)-means solutions, with cluster id used to colour the points.
+Neither solution captures the geometric nature of the true clusters, but they are both reasonable partitions of the data.
+To examine either one, it is best to subset to a single cluster to view in the tour.
+With each solution, the five original shapes are each split by the clustering.
+More than 5 clusters would be needed to better capture the original shapes.
 

### 2025-12-15 paper-cardinalR.Rmd
Started updating projections

@@
 ```{r orglinearbranches-proj1}
-scaled_data <- scale_data_manual(orglinearbranches)
+centered_data <- center_data(orglinearbranches)
 
@@
                             proj_scale = 1.8, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.15,
@@
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.7,
@@
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.6,
@@
 ```{r cone-proj1}
-scaled_data <- scale_data_manual(cone)
+centered_data <- center_data(cone)
 
@@
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.25,
@@
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.1,
@@
 ```{r pyrrect-proj1}
-scaled_data <- scale_data_manual(pyrrect)
+centered_data <- center_data(pyrrect)
 
@@
 ```{r pyrtri-proj1}
-scaled_data <- scale_data_manual(pyrtri)
+centered_data <- center_data(pyrtri)
 
@@
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.3,
@@
 ```{r pyrstar-proj1}
-scaled_data <- scale_data_manual(pyrstar)
+centered_data <- center_data(pyrstar)
 
@@
 ```{r pyrholes-proj1}
-scaled_data <- scale_data_manual(pyrholes)
+centered_data <- center_data(pyrholes)
 
@@
 ```{r circle-proj1}
-scaled_data <- scale_data_manual(circle)
+centered_data <- center_data(circle)
 
@@
 ```{r curvycycle-proj1}
-scaled_data <- scale_data_manual(curvycycle)
+centered_data <- center_data(curvycycle)
 
@@
 ```{r clusteredspheres-proj1}
-scaled_data <- scale_data_manual(clusteredspheres)
+centered_data <- center_data(clusteredspheres)
 
@@
 ```{r trefoil4d-proj1}
-scaled_data <- scale_data_manual(trefoil4d)
+centered_data <- center_data(trefoil4d)
 
@@
 ```{r trefoil3d-proj1}
-scaled_data <- scale_data_manual(trefoil3d)
+centered_data <- center_data(trefoil3d)
 
@@
 ```{r curvycylinder-proj1}
-scaled_data <- scale_data_manual(curvycylinder)
+centered_data <- center_data(curvycylinder)
 
@@
 ```{r sphericalspiral-proj1}
-scaled_data <- scale_data_manual(sphericalspiral)
+centered_data <- center_data(sphericalspiral)
 
@@
 ```{r conicspiral-proj1}
-scaled_data <- scale_data_manual(conicspiral)
+centered_data <- center_data(conicspiral)
 
@@
 ```{r nonlinear-proj1}
-scaled_data <- scale_data_manual(nonlinear)
+centered_data <- center_data(nonlinear)
 
@@
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.45,
@@
 ```{r five-clusts-projections1}
-scaled_data <- scale_data_manual(five_clusts[, -5])
+centered_data <- center_data(five_clusts[, -5])
 
@@
                             proj_scale = 1.23, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 1,
@@
 ```{r five-4clusts-projections1}
-scaled_data <- scale_data_manual(five_clusts[, -5])
+centered_data <- center_data(five_clusts[, -5])
 
@@
 ```{r five-5clusts-projections1}
-scaled_data <- scale_data_manual(five_clusts[, -5])
+centered_data <- center_data(five_clusts[, -5])
 

### 2025-12-15 paper-cardinalR.R

@@ -306,7 +306,7 @@ orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColor
 
 
 ## ----orglinearbranches-proj1--------------------------------------------------
-scaled_data <- scale_data_manual(orglinearbranches)
+centered_data <- center_data(orglinearbranches)
 
 ## First projection
 projection <- cbind(
@@ -315,7 +315,7 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             proj_scale = 1.8, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.15,
                                               axis_scaled = 10, 
                                               axis_pos_x = -0.12, 
@@ -342,7 +342,7 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.7,
                                               axis_scaled = 3, 
                                               axis_pos_x = -0.57, 
@@ -369,7 +369,7 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.6,
                                               axis_scaled = 1.5, 
                                               axis_pos_x = -0.4, 
@@ -426,7 +426,7 @@ cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControl
 
 
 ## ----cone-proj1---------------------------------------------------------------
-scaled_data <- scale_data_manual(cone)
+centered_data <- center_data(cone)
 
 ## First projection
 projection <- cbind(
@@ -435,7 +435,7 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.25,
                                               axis_scaled = 3, 
                                               axis_pos_x = -0.2, 
@@ -462,7 +462,7 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
-                            scaled_data = scaled_data, 
+                            centered_data = centered_data, 
[…]

### 2025-12-15 paper-cardinalR.Rmd
Create alter text for cardinalR..worked well

@@
 <!--script/05_gen_cluster_stat.R-->
-```{r, label = "fig-cluster-stats", fig.width=8, fig.height=6, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
-Several indices consistently suggest that $4–5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics."}
+```{r fig-cluster-stats, fig.width=8, fig.height=6, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
+Several indices consistently suggest that $4-5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics."}
 
@@
 
-```{r}
+```{r find-kmeans-clusts}
 set.seed(1214)

### 2025-12-15 paper-cardinalR.tex
Create alter text for cardinalR..worked well

@@
 \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
-Several indices consistently suggest that $4–5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics.}\label{fig:fig-cluster-stats}
+Several indices consistently suggest that $4-5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics.}\label{fig:fig-cluster-stats}
 \end{figure}

### 2025-12-15 paper-cardinalR.Rmd
Started fixing branching projections

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.15,
+                            axis_param = list(limits = 0.5,
                                               axis_scaled = 10, 
-                                              axis_pos_x = -0.12, 
-                                              axis_pos_y = -0.12, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
                                               threshold = 0.015))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.15, 0.15), 
+  plot_limits = c(-0.6, 0.4), 
   title = "a1", 
@@
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
                                               threshold = 0.01))
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.7,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-1, 1), 
   title = "a3", 

### 2025-12-15 paper-cardinalR.Rmd
Started updating the projections in the application

@@
 proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            proj_scale = 0.3, 
+                            centered_data = centered_data, 
+                            axis_param = list(limits = 0.5,
                                               axis_scaled = 0.8, 
-                                              axis_pos_x = -0.9, 
-                                              axis_pos_y = -0.9, 
-                                              threshold = 0.05))
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.02))
 
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-1.1, 1), 
+  plot_limits = c(-0.5, 0.5), 
   title = "a1", 

### 2025-12-15 paper-cardinalR.R

@@ -1999,20 +1999,20 @@ projection <- cbind(
   c(-0.54999,-0.67673,-0.03631,-0.01193))
 
 proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
+                            proj_scale = 0.3, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 0.5,
                                               axis_scaled = 0.8, 
-                                              axis_pos_x = -0.9, 
-                                              axis_pos_y = -0.9, 
-                                              threshold = 0.05))
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.02))
 
 proj_obj1[["cluster"]] <- as.character(five_clusts$cluster)
 
 five_clusts_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-1.1, 1), 
+  plot_limits = c(-0.5, 0.5), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 4,

### 2025-12-16 paper-cardinalR.Rmd
Updated application projections

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 0.55,
                                               axis_scaled = 0.8, 
-                                              axis_pos_x = -0.9, 
-                                              axis_pos_y = -0.9, 
-                                              threshold = 0.07))
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.05))
 
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-1.1, 1.2), 
+  plot_limits = c(-0.5, 0.67), 
   title = "a2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 0.5,
                                               axis_scaled = 0.8, 
-                                              axis_pos_x = -0.9, 
-                                              axis_pos_y = -0.9, 
-                                              threshold = 0.05))
+                                              axis_pos_x = -0.35, 
+                                              axis_pos_y = -0.35, 
+                                              threshold = 0))
 
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-1.1, 1.1), 
+  plot_limits = c(-0.45, 0.55), 
   title = "a3", 

### 2025-12-16 paper-cardinalR.Rmd
Fixed figure title mathemtical notation

@@
 <!--script/05_gen_cluster_stat.R-->
-```{r fig-cluster-stats, fig.width=8, fig.height=6, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
+```{r fig-cluster-stats, fig.width=8, fig.height=6, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2-10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
 Several indices consistently suggest that $4-5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics."}

### 2025-12-16 paper-cardinalR.tex
Fixed figure title mathemtical notation

@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Cluster validity metrics for solutions with $2–10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
+\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Cluster validity metrics for solutions with $2-10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
 Several indices consistently suggest that $4-5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics.}\label{fig:fig-cluster-stats}

### 2025-12-16 paper-cardinalR.Rmd
Started working on trefoil projections

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.04))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
+  plot_limits = c(-1.2, 1.2), 
   title = "b1", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -0.8, 
+                                              axis_pos_y = -0.8, 
+                                              threshold = 0.04))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
+  plot_limits = c(-1, 1.2), 
   title = "b2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.8,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
+                                              axis_pos_x = -0.8, 
+                                              axis_pos_y = -0.8, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
+  plot_limits = c(-1, 0.8), 
   title = "b3", 

### 2025-12-16 paper-cardinalR.R

@@ -1375,16 +1375,16 @@ projection <- cbind(
 proj_obj1 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.04))
 
 trefoil3d_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
+  plot_limits = c(-1.2, 1.2), 
   title = "b1", 
   cex = 2, 
   axis_text_size = 5,
@@ -1402,16 +1402,16 @@ projection <- cbind(
 proj_obj2 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -0.8, 
+                                              axis_pos_y = -0.8, 
+                                              threshold = 0.04))
 
 trefoil3d_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
+  plot_limits = c(-1, 1.2), 
   title = "b2", 
   cex = 2, 
   axis_text_size = 5,
@@ -1429,16 +1429,16 @@ projection <- cbind(
 proj_obj3 <- get_projection(projection = projection, 
                             proj_scale = 1.2, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.8,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
[…]

### 2025-12-16 paper-cardinalR.Rmd
Updated Table 1

@@
                           "is_bkg"),
-                  type = c("numeric (vector)",
-                           "numeric",
+                  type = c("integer (vector)",
+                           "integer",
                            "numeric (matrix)",

### 2025-12-16 paper-cardinalR.R

@@ -91,8 +91,8 @@ main_tb <- tibble(arg = c("n",
                           "shape",
                           "rotation",
                           "is_bkg"),
-                  type = c("numeric (vector)",
-                           "numeric",
+                  type = c("integer (vector)",
+                           "integer",
                            "numeric (matrix)",
                            "numeric (vector)",
                            "character (vector)",

### 2025-12-16 paper-cardinalR.Rmd
Removed proj scale from the get_projection function

@@
 proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.8, 
                             centered_data = centered_data, 
@@
 proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
                             centered_data = centered_data, 
@@
 proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
                             centered_data = centered_data, 
@@
 proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.2, 
                             centered_data = centered_data, 
@@
 proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 0.3, 
                             centered_data = centered_data, 
@@
 proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
                             centered_data = centered_data, 
@@
 proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
                             centered_data = centered_data, 

### 2025-12-16 scripts/additional_functions.R

@@ -46,7 +46,7 @@ center_data <- function(data) {
 
 # Get projection
 
-get_projection <- function(projection, proj_scale, centered_data, axis_param) {
+get_projection <- function(projection, centered_data, axis_param) {
 
   projected <- as.matrix(centered_data) %*% projection
   projected_df <- projected |>

### 2025-12-16 paper-cardinalR.Rmd
Updated Figure 5

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
+                            axis_param = list(limits = 9,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -7.5, 
+                                              axis_pos_y = -7.5, 
+                                              threshold = 0.6))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
+  plot_limits = c(-9, 9), 
   title = "c1", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 10,
                                               axis_scaled = 1, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              axis_pos_x = -7, 
+                                              axis_pos_y = -7, 
+                                              threshold = 0.4))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
+  plot_limits = c(-9, 9), 
   title = "c2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
+                            axis_param = list(limits = 10,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -7, 
+                                              axis_pos_y = -7, 
+                                              threshold = 0.2))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.62), 
+  plot_limits = c(-9, 9), 
   title = "c3", 

### 2025-12-16 paper-cardinalR.R

@@ -1138,16 +1138,16 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
+                            axis_param = list(limits = 9,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -7.5, 
+                                              axis_pos_y = -7.5, 
+                                              threshold = 0.6))
 
 clusteredspheres_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
+  plot_limits = c(-9, 9), 
   title = "c1", 
   cex = 2, 
   axis_text_size = 5,
@@ -1164,16 +1164,16 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 10,
                                               axis_scaled = 1, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                                              axis_pos_x = -7, 
+                                              axis_pos_y = -7, 
+                                              threshold = 0.4))
 
 clusteredspheres_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
+  plot_limits = c(-9, 9), 
   title = "c2", 
   cex = 2, 
   axis_text_size = 5,
@@ -1190,16 +1190,16 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
+                            axis_param = list(limits = 10,
+                                              axis_scaled = 1, 
[…]

### 2025-12-16 paper-cardinalR.Rmd
Updated Figure 3

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.25,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 3, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.053))
+
+cone_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-0.5, 0.7), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r cone-proj2}
+
+## Second projection
+projection <- cbind(
+    c(0.15591,0.17907,0.03198,0.04496),
+    c(-0.15376,0.10435,0.15766,0.00547))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            centered_data = centered_data, 
+                            axis_param = list(limits = 0.3,
+                                              axis_scaled = 2, 
                                               axis_pos_x = -0.2, 
                                               axis_pos_y = -0.2, 
-                                              threshold = 0.022))
-
-cone_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.25), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r cone-proj2}
-
-## Second projection
-projection <- cbind(
-    c(0.15591,0.17907,0.03198,0.04496),
-    c(-0.15376,0.10435,0.15766,0.00547))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.1,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.08, 
-                                              axis_pos_y = -0.08, 
-                                              threshold = 0.01))
+                                              threshold = 0.02))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.1, 0.12), 
+  plot_limits = c(-0.25, 0.35), 
   title = "a2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.25,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 3, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.016))
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.04))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.2), 
+  plot_limits = c(-0.7, 0.5), 
   title = "a3", 

### 2025-12-16 paper-cardinalR.Rmd
Updated Figure 4 first row

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 2,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -1.6, 
+                                              axis_pos_y = -1.6, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
+  plot_limits = c(-2, 1.8), 
   title = "a1", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.7, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.08))
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 1.6, 
+                                              axis_pos_x = -1.5, 
+                                              axis_pos_y = -1.5, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.6), 
+  plot_limits = c(-1.8, 1.8), 
   title = "a2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 1.5,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
+                                              axis_pos_x = -1.7, 
+                                              axis_pos_y = -1.7, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.75), 
+  plot_limits = c(-2, 2), 
   title = "a3", 

### 2025-12-16 paper-cardinalR.R

@@ -573,16 +573,16 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 2,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -1.6, 
+                                              axis_pos_y = -1.6, 
+                                              threshold = 0.1))
 
 pyrrect_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
+  plot_limits = c(-2, 1.8), 
   title = "a1", 
   cex = 2, 
   axis_text_size = 5,
@@ -599,16 +599,16 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.7, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.08))
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 1.6, 
+                                              axis_pos_x = -1.5, 
+                                              axis_pos_y = -1.5, 
+                                              threshold = 0.1))
 
 pyrrect_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.6), 
+  plot_limits = c(-1.8, 1.8), 
   title = "a2", 
   cex = 2, 
   axis_text_size = 5,
@@ -625,16 +625,16 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 1.5,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
[…]

### 2025-12-16 paper-cardinalR.Rmd
Updated Figure 4

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.3,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.3), 
+  plot_limits = c(-0.7, 1.4), 
   title = "b1", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.3,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.03))
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.06))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.4), 
+  plot_limits = c(-1.2, 1.2), 
   title = "b2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 1.2,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
+                                              axis_pos_x = -1.1, 
+                                              axis_pos_y = -1.1, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.62), 
+  plot_limits = c(-1.3, 1.7), 
   title = "b3", 

### 2025-12-16 paper-cardinalR.Rmd
Updated Figure 4

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -1.3, 
+                                              axis_pos_y = -1.3, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
+  plot_limits = c(-1.5, 1.5), 
   title = "c1", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -1.3, 
+                                              axis_pos_y = -1.3, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.4), 
+  plot_limits = c(-1.6, 1.6), 
   title = "c2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -1.5, 
+                                              axis_pos_y = -1.5, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
+  plot_limits = c(-1.8, 1.6), 
   title = "c3", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.5,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.3, 
+                                              axis_pos_y = -0.3, 
                                               threshold = 0.022))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-0.4, 0.5), 
   title = "d1", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 0.4,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -0.23, 
+                                              axis_pos_y = -0.23, 
+                                              threshold = 0.03))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.75), 
+  plot_limits = c(-0.3, 0.5), 
   title = "d2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.4,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
+                                              axis_pos_x = -0.43, 
+                                              axis_pos_y = -0.43, 
                                               threshold = 0.016))
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
+  plot_limits = c(-0.5, 0.4), 
   title = "d3", 

### 2025-12-16 paper-cardinalR.R

@@ -747,16 +747,16 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -1.3, 
+                                              axis_pos_y = -1.3, 
+                                              threshold = 0.1))
 
 pyrstar_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
+  plot_limits = c(-1.5, 1.5), 
   title = "c1", 
   cex = 2, 
   axis_text_size = 5,
@@ -773,16 +773,16 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -1.3, 
+                                              axis_pos_y = -1.3, 
+                                              threshold = 0.1))
 
 pyrstar_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.4), 
+  plot_limits = c(-1.6, 1.6), 
   title = "c2", 
   cex = 2, 
   axis_text_size = 5,
@@ -799,16 +799,16 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
+                            axis_param = list(limits = 1.5,
[…]

### 2025-12-16 paper-cardinalR.Rmd
Update Figure 7

@@
                             centered_data = centered_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -1.1, 
+                                              axis_pos_y = -1.1, 
+                                              threshold = 0.1))
+
+curvycylinder_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-1.3, 1.1), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycylinder-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            centered_data = centered_data, 
+                            axis_param = list(limits = 2,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -2.2, 
+                                              axis_pos_y = -2.2, 
+                                              threshold = 0.18))
+
+curvycylinder_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-2.6, 2.3), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r curvycylinder-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            centered_data = centered_data, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 1.4, 
+                                              axis_pos_x = -1.2, 
+                                              axis_pos_y = -1.2, 
+                                              threshold = 0.1))
+
+curvycylinder_proj3 <- plot_proj(
+  proj_obj = proj_obj3, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-1.4, 1.4), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$  be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$ be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
+Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
+This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
+
+```{r data-sphericalspiral}
+sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
+```
+
+```{r}
+sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
+```
+
+```{r sphericalspiral-proj1}
+centered_data <- center_data(sphericalspiral)
+
+## First projection
+projection <- cbind(
+  c(0.53506,0.00763,0.00097,-0.10901),
+  c(0.00896,0.52496,0.12597,0.08184))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 0.7,
@@
 
-curvycylinder_proj1 <- plot_proj(
+sphericalspiral_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj2}
+  plot_limits = c(-0.65, 0.7), 
+  title = "b1", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r sphericalspiral-proj2}
 
@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-curvycylinder_proj2 <- plot_proj(
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -0.65, 
+                                              axis_pos_y = -0.65, 
+                                              threshold = 0.06))
+
+sphericalspiral_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj3}
+  plot_limits = c(-0.8, 0.8), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 5,
+  is_color = FALSE)
+
+```
+
+```{r sphericalspiral-proj3}
 
@@
                                               axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-curvycylinder_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$  be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$ be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
-Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
-This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
-
-```{r data-sphericalspiral}
-sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
-```
-
-```{r}
-sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r sphericalspiral-proj1}
-centered_data <- center_data(sphericalspiral)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-sphericalspiral_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphericalspiral-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-sphericalspiral_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphericalspiral-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
+                                              threshold = 0.07))
 

### 2025-12-16 paper-cardinalR.Rmd
Updated Figure 7

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 2.5,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -2, 
+                                              axis_pos_y = -2, 
+                                              threshold = 0.2))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-2.5, 3.5), 
   title = "c1", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 2,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -2.9, 
+                                              axis_pos_y = -2.9, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.5), 
+  plot_limits = c(-3.3, 2), 
   title = "c2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 2.5,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.016))
+                                              axis_pos_x = -2.3, 
+                                              axis_pos_y = -2.3, 
+                                              threshold = 0.2))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.62), 
+  plot_limits = c(-2.8, 3.5), 
   title = "c3", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.2, 
@@
                                               axis_pos_y = -0.3, 
-                                              threshold = 0.022))
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.45, 0.73), 
+  plot_limits = c(-0.5, 1.4), 
   title = "d1", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.45,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.3, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.03))
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.1))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.46), 
+  plot_limits = c(-0.7, 2), 
   title = "d2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.45,
+                            axis_param = list(limits = 0.6,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.016))
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.07))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.35, 0.4), 
+  plot_limits = c(-0.6, 0.8), 
   title = "d3", 

### 2025-12-16 paper-cardinalR.R

@@ -1640,16 +1640,16 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 2.5,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                                              axis_pos_x = -2, 
+                                              axis_pos_y = -2, 
+                                              threshold = 0.2))
 
 conicspiral_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
+  plot_limits = c(-2.5, 3.5), 
   title = "c1", 
   cex = 2, 
   axis_text_size = 5,
@@ -1666,16 +1666,16 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 2,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -2.9, 
+                                              axis_pos_y = -2.9, 
+                                              threshold = 0.1))
 
 conicspiral_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.5), 
+  plot_limits = c(-3.3, 2), 
   title = "c2", 
   cex = 2, 
   axis_text_size = 5,
@@ -1692,16 +1692,16 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 2.5,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.016))
[…]

### 2025-12-17 extra/paper-cardinalR-code.Rmd
Added alt-txt for cone data projections

--- 2025-12-16 paper-cardinalR.Rmd
@@
 ---
-title: "cardinalR: Generating Interesting High-Dimensional Data Structures"
-abstract: >
-  Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
-High-dimensional data is characterized by multiple variables that are dependent or associated in some way, such as linear, nonlinear, clustering or anomalies.
-Here we provide new methods for generating a variety of high-dimensional structures using mathematical functions and statistical distributions organized into the R package cardinalR.
-Several example data sets are also provided.
-These will be useful for researchers to better understand how different analytical methods work and can be improved, with a special focus on nonlinear dimension reduction methods.
-This package enriches the existing toolset of benchmark datasets for evaluating algorithms.
-draft: true
-author:
-  - name: Jayani P. Gamage
-    affiliation: Monash University
-    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
-    url: https://jayanilakshika.netlify.app/
-    orcid: 0000-0002-6265-6481
-    email: jayani.piyadigamage@monash.edu
-  - name: Dianne Cook
-    affiliation: Monash University
-    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
-    url: http://www.dicook.org/
-    email: dicook@monash.edu
-    orcid: 0000-0002-3813-7155
-  - name: Paul Harrison
-    affiliation: Monash University
-    address: MGBP, BDInstitute, VIC 3800 Australia
-    email: paul.harrison@monash.edu
-    orcid: 0000-0002-3980-268X
-  - name: Michael Lydeamore
-    affiliation: Monash University
-    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
-    email: michael.lydeamore@monash.edu
-    orcid: 0000-0001-6515-827X
-  - name: Thiyanga S. Talagala
-    affiliation: University of Sri Jayewardenepura
-    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
-    url: https://thiyanga.netlify.app/
-    email: ttalagala@sjp.ac.lk
-    orcid: 0000-0002-0656-9789
-type: package
-creative_commons: CC BY
-date: "`r Sys.Date()`"
-preamble: >
-  \usepackage{amsmath}
-  \usepackage{array}
-  \usepackage{float}
-output:
-  rjtools::rjournal_article:
-    keep_md: true
-bibliography: paper-cardinalR.bib
-editor_options:
-  chunk_output_type: inline
+title: "Sample cardinalR"
+author: "Chat-GPT"
+date: "2025-09-10"
+format: pdf
 ---
-
-```{r setup, include=FALSE}
-options(repos = "https://cloud.r-project.org") ## set up CRAN mirror
-knitr::opts_chunk$set(
-  echo = FALSE,
-  cache=FALSE, 
-  message=FALSE, 
-  warning=FALSE,
-  out.width = "100%",
-  fig.pos = "!")
-
-```
-
-```{r set-seed}
-set.seed(20240412)
-```
-
-```{r load-libraries}
-library(cardinalR)
-library(tidyverse)
-library(kableExtra)
-library(geozoo)
-library(patchwork)
-library(colorspace)
-library(crosstalk)
-library(fpc)
-library(mclust)
-library(htmltools)
-```
-
-```{r}
-#| label: plot-theme
-theme_set(theme_linedraw() +
-   theme(
-     aspect.ratio = 1,
-     plot.background = element_rect(fill = 'transparent', colour = NA),
-     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-     panel.background = element_rect(fill = 'transparent', 
-                                     colour = NA),
-     panel.grid.major = element_blank(), 
-     panel.grid.minor = element_blank(), 
-     axis.title.x = element_blank(), axis.title.y = element_blank(),
-     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-     legend.background = element_rect(fill = 'transparent', 
-                                      colour = NA),
-     legend.key = element_rect(fill = 'transparent', 
-                               colour = NA),
-     legend.position = "bottom", 
-     legend.title = element_blank(), 
-     legend.text = element_text(size=4),
-     legend.key.height = unit(0.25, 'cm'),
-     legend.key.width = unit(0.25, 'cm'),
-     plot.margin = margin(0, 0, 0, 0)
-   )
-)
-```
-
-```{r}
-#| label: import-scripts
-source("scripts/additional_functions.R")
-```
-
-# Introduction
-
-Generating synthetic datasets with clearly defined geometric properties is useful for evaluating and benchmarking algorithms in various fields, such as machine learning, data mining, and computational biology.
-Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.
-There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas.
-The `geozoo` package [@barret2016] provides functions to generate standard high-dimensional data like cubes, spheres and simplexes, along with some prepared datasets.
-The `snedata` package  [@james2025] provides functions for generating common examples used in dimension reduction publications and to download benchmark data sets.
-The `splatter` package [@luke2017] is designed to simulate complex biological data, capturing field-specific nuances such as batch effects and differential expression.
-The `mlbench` package [@friedrich2024] provides access to benchmark datasets commonly associated with established classification or regression challenges.
-The `surreal` package [@james2024] implements the "Residual (Sur)Realism" algorithm [@leonard2007] to generate datasets that embed hidden images or text into residual plots, providing engaging visual demonstrations for teaching model diagnostics.
-<!--Meanwhile, the `DHARMa` package (@florian2024) adopts a simulation-based approach to create scaled quantile residuals for generalized linear (mixed) models and related frameworks, supporting model diagnostics through intuitive residuals, plots, and tests for common misspecification issues.-->
-
-The current work implemented in the `cardinalR` R package builds on these approaches.
-It provides functions to generate a more extensive set of high-dimensional data structures, allowing users to: (i) construct high-dimensional datasets based on geometric shapes, including the option to enhance dimensionality by adding controlled noise dimensions; (ii) introduce adjustable levels of background noise to these structures; and (iii) combine the shapes to produce multiple clusters.
-The user can control characteristics such as number of dimensions, shape and sample size.
-It is designed to resource researchers with synthetic datasets to evaluate the performance and interpret the fit of NLDR methods, clustering algorithms, and visualization techniques.
-These datasets can also serve as benchmark examples for exploring how different choices of algorithm parameters affect the identification or representation of cluster and manifold structures in high-dimensional spaces.
-
-The motivation for developing this package originated from our own work in studying nonlinear dimension reduction (NLDR) algorithms.
-We wanted to conduct a visualization experiment to understand perception and misperception of a variety of NLDR methods.
-This required simulated datasets with carefully controlled geometric and clustering properties.
-While some existing packages provided useful starting points, none fully supported the creation of flexible, high-dimensional data with the specific structural variations needed for our experiment.
-Developing these generators for research purposes underlies `cardinalR`, which is now a general-purpose package that should be useful for research and teaching.
-
-The example data structures are best viewed using a tour [@As85].
-These show the data as a sequence of low dimensional projections (typically 2D), providing a good sense of the shape in high dimensions.
-The interactive tour plots included in this paper are produced using the software langevitour [@paul2023].
-
-The paper is organized as follows.
-In the next section, we introduce the implementation of the `cardinalR` package on GitHub, including a demonstration of the package's key functions.
-We illustrate how a clustering data structure affects the dimension reductions in the Application section.
-Finally, we give a brief conclusion of the paper and discuss potential opportunities for the use of our data collection.
-
-# Usage
-
-The `cardinalR` package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes (A full list of available shape generators are available at <https://jayanilakshika.github.io/cardinalR/reference/index.html>.), which can then be combined into a single dataset including scaling, rotation, and translation.
-The package is available on CRAN, and the source is available on GitHub at [JayaniLakshika/cardinalR](https://github.com/JayaniLakshika/cardinalR).
-
-The main function, `gen_multicluster()`, is an all-in-one function that includes generating individual shapes, handling scaling and rotating of these shapes, and combining the result into a single unified dataset.
-This function and associated workflow allow flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods.
-Figure `r '\\@ref(fig:workflow)'` illustrates the workflow of `gen_multicluster()`.
-
-```{r workflow, out.width = "80%", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data.
-The user specifies input parameters such as the number of points ($n$), number of clusters ($k$), cluster locations, shapes, scaling, rotations, and optional background noise.
-Each cluster shape is generated by a shape generator, optionally rotated or scaled, and combined into a single dataset.
-Additional background noise can be added, and each observation is labeled by shape.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.", fig.pos="!ht"}
-
-knitr::include_graphics("figures/cardinalR_workflow.png")
-```
-
-Users can control the number of clusters (`k`), and the number of points in each cluster (`n`).
-Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (`shape`), can be scaled to adjust its spread, rotated using custom rotation matrices (`rotation`), and positioned at defined centroids (`loc`).
-The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures.
-
-The following is an example of a three-shape multiclustered dataset. The first shape is Gaussian, the second conical, and the third a cube.
-
-```{r, echo=TRUE}
-clust_data <- gen_multicluster(
-  n = c(200, 300, 500),
-  k = 3,
-  loc = matrix(c(
-    0, 0, 0, 0,
-    5, 9, 0, 0,
-    3, 4, 10, 7
-  ), nrow = 3, byrow = TRUE),
-  scale = c(3, 1, 2),
-  shape = c("gaussian", "cone", "unifcube"),
-  is_bkg = FALSE
-)
-```
-
-Here, the shapes have $200$, $300$, and $500$ points respectively (`n`), are positioned in $4\text{-}D$ space according to a location matrix, `loc`, and stretched according to the `scale`.
-The details of the individual shape generators and the noise elements are contained in the following sections.
-
-# Implementation
-
-The main function of the package is `gen_multicluster()`, which generates datasets consisting of multiple clusters with user-specified characteristics. 
-To maintain consistency across generators, the function identifies the arguments required by each chosen generator function and supplies only those arguments that are valid for that specific generator.
-This design enables the combination of cluster types with differing parameter requirements within the same dataset.
-When clusters are generated with fewer dimensions than others, the function augments the lower-dimensional clusters with additional Gaussian noise variables so that all clusters are represented in the same dimensional space.
-These noise dimensions are drawn independently from normal distributions $X \sim \mathcal{N}(\mu, \sigma^2)$, where the mean ($\mu$) is set to the average of the cluster coordinates and the standard deviation ($\sigma$) defaults to $0.2$.
-
-An optional argument, `is_bkg`, adds background noise drawn from a multivariate normal distribution centered on the dataset’s overall mean with standard deviations matching the observed spread.
-Extra arguments (`...`) can be passed to cluster generators, allowing further control over per-cluster characteristics like radius of the sphere.
-The main arguments of the `gen_multicluster()` function are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:main-tb-html)', '\\@ref(tab:main-tb-pdf)'))`.
-
-```{r}
-main_tb <- tibble(arg = c("n",
-                          "k",
-                          "loc",
-                          "scale",
-                          "shape",
-                          "rotation",
-                          "is_bkg"),
-                  type = c("integer (vector)",
-                           "integer",
-                           "numeric (matrix)",
-                           "numeric (vector)",
-                           "character (vector)",
-                           "numeric (list)",
-                           "boolean"),
-                        exp = c("Number of points in each cluster.",
-                                "Number of clusters.",
-                                "Locations/centroids of clusters.",
-                                "Scaling factors of clusters.",
-                                "Shapes of clusters.",
-                                "Rotation matrices, one per cluster.",
-                                "Background noise should exist or not.")) |> 
-    mutate(arg = gsub("_", "\\\\_", arg))
-```
-
-```{r main-tb-html, eval=knitr::is_html_output()}
-main_tb |> 
-  mutate(arg = paste0("<code>", arg, "</code>")) |> 
-  kable(caption = "The main arguments for `gen_multicluster()`.", col.names = c("Argument", "Type", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
-```
-
-```{r main-tb-pdf, eval=knitr::is_latex_output()}
-main_tb |> 
-    mutate(arg = paste0("\\texttt{", arg, "}")) |> 
-  kable(caption = "The main arguments for \\texttt{gen\\_multicluster()}.", format="latex", col.names = c("Argument", "Type", "Explanation"), booktabs = T, table.pos = "!ht", linesep = "", escape = FALSE)  |>
-  column_spec(1, width = "2cm") |>
-  column_spec(2, width = "3cm") |>
-  column_spec(3, width = "8cm")
-```
-
-## Shape generators
-
-The shape generators form the foundation of the package, providing functions to create synthetic datasets from simple, well-defined geometric forms such as cones, pyramids, spheres, grids, and branching structures.
-Each generator includes the parameter `n`, which specifies the number of points to generate.
-Some functions, such as `gen_unifcube()`, also take the dimension `p`, while others include arguments specific to the geometry (e.g., radius for spheres (`r`), width for bands (`w`)).
-If higher-dimensional data are required, additional noise dimensions can be appended after data generation using any noise generator function.
-This flexibility allows users to construct both low- and high-dimensional datasets from the same underlying structures.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:shape-tb-html)', '\\@ref(tab:shape-tb-pdf)'))` outlines these functions.
-The main arguments of the functions described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:arg-shape-tb-html)', '\\@ref(tab:arg-shape-tb-pdf)'))`.
-
-```{r}
-shape_tb <- tibble(fun = c("gen_expbranches",
-                             "gen_linearbranches",
-                             "gen_curvybranches", 
-                             "gen_orglinearbranches", 
-                             "gen_orgcurvybranches",
-                             "gen_cone",
-                             "gen_gridcube",
-                             "gen_unifcube",
-                            "gen_gaussian",
-                            "gen_longlinear",
-                            "gen_mobius",
-                            "gen_quadratic",
-                             "gen_cubic",
-                            "gen_pyrrect",
-                             "gen_pyrtri",
-                             "gen_pyrstar",
-                             "gen_pyrfrac",
-                            "gen_scurve",
-                            "gen_circle",
-                            "gen_curvycycle",
-                            "gen_unifsphere",
-                            "gen_hollowsphere",
-                            "gen_gridedsphere",
-                            "gen_clusteredspheres",
-                            "gen_hemisphere",
-                            "gen_swissroll",
-                            "gen_trefoil4d",
-                            "gen_trefoil3d",
-                             "gen_crescent",
-                            "gen_curvycylinder",
-                            "gen_sphericalspiral",
-                            "gen_helicalspiral",
-                            "gen_conicspiral",
-                            "gen_nonlinear"),
-                    parm = c("n, k",
-                             "n, k",
-                             "n, k",
-                             "n, p, k",
-                             "n, p, k",
-                             "n, p, h, ratio",
-                             "n, p",
-                             "n, p",
-                             "n, p, s",
-                             "n, p",
-                             "n",
-                             "n",
-                             "n",
-                             "n, p",
-                             "n, p",
-                             "n, p",
-                             "n, p",
-                              "n",
-                             "n, p",
-                             "n, p",
-                             "n, r",
-                             "n, p",
-                             "n",
-                             "n, k, r, loc",
-                             "n, p",
-                             "n, w",
-                             "n, steps",
-                             "n, steps",
-                             "n", 
-                             "n, h", 
-                             "n, spins", 
-                             "n", 
-                             "n, spins", 
-                             "n, hc, non_fac"),
-                      exp = c("Exponential shaped branches.",
-                               "Linear shaped branches.", 
-                               "Curvy shaped branches.",
-                               "Linear shaped branches originated in one point.",
-                               "Curvy shaped branches originated in one point.",
-                               "Cone-shaped structure.",
-                              "Cube with specified grid points along each axes.",
-                              "Cube with uniform points.",
-                              "Multivariate Gaussian cloud.",
-                              "Long linear structure.",
-                              "Möbius strip in $3\text{-}D$.",
-                              "Quadratic pattern in $2\text{-}D$.",
-                                "Cubic pattern in $2\text{-}D$.",
-                              "Rectangular-base, with a sharp or blunted apex.",
-          "Triangular-base, with a sharp or blunted apex.",
-          "Star-shaped base, with a sharp or blunted apex.",
-          "Pyramid with triangular pyramid-shaped holes.",
-          "S-curve in $3\text{-}D$.",
-          "Circle.",
-                                "Curvy cell cycle.",
-                                "Uniform ball.",
-                                "Hollow sphere.",
-                                "Grided sphere.",
-                                "Multiple small spheres within a big sphere.",
-                                "Hemisphere.",
-          "Swissroll structure.",
-          "Trefoil in $4\\text{-}D$.", 
-          "Trefoil in $3\\text{-}D$.",
-          "Crescent pattern.",
-                                "Curvy cylinder.",
-                                "Spherical spiral.",
-                                "Helical spiral.",
-                                "Conic spiral.",
-                                "Nonlinear hyperbola.")) |> 
-   mutate(fun = gsub("_", "\\\\_", fun)) |> 
-   mutate(parm = gsub("_", "\\\\_", parm))
-```
-
-```{r shape-tb-html, eval=knitr::is_html_output()}
-shape_tb |> 
-  mutate(parm = paste0("<code>", parm, "</code>")) |>
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "Overview of shape-generation functions, including their required parameters and a brief description of each geometric structure produced.
-The generators cover branching patterns, spheres, spirals, pyramids, Gaussian clouds, and other nonlinear shapes.", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, table.pos = "!ht")
-```
-
-```{r shape-tb-pdf, eval=knitr::is_latex_output()}
-shape_tb |>
-  mutate(parm = paste0("\\texttt{", parm, "}")) |>
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "Overview of shape-generation functions, including their required parameters and a brief description of each geometric structure produced.
-The generators cover branching patterns, spheres, spirals, pyramids, Gaussian clouds, and other nonlinear shapes.", format="latex", col.names = c("Function", "Arguments", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht", linesep = "")  |>
-  column_spec(1, width = "3.5cm") |>
-  column_spec(2, width = "1.5cm") |>
-  column_spec(3, width = "8cm")
-```
-
-```{r}
-arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "n_vec", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac"), 
-                          type = c("integer", "integer", "integer", "real value",
-                                   "real value", "real value", "real value", "integers", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value"),
-                        exp = c("Number of points.",
-                                "Number of clusters.",
-                                "Number of dimensions.",
-                                "Height.",
-                                "Radius tip to radius base ratio.",
-                                "Variance-covariance matrix.",
-                                "Radius.",
-                                "Sample sizes of the big and small spheres",
-                                "Number of small spheres.",
-                                "Radius of the big and small spheres.",
-                                "How far apart the small spheres are placed.",
-                                "Vertical variation",
-                                "Number of steps for the theta parameter.",
-                                "Number of loops of the spiral.",
-                                "Steepness and vertical scaling of the hyperbola.",
-                                "Strength of this sinusoidal effect.")) |> 
-   mutate(arg = gsub("_", "\\\\_", arg))
-```
-
-```{r arg-shape-tb-html, eval=knitr::is_html_output()}
-arg_branching_tb |> 
-  mutate(arg = paste0("<code>", arg, "</code>")) |>
-  kable(caption = "Argument definitions for the shape generators.
-The table lists each argument, its data type, and a description of its role in controlling geometric structure, dimensionality, scaling, curvature, spacing, and other features of the simulated high-dimensional datasets.", col.names = c("Argument", "Type (positive)", "Explanation"), escape = FALSE, table.pos = "!ht")
-```
-
-```{r arg-shape-tb-pdf, eval=knitr::is_latex_output()}
-arg_branching_tb |> 
-  mutate(arg = paste0("\\texttt{", arg, "}")) |> 
-  kable(caption = "Argument definitions for the shape generators.
-The table lists each argument, its data type, and a description of its role in controlling geometric structure, dimensionality, scaling, curvature, spacing, and other features of the simulated high-dimensional datasets.", format="latex", col.names = c("Argument", "Type (positive)", "Explanation"), booktabs = T, escape = FALSE, table.pos = "!ht")  |>
-  column_spec(1, width = "1cm") |>
-  column_spec(1, width = "2cm") |>
-  column_spec(2, width = "3cm")
-```
-
-### Branching
-
-A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology [@trapnell2014].
-We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, total number of points (`n`) generated across all branches, with points allocated approximately evenly among branches, and number of branches (`k`).
-Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
-
-The simplest structures are approximately linear branches in $2\text{-}D$, generated by the `gen_linearbranches(n, k)` function.
-These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability.
-Mathematically, each branch $i$ is defined as
-
-$$
-X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
-$$
-
-where $(x_{\text{start},i}, y_{\text{start},i})$ is the starting point of branch $i$, $\delta$ controls local jitter, and $s_i$ is the slope, initialized as
-
-$$
-s_i =
-\begin{cases}
-0.5 & i = 1, \\
--0.5 & i = 2, \\
-\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
-\end{cases}
-$$
-
-The jitter term is sampled from a one-sided uniform distribution to introduce directional variability without altering branch orientation.
-
-Branches $1$ and $2$ are initialized with fixed slopes and starting points, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
-
-```{r data-linearbranches}
-linearbranches <- gen_linearbranches(n = 1000, k = 4)
-```
-
-To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in $2\text{-}D$. Each branch follows a quadratic trajectory of the form
-
-$$
-X_1 \sim U(a_i, b_i), \quad X_2 = 0.1 X_1 + s_i X_1^2 + \epsilon, \quad \epsilon \sim U(-\delta, \delta),
-$$
-where $(a_i, b_i)$ defines the domain of the branch, $s_i$ controls curvature, and $\delta$ introduces local jitter.
-For the first two branches, the parameters are fixed to establish reference shapes: $(a_1,b_1,s_1) = (0,1,1), \quad (a_2,b_2,s_2) = (-1,0,-2)$.
-Additional branches are attached iteratively to existing structures.
-Each new branch $i$ starts at a selected point $(x_{\text{start},i}, y_{\text{start},i})$ from the current structure and extends according to
-
-$$
-X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
-$$
-
-where $s_i$ is a scale factor controlling the curvature of branch $i$.
-For the first few initial branches, $s_i$ can be fixed (e.g., $s_1 = 1, s_2 = 2$) to establish reference shapes, while for subsequent branches it is sampled from a predefined set, such as $s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}$, to control curvature magnitude and direction.
-
-```{r data-curvybranches}
-curvybranches <- gen_curvybranches(n = 1000, k = 4)
-```
-
-The `gen_expbranches(n, k)` function creates $k$ exponential branches in $2\text{-}D$, radiating from a central region. Each branch $i$ is defined as
-
-$$
-X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
-$$
-
-where $\sigma_i = (-1)^{i+1}$ alternates the sign of the exponent to produce mirror-symmetric branches.
-The parameter $s_i$ controls the steepness of branch $i$, and $\delta$ introduces small local jitter.
-
-```{r data-expbranches}
-expbranches <- gen_expbranches(n = 1000, k = 4)
-```
-
-High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) and `gen_orgcurvybranches(n, p, k)`.
-For branch
-$i$, the active coordinate pair $(i_1, i_2)$ indexes the selected $2\text{-}D$ subspace.
-When `allow_share = TRUE`, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible $\binom{p}{2}$ combinations are exhausted, after which additional branches may repeat subspaces.
-
-In both cases, branch $i$ is generated according to
-
-$$
-X_{i_1} \sim U(a_i, b_i), \quad
-X_{i_2} = f_i(X_{i_1}) + \epsilon, \quad
-\epsilon \sim N(0, \sigma^2),
-$$
-where $a_i$ and $b_i$ define the domain of the branch and $\epsilon$ introduces smooth variability in the $p\text{-}D$ space. The function $f_i(\cdot)$ determines the branch geometry:
-$$
-f_i(x) =
-\begin{cases}
-s_i x, & \text{linear branches}, \\
--s_i x^2, & \text{curvilinear branches}.
-\end{cases}
-$$
-
-The scale factor $s_i$ controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first $\binom{p}{2}$ branches, $s_i = 1$; for additional branches when $k > \binom{p}{2}$, $s_i$ is randomly drawn from the set $\{1, 1.5, 2, \dots, 8\}$.
-
-Across all branching generators, the scale parameter $s_i$ controls the strength of deviation from linearity, determining slope, curvature, or growth rate depending on the branch geometry.
-
-```{r data-orglinearbranches}
-orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
-```
-
-```{r}
-orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r orglinearbranches-proj1}
-centered_data <- center_data(orglinearbranches)
-
-## First projection
-projection <- cbind(
-    c(-0.083573,-0.019043,0.006574,-0.035774),
-    c(0.005380,0.068541,-0.030810,-0.054717))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.5,
-                                              axis_scaled = 10, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.015))
-
-orglinearbranches_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.4), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r orglinearbranches-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.85, 
-                                              axis_pos_y = -0.85, 
-                                              threshold = 0.01))
-
-orglinearbranches_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 1), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r orglinearbranches-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.85, 
-                                              axis_pos_y = -0.85, 
-                                              threshold = 0.016))
-
-orglinearbranches_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 1), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r data-orgcurvybranches}
-orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
-```
-
-```{r branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `orgcurvybranches` data.
-The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"}
-
-branchfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;",
-    orglinearbranches_lang
-  ),
-  device = "xs"
-)
-
-class(branchfig) <- c(class(branchfig), "htmlwidget")
-
-branchfig
-```
-
-```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ \\texttt{orgcurvybranches} data.
-Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
-These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the $4\\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"}
-
-orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
 
@@
 
-```{r data-bluntedcone}
-cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
-```
-
 Cone-shaped structures appear in particle dispersions, light beams, and tapering processes, where spread decreases along one axis.
@@
 
-```{r}
-cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data.
+Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
+These projections show how the conical geometry is preserved.", fig.pos="H", fig.pos="!ht", fig.width=12, fig.height=4}
+cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
 
-```{r cone-proj1}
+# Add plot label
+
+interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
+  annotation_custom(grid::textGrob(label = label,
+                                   x = unit(position[1], "npc"), y = unit(position[2], "npc"),
+                                   gp = grid::gpar(cex = cex, col=col)))
+}
+
+# Scale high-d data
+
+# Center the data by subtracting the mean of each column
+center_data <- function(data) {
+  center_values <- colMeans(data)
+  data_centered <- sweep(data, 2, center_values, FUN = "-")  # subtract means
+  data_centered
+}
+
+# Function to scale data manually
+# scale_data_manual <- function(data) {
+#   # Step 1: Center the data (mean 0)
+#   data_centered <- center_data(data)
+#
+#   # Step 2: Calculate the standard deviation of each dimension
+#   sds <- apply(data_centered, 2, sd)
+#
+#   # Step 3: Scale each dimension to have the range [0, 1]
+#   data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
+#
+#   # Step 4: Scale dimensions according to their variation
+#   # The dimension with the highest standard deviation is scaled to [-1, 1]
+#   # Other dimensions are scaled to smaller ranges based on their standard deviations
+#   max_sd <- max(sds)
+#
+#   # Normalize the standard deviations to get scaling factors
+#   scaling_factors <- sds / max_sd
+#
+#   for (i in seq_along(scaling_factors)) {
+#     data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
+#   }
+#
+#   # Combine the scaled data with the 'type' column and return as a tibble
+#   data_scaled <- as_tibble(data_scaled)
+#
+#   return(data_scaled)
+# }
+
+# Get projection
+
+get_projection <- function(projection, centered_data, axis_param) {
+
+  projected <- as.matrix(centered_data) %*% projection
+  projected_df <- projected |>
+    tibble::as_tibble(.name_repair = "unique") |>
+    dplyr::rename(c("proj1" = "...1",
+                    "proj2" = "...2")) |>
+    dplyr::mutate(ID = dplyr::row_number())
+
+  limits <- axis_param$limits
+  axis_scaled <- axis_param$axis_scaled
+  axis_pos_x <- axis_param$axis_pos_x
+  axis_pos_y <- axis_param$axis_pos_y
+  threshold <- axis_param$threshold
+
+  axes_obj <- gen_axes(
+    proj = projection * axis_scaled,
+    limits = limits,
+    axis_pos_x = axis_pos_x,
+    axis_pos_y = axis_pos_y,
+    axis_labels = names(centered_data),
+    threshold = threshold)
+
+  axes <- axes_obj$axes
+  circle <- axes_obj$circle
+
+  return(list(projected_df = projected_df,
+              axes = axes,
+              circle = circle))
+
+}
+
+# Plot projection
+plot_proj <- function(proj_obj,
+                      point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
+                      plot_limits, title, cex = 2,
+                      position = c(0.92, 0.92),
+                      axis_text_size = 3,
+                      is_color = FALSE) {
+
+  projected_df <- proj_obj$projected_df
+  axes <- proj_obj$axes
+  circle <- proj_obj$circle
+
+  if(is_color == FALSE) {
+
+    initial_plot <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2),
+        size = as.numeric(point_param[1]),
+        alpha = as.numeric(point_param[2]),
+        color = point_param[3])
+
+  } else {
+
+    projected_df <- projected_df |>
+      dplyr::mutate(cluster = proj_obj$cluster)
+
+    initial_plot <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2,
+          color = cluster),
+        size = as.numeric(point_param[1]),
+        alpha = as.numeric(point_param[2]))
+
+  }
+
+  initial_plot <- initial_plot +
+    geom_segment(
+      data=axes,
+      aes(x=x1, y=y1, xend=x2, yend=y2),
+      colour="grey70") +
+    geom_text(
+      data=axes,
+      aes(x=x2, y=y2),
+      label=rownames(axes),
+      colour="grey50",
+      size = axis_text_size) +
+    geom_path(
+      data=circle,
+      aes(x=c1, y=c2), colour="grey70") +
+    xlim(plot_limits) +
+    ylim(plot_limits) +
+    interior_annotation(title, position, cex = cex)
+
+  initial_plot
+
+}
+
+# Generate axes
+
+gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels, threshold) {
+
+  axis_scale <- limits/6
+
+  if (is.null(axis_pos_x)) {
+
+    axis_pos_x <- -2/3 * limits
+
+  }
+
+  if (is.null(axis_pos_y)) {
+
+    axis_pos_y <- -2/3 * limits
+
+  }
+
+  adj <- function(x, axis_pos) axis_pos + x * axis_scale
+  axes <- data.frame(x1 = adj(0, axis_pos_x),
+                     y1 = adj(0, axis_pos_y),
+                     x2 = adj(proj[, 1], axis_pos_x),
+                     y2 = adj(proj[, 2], axis_pos_y))
+
+  rownames(axes) <- axis_labels
+
+  ## To remove axes
+  axes <- axes |>
+    mutate(distance = sqrt((x2 - x1)^2 + (y2 - y1)^2)) |>
+    filter(distance >= threshold)
+
+  theta <- seq(0, 2 * pi, length = 50)
+  circle <- data.frame(c1 = adj(cos(theta), axis_pos_x),
+                       c2 = adj(sin(theta), axis_pos_y))
+
+  return(list(axes = axes, circle = circle))
+
+}
+
+plot_hbe <- function(error_df) {
+
+  ggplot(error_df,
+         aes(x = a1,
+             y = HBE,
+             colour = method)) +
+    geom_point(size = 0.8) +
+    geom_line(linewidth = 0.3) +
+    ylab("HBE") +
+    xlab(expression(paste("binwidth (", a[1], ")"))) +
+    theme_minimal() +
+    theme(panel.border = element_rect(fill = 'transparent'),
+          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+          axis.ticks.x = element_line(),
+          axis.ticks.y = element_line(),
+          legend.position = "none",
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7),
+          axis.title.x = element_text(size = 7),
+          axis.title.y = element_text(size = 7),
+          plot.margin = margin(0, 0, 0, 0))
+
+}
+
 centered_data <- center_data(cone)
@@
 
-```
-
-```{r cone-proj2}
-
 ## Second projection
@@
   is_color = FALSE)
-
-```
-
-```{r cone-proj3}
 
@@
 
-```
-
-```{r cone, eval=knitr::is_html_output(), fig.cap="Viewing the $3\\text{-}D$ output  of `cone` data.
-Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex.", fig.pos="!ht"}
-
-conefig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;",
-    cone_lang
-  ),
-  device = "xs"
-)
-
-class(conefig) <- c(class(conefig), "htmlwidget")
-
-conefig
-```
-
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data.
-Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
-These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.", fig.pos="!ht", fig.width=12, fig.height=4}
-
 cone_proj1 + cone_proj2 + cone_proj3 +
@@
 ```
-
-### Cube
-
-A cube structure represents uniformly or systematically distributed points within a high-dimensional hypercube, providing a useful framework for assessing how well algorithms preserve uniformity, and boundary properties in high dimensions.
-We provide a set of functions to generate high-dimensional cube structures with flexible configurations, including regular grids, and uniform random points.
-
-The function `gen_gridcube(n, p)` is a wrapper around `geozoo::cube.solid.grid()`.
-It generates a regular lattice of points in $p\text{-}D$, producing a uniform hypercube grid.
-The parameter `n` controls the approximate number of points by determining the grid resolution along each axis.
-
-```{r data-gridcube}
-gridcube <- gen_gridcube(n = 1000, p = 4)
-```
-
-By contrast, `gen_unifcube(n, p)` wraps `geozoo::cube.solid.random()`, producing uniformly distributed points within a $p\text{-}D$ cube.
-To avoid including the cube’s vertices, these points are removed after generation.
-This results in a hypercube filled with random samples rather than structured lattice points.
-
-```{r data-unifcube}
-unifcube <- gen_unifcube(n = 1000, p = 4)
-```
-
-Such cube-based structures are commonly used as benchmarks in Monte Carlo sampling, computational geometry, and density estimation, where assessing how algorithms behave under uniform or grid-like distributions is critical [@devroye1986; @niederreiter1992].
-
-### Gaussian
-
-The `gen_gaussian(n, p, s)` function generates a multivariate Gaussian cloud in $p\text{-}D$, centered at the origin with user-defined covariance structure. For 
-$i=1,\dots,n$, each observation is independently drawn from a multivariate normal distribution,  $X_i \sim N_p(\boldsymbol{0}, s)$, where $s$ is a user-defined $p \times p$ positive-definite matrix.
-
-```{r data-gau}
-gau <- gen_gaussian(n = 1000, p = 4, s = diag(4))
-```
-
-Gaussian clouds are common benchmark structures in statistics and machine learning, used in clustering, classification, and anomaly detection, with applications in image segmentation, speech recognition, and forensic analysis [@geoffrey2000].
-
-### Linear
-
-<!-- The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise.
-Each variable is formed as $X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i$, where $\text{scale}\_i \sim U(-10, 10)$ determines the orientation of the line in each dimension, $\text{shift}\_i \sim U(-300, 300)$ offsets the line to separate dimensions, and $\epsilon \sim N(0, (0.03n)^2)$ introduces Gaussian noise.
--->
-
-The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a single noisy linear trajectory.
-Let $t_i = i - 1, \quad i = 1, \dots, n$, denote a common latent index shared across all dimensions.
-For each dimension $j = 1, \dots, p)$, independent scale and shift parameters are sampled as $a_j \sim U(-10, 10), \qquad b_j \sim U(-300, 300)$.
-Gaussian noise $\varepsilon_{ij} \sim N(0, (0.03n)^2)$ is added independently across observations and dimensions.
-The observed variables are then defined as $X_{ij} = a_j \bigl(t_i + b_j + \varepsilon_{ij}\bigr), \quad i = 1, \dots, n$.cThis construction yields a single elongated linear structure embedded in $p\text{-}D$, with each dimension exhibiting a different orientation, scale, and offset.
-
-```{r data-linear}
-linear <- gen_longlinear(n = 1000, p = 4)
-```
-
-This structure appears in $p\text{-}D$ data when variation is driven by a single factor, such as time-course or sensor measurements, providing a useful test case for trajectory and regression methods [@trapnell2014].
-
-### Möbius
-
-The `gen_mobius()` function is a wrapper around `geozoo::mobius()`, designed to simplify the generation of a Möbius strip in three dimensions for use in high-dimensional diagnostic studies.
-The function returns a tibble with $n$ sampled points forming the surface of a Möbius strip.
-
-```{r data-mobius}
-mobius <- gen_mobius(n = 1000)
-```
-
-The Möbius strip structure can model twisted or cyclic surfaces in physics and engineering, such as conveyor belts, molecular structures, or optical systems with non-orientable geometries [@optica2023].
-
-### Polynomial
-
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in  space.
-To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
-These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
-We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
-
-The first is the quadratic curve of $n$ points in two dimensions.
-This is generated using `gen_quadratic(n, range)`.
-Let $range = [a, b]$.
-The independent variable is defined as $X_1 \sim U(a, b)$, and the response is generated as $X_2 = X_1 - X_1^2 + \varepsilon$, where $\varepsilon \sim U(0, 0.5)$.
-This produces a smooth parabolic arc opening downward, with vertical jitter introduced by the noise term.
-
-```{r data-quadratic}
-quadratic <- gen_quadratic(n = 1000)
-```
-
-The second is the cubic curve of $n$ points in two dimensions.
-This is generated using `gen_cubic(n, range)`.
-Let $range = [a, b]$.
-The independent variable is defined as $X_1 \sim U(a, b)$, and a raw polynomial basis of degree $3$ is applied to construct $X_2 = X_1 + X_1^2 - X_1^3 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$.
-This produces a more complex curvilinear structure than the quadratic case, with both upward and downward turning points.
-
-```{r data-cubic}
-cubic <- gen_cubic(n = 1000)
-```
-
-### Pyramid
-
-A pyramid structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr)"} else { "Figure \\@ref(fig:pyr-proj)"})`) represents data arranged around a central apex and base, useful for exploring how algorithms handle pointed or layered geometries in $p\text{-}D$ space.
-The functions provided allow users to generate pyramids with rectangular, triangular, and star-shaped bases, and sharp or blunted apexes.
-Additionally, it is possible to create a pyramid with a fractal-like internal structure, enabling the study of non-convex and sparse regions.
-
-Let $X_1, \dots, X_p$ denote the coordinates of the generated points.
-For the rectangular, triangular, and star-shaped based pyramid generator functions, the final dimension, $X_p$, encodes the height of each point and is drawn from an exponential distribution capped at the maximum height $h$.
-That is, $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).$ This distribution creates a natural skew toward smaller height values, resulting in a denser concentration of points near the pyramid's apex.
-For the star-shaped base pyramid, the final dimension is drawn from a uniform distribution.
-That is, $X_p = z \sim U(0, h)$.
-
-The remaining dimensions are based on the specific pyramid shape.
-For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`) the base shape is a rectangle whose size shrinks linearly with height.
-Let $l_x$ and $l_y$ denote the half-widths of the rectangular base in the $X_1$ and $X_2$ directions, specified via $l=(l_x,l_y)$, and let $r_t$ denote the half-width at the pyramid tip.
-At height $z\in [0,h]$, the half-widths of the rectangular cross-section are $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.
-The first three coordinates are then defined as $X_1 \sim U(-r_x(z),\ r_x(z)), \quad X_2 \sim U(-r_y(z),\ r_y(z)),\text{ and }X_3 \sim U(-r_x(z),\ r_x(z))$.
-
-```{r data-pyrrect}
-pyrrect <- gen_pyrrect(n = 1000, p = 4)
-```
-
-```{r}
-pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r pyrrect-proj1}
-centered_data <- center_data(pyrrect)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 2,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -1.6, 
-                                              axis_pos_y = -1.6, 
-                                              threshold = 0.1))
-
-pyrrect_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-2, 1.8), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrrect-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 1.6, 
-                                              axis_pos_x = -1.5, 
-                                              axis_pos_y = -1.5, 
-                                              threshold = 0.1))
-
-pyrrect_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1.8, 1.8), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrrect-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -1.7, 
-                                              axis_pos_y = -1.7, 
-                                              threshold = 0.1))
-
-pyrrect_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-2, 2), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the triangular based pyramid, `gen_pyrtri(n, p, h, l, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) b"} else { "Figure \\@ref(fig:pyr-proj) b"})`), let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$.
-That is, $r(z) = r_t + (l-r_t)z/h$.
-A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$.
-The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z)u$, and $X_3 = r(z)v.$
-
-```{r data-pyrtri}
-pyrtri <- gen_pyrtri(n = 1000, p = 4)
-```
-
-```{r}
-pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r pyrtri-proj1}
-centered_data <- center_data(pyrtri)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.1))
-
-pyrtri_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 1.4), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrtri-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -1, 
-                                              axis_pos_y = -1, 
-                                              threshold = 0.06))
-
-pyrtri_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1.2, 1.2), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrtri-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1.2,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -1.1, 
-                                              axis_pos_y = -1.1, 
-                                              threshold = 0.1))
-
-pyrtri_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1.3, 1.7), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the star based pyramid, `gen_pyrstar(n, p, h, rb)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) c"} else { "Figure \\@ref(fig:pyr-proj) c"})`), let the radius at height $z$, $r(z)$, be such that the radius scales linearly from zero (tip) to the base radius $r_b$.
-That is, $r(z) = r_b\left(1 - z/h\right)$.
-Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \pi/3, 2\pi/3, \pi, 4\pi/3, 5\pi/3\}$ and a uniformly random radial scaling factor: $\theta \sim DiscreteUniform\{0, \pi/3, \dots, 5\pi/3\}$, $r_{\text{point}} \sim \sqrt{U(0, 1)}$.
-Then, the first two coordinates are: $X_1 = r(z)r_{\text{point}}\cos(\theta)$, and $X_2 = r(z)r_{\text{point}}\sin(\theta)$.
-
-```{r data-pyrstar}
-pyrstar <- gen_pyrstar(n = 1000, p = 4)
-```
-
-```{r}
-pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r pyrstar-proj1}
-centered_data <- center_data(pyrstar)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -1.3, 
-                                              axis_pos_y = -1.3, 
-                                              threshold = 0.1))
-
-pyrstar_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1.5, 1.5), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrstar-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -1.3, 
-                                              axis_pos_y = -1.3, 
-                                              threshold = 0.1))
-
-pyrstar_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1.6, 1.6), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrstar-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -1.5, 
-                                              axis_pos_y = -1.5, 
-                                              threshold = 0.1))
-
-pyrstar_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1.8, 1.6), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For rectangular and triangular pyramids, the remaining dimensions $X_4$ to $X_{p-1}$, and for star-based pyramids $X_3$ to $X_{p-1}$, are treated as noise.
-
-Finally, for the Sierpinski-like pyramid, `gen_pyrfrac(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) d"} else { "Figure \\@ref(fig:pyr-proj) d"})`), let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points.
-The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.
-Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex.
-At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = 1/2(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$.
-This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points.
-The points remain bounded inside the convex hull of the simplex.
-The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.
-
-```{r data-pyrholes}
-pyrholes <- gen_pyrfrac(n = 1000, p = 4)
-```
-
-```{r}
-pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r pyrholes-proj1}
-centered_data <- center_data(pyrholes)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.5,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
-                                              threshold = 0.022))
-
-pyrholes_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.4, 0.5), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrholes-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.4,
-                                              axis_scaled = 1, 
-                                              axis_pos_x = -0.23, 
-                                              axis_pos_y = -0.23, 
-                                              threshold = 0.03))
-
-pyrholes_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.5), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r pyrholes-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.4,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.43, 
-                                              axis_pos_y = -0.43, 
-                                              threshold = 0.016))
-
-pyrholes_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.4), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Pyramid structures mimic tapering or layered geometries seen in architecture, crystals, and fractal-like natural patterns [@kirkby1983].
-
-```{r pyr, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$  `pyrrect`, `pyrtri`, `pyrstar`, and `pyrholes` data.
-The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
-`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.", fig.alt="Interactive langevitour of four $4\\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing dense bases, triangular spreads, branching points, and hollow regions.", fig.pos="!ht"}
-
-pyrfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;",
-    pyrrect_lang, pyrtri_lang, pyrstar_lang, pyrholes_lang
-  ),
-  device = "xs"
-)
-
-class(pyrfig) <- c(class(pyrfig), "htmlwidget")
-
-pyrfig
-```
-
-```{r pyr-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the \\texttt{pyrrect} (a1-a3), \\texttt{pyrtri} (b1-b3), \\texttt{pyrstar} (c1-c3), and \\texttt{pyrholes} (d1-d3) data.
-The \\texttt{pyrrrect} structure forms a dense rectangular base tapering to a narrow tip, while \\texttt{pytri} shows a more triangular spread with sharper edges.
-\\texttt{pyrstar} extends into multiple pointed branches radiating from a common core, and \\texttt{pyrholes} reveals hollow or open regions within an otherwise compact shape.
-These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.", fig.pos="!ht"}
-
-pyrrect_proj1 + pyrrect_proj2 + pyrrect_proj3 +
-pyrtri_proj1 + pyrtri_proj2 + pyrtri_proj3 +
-pyrstar_proj1 + pyrstar_proj2 + pyrstar_proj3 +
-pyrholes_proj1 + pyrholes_proj2 + pyrholes_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-### S-curve
-
-The S-curve is a smooth, non-linear manifold in $3\text{-}D$ space.
-Using `gen_scurve(n)`, it is defined by $X_1 = \sin(\theta), \quad X_2 \sim U(0, 2), \quad X_3 = \text{sign}(\theta)(\cos(\theta) - 1), \quad \theta \sim U(-3\pi/2, 3\pi/2).$
-
-This follows the `s_curve()` function from snedata [@james2025], itself adapted from *scikit-learn*, but differs by returning a tibble with standardized names (`x1`, `x2`, `x3`), excluding the color variable, and omitting built-in noise (which can be added separately).
-S-curve is commonly used in manifold learning and dimension reduction as benchmarks for unfolding curved structure.
-
-```{r data-scurve}
-scurve <- gen_scurve(n = 1000)
-```
-
-### Sphere
-
-Sphere-shaped structures are useful for evaluating how dimension reduction and clustering algorithms handle curved, symmetric manifolds in high-dimensional spaces.
-Throughout this section, we follow the standard mathematical terminology: a *sphere* refers to the hollow $(p-1)$-dimensional surface in $\mathbb{R}^p$, while a *ball* refers to the filled interior region.
-The functions generate a variety of spherical forms, including simple circles, uniform and hollow spheres, grid-based spheres, and complex arrangements like clustered spheres within a larger sphere.
-The first few coordinates define the main geometric form (circle, cycle, sphere, or hemisphere), while higher-dimensional embeddings are achieved by adding noise dimensions.
-Such spherical or hemispherical structures frequently appear in physical and biological systems, for example in models of celestial bodies, molecular shells, or cell membranes [@tinkham2003; @alberts2014].
-
-The simplest case, `gen_circle(n, p)` creates a unit circle in two dimensions, with the remaining dimensions forming sinusoidal extensions of the angular parameter at progressively smaller scales (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) a"} else { "Figure \\@ref(fig:sphere-proj) a"})`).
-Let a latent angle variable $\theta \sim U(0, 2\pi)$.
-Coordinates in the first two dimensions represent a perfect circle on the plane: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$ For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
-The first component is a scaling factor that decreases with the dimension index, defined as $\text{scale}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$.
-The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2)\pi/2p$.
-Each additional dimension is computed as: $X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p$.
-
-```{r data-circle}
-circle <- gen_circle(n = 1000, p = 4)
-```
-
-```{r}
-circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
-```
-
-```{r circle-proj1}
-centered_data <- center_data(circle)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-circle_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.8), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r circle-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-circle_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.7), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r circle-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.016))
-
-circle_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.53), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the one-dimensional nonlinear cycle embedded in $p\text{-}D$ space, `gen_curvycycle(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) b"} else { "Figure \\@ref(fig:sphere-proj) b"})`), let a latent angle variable $\theta \sim U(0, 2\pi)$.
-The first three dimensions define a non-circular closed curve, referred to as a `"curvy cycle"`.
-In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin.
-Additionally, $X_3 = 1/3\cos(3\theta)$ introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.
-
-Together, these define a periodic, non-trivial, closed curve in $3\text{-}D$ with internal folds that produce a more complex geometry than a standard circle or ellipse.
-For dimensions $X_4$ through $X_p$, additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
-The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from $4$ to $p$, which means that the amplitude decreases as the dimension increases.
-Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j\sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2)\pi/(2p)$.
-
-```{r data-curvycycle}
-curvycycle <- gen_curvycycle(n = 1000, p = 4)
-```
-
-```{r}
-curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
-```
-
-```{r curvycycle-proj1}
-centered_data <- center_data(curvycycle)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-curvycycle_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycycle-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.04))
-
-curvycycle_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycycle-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.016))
-
-curvycycle_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Building on simple circular structures, the `gen_unifsphere(n, r)` function function extends the idea to three dimensions by generating $n$ observations approximately uniformly distributed on the surface of a sphere of radius $r$..
-Each observation is computed from spherical coordinates, with $u \sim U(-1, 1)$ representing $\cos(\phi)$ and $\theta \sim U(0, 2\pi)$ the azimuthal angle.
-Cartesian coordinates are then defined as $$X_1 = r\sqrt{1 - u^2}\cos(\theta), \quad X_2 = r\sqrt{1 - u^2}\sin(\theta),\text{ and }X_3 = ru,$$ ensuring uniform distribution on the surface (not within) of the sphere.
-
-```{r data-unifsphere}
-unifsphere <- gen_unifsphere(n = 1000, r = 1)
-```
-
-In contrast, the `gen_hollowsphere(n, p)` function, a wrapper around `geozoo::sphere.hollow()`, generates $n$ points uniformly distributed only on the surface of the $(p-1)$-dimensional sphere embedded in $\mathbb{R}^p$.
-This results in a hollow shell-like structure with no interior points.
-For example, when $p=3$, `gen_unifsphere()` produces a solid ball in $3\text{-}D$ space, whereas `gen_hollowsphere()` produces only the spherical boundary.
-These paired structures allow controlled experiments to investigate how algorithms behave when data is concentrated throughout the full volume versus constrained to the boundary.
-
-```{r data-hollowsphere}
-hollowsphere <- gen_hollowsphere(n = 1000, p = 4)
-```
-
-In addition, the `gen_gridedsphere(n)` function constructs a $p$-dimensional dataset consisting of approximately $n$ points arranged on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
-Rather than sampling points uniformly, this function creates a deterministic grid in spherical coordinates, using $(p-1)$ angular variables: the first $(p-2)$ angles are taken from $[0, \pi]$, and the final angle from $[0, 2\pi]$.
-The number of grid points along each angular dimension is determined by decomposing $n$ into $(p-1)$ approximately equal integer factors via `gen_nproduct(n, p - 1)`.
-
-Each grid point is subsequently mapped into Cartesian space via the standard hyperspherical-to-Cartesian transformation,
-
-<!-- $$ -->
-<!-- X_i = -->
-<!-- \begin{cases} -->
-<!-- \displaystyle \prod_{j=1}^{i-1} \sin(\theta_j) , \cos(\theta_i), & 1 \le i \le p-1, \\ -->
-<!-- \displaystyle \prod_{j=1}^{p-1} \sin(\theta_j), & i = p, -->
-<!-- \end{cases} -->
-<!-- $$ -->
-
-<!-- with the convention that $\prod_{j=1}^{0} \sin(\theta_j) = 1$. -->
-
-$$
-\begin{aligned}
-X_1 &= \cos(\theta_1), \\
-X_2 &= \sin(\theta_1)\cos(\theta_2), \\
-X_3 &= \sin(\theta_1)\sin(\theta_2)\cos(\theta_3), \\
-&\;\;\vdots \\
-X_{p-1} &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\cos(\theta_{p-1}), \\
-X_p &= \sin(\theta_1)\sin(\theta_2)\cdots \sin(\theta_{p-2})\sin(\theta_{p-1}).
-\end{aligned}
-$$
-
-The result is a deterministic grid of points lying exactly on the surface of the unit $(p-1)$-sphere, without any additional noise dimensions.
-
-For more heterogeneous structures, the `gen_clusteredspheres(n, k, r, loc)` function generates one large sphere of radius $r_1$ and $k$ smaller spheres of radius $r_2$, each centered at a different random location (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) e"} else { "Figure \\@ref(fig:sphere-proj) e"})`).
-A large Uniform ball centered at the origin is created by sampling $n_1$ points uniformly on the surface of a $p\text{-}D$ sphere with a radius of $r_1$.
-The sampling is executed using the function `gen_unifsphere(n_1, r_1)`, which generates the desired points in the specified dimensional space.
-In generation of $k$ smaller Uniform balls, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
-These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
-Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for $3\text{-}D$).
-Each observation is classified by cluster, with labels such as "big" for the large central sphere and "small_1" to "small_k" for the smaller spheres.
-
-```{r data-clusteredspheres}
-clusteredspheres <- gen_clusteredspheres(n_vec = c(1000, 100), k_small = 3, r_vec = c(15, 3),
-                                         sep = 10 / sqrt(3)) |>
-  dplyr::select(-cluster)
-```
-
-```{r}
-clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "260px", height = "170px")
-```
-
-```{r clusteredspheres-proj1}
-centered_data <- center_data(clusteredspheres)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097),
-  c(0.00896,0.52496,0.12597))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 9,
-                                              axis_scaled = 1, 
-                                              axis_pos_x = -7.5, 
-                                              axis_pos_y = -7.5, 
-                                              threshold = 0.6))
-
-clusteredspheres_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-9, 9), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r clusteredspheres-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278),
-  c(-0.50493,0.14388,0.14615))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 10,
-                                              axis_scaled = 1, 
-                                              axis_pos_x = -7, 
-                                              axis_pos_y = -7, 
-                                              threshold = 0.4))
-
-clusteredspheres_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-9, 9), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r clusteredspheres-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416),
-  c(0.50329,0.07199,-0.16064))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 10,
-                                              axis_scaled = 1, 
-                                              axis_pos_x = -7, 
-                                              axis_pos_y = -7, 
-                                              threshold = 0.2))
-
-clusteredspheres_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-9, 9), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Finally, the `gen_hemisphere(n, p)` function restricts sampling to a hemisphere of a $4\text{-}D$ sphere (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) f"} else { "Figure \\@ref(fig:sphere-proj) f"})`).
-Using spherical coordinates, the azimuthal angle $\theta_1 \sim U(0, \pi)$ in the $(x_1, x_2)$ plane, while the elevation angle $\theta_2 \sim U(0, \pi)$ in the $(x_2, x_3)$ plane.
-Additionally, $\theta_3 \sim U(0, \pi/2)$ in the $(x_3, x_4)$ plane, ensuring that the points remain restricted to a hemisphere.
-The coordinates are transformed into $4\text{-}D$ Cartesian space: $$X_1 = \sin(\theta_1)\cos(\theta_2), \quad X_2 = \sin(\theta_1)\sin(\theta_2), \\\quad X_3 = \cos(\theta_1)\cos(\theta_3), \quad X_4 = \cos(\theta_1)\sin(\theta_3).$$ This produces points on one side of a $4\text{-}D$ unit sphere, effectively generating a $4\text{-}D$ hemisphere.
-
-```{r sphere, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `circle`, `curvycycle` data, and $3\\text{-}D$ `clusteredspheres` data.
-The `circle` structure forms a smooth, closed loop, while `curvycycle` shows a wavy, continuous pattern forming a twisted ring.
-The `clusteredspheres` dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.", fig.alt="Interactive langevitour of three datasets: circle and curvycycle in $4\\text{-}D$ showing closed-loop and wavy twisted-ring structures, and clusteredspheres in 3-D showing multiple compact spherical clusters with slight overlap in projections.", fig.pos="!ht", layout="l-body"}
-
-spherefig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;",
-    circle_lang, curvycycle_lang, clusteredspheres_lang),
-  #widths=c(3,3,3),
-  device = "xs"
-)
-
-class(spherefig) <- c(class(spherefig), "htmlwidget")
-
-spherefig
-```
-
-```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, \\texttt{circle} (a1-a3), \\texttt{curvycycle} (b1-b3), and, $3\\text{-}D$ \\texttt{clusteredspheres} (c1-c3).
-The \\texttt{circle} structure forms a smooth, closed loop, while \\texttt{curvycycle} shows a wavy, continuous pattern forming a twisted ring.
-The \\texttt{clusteredspheres} dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
-These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Three $2\\text{-}D$ projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.", fig.pos="!ht"}
-
-circle_proj1 + circle_proj2 + circle_proj3 +
-curvycycle_proj1 + curvycycle_proj2 + curvycycle_proj3  +
-clusteredspheres_proj1 + clusteredspheres_proj2 + clusteredspheres_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-### Swiss Roll
-
-The Swiss roll is a plane curled into $3\text{-}D$, and is a commonly used example of a nonlinear manifold.
-The `gen_swissroll(n, w)` generates points as $X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad t \sim U(0, 3\pi).$
-
-```{r data-swissroll}
-swissroll <- gen_swissroll(n = 1000, w = c(-1, 1))
-```
-
-Compared with `snedata::swiss_roll()` [@james2025], this implementation (i) samples $t$ over $[0, 3\pi]$ instead of $[1.5\pi, 4.5\pi]$, (ii) allows a flexible vertical range $w = (w_1, w_2)$ rather than fixing $z \in [0, z_{\max}]$, and (iii) returns a tibble with `x1, x2, x3` instead of adding a color variable.
-
-The Swiss roll is a classic benchmark for manifold learning, illustrating how a curved surface can be “unrolled” into lower dimensions.
-Similar spiral-like forms appear in galaxies, protein folding, and coiled materials [@dimitris2002].
-
-### Trefoil knots
-
-<!--https://laustep.github.io/stlahblog/posts/TorusKnot4D.html-->
-
-The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil)"} else { "Figure \\@ref(fig:trefoil-proj)"})`).
-The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.
-
-For the $4\text{-}D$ trefoil knot [@laurent2024], the function `gen_trefoil4d(n, steps)` generates the structure on the $3$-sphere ($S^3 \subset \mathbb{R}^4$) using two angular parameters, $\theta$ and $\phi$.
-A band of thickness around the knot path is controlled by the `steps` argument, while the number of $\theta$ and $\phi$ values is determined by the `steps` and `n` arguments, respectively (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) a"} else { "Figure \\@ref(fig:trefoil-proj) a"})`).
-The coordinates are defined as $$X_1 = \cos(\theta) \cos(\phi), \quad X_2 = \cos(\theta) \sin(\phi), \\\quad X_3 = \sin(\theta) \cos(1.5 \phi),\text{ and }X_4 = \sin(\theta) \sin(1.5 \phi),$$ where $\theta$ parameterizes the band thickness and $\phi$ parameterizes the knot trajectory.
-
-```{r data-trefoil4d}
-trefoil4d <- gen_trefoil4d(n = 500, steps = 5)
-```
-
-```{r}
-trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r trefoil4d-proj1}
-centered_data <- center_data(trefoil4d)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.022))
-
-trefoil4d_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.75, 0.7), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil4d-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-trefoil4d_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.65), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil4d-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.55, 
-                                              axis_pos_y = -0.55, 
-                                              threshold = 0.016))
-
-trefoil4d_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.62), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For the $3\text{-}D$ stereographic projection [@laurent2024], `gen_trefoil3d(n, steps)` maps each point $(X_1, X_2, X_3, X_4) \in \mathbb{R}^4$ to $(X_1', X_2', X_3') \in \mathbb{R}^3\text{ using }X_1' = X_1 / (1 - X_4), \quad X_2' = X_2 / (1 - X_4),\text{ and }X_3' = X_3 / (1 - X_4),$ excluding points where $X_4 = 1$ to avoid division by zero (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil) b"} else { "Figure \\@ref(fig:trefoil-proj) b"})`).
-
-```{r data-trefoil3d}
-trefoil3d <- gen_trefoil3d(n = 500, steps = 5)
-```
-
-The trefoil knot appears in molecular biology (DNA/protein knotting), fluid dynamics (knotted vortices), and physics (topological phases), making it a useful benchmark for testing whether dimension reduction preserves global loops and topology [@witten1985; @arsuaga2002].
-
-```{r}
-trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r trefoil3d-proj1}
-centered_data <- center_data(trefoil3d)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097),
-  c(0.00896,0.52496,0.12597))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -1, 
-                                              axis_pos_y = -1, 
-                                              threshold = 0.04))
-
-trefoil3d_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1.2, 1.2), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil3d-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278),
-  c(-0.50493,0.14388,0.14615))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1, 
-                                              axis_pos_x = -0.8, 
-                                              axis_pos_y = -0.8, 
-                                              threshold = 0.04))
-
-trefoil3d_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 1.2), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil3d-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416),
-  c(0.50329,0.07199,-0.16064))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.8,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.8, 
-                                              axis_pos_y = -0.8, 
-                                              threshold = 0.016))
-
-trefoil3d_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 0.8), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r trefoil, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `trefoil4d` and $3\\text{-}D$ `trefoil3d` data.
-The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
-In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.", fig.alt="Interactive langevitour of two trefoil knot datasets: trefoil4d in $4\\text{-}D$ showing complex twisting and looping patterns, and trefoil3d in 3-D showing a simpler compact knot-like form.", fig.pos="!ht"}
-
-trefoilfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;",
-    trefoil4d_lang, trefoil3d_lang
-  ),
-  device = "xs"
-)
-
-class(trefoilfig) <- c(class(trefoilfig), "htmlwidget")
-
-trefoilfig
-```
-
-```{r trefoil-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, \\texttt{trefoil4d} (a1-a3) and $3\\text{-}D$ \\texttt{trefoil3d} (b1-b3) data.
-The \\texttt{trefoil4d} structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
-In contrast, the \\texttt{trefoil3d} dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
-These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three $2\\text{-}D$ projections from each of two trefoil knot datasets, showing complex $4\\text{-}D$ trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.", fig.pos="!ht"}
-
-trefoil4d_proj1 + trefoil4d_proj2 + trefoil4d_proj3 + 
-  trefoil3d_proj1 + trefoil3d_proj2 + trefoil3d_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-### Trigonometric
-
-Trigonometric-based structures provide flexible ways to simulate complex curved patterns and spirals that often arise in real-world high-dimensional data, such as in biological trajectories, or physical systems (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric)"} else { "Figure \\@ref(fig:triginometric-proj)"})`).
-The main geometry is defined by the first few coordinates: crescents ($p=2$), cylinders, spirals, and helices ($p=4$).
-These structures are particularly valuable for testing how well dimension reduction and clustering algorithms preserve intricate geometric and topological features [@calladine1997; @gershenfeld2000].
-
-First, the `gen_crescent(n, p)` function generates a $p$-dimensional dataset of $n$ observations based on a $2\text{-}D$ crescent-shaped manifold with optional structured high-dimensional noise (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) a"} else { "Figure \\@ref(fig:triginometric-proj) a"})`).
-Let $\{\theta_i\}_{i=1}^n$  be a sequence of $n$ evenly spaced angles on the interval $[\pi/6, 2\pi]$, defined as $\theta_i = \frac{\pi}{6} + (i-1)\frac{2\pi - \pi/6}{n-1}, \quad i = 1,\dots,n$.
-The corresponding $2\text{-}D$ coordinates are defined by: $$X_{i1} = \cos(\theta_i), \quad X_{i2} = \sin(\theta_i).$$
-
-```{r data-crescent}
-crescent <- gen_crescent(n = 1000)
-```
-
-Second, the `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
-The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
-Let $\theta \sim U(0, 3\pi)$ represent a random angle on a circular base and $z \sim U(0, h)$ represent the height along the cylinder.
-The coordinates are defined as: $X_1 = \cos(\theta)$ (Circular base, x-axis), $X_2 = \sin(\theta)$ (Circular base, y-axis), $X_3 = z$ (Linear height), and $X_4 = \sin(z)$ (Nonlinear curvy variation along height).
-
-```{r data-curvycylinder}
-curvycylinder <- gen_curvycylinder(n = 1000, h = 10)
-```
-
-```{r}
-curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r curvycylinder-proj1}
-centered_data <- center_data(curvycylinder)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-curvycylinder_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-curvycylinder_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-curvycylinder_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For a spiraling path on a spherical surface in the first four dimensions, `gen_sphericalspiral(n, p, spins)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) c"} else { "Figure \\@ref(fig:triginometric-proj) c"})`), let $\theta \in [0, 2\pi \times \text{spins}]$  be the azimuthal angle (longitude), controls the number of spiral turns and the $\phi \in [0, \pi]$ be the polar angle (latitude), controls the vertical sweep from the north to the south pole.
-Cartesian coordinates from spherical conversion: $X_1 = \sin(\phi)\cos(\theta)$, $X_2 = \sin(\phi)\sin(\theta)$, $X_3 = \cos(\phi) + \varepsilon$, where $\varepsilon \sim U(-0.5, 0.5)$ introduces vertical jitter, and $X_4 = \theta / \max(\theta)$: a normalized progression along the spiral path.
-This generates a spherical spiral curve embedded in $4\text{-}D$ space, combining both circular and vertical movement, with gentle curvature and non-linear progression.
-
-```{r data-sphericalspiral}
-sphericalspiral <- gen_sphericalspiral(n = 1000, spins = 1)
-```
-
-```{r}
-sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r sphericalspiral-proj1}
-centered_data <- center_data(sphericalspiral)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-sphericalspiral_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "b1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphericalspiral-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
-
-sphericalspiral_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.6), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r sphericalspiral-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.016))
-
-sphericalspiral_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "b3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-For a helical spiral in four dimensions, `gen_helicalspiral(n, p)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) d"} else { "Figure \\@ref(fig:triginometric-proj) d"})`), let $\theta \in [0, 5\pi/4]$  be a sequence of angles controlling rotation around a circle.
-Cartesian coordinates; $X_1 = \cos(\theta)$: circular trajectory along the x-axis, $X_2 = \sin(\theta)$: circular trajectory along the y-axis, $X_3 = 0.05\theta + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.5, 0.5)$: linear progression (height) with vertical jitter, simulating a helix, and $X_4 = 0.1\sin(\theta)$: oscillates with $\theta$, representing a periodic "wobble" along the fourth dimension.
-
-```{r data-helicalspiral}
-helicalspiral <- gen_helicalspiral(n = 1000)
-```
-
-Similarly, the `gen_conicspiral(n, p, spins)` function generates a dataset of $n$ points forming a conical spiral in the first four dimensions of $p\text{-}D$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) e"} else { "Figure \\@ref(fig:triginometric-proj) e"})`).
-The geometry combines radial expansion, vertical elevation, and spiral deformation, simulating a structure that fans out like a $3\text{-}D$ conic helix.
-The shape is defined by parameter $\theta \in [0, 2\pi\text{spins}]$, controlling the angular progression of the spiral.
-The Archimedean spiral in the horizontal plane is represented by; $X_1 = \theta\cos(\theta)$ for radial expansion in $x$, and $X_2 = \theta\sin(\theta)$ for radial expansion in $y$.
-The growth pattern resembles a cone, with the height increasing according to $X_3 = 2\theta / \max(\theta) + \varepsilon_3$, with $\varepsilon_3 \sim U(-0.1, 0.6).$ Spiral modulation in the fourth dimension is represented by $X_4 = \theta\sin(2\theta) + \varepsilon_4$, with $\varepsilon_4 \sim U(-0.1, 0.6)$ which simulates a twisting helical component in a non-radial dimension.
-
-```{r data-conicspiral}
-conicspiral <- gen_conicspiral(n = 1000, spins = 1)
-```
-
-```{r}
-conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r conicspiral-proj1}
-centered_data <- center_data(conicspiral)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.022))
-
-conicspiral_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.65, 0.7), 
-  title = "c1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r conicspiral-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
-                                              threshold = 0.01))
-
-conicspiral_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.8, 0.5), 
-  title = "c2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r conicspiral-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.45, 
-                                              axis_pos_y = -0.45, 
-                                              threshold = 0.016))
-
-conicspiral_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.6, 0.62), 
-  title = "c3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-Finally, the `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
-The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{non\_fac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
-
-```{r data-nonlinear}
-nonlinear <- gen_nonlinear(n = 1000, hc = 1, non_fac = 0.5)
-```
-
-```{r}
-nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
-```
-
-```{r nonlinear-proj1}
-centered_data <- center_data(nonlinear)
-
-## First projection
-projection <- cbind(
-  c(0.53506,0.00763,0.00097,-0.10901),
-  c(0.00896,0.52496,0.12597,0.08184))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
-                                              threshold = 0.022))
-
-nonlinear_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.45, 0.73), 
-  title = "d1", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r nonlinear-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.45,
-                                              axis_scaled = 1.3, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.03))
-
-nonlinear_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.3, 0.46), 
-  title = "d2", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r nonlinear-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.45,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.016))
-
-nonlinear_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.35, 0.4), 
-  title = "d3", 
-  cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
-
-```
-
-```{r triginometric, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `crescent`, `curvycylinder`, `sphericalspiral`, `helicalspiral`, `conicspiral`, and `nonlinear` data.
-The `curvycylinder` shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
-The `sphericalspiral` forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
-The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.", fig.alt="Interactive langevitour of six datasets in $4\\text{-}D$: crescent, curvycylinder, sphericalspiral, helicalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral paths, conical helices, and warped $2\\text{-}D$ surfaces with smooth and sharp variations.", fig.pos="!ht"}
-
-triginometricfig <- bscols(
-  htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;",
-    curvycylinder_lang, sphericalspiral_lang, conicspiral_lang, nonlinear_lang
-  ),
-  device = "xs"
-)
-
-class(triginometricfig) <- c(class(triginometricfig), "htmlwidget")
-
-triginometricfig
-```
-
-```{r triginometric-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the \\texttt{curvycylinder} (a1-a3), \\texttt{sphericalspiral} (b1-b3), \\texttt{conicspiral} (c1-c3), and \\texttt{nonlinear} (d1-d3) data.
-The \\texttt{curvycylinder} shows a cylindrical manifold with a nonlinear twist along its height, producing smooth, continuous curvature.
-The \\texttt{sphericalspiral} forms a spiral path on a spherical surface, combining circular and vertical motion in a helical form.
-The \\texttt{conicspiral} spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The \\texttt{nonlinear} dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
-Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped $2\\text{-}D$ surfaces, illustrating curvature, density, and continuity.", fig.pos="!ht"}
-
-curvycylinder_proj1 + curvycylinder_proj2 + curvycylinder_proj3 +
-sphericalspiral_proj1 + sphericalspiral_proj2 + sphericalspiral_proj3 +
-conicspiral_proj1 + conicspiral_proj2 + conicspiral_proj3 +
-nonlinear_proj1 + nonlinear_proj2 + nonlinear_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
-```
-
-## Generate a spherical or hyperspherical hole within a structure
-
-The package provides functionality for generating datasets with spherical hole (in $2\text{-}D$/$3\text{-}D$) or, more generally, hyperspherical hole (in higher dimensions).
-These structures are valuable for evaluating how dimension reduction methods and clustering algorithms handle incomplete manifolds or missing regions of the data space.
-A hyperspherical hole introduces topological complexity: the structure remains continuous but contains excluded regions (voids) that algorithms must correctly represent in lower-dimensional embeddings.
-
-The core function `gen_hole(df, anchor, r)` removes points from a dataset that fall within a user-specified hypersphere.
-Formally, given data points ($x \in \mathbb{R}^p$), a center ($a \in \mathbb{R}^p$), and radius ($r > 0$), only points satisfying $\sqrt{\sum_{i=1}^n(x_i-a_i)^2} > r$ are retained.
-The anchor point ($a$) can either be user-specified or default to the dataset mean, and radius ($r$) is controlled by the user, with safeguards to avoid trivial or degenerate cases.
-Because it operates generically on any dataset, spherical or hyperspherical holes can be embedded in a wide range of geometric structures.
-
-Two specialized wrappers illustrate this idea.
-The function `gen_scurvehole(n, r_hole)` generates an S-curve with a spherical hole by applying `gen_hole()` to the output of `gen_scurve()`.
-This structure has been used in prior diagnostic studies of NLDR methods [@tenenbaum2000, @van2007], since it tests the ability of algorithms to capture non-linear manifolds that are not simply connected.
-The second wrapper, `gen_unifcubehole(n, p, r_hole)`, generates uniformly sampled cube data with a hyperspherical hole.
-By embedding a hyperspherical void inside a convex high-dimensional structure, this creates non-convex regions that challenge algorithms in terms of separability and neighborhood preservation.
-
-## Generate noise dimensions
-
-High-dimensional data structures often benefit from the addition of auxiliary noise dimensions, which can be used to assess the robustness of dimensionality reduction and clustering algorithms.
-The functions in this section provide flexible ways to generate random noise dimensions, ranging from purely random Gaussian variables to more structured, wavy patterns that mimic non-linear distortions in high-dimensional space.
-These functions can be applied independently or combined with other geometric structures to create complex simulated datasets.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:noise-tb-html)', '\\@ref(tab:noise-tb-pdf)'))` details these functions.
-
-```{r}
-noise_fun_tb <- tibble(fun = c("gen_noisedims",
-                            "gen_wavydims1",
-                            "gen_wavydims2",
-                            "gen_wavydims3"), 
-                        exp = c("Gaussian noise dimensions with optional mean and standard deviation.",
-                                "Wavy noise dimensions based on a user-specified theta sequence with added jitter.",
-                                "Wavy noise dimensions using polynomial transformations of an existing dimension vector.",
-                                "Wavy noise dimensions using a combination of polynomial and sine transformations based on the first three dimensions of a dataset.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r noise-tb-html, eval=knitr::is_html_output()}
-noise_fun_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR noise dimensions generation functions", col.names = c("Function", "Explanation"), escape = FALSE, table.pos = "!ht") 
-```
-
-```{r noise-tb-pdf, eval=knitr::is_latex_output()}
-noise_fun_tb |> 
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR noise dimensions generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, table.pos = "!ht")  |>
-  column_spec(1, width = "2.5cm") |>
-  column_spec(2, width = "10.5cm")
-```
-
-The `gen_noisedims(n, p, m, s)` function generates $p$ independent Gaussian noise dimensions,
-
-$$
-X_j \sim N(m_j, s_j^2), \quad j = 1, \dots, p,
-$$
-
-with odd-numbered dimensions multiplied by $-1$.
-This does not affect independence, since all noise dimensions are generated independently.
-The sign alternation is included only to avoid consistent directional drift and to ensure a symmetric appearance of noise when visualized or combined with other simulated structures.
-
-For scenarios where noise should follow a smooth wavy pattern, `gen_wavydims1(n, p, theta)` generates dimensions as
-
-$$
-X_j = \alpha_j \theta + \varepsilon_j, \quad \varepsilon_j \sim N(0, \sigma^2), \quad j = 1, \dots, p,
-$$
-
-where each dimension is scaled by a different factor $\alpha_j$, producing structured noise that oscillates along the latent parameter $\theta$, mimicking trends or trajectories observed in real-world data.
-
-The `gen_wavydims2(n, p, x_1)` function extends this approach by applying a non-linear transformation to an existing dimension vector $x_1$:
-
-$$
-X_j = \beta_j \, (-1)^{\lfloor j/2 \rfloor} \, x_1^{k_j} + \varepsilon_j, \quad j = 1, \dots, p,
-$$
-
-where $k_j$ is a randomly chosen polynomial power, $\beta_j$ is a scaling factor, and $\varepsilon_j$ is small uniform noise.
-
-Finally, `gen_wavydims3(n, p, data)` generates noise for datasets with multiple correlated dimensions.
-The first three dimensions are small perturbations of the original coordinates $(X_1, X_2, X_3)$, while higher dimensions are constructed via non-linear combinations, including polynomial and trigonometric transformations, e.g.,
-
-$$
-X_j = f_j(X_1, X_2, X_3) + \varepsilon_j, \quad j > 3,
-$$
-
-producing high-dimensional noise that preserves some geometric correlation with the base structure while introducing additional complexity.
-
-## Rotating shape generators
-
-In $p\text{-}D$ space, a rotation is an orthogonal transformation that changes the orientation of data while preserving its total variance and pairwise distances.
-The function `gen_rotation()` generates such rotation matrices for any dimension, given a list of rotation planes (axis pairs) and angles.
-
-## Multiple cluster examples
-
-By using the shape generators mentioned above, we can create various examples of multiple clusters.
-The package includes some of these examples, which are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:odd-shape-tb-html)', '\\@ref(tab:odd-shape-tb-pdf)'))`.
-
-```{r}
-odd_shapes_tb <- tibble(fun = c("make_mobiusgau",
-                                "make_multigau",
-                                "make_curvygau",
-                                "make_klink_circles",
-                                "make_chain_circles",
-                                "make_klink_curvycycle",
-                                "make_chain_curvycycle",
-                                "make_gaucircles",
-                                "make_gaucurvycycle",
-                                "make_onegrid",
-                                "make_twogrid_overlap",
-                                "make_twogrid_shift",
-                                "make_shape_para"), 
-                        exp = c(
-    "Möbius-like cluster combined with a Gaussian.",
-    "Multiple Gaussian clusters in high-dimensional space.",
-    "Curvilinear cluster with a Gaussian cluster.",
-    "K-link circular clusters (non-linear circular patterns).",
-    "Chain-like circular clusters connected sequentially.",
-    "K-link curvy cycle clusters (curvilinear loop structures).",
-    "Chain-like curvy cycle clusters connected sequentially.",
-    "Circular clusters with a Gaussian cluster in the middle.",
-    "Curvy circular clusters with a Gaussian in the middle.",
-    "Single grid in two dimensions.",
-    "Two overlapping grids.",
-    "Two grids shifted relative to each other.",
-    "Parallel shaped clusters.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r odd-shape-tb-html, eval=knitr::is_html_output()}
-odd_shapes_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR multiple clusters generation functions", col.names = c("Function", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
-```
-
-```{r odd-shape-tb-pdf, eval=knitr::is_latex_output()}
-odd_shapes_tb |> 
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR multiple clusters generation functions", format="latex", col.names = c("Function", "Explanation"), escape = FALSE, booktabs = T, linesep = "", table.pos = "!ht")  |>
-  column_spec(1, width = "3.5cm") |>
-  column_spec(2, width = "8.5cm")
-```
-
-## Additional functions
-
-The package includes various supplementary tools in addition to the shape generating functions mentioned earlier.
-These tools allow users to create background noise, randomize the rows of the data, relocate clusters, generate a vector whose product and sum are approximately equal to a target value, rotate structures, and normalize the data.
-Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:add-tb-html)', '\\@ref(tab:add-tb-pdf)'))` details these functions.
-More detailed explanations are available in [jayanilakshika.github.io/cardinalR/articles/03additionalfun](https://jayanilakshika.github.io/cardinalR/articles/03additionalfun.html).
-
-```{r}
-add_fun_tb <- tibble(fun = c("gen_bkgnoise",
-                            "randomize_rows",
-                            "relocate_clusters",
-                            "gen_nproduct",
-                            "gen_nsum",
-                            "normalize_data"), 
-                        exp = c("Adds background noise.",
-                                "Randomizes the rows of input data.",
-                                "Relocates the clusters.",
-                                "Generates a vector of positive integers whose product is approximately equal to a target value.",
-                                "Generates a vector of positive integers whose summation is approximately equal to a target value.", 
-                                "Normalizes data.")) |> 
-  mutate(fun = gsub("_", "\\\\_", fun))
-```
-
-```{r add-tb-html, eval=knitr::is_html_output()}
-add_fun_tb |> 
-  mutate(fun = paste0("<code>", fun, "</code>")) |> 
-  kable(caption = "cardinalR additional functions", col.names = c("Function", "Explanation"), linesep = "", escape = FALSE, table.pos = "!ht") 
-```
-
-```{r add-tb-pdf, eval=knitr::is_latex_output()}
-add_fun_tb |> 
-  mutate(fun = paste0("\\texttt{", fun, "}")) |> 
-  kable(caption = "cardinalR additional functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T, linesep = "", escape = FALSE, table.pos = "!ht")  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-# Application
-
-This section demonstrates how the package can be used to generate complex high-dimensional datasets, evaluate dimension reduction (DR) and clustering methods.
-The example shows how diverse geometric structures can be simulated and analyzed to assess algorithmic behavior.
-
-To illustrate how high-dimensional clustered data can be generated using `cardinalR`, we generate a dataset with five clusters in $4\text{-}D$, each representing distinct geometric characteristics: a *helical spiral* (elongated and twisted), a *hemisphere* (curved surface), a *uniform cube* (isotropic distribution), a *cone* (density gradient), and a *Gaussian* cluster (compact and spherical) (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
-Each cluster has a unique number of points and scaling factor, representing variation in cluster size and spread across the $4\text{-}D$ space.
-
-```{r gen-five-clust-data, echo=TRUE}
-
-positions <- geozoo::simplex(p=4)$points
-positions <- positions * 0.3
-
-five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
-                       loc = positions,
-                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
-                       shape = c("helicalspiral", "hemisphere", "unifcube", 
-                                 "cone", "gaussian"),
-                       rotation = NULL,
-                       is_bkg = FALSE)
-```
-
-```{r highd-data, eval=knitr::is_html_output(), fig.cap="Viewing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in $4\\text{-}D$: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht"}
-
-langevitour::langevitour(five_clusts[, -5], group = five_clusts$cluster, levelColors = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e'))
-```
-
-```{r five-clusts-projections1}
-centered_data <- center_data(five_clusts[, -5])
-
-## First projection
-projection <- cbind(
-  c(-0.59692,0.46414,0.42436,-0.10048),
-  c(-0.54999,-0.67673,-0.03631,-0.01193))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.5,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.02))
-
-proj_obj1[["cluster"]] <- as.character(five_clusts$cluster)
-
-five_clusts_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-
-```
-
-```{r five-clusts-projections2}
-## Second projection
-projection <- cbind(
-  c(0.34673,-0.35774,0.66272,0.27298),
-  c(0.24023,0.01170,-0.41886,0.72707))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.55,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.05))
-
-proj_obj2[["cluster"]] <- as.character(five_clusts$cluster)
-
-five_clusts_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.5, 0.67), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-```
-
-```{r five-clusts-projections3}
-## Third projection
-projection <- cbind(
-  c(0.47388,-0.20984,-0.56560,0.41644),
-  c(-0.51124,-0.60184,0.06463,0.36627))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.5,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.35, 
-                                              axis_pos_y = -0.35, 
-                                              threshold = 0))
-
-proj_obj3[["cluster"]] <- as.character(five_clusts$cluster)
-
-five_clusts_proj3 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.45, 0.55), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-
-```
-
-```{r highd-proj, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
-The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three $2\\text{-}D$ projections of five $4\\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=4, fig.pos="!ht"}
-
-five_clusts_proj1 + five_clusts_proj2 + five_clusts_proj3 +
-  plot_layout(ncol = 3) 
-```
-
-## Evaluating dimension reduction (DR) methods
-
-We applied six popular DR techniques to the generated dataset: Principal Component Analysis (PCA) [@jolliffe2011], tSNE, uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
-
-<!--scripts/five_clusts/02_gen_embeddings.R-->
-```{r layouts}
-
-tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds")
-
-nldr1 <- tsne_data |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour = '#636363') +
-  interior_annotation("a", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
-
-umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds")
-
-nldr2 <- umap_data |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
-  geom_point(alpha=0.1, size=1, colour = '#d95f02') +
-  interior_annotation("b", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
-
-phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds")
-
-nldr3 <- phate_data |>
-  ggplot(aes(x = PHATE1,
-             y = PHATE2))+
-  geom_point(alpha=0.1, size=1, colour = '#7570b3') +
-  interior_annotation("c", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
-
-trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
-
-nldr4 <- trimap_data |>
-  ggplot(aes(x = TriMAP1,
-             y = TriMAP2))+
-  geom_point(alpha=0.1, size=1, colour = '#e7298a') +
-  interior_annotation("d", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
-
-pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
-
-nldr5 <- pacmap_data |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1, colour = '#66a61e') +
-  interior_annotation("e", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1)
-
-pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds")
-
-nldr6 <- pca_data |>
-  ggplot(aes(x = pca1,
-             y = pca2))+
-  geom_point(alpha=0.1, size=1, colour = '#a6761d') +
-  interior_annotation("f", c(0.08, 0.93)) + 
-  theme(aspect.ratio = 1) 
-```
-
-<!--scripts/five_clusts/03_gen_rmse.R-->
-```{r}
-error_five_clust <- read_rds("data/five_clusts/five_clusts_hbe.rds")
-
-error_five_clust <- error_five_clust |>
-  mutate(a1 = round(a1, 2)) |>
-  filter(b1 >= 5) |>
-  group_by(method, a1) |>
-  filter(HBE == min(HBE)) |>
-  ungroup()
-
-error_plot_five_clust <- plot_hbe(error_five_clust) +
-  scale_x_continuous(breaks = sort(unique(error_five_clust$a1))[c(1, 5, 9, 13, 17, 21, 25)]) +
-  scale_color_manual(values=c("tSNE" = '#636363', "UMAP" = '#d95f02', "PHATE" = '#7570b3', "TriMAP" = '#e7298a', "PaCMAP" = '#66a61e', "PCA" = '#a6761d'))
-```
-
-```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
-Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
-Layout f is universally poor.
-Layouts a and b are universally optimal.
-Layout b shows six well-separated clusters and layout a shows close clusters, thus layout a is the best choice.", fig.alt="Composite plot assessing six NLDR layouts of five clusters data using HBE over varying binwidths, showing tSNE, UMAP, PAHTE, TriMAP, PaCMAP, and PCA.
-Layouts a and b are optimal, f is poor, with color-coded lines matching the scatterplots of each layout.", fig.pos="!ht"}
-
-error_plot_five_clust + wrap_plots(nldr1, nldr2, nldr3, 
-                                    nldr4, nldr5, nldr6, ncol = 2)
-```
-
-To assess their performance, we computed the hexbin error (HBE) between the observed high-dimensional data and the fitted values, defined as the high-dimensional mappings of the bin centroids [@gamage2025c].
-A lower HBE indicates that the method better preserves the high-dimensional structure in its low-dimensional embedding.
-
-As shown in Figure \@ref(fig:fig-nldr-layouts), tSNE (Figure \@ref(fig:fig-nldr-layouts) a) achieved the lowest HBE across bin widths (mostly tiny), indicating high preservation of both local and global structures.
-Its layout displays well-separated clusters with minimal inter-cluster distances, making it the most faithful representation of the underlying data structure.
-UMAP and PaCMAP (Figure \@ref(fig:fig-nldr-layouts) b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \@ref(fig:fig-nldr-layouts) c) show non-linear cluster structures irrespective of the original structure.
-Also, TriMAP (Figure \@ref(fig:fig-nldr-layouts) d) has high HBE, and show three clusters with small distances.
-PCA (Figure \@ref(fig:fig-nldr-layouts) f) failed to capture the non-linear geometry, leading to the highest HBE.
-
-## Benchmarking clustering algorithms
-
-To further evaluate the structure of the generated data, we benchmarked three clustering algorithms: **$k$-means** [Chapter 20 of @boehmke2019], **hierarchical** [@murtagh2012], and **model-based clustering** [@chris2002; @scrucca2023] using the simulated dataset.
-Model-based clustering performed the `"EII"` covariance structure.
-Under this parameterization, clusters are spherical with equal volume and equal shape, and no orientation parameter is estimated.
-Cluster validity statistics were computed using the `cluster.stats()` function from the `fpc` package [@christian2024].
-
-<!--script/05_gen_cluster_stat.R-->
-```{r fig-cluster-stats, fig.width=8, fig.height=6, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2-10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
-Several indices consistently suggest that $4-5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics."}
-
-# Examine the cluster stats
-all_stats <- read_rds("data/five_clusts/cluster_stat.rds")
-all_stats |>
-  pivot_longer(within.cluster.ss:sindex, names_to = "stat", values_to = "value") |>
-  ggplot(aes(x=cl, y=value, colour=method)) +
-    geom_line() +
-    facet_wrap(~stat, ncol=3, scales="free_y") +
-    xlab("Number of clusters") +
-    ylab("") +
-    scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3')) +
-    theme_minimal() +
-    theme(panel.border = element_rect(fill = 'transparent'),
-          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-          axis.ticks.x = element_line(),
-          axis.ticks.y = element_line(),
-          legend.position = "bottom",
-          axis.text.x = element_text(size = 7),
-          axis.text.y = element_text(size = 7),
-          axis.title.x = element_text(size = 7),
-          axis.title.y = element_text(size = 7),
-          plot.margin = margin(0, 0, 0, 0))
-```
-
-Figure \@ref(fig:fig-cluster-stats) shows a selection of cluster metrics for $2-10$ clusters for each of the methods, $k$-means, hierarchical, and model-based.
-As is typical, the suggestion of the best solution varies between cluster statistics.
-Although the metrics differ in their preferences, several show consistent support for a $4-5$ cluster solution.
-Pearson gamma (`pearsongamma`) increases sharply up to five clusters before levelling off, Calinski–Harabasz index (`ch`) increases sharply from 4 to 5 clusters and Dunn (`dunn2`) has a maximum at 5 for two methods and at 4 for $k$-means.
-All of these are interpreted as higher is better.
-With the other three, lower is better.
-WB ratio (`wbratio`) and within-cluster sum of squares (`within.cluster.ss`) steadily decline with number of clusters, possibly elbowing around 5 clusters.
-The S-index (`sindex`) is optimised at 4 clusters for $k$-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.
-Overall, $k$-means performs slightly better than the hierarchical and model-based clustering across most metrics and number of clusters.
-
-```{r find-kmeans-clusts}
-set.seed(1214)
-
-## k-means with  four clusters
-kcl4 <- kmeans(five_clusts[, -5], centers = 4, nstart = 20)$cluster
-
-## k-means with  five clusters
-kcl5 <- kmeans(five_clusts[, -5], centers = 5, nstart = 20)$cluster
-
-```
-
-```{r highd-data-clusters-algo-html, eval=knitr::is_html_output(), fig.cap="Views of the five-cluster synthetic dataset coloured by the $k$-means four- and five-cluster solutions.
-These show how the five true structures: helical spiral, hemisphere, uniform cube, blunted cone, and Gaussian-shaped cluster appear under $k$-means partitions.
-Both the $4$- and $5$-cluster solutions mix colours, revealing substantial misclassification and limited ability of $k$-means to recover the underlying geometric structure.", fig.pos="!ht", layout = "l-body"}
-
-four_km_langevitour <- langevitour::langevitour(
-  five_clusts[, -5],
-  group = kcl4,
-  levelColors = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e'),
-  width = "350px",
-  height = "300px"
-)
-
-five_km_langevitour <- langevitour::langevitour(
-  five_clusts[, -5],
-  group = kcl5,
-  levelColors = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e'),
-  width = "350px",
-  height = "300px"
-)
-
-km_results <- browsable(
-  div(
-    style = "
-      display: flex;
-      justify-content: space-between;
-      gap: 5px;
-    ",
-    div(style = "flex: 1;", four_km_langevitour),
-    div(style = "flex: 1;", five_km_langevitour)
-  )
-)
-
-class(km_results) <- c(class(km_results), "htmlwidget")
-
-km_results
-```
-
-```{r five-4clusts-projections1}
-centered_data <- center_data(five_clusts[, -5])
-
-## First projection
-projection <- cbind(
-  c(-0.59692,0.46414,0.42436,-0.10048),
-  c(-0.54999,-0.67673,-0.03631,-0.01193))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.5,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.02))
-
-proj_obj1[["cluster"]] <- as.character(kcl4)
-
-five_clusts_proj1_1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-
-```
-
-```{r five-4clusts-projections2}
-## Second projection
-projection <- cbind(
-  c(0.34673,-0.35774,0.66272,0.27298),
-  c(0.24023,0.01170,-0.41886,0.72707))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.55,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.05))
-
-proj_obj2[["cluster"]] <- as.character(kcl4)
-
-five_clusts_proj2_1 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.5, 0.67), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-```
-
-```{r five-4clusts-projections3}
-## Third projection
-projection <- cbind(
-  c(0.47388,-0.20984,-0.56560,0.41644),
-  c(-0.51124,-0.60184,0.06463,0.36627))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.5,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.35, 
-                                              axis_pos_y = -0.35, 
-                                              threshold = 0))
-
-proj_obj3[["cluster"]] <- as.character(kcl4)
-
-five_clusts_proj3_1 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.45, 0.55), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-
-```
-
-```{r five-5clusts-projections1}
-centered_data <- center_data(five_clusts[, -5])
-
-## First projection
-projection <- cbind(
-  c(-0.59692,0.46414,0.42436,-0.10048),
-  c(-0.54999,-0.67673,-0.03631,-0.01193))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.5,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.02))
-
-proj_obj1[["cluster"]] <- as.character(kcl5)
-
-five_clusts_proj1_2 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.5, 0.5), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-
-```
-
-```{r five-5clusts-projections2}
-## Second projection
-projection <- cbind(
-  c(0.34673,-0.35774,0.66272,0.27298),
-  c(0.24023,0.01170,-0.41886,0.72707))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.55,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.05))
-
-proj_obj2[["cluster"]] <- as.character(kcl5)
-
-five_clusts_proj2_2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.5, 0.67), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-```
-
-```{r five-5clusts-projections3}
-## Third projection
-projection <- cbind(
-  c(0.47388,-0.20984,-0.56560,0.41644),
-  c(-0.51124,-0.60184,0.06463,0.36627))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.5,
-                                              axis_scaled = 0.8, 
-                                              axis_pos_x = -0.35, 
-                                              axis_pos_y = -0.35, 
-                                              threshold = 0))
-
-proj_obj3[["cluster"]] <- as.character(kcl5)
-
-five_clusts_proj3_2 <- plot_proj(
-  proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.45, 0.55), 
-  title = "a3", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
-  theme(legend.position = "none")
-
-```
-
-```{r highd-proj-clust-algo-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data coloured by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
-The intermixing of colours within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.", fig.width=12, fig.height=8, fig.pos="!ht"}
-
-five_clusts_proj1_1 + five_clusts_proj2_1 + five_clusts_proj3_1 +
-  five_clusts_proj1_2 + five_clusts_proj2_2 + five_clusts_proj3_2 +
-  plot_layout(ncol = 3) 
-```
-
-`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` shows the four- and five-cluster $k$-means solutions, with cluster id used to colour the points.
-Neither solution captures the geometric nature of the true clusters, but they are both reasonable partitions of the data.
-To examine either one, it is best to subset to a single cluster to view in the tour.
-With each solution, the five original shapes are each split by the clustering.
-More than 5 clusters would be needed to better capture the original shapes.
-
-# Conclusion
-
-The `cardinalR` package introduces a flexible framework for generating high-dimensional data structures with well-defined geometric properties.
-It addresses an important need in the evaluation of clustering, machine learning, and DR methods by enabling the construction of customized datasets with interpretable structures, noise characteristics, and clustering arrangements.
-In this way, `cardinalR` complements existing packages such as `geozoo`, `snedata`, and `mlbench`, while extending the scope to higher dimensions and more complex shapes.
-
-The included structures cover a wide range of diagnostic settings.
-Branching shapes facilitate the study of continuity and topological preservation, the S-curve with a hole allows investigation of incomplete manifolds, and clustered spheres assess separability on curved surfaces.
-The Möbius strip introduces challenges from non-orientable geometry, while gridded cubes and pyrholes test spatial regularity and clustering in sparse, non-convex regions.
-
-These structures are designed to support not only algorithm diagnostics, but also teaching high-dimensional concepts, benchmarking reproducibility, and evaluating hyper-parameter sensitivity.
-By allowing users to adjust dimensionality, sample size, noise, and clustering properties, the package promotes transparent experimentation and comparative model evaluation.
-Together, these capabilities make `cardinalR` a versatile tool for generating interpretable, high-dimensional datasets that advance research, teaching, and evaluation of data-analytic methods.
-
-Future extensions of `cardinalR` may include biologically inspired or application-driven data structures that would further broaden its utility in domains such as bioinformatics, forensic science, and spatial analysis.
-
-# Acknowledgements
-
-The source material for this paper is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
-This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.
-These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].

### 2025-12-17 paper-cardinalR.Rmd
Added alt-txt for cone data projections

@@
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
-These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.", fig.pos="!ht", fig.width=12, fig.height=4}
+These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt=" A 2D scatterplot shows the projection of a high-dimensional cone-shaped point cloud.
+The horizontal and vertical axes are unlabeled numeric coordinates, each spanning a roughly symmetric range around zero (for example, from about -1 to 1).
+Hundreds of points form a dense cluster near one end of the plot, corresponding to the cone’s apex, with density gradually decreasing as points extend outward in a flared, cone-like or triangular region.
+There are no colours, shapes, or facets used; all points share the same appearance, and the key visible pattern is the concentrated tip transitioning smoothly into a wider spread of points, illustrating a cone structure in projection.", fig.pos="!ht", fig.width=12, fig.height=4}
 

### 2025-12-17 paper-cardinalR.tex
Updated Table 1
Added alt-txt for cone data projections

@@
 \midrule
-\texttt{n} & numeric (vector) & Number of points in each cluster.\\
-\texttt{k} & numeric & Number of clusters.\\
+\texttt{n} & integer (vector) & Number of points in each cluster.\\
+\texttt{k} & integer & Number of clusters.\\
 \texttt{loc} & numeric (matrix) & Locations/centroids of clusters.\\
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ \texttt{cone} data.
+\includegraphics[width=1\linewidth,alt={ A 2D scatterplot shows the projection of a high-dimensional cone-shaped point cloud.
+The horizontal and vertical axes are unlabeled numeric coordinates, each spanning a roughly symmetric range around zero (for example, from about -1 to 1).
+Hundreds of points form a dense cluster near one end of the plot, corresponding to the cone’s apex, with density gradually decreasing as points extend outward in a flared, cone-like or triangular region.
+There are no colours, shapes, or facets used; all points share the same appearance, and the key visible pattern is the concentrated tip transitioning smoothly into a wider spread of points, illustrating a cone structure in projection.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ \texttt{cone} data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.

### 2025-12-17 paper-cardinalR.R

@@ -521,7 +521,7 @@ cone_proj3 <- plot_proj(
 # conefig
 
 
-## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex. These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the 3-D cone dataset, showing points concentrated near the tip and the linear decrease in radius toward the apex.", fig.pos="!ht", fig.width=12, fig.height=4----
+## ----cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex. These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt=" A 2D scatterplot shows the projection of a high-dimensional cone-shaped point cloud. The horizontal and vertical axes are unlabeled numeric coordinates, each spanning a roughly symmetric range around zero (for example, from about -1 to 1). Hundreds of points form a dense cluster near one end of the plot, corresponding to the cone’s apex, with density gradually decreasing as points extend outward in a flared, cone-like or triangular region. There are no colours, shapes, or facets used; all points share the same appearance, and the key visible pattern is the concentrated tip transitioning smoothly into a wider spread of points, illustrating a cone structure in projection.", fig.pos="!ht", fig.width=12, fig.height=4----
 
 cone_proj1 + cone_proj2 + cone_proj3 +
   plot_layout(ncol = 3, guides = "collect") 

### 2025-12-17 extra/paper-cardinalR-code.Rmd
Updated branch projections

@@
 
-### Cone
+# Application
 
-To simulate a cone-shaped structure in arbitrary dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cone)"} else { "Figure \\@ref(fig:cone-proj)"})`), we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+This section demonstrates how the package can be used to generate complex high-dimensional datasets, evaluate dimension reduction (DR) and clustering methods.
+The example shows how diverse geometric structures can be simulated and analyzed to assess algorithmic behavior.
 
-This function generates $n$ points in $p\text{-}D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip.
-Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, truncated to the interval $[0, h]$, producing a higher density of points near the tip.
-At each height $X_p$, the radius of the cross-section increases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio} \in [0, 1]$ and $r_{\text{max}} = 1$.
+To illustrate how high-dimensional clustered data can be generated using `cardinalR`, we generate a dataset with five clusters in $4\text{-}D$, each representing distinct geometric characteristics: a *helical spiral* (elongated and twisted), a *hemisphere* (curved surface), a *uniform cube* (isotropic distribution), a *cone* (density gradient), and a *Gaussian* cluster (compact and spherical) (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data)"} else { "Figure \\@ref(fig:highd-proj)"})`).
+Each cluster has a unique number of points and scaling factor, representing variation in cluster size and spread across the $4\text{-}D$ space.
 
-For each point, a direction in the first $p-1$ dimensions is sampled uniformly on a $(p-1)$-dimensional hypersphere using generalized spherical coordinates.
-The radial coordinates are scaled by the height-dependent radius $r$, producing the conical taper.
-In three dimensions ($p = 3$), this results in a classical $3\text{-}D$ cone, while for $p > 3$, additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
-
-Cone-shaped structures appear in particle dispersions, light beams, and tapering processes, where spread decreases along one axis.
-They are also used to benchmark clustering and dimensionality reduction methods [@hadsell2006].
-
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data.
-Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
-These projections show how the conical geometry is preserved.", fig.pos="H", fig.pos="!ht", fig.width=12, fig.height=4}
-cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
+```{r highd-proj, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three $2\\text{-}D$ projections of five $4\\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=4, fig.pos="!ht"}
 
@@
 
-centered_data <- center_data(cone)
+positions <- geozoo::simplex(p=4)$points
+positions <- positions * 0.3
+
+five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
+                       loc = positions,
+                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
+                       shape = c("helicalspiral", "hemisphere", "unifcube", 
+                                 "cone", "gaussian"),
+                       rotation = NULL,
+                       is_bkg = FALSE)
+
+centered_data <- center_data(five_clusts[, -5])
 
@@
 projection <- cbind(
-  c(0.22417,-0.02943,-0.02870,0.08418),
-  c(-0.02715,0.18019,0.05053,0.15254))
+  c(-0.59692,0.46414,0.42436,-0.10048),
+  c(-0.54999,-0.67673,-0.03631,-0.01193))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 3, 
+                            axis_param = list(limits = 0.5,
+                                              axis_scaled = 0.8, 
                                               axis_pos_x = -0.4, 
                                               axis_pos_y = -0.4, 
-                                              threshold = 0.053))
+                                              threshold = 0.02))
 
-cone_proj1 <- plot_proj(
+proj_obj1[["cluster"]] <- as.character(five_clusts$cluster)
+
+five_clusts_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.7), 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.5, 0.5), 
   title = "a1", 
   cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
 
@@
 projection <- cbind(
-    c(0.15591,0.17907,0.03198,0.04496),
-    c(-0.15376,0.10435,0.15766,0.00547))
+  c(0.34673,-0.35774,0.66272,0.27298),
+  c(0.24023,0.01170,-0.41886,0.72707))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 2, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.02))
+                            axis_param = list(limits = 0.55,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.05))
 
-cone_proj2 <- plot_proj(
+proj_obj2[["cluster"]] <- as.character(five_clusts$cluster)
+
+five_clusts_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.35), 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.5, 0.67), 
   title = "a2", 
   cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
 
@@
 projection <- cbind(
-    c(-0.13548,0.13832,0.05991,-0.13544),
-    c(0.08479,-0.10719,-0.01572,-0.20123))
+  c(0.47388,-0.20984,-0.56560,0.41644),
+  c(-0.51124,-0.60184,0.06463,0.36627))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.04))
+                            axis_param = list(limits = 0.5,
+                                              axis_scaled = 0.8, 
+                                              axis_pos_x = -0.35, 
+                                              axis_pos_y = -0.35, 
+                                              threshold = 0))
 
-cone_proj3 <- plot_proj(
+proj_obj3[["cluster"]] <- as.character(five_clusts$cluster)
+
+five_clusts_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.5), 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.45, 0.55), 
   title = "a3", 
   cex = 2, 
-  axis_text_size = 5,
-  is_color = FALSE)
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
+  theme(legend.position = "none")
 
-cone_proj1 + cone_proj2 + cone_proj3 +
-  plot_layout(ncol = 3, guides = "collect") 
+five_clusts_proj1 + five_clusts_proj2 + five_clusts_proj3 +
+  plot_layout(ncol = 3) 
 ```
+
+```

### 2025-12-17 paper-cardinalR.Rmd
Updated branch projections

@@
                                               axis_pos_y = -0.5, 
-                                              threshold = 0.015))
+                                              threshold = 0.05))
 
@@
   cex = 2, 
-  axis_text_size = 5,
+  axis_text_size = 4,
   is_color = FALSE)
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
+                            axis_param = list(limits = 0.9,
+                                              axis_scaled = 1, 
                                               axis_pos_x = -0.85, 
                                               axis_pos_y = -0.85, 
-                                              threshold = 0.01))
+                                              threshold = 0.05))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.9,
                                               axis_scaled = 1.5, 
@@
                                               axis_pos_y = -0.85, 
-                                              threshold = 0.016))
+                                              threshold = 0.1))
 
@@
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
-These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="Three $2\\text{-}D$ projections of the $4\\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.", fig.pos="!ht"}
+These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="The figure shows a 2-D scatter plot of projected branch structures in a high-dimensional space.
+The horizontal axis represents one coordinate (first index of the active pair) and the vertical axis represents the other coordinate (second index of the active pair), each spanning a moderate, symmetric range around zero (for example, roughly −2 to 2).
+Points are grouped visually into several distinct linear or gently curving branches, each corresponding to a different active 2-D subspace; these branches appear as separate streaks or arcs radiating through the plane.
+Colour (or another visual attribute such as line/point grouping) is used to distinguish branches that occupy different subspaces, and some branches may overlap or run close together where subspaces are shared.
+The main pattern is a set of well-separated, structured branches with little noise, illustrating how high-dimensional linear or curvy branches project into 2-D coordinate pairs.", fig.pos="!ht"}
 
@@
 ```{r cone, eval=knitr::is_html_output(), fig.cap="Viewing the $3\\text{-}D$ output  of `cone` data.
-Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="Interactive langevitour of the cone dataset in 3-D, showing points concentrated near the tip with radius decreasing linearly toward the apex.", fig.pos="!ht"}
+Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="A 3D scatterplot shows a cone-shaped cloud of points, with two horizontal axes representing the first two coordinates (x and y) and the vertical axis representing height (z).
+The horizontal axes span roughly symmetric ranges centered around zero, while height increases from a low base up to a maximum near the tip of the cone.
+At lower heights the points spread widely in the horizontal plane, and as height increases the radial spread decreases smoothly, forming a narrowing cone.
+Points are evenly distributed around the cone’s central axis, reflecting that directions in the first two dimensions are sampled uniformly on a circle (a 2D hypersphere), and their distance from the center is scaled by a height-dependent radius.
+The resulting pattern is a dense, symmetric conical taper with no distinct outliers or clusters beyond the cone surface.", fig.pos="!ht"}
 

### 2025-12-17 paper-cardinalR.tex
Updated branch projections

@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of the $4\text{-}D$ orglinearbranches dataset, showing different angles of the linear branches to illustrate the branching structure.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ \texttt{orgcurvybranches} data.
+\includegraphics[width=1\linewidth,alt={The figure shows a 2-D scatter plot of projected branch structures in a high-dimensional space.
+The horizontal axis represents one coordinate (first index of the active pair) and the vertical axis represents the other coordinate (second index of the active pair), each spanning a moderate, symmetric range around zero (for example, roughly −2 to 2).
+Points are grouped visually into several distinct linear or gently curving branches, each corresponding to a different active 2-D subspace; these branches appear as separate streaks or arcs radiating through the plane.
+Colour (or another visual attribute such as line/point grouping) is used to distinguish branches that occupy different subspaces, and some branches may overlap or run close together where subspaces are shared.
+The main pattern is a set of well-separated, structured branches with little noise, illustrating how high-dimensional linear or curvy branches project into 2-D coordinate pairs.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ \texttt{orgcurvybranches} data.
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.

### 2025-12-17 paper-cardinalR.R

@@ -319,7 +319,7 @@ proj_obj1 <- get_projection(projection = projection,
                                               axis_scaled = 10, 
                                               axis_pos_x = -0.5, 
                                               axis_pos_y = -0.5, 
-                                              threshold = 0.015))
+                                              threshold = 0.05))
 
 orglinearbranches_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
@@ -327,7 +327,7 @@ orglinearbranches_proj1 <- plot_proj(
   plot_limits = c(-0.6, 0.4), 
   title = "a1", 
   cex = 2, 
-  axis_text_size = 5,
+  axis_text_size = 4,
   is_color = FALSE)
 
 
@@ -341,11 +341,11 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
+                            axis_param = list(limits = 0.9,
+                                              axis_scaled = 1, 
                                               axis_pos_x = -0.85, 
                                               axis_pos_y = -0.85, 
-                                              threshold = 0.01))
+                                              threshold = 0.05))
 
 orglinearbranches_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
@@ -353,7 +353,7 @@ orglinearbranches_proj2 <- plot_proj(
   plot_limits = c(-1, 1), 
   title = "a2", 
   cex = 2, 
-  axis_text_size = 5,
+  axis_text_size = 4,
   is_color = FALSE)
 
 
@@ -367,11 +367,11 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.9,
                                               axis_scaled = 1.5, 
[…]

### 2025-12-17 paper-cardinalR.Rmd
Updated Figure 4 projections

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 1.6, 
+                            axis_param = list(limits = 1.7,
+                                              axis_scaled = 1.8, 
                                               axis_pos_x = -1.5, 
                                               axis_pos_y = -1.5, 
-                                              threshold = 0.1))
+                                              threshold = 0.2))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 1.5, 
+                            axis_param = list(limits = 1.7,
+                                              axis_scaled = 1.8, 
                                               axis_pos_x = -1.7, 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1.5, 
+                            axis_param = list(limits = 1.1,
+                                              axis_scaled = 1.7, 
                                               axis_pos_x = -1, 
                                               axis_pos_y = -1, 
-                                              threshold = 0.06))
+                                              threshold = 0.1))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1.2,
+                            axis_param = list(limits = 1.3,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -1.1, 
-                                              axis_pos_y = -1.1, 
+                                              axis_pos_x = -1.05, 
+                                              axis_pos_y = -1.05, 
                                               threshold = 0.1))
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 1.4,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -1.3, 
-                                              axis_pos_y = -1.3, 
+                                              axis_pos_x = -1.2, 
+                                              axis_pos_y = -1.2, 
                                               threshold = 0.1))
@@
                                               axis_pos_y = -0.43, 
-                                              threshold = 0.016))
+                                              threshold = 0.05))
 

### 2025-12-17 paper-cardinalR.R

@@ -599,11 +599,11 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 1.6, 
+                            axis_param = list(limits = 1.7,
+                                              axis_scaled = 1.8, 
                                               axis_pos_x = -1.5, 
                                               axis_pos_y = -1.5, 
-                                              threshold = 0.1))
+                                              threshold = 0.2))
 
 pyrrect_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
@@ -625,8 +625,8 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 1.5, 
+                            axis_param = list(limits = 1.7,
+                                              axis_scaled = 1.8, 
                                               axis_pos_x = -1.7, 
                                               axis_pos_y = -1.7, 
                                               threshold = 0.1))
@@ -686,11 +686,11 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1.5, 
+                            axis_param = list(limits = 1.1,
+                                              axis_scaled = 1.7, 
                                               axis_pos_x = -1, 
                                               axis_pos_y = -1, 
-                                              threshold = 0.06))
+                                              threshold = 0.1))
 
 pyrtri_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
@@ -712,10 +712,10 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1.2,
+                            axis_param = list(limits = 1.3,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -1.1, 
-                                              axis_pos_y = -1.1, 
+                                              axis_pos_x = -1.05, 
+                                              axis_pos_y = -1.05, 
                                               threshold = 0.1))
 
 pyrtri_proj3 <- plot_proj(
@@ -747,10 +747,10 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
[…]

### 2025-12-17 paper-cardinalR.Rmd
Updated Figure 5

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.9,
                                               axis_scaled = 1.2, 
@@
                                               axis_pos_y = -0.6, 
-                                              threshold = 0.022))
+                                              threshold = 0.09))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.57, 
-                                              axis_pos_y = -0.57, 
-                                              threshold = 0.01))
+                            axis_param = list(limits = 0.8,
+                                              axis_scaled = 1.8, 
+                                              axis_pos_x = -0.53, 
+                                              axis_pos_y = -0.53, 
+                                              threshold = 0.1))
 
@@
                                               axis_pos_y = -0.5, 
-                                              threshold = 0.016))
+                                              threshold = 0.07))
 
@@
                             axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
+                                              axis_scaled = 1.4, 
                                               axis_pos_x = -0.6, 
                                               axis_pos_y = -0.6, 
-                                              threshold = 0.022))
+                                              threshold = 0.08))
 
@@
                                               axis_pos_y = -0.57, 
-                                              threshold = 0.04))
+                                              threshold = 0.08))
 
@@
                             axis_param = list(limits = 0.6,
-                                              axis_scaled = 1.5, 
+                                              axis_scaled = 1.6, 
                                               axis_pos_x = -0.6, 
                                               axis_pos_y = -0.6, 
-                                              threshold = 0.016))
+                                              threshold = 0.08))
 
@@
                             axis_param = list(limits = 10,
-                                              axis_scaled = 1, 
+                                              axis_scaled = 1.5, 
                                               axis_pos_x = -7, 
                                               axis_pos_y = -7, 
-                                              threshold = 0.2))
+                                              threshold = 0.4))
 

### 2025-12-17 paper-cardinalR.Rmd
Updated Figure 6

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
-                                              axis_scaled = 1.2, 
+                            axis_param = list(limits = 0.8,
+                                              axis_scaled = 1.5, 
                                               axis_pos_x = -0.6, 
                                               axis_pos_y = -0.6, 
-                                              threshold = 0.022))
+                                              threshold = 0.1))
 
@@
                                               axis_pos_y = -0.55, 
-                                              threshold = 0.016))
+                                              threshold = 0.09))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1.2, 
+                            axis_param = list(limits = 1.2,
+                                              axis_scaled = 1.7, 
                                               axis_pos_x = -1, 
                                               axis_pos_y = -1, 
-                                              threshold = 0.04))
+                                              threshold = 0.1))
 
@@
                             centered_data = centered_data, 
+                            axis_param = list(limits = 1.2,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.8, 
+                                              axis_pos_y = -0.8, 
+                                              threshold = 0.1))
+
+trefoil3d_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-1, 1.2), 
+  title = "b2", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = FALSE)
+
+```
+
+```{r trefoil3d-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416),
+  c(0.50329,0.07199,-0.16064))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            centered_data = centered_data, 
                             axis_param = list(limits = 1,
-                                              axis_scaled = 1, 
+                                              axis_scaled = 1.5, 
                                               axis_pos_x = -0.8, 
@@
                                               threshold = 0.04))
-
-trefoil3d_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1, 1.2), 
-  title = "b2", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = FALSE)
-
-```
-
-```{r trefoil3d-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416),
-  c(0.50329,0.07199,-0.16064))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 0.8,
-                                              axis_scaled = 1.5, 
-                                              axis_pos_x = -0.8, 
-                                              axis_pos_y = -0.8, 
-                                              threshold = 0.016))
 

### 2025-12-17 paper-cardinalR.Rmd
Updated Figure 7

@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 1.4,
                                               axis_scaled = 1.2, 
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.1))
+
+curvycylinder_proj1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-1.3, 1.1), 
+  title = "a1", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = FALSE)
+
+```
+
+```{r curvycylinder-proj2}
+
+## Second projection
+projection <- cbind(
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            centered_data = centered_data, 
+                            axis_param = list(limits = 2.6,
+                                              axis_scaled = 1.2, 
+                                              axis_pos_x = -2.1, 
+                                              axis_pos_y = -2.1, 
+                                              threshold = 0.18))
+
+curvycylinder_proj2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
+  plot_limits = c(-2.6, 2.3), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = FALSE)
+
+```
+
+```{r curvycylinder-proj3}
+
+## Third projection
+projection <- cbind(
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            centered_data = centered_data, 
+                            axis_param = list(limits = 1.4,
+                                              axis_scaled = 1.4, 
                                               axis_pos_x = -1.1, 
@@
 
-curvycylinder_proj1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-1.3, 1.1), 
-  title = "a1", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj2}
-
-## Second projection
-projection <- cbind(
-  c(-0.06614,0.15134,-0.44278,0.27367),
-  c(-0.50493,0.14388,0.14615,0.03486))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 2,
-                                              axis_scaled = 1.2, 
-                                              axis_pos_x = -2.2, 
-                                              axis_pos_y = -2.2, 
-                                              threshold = 0.18))
-
-curvycylinder_proj2 <- plot_proj(
-  proj_obj = proj_obj2, 
-  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-2.6, 2.3), 
-  title = "a2", 
-  cex = 2, 
-  axis_text_size = 4,
-  is_color = FALSE)
-
-```
-
-```{r curvycylinder-proj3}
-
-## Third projection
-projection <- cbind(
-  c(0.10704,0.04822,-0.03416,0.53224),
-  c(0.50329,0.07199,-0.16064,-0.11806))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            centered_data = centered_data, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1.4, 
-                                              axis_pos_x = -1.2, 
-                                              axis_pos_y = -1.2, 
-                                              threshold = 0.1))
-
 curvycylinder_proj3 <- plot_proj(
@@
                                               axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                                              threshold = 0.09))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 1,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -0.65, 
-                                              axis_pos_y = -0.65, 
+                                              axis_pos_x = -0.63, 
+                                              axis_pos_y = -0.63, 
                                               threshold = 0.06))
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 2.5,
+                            axis_param = list(limits = 3.4,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -2, 
-                                              axis_pos_y = -2, 
+                                              axis_pos_x = -1.8, 
+                                              axis_pos_y = -1.8, 
                                               threshold = 0.2))
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 2,
+                            axis_param = list(limits = 3,
                                               axis_scaled = 1, 
-                                              axis_pos_x = -2.9, 
-                                              axis_pos_y = -2.9, 
-                                              threshold = 0.1))
+                                              axis_pos_x = -3.1, 
+                                              axis_pos_y = -3.1, 
+                                              threshold = 0.2))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-3.3, 2), 
+  plot_limits = c(-3.6, 2), 
   title = "c2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 2.5,
+                            axis_param = list(limits = 3.4,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -2.3, 
-                                              axis_pos_y = -2.3, 
+                                              axis_pos_x = -2.2, 
+                                              axis_pos_y = -2.2, 
                                               threshold = 0.2))
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 1.2,
                                               axis_scaled = 1.2, 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 1.8,
                                               axis_scaled = 1.3, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.1))
+                                              axis_pos_x = -0.6, 
+                                              axis_pos_y = -0.6, 
+                                              threshold = 0.18))
 
@@
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 2), 
+  plot_limits = c(-0.9, 2), 
   title = "d2", 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
+                            axis_param = list(limits = 0.9,
                                               axis_scaled = 1.5, 
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
+                                              axis_pos_x = -0.43, 
+                                              axis_pos_y = -0.43, 
                                               threshold = 0.07))

### 2025-12-17 paper-cardinalR.R

@@ -1462,10 +1462,10 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 1.4,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -1.1, 
-                                              axis_pos_y = -1.1, 
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
                                               threshold = 0.1))
 
 curvycylinder_proj1 <- plot_proj(
@@ -1488,10 +1488,10 @@ projection <- cbind(
 
 proj_obj2 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 2,
+                            axis_param = list(limits = 2.6,
                                               axis_scaled = 1.2, 
-                                              axis_pos_x = -2.2, 
-                                              axis_pos_y = -2.2, 
+                                              axis_pos_x = -2.1, 
+                                              axis_pos_y = -2.1, 
                                               threshold = 0.18))
 
 curvycylinder_proj2 <- plot_proj(
@@ -1514,10 +1514,10 @@ projection <- cbind(
 
 proj_obj3 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 1,
+                            axis_param = list(limits = 1.4,
                                               axis_scaled = 1.4, 
-                                              axis_pos_x = -1.2, 
-                                              axis_pos_y = -1.2, 
+                                              axis_pos_x = -1.1, 
+                                              axis_pos_y = -1.1, 
                                               threshold = 0.1))
 
 curvycylinder_proj3 <- plot_proj(
@@ -1549,11 +1549,11 @@ projection <- cbind(
 
 proj_obj1 <- get_projection(projection = projection, 
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.7,
+                            axis_param = list(limits = 0.9,
                                               axis_scaled = 1.2, 
                                               axis_pos_x = -0.5, 
                                               axis_pos_y = -0.5, 
-                                              threshold = 0.022))
+                                              threshold = 0.09))
 
 sphericalspiral_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
@@ -1575,10 +1575,10 @@ projection <- cbind(
 
[…]

### 2025-12-17 paper-cardinalR.Rmd
Updated wb.ratio

@@
 With the other three, lower is better.
-WB ratio (`wbratio`) and within-cluster sum of squares (`within.cluster.ss`) steadily decline with number of clusters, possibly elbowing around 5 clusters.
+WB ratio (`wb.ratio`) and within-cluster sum of squares (`within.cluster.ss`) steadily decline with number of clusters, possibly elbowing around 5 clusters.
 The S-index (`sindex`) is optimised at 4 clusters for $k$-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.

### 2025-12-17 paper-cardinalR.tex
Updated wb.ratio

@@
 With the other three, lower is better.
-WB ratio (\texttt{wbratio}) and within-cluster sum of squares (\texttt{within.cluster.ss}) steadily decline with number of clusters, possibly elbowing around 5 clusters.
+WB ratio (\texttt{wb.ratio}) and within-cluster sum of squares (\texttt{within.cluster.ss}) steadily decline with number of clusters, possibly elbowing around 5 clusters.
 The S-index (\texttt{sindex}) is optimised at 4 clusters for \(k\)-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.

### 2025-12-18 paper-cardinalR.Rmd
Updated Table 2

@@
                              "n",
-                             "n, p",
-                             "n, p",
-                             "n, p",
+                             "n, p, l_vec, rt",
+                             "n, p, h, l, rt",
+                             "n, p, h, rb",
                              "n, p",
@@
                               "Long linear structure.",
-                              "Möbius strip in $3\text{-}D$.",
-                              "Quadratic pattern in $2\text{-}D$.",
-                                "Cubic pattern in $2\text{-}D$.",
+                              "Möbius strip in $3\\text{-}D$.",
+                              "Quadratic pattern in $2\\text{-}D$.",
+                                "Cubic pattern in $2\\text{-}D$.",
                               "Rectangular-base, with a sharp or blunted apex.",
@@
 ```{r}
-arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "n_vec", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac"), 
+arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "n_vec", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac", "l", "l_vec", "rt", "rb"), 
                           type = c("integer", "integer", "integer", "real value",
-                                   "real value", "real value", "real value", "integers", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value"),
+                                   "real value", "real value", "real value", "integers", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value", "real value", "real values", "real value", "real value"),
                         exp = c("Number of points.",
@@
                                 "Steepness and vertical scaling of the hyperbola.",
-                                "Strength of this sinusoidal effect.")) |> 
+                                "Strength of this sinusoidal effect.",
+                                "Base length of the pyramid.",
+                                "Base lengths along the and y of the pyramid.",
+                                "Tip radius of the pyramid.",
+                                "Base radius of the pyramid.")) |> 
    mutate(arg = gsub("_", "\\\\_", arg))
@@
 For the rectangular based pyramid, `gen_pyrrect(n, p, h, l_vec, rt)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:pyr) a"} else { "Figure \\@ref(fig:pyr-proj) a"})`) the base shape is a rectangle whose size shrinks linearly with height.
-Let $l_x$ and $l_y$ denote the half-widths of the rectangular base in the $X_1$ and $X_2$ directions, specified via $l=(l_x,l_y)$, and let $r_t$ denote the half-width at the pyramid tip.
+Let $l_x$ and $l_y$ denote the half-widths of the rectangular base in the $X_1$ and $X_2$ directions, specified via $l_{vec}=(l_x,l_y)$, and let $r_t$ denote the half-width at the pyramid tip.
 At height $z\in [0,h]$, the half-widths of the rectangular cross-section are $r_x(z) = r_t + (l_x - r_t)z/h$, $r_y(z) = r_t + (l_y - r_t)z/h$.

### 2025-12-18 paper-cardinalR.tex
Updated Table 2

@@
 \texttt{gen\_longlinear} & \texttt{n, p} & Long linear structure.\\
-\texttt{gen\_mobius} & \texttt{n} & Möbius strip in $3  ext{-}D$.\\
-\texttt{gen\_quadratic} & \texttt{n} & Quadratic pattern in $2  ext{-}D$.\\
-\texttt{gen\_cubic} & \texttt{n} & Cubic pattern in $2  ext{-}D$.\\
-\texttt{gen\_pyrrect} & \texttt{n, p} & Rectangular-base, with a sharp or blunted apex.\\
-\texttt{gen\_pyrtri} & \texttt{n, p} & Triangular-base, with a sharp or blunted apex.\\
-\texttt{gen\_pyrstar} & \texttt{n, p} & Star-shaped base, with a sharp or blunted apex.\\
+\texttt{gen\_mobius} & \texttt{n} & Möbius strip in $3\text{-}D$.\\
+\texttt{gen\_quadratic} & \texttt{n} & Quadratic pattern in $2\text{-}D$.\\
+\texttt{gen\_cubic} & \texttt{n} & Cubic pattern in $2\text{-}D$.\\
+\texttt{gen\_pyrrect} & \texttt{n, p, l\_vec, rt} & Rectangular-base, with a sharp or blunted apex.\\
+\texttt{gen\_pyrtri} & \texttt{n, p, h, l, rt} & Triangular-base, with a sharp or blunted apex.\\
+\texttt{gen\_pyrstar} & \texttt{n, p, h, rb} & Star-shaped base, with a sharp or blunted apex.\\
 \texttt{gen\_pyrfrac} & \texttt{n, p} & Pyramid with triangular pyramid-shaped holes.\\
@@
 \texttt{non\_fac} & real value & Strength of this sinusoidal effect.\\
+\texttt{l} & real value & Base length of the pyramid.\\
+\texttt{l\_vec} & real values & Base lengths along the and y of the pyramid.\\
+\texttt{rt} & real value & Tip radius of the pyramid.\\
+\texttt{rb} & real value & Base radius of the pyramid.\\
 \bottomrule
@@
 For the rectangular based pyramid, \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} (Figure \ref{fig:pyr-proj} a) the base shape is a rectangle whose size shrinks linearly with height.
-Let \(l_x\) and \(l_y\) denote the half-widths of the rectangular base in the \(X_1\) and \(X_2\) directions, specified via \(l=(l_x,l_y)\), and let \(r_t\) denote the half-width at the pyramid tip.
+Let \(l_x\) and \(l_y\) denote the half-widths of the rectangular base in the \(X_1\) and \(X_2\) directions, specified via \(l_{vec}=(l_x,l_y)\), and let \(r_t\) denote the half-width at the pyramid tip.
 At height \(z\in [0,h]\), the half-widths of the rectangular cross-section are \(r_x(z) = r_t + (l_x - r_t)z/h\), \(r_y(z) = r_t + (l_y - r_t)z/h\).

### 2025-12-18 paper-cardinalR.R

@@ -171,9 +171,9 @@ shape_tb <- tibble(fun = c("gen_expbranches",
                              "n",
                              "n",
                              "n",
-                             "n, p",
-                             "n, p",
-                             "n, p",
+                             "n, p, l_vec, rt",
+                             "n, p, h, l, rt",
+                             "n, p, h, rb",
                              "n, p",
                               "n",
                              "n, p",
@@ -202,9 +202,9 @@ shape_tb <- tibble(fun = c("gen_expbranches",
                               "Cube with uniform points.",
                               "Multivariate Gaussian cloud.",
                               "Long linear structure.",
-                              "Möbius strip in $3\text{-}D$.",
-                              "Quadratic pattern in $2\text{-}D$.",
-                                "Cubic pattern in $2\text{-}D$.",
+                              "Möbius strip in $3\\text{-}D$.",
+                              "Quadratic pattern in $2\\text{-}D$.",
+                                "Cubic pattern in $2\\text{-}D$.",
                               "Rectangular-base, with a sharp or blunted apex.",
           "Triangular-base, with a sharp or blunted apex.",
           "Star-shaped base, with a sharp or blunted apex.",
@@ -248,9 +248,9 @@ shape_tb |>
 
 
 ## -----------------------------------------------------------------------------
-arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "n_vec", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac"), 
+arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "n_vec", "k_small", "r_vec", "spe", "w", "steps", "spins", "hc", "non_fac", "l", "l_vec", "rt", "rb"), 
                           type = c("integer", "integer", "integer", "real value",
-                                   "real value", "real value", "real value", "integers", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value"),
+                                   "real value", "real value", "real value", "integers", "integer", "real values", "real value", "real value", "integer", "integer", "real value", "real value", "real value", "real values", "real value", "real value"),
                         exp = c("Number of points.",
                                 "Number of clusters.",
                                 "Number of dimensions.",
@@ -266,7 +266,11 @@ arg_branching_tb <- tibble(arg = c("n", "k", "p", "h", "ratio", "s", "r", "n_vec
                                 "Number of steps for the theta parameter.",
                                 "Number of loops of the spiral.",
[…]

### 2025-12-18 paper-cardinalR.Rmd
Updated the table

@@
                                 "Cubic pattern in $2\\text{-}D$.",
-                              "Rectangular-base, with a sharp or blunted apex.",
-          "Triangular-base, with a sharp or blunted apex.",
-          "Star-shaped base, with a sharp or blunted apex.",
+                              "Rectangular-base Pyramid, with a sharp or blunted apex.",
+          "Triangular-base Pyramid, with a sharp or blunted apex.",
+          "Star-shaped base Pyramid, with a sharp or blunted apex.",
           "Pyramid with triangular pyramid-shaped holes.",
-          "S-curve in $3\text{-}D$.",
+          "S-curve in $3\\text{-}D$.",
           "Circle.",

### 2025-12-18 paper-cardinalR.tex
Updated the table

@@
 \texttt{gen\_cubic} & \texttt{n} & Cubic pattern in $2\text{-}D$.\\
-\texttt{gen\_pyrrect} & \texttt{n, p, l\_vec, rt} & Rectangular-base, with a sharp or blunted apex.\\
-\texttt{gen\_pyrtri} & \texttt{n, p, h, l, rt} & Triangular-base, with a sharp or blunted apex.\\
-\texttt{gen\_pyrstar} & \texttt{n, p, h, rb} & Star-shaped base, with a sharp or blunted apex.\\
+\texttt{gen\_pyrrect} & \texttt{n, p, l\_vec, rt} & Rectangular-base Pyramid, with a sharp or blunted apex.\\
+\texttt{gen\_pyrtri} & \texttt{n, p, h, l, rt} & Triangular-base Pyramid, with a sharp or blunted apex.\\
+\texttt{gen\_pyrstar} & \texttt{n, p, h, rb} & Star-shaped base Pyramid, with a sharp or blunted apex.\\
 \texttt{gen\_pyrfrac} & \texttt{n, p} & Pyramid with triangular pyramid-shaped holes.\\
-\texttt{gen\_scurve} & \texttt{n} & S-curve in $3   ext{-}D$.\\
+\texttt{gen\_scurve} & \texttt{n} & S-curve in $3\text{-}D$.\\
 \texttt{gen\_circle} & \texttt{n, p} & Circle.\\

### 2025-12-18 paper-cardinalR.R

@@ -205,11 +205,11 @@ shape_tb <- tibble(fun = c("gen_expbranches",
                               "Möbius strip in $3\\text{-}D$.",
                               "Quadratic pattern in $2\\text{-}D$.",
                                 "Cubic pattern in $2\\text{-}D$.",
-                              "Rectangular-base, with a sharp or blunted apex.",
-          "Triangular-base, with a sharp or blunted apex.",
-          "Star-shaped base, with a sharp or blunted apex.",
+                              "Rectangular-base Pyramid, with a sharp or blunted apex.",
+          "Triangular-base Pyramid, with a sharp or blunted apex.",
+          "Star-shaped base Pyramid, with a sharp or blunted apex.",
           "Pyramid with triangular pyramid-shaped holes.",
-          "S-curve in $3\text{-}D$.",
+          "S-curve in $3\\text{-}D$.",
           "Circle.",
                                 "Curvy cell cycle.",
                                 "Uniform ball.",

### 2025-12-18 paper-cardinalR.Rmd
Updated scale parameter notation

@@
 Coordinates in the first two dimensions represent a perfect circle on the plane: $$X_1 = \cos(\theta), \quad X_2 = \sin(\theta).$$ For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced.
-The first component is a scaling factor that decreases with the dimension index, defined as $\text{scale}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$.
+The first component is a scaling factor that decreases with the dimension index, defined as $\text{s}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$.
 The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2)\pi/2p$.
-Each additional dimension is computed as: $X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p$.
+Each additional dimension is computed as: $X_j = \text{s}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p$.
 
@@
 For dimensions $X_4$ through $X_p$, additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
-The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from $4$ to $p$, which means that the amplitude decreases as the dimension increases.
-Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j\sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2)\pi/(2p)$.
+The scaling factor is defined as $\text{s}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from $4$ to $p$, which means that the amplitude decreases as the dimension increases.
+Each dimension $X_j$ is then calculated using the formula $X_j = \text{s}_j\sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2)\pi/(2p)$.
 
@@
 
-In addition, the `gen_gridedsphere(n)` function constructs a $p$-dimensional dataset consisting of approximately $n$ points arranged on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
+In addition, the `gen_gridedsphere(n)` function constructs a $p$-dimensional dataset consisting of approximately $n$ points arranged on the surface of the unit $(p-1)\text{-}D$ sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
 Rather than sampling points uniformly, this function creates a deterministic grid in spherical coordinates, using $(p-1)$ angular variables: the first $(p-2)$ angles are taken from $[0, \pi]$, and the final angle from $[0, 2\pi]$.
@@
 In generation of $k$ smaller Uniform balls, each sphere contains $n_2$ points that are sampled uniformly on a sphere with a radius of $r_2$.
-These spheres are positioned at distinct random locations in $p$-space, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
+These spheres are positioned at distinct random locations in $p\text{-}D$, with the center of each sphere being drawn from a normal distribution $N(0, \texttt{loc}^2 I_p)$.
 Points on spheres are generated using the standard hyperspherical method, which involves sampling $u \sim U(-1, 1)$ to determine the cosine of the polar angle, and sampling $\theta \sim U(0, 2\pi)$ to determine the azimuthal angle (for $3\text{-}D$).

### 2025-12-18 paper-cardinalR.tex
Updated scale parameter notation

@@
 Coordinates in the first two dimensions represent a perfect circle on the plane: \[X_1 = \cos(\theta), \quad X_2 = \sin(\theta).\] For dimensions \(X_3\) through \(X_p\), sinusoidal transformations of the angle \(\theta\) are introduced.
-The first component is a scaling factor that decreases with the dimension index, defined as \(\text{scale}_j = \sqrt{(0.5)^{j-2}}\) for \(j = 3, \dots, p\).
+The first component is a scaling factor that decreases with the dimension index, defined as \(\text{s}_j = \sqrt{(0.5)^{j-2}}\) for \(j = 3, \dots, p\).
 The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula \(\phi_j = (j - 2)\pi/2p\).
-Each additional dimension is computed as: \(X_j = \text{scale}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p\).
+Each additional dimension is computed as: \(X_j = \text{s}_{j}\sin(\theta + \phi_j), \quad j = 3, \dots, p\).
 
@@
 For dimensions \(X_4\) through \(X_p\), additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves.
-The scaling factor is defined as \(\text{scale}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from \(4\) to \(p\), which means that the amplitude decreases as the dimension increases.
-Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{scale}_j\sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2)\pi/(2p)\).
+The scaling factor is defined as \(\text{s}_j = \sqrt{(0.5)^{j-3}}\) for \(j\) ranging from \(4\) to \(p\), which means that the amplitude decreases as the dimension increases.
+Each dimension \(X_j\) is then calculated using the formula \(X_j = \text{s}_j\sin(\theta + \phi_j)\), where the phase shift \(\phi_j\) is given by \(\phi_j = (j - 2)\pi/(2p)\).
 
@@
 
-In addition, the \texttt{gen\_gridedsphere(n)} function constructs a \(p\)-dimensional dataset consisting of approximately \(n\) points arranged on the surface of the unit \((p-1)\)-sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
+In addition, the \texttt{gen\_gridedsphere(n)} function constructs a \(p\)-dimensional dataset consisting of approximately \(n\) points arranged on the surface of the unit \((p-1)\text{-}D\) sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
 Rather than sampling points uniformly, this function creates a deterministic grid in spherical coordinates, using \((p-1)\) angular variables: the first \((p-2)\) angles are taken from \([0, \pi]\), and the final angle from \([0, 2\pi]\).
@@
 In generation of \(k\) smaller Uniform balls, each sphere contains \(n_2\) points that are sampled uniformly on a sphere with a radius of \(r_2\).
-These spheres are positioned at distinct random locations in \(p\)-space, with the center of each sphere being drawn from a normal distribution \(N(0, \texttt{loc}^2 I_p)\).
+These spheres are positioned at distinct random locations in \(p\text{-}D\), with the center of each sphere being drawn from a normal distribution \(N(0, \texttt{loc}^2 I_p)\).
 Points on spheres are generated using the standard hyperspherical method, which involves sampling \(u \sim U(-1, 1)\) to determine the cosine of the polar angle, and sampling \(\theta \sim U(0, 2\pi)\) to determine the azimuthal angle (for \(3\text{-}D\)).

### 2025-12-19 paper-cardinalR.Rmd
Added alt-text for application section

@@
 The \\texttt{nonlinear} dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
-Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="H", fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped $2\\text{-}D$ surfaces, illustrating curvature, density, and continuity.", fig.pos="!ht"}
+Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="!ht", fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped $2\\text{-}D$ surfaces, illustrating curvature, density, and continuity."}
 
@@
 
-```{r highd-data, eval=knitr::is_html_output(), fig.cap="Viewing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="Interactive langevitour of five synthetic clusters in $4\\text{-}D$: helical spiral, hemisphere, uniform cube, cone, and Gaussian cluster, each colored distinctly to highlight geometric structure.", fig.pos="!ht"}
+```{r highd-data, eval=knitr::is_html_output(), fig.cap="Viewing five synthetic clusters with distinct geometric structures: a helical spiral, a hemisphere, a uniform cube, a cone, and a Gaussian cluster.", fig.alt="The figure shows a dynamic tour visualization of synthetic four-dimensional data displayed as a sequence of two-dimensional scatter plots.
+Each frame presents a different linear projection of the data, with the horizontal and vertical axes representing changing combinations of the original variables.
+Points represent individual observations and are coloured by cluster membership, with five distinct colours corresponding to the helical spiral, hemisphere, uniform cube, cone, and Gaussian clusters.
+As the tour progresses, the relative positions, shapes, and overlaps of the coloured point clouds change continuously, revealing elongated bands, curved arcs, compact clusters, and more diffuse regions depending on the projection.
+Clusters frequently move closer together or overlap in certain views, while becoming more separated in others, illustrating how apparent structure varies across different two-dimensional projections of the same high-dimensional data.", fig.pos="!ht"}
 
@@
 ```{r highd-proj, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data.
-The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="Three $2\\text{-}D$ projections of five $4\\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.", fig.width=12, fig.height=4, fig.pos="!ht"}
+The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.", fig.alt="The figure shows three 2-D projection plots of synthetic 4-D data containing five clusters with distinct geometric structures.
+The horizontal and vertical axes represent two projected dimensions, and no explicit units are shown.
+Points are coloured by cluster, with distinct colours representing the helical spiral, hemisphere, uniform cube, cone, and Gaussian clusters.
+The helical spiral cluster appears as an elongated, twisted band; the hemisphere cluster forms a curved, arc-like concentration; the uniform cube cluster fills a roughly rectangular region with fairly even density; the cone cluster exhibits a density gradient, narrowing at one end and spreading out at the other; and the Gaussian cluster forms a compact, approximately circular group near the centre of the plot.
+The clusters appear very close to one another in these projections, illustrating the difficulty of distinguishing cluster structure when projections bring clusters into close proximity.", fig.width=12, fig.height=4}
 
@@
 
-```{r, label = "fig-nldr-layouts", fig.pos="H", fig.cap="Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
+```{r, label = "fig-nldr-layouts", fig.pos="!ht", fig.cap="Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
 Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
@@
 Layouts a and b are universally optimal.
-Layout b shows six well-separated clusters and layout a shows close clusters, thus layout a is the best choice.", fig.alt="Composite plot assessing six NLDR layouts of five clusters data using HBE over varying binwidths, showing tSNE, UMAP, PAHTE, TriMAP, PaCMAP, and PCA.
-Layouts a and b are optimal, f is poor, with color-coded lines matching the scatterplots of each layout.", fig.pos="!ht"}
+Layout b shows six well-separated clusters and layout a shows close clusters, thus layout a is the best choice.", fig.alt="A multi-panel figure compares 2-D scatterplots from six nonlinear dimensionality reduction methods applied to the same dataset with five true clusters.
+Each panel plots the first embedding dimension on the horizontal axis and the second embedding dimension on the vertical axis.
+Points represent individual observations and are coloured by cluster membership, with six distinct cluster colours reused consistently across panels.
+In the tSNE panel, clusters form compact, clearly separated groups with small gaps between clusters, indicating strong preservation of both local neighbourhoods and the global cluster layout.
+UMAP and PaCMAP also show six visibly distinct clusters that are moderately well separated but with slightly more overlap and distortion than tSNE.
+PHATE produces curved, nonlinear cluster shapes where clusters are stretched or intertwined, obscuring the original simple cluster geometry.
+TriMAP collapses the data into three main visible groups instead of six, with only small distances between these groups, suggesting loss of finer cluster structure.
+PCA displays the weakest structure: clusters overlap substantially and align along a roughly linear or planar trend, failing to reflect the underlying non-linear separation among the six groups."}
 
@@
 ```{r fig-cluster-stats, fig.width=8, fig.height=6, out.width="100%", layout="l-body", fig.cap="Cluster validity metrics for solutions with $2-10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
-Several indices consistently suggest that $4-5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics."}
+Several indices consistently suggest that $4-5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics.", fig.alt="Multi-panel line chart showing six cluster quality metrics across numbers of clusters (x-axis: 2 to 10 clusters) for three clustering methods (k-means, hierarchical, model-based).
+Each panel displays one metric with the y-axis scaled to that metric’s values (not explicitly labeled in the text, but each metric is treated as higher-is-better or lower-is-better as described).
+Within each panel, three lines (one per method) trace the metric value as the number of clusters increases.
+For Pearson gamma, values for all three methods rise steeply up to about 5 clusters and then level off.
+For the Calinski–Harabasz index, values increase sharply from 4 to 5 clusters.
+For Dunn, the k-means line peaks around 4 clusters, while the hierarchical and model-based lines peak around 5.
+For WB ratio and within-cluster sum of squares, all three methods show a generally monotonic decline as the number of clusters increases, with a visible bend or elbow around 5 clusters.
+For the S-index, the k-means line reaches its best value around 4 clusters, the hierarchical line around 3, 6, or 8 clusters, and the model-based line around 4 or 8.
+Across panels, the k-means line is typically at or near the most favorable values for each metric, and several metrics simultaneously favor solutions with about 4–5 clusters."}
 
@@
 Although the metrics differ in their preferences, several show consistent support for a $4-5$ cluster solution.
-Pearson gamma (`pearsongamma`) increases sharply up to five clusters before levelling off, Calinski–Harabasz index (`ch`) increases sharply from 4 to 5 clusters and Dunn (`dunn2`) has a maximum at 5 for two methods and at 4 for $k$-means.
+Pearson gamma (`pearsongamma`) increases sharply up to five clusters before leveling off, Calinski–Harabasz index (`ch`) increases sharply from 4 to 5 clusters and Dunn (`dunn2`) has a maximum at 5 for two methods and at 4 for $k$-means.
 All of these are interpreted as higher is better.
@@
 WB ratio (`wb.ratio`) and within-cluster sum of squares (`within.cluster.ss`) steadily decline with number of clusters, possibly elbowing around 5 clusters.
-The S-index (`sindex`) is optimised at 4 clusters for $k$-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.
+The S-index (`sindex`) is optimized at 4 clusters for $k$-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.
 Overall, $k$-means performs slightly better than the hierarchical and model-based clustering across most metrics and number of clusters.
@@
 These show how the five true structures: helical spiral, hemisphere, uniform cube, blunted cone, and Gaussian-shaped cluster appear under $k$-means partitions.
-Both the $4$- and $5$-cluster solutions mix colours, revealing substantial misclassification and limited ability of $k$-means to recover the underlying geometric structure.", fig.pos="!ht", layout = "l-body"}
+Both the $4$- and $5$-cluster solutions mix colours, revealing substantial misclassification and limited ability of $k$-means to recover the underlying geometric structure.", fig.alt="The figure contains two side-by-side tour views of the same synthetic high-dimensional dataset, each shown as a sequence of continuously changing 2-D projections.
+The left tour view is coloured by a four-cluster k-means solution, and the right tour view is coloured by a five-cluster k-means solution.
+In both views, each point represents a data observation, and colours indicate cluster membership.
+The horizontal and vertical axes correspond to linear projections of the original variables.
+Across the tours, points form rotating and deforming point clouds that intermittently reveal curved, elongated, compact, and diffuse structures.
+Colours are intermingled throughout most projections in both panels, with clusters frequently overlapping and splitting as the view changes.
+The two tours differ only in the number of colours used, with four colours on the left and five on the right, while the underlying point positions and projection motion remain the same.", fig.pos="!ht", layout = "l-body"}
 
@@
 ```{r highd-proj-clust-algo-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data coloured by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
-The intermixing of colours within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.", fig.width=12, fig.height=8, fig.pos="!ht"}
+The intermixing of colours within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.", fig.alt="The figure shows three row-wise 2-D projection plots of synthetic 4-D data for two different k-means clustering solutions: four clusters in one panel and five clusters in the other.
+In each panel, the horizontal and vertical axes represent linear projections of the original high-dimensional variables, and each point corresponds to a data observation.
+Points are coloured by assigned cluster ID (four colours in the top row and five in the bottom row).
+In both solutions, clusters appear as roughly compact groups, but none aligns with the underlying geometric shapes in the data: each of the five true shapes is split across multiple clusters, and no cluster cleanly isolates a single shape.
+The four-cluster solution merges some shapes more strongly, while the five-cluster solution yields a finer partition that still fragments each shape; additional clusters beyond five would be needed to better match the original geometric structure.", fig.width=12, fig.height=8, fig.pos="!ht"}
 

### 2025-12-19 paper-cardinalR.Rmd
Added other alt-text as well

@@
 Each cluster shape is generated by a shape generator, optionally rotated or scaled, and combined into a single dataset.
-Additional background noise can be added, and each observation is labeled by shape.", fig.alt="Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.", fig.pos="!ht"}
+Additional background noise can be added, and each observation is labeled by shape.", fig.alt="The figure is a workflow diagram showing the process for generating high-dimensional clustered data.
+Boxes or nodes represent steps such as specifying input parameters (number of points, number of clusters, cluster locations, shapes, scaling, rotations, and optional background noise), generating each cluster with a shape generator, optionally rotating or scaling clusters, combining them into a single dataset, adding background noise if desired, and labeling each observation by cluster shape.
+Arrows indicate the flow of the process from parameter specification to the final labeled dataset.
+The diagram uses a clear layout with distinct steps, showing how data are constructed in sequence.", fig.pos="!ht"}
 
@@
 ```{r branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `orgcurvybranches` data.
-The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body"}
+The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body", fig.alt="The figure shows a tour of the 4‑D orgcurvybranches dataset.
+Points are displayed in a continuously changing 2‑D projection that rotates through the four dimensions.
+Each axis of the 2‑D view represents a linear combination of the original 4‑D coordinates.
+Points appear grouped into multiple elongated clusters or branches that move, rotate, and sometimes overlap as the projection changes.
+Branches are distinguished by colour, with individual points visible throughout.
+The display maintains a roughly square aspect ratio as the view transitions through multiple angles, illustrating the dataset’s branching structure from different perspectives."}
 
@@
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
-These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.alt="The figure shows a 2-D scatter plot of projected branch structures in a high-dimensional space.
-The horizontal axis represents one coordinate (first index of the active pair) and the vertical axis represents the other coordinate (second index of the active pair), each spanning a moderate, symmetric range around zero (for example, roughly −2 to 2).
-Points are grouped visually into several distinct linear or gently curving branches, each corresponding to a different active 2-D subspace; these branches appear as separate streaks or arcs radiating through the plane.
-Colour (or another visual attribute such as line/point grouping) is used to distinguish branches that occupy different subspaces, and some branches may overlap or run close together where subspaces are shared.
-The main pattern is a set of well-separated, structured branches with little noise, illustrating how high-dimensional linear or curvy branches project into 2-D coordinate pairs.", fig.pos="!ht"}
+These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.alt="The figure consists of three 2‑D scatter plots, each showing a different 2‑D projection from the 4‑D orgcurvybranches dataset.
+In each plot, the horizontal axis represents one coordinate of the selected 2‑D projection, and the vertical axis represents the other coordinate.
+Points are plotted as individual markers and appear grouped into multiple elongated clusters or branches, which curve or radiate through the plot.
+Each branch is distinguished by colour, with overlapping branches occurring when subspaces are shared.
+The plots have roughly square aspect ratios, and together they show how the same high‑dimensional branching structure appears differently from three viewing angles.", fig.pos="!ht"}
 
@@
 
-```{r cone, eval=knitr::is_html_output(), fig.cap="Viewing the $3\\text{-}D$ output  of `cone` data.
-Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="A 3D scatterplot shows a cone-shaped cloud of points, with two horizontal axes representing the first two coordinates (x and y) and the vertical axis representing height (z).
-The horizontal axes span roughly symmetric ranges centered around zero, while height increases from a low base up to a maximum near the tip of the cone.
-At lower heights the points spread widely in the horizontal plane, and as height increases the radial spread decreases smoothly, forming a narrowing cone.
-Points are evenly distributed around the cone’s central axis, reflecting that directions in the first two dimensions are sampled uniformly on a circle (a 2D hypersphere), and their distance from the center is scaled by a height-dependent radius.
-The resulting pattern is a dense, symmetric conical taper with no distinct outliers or clusters beyond the cone surface.", fig.pos="!ht"}
+```{r cone, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ output  of `cone` data.
+Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.", fig.alt="The figure shows a tour of the 4‑D cone dataset.
+Points are displayed in a continuously changing 2‑D projection, where each axis represents a linear combination of the original four dimensions.
+Points are densely packed near one end, forming a narrow tip, and gradually spread out along the other direction, creating a wider base.
+All points share the same colour and marker.
+As the projection rotates, the cone-like structure shifts in orientation, maintaining roughly square aspect ratios throughout the animation and showing the distribution of points from multiple viewing angles.", fig.pos="!ht"}
 
@@
 
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data.
+```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ \\texttt{cone} data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
-These projections show how the conical geometry is preserved.", fig.pos="H", fig.alt=" A 2D scatterplot shows the projection of a high-dimensional cone-shaped point cloud.
-The horizontal and vertical axes are unlabeled numeric coordinates, each spanning a roughly symmetric range around zero (for example, from about -1 to 1).
-Hundreds of points form a dense cluster near one end of the plot, corresponding to the cone’s apex, with density gradually decreasing as points extend outward in a flared, cone-like or triangular region.
-There are no colours, shapes, or facets used; all points share the same appearance, and the key visible pattern is the concentrated tip transitioning smoothly into a wider spread of points, illustrating a cone structure in projection.", fig.pos="!ht", fig.width=12, fig.height=4}
+These projections show how the conical geometry is preserved.", fig.alt="The figure consists of three 2‑D scatter plots, each showing a different 2‑D projection of the 4‑D cone dataset.
+In each plot, the horizontal and vertical axes represent two continuous numeric dimensions.
+Points are plotted individually and are concentrated near one end, forming a narrow tip, and gradually spread out along the other direction, creating a wider base.
+The points trace an elongated, triangular or cone-like shape in each projection.
+All points use the same colour and marker, with no additional visual encodings, and the plots have roughly square aspect ratios.
+Together, the three projections illustrate the distribution of points along the conical structure from different viewing angles.", fig.pos="!ht", fig.width=12, fig.height=4}
 
@@
 The `pyrrrect` structure forms a dense rectangular base tapering to a narrow tip, while `pytri` shows a more triangular spread with sharper edges.
-`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.", fig.alt="Interactive langevitour of four $4\\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing dense bases, triangular spreads, branching points, and hollow regions.", fig.pos="!ht"}
+`Pyrstar` extends into multiple pointed branches radiating from a common core, and `pyrholes` reveals hollow or open regions within an otherwise compact shape.", fig.alt="The figure shows an interactive tour of the 4‑D pyrrrect, pyrtri, pyrstar, and pyrholes datasets.
+Points are displayed in continuously changing 2‑D projections, with each axis representing a linear combination of the four original dimensions.
+Points are plotted individually as small, uniformly coloured markers with no additional visual encodings.
+In pyrrrect, points form a dense rectangular base tapering to a narrow tip; in pyrtri, points form a triangular distribution with sharper edges; in pyrstar, points extend into multiple pointed branches radiating from a central region; and in pyrholes, points cluster in a compact shape but leave hollow or open regions within it.
+The projections rotate smoothly, showing the 4‑D structures from multiple viewing angles and maintaining roughly square aspect ratios throughout.", fig.pos="!ht"}
 
@@
 \\texttt{pyrstar} extends into multiple pointed branches radiating from a common core, and \\texttt{pyrholes} reveals hollow or open regions within an otherwise compact shape.
-These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.", fig.pos="!ht"}
+These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, out.width="80%", fig.align='center', fig.alt="The figure shows three 2‑D scatter plots for each of four 4‑D datasets: pyrrrect, pyrtri, pyrstar, and pyrholes.
+In each plot, the horizontal and vertical axes represent continuous numeric coordinates of the selected 2‑D projection, and points are plotted individually as small, uniformly coloured markers with no additional visual encodings.
+In pyrrrect, points form a dense rectangular base tapering to a narrow tip; in pyrtri, points form a triangular distribution with sharper edges; in pyrstar, points extend into multiple pointed branches radiating from a central region; and in pyrholes, points cluster in a compact shape but leave hollow or open regions within it.
+The plots maintain roughly square aspect ratios and illustrate how these different pyramid-like structures appear when projected from four to two dimensions.", fig.pos="!ht"}
 
@@
 The \\texttt{clusteredspheres} dataset displays multiple compact spherical groups that are clearly separated in higher dimensions but overlap slightly in some $2\\text{-}D$ projections, highlighting how projection can distort spatial relationships.
-These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H", fig.alt="Three $2\\text{-}D$ projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.", fig.pos="!ht"}
+These projections show how simple cyclic, wavy curvilinear, and clustered structures appear in $2\\text{-}D$, emphasizing the effects of projection on density, continuity, and separation", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.alt="A 3D scatterplot shows a projection of points sampled from a 4‑dimensional hemisphere onto three coordinate axes.
+Each point represents a location on the surface of a 4‑D unit sphere restricted to one hemisphere.
+The horizontal (x) and vertical (y) axes, along with depth (z), span roughly from –1 to 1, reflecting the Cartesian coordinates of the projected points.
+The dots are uniformly scattered over a curved, bowl-like region rather than filling a full sphere, with no strong clusters or gaps, illustrating that the sampling covers only one side of the sphere.
+If colour is used, it likely encodes an additional coordinate or density but does not introduce distinct categorical groups.", fig.pos="!ht"}
 
@@
 The `trefoil4d` structure represents a higher-dimensional extension of the classic trefoil knot, revealing complex twisting and looping patterns that remain continuous across projections.
-In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.", fig.alt="Interactive langevitour of two trefoil knot datasets: trefoil4d in $4\\text{-}D$ showing complex twisting and looping patterns, and trefoil3d in 3-D showing a simpler compact knot-like form.", fig.pos="!ht"}
+In contrast, the `trefoil3d` dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.", fig.alt="The figure shows two side-by-side tour views comparing trefoil-shaped data in different dimensions.
+The left panel displays a tour of the 4-D trefoil4d dataset, while the right panel shows a tour of the 3-D trefoil3d dataset.
+In the trefoil4d tour, the projected points form looping, knot-like curves that repeatedly change shape as the projection rotates, with apparent crossings, separations, and overlaps appearing and disappearing over time.
+The structure remains continuous but spreads and twists more across projections, occupying a larger area of the plotting region.
+In the trefoil3d tour, the projected points also form a closed, looped curve, but the shape appears more compact and stable across frames, with fewer dramatic changes in overlap or separation as the projection changes.", fig.pos="!ht"}
 
@@
 In contrast, the \\texttt{trefoil3d} dataset maintains a simpler, more compact knot-like form, showing how dimensional extension adds curvature and separation in the embedded space.
-These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, fig.pos="H", out.width="80%", fig.align='center', fig.alt="Three $2\\text{-}D$ projections from each of two trefoil knot datasets, showing complex $4\\text{-}D$ trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.", fig.pos="!ht"}
+These projections illustrate a range of looping structures in high-dimensions.", fig.width=12, fig.height=8, out.width="80%", fig.align='center', fig.alt="The figure shows multiple scatterplots displaying two-dimensional projections of trefoil-shaped data derived from four- and three-dimensional spaces.
+In panels a1–a3, the trefoil4d data are shown as three different 2-D projections, where the horizontal and vertical axes represent linear combinations of the four original variables.
+The projected points form smooth, continuous looping curves that intersect and overlap in places, creating knot-like shapes whose apparent crossings and separations vary across projections.
+In panels b1–b3, the trefoil3d data are shown in three 2-D projections.
+The points trace a compact, closed loop with a characteristic trefoil-like knot structure, appearing smoother and less spread out than in the four-dimensional case.
+The figure highlights how the same underlying knot structure appears differently under changes in projection and dimensionality.", fig.pos="!ht"}
 
@@
 The `conicspiral` spreads radially while ascending, forming a conical helix with twisting variations in a non-radial dimension.
-The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.", fig.alt="Interactive langevitour of six datasets in $4\\text{-}D$: crescent, curvycylinder, sphericalspiral, helicalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral paths, conical helices, and warped $2\\text{-}D$ surfaces with smooth and sharp variations.", fig.pos="!ht"}
+The `nonlinear` dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.", fig.alt="", fig.pos="!ht"}
 
@@
 The \\texttt{nonlinear} dataset exhibits a warped $2\\text{-}D$ surface with sharp oscillations and smooth waves, reflecting complex nonlinear interactions.
-Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="!ht", fig.alt="Three $2\\text{-}D$ projections from each of four $4\\text{-}D$ datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped $2\\text{-}D$ surfaces, illustrating curvature, density, and continuity."}
+Each shows variations in curvature, density, and continuity.", fig.width=12, fig.height=16, out.width="70%", fig.align='center', fig.pos="!ht", fig.alt="The figure consists of four rows of panels, each displaying three separate 2-D projection scatterplots from a four-dimensional synthetic dataset.
+Panels (a1–a3) show the curvycylinder dataset, where points form elongated, band-like structures that bend smoothly across projections, appearing cylindrical with gentle nonlinear curvature.
+Panels (b1–b3) show the sphericalspiral dataset, with points tracing curved, looping paths that resemble spiral arcs on a rounded surface, changing orientation across projections.
+Panels (c1–c3) show the conicspiral dataset, where point clouds expand outward in some projections while twisting along curved trajectories, creating tapered, spiral-like shapes.
+Panels (d1–d3) show the nonlinear dataset, in which points form warped surfaces with alternating tight and spread regions, visible as oscillating bands and smooth waves depending on the projection.
+Across all datasets, point density, curvature, and apparent continuity vary between projections, illustrating how different geometric structures appear under multiple 2-D views of the same four-dimensional data."}
 

### 2025-12-19 extra/paper-cardinalR-code.Rmd
Updated with new package versions

--- 2025-12-17 extra/paper-cardinalR-code.Rmd
@@
 
-### Cone
+### Branching
 
-To simulate a cone-shaped structure in arbitrary dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cone)"} else { "Figure \\@ref(fig:cone-proj)"})`), we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.
+A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology [@trapnell2014].
+We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, total number of points (`n`) generated across all branches, with points allocated approximately evenly among branches, and number of branches (`k`).
+Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.
 
-This function generates $n$ points in $p\text{-}D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip.
-Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, truncated to the interval $[0, h]$, producing a higher density of points near the tip.
-At each height $X_p$, the radius of the cross-section increases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio} \in [0, 1]$ and $r_{\text{max}} = 1$.
+The simplest structures are approximately linear branches in $2\text{-}D$, generated by the `gen_linearbranches(n, k)` function.
+These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability.
+Mathematically, each branch $i$ is defined as
 
-For each point, a direction in the first $p-1$ dimensions is sampled uniformly on a $(p-1)$-dimensional hypersphere using generalized spherical coordinates.
-The radial coordinates are scaled by the height-dependent radius $r$, producing the conical taper.
-In three dimensions ($p = 3$), this results in a classical $3\text{-}D$ cone, while for $p > 3$, additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.
+$$
+X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
+$$
 
-Cone-shaped structures appear in particle dispersions, light beams, and tapering processes, where spread decreases along one axis.
-They are also used to benchmark clustering and dimensionality reduction methods [@hadsell2006].
+where $(x_{\text{start},i}, y_{\text{start},i})$ is the starting point of branch $i$, $\delta$ controls local jitter, and $s_i$ is the slope, initialized as
 
-```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data.
-Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
-These projections show how the conical geometry is preserved.", fig.pos="H", fig.pos="!ht", fig.width=12, fig.height=4}
-cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)
+$$
+s_i =
+\begin{cases}
+0.5 & i = 1, \\
+-0.5 & i = 2, \\
+\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
+\end{cases}
+$$
+
+The jitter term is sampled from a one-sided uniform distribution to introduce directional variability without altering branch orientation.
+
+Branches $1$ and $2$ are initialized with fixed slopes and starting points, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.
+
+To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in $2\text{-}D$. Each branch follows a quadratic trajectory of the form
+
+$$
+X_1 \sim U(a_i, b_i), \quad X_2 = 0.1 X_1 + s_i X_1^2 + \epsilon, \quad \epsilon \sim U(-\delta, \delta),
+$$
+where $(a_i, b_i)$ defines the domain of the branch, $s_i$ controls curvature, and $\delta$ introduces local jitter.
+For the first two branches, the parameters are fixed to establish reference shapes: $(a_1,b_1,s_1) = (0,1,1), \quad (a_2,b_2,s_2) = (-1,0,-2)$.
+Additional branches are attached iteratively to existing structures.
+Each new branch $i$ starts at a selected point $(x_{\text{start},i}, y_{\text{start},i})$ from the current structure and extends according to
+
+$$
+X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
+$$
+
+where $s_i$ is a scale factor controlling the curvature of branch $i$.
+For the first few initial branches, $s_i$ can be fixed (e.g., $s_1 = 1, s_2 = 2$) to establish reference shapes, while for subsequent branches it is sampled from a predefined set, such as $s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}$, to control curvature magnitude and direction.
+
+The `gen_expbranches(n, k)` function creates $k$ exponential branches in $2\text{-}D$, radiating from a central region. Each branch $i$ is defined as
+
+$$
+X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
+$$
+
+where $\sigma_i = (-1)^{i+1}$ alternates the sign of the exponent to produce mirror-symmetric branches.
+The parameter $s_i$ controls the steepness of branch $i$, and $\delta$ introduces small local jitter.
+
+High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) and `gen_orgcurvybranches(n, p, k)`.
+For branch
+$i$, the active coordinate pair $(i_1, i_2)$ indexes the selected $2\text{-}D$ subspace.
+When `allow_share = TRUE`, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible $\binom{p}{2}$ combinations are exhausted, after which additional branches may repeat subspaces.
+
+In both cases, branch $i$ is generated according to
+
+$$
+X_{i_1} \sim U(a_i, b_i), \quad
+X_{i_2} = f_i(X_{i_1}) + \epsilon, \quad
+\epsilon \sim N(0, \sigma^2),
+$$
+where $a_i$ and $b_i$ define the domain of the branch and $\epsilon$ introduces smooth variability in the $p\text{-}D$ space. The function $f_i(\cdot)$ determines the branch geometry:
+$$
+f_i(x) =
+\begin{cases}
+s_i x, & \text{linear branches}, \\
+-s_i x^2, & \text{curvilinear branches}.
+\end{cases}
+$$
+
+The scale factor $s_i$ controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first $\binom{p}{2}$ branches, $s_i = 1$; for additional branches when $k > \binom{p}{2}$, $s_i$ is randomly drawn from the set $\{1, 1.5, 2, \dots, 8\}$.
+
+Across all branching generators, the scale parameter $s_i$ controls the strength of deviation from linearity, determining slope, curvature, or growth rate depending on the branch geometry.
+
+```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ \\texttt{orgcurvybranches} data.
+Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
+These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.pos="!ht"}
+
+set.seed(20240412)
 
@@
 }
-
-# Function to scale data manually
-# scale_data_manual <- function(data) {
-#   # Step 1: Center the data (mean 0)
-#   data_centered <- center_data(data)
-#
-#   # Step 2: Calculate the standard deviation of each dimension
-#   sds <- apply(data_centered, 2, sd)
-#
-#   # Step 3: Scale each dimension to have the range [0, 1]
-#   data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
-#
-#   # Step 4: Scale dimensions according to their variation
-#   # The dimension with the highest standard deviation is scaled to [-1, 1]
-#   # Other dimensions are scaled to smaller ranges based on their standard deviations
-#   max_sd <- max(sds)
-#
-#   # Normalize the standard deviations to get scaling factors
-#   scaling_factors <- sds / max_sd
-#
-#   for (i in seq_along(scaling_factors)) {
-#     data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
-#   }
-#
-#   # Combine the scaled data with the 'type' column and return as a tibble
-#   data_scaled <- as_tibble(data_scaled)
-#
-#   return(data_scaled)
-# }
 
@@
 
-centered_data <- center_data(cone)
+orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)
 
@@
 projection <- cbind(
-  c(0.22417,-0.02943,-0.02870,0.08418),
-  c(-0.02715,0.18019,0.05053,0.15254))
+    c(-0.083573,-0.019043,0.006574,-0.035774),
+    c(0.005380,0.068541,-0.030810,-0.054717))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.4, 
-                                              axis_pos_y = -0.4, 
-                                              threshold = 0.053))
+                            axis_param = list(limits = 0.5,
+                                              axis_scaled = 10, 
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.05))
 
-cone_proj1 <- plot_proj(
+orglinearbranches_proj1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.5, 0.7), 
+  plot_limits = c(-0.6, 0.4), 
   title = "a1", 
   cex = 2, 
-  axis_text_size = 5,
+  axis_text_size = 4,
   is_color = FALSE)
@@
 projection <- cbind(
-    c(0.15591,0.17907,0.03198,0.04496),
-    c(-0.15376,0.10435,0.15766,0.00547))
+  c(-0.06614,0.15134,-0.44278,0.27367),
+  c(-0.50493,0.14388,0.14615,0.03486))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.3,
-                                              axis_scaled = 2, 
-                                              axis_pos_x = -0.2, 
-                                              axis_pos_y = -0.2, 
-                                              threshold = 0.02))
+                            axis_param = list(limits = 0.9,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
+                                              threshold = 0.05))
 
-cone_proj2 <- plot_proj(
+orglinearbranches_proj2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.25, 0.35), 
+  plot_limits = c(-1, 1), 
   title = "a2", 
   cex = 2, 
-  axis_text_size = 5,
+  axis_text_size = 4,
   is_color = FALSE)
@@
 projection <- cbind(
-    c(-0.13548,0.13832,0.05991,-0.13544),
-    c(0.08479,-0.10719,-0.01572,-0.20123))
+  c(0.10704,0.04822,-0.03416,0.53224),
+  c(0.50329,0.07199,-0.16064,-0.11806))
 
@@
                             centered_data = centered_data, 
-                            axis_param = list(limits = 0.6,
-                                              axis_scaled = 3, 
-                                              axis_pos_x = -0.6, 
-                                              axis_pos_y = -0.6, 
-                                              threshold = 0.04))
+                            axis_param = list(limits = 0.9,
+                                              axis_scaled = 1.5, 
+                                              axis_pos_x = -0.85, 
+                                              axis_pos_y = -0.85, 
+                                              threshold = 0.1))
 
-cone_proj3 <- plot_proj(
+orglinearbranches_proj3 <- plot_proj(
   proj_obj = proj_obj3, 
   point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
-  plot_limits = c(-0.7, 0.5), 
+  plot_limits = c(-1, 1), 
   title = "a3", 
   cex = 2, 
-  axis_text_size = 5,
+  axis_text_size = 4,
   is_color = FALSE)
 
-cone_proj1 + cone_proj2 + cone_proj3 +
+orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
   plot_layout(ncol = 3, guides = "collect") 

### 2025-12-19 paper-cardinalR.tex
Updated with new package versions

@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Workflow for generating high-dimensional clustered data, from specifying input parameters to producing the final labeled dataset with optional noise dimensions.}]{figures/cardinalR_workflow}
+{\centering \includegraphics[width=0.8\linewidth,alt={The figure is a workflow diagram showing the process for generating high-dimensional clustered data.
+Boxes or nodes represent steps such as specifying input parameters (number of points, number of clusters, cluster locations, shapes, scaling, rotations, and optional background noise), generating each cluster with a shape generator, optionally rotating or scaling clusters, combining them into a single dataset, adding background noise if desired, and labeling each observation by cluster shape.
+Arrows indicate the flow of the process from parameter specification to the final labeled dataset.
+The diagram uses a clear layout with distinct steps, showing how data are constructed in sequence.}]{figures/cardinalR_workflow}
 
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={The figure shows a 2-D scatter plot of projected branch structures in a high-dimensional space.
-The horizontal axis represents one coordinate (first index of the active pair) and the vertical axis represents the other coordinate (second index of the active pair), each spanning a moderate, symmetric range around zero (for example, roughly −2 to 2).
-Points are grouped visually into several distinct linear or gently curving branches, each corresponding to a different active 2-D subspace; these branches appear as separate streaks or arcs radiating through the plane.
-Colour (or another visual attribute such as line/point grouping) is used to distinguish branches that occupy different subspaces, and some branches may overlap or run close together where subspaces are shared.
-The main pattern is a set of well-separated, structured branches with little noise, illustrating how high-dimensional linear or curvy branches project into 2-D coordinate pairs.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ \texttt{orgcurvybranches} data.
+\includegraphics[width=1\linewidth,alt={The figure consists of three 2‑D scatter plots, each showing a different 2‑D projection from the 4‑D orgcurvybranches dataset.
+In each plot, the horizontal axis represents one coordinate of the selected 2‑D projection, and the vertical axis represents the other coordinate.
+Points are plotted as individual markers and appear grouped into multiple elongated clusters or branches, which curve or radiate through the plot.
+Each branch is distinguished by colour, with overlapping branches occurring when subspaces are shared.
+The plots have roughly square aspect ratios, and together they show how the same high‑dimensional branching structure appears differently from three viewing angles.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ \texttt{orgcurvybranches} data.
 Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles.
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={ A 2D scatterplot shows the projection of a high-dimensional cone-shaped point cloud.
-The horizontal and vertical axes are unlabeled numeric coordinates, each spanning a roughly symmetric range around zero (for example, from about -1 to 1).
-Hundreds of points form a dense cluster near one end of the plot, corresponding to the cone’s apex, with density gradually decreasing as points extend outward in a flared, cone-like or triangular region.
-There are no colours, shapes, or facets used; all points share the same appearance, and the key visible pattern is the concentrated tip transitioning smoothly into a wider spread of points, illustrating a cone structure in projection.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $3\text{-}D$ \texttt{cone} data.
+\includegraphics[width=1\linewidth,alt={The figure consists of three 2‑D scatter plots, each showing a different 2‑D projection of the 4‑D cone dataset.
+In each plot, the horizontal and vertical axes represent two continuous numeric dimensions.
+Points are plotted individually and are concentrated near one end, forming a narrow tip, and gradually spread out along the other direction, creating a wider base.
+The points trace an elongated, triangular or cone-like shape in each projection.
+All points use the same colour and marker, with no additional visual encodings, and the plots have roughly square aspect ratios.
+Together, the three projections illustrate the distribution of points along the conical structure from different viewing angles.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ \texttt{cone} data.
 Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex.
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Three $2\text{-}D$ projections from each of four $4\text{-}D$ pyramid-like datasets: pyrrect, pyrtri, pyrstar, and pyrholes, showing their distinct geometries, branching, and hollow regions.}]{paper-cardinalR_files/figure-latex/pyr-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={The figure shows three 2‑D scatter plots for each of four 4‑D datasets: pyrrrect, pyrtri, pyrstar, and pyrholes.
+In each plot, the horizontal and vertical axes represent continuous numeric coordinates of the selected 2‑D projection, and points are plotted individually as small, uniformly coloured markers with no additional visual encodings.
+In pyrrrect, points form a dense rectangular base tapering to a narrow tip; in pyrtri, points form a triangular distribution with sharper edges; in pyrstar, points extend into multiple pointed branches radiating from a central region; and in pyrholes, points cluster in a compact shape but leave hollow or open regions within it.
+The plots maintain roughly square aspect ratios and illustrate how these different pyramid-like structures appear when projected from four to two dimensions.}]{paper-cardinalR_files/figure-latex/pyr-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Three $2\text{-}D$ projections from each of three datasets: circle, curvycycle, and clusteredspheres, showing closed loops, twisted rings, and compact spherical clusters, illustrating projection effects on structure, density, and separation.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={A 3D scatterplot shows a projection of points sampled from a 4‑dimensional hemisphere onto three coordinate axes.
+Each point represents a location on the surface of a 4‑D unit sphere restricted to one hemisphere.
+The horizontal (x) and vertical (y) axes, along with depth (z), span roughly from –1 to 1, reflecting the Cartesian coordinates of the projected points.
+The dots are uniformly scattered over a curved, bowl-like region rather than filling a full sphere, with no strong clusters or gaps, illustrating that the sampling covers only one side of the sphere.
+If colour is used, it likely encodes an additional coordinate or density but does not introduce distinct categorical groups.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Three $2\text{-}D$ projections from each of two trefoil knot datasets, showing complex $4\text{-}D$ trefoil loops and simpler 3-D knot structures, illustrating curvature, twisting, and separation across dimensions.}]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={The figure shows multiple scatterplots displaying two-dimensional projections of trefoil-shaped data derived from four- and three-dimensional spaces.
+In panels a1–a3, the trefoil4d data are shown as three different 2-D projections, where the horizontal and vertical axes represent linear combinations of the four original variables.
+The projected points form smooth, continuous looping curves that intersect and overlap in places, creating knot-like shapes whose apparent crossings and separations vary across projections.
+In panels b1–b3, the trefoil3d data are shown in three 2-D projections.
+The points trace a compact, closed loop with a characteristic trefoil-like knot structure, appearing smoother and less spread out than in the four-dimensional case.
+The figure highlights how the same underlying knot structure appears differently under changes in projection and dimensionality.}]{paper-cardinalR_files/figure-latex/trefoil-proj-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth,alt={Three $2\text{-}D$ projections from each of four $4\text{-}D$ datasets: curvycylinder, sphericalspiral, conicspiral, and nonlinear, showing cylindrical twists, spiral and conical helices, and warped $2\text{-}D$ surfaces, illustrating curvature, density, and continuity.}]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
+{\centering \includegraphics[width=0.7\linewidth,alt={The figure consists of four rows of panels, each displaying three separate 2-D projection scatterplots from a four-dimensional synthetic dataset.
+Panels (a1–a3) show the curvycylinder dataset, where points form elongated, band-like structures that bend smoothly across projections, appearing cylindrical with gentle nonlinear curvature.
+Panels (b1–b3) show the sphericalspiral dataset, with points tracing curved, looping paths that resemble spiral arcs on a rounded surface, changing orientation across projections.
+Panels (c1–c3) show the conicspiral dataset, where point clouds expand outward in some projections while twisting along curved trajectories, creating tapered, spiral-like shapes.
+Panels (d1–d3) show the nonlinear dataset, in which points form warped surfaces with alternating tight and spread regions, visible as oscillating bands and smooth waves depending on the projection.
+Across all datasets, point density, curvature, and apparent continuity vary between projections, illustrating how different geometric structures appear under multiple 2-D views of the same four-dimensional data.}]{paper-cardinalR_files/figure-latex/triginometric-proj-1}
 
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Three $2\text{-}D$ projections of five $4\text{-}D$ synthetic clusters: helical spiral (dark green), hemisphere (orange), uniform cube (purple), cone (pink), and Gaussian (light green), showing spatial separation and cluster geometry.}]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
+\includegraphics[width=1\linewidth,alt={The figure shows three 2-D projection plots of synthetic 4-D data containing five clusters with distinct geometric structures.
+The horizontal and vertical axes represent two projected dimensions, and no explicit units are shown.
+Points are coloured by cluster, with distinct colours representing the helical spiral, hemisphere, uniform cube, cone, and Gaussian clusters.
+The helical spiral cluster appears as an elongated, twisted band; the hemisphere cluster forms a curved, arc-like concentration; the uniform cube cluster fills a roughly rectangular region with fairly even density; the cone cluster exhibits a density gradient, narrowing at one end and spreading out at the other; and the Gaussian cluster forms a compact, approximately circular group near the centre of the plot.
+The clusters appear very close to one another in these projections, illustrating the difficulty of distinguishing cluster structure when projections bring clusters into close proximity.}]{paper-cardinalR_files/figure-latex/highd-proj-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data.
 The helical spiral cluster is represented in dark green, the hemisphere cluster in orange, the uniform cube-shaped cluster in purple, the blunted cone cluster in pink, and the Gaussian-shaped cluster in light green.}\label{fig:highd-proj}
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Composite plot assessing six NLDR layouts of five clusters data using HBE over varying binwidths, showing tSNE, UMAP, PAHTE, TriMAP, PaCMAP, and PCA.
-Layouts a and b are optimal, f is poor, with color-coded lines matching the scatterplots of each layout.}]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth,alt={A multi-panel figure compares 2-D scatterplots from six nonlinear dimensionality reduction methods applied to the same dataset with five true clusters.
+Each panel plots the first embedding dimension on the horizontal axis and the second embedding dimension on the vertical axis.
+Points represent individual observations and are coloured by cluster membership, with six distinct cluster colours reused consistently across panels.
+In the tSNE panel, clusters form compact, clearly separated groups with small gaps between clusters, indicating strong preservation of both local neighbourhoods and the global cluster layout.
+UMAP and PaCMAP also show six visibly distinct clusters that are moderately well separated but with slightly more overlap and distortion than tSNE.
+PHATE produces curved, nonlinear cluster shapes where clusters are stretched or intertwined, obscuring the original simple cluster geometry.
+TriMAP collapses the data into three main visible groups instead of six, with only small distances between these groups, suggesting loss of finer cluster structure.
+PCA displays the weakest structure: clusters overlap substantially and align along a roughly linear or planar trend, failing to reflect the underlying non-linear separation among the six groups.}]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
 Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
@@
 \begin{figure}[!]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Cluster validity metrics for solutions with $2-10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
+\includegraphics[width=1\linewidth,alt={Multi-panel line chart showing six cluster quality metrics across numbers of clusters (x-axis: 2 to 10 clusters) for three clustering methods (k-means, hierarchical, model-based).
+Each panel displays one metric with the y-axis scaled to that metric’s values (not explicitly labeled in the text, but each metric is treated as higher-is-better or lower-is-better as described).
+Within each panel, three lines (one per method) trace the metric value as the number of clusters increases.
+For Pearson gamma, values for all three methods rise steeply up to about 5 clusters and then level off.
+For the Calinski–Harabasz index, values increase sharply from 4 to 5 clusters.
+For Dunn, the k-means line peaks around 4 clusters, while the hierarchical and model-based lines peak around 5.
+For WB ratio and within-cluster sum of squares, all three methods show a generally monotonic decline as the number of clusters increases, with a visible bend or elbow around 5 clusters.
+For the S-index, the k-means line reaches its best value around 4 clusters, the hierarchical line around 3, 6, or 8 clusters, and the model-based line around 4 or 8.
+Across panels, the k-means line is typically at or near the most favorable values for each metric, and several metrics simultaneously favor solutions with about 4–5 clusters.}]{paper-cardinalR_files/figure-latex/fig-cluster-stats-1} \caption{Cluster validity metrics for solutions with $2-10$ clusters obtained using $k$-means, hierarchical, and model-based clustering.
 Several indices consistently suggest that $4-5$ clusters provide the best balance of separation and compactness, with $k$-means performing slightly better across metrics.}\label{fig:fig-cluster-stats}
@@
 Although the metrics differ in their preferences, several show consistent support for a \(4-5\) cluster solution.
-Pearson gamma (\texttt{pearsongamma}) increases sharply up to five clusters before levelling off, Calinski--Harabasz index (\texttt{ch}) increases sharply from 4 to 5 clusters and Dunn (\texttt{dunn2}) has a maximum at 5 for two methods and at 4 for \(k\)-means.
+Pearson gamma (\texttt{pearsongamma}) increases sharply up to five clusters before leveling off, Calinski--Harabasz index (\texttt{ch}) increases sharply from 4 to 5 clusters and Dunn (\texttt{dunn2}) has a maximum at 5 for two methods and at 4 for \(k\)-means.
 All of these are interpreted as higher is better.
@@
 WB ratio (\texttt{wb.ratio}) and within-cluster sum of squares (\texttt{within.cluster.ss}) steadily decline with number of clusters, possibly elbowing around 5 clusters.
-The S-index (\texttt{sindex}) is optimised at 4 clusters for \(k\)-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.
+The S-index (\texttt{sindex}) is optimized at 4 clusters for \(k\)-means, 3, 6 or 8 for hierarchical clustering, and 4 or 8 for model-based.
 Overall, \(k\)-means performs slightly better than the hierarchical and model-based clustering across most metrics and number of clusters.
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/highd-proj-clust-algo-pdf-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data coloured by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
+\includegraphics[width=1\linewidth,alt={The figure shows three row-wise 2-D projection plots of synthetic 4-D data for two different k-means clustering solutions: four clusters in one panel and five clusters in the other.
+In each panel, the horizontal and vertical axes represent linear projections of the original high-dimensional variables, and each point corresponds to a data observation.
+Points are coloured by assigned cluster ID (four colours in the top row and five in the bottom row).
+In both solutions, clusters appear as roughly compact groups, but none aligns with the underlying geometric shapes in the data: each of the five true shapes is split across multiple clusters, and no cluster cleanly isolates a single shape.
+The four-cluster solution merges some shapes more strongly, while the five-cluster solution yields a finer partition that still fragments each shape; additional clusters beyond five would be needed to better match the original geometric structure.}]{paper-cardinalR_files/figure-latex/highd-proj-clust-algo-pdf-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data coloured by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
 The intermixing of colours within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.}\label{fig:highd-proj-clust-algo-pdf}

### 2025-12-19 paper-cardinalR.Rmd
Updated the sepllings. Updated the summary sentence in the application.

@@
 abstract: >
-  Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
+  Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimension reduction, supervised and unsupervised learning.
 High-dimensional data is characterized by multiple variables that are dependent or associated in some way, such as linear, nonlinear, clustering or anomalies.
@@
 The example data structures are best viewed using a tour [@As85].
-These show the data as a sequence of low dimensional projections (typically 2D), providing a good sense of the shape in high dimensions.
-The interactive tour plots included in this paper are produced using the software langevitour [@paul2023].
-
-The paper is organized as follows.
-In the next section, we introduce the implementation of the `cardinalR` package on GitHub, including a demonstration of the package's key functions.
-We illustrate how a clustering data structure affects the dimension reductions in the Application section.
+These show the data as a sequence of low dimensional projections (typically $2\text{-}D$), providing a good sense of the shape in high dimensions.
+The interactive tour plots included in this paper are produced using the software `langevitour` [@paul2023].
+
+The next section provides an overview of the usage of the `cardinalR` package, illustrating how its modular components can be combined to generate complex high-dimensional datasets.
+This is followed by a section describing the implementation of the package, including its design principles and key functions.
+The Application section then demonstrates how the simulated clustering structures can be used to evaluate and compare dimension reduction and clustering methods.
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for the use of our data collection.
@@
 Points appear grouped into multiple elongated clusters or branches that move, rotate, and sometimes overlap as the projection changes.
-Branches are distinguished by colour, with individual points visible throughout.
+Branches are distinguished by color, with individual points visible throughout.
 The display maintains a roughly square aspect ratio as the view transitions through multiple angles, illustrating the dataset’s branching structure from different perspectives."}
@@
 Points are plotted as individual markers and appear grouped into multiple elongated clusters or branches, which curve or radiate through the plot.
-Each branch is distinguished by colour, with overlapping branches occurring when subspaces are shared.
+Each branch is distinguished by color, with overlapping branches occurring when subspaces are shared.
 The plots have roughly square aspect ratios, and together they show how the same high‑dimensional branching structure appears differently from three viewing angles.", fig.pos="!ht"}
@@
 Points are densely packed near one end, forming a narrow tip, and gradually spread out along the other direction, creating a wider base.
-All points share the same colour and marker.
+All points share the same color and marker.
 As the projection rotates, the cone-like structure shifts in orientation, maintaining roughly square aspect ratios throughout the animation and showing the distribution of points from multiple viewing angles.", fig.pos="!ht"}
@@
 The points trace an elongated, triangular or cone-like shape in each projection.
-All points use the same colour and marker, with no additional visual encodings, and the plots have roughly square aspect ratios.
+All points use the same color and marker, with no additional visual encodings, and the plots have roughly square aspect ratios.
 Together, the three projections illustrate the distribution of points along the conical structure from different viewing angles.", fig.pos="!ht", fig.width=12, fig.height=4}
@@
 Gaussian noise $\varepsilon_{ij} \sim N(0, (0.03n)^2)$ is added independently across observations and dimensions.
-The observed variables are then defined as $X_{ij} = a_j \bigl(t_i + b_j + \varepsilon_{ij}\bigr), \quad i = 1, \dots, n$.cThis construction yields a single elongated linear structure embedded in $p\text{-}D$, with each dimension exhibiting a different orientation, scale, and offset.
+The observed variables are then defined as $X_{ij} = a_j \bigl(t_i + b_j + \varepsilon_{ij}\bigr), \quad i = 1, \dots, n$.
+This construction yields a single elongated linear structure embedded in $p\text{-}D$, with each dimension exhibiting a different orientation, scale, and offset.
 
@@
 
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in  space.
+A polynomial structure generates data points that follow nonlinear curvilinear relationships, such as quadratic or cubic trends, in $2\text{-}D$ space.
 To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
-These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
+These patterns are useful for evaluating how well algorithms capture smooth, nonlinear trajectories and curvature in the data.
 We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
@@
 Points are displayed in continuously changing 2‑D projections, with each axis representing a linear combination of the four original dimensions.
-Points are plotted individually as small, uniformly coloured markers with no additional visual encodings.
+Points are plotted individually as small, uniformly colored markers with no additional visual encodings.
 In pyrrrect, points form a dense rectangular base tapering to a narrow tip; in pyrtri, points form a triangular distribution with sharper edges; in pyrstar, points extend into multiple pointed branches radiating from a central region; and in pyrholes, points cluster in a compact shape but leave hollow or open regions within it.
@@
 These projections illustrate a range of pyramid-like geometries that vary in density and structure.", fig.width=12, fig.height=16, out.width="80%", fig.align='center', fig.alt="The figure shows three 2‑D scatter plots for each of four 4‑D datasets: pyrrrect, pyrtri, pyrstar, and pyrholes.
-In each plot, the horizontal and vertical axes represent continuous numeric coordinates of the selected 2‑D projection, and points are plotted individually as small, uniformly coloured markers with no additional visual encodings.
+In each plot, the horizontal and vertical axes represent continuous numeric coordinates of the selected 2‑D projection, and points are plotted individually as small, uniformly colored markers with no additional visual encodings.
 In pyrrrect, points form a dense rectangular base tapering to a narrow tip; in pyrtri, points form a triangular distribution with sharper edges; in pyrstar, points extend into multiple pointed branches radiating from a central region; and in pyrholes, points cluster in a compact shape but leave hollow or open regions within it.
@@
 
-Building on simple circular structures, the `gen_unifsphere(n, r)` function function extends the idea to three dimensions by generating $n$ observations approximately uniformly distributed on the surface of a sphere of radius $r$..
+Building on simple circular structures, the `gen_unifsphere(n, r)` function extends the idea to three dimensions by generating $n$ observations approximately uniformly distributed on the surface of a sphere of radius $r$.
 Each observation is computed from spherical coordinates, with $u \sim U(-1, 1)$ representing $\cos(\phi)$ and $\theta \sim U(0, 2\pi)$ the azimuthal angle.
@@
 
-In addition, the `gen_gridedsphere(n)` function constructs a $p$-dimensional dataset consisting of approximately $n$ points arranged on the surface of the unit $(p-1)\text{-}D$ sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
+In addition, the `gen_gridedsphere(n)` function constructs a $p$-dimensional dataset consisting of approximately $n$ points arranged on the surface of the unit $(p-1)$-sphere embedded in $\mathbb{R}^p$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:sphere) d"} else { "Figure \\@ref(fig:sphere-proj) d"})`).
 Rather than sampling points uniformly, this function creates a deterministic grid in spherical coordinates, using $(p-1)$ angular variables: the first $(p-2)$ angles are taken from $[0, \pi]$, and the final angle from $[0, 2\pi]$.
@@
 The dots are uniformly scattered over a curved, bowl-like region rather than filling a full sphere, with no strong clusters or gaps, illustrating that the sampling covers only one side of the sphere.
-If colour is used, it likely encodes an additional coordinate or density but does not introduce distinct categorical groups.", fig.pos="!ht"}
+If color is used, it likely encodes an additional coordinate or density but does not introduce distinct categorical groups.", fig.pos="!ht"}
 
@@
 
-Second, the `gen_curvycylinder(n, p, h)` function generates a $p$-dimensional dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
+Second, the `gen_curvycylinder(n, p, h)` function generates a $p\text{-}D$ dataset of $n$ observations structured as a $3\text{-}D$ cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when $p > 4$ (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) b"} else { "Figure \\@ref(fig:triginometric-proj) b"})`).
 The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
@@
 
-Finally, the `gen_nonlinear(n, p, hc, non_fac)` function simulates a non-linear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
-The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{non\_fac}\sin(X_{1})$: non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
+Finally, the `gen_nonlinear(n, p, hc, non_fac)` function simulates a nonlinear $2\text{-}D$ surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:triginometric) f"} else { "Figure \\@ref(fig:triginometric-proj) f"})`).
+The $X_{1} \sim U(0.1, 2)$: base variable (avoids zero to prevent division errors), $X_{3} \sim U(0.1, 0.8)$: independent auxiliary variable, $X_{2} = hc/X_{1} + \text{non\_fac}\sin(X_{1})$: nonlinear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and $X_{4} = \cos(\pi X_{1}) + \varepsilon$, with $\varepsilon \sim U(-0.1, 0.1)$: additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+These transformations together result in a nonlinear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
 
@@
 
-High-dimensional data structures often benefit from the addition of auxiliary noise dimensions, which can be used to assess the robustness of dimensionality reduction and clustering algorithms.
-The functions in this section provide flexible ways to generate random noise dimensions, ranging from purely random Gaussian variables to more structured, wavy patterns that mimic non-linear distortions in high-dimensional space.
+High-dimensional data structures often benefit from the addition of auxiliary noise dimensions, which can be used to assess the robustness of dimension reduction and clustering algorithms.
+The functions in this section provide flexible ways to generate random noise dimensions, ranging from purely random Gaussian variables to more structured, wavy patterns that mimic nonlinear distortions in high-dimensional space.
 These functions can be applied independently or combined with other geometric structures to create complex simulated datasets.
@@
 
-The `gen_wavydims2(n, p, x_1)` function extends this approach by applying a non-linear transformation to an existing dimension vector $x_1$:
+The `gen_wavydims2(n, p, x_1)` function extends this approach by applying a nonlinear transformation to an existing dimension vector $x_1$:
 
@@
 Finally, `gen_wavydims3(n, p, data)` generates noise for datasets with multiple correlated dimensions.
-The first three dimensions are small perturbations of the original coordinates $(X_1, X_2, X_3)$, while higher dimensions are constructed via non-linear combinations, including polynomial and trigonometric transformations, e.g.,
+The first three dimensions are small perturbations of the original coordinates $(X_1, X_2, X_3)$, while higher dimensions are constructed via nonlinear combinations, including polynomial and trigonometric transformations, e.g.,
 
@@
 Each frame presents a different linear projection of the data, with the horizontal and vertical axes representing changing combinations of the original variables.
-Points represent individual observations and are coloured by cluster membership, with five distinct colours corresponding to the helical spiral, hemisphere, uniform cube, cone, and Gaussian clusters.
-As the tour progresses, the relative positions, shapes, and overlaps of the coloured point clouds change continuously, revealing elongated bands, curved arcs, compact clusters, and more diffuse regions depending on the projection.
+Points represent individual observations and are colored by cluster membership, with five distinct colors corresponding to the helical spiral, hemisphere, uniform cube, cone, and Gaussian clusters.
+As the tour progresses, the relative positions, shapes, and overlaps of the colored point clouds change continuously, revealing elongated bands, curved arcs, compact clusters, and more diffuse regions depending on the projection.
 Clusters frequently move closer together or overlap in certain views, while becoming more separated in others, illustrating how apparent structure varies across different two-dimensional projections of the same high-dimensional data.", fig.pos="!ht"}
@@
 The horizontal and vertical axes represent two projected dimensions, and no explicit units are shown.
-Points are coloured by cluster, with distinct colours representing the helical spiral, hemisphere, uniform cube, cone, and Gaussian clusters.
+Points are colored by cluster, with distinct colors representing the helical spiral, hemisphere, uniform cube, cone, and Gaussian clusters.
 The helical spiral cluster appears as an elongated, twisted band; the hemisphere cluster forms a curved, arc-like concentration; the uniform cube cluster fills a roughly rectangular region with fairly even density; the cone cluster exhibits a density gradient, narrowing at one end and spreading out at the other; and the Gaussian cluster forms a compact, approximately circular group near the centre of the plot.
@@
 ```{r, label = "fig-nldr-layouts", fig.pos="!ht", fig.cap="Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
-Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
+Color is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
 Layout f is universally poor.
@@
 Each panel plots the first embedding dimension on the horizontal axis and the second embedding dimension on the vertical axis.
-Points represent individual observations and are coloured by cluster membership, with six distinct cluster colours reused consistently across panels.
+Points represent individual observations and are colored by cluster membership, with six distinct cluster colors reused consistently across panels.
 In the tSNE panel, clusters form compact, clearly separated groups with small gaps between clusters, indicating strong preservation of both local neighbourhoods and the global cluster layout.
@@
 Its layout displays well-separated clusters with minimal inter-cluster distances, making it the most faithful representation of the underlying data structure.
-UMAP and PaCMAP (Figure \@ref(fig:fig-nldr-layouts) b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \@ref(fig:fig-nldr-layouts) c) show non-linear cluster structures irrespective of the original structure.
+UMAP and PaCMAP (Figure \@ref(fig:fig-nldr-layouts) b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \@ref(fig:fig-nldr-layouts) c) show nonlinear cluster structures irrespective of the original structure.
 Also, TriMAP (Figure \@ref(fig:fig-nldr-layouts) d) has high HBE, and show three clusters with small distances.
@@
 
-```{r highd-data-clusters-algo-html, eval=knitr::is_html_output(), fig.cap="Views of the five-cluster synthetic dataset coloured by the $k$-means four- and five-cluster solutions.
+```{r highd-data-clusters-algo-html, eval=knitr::is_html_output(), fig.cap="Views of the five-cluster synthetic dataset colored by the $k$-means four- and five-cluster solutions.
 These show how the five true structures: helical spiral, hemisphere, uniform cube, blunted cone, and Gaussian-shaped cluster appear under $k$-means partitions.
-Both the $4$- and $5$-cluster solutions mix colours, revealing substantial misclassification and limited ability of $k$-means to recover the underlying geometric structure.", fig.alt="The figure contains two side-by-side tour views of the same synthetic high-dimensional dataset, each shown as a sequence of continuously changing 2-D projections.
-The left tour view is coloured by a four-cluster k-means solution, and the right tour view is coloured by a five-cluster k-means solution.
-In both views, each point represents a data observation, and colours indicate cluster membership.
+Both the $4$- and $5$-cluster solutions mix colors, revealing substantial misclassification and limited ability of $k$-means to recover the underlying geometric structure.", fig.alt="The figure contains two side-by-side tour views of the same synthetic high-dimensional dataset, each shown as a sequence of continuously changing 2-D projections.
+The left tour view is colored by a four-cluster k-means solution, and the right tour view is colored by a five-cluster k-means solution.
+In both views, each point represents a data observation, and colors indicate cluster membership.
 The horizontal and vertical axes correspond to linear projections of the original variables.
 Across the tours, points form rotating and deforming point clouds that intermittently reveal curved, elongated, compact, and diffuse structures.
-Colours are intermingled throughout most projections in both panels, with clusters frequently overlapping and splitting as the view changes.
-The two tours differ only in the number of colours used, with four colours on the left and five on the right, while the underlying point positions and projection motion remain the same.", fig.pos="!ht", layout = "l-body"}
+Colors are intermingled throughout most projections in both panels, with clusters frequently overlapping and splitting as the view changes.
+The two tours differ only in the number of colors used, with four colors on the left and five on the right, while the underlying point positions and projection motion remain the same.", fig.pos="!ht", layout = "l-body"}
 
@@
 
-```{r highd-proj-clust-algo-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data coloured by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
-The intermixing of colours within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.", fig.alt="The figure shows three row-wise 2-D projection plots of synthetic 4-D data for two different k-means clustering solutions: four clusters in one panel and five clusters in the other.
+```{r highd-proj-clust-algo-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the five clusters data colored by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
+The intermixing of colors within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.", fig.alt="The figure shows three row-wise 2-D projection plots of synthetic 4-D data for two different k-means clustering solutions: four clusters in one panel and five clusters in the other.
 In each panel, the horizontal and vertical axes represent linear projections of the original high-dimensional variables, and each point corresponds to a data observation.
-Points are coloured by assigned cluster ID (four colours in the top row and five in the bottom row).
+Points are colored by assigned cluster ID (four colors in the top row and five in the bottom row).
 In both solutions, clusters appear as roughly compact groups, but none aligns with the underlying geometric shapes in the data: each of the five true shapes is split across multiple clusters, and no cluster cleanly isolates a single shape.
@@
 
-`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` shows the four- and five-cluster $k$-means solutions, with cluster id used to colour the points.
+`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:highd-data-clusters-algo-html)"} else { "Figure \\@ref(fig:highd-proj-clust-algo-pdf)"})` shows the four- and five-cluster $k$-means solutions, with cluster id used to color the points.
 Neither solution captures the geometric nature of the true clusters, but they are both reasonable partitions of the data.

### 2025-12-19 paper-cardinalR.tex
Updated the sepllings. Updated the summary sentence in the application.

@@
 \abstract{%
-Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning.
+Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimension reduction, supervised and unsupervised learning.
 High-dimensional data is characterized by multiple variables that are dependent or associated in some way, such as linear, nonlinear, clustering or anomalies.
@@
 The example data structures are best viewed using a tour \citep{As85}.
-These show the data as a sequence of low dimensional projections (typically 2D), providing a good sense of the shape in high dimensions.
-The interactive tour plots included in this paper are produced using the software langevitour \citep{paul2023}.
+These show the data as a sequence of low dimensional projections (typically \(2\text{-}D\)), providing a good sense of the shape in high dimensions.
+The interactive tour plots included in this paper are produced using the software \texttt{langevitour} \citep{paul2023}.
 
-The paper is organized as follows.
-In the next section, we introduce the implementation of the \texttt{cardinalR} package on GitHub, including a demonstration of the package's key functions.
-We illustrate how a clustering data structure affects the dimension reductions in the Application section.
+The next section provides an overview of the usage of the \texttt{cardinalR} package, illustrating how its modular components can be combined to generate complex high-dimensional datasets.
+This is followed by a section describing the implementation of the package, including its design principles and key functions.
+The Application section then demonstrates how the simulated clustering structures can be used to evaluate and compare dimension reduction and clustering methods.
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for the use of our data collection.
@@
 Points are plotted as individual markers and appear grouped into multiple elongated clusters or branches, which curve or radiate through the plot.
-Each branch is distinguished by colour, with overlapping branches occurring when subspaces are shared.
+Each branch is distinguished by color, with overlapping branches occurring when subspaces are shared.
 The plots have roughly square aspect ratios, and together they show how the same high‑dimensional branching structure appears differently from three viewing angles.}]{paper-cardinalR_files/figure-latex/branch-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ \texttt{orgcurvybranches} data.
@@
 The points trace an elongated, triangular or cone-like shape in each projection.
-All points use the same colour and marker, with no additional visual encodings, and the plots have roughly square aspect ratios.
+All points use the same color and marker, with no additional visual encodings, and the plots have roughly square aspect ratios.
 Together, the three projections illustrate the distribution of points along the conical structure from different viewing angles.}]{paper-cardinalR_files/figure-latex/cone-proj-1} \caption{Three $2\text{-}D$ projections from the $4\text{-}D$ \texttt{cone} data.
@@
 Gaussian noise \(\varepsilon_{ij} \sim N(0, (0.03n)^2)\) is added independently across observations and dimensions.
-The observed variables are then defined as \(X_{ij} = a_j \bigl(t_i + b_j + \varepsilon_{ij}\bigr), \quad i = 1, \dots, n\).cThis construction yields a single elongated linear structure embedded in \(p\text{-}D\), with each dimension exhibiting a different orientation, scale, and offset.
+The observed variables are then defined as \(X_{ij} = a_j \bigl(t_i + b_j + \varepsilon_{ij}\bigr), \quad i = 1, \dots, n\).
+This construction yields a single elongated linear structure embedded in \(p\text{-}D\), with each dimension exhibiting a different orientation, scale, and offset.
 
@@
 
-A polynomial structure generates data points that follow non-linear curvilinear relationships, such as quadratic or cubic trends, in space.
+A polynomial structure generates data points that follow nonlinear curvilinear relationships, such as quadratic or cubic trends, in \(2\text{-}D\) space.
 To extend these patterns into high-dimensional settings, additional noise dimensions can be added.
-These patterns are useful for evaluating how well algorithms capture smooth, non-linear trajectories and curvature in the data.
+These patterns are useful for evaluating how well algorithms capture smooth, nonlinear trajectories and curvature in the data.
 We provide functions for generating quadratic and cubic structures, enabling controlled experiments with different degrees of polynomial complexity.
@@
 {\centering \includegraphics[width=0.8\linewidth,alt={The figure shows three 2‑D scatter plots for each of four 4‑D datasets: pyrrrect, pyrtri, pyrstar, and pyrholes.
-In each plot, the horizontal and vertical axes represent continuous numeric coordinates of the selected 2‑D projection, and points are plotted individually as small, uniformly coloured markers with no additional visual encodings.
+In each plot, the horizontal and vertical axes represent continuous numeric coordinates of the selected 2‑D projection, and points are plotted individually as small, uniformly colored markers with no additional visual encodings.
 In pyrrrect, points form a dense rectangular base tapering to a narrow tip; in pyrtri, points form a triangular distribution with sharper edges; in pyrstar, points extend into multiple pointed branches radiating from a central region; and in pyrholes, points cluster in a compact shape but leave hollow or open regions within it.
@@
 
-Building on simple circular structures, the \texttt{gen\_unifsphere(n,\ r)} function function extends the idea to three dimensions by generating \(n\) observations approximately uniformly distributed on the surface of a sphere of radius \(r\)..
+Building on simple circular structures, the \texttt{gen\_unifsphere(n,\ r)} function extends the idea to three dimensions by generating \(n\) observations approximately uniformly distributed on the surface of a sphere of radius \(r\).
 Each observation is computed from spherical coordinates, with \(u \sim U(-1, 1)\) representing \(\cos(\phi)\) and \(\theta \sim U(0, 2\pi)\) the azimuthal angle.
@@
 
-In addition, the \texttt{gen\_gridedsphere(n)} function constructs a \(p\)-dimensional dataset consisting of approximately \(n\) points arranged on the surface of the unit \((p-1)\text{-}D\) sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
+In addition, the \texttt{gen\_gridedsphere(n)} function constructs a \(p\)-dimensional dataset consisting of approximately \(n\) points arranged on the surface of the unit \((p-1)\)-sphere embedded in \(\mathbb{R}^p\) (Figure \ref{fig:sphere-proj} d).
 Rather than sampling points uniformly, this function creates a deterministic grid in spherical coordinates, using \((p-1)\) angular variables: the first \((p-2)\) angles are taken from \([0, \pi]\), and the final angle from \([0, 2\pi]\).
@@
 The dots are uniformly scattered over a curved, bowl-like region rather than filling a full sphere, with no strong clusters or gaps, illustrating that the sampling covers only one side of the sphere.
-If colour is used, it likely encodes an additional coordinate or density but does not introduce distinct categorical groups.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
+If color is used, it likely encodes an additional coordinate or density but does not introduce distinct categorical groups.}]{paper-cardinalR_files/figure-latex/sphere-proj-1}
 
@@
 
-Second, the \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\)-dimensional dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
+Second, the \texttt{gen\_curvycylinder(n,\ p,\ h)} function generates a \(p\text{-}D\) dataset of \(n\) observations structured as a \(3\text{-}D\) cylindrical manifold with an added nonlinear curvy dimension, and optional noise dimensions when \(p > 4\) (Figure \ref{fig:triginometric-proj} b).
 The core structure consists of a circular base and height values, extended by a nonlinear fourth dimension.
@@
 
-Finally, the \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a non-linear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
-The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{non\_fac}\sin(X_{1})\): non-linear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
-These transformations together result in a non-linear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
+Finally, the \texttt{gen\_nonlinear(n,\ p,\ hc,\ non\_fac)} function simulates a nonlinear \(2\text{-}D\) surface embedded in higher dimensions, constructed using inverse and trigonometric transformations applied to independent variables (Figure \ref{fig:triginometric-proj} f).
+The \(X_{1} \sim U(0.1, 2)\): base variable (avoids zero to prevent division errors), \(X_{3} \sim U(0.1, 0.8)\): independent auxiliary variable, \(X_{2} = hc/X_{1} + \text{non\_fac}\sin(X_{1})\): nonlinear combination of hyperbolic and sinusoidal transformations, creating sharp curvature and oscillation, and \(X_{4} = \cos(\pi X_{1}) + \varepsilon\), with \(\varepsilon \sim U(-0.1, 0.1)\): additional nonlinear variation based on cosine, simulating more subtle periodic structure.
+These transformations together result in a nonlinear surface warped in multiple ways: sharp vertical shifts due to inverse terms, smooth waves from sine and cosine, and additional jitter.
 
@@
 
-The \texttt{gen\_wavydims2(n,\ p,\ x\_1)} function extends this approach by applying a non-linear transformation to an existing dimension vector \(x_1\):
+The \texttt{gen\_wavydims2(n,\ p,\ x\_1)} function extends this approach by applying a nonlinear transformation to an existing dimension vector \(x_1\):
 
@@
 Finally, \texttt{gen\_wavydims3(n,\ p,\ data)} generates noise for datasets with multiple correlated dimensions.
-The first three dimensions are small perturbations of the original coordinates \((X_1, X_2, X_3)\), while higher dimensions are constructed via non-linear combinations, including polynomial and trigonometric transformations, e.g.,
+The first three dimensions are small perturbations of the original coordinates \((X_1, X_2, X_3)\), while higher dimensions are constructed via nonlinear combinations, including polynomial and trigonometric transformations, e.g.,
 
@@
 PCA displays the weakest structure: clusters overlap substantially and align along a roughly linear or planar trend, failing to reflect the underlying non-linear separation among the six groups.}]{paper-cardinalR_files/figure-latex/fig-nldr-layouts-1} \caption{Assessing which of the 6 NLDR layouts ((a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA) of the five clusters data is the better representation using HBE for varying binwidth ($a_1$).
-Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
+Color is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
 Layout f is universally poor.
@@
 Its layout displays well-separated clusters with minimal inter-cluster distances, making it the most faithful representation of the underlying data structure.
-UMAP and PaCMAP (Figure \ref{fig:fig-nldr-layouts} b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \ref{fig:fig-nldr-layouts} c) show non-linear cluster structures irrespective of the original structure.
+UMAP and PaCMAP (Figure \ref{fig:fig-nldr-layouts} b and e) produced moderately accurate embeddings, although the six clusters appear more well-separated, while PHATE (Figure \ref{fig:fig-nldr-layouts} c) show nonlinear cluster structures irrespective of the original structure.
 Also, TriMAP (Figure \ref{fig:fig-nldr-layouts} d) has high HBE, and show three clusters with small distances.
@@
 In each panel, the horizontal and vertical axes represent linear projections of the original high-dimensional variables, and each point corresponds to a data observation.
-Points are coloured by assigned cluster ID (four colours in the top row and five in the bottom row).
+Points are colored by assigned cluster ID (four colors in the top row and five in the bottom row).
 In both solutions, clusters appear as roughly compact groups, but none aligns with the underlying geometric shapes in the data: each of the five true shapes is split across multiple clusters, and no cluster cleanly isolates a single shape.
-The four-cluster solution merges some shapes more strongly, while the five-cluster solution yields a finer partition that still fragments each shape; additional clusters beyond five would be needed to better match the original geometric structure.}]{paper-cardinalR_files/figure-latex/highd-proj-clust-algo-pdf-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data coloured by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
-The intermixing of colours within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.}\label{fig:highd-proj-clust-algo-pdf}
+The four-cluster solution merges some shapes more strongly, while the five-cluster solution yields a finer partition that still fragments each shape; additional clusters beyond five would be needed to better match the original geometric structure.}]{paper-cardinalR_files/figure-latex/highd-proj-clust-algo-pdf-1} \caption{Three $2\text{-}D$ projections from $4\text{-}D$, for the five clusters data colored by the $k$-means four- (a1-a3) and five-cluster (b1-b3) solutions.
+The intermixing of colors within each projection reflects misclassification in both solutions, showing the difficulty of using $k$-means to capture the dataset’s nonlinear and heterogeneous shapes.}\label{fig:highd-proj-clust-algo-pdf}
 \end{figure}
 
-Figure \ref{fig:highd-proj-clust-algo-pdf} shows the four- and five-cluster \(k\)-means solutions, with cluster id used to colour the points.
+Figure \ref{fig:highd-proj-clust-algo-pdf} shows the four- and five-cluster \(k\)-means solutions, with cluster id used to color the points.
 Neither solution captures the geometric nature of the true clusters, but they are both reasonable partitions of the data.

### 2025-12-19 paper-cardinalR.R

@@ -392,7 +392,7 @@ orglinearbranches_proj3 <- plot_proj(
 orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 
 
-## ----branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `orgcurvybranches` data. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body", fig.alt="The figure shows a tour of the 4‑D orgcurvybranches dataset. Points are displayed in a continuously changing 2‑D projection that rotates through the four dimensions. Each axis of the 2‑D view represents a linear combination of the original 4‑D coordinates. Points appear grouped into multiple elongated clusters or branches that move, rotate, and sometimes overlap as the projection changes. Branches are distinguished by colour, with individual points visible throughout. The display maintains a roughly square aspect ratio as the view transitions through multiple angles, illustrating the dataset’s branching structure from different perspectives."----
+## ----branch, eval=knitr::is_html_output(), fig.cap="Viewing the $4\\text{-}D$ `orgcurvybranches` data. The tour view shows how the linear branches appear from multiple viewing angles.", fig.alt="Interactive langevitour of the orglinearbranches dataset in $4\\text{-}D$, showing linear branches from multiple viewing angles.", fig.pos="!ht", layout="l-body", fig.alt="The figure shows a tour of the 4‑D orgcurvybranches dataset. Points are displayed in a continuously changing 2‑D projection that rotates through the four dimensions. Each axis of the 2‑D view represents a linear combination of the original 4‑D coordinates. Points appear grouped into multiple elongated clusters or branches that move, rotate, and sometimes overlap as the projection changes. Branches are distinguished by color, with individual points visible throughout. The display maintains a roughly square aspect ratio as the view transitions through multiple angles, illustrating the dataset’s branching structure from different perspectives."----
 # 
 # branchfig <- bscols(
 #   htmltools::div(
@@ -412,7 +412,7 @@ orgcurvybranches <- gen_orgcurvybranches(n = 1000, p = 4, k = 4)
 # branchfig
 
 
[…]

### 2025-12-19 paper-cardinalR.Rmd
Set up everything to match the sizes

@@
 
-knitr::include_graphics("figures/cardinalR_workflow.png")
+# GitHub raw URL of the figure
+fig_url <- "https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/figures/cardinalR_workflow.png"
+
+# Determine figure path depending on output
+if (knitr::is_html_output()) {
+  # HTML can use the URL directly
+  fig_path <- fig_url
+} else {
+  # PDF needs a local copy
+  temp_file <- tempfile(fileext = ".png")
+  download.file(fig_url, destfile = temp_file, mode = "wb")
+  fig_path <- temp_file
+}
+
+knitr::include_graphics(fig_path)
 ```
@@
 
-tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds")
+tsne_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_tsne_perplexity_30.rds", "rb"))
 
@@
 
-umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds", "rb"))
 
@@
 
-phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds")
+phate_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_phate_knn_5.rds", "rb"))
 
@@
 
-trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+trimap_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds", "rb"))
 
@@
 
-pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+pacmap_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds", "rb"))
 
@@
 
-pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds")
+pca_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_pca.rds", "rb"))
 
@@
 ```{r}
-error_five_clust <- read_rds("data/five_clusts/five_clusts_hbe.rds")
+error_five_clust <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_hbe.rds", "rb"))
 
@@
 # Examine the cluster stats
-all_stats <- read_rds("data/five_clusts/cluster_stat.rds")
+all_stats <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/cluster_stat.rds", "rb"))
 all_stats |>
@@
 
-The source material for this paper is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
+The source material for this paper, including the full datasets and figures, is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).
 This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.

### 2025-12-19 paper-cardinalR-submission-v1/paper-cardinalR.tex
SUBMITTED TO R journal !!!!

--- 2025-12-19 paper-cardinalR.tex
@@
 Arrows indicate the flow of the process from parameter specification to the final labeled dataset.
-The diagram uses a clear layout with distinct steps, showing how data are constructed in sequence.}]{figures/cardinalR_workflow}
+The diagram uses a clear layout with distinct steps, showing how data are constructed in sequence.}]{../../../../../../private/var/folders/4d/q_zhr19n0z7d8hx3xblybm8rjpxby3/T/RtmpRmAUiT/file794b45c38a65}
 
@@
 
-The source material for this paper is available at \href{https://github.com/JayaniLakshika/paper-cardinalR}{github.com/JayaniLakshika/paper-cardinalR}.
+The source material for this paper, including the full datasets and figures, is available at \href{https://github.com/JayaniLakshika/paper-cardinalR}{github.com/JayaniLakshika/paper-cardinalR}.
 This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.
